### 2018-04-22 Defense/defense.tex
initial project version

\documentclass{article}
\title{Defense of decisions}
\date{2018-04-17}
\author{Jomar Alcantara}

\begin{document}
	Question - Why focus on just language rather than speech and language given you are collecting both?
	Answer - Orimaye et al, 2017 suggested that distinction of healthy controls and those with dementia could be done solely with processing of transcripts. Given the abundance of sources of data based on transcripts and the potential burden of processing two types of data means that to keep within the scope of the project and the focus on Natural Language Processing I made the decision to see whether this could be done with just the processing of language. BETTER ANSWER?

	Question - Why did you pick the FCSRDT as one of your test battery?
	Answer - Research from the University of Lancaster suggested that this was the most effective test for differentiating between controls, MCI and full blown AD.
\end{document}

### 2018-04-22 LiteratureReview/LiteratureReview.tex
initial project version

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Document Class
\documentclass[a4paper]{article}

%% Packages
\input{settings/packages}

%% Page Settings
\input{settings/page}

%% Own Commands
\input{settings/macros}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin {document}

\tableofcontents

\section{Introduction to Dementia}

Identification of language impairment is important in Dementia because it aids diagnosis of specific types of dementia, which in turn can alter the prognosis and change the management of the degenerative disorder. As these differences in language are quite subtle, the varying subtypes of dementia are frequently misdiagnosed.

According to a recent report commissioned by the Alzheimer's Society, they estimate the prevalence of Dementia in the UK at approximately 815,000 people. This represents 1 in 14 of the population aged 65 or over. This report also estimates an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis. They also estimate that the overall impact of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually.

\section{Assessment of Language function in Dementia}

Language can be defined as the ability to encode ideas into words and/or symbols for the means of communication. Difficulty in language, both spoken and written, are often described as symptoms of various types of Alzheimers' disease (AD). One of the challenges when using speech as a predictor is recognising whether there is a problem with language production or with the motor skills. Impairments in speech that arise from any process that disrupts the neuraxis from the cortex to muscle and encompass dysarthria (disturbance in articulation) and dysphonia (disturbance in the production in vocal sounds). However it is important to use language as a tool to aid the diagnosis of Dementia, as it provides vital clues to aid a clinician in differentiating in the different types of Dementia, which will in turn aid a clinician in an attempt to manage an individual case of Dementia. Due to the subtle nature of the languages changes that are experienced in those who suffer with different types of Dementia, it is often misdiagnosed.

A common way for clinicians' to diagnose dementia is through use of the Mini-Mental State Examination. The Mini-Mental State Examination (MMSE) has shown to be particularly effective at differentiating betweeen Dementia and other psychological disorders such as clinical depression, schitzophrenia and personality disorder and is currently the most widely used measure of diagnosing Dementia within clinical psychology.  However, it is not without it's faults...

\section{Current ways of diagnosing Dementia using Language}

Mini Mental State Examination (MMSE)
Addenbrooke's Cognitive Examination (ACE-III)
Dementia Rating Scale (DRS)
Western Aphasia Battery (WAB)
Boston Diagnostic Aphasia Examination (BDAE)
Boston Naming Test (BNT)

\section{Existing studies that have used Natural Language Processing to help in diagnosing psychological difficulties}

Psychology - Schitzophrenia

\section{What progress has been made in this area already?}

\section{State of literature into Natural Language processing techniques}

\end {document}

### 2018-04-22 Workplan/Workplan.tex
initial project version

\documentclass{article}

\title{Workplan}
\date{2018-04-17}
\author{Jomar Alcantara}

\begin{document}
	\maketitle
	%\newpage
	\section{Interest Statement}
	I am interested in whether technology can be used to aid the early diagnosis of dementia specifically by analysing speech and language in a natural environment. As part of this, I am exploring the use of technologies such as the Amazon Echo or Google Home kit to capture speech and language. I am using Natural Language Processing to generate features from the captured speech such that we can categorise people into specific categories and/or predict test results on scores using machine learning.
	\section{Literature Review}
	\subsection{Dementia - Context}
	What is the Problem? 
	Why is this important?
	What is the potential impact?
	\subsection{How do Psychologists assess dementia}
	What tools do psychologists use to diagnose dementia? 
	Mini mental state examination (MMSE), 
	Addenbrooke's Cognitive Examination(ACE-III)
	(WAIS?)
	\subsection{How do Psychologists understand how language is affected by dementia}
	What language deficits are there and how are they characterised?
	What speech deficits are there and how are they characterised?
	\subsection{How is Natural Language Processing best positioned to help in this area?}
	\subsection{Are there any other technologies that can be used to help diagnosis?}

	\subsection{What Research has been done in the area so far?}
	\subsection{Conclusion}

	\section{Training needs analysis}
	\subsection{Psychology}
	Training in delivering the WAIS or psychological tools that might be needed 
	\subsection{Programming}
	\begin{itemize}
		\item Online courses in Python Programming
		\item Online courses in Machine Learning
	\end{itemize}

	\subsection{Natural Language Processing}
	Online courses in Natural Language Processing
	Specific training in Natural Language Processing - Summer Schools?

	\subsection{Signal Processing}
	MATLAB? Speech processing?

	\subsection{Alexa Skills Kit and Amazon Web Services}
	\begin{itemize}
		\item \textbf{Training in Alexa Skills Kit - Booked}
		\item Training in Amazon Web Services
	\end{itemize}

	\section{Data collection}
	\subsection{DementiaBank and AphasiaBank}
	Application to be submitted.

	\subsection {Three Authors, P.D. James, Iris Murdoch and Agatha Christie}	
	Data to be downloaded

	\subsection{Presidents Speeches}

	\section{Proposed Timeline, Key Dates and Deliverables}
	\subsection{April 2018}
	Wednesday, 18th April - Workplan (First Draft Agreed)
	\subsection{May 2018}
	Wednesday, 9th May - Alexa Skills Training (All Day)
	\subsection{June 2018}
	Friday 15th June - First Draft of Literature to be handed in.
	Monday 18th June - Monday 25th June - Annual Leave
	\subsection{July 2018}

	\subsection{August 2018}

	\subsection{September 2018}

	\subsection{January 2019}
	Friday 18th January - Complete first rough pass of capturing and processing speech.

	\subsection{February 2019}
	Friday, 1st February - Viva and Qualifying Report Due
\end{document}

### 2018-04-22 Python/Processor.py
added placeholder file

@@ -1,0 +1,1 @@
+# Placeholder for Processing transcribed data

### 2018-04-22 QualifyingReport/QualifyingReport.tex
reorganised folders and added folder for qualifying report

\documentclass{article}

\title{Qualifying Report}
\date{2018-04-22}
\author{Jomar Alcantara}

\begin{document}
	\section{Overview of Problem}
	\section{Literature Review}
	\section{Work already carried out}
	\section{Work Plan for the future}
	\section{Proposed Timetable}
	\section{Provision Table of Contents}
\end{document}

### 2018-04-22 Workplan/Workplan.tex
added Qualifying Report

@@
 	\subsection{Presidents Speeches}
+	Data to be downloaded
+
+	\subsection{Experimental Design}
+	Battery of tests to differentiate between Controls, those with MCI, those with AD. 
+	\begin{itemize}
+		\item FCSRDT - As most effective 
+	\end{itemize}
 
@@
 
+	\subsection{October 2018}
+
+	\subsection{November 2018}
+
+	\subsection{December 2018}
+
 	\subsection{January 2019}
+	Friday 11th January - Rough Draft of Qualifying Due
 	Friday 18th January - Complete first rough pass of capturing and processing speech.

### 2018-04-23 QualifyingReport/QualifyingReport.tex
added Qualifying Report
workplan updated

@@
 \begin{document}
+	\maketitle
+	\newpage
 	\section{Overview of Problem}
+	I am interested in whether technology can be used to aid the early diagnosis of dementia specifically by analysing speech and language in a natural environment.
+As part of this, I am exploring the use of technologies such as the Amazon Echo or Google Home kit to capture speech and language.
+I am using Natural Language Processing to generate features from the captured speech such that we can categorise people into specific categories and/or predict test results on scores using machine learning.
+
 	\section{Literature Review}
+
 	\section{Work already carried out}
+
 	\section{Work Plan for the future}
+
 	\section{Proposed Timetable}
+	\subsection{May 2019}
+
+	\subsection{June 2019}
+
+	\subsection{July 2019}
+
+	\subsection{August 2019}
+
+	\subsection{September 2019}
+
+	\subsection{October 2019}
+
+	\subsection{November 2019}
+
+	\subsection{December 2019}
+
+	\subsection{January 2020}
+
+	\subsection{February 2020}
+
+	\subsection{March 2020}
+
+	\subsection{April 2020}
+
+	\subsection{May 2020}
+
+	\subsection{June 2020}
+
+	\subsection{July 2020}
+
+	\subsection{August 2020}
+
+	\subsection{September 2020}
+
+	\subsection{October 2020}
+
+	\subsection{November 2020}
+
+	\subsection{December 2020}
+
+	\subsection{January 2021}
+
+	\subsection{February 2021}
+
+	\subsection{March 2021}
+
+	\subsection{April 2021}
 	\section{Provision Table of Contents}

### 2018-04-23 Workplan/Workplan.tex
workplan updated

@@
 	\maketitle
-	%\newpage
 	\section{Interest Statement}
@@
 	Addenbrooke's Cognitive Examination(ACE-III)
-	(WAIS?)
+	Wechsler Adult Intelligence Scale(WAIS)
+	The Free and Cued Selective Reminding Test (FCSRT)
 	\subsection{How do Psychologists understand how language is affected by dementia}
@@
 	Online courses in Natural Language Processing
-	Specific training in Natural Language Processing - Summer Schools?
+
+	\begin{itemize}
+		\item Jurafsky and Martin - Speech and Language Processing 
+		\item Manning and Schutze - Statistical NLP 
+		\item Manning and Raghavan - Intro to IR
+	\end{itemize}
 
@@
 	\subsection{June 2018}
-	Friday 15th June - First Draft of Literature to be handed in.
+	Friday 15th June - First Draft of Literature Review to be handed in.
+	\newline
 	Monday 18th June - Monday 25th June - Annual Leave
@@
 	\subsection{February 2019}
-	Friday, 1st February - Viva and Qualifying Report Due
+	Friday, 1st February - Qualifying Report Due
+
+	\subsection{March 2018}
+
+	\subsection{April 2018}
+	Monday, 1st April - Viva Due by this point.
 \end{document}

### 2018-04-24 LiteratureReview/LiteratureReview.tex
added Qualifying Report
workplan updated
updated structure on Literature Review

@@
 
+\title{Literature Review}
+\date{2018-04-22}
+\author{Jomar Alcantara}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin {document}
+\maketitle
+\newpage
+\tableofcontents
+\newpage
 
-\tableofcontents
-
-\section{Introduction to Dementia}
-
+\section{Introduction to Dementia and the Literature Review}
+\par
 Identification of language impairment is important in Dementia because it aids diagnosis of specific types of dementia, which in turn can alter the prognosis and change the management of the degenerative disorder.
 As these differences in language are quite subtle, the varying subtypes of dementia are frequently misdiagnosed.
-
+\newline
+\par
 According to a recent report commissioned by the Alzheimer's Society, they estimate the prevalence of Dementia in the UK at approximately 815,000 people.
@@
 They also estimate that the overall impact of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually.
+\newline
+\par
+The purpose of this review is to seek to understand how dementia impacts speech and language production from a clinical psychology perspective. This review goes on to
+
+% Describe how you have conducted the search. What did you search, how did you search, how did you filter.
+
+\section{Description of Dementia}
+
+\subsection{Alzheimer's Disease}
+\subsection{Vascular Dementia}
+\subsection{Dementia with Lewy Bodies}
+\subsection{Parkinson's disease dementia}
+\subsection{Fronto-temporal dementia}
+\subsection{Mild Cognitive Impairment}
 
@@
 
+\section{Conclusion}
+Given the burden on the diagnosis of dementia on clinicians, it appears to be useful to find some non-invasive protocols for the early diagnosis of dementia. 
 \end {document}

### 2018-04-25 Python/Processor.py
added function to feature generator

@@ -1,1 +1,8 @@
 # Placeholder for Processing transcribed data
+
+import nltk
+
+
+# Calculate lexical_diversity of the data
+def lexical_diversity(text):
+    return len(set(text)) / len(text)

### 2018-04-26 LiteratureReview/LiteratureReview.tex
reorganised workplan and added more to literature review

@@
 \title{Literature Review}
-\date{2018-04-22}
+\date{2018-04-26}
 \author{Jomar Alcantara}
@@
 \subsection{Fronto-temporal dementia}
+\subsection{Primary Progressive Aphasia}
 \subsection{Mild Cognitive Impairment}
@@
 \section{Current ways of diagnosing Dementia using Language}
+Battery of tests to differentiate between Controls, those with MCI, those with AD. 
+\begin{itemize}
+	\item FCSRDT - As most effective 
+	\item Mini Mental State Examination (MMSE)
+	\item Addenbrooke's Cognitive Examination (ACE-III)
+	\item Dementia Rating Scale (DRS)
+	\item Western Aphasia Battery (WAB)
+	\item Boston Diagnostic Aphasia Examination (BDAE)
+	\item Boston Naming Test (BNT)
+\end{itemize}
 
-Mini Mental State Examination (MMSE)
-Addenbrooke's Cognitive Examination (ACE-III)
-Dementia Rating Scale (DRS)
-Western Aphasia Battery (WAB)
-Boston Diagnostic Aphasia Examination (BDAE)
-Boston Naming Test (BNT)
+\section{Existing studies that have used Natural Language Processing to help in diagnosing psychological difficulties, particularly language difficulties}
 
-\section{Existing studies that have used Natural Language Processing to help in diagnosing psychological difficulties}
-
-Psychology - Schitzophrenia
-
-\section{What progress has been made in this area already?}
+\subsection{Nun Study - Snowdon (1996)}
+\subsection{Comparing Presidents Speeches - Berisha et al (2015)}
 

### 2018-04-26 Workplan/Workplan.tex
updated structure on Literature Review
reorganised workplan and added more to literature review

@@
 
-	\section{Training needs analysis}
+	\section{Stages of research}
+	\section{Stage 1 - Data capture}
+	\subsection{Training needs analysis}
 	\subsection{Psychology}
@@
 
+	\section{Data collection}
+	\subsection{DementiaBank and AphasiaBank}
+	\textbf{Holland Corpus}
+	\newline 
+	Consists of 2 participants a 62 year old man with mild AD and a 78 year old woman with more advanced AD.
+	\newline
+	\textbf{Kempler Corpus}
+	\newline
+	Consists of 6 participants with a range of 65 to 86 (mean, sd)
+	\newline
+	\textbf{Pitt Corpus}
+	\newline
+	Consists of 104 control participants, 208 with dementia and 85 with an unknown diagnosis.
+	\newline
+	\textbf{DePaul Corpus}
+	\newline
+	Consists of 1 participant, 66 year old female with PPA.
+	\newline
+	\textbf{Hopkins Corpus}
+	\newline
+	Consists of 36 participants, 23 males and 13 females
+
+	\subsection {Three Authors, P.D. James, Iris Murdoch and Agatha Christie}	
+	Data to be downloaded
+
+	\subsection{Presidents Speeches}
+	Data to be downloaded
+
+	\subsection{Alexa Skills Kit and Amazon Web Services}
+	\begin{itemize}
+		\item \textbf{Training in Alexa Skills Kit - Booked}
+		\item Training in Amazon Web Services
+	\end{itemize}
+
+	\section{Stage 2 - Data Processing, Feature Construction}
 	\subsection{Natural Language Processing}
@@
 
-	\subsection{Signal Processing}
-	MATLAB? Speech processing?
-
-	\subsection{Alexa Skills Kit and Amazon Web Services}
-	\begin{itemize}
-		\item \textbf{Training in Alexa Skills Kit - Booked}
-		\item Training in Amazon Web Services
-	\end{itemize}
-
-	\section{Data collection}
-	\subsection{DementiaBank and AphasiaBank}
-	Application to be submitted.
-
-	\subsection {Three Authors, P.D. James, Iris Murdoch and Agatha Christie}	
-	Data to be downloaded
-
-	\subsection{Presidents Speeches}
-	Data to be downloaded
-
-	\subsection{Experimental Design}
-	Battery of tests to differentiate between Controls, those with MCI, those with AD. 
-	\begin{itemize}
-		\item FCSRDT - As most effective 
-	\end{itemize}
+	\section{Stage 3 - Data Analysis and Machine Learning}
+	\subsection{Application of Machine Learning Algorithms}
 

### 2018-04-26 LiteratureReview/LiteratureReview.tex
continue to work on literature review and make amendments to other documents

@@
 
-\section{Introduction to Dementia and the Literature Review}
-\par
-Identification of language impairment is important in Dementia because it aids diagnosis of specific types of dementia, which in turn can alter the prognosis and change the management of the degenerative disorder.
-As these differences in language are quite subtle, the varying subtypes of dementia are frequently misdiagnosed.
-\newline
+\section{Introduction}
+
 \par
@@
 They also estimate that the overall impact of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually.
+Globally, this picture is a lot bleaker.
+A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050(CITE).
+The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillon US dollars.
 \newline
 \par
-The purpose of this review is to seek to understand how dementia impacts speech and language production from a clinical psychology perspective. This review goes on to
+The purpose of this review is to seek to understand how dementia impacts speech and language production from a psychological and linguistic perspective with a view to potentially using natural language processing and technology to aid in the early diagnosis of dementia and/or Mild Cognitive Impairment.
+\newline
+\par
+A search of the literature was conducted mainly from a psychological perspective as this would inform the theoretical side of things.
+Next, a search of the literature from a Computer Science perspective was also made with a view to exploring the how domain of machine learning and more specifically natural language processing can be applied to this problem.
+Finally a search was carried out to see what work has already been done in this area.
 
-% Describe how you have conducted the search. What did you search, how did you search, how did you filter.
+All stages of this literature review were completed by myself.
 
@@
 \section{Conclusion}
+\par
+Identification of language impairment is important in Dementia because it aids diagnosis of specific types of dementia, which in turn can alter the prognosis and change the management of the degenerative disorder.
+As these differences in language are quite subtle, the varying subtypes of dementia are frequently misdiagnosed.
+\newline
+\par 
 Given the burden on the diagnosis of dementia on clinicians, it appears to be useful to find some non-invasive protocols for the early diagnosis of dementia.
+It has already been shown that analysis of speech and language has shown markers that pre-date the official diagnosis of dementia (Snowdon et al, 1996; Berisha et al, 2015).
 \end {document}

### 2018-04-26 QualifyingReport/QualifyingReport.tex
updated structure on Literature Review
continue to work on literature review and make amendments to other documents

@@
 
+	\subsection{April 2018}
+
+	Thursday, 26th April - Workplan (First Draft Agreed)
+
 	\section{Work Plan for the future}
@@
 	\section{Proposed Timetable}
+
+	\subsection{May 2018}
+	Wednesday, 9th May - Alexa Skills Training (All Day)
+	\subsection{June 2018}
+	Friday 15th June - First Draft of Literature Review to be handed in.
+	\newline
+	Monday 18th June - Monday 25th June - Annual Leave
+	\subsection{July 2018}
+
+	\subsection{August 2018}
+
+	\subsection{September 2018}
+
+	\subsection{October 2018}
+
+	\subsection{November 2018}
+
+	\subsection{December 2018}
+
+	\subsection{January 2019}
+	Friday 11th January - Rough Draft of Qualifying Due
+	Friday 18th January - Complete first rough pass of capturing and processing speech.
+
+	\subsection{February 2019}
+	Friday, 1st February - Qualifying Report Due
+
+	\subsection{March 2019}
+
+	\subsection{April 2019}
+	Monday, 1st April - Viva Due by this point.
 	\subsection{May 2019}
@@
 	\section{Provision Table of Contents}
+	\subsection{Chapter 1: Introduction}
+	\subsection{Chapter 2: Literature Review}
+	\subsection{Chapter 3: New theories, Description of experiments}
+	\subsection{Chapter 4: Results}
+	\subsection{Chapter 5: Discussion}
+	\subsection{Chapter 6: Conclusion}
 \end{document}

### 2018-04-26 Workplan/Workplan.tex
continue to work on literature review and make amendments to other documents

@@
 	\subsection{April 2018}
-	Wednesday, 18th April - Workplan (First Draft Agreed)
+	Thursday, 26th April - Workplan (First Draft Agreed)
 	\subsection{May 2018}
@@
 	\subsection{August 2018}
+	Friday 10th August - First draft of experimental proposal to be submitted.
 
 	\subsection{September 2018}
-
+	Friday 21st September - Final draft of experimental proposal to be submitted. 
+	Friday 21st September - NHS Ethics forms to be sorted  - Takes 60 days!
 	\subsection{October 2018}

### 2018-04-28 LiteratureReview/LiteratureReview.tex
updated Lit Review

@@
 \section{Description of Dementia}
+
+Language - DSM-V
+'Major' Has significant difficulties with expressive or receptive language.
+Often uses general-use phrases such as "that thing" and "you know what I mean," and prefers general pronouns rather than names.
+With severe impairment, may not even recall names of closer friends and family.
+Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur.
+Stereotypy of speech occurs; echolalia and automatic speech typically precede mutism.
+
+'Mild' Has noticeable word-finding difficulty.
+May substitute general for specific terms.
+May avoid use of specific names of acquaintances.
+Grammatical errors involve subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc.
 

### 2018-04-30 QualifyingReport/QualifyingReport.tex
added more to overview of problem

@@
 	\section{Overview of Problem}
-	I am interested in whether technology can be used to aid the early diagnosis of dementia specifically by analysing speech and language in a natural environment.
-As part of this, I am exploring the use of technologies such as the Amazon Echo or Google Home kit to capture speech and language.
-I am using Natural Language Processing to generate features from the captured speech such that we can categorise people into specific categories and/or predict test results on scores using machine learning.
+
+	Dementia has been identified as one of those fast growing difficulties facing the world.
+A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050(CITE).
+The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillon US dollars.
+	\newline
+	\par
+	A lot of work has gone into language analysis based on clinician diagnosis.
+This takes the form of tasks such as Free Cued Selective Recall Task, the Boston Cogntive Assessment ,and the Mini Mental State Examination.
+All of which are quite targeted in the language required of the person being examined.
+A step towards the ambient detection of AD and/or MCI in population requires that we look for similar features in language used naturally as part of general conversation.
+Part of this work therefore, is to see whether non prescriptive tasks (that is, tasks that do not have a pre determined right answer), can elicit the same or comparable features in language such that this can be detected ambiently.
+The next part of this work looks at how technology, particularly the use of smart home devices such as the Amazon Echo and Google Home families of devices.
+	\newline
+	\par
+	The potential impact of this research is immense.
+Research has shown (has it?) that early diagnosis of people with AD or MCI improves people's quality of life and can slow the progress of the disease.
+Equally, early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
 
@@
 	\section{Work Plan for the future}
+	\subsection{Proposed Experiments}
 
@@
 
-	\subsection{April 2021}
 	\section{Provision Table of Contents}

### 2018-04-30 LiteratureReview/LiteratureReview.tex
added bibliographies to each document

@@
 Globally, this picture is a lot bleaker.
-A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050(CITE).
+A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
 The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillon US dollars.
@@
 \section{Current ways of diagnosing Dementia using Language}
+\subsection{Free Cued Selective Recall Test}
+The Free Cued Selective Recall Test (FCSRT) was created by Grober and his team \cite{Grober1988}.
+
 Battery of tests to differentiate between Controls, those with MCI, those with AD. 
@@
 It has already been shown that analysis of speech and language has shown markers that pre-date the official diagnosis of dementia (Snowdon et al, 1996; Berisha et al, 2015).
+
+\bibliographystyle{unsrt}
+\bibliography{LiteratureReview}
 \end {document}

### 2018-04-30 QualifyingReport/QualifyingReport.tex
added bibliographies to each document

@@
 
-	\subsection{April 2018}
+	\subsection{Literature Review - April 2018 to July 2018}
+	A literature review was carried out from several perspectives in order to understand both an applied and theoretical perspective.
+Thus a search for papers was conducted in both the psychology and computer science domains, as well as papers which intersected these areas and were relevant to this particular problem.
 
-	Thursday, 26th April - Workplan (First Draft Agreed)
+	\subsection{Skills development - April 2018 to March 2019}
 
@@
 	\subsection{Chapter 6: Conclusion}
+
+	\bibliographystyle{unsrt}
+	\bibliography{QualifyingReport}
 \end{document}

### 2018-05-02 LiteratureReview/LiteratureReview.tex
edited Literature Review and Qualifying Report
changes to Literature Review and QR

@@
 
+Emery \cite{Emery2000}, states that you can see language decline as Hierarchical.
+According to a semiotic model there are four ranks of language based on their cognitive complexity.
+These ranks are Phonology, Morphology, Syntax and Semantics.
+She conducted a review studies which looked at language decline from the perspective of these hierarchical ranks.
+Her conclusion is that language decline is related to the complexity of the language task given to a participant and that language decline is hierarchical in that the language forms we learn last (the most complex language forms) are the first to deteriorate.
+
 A common way for clinicians' to diagnose dementia is through use of the Mini-Mental State Examination.
@@
 The Free Cued Selective Recall Test (FCSRT) was created by Grober and his team \cite{Grober1988}.
+
+\subsection{Mini-Mental State Examination}
 
@@
 Given the burden on the diagnosis of dementia on clinicians, it appears to be useful to find some non-invasive protocols for the early diagnosis of dementia.
-It has already been shown that analysis of speech and language has shown markers that pre-date the official diagnosis of dementia (Snowdon et al, 1996; Berisha et al, 2015).
+It has already been shown that analysis of speech and language has shown markers that pre-date the official diagnosis of dementia (Snowdon et al, 1996;)\cite{Berisha2015}.
 

### 2018-05-02 QualifyingReport/QualifyingReport.tex
edited Literature Review and Qualifying Report
changes to Literature Review and QR

@@
 \title{Qualifying Report}
-\date{2018-04-22}
+\date{2018-05-02}
 \author{Jomar Alcantara}
@@
 	\section{Overview of Problem}
-
 	Dementia has been identified as one of those fast growing difficulties facing the world.
 A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050(CITE).
-The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillon US dollars.
+The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
 	\newline
 	\par
-	A lot of work has gone into language analysis based on clinician diagnosis.
-This takes the form of tasks such as Free Cued Selective Recall Task, the Boston Cogntive Assessment ,and the Mini Mental State Examination.
-All of which are quite targeted in the language required of the person being examined.
-A step towards the ambient detection of AD and/or MCI in population requires that we look for similar features in language used naturally as part of general conversation.
-Part of this work therefore, is to see whether non prescriptive tasks (that is, tasks that do not have a pre determined right answer), can elicit the same or comparable features in language such that this can be detected ambiently.
-The next part of this work looks at how technology, particularly the use of smart home devices such as the Amazon Echo and Google Home families of devices.
+	A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cogntive Impairment (MCI).
+As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages and thus it is important that we utilise technology wherever possible to aid clinicians.
+At the present time diagnosis is typically conducted at memory clinics by trained clinicians.
+I theorise that we might be able to use diagnose patients in their own homes utilising smart-home technology.
+	\newline
+	\par
+	Language analysis is an important part of assessing those suspected of suffering with AD or MCI and this language analysis in part informs an official clinical diagnosis.
+This analysis often takes the form of tasks such as Free Cued Selective Recall Task, the Boston Cognitive Assessment, and the Mini Mental State Examination which are all quite targeted in the language required of the person being assessed.
+	\newline
+	\par
+	A step towards the ambient detection of AD and/or MCI in population requires that we look for similar features in language used naturally as part of every day conversation.
+Part of this work therefore is to see whether spontaneous discourse, such a semi-structured interview is a task that has the ability to put pressure both the cognitive and linguistic systems as it requires the ability to understand speech, formulating a coherent answer and then responding in an appropriate way.
+Berisha \cite{Berisha2015}, has shown through language analysis in this area that there are marked differences in this process between those who have a diagnosis of AD and healthy controls.
+The next part of this work looks at how technology, particularly the use of smart home devices such as the Amazon Echo and Google Home families of devices can be used to capture data that can then be analysed.
 	\newline
@@
 	The potential impact of this research is immense.
-Research has shown (has it?) that early diagnosis of people with AD or MCI improves people's quality of life and can slow the progress of the disease.
+Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can slow the progress of the disease.
 Equally, early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
@@
 	\subsection{Skills development - April 2018 to March 2019}
+	\textbf{IS-4001 Research Skills and Professional Development Module - Aston University - 3rd May, 10th May, 16th May - 40 Hours} 
+	Summary Here
+	\newline
+	\textbf{Alexa Skills Training Day - Amazon - May 9th 2018 - 7 Hours}
+	Summary Here
+	\newline
+	\textbf{Deep Learning Online Course - deeplearning.ai - Dates - 70 Hours}
+	Summary Here
 
@@
 	Monday 18th June - Monday 25th June - Annual Leave
-	\subsection{July 2018}
-
-	\subsection{August 2018}
-
-	\subsection{September 2018}
-
-	\subsection{October 2018}
-
-	\subsection{November 2018}
-
-	\subsection{December 2018}
-
 	\subsection{January 2019}
@@
 	Friday, 1st February - Qualifying Report Due
-
-	\subsection{March 2019}
-
 	\subsection{April 2019}
 	Monday, 1st April - Viva Due by this point.
-	\subsection{May 2019}
 
-	\subsection{June 2019}
+	\section{Reflection on my experience so far}
 
-	\subsection{July 2019}
-
-	\subsection{August 2019}
-
-	\subsection{September 2019}
-
-	\subsection{October 2019}
-
-	\subsection{November 2019}
-
-	\subsection{December 2019}
-
-	\subsection{January 2020}
-
-	\subsection{February 2020}
-
-	\subsection{March 2020}
-
-	\subsection{April 2020}
-
-	\subsection{May 2020}
-
-	\subsection{June 2020}
-
-	\subsection{July 2020}
-
-	\subsection{August 2020}
-
-	\subsection{September 2020}
-
-	\subsection{October 2020}
-
-	\subsection{November 2020}
-
-	\subsection{December 2020}
-
-	\subsection{January 2021}
-
-	\subsection{February 2021}
-
-	\subsection{March 2021}
-
-	\section{Provision Table of Contents}
+	\section{Provisional Table of Contents}
 	\subsection{Chapter 1: Introduction}

### 2018-05-03 QualifyingReport/QualifyingReport.tex
updated LR and QR
added LiteratureReview3000

@@
 	\par
-	A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cogntive Impairment (MCI).
+	In 2009, the Department of Health in it's National Dementia strategy made 'early diagnosis and support' one of it's key themes as part of this strategy.
+A lot of work, therefore, has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cogntive Impairment (MCI).
 As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages and thus it is important that we utilise technology wherever possible to aid clinicians.
@@
 	\subsection{Skills development - April 2018 to March 2019}
-	\textbf{IS-4001 Research Skills and Professional Development Module - Aston University - 3rd May, 10th May, 16th May - 40 Hours} 
+	\textbf{IS-4001 Research Skills and Professional Development Module - Aston University - 3rd May, 10th May, 16th May - 45 Hours} 
 	Summary Here
@@
 	\subsection{Proposed Experiments}
+	\subsection{Publication plan}
 

### 2018-05-10 QualifyingReport/QualifyingReport.tex
updated Qualifying Report to add Day 1 of IS4001

@@
 \title{Qualifying Report}
-\date{2018-05-02}
+\date{2018-05-10}
 \author{Jomar Alcantara}
@@
 	\subsection{Skills development - April 2018 to March 2019}
-	\textbf{IS-4001 Research Skills and Professional Development Module - Aston University - 3rd May, 10th May, 16th May - 45 Hours} 
-	Summary Here
+	\textbf{IS-4001 Research Skills and Professional Development Module - Aston University - 3rd May, 10th May, 16th May - 45 Hours} \newline
+	\par
+	Day 1 consisted of an introduction to the module (IS4001), a rough guide to how to start a PhD and a look at how to conduct and write a literature review.
+This looked at the concept of a PhD and the meaning of original research.
+This also talked about the supervisor relationship and understanding the role of the supervisor in the PhD.
+In addition, we explored the idea of our expectations around supervision.
+We then went on to look at what skills, qualities and attitudes are necessary to be a good researcher including the idea of what consitutes ethical research and finally we looked at the obstacles to progress on a PhD and what we can do as students.
 	\newline
+	\par
+	In the afternoon, we looked at how a way of conducting a literature review.
+This talked about what a literature review is, along with the different types of literature review (Narrative, Systematic and Meta-Analysis).
+We talked about the function of a literature review, and how this fits in with the writing up of a thesis.
+We then explored the stages that the literature review process contains, from defining the scope of the research, to coming up with a plan on how to collect sources of information to synthesising and reporting the findings of the review.
+This part of the lecture looked in detail at each of these areas and included some examples to put the theory of the literature into context.
+Finally, we talked about the the process of a literature review in the context of the written assignment.
+	\newline
+	\par
 	\textbf{Alexa Skills Training Day - Amazon - May 9th 2018 - 7 Hours}
-	Summary Here
+	\newline
+	\par
+	Test
 	\newline

### 2018-05-14 LiteratureReview/LiteratureReview.tex
updated LR and QR
updated

@@
 The purpose of this review is to seek to understand how dementia impacts speech and language production from a psychological and linguistic perspective with a view to potentially using natural language processing and technology to aid in the early diagnosis of dementia and/or Mild Cognitive Impairment.
+\newline
+\par 
+Aims and Objectives:
 \newline
@@
 The Free Cued Selective Recall Test (FCSRT) was created by Grober and his team \cite{Grober1988}.
+This is a test that primarily tests memory but also tests a subjects ability to name objects.
+Four items are presented at a time on a card and each participant is asked to name and recall with and without cues.
 
@@
 
+\section{stuff to add}
+Annotated Bibliography for the project "Ambient Monitoring of Speech for Evidence of Cognitive Decline".
+
+Orimaye, S. O., Wong, J. S-M., Golden, K. J., Wong, C. P., & Soyiri, I. N. (2017) Predicting probably Alzheimer's Disease using linguistic deficits and biomarkers.
+
+There is an increasing body of literature which supports the use of machine learning to analyse speech to evidence cognitive decline.
+Orimaye et al (2017) investigate the use of machine learning algorithms to detect differences in syntactic, lexical and n-gram linguistic biomarkers to distinguish between those with probably AD and healthy controls, The authors found significant differences in the uses of all three types of biomarkers for those with AD and healthy controls.
+Their results show that the top 1000 n-gram features plus the twenty-three syntactic and lexical features was the most successful at differentiating the two groups (0.93 AUC)).
+However, there were some limitations of the study.
+The context of the audio content was very specific, limited to one specific description task, and therefore using general speech as proposed in this project would potentially use a different set of linguistic features which may or may not have similar predictive power.
+In addition, the transcriptions were encoded using the CHAT format which is a framework for manually annotating speech and the challenge would be finding a way to automate this.
+This paper is noteworthy as it provides evidence that using machine learning to identify participants with MCI is possible and it provides some ideas of potential features which can be used.
+Unique to this paper is the use of n-gram features, which could be explored in the proposed project.
+
+Asgari, M., Kaye, J., & Dodge, H. (2017) Predicting mild cognitive impairment from spontaneous spoken utterances. Alzheimers & Dementia: Translational Research & Clinical Interventions 3, 219 - 228
+
+Asgari, Kaye and Dodge (2017) also looked at the linguistic characteristics of older adults with mild cognitive impairment (MCI) vs healthy controls where they hypothesised that they would be able to predict those with MCI, a distinguishing characteristic of Alzheimer's disease and other variations of dementia.
+Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words.
+They then applied support vector machines (SVM's) and random forest classification algorithms to investigate whether machine learning could be used to distinguish between those with MCI and healthy controls.
+They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84%.
+The authors report that this method is highly reliant on high-fidelity transcription of the conversations which is labour intensive, but they anticipate that technology is advancing sufficiently quickly that automated high-quality transcriptions are possible in the near future.
+This paper provides a different perspective in how to tackle the problem deriving language features from speech using a different framework.
+The challenges both Orimaye et al and these authors experienced were around the automation of transcription and this would potentially be an area to explore in the proposed project.
+
+Pennebaker, J.W., Boyd, R. L., Jordan, K and Blackburn, K., (2015) The development and psychometric properties of LIWC2015
+
+Asgari et al (2017) used the Linguistic Inquiry and Word Count (LIWC) methodology, developed by Pennebaker et al (2007) to generate features for their paper.
+Pennebaker, Boyd, Jordan and Blackburn (2015) have since developed their framework.
+The LIWC is designed to categorize and evaluate the various emotional, cognitive and structural components which are present within samples of speech and/or written text.
+This method, Linquistic Inquiry & Word Count (LIWC) has evolved from it's initial incarnation in 1993 to it's last update in 2015.
+The premise is the use of words from particular categories provide an insight into the psychological processes and/or diagnoses an individual has.
+The latest version of this methodology uses a dictionary of 6400 words, word stems and select emoticons and assigns these constructs to various categories.
+This has been updated to include modern uses of language such as 'text speak'.
+This manual goes summarises the process of constructing the framework as well as discussing the reliability and validity of the measure and a review of studies using the LIWC was conducted by Tausczik and Pennebaker (2010) support the notion that the LIWC is valid across multiple psychological domains.
+This paper is useful as it provides a useful and validated framework from which the proposed project can potentially derive language features.
+
+Schwartz, A.
+H., Eichstaedt, J., Kern, M.
+L., Park, G., Sap.
+M., Stillwell.
+D., .
+.
+.
+Ungar.
+L.
+(2014)  Towards Assessing Changes in Degree of Depression through Facebook.
+Workshop on Computational Linguistics and Clinical Psychology: From Linguistic Signal to Clinical Reality, 118-125
+Currently formal diagnosis of depression is expensive and time-consuming involving both the use of questionnaires and the use of a trained professional to assess an individual.
+Schwartz et al, explored the use of language as an aid to diagnose depression in a naturalistic setting i.e.
+facebook status updates.
+The authors used both the LIWC (described above) and n-grams as features to distinguish those with depression from healthy individuals and were able to track participants levels' of depression through language successfully.Whilst the authors were able to use a reasonably sound method of measure degree of depression and have evidenced a methodology that is able to track a levels of depression over time, it would be more helpful to use a more clinically relevant measure of depression and anxiety i.e., PHQ-9 (Kroenke, 2001) and GAD-7 (Spitzer et al, 2006) to add validity to their findings.
+This paper is interesting as it uses both LIWC and n-grams as a method of deriving features from text / speech for depression which, as described above, have also been used as features to distinguish MCI from healthy participants.
+Given the comorbity of depression and dementia (Meyers, 1998), it would be interesting for the proposed project to further analyse the link between dementia and clinical depression via the use of language.
+This paper also supports the idea that language changes over time and is a marker for deterioration in both those with depression and dementia.
+
 \section{Conclusion}

### 2018-05-14 LiteratureReview/LiteratureReview3000.tex
added LiteratureReview3000
updated

--- 2018-05-02 LiteratureReview/LiteratureReview.tex
@@
 \newpage
-
 \section{Introduction}

### 2018-05-14 QualifyingReport/QualifyingReport.tex
updated

--- 2018-05-03 QualifyingReport/QualifyingReport.tex
@@
 Part of this work therefore is to see whether spontaneous discourse, such a semi-structured interview is a task that has the ability to put pressure both the cognitive and linguistic systems as it requires the ability to understand speech, formulating a coherent answer and then responding in an appropriate way.
-Berisha \cite{Berisha2015}, has shown through language analysis in this area that there are marked differences in this process between those who have a diagnosis of AD and healthy controls.
+Berisha et al \cite{Berisha2015}, has shown through language analysis in this area that there are marked differences in this process between those who have a diagnosis of AD and healthy controls.
 The next part of this work looks at how technology, particularly the use of smart home devices such as the Amazon Echo and Google Home families of devices can be used to capture data that can then be analysed.
@@
 
+	%Complete appendices for each of the assessment. How do you do appendices in LaTeX
+
 	\bibliographystyle{unsrt}

### 2018-05-14 Workplan/Workplan.tex
updated Lit Review
changes to Literature Review and QR
updated

@@
 	\subsection{July 2018}
+	Friday 13th July - Written Assignment (IS4001) Due
+	Friday 13th July - Project Plan (IS4001) Due
 
@@
 	Friday 18th January - Complete first rough pass of capturing and processing speech.
+	Thursday, 31st January - Qualifying Report Due
 
 	\subsection{February 2019}
-	Friday, 1st February - Qualifying Report Due
 
-	\subsection{March 2018}
+	\subsection{March 2019}
 
-	\subsection{April 2018}
+	\subsection{April 2019}
 	Monday, 1st April - Viva Due by this point.

### 2018-05-14 QualifyingReport/QualifyingReport.tex

--- 2018-05-10 QualifyingReport/QualifyingReport.tex
@@
 
-	\subsection{May 2018}
-	Wednesday, 9th May - Alexa Skills Training (All Day)
 	\subsection{June 2018}
@@
 
+	%Complete appendices for each of the assessment. How do you do appendices in LaTeX
+
 	\bibliographystyle{unsrt}
 	\bibliography{QualifyingReport}
+
 \end{document}

### 2018-05-14 QualifyingReport/QualifyingReport.tex
fixed conflicts
added more searches for literature review

@@
 Thus a search for papers was conducted in both the psychology and computer science domains, as well as papers which intersected these areas and were relevant to this particular problem.
+	\newline
+	SEARCH - ProQuest (PsychArticles) - Search Terms (Language AND Decline AND DEMENTIA) - 1484 Results
+	SEARCH - ProQuest (PsychArticles) - Search Terms (Language AND Decline AND DEMENTIA AND Speech) - 486 Results
+	SEARCH - Web Of Science - Search Terms (Language AND Decline AND DEMENTIA AND Speech) - 151 Results
+	SEARCH - Web of Science - Search Terms (Language AND Decline AND DEMENTIA) - 1207 Results
+	SEARCH - Scopus - Search Terms (Language AND Decline AND DEMENTIA AND Speech) - 91 Results 
+	SEARCH - Scopus - Search Terms (Languge AND DEMENTIA AND Decline) - 791 Results
 
@@
 Finally, we talked about the the process of a literature review in the context of the written assignment.
+	\newline
+	\par
+	Test
 	\newline

### 2018-05-14 QualifyingReport/QualifyingReport.tex
updated QR to contain more detail on IS4001

@@
 \title{Qualifying Report}
-\date{2018-05-10}
+\date{2018-05-22}
 \author{Jomar Alcantara}
@@
 	SEARCH - Scopus - Search Terms (Languge AND DEMENTIA AND Decline) - 791 Results
+	\newline
+	Studies that met the following criteria were included: articles that addressed the association between language and all forms of dementia.
+Studies with one or more of the following characteristics were not included in this literature review: articles and dissertations that did not meet the selected inclusion in different databases and those that were not available in it's original form.
 
@@
 We then went on to look at what skills, qualities and attitudes are necessary to be a good researcher including the idea of what consitutes ethical research and finally we looked at the obstacles to progress on a PhD and what we can do as students.
-	\newline
-	\par
 In the afternoon, we looked at how a way of conducting a literature review.
@@
 	\par
-	Test
+	Day 2 consisted of a presentation on Effective Presentation skills, what consitutes a perfect poster presentation, research project planning and management and how to work effectively in research teams.
+In the presentation on effective presentation skills we talked about different ways in which we could plan a presentation and some of the key concepts in terms of communication and engaging the audience.
+We also talked about different presentation styles.
+This was with a view to thinking about our presentation assessment.
+We then went on to talk about what constitutes a poster presentation, and what the general features and conventions of poster presentations were.
+We then went through some examples of what made poster presentations good and bad.
+In the afternoon we went through how to plan and manage a research project effectively.
+Our goals were to see our research projects as a series of tasks which needed to be completed in a sequence, and to be able to plan this in a way that makes good use of time and energy as well a mechanism for keeping track of progress.
+We used the Association of Project Managers (APM) protocol for managing a project and we looked at the notion of having a Gantt chart to have a graphical representation of research project.
+We then talked about these processes in relation to the research plan which needs to be produced for assessment.
+Finally we talked about how to work and communicate effectively in teams.
+We talked about how to work effectively within a team, and specified the difference between groups and teams.
+We talked about the different roles that people have within teams, and the idea that all teams have individuals that take on these distinct roles.
+We talked about how to identify our own-strengths within and how we might facilitate good teamwork withing a research team using our strengths.
+Finally we talked about some of the the results of poor communication within a team.
+	\newline
+	\par
+	The assessments for this module were a 3,000 word literature review with a focus on process rather than content (See Appendix A), a ten minute presentation based on my research project (See Appendix B) and the submission of a project plan spanning the first year of my research(See Appendix C).
 	\newline

### 2018-05-14 QualifyingReport/QualifyingReport.tex
Updated Qualifying Report - 14-05-2018

@@
 	\newline
-	SEARCH - ProQuest (PsychArticles) - Search Terms (Language AND Decline AND DEMENTIA) - 1484 Results
-	SEARCH - ProQuest (PsychArticles) - Search Terms (Language AND Decline AND DEMENTIA AND Speech) - 486 Results
-	SEARCH - Web Of Science - Search Terms (Language AND Decline AND DEMENTIA AND Speech) - 151 Results
-	SEARCH - Web of Science - Search Terms (Language AND Decline AND DEMENTIA) - 1207 Results
-	SEARCH - Scopus - Search Terms (Language AND Decline AND DEMENTIA AND Speech) - 91 Results 
+	SEARCH - ProQuest (PsychArticles) - Search Terms (Language AND Decline AND DEMENTIA) - 1484 Results \newline
+	SEARCH - ProQuest (PsychArticles) - Search Terms (Language AND Decline AND DEMENTIA AND Speech) - 486 Results \newline
+	SEARCH - Web Of Science - Search Terms (Language AND Decline AND DEMENTIA AND Speech) - 151 Results\newline
+	SEARCH - Web of Science - Search Terms (Language AND Decline AND DEMENTIA) - 1207 Results\newline
+	SEARCH - Scopus - Search Terms (Language AND Decline AND DEMENTIA AND Speech) - 91 Results\newline 
 	SEARCH - Scopus - Search Terms (Languge AND DEMENTIA AND Decline) - 791 Results
@@
 
+	\appendix
+	\section{Title of Appendix A}
+
+	Text of Appendix A is Here
+
+	\section{Title of Appendix B}
+
+	Text of Appendix B is Here
+
 \end{document}

### 2018-05-22 Python/LPTHW/ex1.py
added Python practice code

@@ -1,0 +1,7 @@
+print("Hello World!")
+print("Hello Again")
+print("I like typing this.")
+print("This is fun.")
+print("Yay! Printing.")
+print("I'd rather you 'not'.")
+print('I "said" do not touch this.')

### 2018-05-22 Python/LPTHW/ex3.py

@@ -1,0 +1,23 @@
+print("I will now count my chickens:")
+
+print("Hens", 25 + 30 / 6)
+print("Roosters", 100 - 25 * 3 % 4)
+
+print("I will now count the eggs:")
+
+print(3 + 2 + 1 - 5 + 4 % 2 - 1 / 4 + 6)
+
+print("Is it true that 3 + 2 < 5 - 7?")
+
+print(3 + 2 < 5 - 7)
+
+print("What is 3 + 2?", 3 + 2)
+print("What is 5 - 7?", 5 - 7)
+
+print("Oh that's why it's False.")
+
+print("How about some more.")
+
+print("Is it greater?", 5 > -2)
+print("Is it greater or equal?", 5 >= -2)
+print("Is it less or equal?", 5 <=2)

### 2018-05-22 Python/LPTHW/ex4.py

@@ -1,0 +1,17 @@
+cars = 100
+space_in_a_car = 4.0
+drivers = 30
+passengers = 90
+cars_not_driven = cars - drivers
+cars_driven = drivers
+carpool_capacity = cars_driven * space_in_a_car
+average_passengers_per_car = passengers / cars_driven
+
+
+print("There are", cars, "cars available.")
+print("Thera are only", drivers, "drivers available.")
+print("There will be", cars_not_driven, "empty cars today.")
+print("We can transport", carpool_capacity, "people today.")
+print("We have", passengers, "to carpool today.")
+print("We need to put about", average_passengers_per_car,
+      "in each car.")

### 2018-05-22 Python/LPTHW/ex5.py

@@ -1,0 +1,18 @@
+my_name = 'Zed A. Shaw'
+my_age = 35 # not a lie
+my_height = 74 # inches
+my_weight = 180 # lbs
+my_eyes = 'Blue'
+my_teeth = 'White'
+my_hair = 'Brown'
+
+print(f"Let's talk about {my_name}.")
+print(f"He's {my_height} inches tall.")
+print(f"He's {my_weight} pounds heavy.")
+print("Actually that's not too heavy.")
+print(f"He's got {my_eyes} eyes and {my_hair} hair.")
+print(f"His hair are usually {my_teeth} depending on the coffee.")
+
+# this line is tricky, try to get it exactly right.
+total = my_age + my_height + my_weight
+print(f"If I add {my_age}, {my_height}, and {my_weight} I get {total}.")

### 2018-05-22 Python/LPTHW/ex6.py

@@ -1,0 +1,23 @@
+types_of_people = 10
+x = f"There are {types_of_people} types of people."
+
+binary = "binary"
+do_not = "don't"
+y = f"Those who know {binary} and those who {do_not}."
+
+print(x)
+print(y)
+
+print(f"I said: {x} ")
+print(f"I also said: '{y}'")
+
+hilarious = False
+joke_evaluation = "Isn't that joke so funny?! {}"
+
+print(joke_evaluation.format(hilarious))
+
+w = "This is the left side of..."
+e = "a string with a right side."
+
+print(w + e)
+

### 2018-05-22 Python/Sentdex/linearRegression.py

@@ -1,0 +1,29 @@
+from statistics import mean
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib import style
+style.use('fivethirtyeight')
+
+
+xs = np.array([1, 2, 3, 4, 5, 6], dtype=np.float)
+ys = np.array([5, 4, 6, 5, 6, 7], dtype=np.float)
+
+def best_fit_slope_and_intercept(xs, ys):
+    m = ( ((mean(xs) * mean(ys)) - mean(xs*ys)) / 
+          ((mean(xs) * mean(xs)) - mean(xs*xs)))
+    b = mean(ys) - m * mean(xs)
+    return m, b
+
+
+m, b = best_fit_slope_and_intercept(xs, ys)
+print(m, b)
+
+regression_line = [(m*x)+b for x in xs]
+
+predict_x = 8
+predict_y = (m*predict_x+b)
+
+plt.scatter(xs, ys)
+plt.scatter(predict_x, predict_y, color='g')
+plt.plot(xs, regression_line)
+plt.show()

### 2018-05-22 Python/Sentdex/regression1.py

@@ -1,0 +1,73 @@
+# Load Libraries
+import pandas as pd
+import numpy as np
+import math, datetime
+import quandl
+from sklearn import preprocessing, cross_validation, svm
+from sklearn.linear_model import LinearRegression
+import matplotlib.pyplot as plt
+from matplotlib import style
+import pickle
+
+style.use('ggplot')
+
+# Get Data
+df = quandl.get('WIKI/GOOGL')
+
+# Data Filtering and Adding Features
+df = df[['Adj. Open', 'Adj. High','Adj. Low','Adj. Close','Adj. Volume',]]
+df['HL_PCT'] = (df['Adj. High'] - df['Adj. Close']) / df['Adj. Close'] * 100.0
+df['PCT_change'] = (df['Adj. Close'] - df['Adj. Open']) / df['Adj. Open'] * 100.0
+df = df[['Adj. Close', 'HL_PCT', 'PCT_change', 'Adj. Volume']]
+
+forecast_col = 'Adj. Close'
+df.fillna(-99999, inplace=True)
+
+forecast_out = int(math.ceil(0.01*len(df)))
+
+df['label'] = df[forecast_col].shift(-forecast_out)
+
+X = np.array(df.drop(['label'], 1))
+X = preprocessing.scale(X)
+X = X[:-forecast_out]
+X_lately = X[-forecast_out:]
+
+df.dropna(inplace=True)
+y = np.array(df['label'])
+
+X_train, X_test, y_train, y_test = cross_validation.train_test_split(X, y, test_size=0.2)
+
+# clf = LinearRegression(n_jobs=-1)
+# clf.fit(X_train, y_train)
+# with open('linearregression.pickle', 'wb') as f:
+#     pickle.dump(clf, f)
+
+pickle_in = open('linearregression.pickle', 'rb')
+clf=pickle.load(pickle_in)
+
+accuracy = clf.score(X_test, y_test)
+
+#print(accuracy)
+
+forecast_set = clf.predict(X_lately)
+
+print(forecast_set, accuracy, forecast_out)
+
+df['Forecast'] = np.nan
+
+last_date = df.iloc[-1].name
+last_unix = last_date.timestamp()
+one_day = 86400
+next_unix = last_unix + one_day
+
+for i in forecast_set:
+    next_date = datetime.datetime.fromtimestamp(next_unix)
+    next_unix += one_day
[…]

### 2018-06-07 LiteratureReview/LiteratureReview3000.tex
made changes to literature review

@@
 \title{Literature Review}
-\date{2018-04-26}
+\date{2018-06-06}
 \author{Jomar Alcantara}
@@
 \newline
-\par
-The purpose of this review is to seek to understand how dementia impacts speech and language production from a psychological and linguistic perspective with a view to potentially using natural language processing and technology to aid in the early diagnosis of dementia and/or Mild Cognitive Impairment.
+
+According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with Mild dementia suffer from noticeable word-find difficulty.
+They may substitute general terms for more specific terms.
+They may avoid the use of specific names of acquaintances.
+There may be grammatical errors involving subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc.
+Those who have progressed from Mild to Major depression also have difficulties with expressive or receptive language.
+They will often use general-use phrases such as "that thing" and "you know what I mean," and prefers general pronouns rather than names.
+With severe impairment, sufferers may not even recall names of closer friends and family.
+Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur.
+Stereotypy of speech occurs; echolalia and automatic speech typically precede mutism.
 \newline
 \par
-A search of the literature was conducted mainly from a psychological perspective as this would inform the theoretical side of things.
-Next, a search of the literature from a Computer Science perspective was also made with a view to exploring the how domain of machine learning and more specifically natural language processing can be applied to this problem.
-Finally a search was carried out to see what work has already been done in this area.
+The purpose of this review is to seek to understand how dementia impacts speech and language production from a psychological and linguistic perspective.
+A search of the literature was conducted mainly from a psychological perspective using ProQuest (PsychArticles), SCOPUS, Web of Science.
+The following results were found.
+
+\begin{center}
+	\begin{tabular}{ | l | l | l | p{3cm} |}
+		\hline
+		Database & Number of Results & Search Terms  \\ \hline
+		ProQuest(PsychArticles) & 1484 Results & Language AND Decline AND Dementia \\ \hline
+		ProQuest(PsychArticles) & 486 Results  & Language AND Decline AND Dementia AND Speech \\ \hline
+		Web of Science & 1207 Results  & Language AND Decline AND Dementia   \\ \hline
+		Web of Science & 151 Results  & Language AND Decline AND Dementia AND Speech  \\ \hline
+		Scopus & 791 Results & Language AND Decline AND Dementia  \\ \hline
+		Scopus & 91 Results & Language AND Decline AND Dementia AND Speech   \\ 
+		\hline
+	\end{tabular}
+\end{center}
 
 All stages of this literature review were completed by myself.
-
-\section{Description of Dementia}
-
-Language - DSM-V
-'Major' Has significant difficulties with expressive or receptive language.
-Often uses general-use phrases such as "that thing" and "you know what I mean," and prefers general pronouns rather than names.
-With severe impairment, may not even recall names of closer friends and family.
-Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur.
-Stereotypy of speech occurs; echolalia and automatic speech typically precede mutism.
-
-'Mild' Has noticeable word-finding difficulty.
-May substitute general for specific terms.
-May avoid use of specific names of acquaintances.
-Grammatical errors involve subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc.
-
-\subsection{Alzheimer's Disease}
-\subsection{Vascular Dementia}
-\subsection{Dementia with Lewy Bodies}
-\subsection{Parkinson's disease dementia}
-\subsection{Fronto-temporal dementia}
-\subsection{Primary Progressive Aphasia}
-\subsection{Mild Cognitive Impairment}
 
@@
 
-Emery \cite{Emery2000}, states that you can see language decline as Hierarchical.
+Emery (2000)  \cite{Emery2000}, takes the perspective of semiotics, and the idea that language and the processes involved in language are hierarchical.
+She proposes that language goes from simple units of construction, and build layers of complexity and sophistication.
+There are four levels of language according to this perspective: Phonology, Morphology, Syntax and Semantics.
+Emery conducted a review of research that looked at how AD impacted on each level of language.
+She found that people with AD generally had intact Phonology and Morphology but more impaired Syntax and Semantics.
+
+Emery states that you can see language decline as Hierarchical.
 According to a semiotic model there are four ranks of language based on their cognitive complexity.
@@
 
-A common way for clinicians' to diagnose dementia is through use of the Mini-Mental State Examination.
-The Mini-Mental State Examination (MMSE) has shown to be particularly effective at differentiating betweeen Dementia and other psychological disorders such as clinical depression, schitzophrenia and personality disorder and is currently the most widely used measure of diagnosing Dementia within clinical psychology.
-However, it is not without it's faults...
-
-\section{Current ways of diagnosing Dementia using Language}
-\subsection{Free Cued Selective Recall Test}
-The Free Cued Selective Recall Test (FCSRT) was created by Grober and his team \cite{Grober1988}.
-This is a test that primarily tests memory but also tests a subjects ability to name objects.
-Four items are presented at a time on a card and each participant is asked to name and recall with and without cues.
-
-\subsection{Mini-Mental State Examination}
-
-Battery of tests to differentiate between Controls, those with MCI, those with AD. 
-\begin{itemize}
-	\item FCSRDT - As most effective 
-	\item Mini Mental State Examination (MMSE)
-	\item Addenbrooke's Cognitive Examination (ACE-III)
-	\item Dementia Rating Scale (DRS)
-	\item Western Aphasia Battery (WAB)
-	\item Boston Diagnostic Aphasia Examination (BDAE)
-	\item Boston Naming Test (BNT)
-\end{itemize}
-
-\section{Existing studies that have used Natural Language Processing to help in diagnosing psychological difficulties, particularly language difficulties}
-
-\subsection{Nun Study - Snowdon (1996)}
-\subsection{Comparing Presidents Speeches - Berisha et al (2015)}
-
-\section{State of literature into Natural Language processing techniques}
+\section{Connected Speech}
 
@@
 Annotated Bibliography for the project "Ambient Monitoring of Speech for Evidence of Cognitive Decline".
-
-Orimaye, S. O., Wong, J. S-M., Golden, K. J., Wong, C. P., & Soyiri, I. N. (2017) Predicting probably Alzheimer's Disease using linguistic deficits and biomarkers.
-
+\newline
+\par 
 There is an increasing body of literature which supports the use of machine learning to analyse speech to evidence cognitive decline.
@@
 Unique to this paper is the use of n-gram features, which could be explored in the proposed project.
-
-Asgari, M., Kaye, J., & Dodge, H. (2017) Predicting mild cognitive impairment from spontaneous spoken utterances. Alzheimers & Dementia: Translational Research & Clinical Interventions 3, 219 - 228
-
+\newline
+\par
 Asgari, Kaye and Dodge (2017) also looked at the linguistic characteristics of older adults with mild cognitive impairment (MCI) vs healthy controls where they hypothesised that they would be able to predict those with MCI, a distinguishing characteristic of Alzheimer's disease and other variations of dementia.
@@
 They then applied support vector machines (SVM's) and random forest classification algorithms to investigate whether machine learning could be used to distinguish between those with MCI and healthy controls.
-They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84%.
+They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84.
 The authors report that this method is highly reliant on high-fidelity transcription of the conversations which is labour intensive, but they anticipate that technology is advancing sufficiently quickly that automated high-quality transcriptions are possible in the near future.
@@
 The challenges both Orimaye et al and these authors experienced were around the automation of transcription and this would potentially be an area to explore in the proposed project.
-
-Pennebaker, J.W., Boyd, R. L., Jordan, K and Blackburn, K., (2015) The development and psychometric properties of LIWC2015
-
 Asgari et al (2017) used the Linguistic Inquiry and Word Count (LIWC) methodology, developed by Pennebaker et al (2007) to generate features for their paper.
@@
 The LIWC is designed to categorize and evaluate the various emotional, cognitive and structural components which are present within samples of speech and/or written text.
-This method, Linquistic Inquiry & Word Count (LIWC) has evolved from it's initial incarnation in 1993 to it's last update in 2015.
+This method, Linquistic Inquiry and Word Count (LIWC) has evolved from it's initial incarnation in 1993 to it's last update in 2015.
 The premise is the use of words from particular categories provide an insight into the psychological processes and/or diagnoses an individual has.
@@
 This paper is useful as it provides a useful and validated framework from which the proposed project can potentially derive language features.
-
-Schwartz, A.
-H., Eichstaedt, J., Kern, M.
-L., Park, G., Sap.
-M., Stillwell.
-D., .
-.
-.
-Ungar.
-L.
-(2014)  Towards Assessing Changes in Degree of Depression through Facebook.
-Workshop on Computational Linguistics and Clinical Psychology: From Linguistic Signal to Clinical Reality, 118-125
+\newline
+\par
 Currently formal diagnosis of depression is expensive and time-consuming involving both the use of questionnaires and the use of a trained professional to assess an individual.
@@
 \bibliographystyle{unsrt}
-\bibliography{LiteratureReview}
+\bibliography{LiteratureReview3000}
 \end {document}

### 2018-06-07 Python/nltk/nltk_ch1.py

@@ -166,3 +166,7 @@ print(fdist1.most_common(50))
 print(fdist1["whale"])
 
 fdist1.plot(50, cumulative=True)
+
+V = set(text1)
+long_words = [w for w in V if len(w) > 15]
+print(sorted(long_words))

### 2018-06-07 LiteratureReview/LiteratureReview3000.tex
added more to literature review

@@
 \newline
-
+\par
 According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with Mild dementia suffer from noticeable word-find difficulty.
@@
 Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur.
-Stereotypy of speech occurs; echolalia and automatic speech typically precede mutism.
-\newline
+Stereotypy of speech occurs; echolalia and automatic speech typically precede mutism.\newline
 \par
+% In the paragraph below, cite Goodglass and Kaplan (1983) and/or Goodglass (2000)
+In assessing whether a client has AD, part of a clinicans job is to assess langugage capability.
+As described above, AD has a significant impact on language and a clinician has access to tools designed to test lanaguage capabilities.
+An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{} .
+In this task participants are asked to describe a picture presented to them in as much detail as possible.
+The picture itself depicts a familiar domestic scene and would not require participants to use any vocabulary beyond that learned in childhood.
+It was originally designed to assess Aphasia, but has shown itself to be useful in the diagnosis of AD as well \cite{GilesPattersonHodges1995}\newline
+\par
+% This paragraph is to look at the development of Machine Learning algos
+Test\newline
+\par
+It is clear from both the clinical diagnostic criteria and supporting research that language is impacted in those with AD.
 The purpose of this review is to seek to understand how dementia impacts speech and language production from a psychological and linguistic perspective.
-A search of the literature was conducted mainly from a psychological perspective using ProQuest (PsychArticles), SCOPUS, Web of Science.
+An understanding of this would allow the development of  A search of the literature was conducted mainly from a psychological perspective using ProQuest (PsychArticles), SCOPUS, Web of Science.
 The following results were found.

### 2018-06-27 supervisionMeetingOne/supervisionNotes.tex
added supervisor meeting

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Document Class
\documentclass[a4paper]{article}
\title{Supervision Meeting Notes}
\date{2018-06-27}
\author{Jomar Alcantara}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin {document}
\maketitle
\section{My Progress so far}
In the past two months, I have attended one training course on the development of applications for the Amazon Echo. This was a beginners guide to using the Amazon Echo and it's API as a tool for development. I have also attended the research module IS4001. This is part of the introduction to completing a Ph.D and consists of a 3000 Literature Review, 10 minute presentation and Organisational Plan. Finally, I attended the Alzheimer's Society Annual Conference in London. This allowed me to get an idea of the context in which my research is placed. It also allows me to understand what research is taking place in this area. At this conference I encountered ESCAPE (Nicholas Firth, See below) and this tool allows me to capture the transcribed speech of everything captured by the Amazon Echo. \newline
\par 
I have started a trial of the Amazon Echo tool (Echo SCraper and ClAssifier of Persons) using the Amazon Echo in my own home. The goal is to test the effectiveness and accuracy of the software at tracking what I'm saying, as well as looking at the richness of the text generated by my use. I'm going to do this over the course of a month. \newline
\par 
As such I have conceptualised a datapipe line. This is currently based on the assumption that I will be using the Amazon Echo to collect data. Firstly, I will use ESCAPE to capture and transcribe the speech and output to a data file. I will then split each interaction into sentences and run this through the stanford parser to tag for various features. This will then be converted into features that can then be used in a Machine Learning algorithm to classify AD or Not AD. \newline
\par
I have collected a number of datasets, most notably the DementiaBank corpus of cognitive interviews. This has a total of xxx Participants with Dementia as well as controls who complete the Boston Cookie Theft exercise. I have access to other data such as the Presidents Speeches and Conferences archive. Currently I'm in the process of detagging these interviews as these are marked up in a different format. \newline
\par
I have conducted literature searches using Web of Science and Scopus which have highlighted a number of papers to read in the relevant areas. I have collected a number of these papers to read through which will contribute to my literature review. Periodically I conduct new searches and add to these papers. 
\section{What I intend to do over the next two months - end of August}
In the next two months I plan to complete and submit an initial literature review as well an initial experimental idea formalised and finally I would like to submit a first work-in-progress draft of the Qualifying Report. I also intend to look into NHS ethics
\section{Further down the line}
It's my intention to attend the 2019 Alzheimer's Society Conference and as part of this I would like to submit a paper/poster presentation detailing my research so far. In addition, I will be attending the Natural Language Processing conference in Brussels (October / November 2018).

\section{Questions and thoughts}
Question - Is there any training you can recommend on Natural Language Processing? Particularly the Stanford Parser.
\newline
Question - The literature points to voice patterns as an area to explore, is this necessary?
\newline
Thought - What if the language used in interacting with the Amazon Echo is not diverse enough? Test to see with my data? Vs DementiaBank
\newline

\end {document}

### 2018-06-27 Python/nltk/nltk_ch1.py

@@ -1,4 +1,6 @@
 from nltk.book import *
+import nltk
+
 
 # concordance shows every occurance of the word passed to the function as well as some context.
 text1.concordance("monstrous")
@@ -170,3 +172,13 @@ fdist1.plot(50, cumulative=True)
 V = set(text1)
 long_words = [w for w in V if len(w) > 15]
 print(sorted(long_words))
+
+fdist5 = FreqDist(text5)
+print(sorted(w for w in set(text5) if len(w) > 7 and fdist5[w] > 7))
+print(list(nltk.bigrams(['more', 'is', 'said', 'than', 'done'])))
+
+print(text4.collocations())
+print(text6.collocations())
+print(text8.collocations())
+
+print([len(w) for w in text1])

### 2018-06-27 Python/nltk/nltk_ch1_ex.py

@@ -1,0 +1,76 @@
+import nltk
+from nltk.book import *
+
+    ☼ Try using the Python interpreter as a calculator, and typing expressions like 12 / (4 + 1).
+
+    ☼ Given an alphabet of 26 letters, there are 26 to the power 10, or 26 ** 10, ten-letter strings we can form. That works out to 141167095653376. How many hundred-letter strings are possible?
+
+    ☼ The Python multiplication operation can be applied to lists. What happens when you type ['Monty', 'Python'] * 20, or 3 * sent1?
+
+    ☼ Review 1 on computing with language. How many words are there in text2? How many distinct words are there?
+
+    ☼ Compare the lexical diversity scores for humor and romance fiction in 1.1. Which genre is more lexically diverse?
+
+    ☼ Produce a dispersion plot of the four main protagonists in Sense and Sensibility: Elinor, Marianne, Edward, and Willoughby. What can you observe about the different roles played by the males and females in this novel? Can you identify the couples?
+
+    ☼ Find the collocations in text5.
+
+    ☼ Consider the following Python expression: len(set(text4)). State the purpose of this expression. Describe the two steps involved in performing this computation.
+
+    ☼ Review 2 on lists and strings.
+        Define a string and assign it to a variable, e.g., my_string = 'My String' (but put something more interesting in the string). Print the contents of this variable in two ways, first by simply typing the variable name and pressing enter, then by using the print statement.
+        Try adding the string to itself using my_string + my_string, or multiplying it by a number, e.g., my_string * 3. Notice that the strings are joined together without any spaces. How could you fix this?
+
+    ☼ Define a variable my_sent to be a list of words, using the syntax my_sent = ["My", "sent"] (but with your own words, or a favorite saying).
+        Use ' '.join(my_sent) to convert this into a string.
+        Use split() to split the string back into the list form you had to start with.
+
+    ☼ Define several variables containing lists of words, e.g., phrase1, phrase2, and so on. Join them together in various combinations (using the plus operator) to form whole sentences. What is the relationship between len(phrase1 + phrase2) and len(phrase1) + len(phrase2)?
+
+    ☼ Consider the following two expressions, which have the same value. Which one will typically be more relevant in NLP? Why?
+        "Monty Python"[6:12]
+        ["Monty", "Python"][1]
+
+    ☼ We have seen how to represent a sentence as a list of words, where each word is a sequence of characters. What does sent1[2][2] do? Why? Experiment with other index values.
+
[…]

### 2018-07-02 supervisionMeetingOne/supervisionNotes.tex
added new files

@@
 I have collected a number of datasets, most notably the DementiaBank corpus of cognitive interviews.
-This has a total of xxx Participants with Dementia as well as controls who complete the Boston Cookie Theft exercise.
+This has a total of 240 Participants with Dementia as well as controls who complete the Boston Cookie Theft exercise.
 I have access to other data such as the Presidents Speeches and Conferences archive.
@@
 I have collected a number of these papers to read through which will contribute to my literature review.
-Periodically I conduct new searches and add to these papers.
+Periodically I conduct new searches and add to these papers and I have searches set up to automatically update me on papers.
 \section{What I intend to do over the next two months - end of August}
 In the next two months I plan to complete and submit an initial literature review as well an initial experimental idea formalised and finally I would like to submit a first work-in-progress draft of the Qualifying Report.
-I also intend to look into NHS ethics
+I also intend to look into the NHS ethics procedure as it's likely that I will need to be able to submit this before
+\newline
+I would like to look into how to measure the complexity of a sentence as I think a predictor could be how complex a sentence is.
+This certainly fits in with the idea that generating sentences increases cognitive load.
+Are the sentences required by Alexa to activate commands complex enough?
+\newline
+I would like to explore further the idea of Deep Learning for Natural Language Processing - the idea of pre-processing (One-Hot Encoding) vs vector representations.
 \section{Further down the line}
@@
 \newline
-Question - The literature points to voice patterns as an area to explore, is this necessary?
+Question - The literature points to voice patterns as an area to explore, is this necessary? How might this work? Pauses might be useful to measure?
 \newline
@@
 \newline
-
+Thought - Is there a stage before collecting data with Amazon Echo where we just look at natural language in terms of speech between controls and those with dementia.
+Cookie Theft Picture + Semi Structured Interview?
 \end {document}

### 2018-07-05 LiteratureReview/LiteratureReview3000.tex
added supervisor meeting
updated literature review

@@
 \section{Introduction}
-
 \par
-According to a recent report commissioned by the Alzheimer's Society, they estimate the prevalence of Dementia in the UK at approximately 815,000 people.
-This represents 1 in 14 of the population aged 65 or over.
+According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of Dementia in the UK at approximately 815,000 people.
+This represents 1 in 14 of the population aged 65 or over and 1 in 79 in the general population.
 This report also estimates an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis.
@@
 \newline
+\par
+Those who are diagnosed with dementia suffer from a number of cogniive, emotional and behavioural problems.
+In terms of cognitive problems, they suffer from deficiencies to memory, planning and organising, speech and language, visuo-spatial tasks and difficulties in orienting in time and place.
+From an emotional perspective, those with dementia can suffer from mental health problems such as depression, anxiety and irritability.
+\newline
+\par
+There has been some work in the use of machine learning (ML) and natural language processing for looking at mental health difficulties.\newline
 \par

### 2018-07-09 LiteratureReview/LiteratureReview3000.tex
Updated literature review and added project plan template

@@
 \title{Literature Review}
-\date{2018-06-06}
+\date{2018-07-13}
 \author{Jomar Alcantara}
@@
 \par
-There has been some work in the use of machine learning (ML) and natural language processing for looking at mental health difficulties.\newline
-\par
 According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with Mild dementia suffer from noticeable word-find difficulty.
@@
 \par
-% This paragraph is to look at the development of Machine Learning algos
-Test\newline
+The study of language decline in Dementia is not a new field of study.
+The focus of this work has been in understanding in which areas of language are impacted by dementia.
+Emery \cite{Emery2000} completed a literature review in this area and structured her review into four levels of language.
+There are four levels of language according to this perspective: Phonology, Morphology, Syntax and Semantics and her review looked at the idea that language and the processes involved in language are hierarchical.
+She proposes that language goes from simple units of construction, and build layers of complexity and sophistication.
+She found that people with AD generally had intact Phonology and Morphology but more impaired Syntax and Semantics.
+Emery states that you can see language decline as Hierarchical.
+According to a semiotic model there are four ranks of language based on their cognitive complexity.
+These ranks are Phonology, Morphology, Syntax and Semantics.
+She conducted a review studies which looked at language decline from the perspective of these hierarchical ranks.
+Her conclusion is that language decline is related to the complexity of the language task given to a participant and that language decline is hierarchical in that the language forms we learn last (the most complex language forms) are the first to deteriorate.\newline
+\par
+A lot of research since Emery's literature review \cite{Emery2000} have looked supported her ideas. \newline
 \par
 It is clear from both the clinical diagnostic criteria and supporting research that language is impacted in those with AD.
-The purpose of this review is to seek to understand how dementia impacts speech and language production from a psychological and linguistic perspective.
-An understanding of this would allow the development of  A search of the literature was conducted mainly from a psychological perspective using ProQuest (PsychArticles), SCOPUS, Web of Science.
-The following results were found.
+However, one of the costs of analysing in this way is the idea that there is a huge burden on trained practitioners, be it clinical psychologists, audio transcribers and text encoders, to facilitate the process of collecting data and analysis.
+The field of machine learning and natural language processing could be used to improve the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline.
+The purpose of this review is to seek to understand what developments have been made in the use of Machine Learning and Natural Language processing to aid the diagnosis of Mild Cognitive Impairment (MCI) and Alzheimer's Disease (AD).
+A search of the literature was conducted mainly from a psychological perspective using ProQuest (PsychArticles), SCOPUS, Web of Science.
+The following results were found (Table 1).
+All papers were the reviewed for relevance by reading the abstract and full text where appropriate.
+Papers were included where there was an attempt to classify participants as MCI, AD or Healthy using speech and/or language by using Machine Learning techniques.
+We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English.
+This resulted in 21 journal articles and/or conference papers which form this review.
 
@@
 		ProQuest(PsychArticles) & 486 Results  & Language AND Decline AND Dementia AND Speech \\ \hline
+		ProQuest(PsychArticles) & 159 Results & Machine Learning AND Dementia AND Language \\ \hline
 		Web of Science & 1207 Results  & Language AND Decline AND Dementia   \\ \hline
 		Web of Science & 151 Results  & Language AND Decline AND Dementia AND Speech  \\ \hline
+		Web of Science & 34 Results & Machine Learning AND Dementia AND Language \\ \hline
 		Scopus & 791 Results & Language AND Decline AND Dementia  \\ \hline
-		Scopus & 91 Results & Language AND Decline AND Dementia AND Speech   \\ 
-		\hline
+		Scopus & 91 Results & Language AND Decline AND Dementia AND Speech   \\ \hline
+		Scopus & 29 Results & Machine Learning AND Dementia AND Langauge \\ \hline
+
 	\end{tabular}
@@
 
-All stages of this literature review were completed by myself.
-
-\section{Assessment of Language function in Dementia}
-
-Language can be defined as the ability to encode ideas into words and/or symbols for the means of communication.
-Difficulty in language, both spoken and written, are often described as symptoms of various types of Alzheimers' disease (AD).
-One of the challenges when using speech as a predictor is recognising whether there is a problem with language production or with the motor skills.
-Impairments in speech that arise from any process that disrupts the neuraxis from the cortex to muscle and encompass dysarthria (disturbance in articulation) and dysphonia (disturbance in the production in vocal sounds).
-However it is important to use language as a tool to aid the diagnosis of Dementia, as it provides vital clues to aid a clinician in differentiating in the different types of Dementia, which will in turn aid a clinician in an attempt to manage an individual case of Dementia.
-Due to the subtle nature of the languages changes that are experienced in those who suffer with different types of Dementia, it is often misdiagnosed.
-
-Emery (2000)  \cite{Emery2000}, takes the perspective of semiotics, and the idea that language and the processes involved in language are hierarchical.
-She proposes that language goes from simple units of construction, and build layers of complexity and sophistication.
-There are four levels of language according to this perspective: Phonology, Morphology, Syntax and Semantics.
-Emery conducted a review of research that looked at how AD impacted on each level of language.
-She found that people with AD generally had intact Phonology and Morphology but more impaired Syntax and Semantics.
-
-Emery states that you can see language decline as Hierarchical.
-According to a semiotic model there are four ranks of language based on their cognitive complexity.
-These ranks are Phonology, Morphology, Syntax and Semantics.
-She conducted a review studies which looked at language decline from the perspective of these hierarchical ranks.
-Her conclusion is that language decline is related to the complexity of the language task given to a participant and that language decline is hierarchical in that the language forms we learn last (the most complex language forms) are the first to deteriorate.
-
 \section{Connected Speech}
-
-\section{stuff to add}
-Annotated Bibliography for the project "Ambient Monitoring of Speech for Evidence of Cognitive Decline".
-\newline
 \par 
@@
 This paper is noteworthy as it provides evidence that using machine learning to identify participants with MCI is possible and it provides some ideas of potential features which can be used.
-Unique to this paper is the use of n-gram features, which could be explored in the proposed project.
-\newline
+Unique to this paper is the use of n-gram features, which could be explored in the proposed project.\newline
 \par
@@
 This manual goes summarises the process of constructing the framework as well as discussing the reliability and validity of the measure and a review of studies using the LIWC was conducted by Tausczik and Pennebaker (2010) support the notion that the LIWC is valid across multiple psychological domains.
-This paper is useful as it provides a useful and validated framework from which the proposed project can potentially derive language features.
-\newline
+This paper is useful as it provides a useful and validated framework from which the proposed project can potentially derive language features.\newline
 \par
@@
 This paper also supports the idea that language changes over time and is a marker for deterioration in both those with depression and dementia.
-
+\newline
+\par
+Fraser, Meltzer and Rudzicz (2015) looked at connected speech using the DementiaBank corpus.
+They found that there were four factors which informed the classification of participants as either healthy or AD.
+These four factors were semantic impairment, acoustic abnormality, syntactic impairment and information impairment.\newline
+\par
 \section{Conclusion}
@@
 It has already been shown that analysis of speech and language has shown markers that pre-date the official diagnosis of dementia (Snowdon et al, 1996;)\cite{Berisha2015}.
-
+A significant amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
+Firstly, the vast majority of the research described above looks at pre-existing datasets like the DementiaBank corpus (CITE).
+Whilst this is useful for 'backtesting the data', it would be useful for this process to be tested live to see whether these results can be replicated now.
+\newline
+\par
+A lot of the research was conducted using the idea of Picture Description, and whilst this is a valid measure of testing the linguistic capability of a participant, there are some flaws with this methodology.
+The picture description task
 \bibliographystyle{unsrt}

### 2018-07-09 ProjectPlan/ProjectPlan.tex
Updated literature review and added project plan template

--- 2018-05-14 Workplan/Workplan.tex
@@
-\documentclass{article}
-
-\title{Workplan}
-\date{2018-04-17}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Document Class
+\documentclass[a4paper]{article}
+\title{Project Plan}
+\date{2018-06-06}
 \author{Jomar Alcantara}
 
-\begin{document}
-	\maketitle
-	\section{Interest Statement}
-	I am interested in whether technology can be used to aid the early diagnosis of dementia specifically by analysing speech and language in a natural environment.
-As part of this, I am exploring the use of technologies such as the Amazon Echo or Google Home kit to capture speech and language.
-I am using Natural Language Processing to generate features from the captured speech such that we can categorise people into specific categories and/or predict test results on scores using machine learning.
-	\section{Literature Review}
-	\subsection{Dementia - Context}
-	What is the Problem? 
-	Why is this important?
-	What is the potential impact?
-	\subsection{How do Psychologists assess dementia}
-	What tools do psychologists use to diagnose dementia? 
-	Mini mental state examination (MMSE), 
-	Addenbrooke's Cognitive Examination(ACE-III)
-	Wechsler Adult Intelligence Scale(WAIS)
-	The Free and Cued Selective Reminding Test (FCSRT)
-	\subsection{How do Psychologists understand how language is affected by dementia}
-	What language deficits are there and how are they characterised?
-	What speech deficits are there and how are they characterised?
-	\subsection{How is Natural Language Processing best positioned to help in this area?}
-	\subsection{Are there any other technologies that can be used to help diagnosis?}
-
-	\subsection{What Research has been done in the area so far?}
-	\subsection{Conclusion}
-
-	\section{Stages of research}
-	\section{Stage 1 - Data capture}
-	\subsection{Training needs analysis}
-	\subsection{Psychology}
-	Training in delivering the WAIS or psychological tools that might be needed 
-	\subsection{Programming}
-	\begin{itemize}
-		\item Online courses in Python Programming
-		\item Online courses in Machine Learning
-	\end{itemize}
-
-	\section{Data collection}
-	\subsection{DementiaBank and AphasiaBank}
-	\textbf{Holland Corpus}
-	\newline 
-	Consists of 2 participants a 62 year old man with mild AD and a 78 year old woman with more advanced AD.
-	\newline
-	\textbf{Kempler Corpus}
-	\newline
-	Consists of 6 participants with a range of 65 to 86 (mean, sd)
-	\newline
-	\textbf{Pitt Corpus}
-	\newline
-	Consists of 104 control participants, 208 with dementia and 85 with an unknown diagnosis.
-	\newline
-	\textbf{DePaul Corpus}
-	\newline
-	Consists of 1 participant, 66 year old female with PPA.
-	\newline
-	\textbf{Hopkins Corpus}
-	\newline
-	Consists of 36 participants, 23 males and 13 females
-
-	\subsection {Three Authors, P.D. James, Iris Murdoch and Agatha Christie}	
-	Data to be downloaded
-
-	\subsection{Presidents Speeches}
-	Data to be downloaded
-
-	\subsection{Alexa Skills Kit and Amazon Web Services}
-	\begin{itemize}
-		\item \textbf{Training in Alexa Skills Kit - Booked}
-		\item Training in Amazon Web Services
-	\end{itemize}
-
-	\section{Stage 2 - Data Processing, Feature Construction}
-	\subsection{Natural Language Processing}
-	Online courses in Natural Language Processing
-
-	\begin{itemize}
-		\item Jurafsky and Martin - Speech and Language Processing 
-		\item Manning and Schutze - Statistical NLP 
-		\item Manning and Raghavan - Intro to IR
-	\end{itemize}
-
-	\section{Stage 3 - Data Analysis and Machine Learning}
-	\subsection{Application of Machine Learning Algorithms}
-
-	\section{Proposed Timeline, Key Dates and Deliverables}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin {document}
+\section{Aims and Objectives}
+\section{Tasks}
 	\subsection{April 2018}
@@
 	Monday, 1st April - Viva Due by this point.
-\end{document}
+\section{Obstacles}
+\end {document}

### 2018-07-09 LiteratureReview/LiteratureReview3000.tex
updated literature review and qualifying report - 09/07/2018

@@
 \par
+There are a number of classical studies which look at language decline in those with dementia.
+One of the most famous pieces of research was by Berisha and Liss (2015) \cite{Berisha2015} who examined speeches and public interviews of former US president Ronald Reagan.
+They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease.
+Another \newline
+\par
 A lot of research since Emery's literature review \cite{Emery2000} have looked supported her ideas. \newline
@@
 All papers were the reviewed for relevance by reading the abstract and full text where appropriate.
+An additonal search through references of shortlisted papers was also conducted.
 Papers were included where there was an attempt to classify participants as MCI, AD or Healthy using speech and/or language by using Machine Learning techniques.
@@
 \begin{center}
-	\begin{tabular}{ | l | l | l | p{3cm} |}
+	\begin{tabular}{ | l | c | r | p{1cm} |}
 		\hline
@@
 		Scopus & 29 Results & Machine Learning AND Dementia AND Langauge \\ \hline
-
 	\end{tabular}
@@
 
-\section{Connected Speech}
+\section{Connected Language and potential features}
+Connected Language and it's study can be defined as self generated discourse.
+This is important as this requires a number of cognitive processes which we know are impacted in those who suffer from dementia.
+One of the common problems in the analysis of connected language is the ability to differentiate between what is a purely language related problem vs a memory problem.
+\newline
 \par
-There is an increasing body of literature which supports the use of machine learning to analyse speech to evidence cognitive decline.
-Orimaye et al (2017) investigate the use of machine learning algorithms to detect differences in syntactic, lexical and n-gram linguistic biomarkers to distinguish between those with probably AD and healthy controls, The authors found significant differences in the uses of all three types of biomarkers for those with AD and healthy controls.
+\subsection(n-grams)
+One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
+An n-gram is a contiguous sequence of n items from a given sample of text or speech.
+The items can be phonemes, syllables, letters, words or base pairs according to the application.
+For example, given the sequence of words "to be or not to be", this extract is said to contain six 1-gram sequences (to, be, or, not, to, be), five 2-gram sequences (to be, be or, or not, not to, to be), four 3-gram sequences(to be or, be or not, or not to, not to be) and so on.
+This is useful as, given a large portion of text or speech, we can predict the probability of a word being close by to a given word.
+One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class.
+\newline
+\par 
+Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences in syntactic, lexical and n-gram linguistic biomarkers to distinguish between those with probably AD and healthy controls, The authors found significant differences in the uses of all three types of biomarkers for those with AD and healthy controls.
 Their results show that the top 1000 n-gram features plus the twenty-three syntactic and lexical features was the most successful at differentiating the two groups (0.93 AUC)).
@@
 \par
+Lopez-de-Ipina (2018) have since moved on 
+\section {Types of Language Assessment}
+\subsection{Picture Description Tasks}
+One of the most commonly used tasks to measure lexical analysis is the Picture Description task.
+In dementia research, the most common task picture description task is the Cookie Theft picture.
+This does very well in giving a valid dataset of language but suffers somewhat from ecological validity.
+In essence, when completing this task, the participant is generating speech to describe the picture but does not use some of the skills generally used in every day interactions.
+If the goal is to a way
+\subsection{Other types of task}
+
 \section{Conclusion}
@@
 Identification of language impairment is important in Dementia because it aids diagnosis of specific types of dementia, which in turn can alter the prognosis and change the management of the degenerative disorder.
-As these differences in language are quite subtle, the varying subtypes of dementia are frequently misdiagnosed.
+As these differences in language are quite subtle, the varying subtypes of dementia are frequently misdiagnosed and can sometimes be missed altogether.
 \newline
@@
 \par
-A lot of the research was conducted using the idea of Picture Description, and whilst this is a valid measure of testing the linguistic capability of a participant, there are some flaws with this methodology.
-The picture description task
+Future research should be directed towards developing non-intrusive ways of detecting subtle changes in language based in the home, such that any deterioration that could indicate the presence of MCI or AD and be flagged up for further review via referral.
+Given the amount of processing power and the sophistication of machine learning, in particular deep learning algorithms.
+Further, despite the quality of datasets being used to 'backtest' these algorithms, further research should look at generating further datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which we could claim are limited in scope.
+It would be worth looking at how we can utilse these technologies further to aid detect the earliest signs of cognitive impairment.
+
 \bibliographystyle{unsrt}

### 2018-07-09 ProjectPlan/ProjectPlan.tex
updated literature review and qualifying report - 09/07/2018

@@
 \section{Aims and Objectives}
+According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of Dementia in the UK at approximately 815,000 people.
+This represents 1 in 14 of the population aged 65 or over and 1 in 79 in the general population.
+This report also estimates an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis.
+They also estimate that the overall impact of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually.
+Globally, this picture is a lot bleaker.
+A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillon US dollars.
+\newline
 \section{Tasks}

### 2018-07-10 LiteratureReview/LiteratureReview3000.tex
updated intro LR3k

@@
 \par
+Alzheimer's Disease (AD) is a neurodegenerative disease in which the hallmark clinical symptoms are cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits.
+In addition, these symptoms are often accompanied by emotional difficulties such as depression and irritability and behavioural difficulties.
+\newline
+\par
 According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of Dementia in the UK at approximately 815,000 people.
@@
 \par
-Those who are diagnosed with dementia suffer from a number of cogniive, emotional and behavioural problems.
-In terms of cognitive problems, they suffer from deficiencies to memory, planning and organising, speech and language, visuo-spatial tasks and difficulties in orienting in time and place.
-From an emotional perspective, those with dementia can suffer from mental health problems such as depression, anxiety and irritability.
-\newline
+Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent. \newline
 \par

### 2018-07-10 LiteratureReview/LiteratureReview3000.tex
Updated Literature Review - Introduction Done

@@
 \par
-Alzheimer's Disease (AD) is a neurodegenerative disease in which the hallmark clinical symptoms are cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits.
+Alzheimer's Disease (AD) is a neurodegenerative disease in which the brain develops neurofibrillary tangles and neuritic plaques along with the loss of cortical neurons and synapses.
+The hallmark clinical symptoms are cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits\cite{AmericanPsychiatricAssociation2013}.
 In addition, these symptoms are often accompanied by emotional difficulties such as depression and irritability and behavioural difficulties.
+Whilst there are several variants of Dementia, AD remains the most common type of Dementia and will be the focus of this review.
 \newline
 \par
-According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of Dementia in the UK at approximately 815,000 people.
+According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
 This represents 1 in 14 of the population aged 65 or over and 1 in 79 in the general population.
-This report also estimates an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis.
-They also estimate that the overall impact of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually.
+They estimate an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
 Globally, this picture is a lot bleaker.
-A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
-The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillon US dollars.
+A recent report suggests that in 2015 there were 46 million people with a diagnosis of AD and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
 \newline
 \par
-Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent. \newline
+Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent.
+Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time until a threshold is reached and clinical symptoms become apparent \cite{Nestor2006}.
+If this is the case, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
+Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on the decline in the use of language in those with AD.
+\newline
 \par
-According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with Mild dementia suffer from noticeable word-find difficulty.
-They may substitute general terms for more specific terms.
-They may avoid the use of specific names of acquaintances.
+One of the most famous pieces of research was by Berisha and Liss (2015) \cite{Berisha2015} who examined speeches and public interviews of former US president Ronald Reagan.
+They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease.
+Another classical study by Snowdon et al (1996) \cite{Snowdon1996} looked at whether linguistic ability in early life was asoociated with cognitive function and AD in later life.
+They found that idea density (defined as the number of expressed propositions divided by the number of words) was a key predictor in predicting whether nuns would go on to develop AD in later life.
+They found that those who would go on to develop AD all had low idea density in early life and they found no AD present in those with high idea density.
+\newline
+\par
+The range of language deficits in those who suffer with AD are wide and varied and differ as the disease develops.
+According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with Mild dementia suffer from noticeable word finding difficulty.
+They may substitute general terms for more specific terms and may avoid the use of specific names of acquaintances.
 There may be grammatical errors involving subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc.
 Those who have progressed from Mild to Major depression also have difficulties with expressive or receptive language.
-They will often use general-use phrases such as "that thing" and "you know what I mean," and prefers general pronouns rather than names.
+They will often use general-use phrases such as "that thing" and "you know what I mean" and prefers general pronouns rather than names.
 With severe impairment, sufferers may not even recall names of closer friends and family.
 Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur.
-Stereotypy of speech occurs; echolalia and automatic speech typically precede mutism.\newline
+Echolalia (meaningless repetition of another person's spoken words) and automatic speech typically precede mutism.
+With the wide range of deficits someone with AD can suffer, it makes sense to try to categorise these deficits in some way.\newline
 \par
-% In the paragraph below, cite Goodglass and Kaplan (1983) and/or Goodglass (2000)
-In assessing whether a client has AD, part of a clinicans job is to assess langugage capability.
-As described above, AD has a significant impact on language and a clinician has access to tools designed to test lanaguage capabilities.
-An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{} .
-In this task participants are asked to describe a picture presented to them in as much detail as possible.
-The picture itself depicts a familiar domestic scene and would not require participants to use any vocabulary beyond that learned in childhood.
-It was originally designed to assess Aphasia, but has shown itself to be useful in the diagnosis of AD as well \cite{GilesPattersonHodges1995}\newline
-\par
-The study of language decline in Dementia is not a new field of study.
-The focus of this work has been in understanding in which areas of language are impacted by dementia.
-Emery \cite{Emery2000} completed a literature review in this area and structured her review into four levels of language.
-There are four levels of language according to this perspective: Phonology, Morphology, Syntax and Semantics and her review looked at the idea that language and the processes involved in language are hierarchical.
+Emery \cite{Emery2000} completed a literature review in this area and she looked at four levels of language: Phonology, Morphology, Syntax and Semantics.
+Her review looked at the idea that language and the processes involved in language are hierarchical.
 She proposes that language goes from simple units of construction, and build layers of complexity and sophistication.
 She found that people with AD generally had intact Phonology and Morphology but more impaired Syntax and Semantics.
-Emery states that you can see language decline as Hierarchical.
-According to a semiotic model there are four ranks of language based on their cognitive complexity.
-These ranks are Phonology, Morphology, Syntax and Semantics.
-She conducted a review studies which looked at language decline from the perspective of these hierarchical ranks.
-Her conclusion is that language decline is related to the complexity of the language task given to a participant and that language decline is hierarchical in that the language forms we learn last (the most complex language forms) are the first to deteriorate.\newline
-\par
-There are a number of classical studies which look at language decline in those with dementia.
-One of the most famous pieces of research was by Berisha and Liss (2015) \cite{Berisha2015} who examined speeches and public interviews of former US president Ronald Reagan.
-They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease.
-Another \newline
-\par
-A lot of research since Emery's literature review \cite{Emery2000} have looked supported her ideas. \newline
+Emery concludes that language decline is hierarchical and is related to the complexity of the language task given to a participant.
+She and that language decline is hierarchical in that the language forms we learn last (the most complex language forms) are the first to deteriorate.\newline
 \par
 It is clear from both the clinical diagnostic criteria and supporting research that language is impacted in those with AD.
-However, one of the costs of analysing in this way is the idea that there is a huge burden on trained practitioners, be it clinical psychologists, audio transcribers and text encoders, to facilitate the process of collecting data and analysis.
-The field of machine learning and natural language processing could be used to improve the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline.
-The purpose of this review is to seek to understand what developments have been made in the use of Machine Learning and Natural Language processing to aid the diagnosis of Mild Cognitive Impairment (MCI) and Alzheimer's Disease (AD).
-A search of the literature was conducted mainly from a psychological perspective using ProQuest (PsychArticles), SCOPUS, Web of Science.
+However, one of the costs of analysing language is that there is a huge burden on trained practitioners, be it clinical psychologists, audio transcribers and text encoders to facilitate the process of collecting data and analysis.
+The field of machine learning and natural language processing has been suggested as a way toimprove the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline.
+\newline
+\par  
+The purpose of this review is to seek to understand what developments have been made in the use of Machine Learning and Natural Language processing to aid the diagnosis of Mild Cognitive Impairment (MCI) and AD.
+A search of the literature was conducted using ProQuest (PsychArticles), SCOPUS, Web of Science.
 The following results were found (Table 1).
-All papers were the reviewed for relevance by reading the abstract and full text where appropriate.
+All papers were then reviewed for relevance by reading the abstract and full text where appropriate and a shortlist was compiled.
 An additonal search through references of shortlisted papers was also conducted.
-Papers were included where there was an attempt to classify participants as MCI, AD or Healthy using speech and/or language by using Machine Learning techniques.
+Papers were included where researchers used machine learning to classify participants as MCI, AD or Healthy using language.
 We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English.
@@
 
+\section {Types of Language Assessment}
+One of the key debates when looking at how to analyse language is the type of task provided to elicit language.
+In the literature  researchers have primarily focused on Picture Description tasks but have also suggested other ways in which we might collect data.
+\newline
+\par
+\subsection{Picture Description Tasks}
+One of the most commonly used tasks to measure lexical analysis is the Picture Description task.
+In dementia research, the most common task picture description task is the Cookie Theft picture.
+This does very well in giving a valid dataset of language but suffers somewhat from ecological validity.
+In essence, when completing this task, the participant is generating speech to describe the picture but does not use some of the skills generally used in every day interactions.
+If the goal is to a way
+
+In assessing whether a client has AD, part of a clinicans job is to assess langugage capability.
+As described above, AD has a significant impact on language and a clinician has access to tools designed to test lanaguage capabilities.
+An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Kaplan2010} .
+In this task participants are asked to describe a picture presented to them in as much detail as possible.
+The picture itself depicts a familiar domestic scene and would not require participants to use any vocabulary beyond that learned in childhood.
+It was originally designed to assess Aphasia, but has shown itself to be useful in the diagnosis of AD as well \cite{Giles1996}\newline
+\par
+\subsection{Other types of task}
 \section{Connected Language and potential features}
@@
 Lopez-de-Ipina (2018) have since moved on
-\section {Types of Language Assessment}
-\subsection{Picture Description Tasks}
-One of the most commonly used tasks to measure lexical analysis is the Picture Description task.
-In dementia research, the most common task picture description task is the Cookie Theft picture.
-This does very well in giving a valid dataset of language but suffers somewhat from ecological validity.
-In essence, when completing this task, the participant is generating speech to describe the picture but does not use some of the skills generally used in every day interactions.
-If the goal is to a way
-\subsection{Other types of task}
 

### 2018-07-10 LiteratureReview/LiteratureReview3000.tex
updated LR3k

@@
 However, one of the costs of analysing language is that there is a huge burden on trained practitioners, be it clinical psychologists, audio transcribers and text encoders to facilitate the process of collecting data and analysis.
-The field of machine learning and natural language processing has been suggested as a way toimprove the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline.
+The field of machine learning and natural language processing has been suggested as a way toimprove the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline \cite{Boschi2017}.
 \newline
@@
 \section {Types of Language Assessment}
-One of the key debates when looking at how to analyse language is the type of task provided to elicit language.
+One of the key debates when looking at how to analyse language is the type of task provided to elicit language production in participants.
 In the literature researchers have primarily focused on Picture Description tasks but have also suggested other ways in which we might collect data.
@@
 \subsection{Picture Description Tasks}
-One of the most commonly used tasks to measure lexical analysis is the Picture Description task.
-In dementia research, the most common task picture description task is the Cookie Theft picture.
-This does very well in giving a valid dataset of language but suffers somewhat from ecological validity.
-In essence, when completing this task, the participant is generating speech to describe the picture but does not use some of the skills generally used in every day interactions.
-If the goal is to a way
-
-In assessing whether a client has AD, part of a clinicans job is to assess langugage capability.
-As described above, AD has a significant impact on language and a clinician has access to tools designed to test lanaguage capabilities.
-An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Kaplan2010} .
+One of the most commonly used tasks to measure language is the Picture Description task.
+An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Kaplan2010}.
 In this task participants are asked to describe a picture presented to them in as much detail as possible.
 The picture itself depicts a familiar domestic scene and would not require participants to use any vocabulary beyond that learned in childhood.
-It was originally designed to assess Aphasia, but has shown itself to be useful in the diagnosis of AD as well \cite{Giles1996}\newline
+It was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of AD as well \cite{Giles1996}\newline
 \par
-\subsection{Other types of task}
+The picture description task does a fine job of eliciting descriptive language, however because of the limited content has limited use.
+The task in itself just a descriptive task, and therefore elicits a certain type of language.
+There is some disagreement as to the benefits of this using this methodology.
+This task is reported as being useful to lexico-semantic disorders \cite{Boschi2017, Sajjadi2012} as the language being generated is primarly nouns and deixis (words to identify items and words to put those items into context).
+However, Ash \cite{Ash2012}felt that there was no difference in using this task vs Story Narration (described below).
+In explaining the differences, it is worth noting that these researchers were using differing variables and this could explain their different perspectives.
+\newline
+\par
+In terms of Machine Learning research in this area, a number of researchers have used transcripts based on picture description tasks \cite{Orimaye2017, Konig2018, Mueller2018a, Fraser2015} and have successfully extracted linguistic features that could differentiate between AD and controls.\newline
+\par
+\subsection{Narrative description task}
+The story narration task is designed to study a participant's ability to describe and elaborate on a story which is depicted using a series of pictures.
+The stories depicted are usually based on children's books or famous stories such as Cinderella.
+\cite{Fraser2014}\newline
+\par
+This task requires ordering the story in a structured and coherent framework.
+It also requires comprehension and understanding of the stories characters and the events depicted, as well as an awareness of a character's goals and internal responses to given events.
+This task is particularly useful, as the procedure reduces the demands on memory and is therefore able rule out memory as a confounding variable for any results observed.
+As noted above, Ash \cite{Ash2012} felt that this task was interchangable with the Picture Description task and because this task requires elaboration rather than simple description, is a sturdier test of lexical and semantic abilities as well as syntactic complexity.
+\cite{DeLira2011} \newline
+\par
+Given the relative strengths of the Narrative description task vs Picture description task, there is no research that has used Machine Learning to analyse features from Narrative picture tasks.
+This could be down to the availability of data and the absence of any meaningful sets of transcripts of participants performing this task.
+However, this could be an interesting direction to take research in the future to see if features generated from this task could be used to predict MCI or AD.
+\newline
+\par
+\subsection{Interviews}
+Interviews can also be used to elicit language, the idea of employing questions to guide a conversation between speakers.
+There are three types of interviews: unstructured, structured and semi-structured.
+Structured interviews tend to produce very limited speech and therefore has never been used in this area \cite{Boschi2017}.
+Unstructured interviews are open ended and generally do not conform to any particular pattern.
+They use generic themes such as family or hobbies to guide the conversation.
+Whilst this is the most ecologically valid form of conversation and therefore language generation, it's unstructured nature means that the protocol cannot be consistent and therefore reproduced.
+Semi-structured interviews are therefore preferred other forms of interview as a middle ground.
+The semi structured nature of these interviews means that there is some replicability but does not constrain the participant in answering questions.
+\newline
+\par
+The analysis of interviews can be difficult to analyse as both the content can vary even between participants.
+It is also difficult to measure as there are no pre-defined task goals in comparison to the other two methods.
+Nevertheless, this is the most naturalistic setting for looking at language production and can be used to look at the syntactic and semantic parts of language generation \cite{Sajjadi2012}.
+There have been some attempts to use interviews to to assess language production in AD \cite{}.\newline
+\par
 \section{Connected Language and potential features}

### 2018-07-10 Python/nltk/nltk_ch1_ex.py

@@ -1,76 +1,74 @@
 import nltk
 from nltk.book import *
 
-    ☼ Try using the Python interpreter as a calculator, and typing expressions like 12 / (4 + 1).
+# ☼ Try using the Python interpreter as a calculator, and typing expressions like 12 / (4 + 1).
 
-    ☼ Given an alphabet of 26 letters, there are 26 to the power 10, or 26 ** 10, ten-letter strings we can form. That works out to 141167095653376. How many hundred-letter strings are possible?
+# ☼ Given an alphabet of 26 letters, there are 26 to the power 10, or 26 ** 10, ten-letter strings we can form. That works out to 141167095653376. How many hundred-letter strings are possible?
 
-    ☼ The Python multiplication operation can be applied to lists. What happens when you type ['Monty', 'Python'] * 20, or 3 * sent1?
+# ☼ The Python multiplication operation can be applied to lists. What happens when you type ['Monty', 'Python'] * 20, or 3 * sent1?
 
-    ☼ Review 1 on computing with language. How many words are there in text2? How many distinct words are there?
+# ☼ Review 1 on computing with language. How many words are there in text2? How many distinct words are there?
 
-    ☼ Compare the lexical diversity scores for humor and romance fiction in 1.1. Which genre is more lexically diverse?
+# ☼ Compare the lexical diversity scores for humor and romance fiction in 1.1. Which genre is more lexically diverse?
 
-    ☼ Produce a dispersion plot of the four main protagonists in Sense and Sensibility: Elinor, Marianne, Edward, and Willoughby. What can you observe about the different roles played by the males and females in this novel? Can you identify the couples?
+# ☼ Produce a dispersion plot of the four main protagonists in Sense and Sensibility: Elinor, Marianne, Edward, and Willoughby. What can you observe about the different roles played by the males and females in this novel? Can you identify the couples?
 
-    ☼ Find the collocations in text5.
+# ☼ Find the collocations in text5.
 
-    ☼ Consider the following Python expression: len(set(text4)). State the purpose of this expression. Describe the two steps involved in performing this computation.
+# ☼ Consider the following Python expression: len(set(text4)). State the purpose of this expression. Describe the two steps involved in performing this computation.
 
-    ☼ Review 2 on lists and strings.
-        Define a string and assign it to a variable, e.g., my_string = 'My String' (but put something more interesting in the string). Print the contents of this variable in two ways, first by simply typing the variable name and pressing enter, then by using the print statement.
[…]

### 2018-07-11 LiteratureReview/LiteratureReview3000.tex
updated LR3k

@@
 \par  
-The purpose of this review is to seek to understand what developments have been made in the use of Machine Learning and Natural Language processing to aid the diagnosis of Mild Cognitive Impairment (MCI) and AD.
+The purpose of this review is to seek to understand what techniques have been used when considering the application of Machine Learning and Natural Language processing to aid the diagnosis of Mild Cognitive Impairment (MCI) and AD.
 A search of the literature was conducted using ProQuest (PsychArticles), SCOPUS, Web of Science.
@@
 We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English.
-This resulted in 21 journal articles and/or conference papers which form this review.
+This resulted in 17 journal articles and/or conference papers which form this review.
 
@@
 \par
-In terms of Machine Learning research in this area, a number of researchers have used transcripts based on picture description tasks \cite{Orimaye2017, Konig2018, Mueller2018a, Fraser2015} and have successfully extracted linguistic features that could differentiate between AD and controls.\newline
+In terms of Machine Learning research in this area, a number of researchers have used transcripts based on picture description tasks \cite{Zimmerer2016, Orimaye2017, Mueller2018a, Fraser2015} and have successfully extracted linguistic features that could differentiate between AD and controls.\newline
 \par
@@
 \par
-Given the relative strengths of the Narrative description task vs Picture description task, there is no research that has used Machine Learning to analyse features from Narrative picture tasks.
+Given the relative strengths of the Narrative description task vs Picture description task, there are few pieces of research research that have used Machine Learning to analyse features from Narrative picture tasks \cite{Fraser2014}.
 This could be down to the availability of data and the absence of any meaningful sets of transcripts of participants performing this task.
@@
 Nevertheless, this is the most naturalistic setting for looking at language production and can be used to look at the syntactic and semantic parts of language generation \cite{Sajjadi2012}.
-There have been some attempts to use interviews to to assess language production in AD \cite{}.\newline
+There have been some attempts to use interviews to to assess language production in AD with promising results \cite{Asgari2017, Guinn2015}.\newline
 \par
-\section{Connected Language and potential features}
-Connected Language and it's study can be defined as self generated discourse.
-This is important as this requires a number of cognitive processes which we know are impacted in those who suffer from dementia.
-One of the common problems in the analysis of connected language is the ability to differentiate between what is a purely language related problem vs a memory problem.
+\section{How do we analyse language?}
+Part of the reason we need to pay attention to how we ask participants to generate data  is understanding how we wish to analyse the data afterwards.
+As discussed above, the different methodologies to collect data generate different types of data.
+There are two main approaches which we have looked at to analyse language, using frequencies of words and combinations of words and measures of syntax and semantics.
+There are other less common methods of analysing language but these are beyond the scope of this review.
 \newline
 \par
-\subsection(n-grams)
+\subsection{On using frequencies of words and non-words}
+The basic premise behind using frequencies to measure changes in language is that certain types of words and non words are more prominent in those suffering with AD than healthy controls.\newline 
+\par
 One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
@@
 \par 
-Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences in syntactic, lexical and n-gram linguistic biomarkers to distinguish between those with probably AD and healthy controls, The authors found significant differences in the uses of all three types of biomarkers for those with AD and healthy controls.
-Their results show that the top 1000 n-gram features plus the twenty-three syntactic and lexical features was the most successful at differentiating the two groups (0.93 AUC)).
-However, there were some limitations of the study.
-The context of the audio content was very specific, limited to one specific description task, and therefore using general speech as proposed in this project would potentially use a different set of linguistic features which may or may not have similar predictive power.
-In addition, the transcriptions were encoded using the CHAT format which is a framework for manually annotating speech and the challenge would be finding a way to automate this.
-This paper is noteworthy as it provides evidence that using machine learning to identify participants with MCI is possible and it provides some ideas of potential features which can be used.
-Unique to this paper is the use of n-gram features, which could be explored in the proposed project.\newline
+Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences in syntactic, lexical and n-gram linguistic biomarkers to distinguish between those with a diagnosis of AD and healthy controls, they found significant differences in the uses of all three types of biomarkers but their main finding supported n-grams as the most significant predictor.
+Because Orimaye used the picture description task as their method of data collection, the n-grams generated are only specific to the task given and cannot be generalised.
+A problem with using n-grams for  \newline
 \par
-Asgari, Kaye and Dodge (2017) also looked at the linguistic characteristics of older adults with mild cognitive impairment (MCI) vs healthy controls where they hypothesised that they would be able to predict those with MCI, a distinguishing characteristic of Alzheimer's disease and other variations of dementia.
-Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words.
-They then applied support vector machines (SVM's) and random forest classification algorithms to investigate whether machine learning could be used to distinguish between those with MCI and healthy controls.
-They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84.
-The authors report that this method is highly reliant on high-fidelity transcription of the conversations which is labour intensive, but they anticipate that technology is advancing sufficiently quickly that automated high-quality transcriptions are possible in the near future.
-This paper provides a different perspective in how to tackle the problem deriving language features from speech using a different framework.
-The challenges both Orimaye et al and these authors experienced were around the automation of transcription and this would potentially be an area to explore in the proposed project.
-Asgari et al (2017) used the Linguistic Inquiry and Word Count (LIWC) methodology, developed by Pennebaker et al (2007) to generate features for their paper.
-Pennebaker, Boyd, Jordan and Blackburn (2015) have since developed their framework.
-The LIWC is designed to categorize and evaluate the various emotional, cognitive and structural components which are present within samples of speech and/or written text.
-This method, Linquistic Inquiry and Word Count (LIWC) has evolved from it's initial incarnation in 1993 to it's last update in 2015.
-The premise is the use of words from particular categories provide an insight into the psychological processes and/or diagnoses an individual has.
-The latest version of this methodology uses a dictionary of 6400 words, word stems and select emoticons and assigns these constructs to various categories.
-This has been updated to include modern uses of language such as 'text speak'.
-This manual goes summarises the process of constructing the framework as well as discussing the reliability and validity of the measure and a review of studies using the LIWC was conducted by Tausczik and Pennebaker (2010) support the notion that the LIWC is valid across multiple psychological domains.
-This paper is useful as it provides a useful and validated framework from which the proposed project can potentially derive language features.\newline
-\par
-Currently formal diagnosis of depression is expensive and time-consuming involving both the use of questionnaires and the use of a trained professional to assess an individual.
-Schwartz et al, explored the use of language as an aid to diagnose depression in a naturalistic setting i.e.
-facebook status updates.
-The authors used both the LIWC (described above) and n-grams as features to distinguish those with depression from healthy individuals and were able to track participants levels' of depression through language successfully.Whilst the authors were able to use a reasonably sound method of measure degree of depression and have evidenced a methodology that is able to track a levels of depression over time, it would be more helpful to use a more clinically relevant measure of depression and anxiety i.e., PHQ-9 (Kroenke, 2001) and GAD-7 (Spitzer et al, 2006) to add validity to their findings.
-This paper is interesting as it uses both LIWC and n-grams as a method of deriving features from text / speech for depression which, as described above, have also been used as features to distinguish MCI from healthy participants.
-Given the comorbity of depression and dementia (Meyers, 1998), it would be interesting for the proposed project to further analyse the link between dementia and clinical depression via the use of language.
-This paper also supports the idea that language changes over time and is a marker for deterioration in both those with depression and dementia.
+Asgari, Kaye and Dodge (2017) \cite{Asgari2017} used another form of word frequency measurement.
+Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words \cite{Pennebaker2015}.
+They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
 \newline
 \par
-Fraser, Meltzer and Rudzicz (2015) looked at connected speech using the DementiaBank corpus.
+\subsection{Measures of Syntactic and Semantic Complexity}
+Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
 They found that there were four factors which informed the classification of participants as either healthy or AD.
@@
 \par
-Lopez-de-Ipina (2018) have since moved on
-
-\section{Conclusion}
+Zimmerer (2016) \cite{Zimmerer2016} looking at whether language was more formulaic \newline
+\par
+\section{Conclusions and Future Work}
 \par
@@
 As these differences in language are quite subtle, the varying subtypes of dementia are frequently misdiagnosed and can sometimes be missed altogether.
+\newline
+\par
+One of the critisicms of the literature to date is the features that are being used to measure language.
+For example, Zimmerer (2016) \cite{Zimmerer2016} describes connectivity in such as a way that correlates directly with what Mueller (2018) \cite{Mueller2018a} calls Fluency.
+Whilst these are very nuanced measures which differ slightly in the form they take, the sheer range of measures and features being produced make it difficult to conceptualise as part of the larger problem.
+Some work needs to be done in producing a consistent list of measures that are validated using existing datasets and can be used for future research.
 \newline
@@
 Given the amount of processing power and the sophistication of machine learning, in particular deep learning algorithms.
-Further, despite the quality of datasets being used to 'backtest' these algorithms, further research should look at generating further datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which we could claim are limited in scope.
+Further, despite the quality of datasets being used to 'backtest' these algorithms,further research should look at generating further datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which we could claim are limited in scope.
 It would be worth looking at how we can utilse these technologies further to aid detect the earliest signs of cognitive impairment.

### 2018-07-12 LiteratureReview/LiteratureReview3000.tex
completed LR3k

@@
 \documentclass[a4paper]{article}
-
 %% Packages
 \input{settings/packages}
-
 %% Page Settings
 \input{settings/page}
-
 %% Own Commands
 \input{settings/macros}
-
-\title{Literature Review}
+\title{How can machine learning and natural language processing techniques aid the process of diagnosing Mild Cognitive Impairment and Alzheimer's Disease?}
 \date{2018-07-13}
 \author{Jomar Alcantara}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{center}
-	\begin{tabular}{ | l | c | r | p{1cm} |}
+\begin{table}
+	\begin{tabular}{ | c | c | c | p{1cm} |}
 		\hline
@@
 	\end{tabular}
-\end{center}
+	\caption{\label{tab:table-name}Search Terms and Number of Results.}
+\end{table}
 
@@
 This is useful as, given a large portion of text or speech, we can predict the probability of a word being close by to a given word.
-One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class.
 \newline
 \par
+A number of researchers have used n-grams as features.
+One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class.
 Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences in syntactic, lexical and n-gram linguistic biomarkers to distinguish between those with a diagnosis of AD and healthy controls, they found significant differences in the uses of all three types of biomarkers but their main finding supported n-grams as the most significant predictor.
 Because Orimaye used the picture description task as their method of data collection, the n-grams generated are only specific to the task given and cannot be generalised.
-A problem with using n-grams for  \newline
+\newline
 \par
@@
 \subsection{Measures of Syntactic and Semantic Complexity}
+Another approach to linguistic analysis in this field is the idea of measuing syntactic and semantic complexity.
+According to Emery (2000) \cite{Emery2000} in which she states that Semantic and Syntactic skills deteriorate first in people with MCI and AD.
+If this is true, then psychological measures of semantic and syntactic skills should be able to pick up signs of deterioration and act as markers for possible MCI and AD.
+An example of these measures is the concept of idea density.
+Formally, idea density is defined as the average number of propositions per sentence \cite{Kintsch1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}.
+Type to Token Ratio, Brunet's Index and Honore's Statistic are two more examples of validated measures which are used to measure the lexical diversity of a given piece of text.
+This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2005}.
+\newline
+\par
 Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
 They found that there were four factors which informed the classification of participants as either healthy or AD.
-These four factors were semantic impairment, acoustic abnormality, syntactic impairment and information impairment.\newline
+These four factors were semantic impairment, acoustic abnormality, syntactic impairment and information impairment.
+Zimmerer (2016) \cite{Zimmerer2016} looking at whether language was more formulaic in those suffering from AD.
+He proposed that those who suffer from AD rely on formulatic sentences for example 'Noun Verb Noun' is a formulaic sentence structure and by relying on this, AD suffers reduce the cognitive load of trying to come up with a sentence structure.
+\newline
 \par
-Zimmerer (2016) \cite{Zimmerer2016} looking at whether language was more formulaic \newline
+It appears that both frequency based approaches and approaches that use measures of linguistic complexity have both been shown to be effective in various experimental settings.
+The fact that there are two vastly different perspectives to tackling this question, and even within these perspectives there are a range of various methodologies leaves the field with a sense of confusion as to the most effective way to solving the problem.
+\newline
 \par
@@
 As these differences in language are quite subtle, the varying subtypes of dementia are frequently misdiagnosed and can sometimes be missed altogether.
+Given the burden on the diagnosis of dementia on clinicians, it appears to be useful to find some non-invasive protocols for the early diagnosis of dementia.
+It has already been shown that analysis of speech and language has shown markers that pre-date the official diagnosis of dementia \cite{Snowdon1996, Berisha2015}.
+A significant amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
+Firstly, the vast majority of the research described above looks at pre-existing datasets like the DementiaBank corpus.
+Whilst this is useful for 'backtesting the data', it would be useful for this process to be tested live to see whether these results can be replicated now.
 \newline
@@
 \par
-Given the burden on the diagnosis of dementia on clinicians, it appears to be useful to find some non-invasive protocols for the early diagnosis of dementia.
-It has already been shown that analysis of speech and language has shown markers that pre-date the official diagnosis of dementia (Snowdon et al, 1996;)\cite{Berisha2015}.
-A significant amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
-Firstly, the vast majority of the research described above looks at pre-existing datasets like the DementiaBank corpus (CITE).
-Whilst this is useful for 'backtesting the data', it would be useful for this process to be tested live to see whether these results can be replicated now.
-\newline
-\par
-Future research should be directed towards developing non-intrusive ways of detecting subtle changes in language based in the home, such that any deterioration that could indicate the presence of MCI or AD and be flagged up for further review via referral.
-Given the amount of processing power and the sophistication of machine learning, in particular deep learning algorithms.
-Further, despite the quality of datasets being used to 'backtest' these algorithms,further research should look at generating further datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which we could claim are limited in scope.
-It would be worth looking at how we can utilse these technologies further to aid detect the earliest signs of cognitive impairment.
-
+Future research should be directed towards developing non-intrusive ways of detecting subtle changes in language based in the home, such that any deterioration that could indicate the presence of MCI or AD and be flagged up early in the progression of any potential cogntive impairment for further review via referral.
+Given the amount of processing power and the sophistication of machine learning, in particular deep learning algorithms, machine learning approaches seem to be the most logical approach for achieving this aim.
+Further, despite the quality of datasets being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which we could claim are limited in scope.
+This area of research is extremely promising in its early results and the impact of successful research would be life changing for both individuals and the health of the worlds aging population in general.
 \bibliographystyle{unsrt}

### 2018-07-12 ProjectPlan/ProjectPlan.tex
completed ProjectPlan and Gantt Char

--- 2018-07-09 ProjectPlan/ProjectPlan.tex
@@
 \title{Project Plan}
-\date{2018-06-06}
+\date{2018-07-13}
 \author{Jomar Alcantara}
@@
 \begin {document}
-\section{Aims and Objectives}
+\maketitle
+\section{Background, Aims and Objectives}
+According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of Dementia in the UK at approximately 815,000 people.
+This represents 1 in 14 of the population aged 65 or over and 1 in 79 in the general population.
+This report also estimates an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis.
+They also estimate that the overall impact of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually.
+Globally, this picture is a lot bleaker.
+A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050.
+The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillon US dollars.\newline
+\par
+The aim of the research I am conducting is to investigate the use of technology, particularly the use of Machine Learning and Natural Language Processing techniques to detect mild cognitive impairment (MCI) and / or Alzheimer's Disease (AD).
+The objectives are: To design a framework and develop.
+To develop a proof of concept design.
+To investigate the application of my framework to be used in a more naturalistic setting i.e.
+in a person's home, rather than in a healthcare setting.
 \section{Tasks}
-	\subsection{April 2018}
-	Thursday, 26th April - Workplan (First Draft Agreed)
-	\subsection{May 2018}
-	Wednesday, 9th May - Alexa Skills Training (All Day)
-	\subsection{June 2018}
-	Friday 15th June - First Draft of Literature Review to be handed in.
+\subsection{Literature Review}
+The Literature Review is divided into three sections as my research is interdisciplinary.
+I will be looking at the psychological perspective of language testing and analysis for MCI and/or AD as this will provide the theoretical foundation for any work I do with Machine Learning.
+I will then will look at the different possible ways to measure for language difficulties taking into account approaches that use frequencies, approaches that use measures of semantic and syntactic complexity and other more novel approaches.
+Finally, I will look at current developments in Machine Learning as it pertains to this area of research including theoretical approaches as well as previous research applying research in this specific area.\newline
+\par
+\subsection{1st Year PhD Requirements}
+There are a number of requirements for the first year of my Ph.D Programme.
+The first part of this is the IS4001 module which consists of three assessments.
+These are: A presentation, A project plan and written report and a 3000 word literature review.
+In addition, I must complete a qualifying report which is a report in which I talk about my progress and how I have developed as a researcher in the first year.
+This also looks at any papers I have submitted / published and a literature review in the area I'm studying.
+Finally, I must defend my qualifying report at a Viva.
 \newline
-	Monday 18th June - Monday 25th June - Annual Leave
-	\subsection{July 2018}
-	Friday 13th July - Written Assignment (IS4001) Due
-	Friday 13th July - Project Plan (IS4001) Due
-
-	\subsection{August 2018}
-	Friday 10th August - First draft of experimental proposal to be submitted.
-
-	\subsection{September 2018}
-	Friday 21st September - Final draft of experimental proposal to be submitted. 
-	Friday 21st September - NHS Ethics forms to be sorted  - Takes 60 days!
-	\subsection{October 2018}
-
-	\subsection{November 2018}
-
-	\subsection{December 2018}
-
-	\subsection{January 2019}
-	Friday 11th January - Rough Draft of Qualifying Due
-	Friday 18th January - Complete first rough pass of capturing and processing speech.
-	Thursday, 31st January - Qualifying Report Due
-
-	\subsection{February 2019}
-
-	\subsection{March 2019}
-
-	\subsection{April 2019}
-	Monday, 1st April - Viva Due by this point.
+\par
+\subsection{Skills Training and Knowledge Gathering}
+Under this section, I have identified a number of knowledge gaps where I need to work on.
+The main knowledge gap is in learning a new programming language and the extensions I will need to learn and use in my research.
+Given this is a fundamental part of my research, I have allowed six months to get up to speed.
+I also plan on attending two key conferences in my field of researched.\newline
+\subsection{Research Planning}
+My research planning is going be based on identifying gaps in the literature and planning some experiments to test proof of concept.
+Based on my literature review, I will be designing an experiment which means the objectives I have listed above.
+As part of the research planning stage, I intend to submit my research proposal for NHS ethics approval in order to join the clinical research network Dementias and Neurodegneration.
+This process can take up to 60 days for the committee to initially look at the research proposal and I have allowed time for the research proposal to be rejected and resubmitted.
+\newline
+\par
+\subsection{Sample Dataset Collection and Cleaning Model Conceptualisation and Production}
+I have identified a number of pre-existing datasets that I can use to develop my models and to test the data pipeline.
+However these datasets are owned by other research groups.
+There is an application process in order to obtain some of the datasets and they require approval of the research group before the data can be released.
+Using this data I plan to develop a model and test methodologies on these datasets.
+This is mainly a replication of previous research but allows me to validate my model and also allows me to smooth out any difficulties in the processing of large sets of data before I start to collect my own data.
+\newline
+\par
 \section{Obstacles}
+There are a number obstacles that may delay my progress.
+The most obvious example is my application for NHS Ethics Approval.
+The average time take for this to be accepted is 60 days.
+I have allowed 90 days in my project plan as my research proposal may require amending and resubmitting.
+Another obstacle is in the acquisition of datasets.
+My experience with acquiring other datasets is that applications can take up to 2 weeks to be approved.
+I have therefore allowed up to a month for each dataset including the application process.
+\newline
+\par
+In terms of managing my own time.
+I have assumed working eight hour days from Monday to Friday and for each task, I have estimated the time I would need to complete the task and then added on an additional 33\% to 50\% as contingency.
+This should allow me the ability to 'catch up' if I fall behind at any stage.
 \end {document}

### 2018-07-13 ExperimentProposals/Experiments.tex
tidy up of folders and moved LR3K files to archive

--- 2018-05-14 QualifyingReport/QualifyingReport.tex
@@
 
-\title{Qualifying Report}
-\date{2018-05-10}
+\title{Experimental Report}
+\date{2018-04-22}
 \author{Jomar Alcantara}
@@
 	\maketitle
-	\newpage
-	\section{Overview of Problem}
-	Dementia has been identified as one of those fast growing difficulties facing the world.
-A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050(CITE).
-The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
-	\newline
-	\par
-	In 2009, the Department of Health in it's National Dementia strategy made 'early diagnosis and support' one of it's key themes as part of this strategy.
-A lot of work, therefore, has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cogntive Impairment (MCI).
-As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages and thus it is important that we utilise technology wherever possible to aid clinicians.
-At the present time diagnosis is typically conducted at memory clinics by trained clinicians.
-I theorise that we might be able to use diagnose patients in their own homes utilising smart-home technology.
-	\newline
-	\par
-	Language analysis is an important part of assessing those suspected of suffering with AD or MCI and this language analysis in part informs an official clinical diagnosis.
-This analysis often takes the form of tasks such as Free Cued Selective Recall Task, the Boston Cognitive Assessment, and the Mini Mental State Examination which are all quite targeted in the language required of the person being assessed.
-	\newline
-	\par
-	A step towards the ambient detection of AD and/or MCI in population requires that we look for similar features in language used naturally as part of every day conversation.
-Part of this work therefore is to see whether spontaneous discourse, such a semi-structured interview is a task that has the ability to put pressure both the cognitive and linguistic systems as it requires the ability to understand speech, formulating a coherent answer and then responding in an appropriate way.
-Berisha et al \cite{Berisha2015}, has shown through language analysis in this area that there are marked differences in this process between those who have a diagnosis of AD and healthy controls.
-The next part of this work looks at how technology, particularly the use of smart home devices such as the Amazon Echo and Google Home families of devices can be used to capture data that can then be analysed.
-	\newline
-	\par
-	The potential impact of this research is immense.
-Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can slow the progress of the disease.
-Equally, early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
-
-	\section{Literature Review}
-
-	\section{Work already carried out}
-
-	\subsection{Literature Review - April 2018 to July 2018}
-	A literature review was carried out from several perspectives in order to understand both an applied and theoretical perspective.
-Thus a search for papers was conducted in both the psychology and computer science domains, as well as papers which intersected these areas and were relevant to this particular problem.
-	\newline
-	SEARCH - ProQuest (PsychArticles) - Search Terms (Language AND Decline AND DEMENTIA) - 1484 Results
-	SEARCH - ProQuest (PsychArticles) - Search Terms (Language AND Decline AND DEMENTIA AND Speech) - 486 Results
-	SEARCH - Web Of Science - Search Terms (Language AND Decline AND DEMENTIA AND Speech) - 151 Results
-	SEARCH - Web of Science - Search Terms (Language AND Decline AND DEMENTIA) - 1207 Results
-	SEARCH - Scopus - Search Terms (Language AND Decline AND DEMENTIA AND Speech) - 91 Results 
-	SEARCH - Scopus - Search Terms (Languge AND DEMENTIA AND Decline) - 791 Results
-
-	\subsection{Skills development - April 2018 to March 2019}
-	\textbf{IS-4001 Research Skills and Professional Development Module - Aston University - 3rd May, 10th May, 16th May - 45 Hours} \newline
-	\par
-	Day 1 consisted of an introduction to the module (IS4001), a rough guide to how to start a PhD and a look at how to conduct and write a literature review.
-This looked at the concept of a PhD and the meaning of original research.
-This also talked about the supervisor relationship and understanding the role of the supervisor in the PhD.
-In addition, we explored the idea of our expectations around supervision.
-We then went on to look at what skills, qualities and attitudes are necessary to be a good researcher including the idea of what consitutes ethical research and finally we looked at the obstacles to progress on a PhD and what we can do as students.
-	\newline
-	\par
-	In the afternoon, we looked at how a way of conducting a literature review.
-This talked about what a literature review is, along with the different types of literature review (Narrative, Systematic and Meta-Analysis).
-We talked about the function of a literature review, and how this fits in with the writing up of a thesis.
-We then explored the stages that the literature review process contains, from defining the scope of the research, to coming up with a plan on how to collect sources of information to synthesising and reporting the findings of the review.
-This part of the lecture looked in detail at each of these areas and included some examples to put the theory of the literature into context.
-Finally, we talked about the the process of a literature review in the context of the written assignment.
-	\newline
-	\par
-	Test
-	\newline
-	\par
-	\textbf{Alexa Skills Training Day - Amazon - May 9th 2018 - 7 Hours}
-	\newline
-	\par
-	Test
-	\newline
-	\textbf{Deep Learning Online Course - deeplearning.ai - Dates - 70 Hours}
-	Summary Here
-
-	\section{Work Plan for the future}	
-	\subsection{Proposed Experiments}
-	\subsection{Publication plan}
-
-	\section{Proposed Timetable}
-
-	\subsection{June 2018}
-	Friday 15th June - First Draft of Literature Review to be handed in.
-	\newline
-	Monday 18th June - Monday 25th June - Annual Leave
-	\subsection{January 2019}
-	Friday 11th January - Rough Draft of Qualifying Due
-	Friday 18th January - Complete first rough pass of capturing and processing speech.
-
-	\subsection{February 2019}
-	Friday, 1st February - Qualifying Report Due
-	\subsection{April 2019}
-	Monday, 1st April - Viva Due by this point.
-
-	\section{Reflection on my experience so far}
-
-	\section{Provisional Table of Contents}
-	\subsection{Chapter 1: Introduction}
-	\subsection{Chapter 2: Literature Review}
-	\subsection{Chapter 3: New theories, Description of experiments}
-	\subsection{Chapter 4: Results}
-	\subsection{Chapter 5: Discussion}
-	\subsection{Chapter 6: Conclusion}
-
-	%Complete appendices for each of the assessment. How do you do appendices in LaTeX
 
 \bibliographystyle{unsrt}
-	\bibliography{QualifyingReport}
+\bibliography{ExperimentReport}
 

### 2018-07-13 LiteratureReview/LiteratureReview.tex
updated original literature review

--- 2018-04-30 LiteratureReview/LiteratureReview.tex
@@
 \newpage
-
 \section{Introduction}
-
 \par
-According to a recent report commissioned by the Alzheimer's Society, they estimate the prevalence of Dementia in the UK at approximately 815,000 people.
-This represents 1 in 14 of the population aged 65 or over.
-This report also estimates an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis.
-They also estimate that the overall impact of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually.
-Globally, this picture is a lot bleaker.
-A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
-The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillon US dollars.
+Alzheimer's Disease (AD) is a neurodegenerative disease in which the brain develops neurofibrillary tangles and neuritic plaques along with the loss of cortical neurons and synapses.
+The hallmark clinical symptoms are cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits\cite{AmericanPsychiatricAssociation2013}.
+In addition, these symptoms are often accompanied by emotional difficulties such as depression and irritability and behavioural difficulties.
+Whilst there are several variants of Dementia, AD remains the most common type of Dementia and will be the focus of this review.
 \newline
 \par
-The purpose of this review is to seek to understand how dementia impacts speech and language production from a psychological and linguistic perspective with a view to potentially using natural language processing and technology to aid in the early diagnosis of dementia and/or Mild Cognitive Impairment.
+According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
+This represents 1 in 14 of the population aged 65 or over and 1 in 79 in the general population.
+They estimate an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
+Globally, this picture is a lot bleaker.
+A recent report suggests that in 2015 there were 46 million people with a diagnosis of AD and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
+\newline
+\par
+Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent.
+Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time until a threshold is reached and clinical symptoms become apparent \cite{Nestor2006}.
+If this is the case, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
+Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on the decline in the use of language in those with AD.
+\newline
+\par
+One of the most famous pieces of research was by Berisha and Liss (2015) \cite{Berisha2015} who examined speeches and public interviews of former US president Ronald Reagan.
+They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease.
+It turned out later to be the case that he had Alzheimer's Disease.
+Another classical study by Snowdon et al (1996) \cite{Snowdon1996} looked at whether linguistic ability in early life was associated with cognitive function and AD in later life.
+They found that idea density (defined as the number of expressed propositions divided by the number of words) was a key predictor in predicting whether nuns would go on to develop AD in later life.
+They found that those who would go on to develop AD all had low idea density in early life and they found no AD present in those with high idea density in early life.
+As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are wide and varied and differ as the disease develops.
+\newline
+\par
+According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with Mild dementia suffer from noticeable word finding difficulty.
+They may substitute general terms for more specific terms and may avoid the use of specific names of acquaintances.
+There may be grammatical errors involving subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc.
+Those who have progressed from Mild to Major depression also have difficulties with expressive or receptive language.
+They will often use general-use phrases such as "that thing" and "you know what I mean" and prefers general pronouns rather than names.
+With severe impairment, sufferers may not even recall names of closer friends and family.
+Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur.
+Echolalia (meaningless repetition of another person's spoken words) and automatic speech typically precede mutism.
+With the wide range of deficits someone with AD can suffer, it makes sense to try to categorise these deficits in some way.\newline
+\par
+Emery \cite{Emery2000} completed a literature review looking at all the potential language deficits that could exist in those with AD and / or MCI.
+She divided these deficits into four levels of language: Phonology, Morphology, Syntax and Semantics.
+She proposed that language and the processes involved in language are hierarchical in nature and that language moves from simple units of construction (Phonology and Morphology), and build layers of complexity and sophistication (Syntax and Semantics).
+She found that people with AD generally had intact Phonology and Morphology but more impaired Syntax and Semantics.
+She stated that the language forms we learn last are the first to deteriorate.
+We generally learn language in small simple units initially and build syntax and complexity as we are more comfortable with language.\newline
+\par
+It is clear from both the clinical diagnostic criteria and supporting research that language is impacted in those with AD.
+However, one of the costs of analysing language is that there is a huge burden on trained practitioners, be it clinical psychologists, audio transcribers and text encoders to facilitate the process of collecting data and analysis.
+The field of machine learning and natural language processing has been suggested as a way to improve the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline \cite{Boschi2017}.
+\newline
+\par  
+The purpose of this review is to seek to understand what techniques have been used when considering the application of Machine Learning and Natural Language processing to aid the diagnosis of Mild Cognitive Impairment (MCI) and AD.
+A search of the literature was conducted using ProQuest (PsychArticles), SCOPUS, Web of Science.
+The following results were found (Table 1).
+All papers were then reviewed for relevance by reading the abstract and full text where appropriate and a shortlist was compiled.
+An additonal search through references of shortlisted papers was also conducted and any papers who upon further review appeared relevant were added to the shortlist.
+Papers were included where researchers used machine learning to classify participants as MCI, AD or Healthy using language.
+We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English.
+This resulted in 17 journal articles and/or conference papers which form this review.
+
+\begin{table}
+	\begin{tabular}{ | c | c | c | p{1cm} |}
+		\hline
+		Database & Number of Results & Search Terms  \\ \hline
+		ProQuest(PsychArticles) & 1484 Results & Language AND Decline AND Dementia \\ \hline
+		ProQuest(PsychArticles) & 486 Results  & Language AND Decline AND Dementia AND Speech \\ \hline
+		ProQuest(PsychArticles) & 159 Results & Machine Learning AND Dementia AND Language \\ \hline
+		Web of Science & 1207 Results  & Language AND Decline AND Dementia   \\ \hline
+		Web of Science & 151 Results  & Language AND Decline AND Dementia AND Speech  \\ \hline
+		Web of Science & 34 Results & Machine Learning AND Dementia AND Language \\ \hline
+		Scopus & 791 Results & Language AND Decline AND Dementia  \\ \hline
+		Scopus & 91 Results & Language AND Decline AND Dementia AND Speech   \\ \hline
+		Scopus & 29 Results & Machine Learning AND Dementia AND Langauge \\ \hline
+	\end{tabular}
+	\caption{\label{tab:table-name}Search Terms and Number of Results.}
+\end{table}
+\section {Types of Language Assessment}
+One of the key debates when looking at how to analyse language is the type of task provided to elicit language production in participants.
+In the literature researchers have primarily focused on Picture Description tasks but have also suggested other ways in which we might collect data.
+\newline
+\par
+\subsection{Picture Description Tasks}
+One of the most commonly used tasks to measure language is the Picture Description task.
+An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Kaplan2010}.
+In this task participants are asked to describe a picture presented to them in as much detail as possible.
+The picture itself depicts a familiar domestic scene and would not require participants to use any vocabulary beyond that learned in childhood.
+It was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}\newline
+\par
+The picture description task does a fine job of eliciting descriptive language, however because of the limited content has limited use.
+The task in itself just a descriptive task, and therefore elicits a certain type of language.
+There is some disagreement as to the benefits of this using this methodology.
+This task is reported as being useful to lexico-semantic disorders \cite{Boschi2017, Sajjadi2012} as the language being generated is primarly nouns and deixis (words to identify items and words to put those items into context).
+However, Ash \cite{Ash2012}felt that there was no difference in using this task vs Story Narration (described below).
+In explaining the differences, it is worth noting that these researchers were using differing variables and this could explain their different perspectives.
+\newline
+\par
+In terms of Machine Learning research in this area, a number of researchers have used transcripts based on picture description tasks \cite{Zimmerer2016, Orimaye2017, Mueller2018a, Fraser2015} and have successfully extracted linguistic features that could differentiate between AD and controls.\newline
+\par
+\subsection{Narrative description task}
+The story narration task is designed to study a participant's ability to describe and elaborate on a story which is depicted using a series of pictures.
+The stories depicted are usually based on children's books or famous stories such as Cinderella.
+\cite{Fraser2014} This task requires ordering the story in a structured and coherent framework.
+It also requires comprehension and understanding of the stories characters and the events depicted, as well as an awareness of a character's goals and internal responses to given events.
+This task is particularly useful, as the procedure reduces the demands on memory and is therefore able rule out memory as a confounding variable for any results observed.
+As noted above, Ash \cite{Ash2012} felt that this task was interchangable with the Picture Description task and because this task requires elaboration rather than simple description, is a sturdier test of lexical and semantic abilities as well as syntactic complexity.
+\cite{DeLira2011} \newline
+\par
+Given the relative strengths of the Narrative description task vs Picture description task, there are few pieces of research research that have used Machine Learning to analyse features from Narrative picture tasks \cite{Fraser2014}.
+This could be down to the availability of data and the absence of any meaningful sets of transcripts of participants performing this task.
+However, this could be an interesting direction to take research in the future to see if features generated from this task could be used to predict MCI or AD.
+\newline
+\par
+\subsection{Interviews}
+Interviews can also be used to elicit language, the idea of employing questions to guide a conversation between speakers.
+There are three types of interviews: unstructured, structured and semi-structured.
+Structured interviews tend to produce very limited speech and therefore has never been used in this area \cite{Boschi2017}.
+Unstructured interviews are open ended and generally do not conform to any particular pattern.
+They use generic themes such as family or hobbies to guide the conversation.
+Whilst this is the most ecologically valid form of conversation and therefore language generation, it's unstructured nature means that the protocol cannot be consistent and therefore reproduced.
+Semi-structured interviews are therefore preferred other forms of interview as a middle ground.
+The semi structured nature of these interviews means that there is some replicability but does not constrain the participant in answering questions.
+\newline
+\par
+The analysis of interviews can be difficult to analyse as both the content can vary even between participants.
+It is also difficult to measure as there are no pre-defined task goals in comparison to the other two methods.
+Nevertheless, this is the most naturalistic setting for looking at language production and can be used to look at the syntactic and semantic parts of language generation \cite{Sajjadi2012}.
+There have been some attempts to use interviews to to assess language production in AD with promising results \cite{Asgari2017, Guinn2015}.\newline
+\par
+\section{How do we analyse language?}
+Part of the reason we need to pay attention to how we ask participants to generate data is understanding how we wish to analyse the data afterwards.
+As discussed above, the different methodologies to collect data generate different types of language.
+There are two main approaches which we have looked at to analyse language, using frequencies of words and combinations of words and measures of syntax and semantics.
+There are other less common methods of analysing language but these are beyond the scope of this review.
+\newline
+\par
+\subsection{Using frequencies of words and non-words}
+The basic premise behind using frequencies to measure changes in language is that certain types of words and non words are more prominent in those suffering with AD than healthy controls.
+For example, Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in diaglogue in which a speaker provides responsees do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls..\newline
+\par
+One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
+An n-gram is a contiguous sequence of n items from a given sample of text or speech.
+The items can be phonemes, syllables, letters, words or base pairs according to the application.
+For example, given the sequence of words "to be or not to be", this extract is said to contain six 1-gram sequences (to, be, or, not, to, be), five 2-gram sequences (to be, be or, or not, not to, to be), four 3-gram sequences(to be or, be or not, or not to, not to be) and so on.
+This is useful as, given a large portion of text or speech, we can predict the probability of a word being close by to a given word.
+\newline
+\par
+A number of researchers have used n-grams as features.
+One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class.
+Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls.
+Their main finding supported n-grams as the most significant predictor.
+One of the criticisms is the use of picture description tasks and n-grams.
+Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
+\newline
+\par
+Asgari, Kaye and Dodge (2017) \cite{Asgari2017} used another form of word frequency measurement.
+Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words \cite{Pennebaker2015}.
+They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
+\newline
+\par
+\subsection{Measures of Syntactic and Semantic Complexity}
+Another approach to linguistic analysis in this field is the idea of measuing syntactic and semantic complexity.
+According to Emery (2000) \cite{Emery2000} in which she states that Semantic and Syntactic skills deteriorate first in people with MCI and AD.
+If this is true, then psychological measures of semantic and syntactic skills should be able to pick up signs of deterioration and act as markers for possible MCI and AD.
+An example of these measures is the concept of idea density.
+Formally, idea density is defined as the average number of propositions per sentence \cite{Kintsch1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}.
+Type to Token Ratio, Brunet's Index and Honore's Statistic are two more examples of validated measures which are used to measure the lexical diversity of a given piece of text.
+This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2005}.
+\newline
+\par
+Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
+They found that there were four factors which informed the classification of participants as either healthy or AD.
+These four factors were semantic impairment, acoustic abnormality, syntactic impairment and information impairment and were based on existing measures of semantic and syntactic complexity.
+Zimmerer (2016) \cite{Zimmerer2016} looked at whether language was more formulaic in those suffering from AD.
+He proposed that those who suffer from AD rely on formulatic sentences, for example 'Noun-Verb-Noun', and this is done to reduce language complexity.
+He noticed a significant difference in the use of formulaic sentences between AD and Healthy Controls.
+\newline
+\par
+It appears that both frequency based approaches and approaches that use measures of linguistic complexity have both been shown to be effective in various experimental settings.
+The fact that there are two vastly different perspectives to tackling this question, and even within these perspectives there are a range of various methodologies leaves the field with a sense of confusion as to the most effective way to solving the problem.
+\newline
+\par
+\section{Conclusions and Future Work}
+\par
+Identification of language impairment is key idea in the diagnosis of AD and early diagnosis can alter the prognosis and change the management of this degenerative disorder as well as provide new opportunities to study the progression of the disease.
+Given the burden on the diagnosis of AD on professionals there has been a call to use technology to potentially ease this burden \cite{Boschi2017}.
+It has already been shown that analysis of speech and language has shown markers that pre-date the official diagnosis of dementia \cite{Snowdon1996, Berisha2015}.
+A significant amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
+\newline
+\par
+One of the areas for research to study is a careful examination of the features that are being used to measure language deterioration.
+For example, Zimmerer (2016) \cite{Zimmerer2016} describes connectivity in such as a way that correlates directly with what Mueller (2018) \cite{Mueller2018a} calls Fluency.
+Whilst these are very nuanced measures which differ slightly in the form they take, the sheer range of measures and features being produced make it difficult to conceptualise as part of the larger problem.
+Some work needs to be done in producing a consistent list of measures that are validated using existing datasets and can be used for future research moving forward.
+\newline
+\par
+Future research should also be directed towards developing non-intrusive ways of detecting subtle changes in language based in the home, such that any deterioration that could indicate the presence of MCI or AD and be flagged up early in the progression of any potential cogntive impairment for further review via referral.
+Machine learning approaches seem to be the most logical approach for achieving this aim.
+Further, despite the quality of datasets being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which we could claim are limited in scope.
+\newline
+\par
+This area of research is extremely promising in its early results and the impact of successful research would be life changing for both individuals and the health of the worlds aging population in general.
 \newline
@@
 
-\section{Description of Dementia}
-
-Language - DSM-V
-'Major' Has significant difficulties with expressive or receptive language.
-Often uses general-use phrases such as "that thing" and "you know what I mean," and prefers general pronouns rather than names.
-With severe impairment, may not even recall names of closer friends and family.
-Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur.
-Stereotypy of speech occurs; echolalia and automatic speech typically precede mutism.
-
-'Mild' Has noticeable word-finding difficulty.
-May substitute general for specific terms.
-May avoid use of specific names of acquaintances.
-Grammatical errors involve subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc.
-
-\subsection{Alzheimer's Disease}
-\subsection{Vascular Dementia}
-\subsection{Dementia with Lewy Bodies}
-\subsection{Parkinson's disease dementia}
-\subsection{Fronto-temporal dementia}
-\subsection{Primary Progressive Aphasia}
-\subsection{Mild Cognitive Impairment}
-
-\section{Assessment of Language function in Dementia}
-
-Language can be defined as the ability to encode ideas into words and/or symbols for the means of communication.
-Difficulty in language, both spoken and written, are often described as symptoms of various types of Alzheimers' disease (AD).
-One of the challenges when using speech as a predictor is recognising whether there is a problem with language production or with the motor skills.
-Impairments in speech that arise from any process that disrupts the neuraxis from the cortex to muscle and encompass dysarthria (disturbance in articulation) and dysphonia (disturbance in the production in vocal sounds).
-However it is important to use language as a tool to aid the diagnosis of Dementia, as it provides vital clues to aid a clinician in differentiating in the different types of Dementia, which will in turn aid a clinician in an attempt to manage an individual case of Dementia.
-Due to the subtle nature of the languages changes that are experienced in those who suffer with different types of Dementia, it is often misdiagnosed.
-
-A common way for clinicians' to diagnose dementia is through use of the Mini-Mental State Examination.
-The Mini-Mental State Examination (MMSE) has shown to be particularly effective at differentiating betweeen Dementia and other psychological disorders such as clinical depression, schitzophrenia and personality disorder and is currently the most widely used measure of diagnosing Dementia within clinical psychology.
-However, it is not without it's faults...
-
 \section{Current ways of diagnosing Dementia using Language}
@@
 The Free Cued Selective Recall Test (FCSRT) was created by Grober and his team \cite{Grober1988}.
+This is a test that primarily tests memory but also tests a subjects ability to name objects.
+Four items are presented at a time on a card and each participant is asked to name and recall with and without cues.
+
+\subsection{Mini-Mental State Examination}
 
@@
 
+\section{stuff to add}
 \section{Conclusion}

### 2018-07-19 Python/Kaggle/HomeCredit/model1.py
added data from Reagan and Bush and toy tensorflow model

@@ -1,13 +1,0 @@
-import numpy as np 
-import pandas as pd 
-import matplotlib as plt
-
-train = pd.read_csv('~/Documents/Data/HomeCredit/application_train.csv')
-test = pd.read_csv('~/Documents/Data/HomeCredit/application_test.csv')
-
-pd.set_option('display.max_colwidth', -1)
-# print(train.head())
-# print('-' * 40)
-# print(test.head())
-
-print(list(train))

### 2018-07-19 Python/LPTHW/ex3.py

@@ -1,23 +1,0 @@
-print("I will now count my chickens:")
-
-print("Hens", 25 + 30 / 6)
-print("Roosters", 100 - 25 * 3 % 4)
-
-print("I will now count the eggs:")
-
-print(3 + 2 + 1 - 5 + 4 % 2 - 1 / 4 + 6)
-
-print("Is it true that 3 + 2 < 5 - 7?")
-
-print(3 + 2 < 5 - 7)
-
-print("What is 3 + 2?", 3 + 2)
-print("What is 5 - 7?", 5 - 7)
-
-print("Oh that's why it's False.")
-
-print("How about some more.")
-
-print("Is it greater?", 5 > -2)
-print("Is it greater or equal?", 5 >= -2)
-print("Is it less or equal?", 5 <=2)

### 2018-07-19 Python/LPTHW/ex4.py

@@ -1,17 +1,0 @@
-cars = 100
-space_in_a_car = 4.0
-drivers = 30
-passengers = 90
-cars_not_driven = cars - drivers
-cars_driven = drivers
-carpool_capacity = cars_driven * space_in_a_car
-average_passengers_per_car = passengers / cars_driven
-
-
-print("There are", cars, "cars available.")
-print("Thera are only", drivers, "drivers available.")
-print("There will be", cars_not_driven, "empty cars today.")
-print("We can transport", carpool_capacity, "people today.")
-print("We have", passengers, "to carpool today.")
-print("We need to put about", average_passengers_per_car,
-      "in each car.")

### 2018-07-19 Python/LPTHW/ex5.py

@@ -1,18 +1,0 @@
-my_name = 'Zed A. Shaw'
-my_age = 35 # not a lie
-my_height = 74 # inches
-my_weight = 180 # lbs
-my_eyes = 'Blue'
-my_teeth = 'White'
-my_hair = 'Brown'
-
-print(f"Let's talk about {my_name}.")
-print(f"He's {my_height} inches tall.")
-print(f"He's {my_weight} pounds heavy.")
-print("Actually that's not too heavy.")
-print(f"He's got {my_eyes} eyes and {my_hair} hair.")
-print(f"His hair are usually {my_teeth} depending on the coffee.")
-
-# this line is tricky, try to get it exactly right.
-total = my_age + my_height + my_weight
-print(f"If I add {my_age}, {my_height}, and {my_weight} I get {total}.")

### 2018-07-19 LiteratureReview/MeasuresOfLinguisticComplexity.tex

--- 2018-05-02 LiteratureReview/LiteratureReview.tex
@@
 \documentclass[a4paper]{article}
-
 %% Packages
 \input{settings/packages}
-
 %% Page Settings
 \input{settings/page}
-
 %% Own Commands
 \input{settings/macros}
-
-\title{Literature Review}
-\date{2018-04-26}
+\title{Measures of Linguistic Complexity: A literature Review}
+\date{2018-07-13}
 \author{Jomar Alcantara}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newpage
+\section{Introduction}
+\section{Measures of Lexical Richness}
+\subsection{Type token ratio(TTR)}
+Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
+\begin{equation} \label{x1}
+    TTR = number of unique words / total number of words.
+\end{equation}
 
-\section{Introduction}
-
-\par
-According to a recent report commissioned by the Alzheimer's Society, they estimate the prevalence of Dementia in the UK at approximately 815,000 people.
-This represents 1 in 14 of the population aged 65 or over.
-This report also estimates an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis.
-They also estimate that the overall impact of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually.
-Globally, this picture is a lot bleaker.
-A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
-The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillon US dollars.
 \newline
 \par
-The purpose of this review is to seek to understand how dementia impacts speech and language production from a psychological and linguistic perspective with a view to potentially using natural language processing and technology to aid in the early diagnosis of dementia and/or Mild Cognitive Impairment.
-\newline
-\par
-A search of the literature was conducted mainly from a psychological perspective as this would inform the theoretical side of things.
-Next, a search of the literature from a Computer Science perspective was also made with a view to exploring the how domain of machine learning and more specifically natural language processing can be applied to this problem.
-Finally a search was carried out to see what work has already been done in this area.
-
-All stages of this literature review were completed by myself.
-
-\section{Description of Dementia}
-
-Language - DSM-V
-'Major' Has significant difficulties with expressive or receptive language.
-Often uses general-use phrases such as "that thing" and "you know what I mean," and prefers general pronouns rather than names.
-With severe impairment, may not even recall names of closer friends and family.
-Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur.
-Stereotypy of speech occurs; echolalia and automatic speech typically precede mutism.
-
-'Mild' Has noticeable word-finding difficulty.
-May substitute general for specific terms.
-May avoid use of specific names of acquaintances.
-Grammatical errors involve subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc.
-
-\subsection{Alzheimer's Disease}
-\subsection{Vascular Dementia}
-\subsection{Dementia with Lewy Bodies}
-\subsection{Parkinson's disease dementia}
-\subsection{Fronto-temporal dementia}
-\subsection{Primary Progressive Aphasia}
-\subsection{Mild Cognitive Impairment}
-
-\section{Assessment of Language function in Dementia}
-
-Language can be defined as the ability to encode ideas into words and/or symbols for the means of communication.
-Difficulty in language, both spoken and written, are often described as symptoms of various types of Alzheimers' disease (AD).
-One of the challenges when using speech as a predictor is recognising whether there is a problem with language production or with the motor skills.
-Impairments in speech that arise from any process that disrupts the neuraxis from the cortex to muscle and encompass dysarthria (disturbance in articulation) and dysphonia (disturbance in the production in vocal sounds).
-However it is important to use language as a tool to aid the diagnosis of Dementia, as it provides vital clues to aid a clinician in differentiating in the different types of Dementia, which will in turn aid a clinician in an attempt to manage an individual case of Dementia.
-Due to the subtle nature of the languages changes that are experienced in those who suffer with different types of Dementia, it is often misdiagnosed.
-
-Emery \cite{Emery2000}, states that you can see language decline as Hierarchical.
-According to a semiotic model there are four ranks of language based on their cognitive complexity.
-These ranks are Phonology, Morphology, Syntax and Semantics.
-She conducted a review studies which looked at language decline from the perspective of these hierarchical ranks.
-Her conclusion is that language decline is related to the complexity of the language task given to a participant and that language decline is hierarchical in that the language forms we learn last (the most complex language forms) are the first to deteriorate.
-
-A common way for clinicians' to diagnose dementia is through use of the Mini-Mental State Examination.
-The Mini-Mental State Examination (MMSE) has shown to be particularly effective at differentiating betweeen Dementia and other psychological disorders such as clinical depression, schitzophrenia and personality disorder and is currently the most widely used measure of diagnosing Dementia within clinical psychology.
-However, it is not without it's faults...
-
-\section{Current ways of diagnosing Dementia using Language}
-\subsection{Free Cued Selective Recall Test}
-The Free Cued Selective Recall Test (FCSRT) was created by Grober and his team \cite{Grober1988}.
-
-\subsection{Mini-Mental State Examination}
-
-Battery of tests to differentiate between Controls, those with MCI, those with AD. 
-\begin{itemize}
-	\item FCSRDT - As most effective 
-	\item Mini Mental State Examination (MMSE)
-	\item Addenbrooke's Cognitive Examination (ACE-III)
-	\item Dementia Rating Scale (DRS)
-	\item Western Aphasia Battery (WAB)
-	\item Boston Diagnostic Aphasia Examination (BDAE)
-	\item Boston Naming Test (BNT)
-\end{itemize}
-
-\section{Existing studies that have used Natural Language Processing to help in diagnosing psychological difficulties, particularly language difficulties}
-
-\subsection{Nun Study - Snowdon (1996)}
-\subsection{Comparing Presidents Speeches - Berisha et al (2015)}
-
-\section{State of literature into Natural Language processing techniques}
-
+\subsection{Brunet's Index(W)}
+\subsection{Honore's Scale}
 \section{Conclusion}
-\par
-Identification of language impairment is important in Dementia because it aids diagnosis of specific types of dementia, which in turn can alter the prognosis and change the management of the degenerative disorder.
-As these differences in language are quite subtle, the varying subtypes of dementia are frequently misdiagnosed.
-\newline
-\par 
-Given the burden on the diagnosis of dementia on clinicians, it appears to be useful to find some non-invasive protocols for the early diagnosis of dementia.
-It has already been shown that analysis of speech and language has shown markers that pre-date the official diagnosis of dementia (Snowdon et al, 1996;)\cite{Berisha2015}.
-
 \bibliographystyle{unsrt}

### 2018-07-19 workingNotebook.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Document Class
\documentclass[a4paper]{book}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin {document}
\section {Things to do}
\begin {itemize}
    \item Learn how to use AWS or Google instances to run deep learning models
\end {itemize}
\section {Ideas to explore}
What if you extract sentences? So you have x list of sentences by controls as data. Each observation is 1 sentence. Each sentence the model reads adds to the probability AD or Not, MCI or Not. \newline
Can you hold meaning / context from previous sentences for repetition? Is this where LSTM's come in? \newline
Me collecting data? What does this look like? Then we can produce probability and confidence? This can then be tackled longditunially. \newline
Does this depend on language too much? Is there sufficient research to support acoustic analysis?
Experiment using this? In theory we migth be able to spot cognitive decline \newline
Read Berisha and Liss \newline
Can I write a paper comparing Reagan, Bush and Trump (use both original Berisha and Liss model and new deep learning model?)
If deficiencies in language are idiosyncratic, how can we generalise? How can we track changes over time?

\section {Bits of writing to go in QR, or Thesis}
\subsection {Potential Rationale for using deep learning over traditional machine learning methods.}
The recent resurgence in the use of neural networks and deep learning is because of it's significantly improved performance on many problems and it's ability to scale from small to large datasets. Another key benefit of deep learning is it's ability to automate the process of feature engineering. Feature engineering, with this particular problem is an interesting debate as there are currently numerous ways in which to try to generate features.

\subsection {On the decline of those with dementia} Whilst there are a wide number of factors that are involved in language generation, and therefore there will be an expected amount of variability between subjects. There is also an observation that as people get older, our language skills do decline. The consensus that this decline is typically exacerbated by the presence of dementia. 
\end {document}

### 2018-07-25 Python/PythonCrashCourse/cars.py
work done 24-07-18

@@ -18,3 +18,9 @@ cars.reverse()
 print(cars)
 
 print(len(cars))
+
+for car in cars:
+    if car == 'bmw':
+        print(car.upper())
+    else:
+        print(car.title())

### 2018-07-25 Python/PythonCrashCourse/foods.py

@@ -1,0 +1,20 @@
+my_foods = ['pizza', 'falafel', 'carrot cake']
+friend_foods = my_foods[:]
+
+print("My favourite foods are:")
+print(my_foods)
+
+print("My friend's favourite foods are:")
+print(friend_foods)
+
+my_foods = ['pizza', 'falafel', 'carrot cake']
+friend_foods = my_foods[:]
+
+my_foods.append('cannoli')
+friend_foods.append('ice cream')
+
+print("My favourite foods are:")
+print(my_foods)
+
+print("My friend's favourite foods are:")
+print(friend_foods)

### 2018-07-25 Python/PythonCrashCourse/players.py

@@ -1,0 +1,10 @@
+players = ['charles', 'martina', 'michael', 'florence', 'eli']
+print(players[0:3])
+print(players[1:4])
+print(players[:4])
+print(players[2:])
+print(players[-3:])
+
+print("Here are the first three players on my team.")
+for player in players[:3]:
+    print(player.title())

### 2018-07-25 Python/PythonCrashCourse/slices.py

@@ -1,0 +1,5 @@
+ingredients = ['onion', 'mushrooms', 'ham', 'pepperoni', 'peppers', 'bacon']
+print("The first three items on this list are:")
+print(ingredients[0:3])
+print(ingredients[2:5])
+print(ingredients[-3:])

### 2018-07-25 Python/PythonCrashCourse/alien.py
added work from 25-07-18 am

@@ -1,0 +1,44 @@
+alien_0 = {'colour' : 'green', 'points' : 5}
+
+print(alien_0['colour'])
+print(alien_0['points'])
+print(alien_0)
+
+new_points = alien_0['points']
+print("You just earned " + str(new_points) + " points!")
+
+alien_0['x_position'] = 0
+alien_0['y_position'] = 25
+print(alien_0)
+
+alien_0 = {}
+alien_0['colour'] = 'green'
+alien_0['points'] = 5
+print(alien_0)
+
+print("The colour of alien_0 is: " + alien_0['colour'] + ".")
+alien_0['colour'] = 'yellow'
+print("The colour of alien_0 is: " + alien_0['colour'] + ".")
+
+alien_0 = {'x_position': 0, 'y_position': 25, 'speed': 'medium'}
+print("Original x-position: " + str(alien_0['x_position']))
+
+# Move alien to the right
+# Determine how far to move the alien based on it's current speed.
+if alien_0['speed'] == 'slow':
+    x_increment = 1
+elif alien_0['speed'] == 'medium':
+    x_increment = 2
+else:
+    x_increment = 3
+
+# The new position of the alien is the old position + the increment.
+alien_0['x_position'] = alien_0['x_position'] + x_increment
+
+print("New x-position: " + str(alien_0['x_position']))
+
+alien_0 = {'colour': 'green', 'points': 5}
+print(alien_0)
+
+del alien_0['points']
+print(alien_0)

### 2018-07-25 Python/PythonCrashCourse/alien_colours1.py

@@ -1,0 +1,9 @@
+alien_colour = 'green'
+if alien_colour == 'green':
+    print("You just earned 5 points!")
+
+print("\nSecond Test")
+
+alien_colour = 'blue'
+if alien_colour == 'green':
+    print("You just earned 5 points!")

### 2018-07-25 Python/PythonCrashCourse/alien_colours2.py

@@ -1,0 +1,13 @@
+alien_colour = 'green'
+if alien_colour == 'green':
+    print("You just earned 5 points!")
+else:
+    print("You just earned 10 points!")
+
+print("\nSecond Test")
+
+alien_colour = 'blue'
+if alien_colour == 'green':
+    print("You just earned 5 points!")
+else:
+    print("You just earned 10 points!")

### 2018-07-25 Python/PythonCrashCourse/alien_colours3.py

@@ -1,0 +1,26 @@
+alien_colour = 'green'
+if alien_colour == 'green':
+    print("You just earned 5 points!")
+elif alien_colour == 'yellow':
+    print("You just earned 10 points!")
+else:
+    print("You just earned 15 points!")
+
+print("\nSecond Test")
+
+alien_colour = 'blue'
+if alien_colour == 'green':
+    print("You just earned 5 points!")
+elif alien_colour == 'yellow':
+    print("You just earned 10 points!")
+else:
+    print("You just earned 15 points!")
+
+print("\nThird Test")
+alien_colour = 'yellow'
+if alien_colour == 'green':
+    print("You just earned 5 points!")
+elif alien_colour == 'yellow':
+    print("You just earned 10 points!")
+else:
+    print("You just earned 15 points!")

### 2018-07-25 Python/PythonCrashCourse/aliens.py

@@ -1,0 +1,26 @@
+alien_0 = {'colour': 'green', 'points': 5}
+alien_1 = {'colour': 'yellow', 'points': 10}
+alien_2 = {'colour': 'red', 'points': 15}
+
+aliens = [alien_0, alien_1, alien_2]
+
+print(aliens)
+
+aliens = []
+
+for alien_number in range(30):
+    new_alien = {'colour': 'green', 'points': 5, 'speed': 'slow'}
+    aliens.append(new_alien)
+
+for alien in aliens[0:3]:
+    if alien['colour'] == 'green':
+        alien['colour'] = 'yellow'
+        alien['points'] = 10
+        alien['speed'] = 'medium'
+
+for alien in aliens[:5]:
+    print(alien)
+
+print("...")
+
+print("Total number of aliens " + str(len(aliens)))

### 2018-07-25 Python/PythonCrashCourse/amusement_park.py

@@ -1,0 +1,21 @@
+age = 12
+
+if age < 4:
+    print("Your admission cost is $0")
+elif age < 18:
+    print("Your admission cost is $5")
+else:
+    print("Your admission cost is $10")
+
+# More efficient version
+age = 12
+if age < 4:
+    price = 0
+elif age < 18:
+    price = 5
+elif age < 65:
+    price = 10
+else:
+    price = 5
+
+print("Your admission cost is $" + str(price) + ".")

### 2018-07-25 Python/PythonCrashCourse/checking_usernames.py

@@ -1,0 +1,8 @@
+current_users = ['admin', 'joe', 'steve', 'eloise', 'jenny']
+new_users = ['alice', 'STEVE', 'bob', 'charles', 'eloise']
+
+for user in new_users:
+    if user.lower() in current_users:
+        print("This username: " + user + " is not available.")
+    else: 
+        print("This username: " + user + " is available.")

### 2018-07-25 Python/PythonCrashCourse/conditional_tests.py

@@ -1,0 +1,34 @@
+age = 17
+print("Is age > 16, I predict True.")
+print(age > 16)
+
+print("Is age > 21, I predict False.")
+print(age > 21)
+
+print("Is age 17, I predict True.")
+print(age == 17)
+
+print("Is age greater or equal to 18, I predict False. ")
+print(age >= 18)
+
+car = 'Ford'
+print("Is car == 'ford', I predict False")
+print(car == 'ford')
+
+print("Is car.lower() == 'ford', I predict true")
+print(car.lower() == 'ford')
+
+print("Is car == 'Audi', I predict False")
+print(car == 'Audi')
+
+age1 = 12 
+age2 = 15
+
+print("Is age 1 > age 2, I predict False")
+print(age1 > age2)
+
+print("Is age2 > age 1, I predict True")
+print(age2 > age1)
+
+print("Is the sum of age 2 and age 1 < 30, I predict True")
+print(age1 + age2 < 30)

### 2018-07-25 Python/PythonCrashCourse/favourite_fruit.py

@@ -1,0 +1,8 @@
+fruits = ['Apple', 'Banana', 'Strawberry', 'Raspberry', 'Grape']
+
+if 'Apple' in fruits:
+    print("You really like Apples")
+if 'Strawberry' in fruits:
+    print("You really like Strawberries")
+if 'Raspberry' in fruits:
+    print("You really like Raspberries")

### 2018-07-25 Python/PythonCrashCourse/favourite_languages.py

@@ -1,0 +1,42 @@
+favourite_languages = {
+    'jen': 'python',
+    'sarah': 'c',
+    'edward': 'ruby',
+    'phil': 'python',
+    }
+
+print("Sarah's favourite language is " + 
+    favourite_languages['sarah'].title() + 
+    ".")
+
+for name, language in favourite_languages.items():
+    print(name.title() + "'s favouite language is: " + language.title() + ".")
+
+for name in favourite_languages.keys():
+    print(name.title())
+
+for name in favourite_languages:
+    print(name.title())
+
+friends = ['phil', 'sarah']
+for name in favourite_languages.keys():
+    print(name.title())
+
+    if name in friends:
+        print(" Hi " + name.title() + 
+            ", I see your favourite language is " +
+            favourite_languages[name].title() + "!")
+
+if 'erin' not in favourite_languages.keys():
+    print("Erin, please take our poll!")
+
+for name in sorted(favourite_languages.keys()):
+    print(name.title() + ", thank you for taking the poll.")
+
+print("The following languages have been mentioned:")
+for language in favourite_languages.values():
+    print(language.title())
+
+print("The following languages have been mentioned:")
+for language in set(favourite_languages.values()):
+    print(language.title())

### 2018-07-25 Python/PythonCrashCourse/favourite_numbers.py

@@ -1,0 +1,7 @@
+favourite_numbers = {'Joe': 7, 'Lisa': 42, 'Bob': 1, 'Klay': 11, 'Jenny':12}
+
+print("Joe's favourite number is " + str(favourite_numbers['Joe']))
+print("Lisa's favourite number is " + str(favourite_numbers['Lisa']))
+print("Bob's favourite number is " + str(favourite_numbers['Bob']))
+print("Klay's favourite number is " + str(favourite_numbers['Klay']))
+print("Jenny's favourite number is " + str(favourite_numbers['Jenny']))

### 2018-07-25 Python/PythonCrashCourse/glossary.py

@@ -1,0 +1,11 @@
+glossary = {'Print': 'Prints a statement',
+            'if..else': 'conditional statement',
+            'for' : 'A loop',
+            '.append': 'a function that appends an item to a list', 
+            '.sort': 'a function that sorts a list into alphabetical order.'}
+
+print("Print statement: " + glossary['Print'])
+print("If statement: " + glossary['if..else'])
+print("For statement: " + glossary['for'])
+print(".append function: " + glossary['.append'])
+print(".sort function: " + glossary['.sort'])

### 2018-07-25 Python/PythonCrashCourse/glossary2.py

@@ -1,0 +1,13 @@
+glossary = {'Print': 'Prints a statement',
+            'if..else': 'conditional statement',
+            'for' : 'A loop',
+            '.append': 'a function that appends an item to a list', 
+            '.sort': 'a function that sorts a list into alphabetical order.',
+            '.items': 'a function that gets all keys and values from a dictionary.',
+            '.keys': 'a function that gets just the keys from a dictionary.',
+            '.values': 'a function that gets just the values from a dictionary.',
+            '.sorted': 'a function that temporarily sorts a list',
+            '.title': 'a function that takes a string and turns it into a capitalised string.'}
+
+for k, v in glossary.items():
+    print(k.title() + ": " + v)

### 2018-07-25 Python/PythonCrashCourse/hello_admin.py

@@ -1,0 +1,10 @@
+# usernames = []
+usernames = ['admin', 'joe', 'steve', 'jenny', 'liz']
+if usernames:
+    for name in usernames:
+        if name == 'admin':
+            print("Hello admin, would you like a status report?")
+        else:
+            print("Hello " + name + ", you are now logged in.")
+else:
+    print("We need to find some users.")

### 2018-07-25 Python/PythonCrashCourse/magic_number.py

@@ -1,0 +1,4 @@
+answer = 17
+
+if answer != 42:
+    print("That is not the correct answer. Try again!")

### 2018-07-25 Python/PythonCrashCourse/more_conditional_tests.py

@@ -1,0 +1,39 @@
+str1 = 'bmw'
+str2 = 'BMW'
+
+print("Is str1 == str2, I predict False")
+print(str1 == str2)
+
+print("Is str1.upper == str2, I predict True")
+print(str1.upper() == str2)
+
+print("Is str1 == str2.lower(), I predict True")
+print(str1 == str2.lower())
+
+age1 = 12 
+age2 = 15
+
+print("Is age 1 > age 2, I predict False")
+print(age1 > age2)
+
+print("Is age2 > age 1, I predict True")
+print(age2 > age1)
+
+print("Is the sum of age 2 and age 1 < 30, I predict True")
+print(age1 + age2 < 30)
+
+print("Is age1 or age2 < 14, I predict True")
+print(age1 < 14 or age2 < 14)
+
+print("Is age1 and age 2 > 13, I predict False")
+print(age1 > 13 and age2 > 13)
+
+nums = []
+for i in range(1, 11):
+    nums.append(i**4)
+
+print('Is 7 a multiple of 4, I predict False')
+print(4 in nums)
+
+print('Is 16 a multiple of 4, I predict True')
+print(16 in nums)

### 2018-07-25 Python/PythonCrashCourse/ordinal_numbers.py

@@ -1,0 +1,11 @@
+nums = []
+for i in range (1, 11):
+    nums.append(i)
+    if i == 1:
+        print(str(i) + "st")
+    elif i==2:
+        print(str(i) + "nd")
+    elif i==3:
+        print(str(i) + "rd")
+    else:
+        print(str(i) + "th")

### 2018-07-25 Python/PythonCrashCourse/person.py

@@ -1,0 +1,6 @@
+joe = {'firstname': 'joe', 'lastname': 'alcantara', 'age': 41, 'city': 'birmingham'}
+
+print("First name: " + joe['firstname'].title())
+print("Last name: " + joe['lastname'].title())
+print("Age: " + str(joe['age']))
+print("City: " + joe['city'].title() )

### 2018-07-25 Python/PythonCrashCourse/pizza.py

@@ -1,0 +1,12 @@
+# Store information about a pizza being ordered.
+pizza = {
+    'crust': 'thick',
+    'toppings': ['mushrooms', 'extra cheese'],
+    }
+
+# Summarise the order.
+print("You ordered a " + pizza['crust'] + "-crust pizza" + 
+    " with the following toppings:")
+
+for topping in pizza['toppings']:
+    print("\t" + topping)

### 2018-07-25 Python/PythonCrashCourse/polling.py

@@ -1,0 +1,49 @@
+favourite_languages = {
+    'jen': 'python',
+    'sarah': 'c',
+    'edward': 'ruby',
+    'phil': 'python',
+    }
+
+print("Sarah's favourite language is " + 
+    favourite_languages['sarah'].title() + 
+    ".")
+
+for name, language in favourite_languages.items():
+    print(name.title() + "'s favouite language is: " + language.title() + ".")
+
+for name in favourite_languages.keys():
+    print(name.title())
+
+for name in favourite_languages:
+    print(name.title())
+
+friends = ['phil', 'sarah']
+for name in favourite_languages.keys():
+    print(name.title())
+
+    if name in friends:
+        print(" Hi " + name.title() + 
+            ", I see your favourite language is " +
+            favourite_languages[name].title() + "!")
+
+if 'erin' not in favourite_languages.keys():
+    print("Erin, please take our poll!")
+
+for name in sorted(favourite_languages.keys()):
+    print(name.title() + ", thank you for taking the poll.")
+
+print("The following languages have been mentioned:")
+for language in favourite_languages.values():
+    print(language.title())
+
+print("The following languages have been mentioned:")
+for language in set(favourite_languages.values()):
+    print(language.title())
+
+new_pollsters = ['jen', 'paul', 'phil', 'louisa', 'erin']
+for user in new_pollsters:
+    if user not in favourite_languages.keys():
+        print(user.title() + ", please can you take the poll?")
+    else: 
+        print(user.title() + ", thanks for taking the poll.")

### 2018-07-25 Python/PythonCrashCourse/rivers.py

@@ -1,0 +1,10 @@
+rivers = {'Nile': 'Egypt', 'Thames': "England", 'Seine': 'France'}
+
+for k, v in rivers.items():
+    print("The river " + k + " is in " + v)
+
+for k in rivers.keys():
+    print(k)
+
+for v in rivers.values():
+    print(v)

### 2018-07-25 Python/PythonCrashCourse/stages_of_life.py

@@ -1,0 +1,14 @@
+age = 66
+
+if age < 2:
+    print("You are a baby!")
+elif age < 4:
+    print("You are a toddler!")
+elif age < 13:
+    print("You are a kid!")
+elif age < 20:
+    print("You are a teenager!")
+elif age < 65:
+    print("You are an adult!")
+else:
+    print("You are an elder!")

### 2018-07-25 Python/PythonCrashCourse/toppings.py

@@ -2,3 +2,43 @@ requested_topping = 'mushrooms'
 
 if requested_topping != 'anchovies':
     print("Hold the anchovies!")
+
+requested_toppings = ['mushrooms', 'extra cheese']
+if 'mushrooms' in requested_toppings:
+    print('Adding Mushrooms')
+if 'pepperoni' in requested_toppings:
+    print('Adding Pepperoni')
+if 'extra cheese' in requested_toppings:
+    print('Adding Extra Cheese')
+
+print("\nFinished making your pizza!")
+
+requested_toppings = ['mushrooms', 'green peppers', 'extra cheese']
+for topping in requested_toppings:
+    if topping == 'green peppers':
+        print("Sorry we are out of green peppers right now.")
+    else:
+        print("Adding " + topping + ".")
+
+print("\nFinished making your pizza!")
+
+requested_toppings = []
+
+if requested_toppings:
+    for topping in requested_toppings:
+        print("Adding " + topping + ".")
+    print("\nFinished making your pizza!")
+else:
+    print("Are you sure you want a plain pizza?")
+
+available_toppings = ['mushrooms', 'olives', 'green peppers',
+                        'pepperoni', 'pineapple', 'extra cheese']
+requested_toppings = ['mushrooms', 'french fries', 'extra cheese']
+
+for topping in requested_toppings:
+    if topping in available_toppings:
+        print("Adding " + topping + ".")
+    else:
+        print("Sorry, we don't have " + topping + ".")
+
+print("\nFinished making your pizza.")

### 2018-07-25 Python/PythonCrashCourse/voting.py

@@ -1,0 +1,7 @@
+age = 17
+if age >= 18:
+    print("You are old enough to vote!")
+    print("Have you registered to vote yet?")
+else:
+    print("Sorry you're too young to vote.")
+    print("Please register to vote as soon as you turn 18.")

### 2018-07-25 Python/PythonCrashCourse/cities.py
Chapter 6 done

@@ -1,0 +1,23 @@
+cities = {
+    'london': {
+        'country': 'england',
+        'population': 8000000,
+        'famous site': 'tower of london',
+        },
+    'seattle': {
+        'country': 'united states',
+        'population': 5000000,
+        'famous site': 'space needle',
+        },
+    'paris': {
+        'country': 'france',
+        'population': 7000000,
+        'famous site': 'le louvre',
+        },
+    }
+
+for k, v in cities.items():
+    print(k.title())
+    print("Country: " + v['country'].title())
+    print("Population: " + str(v['population']))
+    print("Famous Site: " + v['famous site'].title())

### 2018-07-25 Python/PythonCrashCourse/favourite_languages.py

@@ -40,3 +40,15 @@ for language in favourite_languages.values():
 print("The following languages have been mentioned:")
 for language in set(favourite_languages.values()):
     print(language.title())
+
+favourite_languages = {
+    'jen': ['python', 'ruby'],
+    'sarah': ['C'],
+    'edward': ['ruby', 'go'],
+    'phil': ['python', 'haskell'],
+    }
+
+for name, languages in favourite_languages.items():
+    print("\n" + name.title() + "'s favourite languages are:")
+    for language in languages:
+        print("\t" + language.title())

### 2018-07-25 Python/PythonCrashCourse/favourite_numbers2.py

@@ -1,0 +1,6 @@
+favourite_numbers = {'Joe': [7], 'Lisa': [42, 69], 'Bob': [1, 13], 'Klay': [11, 23], 'Jenny': [12]}
+
+for k, v in favourite_numbers.items():
+    print(k + "'s favourite numbers are")
+    for i in v:
+        print("\t" + str(i))

### 2018-07-25 Python/PythonCrashCourse/favourite_places.py

@@ -1,0 +1,8 @@
+favourite_places = {'joe': ['London', 'Chicago', 'Seattle'],
+                'fred': ['Tokyo', 'London', 'Paris'],
+                'donald': ['Moscow', 'Washington', 'Wonderland']}
+
+for k, v in favourite_places.items():
+    print(k.title())
+    for item in v:
+        print("\t" + item)

### 2018-07-25 Python/PythonCrashCourse/many_users.py

@@ -1,0 +1,21 @@
+users = {
+    'aeinstein': {
+        'first': 'albert',
+        'last': 'einstein',
+        'location': 'princeton',
+        },
+    'mcurie': {
+        'first': 'marie',
+        'last': 'curie',
+        'location': 'paris',
+        },
+
+    }
+
+for username, user_info in users.items():
+    print("\nUsername: " + username)
+    full_name = user_info['first'] + " " + user_info['last']
+    location = user_info['location']
+
+    print("\tFull name: " + full_name.title())
+    print("\tLocation: " + location.title())

### 2018-07-25 Python/PythonCrashCourse/people.py

@@ -1,0 +1,11 @@
+joe = {'firstname': 'joe', 'lastname': 'alcantara', 'age': 41, 'city': 'birmingham'}
+italia = {'firstname': 'italia', 'lastname': 'ricci', 'age': 37, 'city' : 'toronto'}
+hailey = {'firstname': 'hailey', 'lastname': 'little', 'age': 21, 'city' : 'los angeles'}
+
+people = [joe, italia, hailey]
+
+for person in people:
+    full_name = person['firstname'] + ' ' + person['lastname']
+    print("Full Name: " + full_name)
+    print("Age: " + str(person['age']))
+    print("From: " + person['city'])

### 2018-07-25 Python/PythonCrashCourse/pets.py

@@ -1,0 +1,7 @@
+kentavious = {'species': 'cat', 'owner': 'joe'}
+rocco = {'species': 'dog', 'owner': 'klay'}
+jess = {'species': 'cat', 'owner': 'marina'}
+
+pets = [kentavious, rocco, jess]
+for pet in pets:
+    print(pet)

### 2018-08-06 Python/PythonCrashCourse/addition.py
work done week beg 30-7-18

@@ -1,0 +1,20 @@
+print("Addition program. \n")
+active = True
+while active == True:
+    num1 = input("Please enter a number : ")
+    num2 = input("Please enter another number: ")
+
+    try:
+        num1 = int(num1)
+        num2 = int(num2)
+    except ValueError:
+        print("You didn't enter a number")
+    else:
+        answer = num1 + num2
+        print("\tThe answer is : " + str(answer))
+
+    response = input("Do you want to continue? (yes / no) ")
+    if response.lower == 'yes':
+        active == True
+    else:
+        break

### 2018-08-06 Python/PythonCrashCourse/admin.py

@@ -1,0 +1,60 @@
+class User():
+    """ First attempt at describing a user """
+    
+    def __init__(self, fname, lname):
+        """ Initialise attributes """
+        self.fname = fname
+        self.lname = lname
+        self.login_att = 0
+
+    def describe_user(self):
+        print("First Name : " + self.fname)
+        print("Last Name : " + self.lname)
+
+    def greet_user(self):
+        print("Hello, " + self.fname.title() + " " + self.lname.title() + ".")
+    
+    def increment_login_attempts(self):
+        self.login_att += 1
+    
+    def reset_login_attempts(self):
+        self.login_att = 0
+
+class Admin(User):
+    """ Models a special type of user - Admin """
+    
+    def __init__(self, fname, lname, *privileges):
+        self.fname = fname
+        self.lname = lname
+        self.login_att = 0
+        self.privileges = privileges
+
+    def list_privileges(self):
+        for privilege in self.privileges:
+            print(privilege)
+
+user1 = User("Joe", "Alcantara")
+user2 = User("Lexi", "Belle")
+user3 = User("Hailey", "Little")
+
+user1.describe_user()
+user1.greet_user()
+user2.describe_user()
+user2.greet_user()
+user3.describe_user()
+user3.greet_user()
+
+user3.increment_login_attempts()
+print(user3.login_att)
+user3.increment_login_attempts()
+print(user3.login_att)
+user3.increment_login_attempts()
+print(user3.login_att)
+user3.increment_login_attempts()
+print(user3.login_att)
+user3.reset_login_attempts()
+print(user3.login_att)
+
+user4 = Admin("Carter", "Cruise", "Can ban user", "Can delete post", "Can add post")
+user4.describe_user()
+user4.list_privileges()

### 2018-08-06 Python/PythonCrashCourse/admin2.py

@@ -1,0 +1,65 @@
+class User():
+    """ First attempt at describing a user """
+    
+    def __init__(self, fname, lname):
+        """ Initialise attributes """
+        self.fname = fname
+        self.lname = lname
+        self.login_att = 0
+
+    def describe_user(self):
+        print("First Name : " + self.fname)
+        print("Last Name : " + self.lname)
+
+    def greet_user(self):
+        print("Hello, " + self.fname.title() + " " + self.lname.title() + ".")
+    
+    def increment_login_attempts(self):
+        self.login_att += 1
+    
+    def reset_login_attempts(self):
+        self.login_att = 0
+
+class Privileges():
+    def __init__(self, *privileges):
+        self.privileges = privileges
+
+    def list_privileges(self):
+        for privilege in self.privileges:
+            print(privilege)
+
+class Admin(User):
+    """ Models a special type of user - Admin """
+    
+    def __init__(self, fname, lname):
+        super().__init__(fname, lname)
+        self.fname = fname
+        self.lname = lname
+        self.login_att = 0
+        self.privileges = Privileges("Can ban user", "Can delete post", "Can add post")
+
+user1 = User("Joe", "Alcantara")
+user2 = User("Lexi", "Belle")
+user3 = User("Hailey", "Little")
+
+user1.describe_user()
+user1.greet_user()
+user2.describe_user()
+user2.greet_user()
+user3.describe_user()
+user3.greet_user()
+
+user3.increment_login_attempts()
+print(user3.login_att)
+user3.increment_login_attempts()
+print(user3.login_att)
+user3.increment_login_attempts()
+print(user3.login_att)
+user3.increment_login_attempts()
+print(user3.login_att)
+user3.reset_login_attempts()
+print(user3.login_att)
+
+user4 = Admin("Carter", "Cruise")
+user4.describe_user()
+user4.privileges.list_privileges()

### 2018-08-06 Python/PythonCrashCourse/alice.py

@@ -1,0 +1,13 @@
+filename = 'alice.txt'
+
+try:
+    with open(filename) as file_object:
+        contents = file_object.read()
+except FileNotFoundError:
+    msg = "Sorry, " + filename + " does not exist."
+    print(msg)
+else:
+    # Count the approximate number of words in the file
+    words = contents.split()
+    num_words = len(words)
+    print("The file " + filename + " has about " + str(num_words) + " words.")

### 2018-08-06 Python/PythonCrashCourse/alienInvasion/alien_invasion.py

@@ -1,0 +1,31 @@
+import sys
+import pygame
+
+from settings import Settings
+
+def run_game():
+    # Initialize game, settings and screen object.
+    pygame.init()
+    ai_settings = Settings()
+    screen = pygame.display.set_mode(
+        (ai_settings.screen_width, ai_settings.screen_height))
+    pygame.display.set_caption("Alien Invasion")
+
+    # Set background colour.
+    bg_colour = (230, 230, 230)
+
+    # Start main loop for the game.
+    while True:
+
+        # Watch for Keyboard and Mouse Events:
+        for event in pygame.event.get():
+            if event.type == pygame.QUIT:
+                sys.exit()
+
+        # Redraw the screen during each pass through the loop
+        screen.fill(ai_settings.bg_colour)
+
+        # Make the most recently drawn screen visible.
+        pygame.display.flip()
+
+run_game()

### 2018-08-06 Python/PythonCrashCourse/alienInvasion/settings.py

@@ -1,0 +1,9 @@
+class Settings():
+    """ A class to store all settings for Alien Invasion """
+
+    def __init__ (self):
+        """ Initialise the game's settings """
+        # Screen Settings
+        self.screen_width = 1200
+        self.screen_height = 800
+        self.bg_colour = (230, 230, 230)

### 2018-08-06 Python/PythonCrashCourse/car.py

@@ -1,0 +1,66 @@
+""" A class that can be used to represent a car """
+class Car():
+    """ A simple attempt to represent a car """
+
+    def __init__ (self, make, model, year):
+        """ Initialise attribute for the car """
+        self.make = make
+        self.model = model 
+        self.year = year
+        self.odometer_reading = 0
+
+    def get_descriptive_name(self):
+        """ Returns a neatly formatted descriptive name """
+        long_name = str(self.year) + " " + self.make + " " + self.model
+        return long_name.title()
+
+    def read_odometer(self):
+        """ Print a statement showing the car's mileage """
+        print("This car has " + str(self.odometer_reading) + " miles on it.")
+
+    def update_odometer(self, mileage):
+        """
+        Set the odometer to a given value.
+        Reject the change if it attempts to roll the odometer back
+        """
+        if mileage >= self.odometer_reading:
+            self.odometer_reading = mileage
+        else:
+            print("You can't roll back an odometer!")
+    
+    def increment_odometer(self, miles):
+        """ Add the given amount to the odometer reading. """
+        self.odometer_reading += miles
+
+class Battery():
+    """ A simple attempt to model a battery for an electric car."""
+
+    def __init__(self, battery_size = 60):
+        """ Initialize the battery's attributes """
+        self.battery_size = battery_size
+
+    def describe_battery(self):
+        """ Print a statement descibing the battery size """
+        print("This car has a " + str(self.battery_size) + "-kWh battery.")
+
+    def get_range(self):
+        """ Print a statement describing the range of the car based on battery size """
+        if self.battery_size == 60:
+            range = 240
+        elif self.battery_size == 85:
+            range = 270
+
+        message = "This car can go approximately " + str(range)
+        message += " miles on a full charge."
+        print(message) 
+
+class ElectricCar(Car):
+    """ Models aspects of a car, specific to an electric car. """
+
+    def __init__(self, make, model, year):
+        """ 
+        Initializes attributes of the parent class.
+        Then initializes attributes specific to an electric car.
+        """
+        super().__init__(make, model, year)
+        self.battery = Battery()

### 2018-08-06 Python/PythonCrashCourse/cats_and_dogs.py

@@ -1,0 +1,13 @@
+filenames = ["cats1.txt", "dogs.txt"]
+
+for filename in filenames:
+    try:
+        with open(filename) as fileobject:
+            contents = fileobject.read()
+    except FileNotFoundError:
+        # print(filename + " - file is missing.")
+        # print("#" * 12)
+        pass
+    else:
+        print(contents)
+        print("*" * 12)

### 2018-08-06 Python/PythonCrashCourse/city-country.py

@@ -1,0 +1,3 @@
+def city(city, country):
+    """ Takes city and country and outputs this neatly """
+    fullString = city.title() + ", " + country_title()

### 2018-08-06 Python/PythonCrashCourse/city_functions.py

@@ -1,0 +1,7 @@
+def city(city, country, population=''):
+    """ Takes city and country and outputs this neatly """
+    if population:
+        fullString = city.title() + ", " + country.title() + " - " + str(population)
+    else:
+        fullString = city.title() + ", " + country.title()
+    return fullString

### 2018-08-06 Python/PythonCrashCourse/division.py

@@ -1,0 +1,16 @@
+print("Give me two numbers and I'll divide them.")
+print("Enter 'q' to quit")
+
+while True:
+    firstNumber = input("\n\tFirst Number : ")
+    if firstNumber == 'q':
+        break
+    secondNumber = input("\n\tSecond Number : ")
+    if secondNumber == 'q':
+        break
+    try:
+        answer = int(firstNumber) / int(secondNumber)
+    except ZeroDivisionError:
+        print("You can't divide by zero!")
+    else:
+        print(answer)

### 2018-08-06 Python/PythonCrashCourse/electric_car.py

@@ -1,6 +1,5 @@
 class Car():
     """ A simple attempt to represent a car """
-
     def __init__(self, make, model, year):
         self.make = make
         self.model = model
@@ -23,17 +22,44 @@ class Car():
     def increment_odometer(self, miles):
         self.odometer_reading += miles
 
+class Battery():
+    """ A simple attempt to represent an electric car battery. """
+    def __init__(self, battery_size=70):
+        self.battery_size = battery_size
+
+    def describe_battery(self):
+        print("This car has a " + str(self.battery_size) + "-kWh battery.")
+
+    def get_range(self):
+        """Print a statement about the range this battery provides."""
+        if self.battery_size == 70:
+            range = 240
+        elif self.battery_size == 85:
+            range = 270
+            
+        message = "This car can go approximately " + str(range)
+        message += " miles on a full charge."
+        print(message)
+    
+    def upgrade_battery(self):
+        """ Upgrades the battery, if not already upgraded """
+        if self.battery_size == 70:
+            self.battery_size = 85
+        else:
+            print("Battery already upgraded!")
+
 class ElectricCar(Car):
     """ Represent aspects of a car, specific to electric vehicles """
     def __init__(self, make, model, year):
         """ Initialize attributes of the parent class """
         super().__init__(make, model, year)
-        self.battery_size = 70
-    
-    def describe_battery(self):
-        """ Print a statement describing battery size. """
-        print("This car has a " + str(self.battery_size) + "-kWh battery.")
+        self.battery = Battery()
 
 my_tesla = ElectricCar('tesla', 'model s', 2016)
 print(my_tesla.get_descriptive_name())
-my_tesla.describe_battery()
+my_tesla.battery.describe_battery()
+my_tesla.battery.get_range()
+my_tesla.battery.upgrade_battery()
+my_tesla.battery.describe_battery()
+my_tesla.battery.get_range()
+my_tesla.battery.upgrade_battery()

### 2018-08-06 Python/PythonCrashCourse/employee.py

@@ -1,0 +1,9 @@
+class Employee():
+    """ A simple model of an employee """
+    def __init__(self, fname, lname, salary):
+        self.fname = fname.title()
+        self.lname = lname.title()
+        self.salary = salary
+
+    def give_raise(self, amount=5000):
+        self.salary += amount

### 2018-08-06 Python/PythonCrashCourse/favourite_number3.py

@@ -1,0 +1,8 @@
+import json
+
+filename = 'favourite.json'
+with open(filename, 'w') as f_obj:
+    fav_num = input("What is your favourite number? ")
+    json.dump(fav_num, f_obj)
+
+

### 2018-08-06 Python/PythonCrashCourse/file_reader.py

@@ -1,0 +1,38 @@
+with open ('pi_digits.txt') as file_object:
+    contents = file_object.read()
+    print(contents)
+
+filename = 'pi_digits.txt'
+with open(filename) as file_object:
+    for line in file_object:
+        print(line.rstrip())
+
+with open(filename) as file_object:
+    lines = file_object.readlines()
+
+for line in lines:
+    print(line.rstrip())
+
+pi_string = ''
+for line in lines:
+    pi_string += line.strip()
+
+print(pi_string)
+print(len(pi_string))
+
+filename = 'pi_million_digits.txt'
+with open(filename) as file_object:
+    lines = file_object.readlines()
+
+pi_string = ''
+for line in lines:
+    pi_string += line.strip()
+
+print(pi_string[:52] + "...")
+print(len(pi_string))
+
+birthday = input("Enter your birthday, in the form mmddyy:")
+if birthday in pi_string:
+    print("Your birthday appears in the first million digits of pi!")
+else: 
+    print("Your birthday doesn't appear in the first million digits of pi.")

### 2018-08-06 Python/PythonCrashCourse/greet_user.py

@@ -1,0 +1,7 @@
+import json
+
+filename = 'username.json'
+
+with open(filename) as f_obj:
+    username = json.load(f_obj)
+    print("Welcome Back, " + username + "!")

### 2018-08-06 Python/PythonCrashCourse/guest.py

@@ -1,0 +1,6 @@
+filename = 'guest.txt'
+
+guest = input("Please enter your name :")
+
+with open(filename, 'w') as file_object:
+    file_object.write(guest)

### 2018-08-06 Python/PythonCrashCourse/guest_book.py

@@ -1,0 +1,12 @@
+filename = 'guest_book.txt'
+guest = 'Blank'
+while guest != '':
+    guest = input("Please enter your name :")
+    if guest == '':
+        break
+    
+    print('Hello, ' + guest + '. Sign our guestbook!')
+    message = (guest.title() + ' was ere.\n')
+
+    with open(filename, 'a') as file_object:
+        file_object.write(message)

### 2018-08-06 Python/PythonCrashCourse/ice_cream_stand.py

@@ -1,0 +1,44 @@
+class Restaurant():
+    """ First attempt to model a restaurant """
+
+    def __init__(self, restaurant_name, cuisine_type):
+        """ Initialise variables for Restaurant """
+        self.restaurant_name = restaurant_name
+        self.cuisine_type = cuisine_type
+
+    def describe_restaurant(self):
+        """ Describes the restaurant """
+        print("Restaurant Name : " + self.restaurant_name)
+        print("Cuisine Type : " + self.cuisine_type)
+
+    def restaurant_open(self):
+        """ Signals that the restaurant is open """
+        print("The " + self.restaurant_name + " restaurant is now open for business")
+
+class IceCreamStand(Restaurant):
+    """ Modelling an ice cream stand """
+
+    def __init__(self, restaurant_name, cuisine_type, *flavours):
+        """ Initialise variables for Ice Cream Stand """
+        self.restaurant_name = restaurant_name
+        self.cuisine_type = cuisine_type
+        self.flavours = flavours
+
+    def menu(self):
+        print("The ice cream flavours on offer are :")
+        for flavour in self.flavours:
+            print("\t" + flavour) 
+
+test = Restaurant("'Nando's", "Chicken")
+test.describe_restaurant()
+test.restaurant_open()
+
+test2 = Restaurant("KFC", "Chicken")
+test3 = Restaurant("Carluccio's", "Italian")
+
+test2.describe_restaurant()
+test3.describe_restaurant()
+
+test4 = IceCreamStand("Joe's", "Ice-Cream", "Vanilla", "Chocolate", "Raspberry")
+test4.describe_restaurant()
+test4.menu()

### 2018-08-06 Python/PythonCrashCourse/learning_python.py

@@ -1,0 +1,20 @@
+with open('learning_python.txt') as file_object:
+    contents = file_object.read()
+    print(contents)
+    print("*" * 8)
+
+with open('learning_python.txt') as file_object:
+    for line in file_object:
+        print(line.rstrip())
+
+with open('learning_python.txt') as file_object:
+    lines = file_object.readlines()
+
+print("*" * 8)
+for line in lines:
+    print(line.rstrip())
+
+print("*" * 8)
+for line in lines:
+    newline = line.replace("Python", "C")
+    print(newline.rstrip())

### 2018-08-06 Python/PythonCrashCourse/my_car.py

@@ -1,0 +1,7 @@
+from car import Car
+
+my_new_car = Car("Audi", "A6", "2016")
+print(my_new_car.get_descriptive_name())
+
+my_new_car.odometer_reading = 23
+my_new_car.read_odometer()

### 2018-08-06 Python/PythonCrashCourse/my_cars.py

@@ -1,0 +1,7 @@
+from car import Car, ElectricCar
+
+my_beetle = Car("volkswagen", "beetle", 2016)
+print(my_beetle.get_descriptive_name())
+
+my_tesla = ElectricCar("tesla", "model s", 2016)
+print(my_tesla.get_descriptive_name())

### 2018-08-06 Python/PythonCrashCourse/my_electric_car.py

@@ -1,0 +1,7 @@
+from car import ElectricCar
+
+my_tesla = ElectricCar("tesla", 'model s', '2016')
+
+print(my_tesla.get_descriptive_name())
+my_tesla.battery.describe_battery()
+my_tesla.battery.get_range()

### 2018-08-06 Python/PythonCrashCourse/name_function.py

@@ -1,0 +1,7 @@
+def get_formatted_name(first, last, middle = ''):
+    """ Generated a neatly formatted name """
+    if middle:
+        full_name = first + " " + middle + " " + last
+    else:
+        full_name = first + " " + last
+    return full_name.title()

### 2018-08-06 Python/PythonCrashCourse/names2.py

@@ -1,0 +1,13 @@
+from name_function import get_formatted_name
+
+print("Enter 'q' at any time to quit.")
+while True:
+    first = input("\nPlease give me a first name : ")
+    if first == 'q':
+        break
+    last = input("\nPlease give me a last name : ")
+    if last == 'q':
+        break
+
+    formatted_name = get_formatted_name(first, last)
+    print("\tNeatly formatted name " + formatted_name + ".")

### 2018-08-06 Python/PythonCrashCourse/number_reader.py

@@ -1,0 +1,7 @@
+import json
+
+filename = 'numbers.json'
+with open(filename) as fileobject:
+    numbers = json.load(fileobject)
+
+print(numbers)

### 2018-08-06 Python/PythonCrashCourse/number_reader2.py

@@ -1,0 +1,7 @@
+import json
+
+filename = 'favourite.json'
+
+with open(filename) as f_obj:
+    fav_num = json.load(f_obj)
+    print("Your favourite number is " + str(fav_num) + ".")

### 2018-08-06 Python/PythonCrashCourse/number_writer.py

@@ -1,0 +1,7 @@
+import json
+
+numbers = [2, 3, 5, 7, 11, 13]
+
+filename = 'numbers.json'
+with open(filename, 'w') as fileobject:
+    json.dump(numbers, fileobject)

### 2018-08-06 Python/PythonCrashCourse/programming_poll.py

@@ -1,0 +1,16 @@
+filename = 'response.txt'
+with open(filename, 'w') as file_object:
+    file_object.write("Reasons why:\n")
+
+active = True
+while active != False:
+    response = input("Why do you like programming?")
+
+    with open(filename, 'a') as file_object:
+        file_object.write(response + "\n")
+
+    willContinue = input ("Would you like to continue? (yes / no) :")
+    if willContinue.lower() == 'yes':
+        active = True
+    else:
+        active = False

### 2018-08-06 Python/PythonCrashCourse/remember_me.py

@@ -1,0 +1,35 @@
+import json
+# Load the username, if it has been stored previously.
+# Otherwise, prompt for the username and store it.
+def get_stored_username():
+    filename = 'username.json'
+    try:
+        with open(filename) as f_obj:
+            username = json.load(f_obj)
+    except FileNotFoundError:
+        return None
+    else:
+        return username
+
+def get_new_username():
+    """ Prompt user for a username """ 
+    username = input("What is your name? ")
+    filename = 'username.json'
+    with open(filename, 'w') as f_obj:
+        json.dump(username, f_obj)
+    return username        
+
+def greet_user():
+    """ Greet user by name """
+    username = get_stored_username()
+    if username:
+        check = input("Is your username - " + username + " ? (y/n)")
+        if check == 'y':
+            print("Welcome back, " + username + "!")
+        elif check == 'n':
+            get_new_username()
+    else:
+        username = get_new_username()
+        print("We'll remember you when you come back, " + username + "!")
+
+greet_user()

### 2018-08-06 Python/PythonCrashCourse/survey.py

@@ -1,0 +1,21 @@
+class AnonymousSurvey():
+    """ Collect anonymous answer to a survey question. """
+
+    def __init__ (self, question):
+        """ Store a question, and prepare to store responses. """
+        self.question = question
+        self.responses = []
+
+    def show_question(self):
+        """ Show the survey question """
+        print(self.question)
+
+    def store_response(self, new_response):
+        """ Store a response to the survey """
+        self.responses.append(new_response)
+
+    def show_results(self):
+        """ Show all the responses that have been given """
+        print("Survey Results")
+        for response in self.responses:
+            print("- " + response)

### 2018-08-06 Python/PythonCrashCourse/test_cities.py

@@ -1,0 +1,20 @@
+import unittest
+
+from city_functions import city
+
+class TestCitiesFunctions (unittest.TestCase):
+    
+    def test_city(self):
+        print("Start Test")
+        formatted_test = city('Santiago', 'Chile')
+        self.assertEqual(formatted_test, 'Santiago, Chile')
+
+    def test_city2(self):
+        formatted_test2 = city('santiago', 'chile')
+        self.assertEqual(formatted_test2, 'Santiago, Chile')
+
+    def test_city_population(self):
+        formatted_test3 = city('santiago', 'chile', '5000000')
+        self.assertEqual(formatted_test3, 'Santiago, Chile - 5000000')
+
+unittest.main()

### 2018-08-06 Python/PythonCrashCourse/test_employee.py

@@ -1,0 +1,18 @@
+import unittest
+from employee import Employee
+
+class TestEmployeeCases(unittest.TestCase):
+    """ Tests for the Employee Classes"""
+    def setUp(self):
+        self.my_employee = Employee("Joe", "Alcantara", 0)
+        self.raises = ['', 5500]
+
+    def test_default_raise(self):
+        self.my_employee.give_raise()
+        self.assertEqual(self.my_employee.salary, 5000)
+    
+    def test_custom_raise(self):
+        self.my_employee.give_raise(5500)
+        self.assertEqual(self.my_employee.salary, 5500)
+
+unittest.main()

### 2018-08-06 Python/PythonCrashCourse/test_name_function.py

@@ -1,0 +1,19 @@
+import unittest
+from name_function import get_formatted_name
+
+class NamesTestCase(unittest.TestCase):
+    """ Tests for name.function.py """
+
+    def test_first_last_name(self):
+        """ Do names like Janis Joplin work """
+        print("*** Start Test ***")
+        formatted_name = get_formatted_name("janis", "joplin")
+        self.assertEqual(formatted_name, 'Janis Joplin')
+
+    def test_first_middle_last(self):
+        """ Do names like Barrack Hussein Obama work """
+        print("*** Start Test ***")
+        formatted_name = get_formatted_name("Barrack", "Obama", "Hussein")
+        self.assertEqual(formatted_name, 'Barrack Hussein Obama')
+
+unittest.main()

### 2018-08-06 Python/PythonCrashCourse/test_survey.py

@@ -1,0 +1,28 @@
+import unittest
+from survey import AnonymousSurvey
+
+class TestAnonymousSurvey(unittest.TestCase):
+    """ Tests for Anonymous Survey """
+    def setUp(self):
+        """
+        Create a survey and a set of responses for use in all test methods.
+        """
+        question = "What language did you first learn to speak"
+        self.my_survey = AnonymousSurvey(question)
+        self.responses = ["English", "Spanish", "Mandarin"]
+
+    def test_store_single_response(self):
+        """ Test that a single response is stored properly. """
+        self.my_survey.store_response(self.responses[0])
+        self.assertIn(self.responses[0], self.my_survey.responses)
+
+    def test_store_three_responses(self):
+        """ Test that three responses are stored properly. """
+        for response in self.responses:
+            self.my_survey.store_response(response)
+
+        for response in self.responses:
+            self.assertIn(response, self.my_survey.responses)
+
+
+unittest.main()

### 2018-08-06 Python/PythonCrashCourse/word_count.py

@@ -1,0 +1,21 @@
+def count_words(filename, word):
+    """ Count the approximate number of words in a file. """
+    try:
+        with open(filename) as fileobject:
+            contents = fileobject.read()
+    except FileNotFoundError:
+        pass
+        # msg = ("Sorry, the file " + filename + " does not exist." )
+        # print(msg)
+    else:
+        # Count the approximate number of words in a file.
+        words = contents.split()
+        word_count = words.count(word)
+        print("The number of time the word " + word + " is in this text is : " + str(word_count))
+        num_words = len(words)
+        print("The file " + filename + " contains approximately " + 
+        str(num_words) + " words.")
+
+filenames = ['alice.txt', 'siddartha.txt', 'moby_dick.txt', 'little_women.txt']
+for filename in filenames:
+    count_words(filename, "the")

### 2018-08-06 Python/PythonForDeepLearning/test2.py

@@ -1,0 +1,13 @@
+from keras.datasets import mnist
+
+(train_images, train_labels), (test_images, test_labels) = mnist.load_data()
+
+print(train_images.ndim)
+print(train_images.shape)
+print(train_images.dtype)
+
+digit = train_images[4]
+
+import matplotlib.pyplot as plt
+plt.imshow(digit, cmap = plt.cm.binary)
+plt.show()

### 2018-08-06 ExperimentProposals/Experiments.tex
work done 19/07-18
work completed 06/08/18am

@@
 
-\title{Experimental Report}
-\date{2018-04-22}
+\title{Comparing the presidential news conferences of Ronald Reagan, George H.W.
+Bush and Donald Trump - A exploration of deep learning for the purposes of tracking language changes over time.\newline \\\large Experimental Proposal 1 }
+\date{2018-07-19}
 \author{Jomar Alcantara}
@@
 \maketitle
+\newpage
+\tableofcontents
+\newpage
+\section{Abstract}
+\section{Introduction}
 
+Berisha and Liss compared the press conferences of Ronald Reagan (RR) and George Herbert Walker Bush (GHWB) and found that there were significant markers found that were able to distinguish between RR and GHWB in terms of cognitive impairment.
+The rationale for comparing these two presidents was that they were the closest in age to each other.
+With the appointment of Donald Trump (DT) as president of the United States, this affords us an opportunity to compare and contrast Berisha and Liss's results with data from a new and arguably more comparible subject in DT.\newline
+\par 
+One of the criticisms of the current literature is the amount of approaches that are taken in the pursuit of features / attributes which can be used to predict cognitive impairment.
+One of the key dependencies that traditional machine learning algorithms have is on the features that are given to the algorithm.
+Therefore, in some sense, the success or failure of a traditional machine learning algorithm lies in part on the quality of the features given to it.
+\subsection{Hypotheses}
+Our aim is to look at the press conferences of RR, GHWB and DT.
+These are widely available, unedited, on a website (CITE).
+These press conferences consists of two parts.
+The first part is a prepared statement which is usually, in part, at least co-written by a member of White House staff, typically a speech writer.
+The second part is a Question and Answer session, in which the president takes questions from members of the press.
+It is this part of the press conference in which we shall focus.
+\subsubsection{Hypothesis 1}
+We expect to find that both traditional forms of machine learning algorithms and deep learning algorithms are able to differentiate between those with cognitive impairment and healthy controls. 
+\subsubsection{Hypothesis 2}
+We expect to replicate the results of Berisha and Liss.
+\subsubsection{Hypothesis 3}
+We expect there to 
+\section{Methodology}
+\section{Results}
+\section{Discussion and Future Work}
+\section{Conclusion}
 \bibliographystyle{unsrt}
 \bibliography{ExperimentReport}
-
 \end{document}

### 2018-08-06 LiteratureReview/LiteratureReview.tex
work completed 06/08/18am

@@
 In addition, these symptoms are often accompanied by emotional difficulties such as depression and irritability and behavioural difficulties.
-Whilst there are several variants of Dementia, AD remains the most common type of Dementia and will be the focus of this review.
+Whilst there are several variants of Dementia, AD remains the most common type of Dementia and will be the primary focus of this review.
 \newline
@@
 If this is the case, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
+One of the most common ways in which clinicians make an early diagnosis of cognitive impairment is through the use of the Mini Mental State Examination \cite{Folstein1975}.
+
  Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on the decline in the use of language in those with AD.
@@
 \newline
+
+\subsection{Aims and Methodology}
 \par 
-The purpose of this review is to seek to understand what techniques have been used when considering the application of Machine Learning and Natural Language processing to aid the diagnosis of Mild Cognitive Impairment (MCI) and AD.
+The purpose of this review is to seek to understand what techniques for assessing language have been used within the field of cognitive and clinical psychology.
+We then go on to look at what techniques have been developed in the field of machine learning (ML), deep learning (DL) and natural language processing (NLP) that might enable the automated analysis of language easier.
+Finally, we look at some studies which have already looked at the intersection of these two domains.
+
+when considering the application of Machine Learning and Natural Language processing to aid the diagnosis of Mild Cognitive Impairment (MCI) and AD.
 A search of the literature was conducted using ProQuest (PsychArticles), SCOPUS, Web of Science.
@@
 All papers were then reviewed for relevance by reading the abstract and full text where appropriate and a shortlist was compiled.
-An additonal search through references of shortlisted papers was also conducted and any papers who upon further review appeared relevant were added to the shortlist.
+An additional search through references of shortlisted papers was also conducted and any papers who upon further review appeared relevant were added to the shortlist.
 Papers were included where researchers used machine learning to classify participants as MCI, AD or Healthy using language.
@@
 \end{table}
+
 \section {Types of Language Assessment}
@@
 \par
+
+\section{Measures of Lexical Richness}
+\subsection{Type token ratio(TTR)}
+Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
+\begin{equation} \label{x1}
+TTR = numberOfUniqueWords / totalNumberOfWords.
+\end{equation}
+\subsection{Brunet's Index(W)} % Don't forget to cite the original article.
+Brunet's Index (W) differentiates itself for TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
+\begin{equation} \label{x2}
+W = N^{V(-0.165)}
+\end{equation}
+where N is the total length of the utterance being measured and V is equal to the total vocabulary being used by the subject.
+Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers.
+\newline
+\par 
+\subsection{Honore's Statistic (R)} % Don't forget to cite the original article.
+Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation: 
+\begin{equation} \label{x3}
+%% Check to see if this is accurate.
+R = (100 \log N) / (1 - V1/V)
+\end{equation}
+where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
+
+\section{State of literature into Natural Language processing techniques}
+
 \section{Conclusions and Future Work}
@@
 \par
+\subsection{Future Work}
 Future research should also be directed towards developing non-intrusive ways of detecting subtle changes in language based in the home, such that any deterioration that could indicate the presence of MCI or AD and be flagged up early in the progression of any potential cogntive impairment for further review via referral.
@@
 \newline
-\par
-A search of the literature was conducted mainly from a psychological perspective as this would inform the theoretical side of things.
-Next, a search of the literature from a Computer Science perspective was also made with a view to exploring the how domain of machine learning and more specifically natural language processing can be applied to this problem.
-Finally a search was carried out to see what work has already been done in this area.
-
-All stages of this literature review were completed by myself.
-
-\section{Current ways of diagnosing Dementia using Language}
-\subsection{Free Cued Selective Recall Test}
-The Free Cued Selective Recall Test (FCSRT) was created by Grober and his team \cite{Grober1988}.
-This is a test that primarily tests memory but also tests a subjects ability to name objects.
-Four items are presented at a time on a card and each participant is asked to name and recall with and without cues.
-
-\subsection{Mini-Mental State Examination}
-
-Battery of tests to differentiate between Controls, those with MCI, those with AD. 
-\begin{itemize}
-	\item FCSRDT - As most effective 
-	\item Mini Mental State Examination (MMSE)
-	\item Addenbrooke's Cognitive Examination (ACE-III)
-	\item Dementia Rating Scale (DRS)
-	\item Western Aphasia Battery (WAB)
-	\item Boston Diagnostic Aphasia Examination (BDAE)
-	\item Boston Naming Test (BNT)
-\end{itemize}
-
-\section{Existing studies that have used Natural Language Processing to help in diagnosing psychological difficulties, particularly language difficulties}
-
-\subsection{Nun Study - Snowdon (1996)}
-\subsection{Comparing Presidents Speeches - Berisha et al (2015)}
-
-\section{State of literature into Natural Language processing techniques}
-
-\section{stuff to add}
-\section{Conclusion}
-\par
-Identification of language impairment is important in Dementia because it aids diagnosis of specific types of dementia, which in turn can alter the prognosis and change the management of the degenerative disorder.
-As these differences in language are quite subtle, the varying subtypes of dementia are frequently misdiagnosed.
-\newline
-\par 
-Given the burden on the diagnosis of dementia on clinicians, it appears to be useful to find some non-invasive protocols for the early diagnosis of dementia.
-It has already been shown that analysis of speech and language has shown markers that pre-date the official diagnosis of dementia (Snowdon et al, 1996;)\cite{Berisha2015}.
 

### 2018-08-09 ExperimentProposals/Experiments.tex
work done 09-08-2018

@@
 \documentclass{article}
-
 \title{Comparing the presidential news conferences of Ronald Reagan, George H.W.
@@
 It is this part of the press conference in which we shall focus.
+\newline
+The recent resurgence in the use of neural networks and deep learning is because of it's significantly improved performance on many problems and it's ability to scale from small to large datasets.
+Another key benefit of deep learning is it's ability to automate the process of feature engineering.
+Feature engineering, with this particular problem is an interesting debate as there are currently numerous ways in which to try to generate features.
+\newline
+\par 
 \subsubsection{Hypothesis 1}
@@
 We expect there to 
+\subsection{EXTRA}
+The objective of interest is whether or not cognitive impairment can be detected and is therefore binary in nature.
+We at this point have chosen not to distinguish between severity of the cognitive impairment.
 \section{Methodology}

### 2018-08-09 LiteratureReview/LiteratureReview.tex
work done 09-08-2018

@@
 \title{Literature Review}
-\date{2018-04-26}
+\date{2018-08-09}
 \author{Jomar Alcantara}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 If this is the case, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
-One of the most common ways in which clinicians make an early diagnosis of cognitive impairment is through the use of the Mini Mental State Examination \cite{Folstein1975}.
-
+One of the most common ways in which clinicians make an early diagnosis of cognitive impairment is through the use of the Mini Mental State Examination (MMSE) \cite{Folstein1975}.
+The MMSE is a brief questionnaire consisting of eleven questions which tests cognitive aspects of mental function and requires only 5-10 minutes to administer \cite{Folstein1975}.
+The MMSE is chosen due to it's effectiveness at assessing a person's cognitive mental state at a specific point in time, as well as being as sensitive to changes as a more detailed and complex assessment such as the Wechler Adult Intelligence Scale \cite{Folstein1975}.
+\newline
+\par 
 Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on the decline in the use of language in those with AD.
+\newline
+\par
+\subsection {On the decline of those with dementia} Whilst there are a wide number of factors that are involved in language generation, and therefore there will be an expected amount of variability between subjects.
+There is also an observation that as people get older, our language skills do decline.
+The consensus that this decline is typically exacerbated by the presence of dementia.
+Given this statement, it seems logical to conclude that one of the key variables that distinguishes language decline in healthy individuals vs those with dementia is the rate of change in which the decline occurs.
 \newline
@@
 The purpose of this review is to seek to understand what techniques for assessing language have been used within the field of cognitive and clinical psychology.
-We then go on to look at what techniques have been developed in the field of machine learning (ML), deep learning (DL) and natural language processing (NLP) that might enable the automated analysis of language easier.
+We then go on to look at what techniques have been developed in the field of machine learning (ML), deep learning (DL) and natural language processing (NLP) that might enable the automated analysis of language easier as well as an obstacles and/or limitations of current technology.
 Finally, we look at some studies which have already looked at the intersection of these two domains.
-
-when considering the application of Machine Learning and Natural Language processing to aid the diagnosis of Mild Cognitive Impairment (MCI) and AD.
+\newline
+\par 
 A search of the literature was conducted using ProQuest (PsychArticles), SCOPUS, Web of Science.
@@
 		Scopus & 29 Results & Machine Learning AND Dementia AND Langauge \\ \hline
+		Scopus & 1292 Results & "Language Deficits" AND Dementia \\ \hline
 	\end{tabular}
@@
 \subsection{Measures of Syntactic and Semantic Complexity}
-Another approach to linguistic analysis in this field is the idea of measuing syntactic and semantic complexity.
+Another approach to linguistic analysis in this field is the idea of measuring syntactic and semantic complexity.
 According to Emery (2000) \cite{Emery2000} in which she states that Semantic and Syntactic skills deteriorate first in people with MCI and AD.
@@
 
+\subsection{Semantics vs Pragmatics}
+When navigating the English language it is necessary to distinguish between what a sentence says in both semantic and pragmatic terms.
+Semantic meaning refers to the meaning of the words in a sentence local only to the given sentence.
+Another way to put this is, semantics considers the meaning of words without taking into account the context in which these words are spoken.
+Pragmatic meaning refers looks at the same sentence in terms of words and grammar but takes into account the situation or context in which these words are spoken.
+\newline
+\par
+%% IS THERE ANY RESEARCH ON THIS?
+
+%Current state-of-the-art diagnostic measures of AD are
+%invasive (CSF analysis), expensive (neuroimaging), and
+%time-consuming (neuropsychological assessment). Furthermore,
+%these measures are limited to specialty clinics and
+%thus have limited accessibility as high-throughput, or frontline,
+%screening and diagnostic tools for AD. More importantly,
+%nonspecialists are often inaccurate at identifying
+%early AD and MCI [5]. Thus, there is an increasing need
+%for additional noninvasive and/or cost-effective tools, allowing
+%frontline identification of subjects in the preclinical or
+%early clinical stages of AD who could be suitable for monitoring
+%in specialty clinics and for early treatment. Implementation
+%of effective screening instruments will allow diagnosis
+%earlier in the course of dementia, even at the point when
+%memory function is still essentially within the normal range.
+%This strategy would enable an earlier, and potentially more
+%effective, prevention and treatment of AD with a special
+%focus to preserve cognitive functions
+
 \section{Conclusions and Future Work}
@@
 \par
+The recent resurgence in the use of neural networks and deep learning is because of it's significantly improved performance on many problems and it's ability to scale from small to large datasets.
+Another key benefit of deep learning is it's ability to automate the process of feature engineering.
+Feature engineering, with this particular problem is an interesting debate as there are currently numerous ways in which to try to generate features.
+\newline
+\par 
 This area of research is extremely promising in its early results and the impact of successful research would be life changing for both individuals and the health of the worlds aging population in general.

### 2018-08-09 workingNotebook.tex
added working notebook
work done 09-08-2018

@@
 \section {Bits of writing to go in QR, or Thesis}
-\subsection {Potential Rationale for using deep learning over traditional machine learning methods.}
-The recent resurgence in the use of neural networks and deep learning is because of it's significantly improved performance on many problems and it's ability to scale from small to large datasets.
-Another key benefit of deep learning is it's ability to automate the process of feature engineering.
-Feature engineering, with this particular problem is an interesting debate as there are currently numerous ways in which to try to generate features.
 
-\subsection {On the decline of those with dementia} Whilst there are a wide number of factors that are involved in language generation, and therefore there will be an expected amount of variability between subjects.
-There is also an observation that as people get older, our language skills do decline.
-The consensus that this decline is typically exacerbated by the presence of dementia.
 \end {document}

### 2018-08-19 LiteratureReview/LiteratureReview.tex
work done over the weekend

@@
 \par
+% IN THIS PARAGRAPH - CITE UK REPORT
+Alzheimer's disease and other forms of dementia is affecting a significant proportion of geriatric population in the world today.
+According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
+This represents 1 in 14 of the population aged 65 or over and 1 in 79 in the general population.
+They estimate an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
+Globally, this picture is a lot bleaker.
+Another report by Alzheimer's Disease International suggests that in 2015 there were 46 million people with a diagnosis of AD and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
+\newline
+\par
 Alzheimer's Disease (AD) is a neurodegenerative disease in which the brain develops neurofibrillary tangles and neuritic plaques along with the loss of cortical neurons and synapses.
-The hallmark clinical symptoms are cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits\cite{AmericanPsychiatricAssociation2013}.
+The hallmark clinical symptoms of AD are cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits\cite{AmericanPsychiatricAssociation2013}.
 In addition, these symptoms are often accompanied by emotional difficulties such as depression and irritability and behavioural difficulties.
@@
 \par
-According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
-This represents 1 in 14 of the population aged 65 or over and 1 in 79 in the general population.
-They estimate an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
-Globally, this picture is a lot bleaker.
-A recent report suggests that in 2015 there were 46 million people with a diagnosis of AD and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
-The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
-\newline
-\par
+
 Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent.
@@
 \section{State of literature into Natural Language processing techniques}
+As we can see, diagnosing dementia through analysis has a large background in terms of psychological research.
+However, a lot of more current research has called for the use of machine learning as a way of assisting in the process of diagnosis \cite{Boschi2017}.
 

### 2018-08-19 Python/PythonCrashCourse/cubes2.py

@@ -1,0 +1,8 @@
+import matplotlib.pyplot as plt
+
+x_values = list(range(1, 1001))
+y_values = [x**3 for x in x_values]
+
+plt.scatter(x_values, y_values, c=y_values, cmap=plt.cm.Reds, edgecolor='none', s=1)
+
+plt.show()

### 2018-08-19 Python/PythonCrashCourse/mpl_squares.py

@@ -1,0 +1,14 @@
+import matplotlib.pyplot as plt
+
+input_values = [1, 2, 3, 4, 5]
+squares = [1, 4, 9, 16, 25]
+plt.plot(input_values, squares, linewidth=5)
+
+# Set chart title and label axes.
+plt.title("Square Numbers", fontsize=24)
+plt.xlabel("Value", fontsize=14)
+plt.ylabel("Square of Value", fontsize=14)
+
+# Set size of tick labels.
+plt.tick_params(axis="Both", labelsize=14)
+plt.show()

### 2018-08-19 Python/PythonCrashCourse/numbers.py

@@ -1,8 +1,0 @@
-for value in range(1, 5):
-    print(value)
-
-for value in range(1, 6):
-    print(value)
-
-numbers = list(range(1, 11))
-print(numbers)

### 2018-08-19 Python/PythonCrashCourse/random_walk.py

@@ -1,0 +1,36 @@
+from random import choice
+
+class RandomWalk():
+    """ A class to generate random walks """
+
+    def __init__(self, num_points=5000):
+        """Initialize attributes of a walk."""
+        self.num_points = num_points
+
+        # All walks start at (0, 0)
+        self.x_values=[0]
+        self.y_values=[0]
+
+    def fill_walk(self):
+        """Calculate all the points in the walk."""
+
+        # Keep taking steps until the walk reaches the desired lengthself.
+        while len(self.x_values) < self.num_points:
+            x_direction = choice([1, -1])
+            x_distance = choice([0, 1, 2, 3, 4])
+            x_step = x_direction * x_distance
+
+            y_direction = choice([1, -1])
+            y_distance = choice([0, 1, 2, 3, 4])
+            y_step = y_direction * y_distance
+
+            # Reject moves that go nowhere
+            if x_step == 0 and y_step == 0:
+                continue
+
+            # Calculate the next x and y values
+            next_x = self.x_values[-1] + x_step
+            next_y = self.y_values[-1] + y_step
+
+            self.x_values.append(next_x)
+            self.y_values.append(next_y)

### 2018-08-19 Python/PythonCrashCourse/rw_visual.py

@@ -1,0 +1,10 @@
+import matplotlib.pyplot as plt
+
+from random_walk import RandomWalk
+
+# Make a random walk, and plot the points.
+rw = RandomWalk()
+rw.fill_walk()
+
+plt.scatter(rw.x_values, rw.y_values, s=10)
+plt.show()

### 2018-08-19 Python/PythonCrashCourse/scatter_squares.py

@@ -1,0 +1,20 @@
+import matplotlib.pyplot as plt
+
+x_values = list(range(1, 1001))
+y_values = [x**2 for x in x_values]
+
+plt.scatter(x_values, y_values, c=y_values, cmap=plt.cm.Blues,
+    edgecolor='none', s=10)
+
+# Set chart title and label axes.
+plt.title("Square Numbers", fontsize=24)
+plt.xlabel("Value", fontsize=14)
+plt.ylabel("Square of Value", fontsize=14)
+
+# Set size of tick labels.
+plt.tick_params(axis='both', which='major', labelsize=14)
+
+# Set the range for each axis.
+plt.axis([0, 1100, 0, 1100000])
+
+plt.savefig('squares_plot.png', bbox_inches='tight')

### 2018-08-21 Python/nltk/chapter3.py
work done 21-08-2018

@@ -1,0 +1,80 @@
+import keras
+from keras.datasets import imdb
+
+(train_data, train_labels), (test_data, test_labels) = imdb.load_data(num_words=10000)
+
+import numpy as np 
+
+def vectorize_sequences(sequences, dimension=11000):
+    results = np.zeros((len(sequences), dimension))
+    for i, sequence in enumerate(sequences):
+        results[i, sequence] = 1. 
+    return results
+
+x_train = vectorize_sequences(train_data)
+x_test = vectorize_sequences(test_data)
+
+y_train = np.asarray(train_labels).astype('float32')
+y_test = np.asarray(test_labels).astype('float32')
+
+from keras import models
+from keras import layers
+
+model = models.Sequential()
+model.add(layers.Dense(32, activation = 'relu', input_shape=(11000, )))
+model.add(layers.Dense(16, activation = 'relu'))
+model.add(layers.Dense(16, activation = 'relu'))
+model.add(layers.Dense(1, activation = 'sigmoid'))
+model.compile(optimizer='rmsprop',
+                loss='mse',
+                metrics=['accuracy'])
+
+x_val = x_train[:11000]
+partial_x_train = x_train[11000:]
+y_val = y_train[:11000]
+partial_y_train = y_train[11000:]
+
+history = model.fit(partial_x_train,
+                    partial_y_train,
+                    epochs=20,
+                    batch_size=512,
+                    validation_data=(x_val, y_val))
+
+import matplotlib.pyplot as plt
+history_dict = history.history
+acc = history.history['acc']
+val_acc = history.history['val_acc']
+loss_values = history_dict['loss']
+val_loss_values = history_dict['val_loss']
+epochs = range(1, len(acc) + 1)
+plt.plot(epochs, loss_values, 'bo', label='Training loss')
+plt.plot(epochs, val_loss_values, 'b', label='Validation loss')
+plt.title('Training and validation loss')
+plt.xlabel('Epochs')
+plt.ylabel('Loss')
+plt.legend()
+plt.show()
+
+plt.clf()
+acc_values = history_dict['acc']
+val_acc_values = history_dict['val_acc']
+plt.plot(epochs, acc, 'bo', label='Training acc')
+plt.plot(epochs, val_acc, 'b', label='Validation acc')
+plt.title('Training and validation accuracy')
+plt.xlabel('Epochs')
+plt.ylabel('Loss')
+plt.legend()
+plt.show()
+
[…]

### 2018-08-21 Python/nltk/chapter5.py

@@ -1,0 +1,31 @@
+from keras import layers
+from keras import models
+model = models.Sequential()
+model.add(layers.Conv2D(32, (3, 3), activation='relu', input_shape=(28, 28, 1)))
+model.add(layers.MaxPooling2D((2, 2)))
+model.add(layers.Conv2D(64, (3, 3), activation='relu'))
+model.add(layers.MaxPooling2D((2, 2)))
+model.add(layers.Conv2D(64, (3, 3), activation='relu'))
+model.add(layers.Flatten())
+model.add(layers.Dense(64, activation='relu'))
+model.add(layers.Dense(10, activation='softmax'))
+
+print(model.summary())
+
+from keras.datasets import mnist
+from keras.utils import to_categorical
+(train_images, train_labels), (test_images, test_labels) = mnist.load_data()
+
+train_images = train_images.reshape((60000, 28, 28, 1))
+train_images = train_images.astype('float32') / 255
+test_images = test_images.reshape((10000, 28, 28, 1))
+test_images = test_images.astype('float32') / 255
+train_labels = to_categorical(train_labels)
+test_labels = to_categorical(test_labels)
+model.compile(optimizer='rmsprop',
+loss='categorical_crossentropy',
+metrics=['accuracy'])
+model.fit(train_images, train_labels, epochs=5, batch_size=64)
+
+test_loss, test_acc = model.evaluate(test_images, test_labels)
+print(test_acc)

### 2018-08-21 Python/nltk/nltk_ch1.py

@@ -1,7 +1,5 @@
 from nltk.book import *
 import nltk
-
-
 # concordance shows every occurance of the word passed to the function as well as some context.
 text1.concordance("monstrous")
 

### 2018-08-27 LiteratureReview/LiteratureReview.tex
work done over the weekend 27082018

@@
 \documentclass[a4paper]{article}
-
 %% Packages
 \input{settings/packages}
-
 %% Page Settings
 \input{settings/page}
-
 %% Own Commands
 \input{settings/macros}
-
 \title{Literature Review}
-\date{2018-08-09}
+\date{2018-08-22}
 \author{Jomar Alcantara}
@@
 % IN THIS PARAGRAPH - CITE UK REPORT
-Alzheimer's disease and other forms of dementia is affecting a significant proportion of geriatric population in the world today.
+Alzheimer's disease and other forms of dementia are affecting a significant proportion of geriatric population in the world today.
 According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
-This represents 1 in 14 of the population aged 65 or over and 1 in 79 in the general population.
-They estimate an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
+This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population.
+From a financial perspective, they estimate an annual spend of £4.3 billion of which approximately £85 million is spent solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
 Globally, this picture is a lot bleaker.
 Another report by Alzheimer's Disease International suggests that in 2015 there were 46 million people with a diagnosis of AD and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
-The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
+The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.\newline
+\par
+Alzheimer's Disease (AD) is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
+However, a definitive diagnosis of Dementia can only produced at post-mortem.
+From a clinical psychology perspective, those who have dementia have cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits\cite{AmericanPsychiatricAssociation2013}.
+In addition, these symptoms are often accompanied by emotional difficulties such as depression and irritability and behavioural difficulties.
 \newline
 \par
-Alzheimer's Disease (AD) is a neurodegenerative disease in which the brain develops neurofibrillary tangles and neuritic plaques along with the loss of cortical neurons and synapses.
-The hallmark clinical symptoms of AD are cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits\cite{AmericanPsychiatricAssociation2013}.
-In addition, these symptoms are often accompanied by emotional difficulties such as depression and irritability and behavioural difficulties.
-Whilst there are several variants of Dementia, AD remains the most common type of Dementia and will be the primary focus of this review.
+Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent.
+Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time and alter cognitive function until a threshold is reached and clinical symptoms become more obvious.
+\cite{Nestor2006}.
+The case of Iris Murdoch illustrates this theory well.
+Le et al \cite{Le2012}found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel.
+If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.\newline
+\par
+One of the most common ways in which clinicians traditionally make an early diagnosis of cognitive impairment is through the use of the Mini Mental State Examination (MMSE) \cite{Folstein1975}.
+The MMSE is a brief questionnaire consisting of eleven questions which tests cognitive aspects of mental function and requires only 5-10 minutes to administer \cite{Folstein1975}.
+The MMSE is chosen due to it's effectiveness at assessing a person's cognitive mental state at a specific point in time, as well as being as sensitive to changes as a more detailed and complex assessment such as the Wechler Adult Intelligence Scale \cite{Folstein1975}.
+Whilst the MMSE is useful as a brief screening tool it has it's limitations.
+The MMSE was not specifically created to screen for dementias and therefore does not interrogate key aspects of cognitive impairment known to be affected in dementia.
+It also has limited value in assessing under educated subjects and a meta-analysis on the effectiveness of the MMSE as a diagnostic tool for dementia to showed that it's accuracy was low (sensitivity between 78.4\% and 85.1\% and specificity between 81.3\% and 87.8\%) \newline
+% Need to cite Should the Mini-Mental State Examination be retired above.
+\par 
+As the MMSE is shown to have low accuracy specifically in the diagnosis of dementia, it becomes necessary for professionals to employ the use of other tools or measures such as the Free Cued Selective Reminding Test (FCSRT) or the Montreal Cognitive Assessment.
+These tests have the benefits of being much more accurate at diagnosing cognitive impairment and discriminating between dementia and other types of cognitive impairment at the cost time and training of psychological professionals such as clinical psychologists in administering these tests.
+Given the burden on these professionals is likely to increase due to a growing elderly population or in the case of developing countries where the clinician / patient ratio is already unsustainable, it seems useful to find less burdensome ways do aid professionals in diagnosing dementia.
 \newline
@@
 
-Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent.
-Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time until a threshold is reached and clinical symptoms become apparent \cite{Nestor2006}.
-If this is the case, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
-One of the most common ways in which clinicians make an early diagnosis of cognitive impairment is through the use of the Mini Mental State Examination (MMSE) \cite{Folstein1975}.
-The MMSE is a brief questionnaire consisting of eleven questions which tests cognitive aspects of mental function and requires only 5-10 minutes to administer \cite{Folstein1975}.
-The MMSE is chosen due to it's effectiveness at assessing a person's cognitive mental state at a specific point in time, as well as being as sensitive to changes as a more detailed and complex assessment such as the Wechler Adult Intelligence Scale \cite{Folstein1975}.
-\newline
-\par 
-Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on the decline in the use of language in those with AD.
+Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on measure decline in the use of language in those with AD.
 \newline

### 2018-08-28 LiteratureReview/LiteratureReview.tex
updated literature review

@@
 \par   
-
+The two main ways in which diagnosis is performed is through assessment of memory and language.
+Tests of memory are classically among the most accurate ways of diagnosing dementia, however they suffer from the same reliance on clinicians to administer these tests in a clinical setting.
+Language however is a lot easier to collect and can be done in more naturalistic settings.
+As with memory, these tests can be done over time and would be able to chart a patients language degeneration over time.
 Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on measure decline in the use of language in those with AD.
@@
 \par
-\subsection {On the decline of those with dementia} Whilst there are a wide number of factors that are involved in language generation, and therefore there will be an expected amount of variability between subjects.
-There is also an observation that as people get older, our language skills do decline.
-The consensus that this decline is typically exacerbated by the presence of dementia.
+There are a wide number of factors that are involved in language degeneration in the elderly, and consequently there will be an expected amount of variability between subjects.
+The consensus is that this decline is typically accelerated by the presence of dementia \cite{Berisha2015}.
 Given this statement, it seems logical to conclude that one of the key variables that distinguishes language decline in healthy individuals vs those with dementia is the rate of change in which the decline occurs.
-\newline
-\par 
-One of the most famous pieces of research was by Berisha and Liss (2015) \cite{Berisha2015} who examined speeches and public interviews of former US president Ronald Reagan.
-They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease.
-It turned out later to be the case that he had Alzheimer's Disease.
-Another classical study by Snowdon et al (1996) \cite{Snowdon1996} looked at whether linguistic ability in early life was associated with cognitive function and AD in later life.
-They found that idea density (defined as the number of expressed propositions divided by the number of words) was a key predictor in predicting whether nuns would go on to develop AD in later life.
-They found that those who would go on to develop AD all had low idea density in early life and they found no AD present in those with high idea density in early life.
-As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are wide and varied and differ as the disease develops.
 \newline
@@
 With the wide range of deficits someone with AD can suffer, it makes sense to try to categorise these deficits in some way.\newline
+\par
+One of the most famous pieces of research on the topic of language decline in dementia was by Berisha and Liss (2015) \cite{Berisha2015} who examined speeches and public interviews of former US president Ronald Reagan.
+They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease.
+It turned out later to be the case that he had Alzheimer's Disease.
+Another classical study by Snowdon et al (1996) \cite{Snowdon1996} looked at whether linguistic ability in early life was associated with cognitive function and AD in later life.
+They found that idea density (defined as the number of expressed propositions divided by the number of words) was a key predictor in predicting whether nuns would go on to develop AD in later life.
+They found that those who would go on to develop AD all had low idea density in early life and they found no AD present in those with high idea density in early life.
+As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are wide and varied and differ as the disease progresses.
+\newline
 \par
@@
 \par
-\section{How do we analyse language?}
+\section{How do we analyse language, issues and debates}
+\subsection{Single Word Language tasks vs Connected Language tasks}
 Part of the reason we need to pay attention to how we ask participants to generate data is understanding how we wish to analyse the data afterwards.
@@
 \par
-\subsection{Using frequencies of words and non-words}
-The basic premise behind using frequencies to measure changes in language is that certain types of words and non words are more prominent in those suffering with AD than healthy controls.
-For example, Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in diaglogue in which a speaker provides responsees do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls..\newline
+\subsection{Semantics vs Pragmatics}
+When navigating the English language it is necessary to distinguish between what a sentence says in both semantic and pragmatic terms.
+Semantic meaning refers to the meaning of the words in a sentence local only to the given sentence.
+Another way to put this is, semantics considers the meaning of words without taking into account the context in which these words are spoken.
+Pragmatic meaning refers looks at the same sentence in terms of words and grammar but takes into account the situation or context in which these words are spoken.
+\newline
 \par
+\section{Semantic Content}
+Another approach to linguistic analysis in this field is the idea of measuring semantic content and complexity.
+According to Emery (2000) \cite{Emery2000} in which she states that Semantic and Syntactic skills deteriorate first in people with MCI and AD.
+If this is true, then psychological measures of semantic and syntactic skills should be able to pick up signs of deterioration and act as markers for possible MCI and AD.
+An example of these measures is the concept of idea density.
+Formally, idea density is defined as the average number of propositions per sentence \cite{Kintsch1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}.
+Type to Token Ratio, Brunet's Index and Honore's Statistic are two more examples of validated measures which are used to measure the lexical diversity of a given piece of text.
+This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2005}.
+\newline
+\par
+\subsection{Picture-related content thematic elements}
+Several studies examined the amount of thematic elements expressed that were directly relevant to picture stimulus in picture description tasks.
+The studies used a variety of phrases to denote these thematic elements, these included 'pictorial themes', 'relevant observations and 'semantic units' but are notionally similar.
+The only difference was the number of thematic elements that 'scored' for the studies in question.
+Nicholas et al identified eight thematic elements of the Cookie Theft picture and used the number of elements as an outcome measure in different groups.
+He found that patients with AD expressed significantly fewer content elements than controls.
+\newline
+\par
+Hier, Hagenlocker and Shindler assessed content using a similar list of thematic elements.
+They divided their participants into early-stage and late-stage AD, as well as including a control group.
+The late-stage AD group produced significantly fewer relevant observations than the early stage group, and the AD group combined produced fewer relevant observations than controls.
+This study was replicated by Vuorinen et al (2001).
+\newline
+\par 
+Smith, Chenery and Murdoch (1989) applied Hier's methodology for constructing pictorial 'themes' with the Picnic Scene from the Western Aphasia Battery (WAB) with a control and patients with moderate to moderately severe AD.
+The authors found no difference in the number of semantic elements produced but did not that the group with moderate to moderately severe AD took more time and more syllables to communicate these elements.
+\newline
+\par
+Sajjadi et al (2012) examined 10 pictorial themes in picture description the Comprehensive Aphasia Test and found that the group with mild AD produced similar themes than controls.
+Bschor et al.
+(2001) examined Cookie Theft picture descriptions at four stages of AD.
+They identified 12 elements of the cookie theft picture and found that whilst each AD group differed significantly from the others and from controls, the measures did not distinguish between MCI and normal controls.
+\newline
+\par
+Finally, a number of studies used composite measures which contained thematic elements and other unspecified information units resulting in a list of 23 possible information units of the Cookie Theft picture.
+The authors felt that this provided a wider, more liberal range of relevant content and thus subtler differences could be noted.
+Studies using these features found some differences between AD and controls, and some could differentiate between different stages of AD.
+\subsection{General Information Units}
+Some studies used a more general concept of content, defining "information units" as "the smallest non redundant meaningful fact or inference," whether or not the information conveyed was specific to the context in which the conversation happened.
+Giles et al, for example, studied adults with minimal, mild or moderate AD vs controls and found that adults with AD produced fewer overall information units than controls.
+\subsection{Conciseness of information}
+Conciseness has been defined as the number of words a speaker uses to express ideas.
+The theory is that people with AD will need more words to convey ideas because of word-finding related behaviours such as circumlocutions and repetitions.
+Conciseness is calculated by dividing the number of ideas expressed by the total number of words in a measure commonly referred to as idea density but also known as lexical index, information content and information unit conciseness index.
+Snowdon et al examined written discourse from the Nun study and found that low idea density in early life was associated with reduced cognitive performance in later life.
+Riley et al extended these findings by concluding that early-life idea density was associated with lower brain weight, higher degree of cerebral atrophy and increased neurofibrillary pathology in later life.
+\newline
+\par 
+Ahmed, de Jager et al examined idea density with patients who had confirmed AD post mortem.
+They found that those with AD produced fewer total semantic units than controls but there was no significant difference between the groups with regards to idea density.
+The study of "empty speech" by Nicholas et al examined conciseness with measures thought to contribute to the "nonspecifity" of discourse in AD, such as empty phrases (defined as common idioms contributing no relevant information), deictic terms (e.g.
+"this", "that" without referents), indefinite terms (e.g.
+"thing" or "stuff"), pronouns without proper noun antecedants, and repetitions.
+In their study they found that AD patients produced more of these behaviours than did controls.
+\newline
+\par 
+\subsection{Efficiency}
+Efficiency is the rate at which meaningful information is conveyed over time, calculated by dividing the total number of information units by the duration in seconds of the speech sample.
+Smith et al, 1989 found that 18 adults with moderately severe AD produced fewer content units per minute on average than controls, and attributed this different to increased circumlocutions and repetitions.
+Murray 2010, used an analogous measure which he called "performance deviations per minute", in which fillers, irrelevant words, revisions or false stars, vague or non-specific vocabulary and inaccurate output (e.g.
+paraphasias) were divided by the total number of minutes in the speech sample; this measure was lower for those with AD than those with depression, and also healthy controls.
+The authors suggested that discourse information measures may help disentangle the similarities in symptoms of early AD versus depression in older adults.
+Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in dialogue in which a speaker provides responsees do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls..\newline
+\subsection{Lexical richness and diversity}
+\subsubsection{Type token ratio(TTR)}
+Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
+\begin{equation} \label{x1}
+TTR = numberOfUniqueWords / totalNumberOfWords.
+\end{equation}
+\subsubsection{Brunet's Index(W)} % Don't forget to cite the original article.
+Brunet's Index (W) differentiates itself for TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
+\begin{equation} \label{x2}
+W = N^{V(-0.165)}
+\end{equation}
+where N is the total length of the utterance being measured and V is equal to the total vocabulary being used by the subject.
+Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers.
+\newline
+\par 
+\subsubsection{Honore's Statistic (R)} % Don't forget to cite the original article.
+Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation: 
+\begin{equation} \label{x3}
+%% Check to see if this is accurate.
+R = (100 \log N) / (1 - V1/V)
+\end{equation}
+where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
+\subsection{Quantity - Total number of words}
+Several studies report that adults with moderate AD produce fewer words than controls on picture description, however other studies found no differences in total words among groups of controls and patients with MCI or AD.
+Murray and Nicholas et al investigated normal controls, patients with AD and older adults with depression and found no group differences in total words.
+In contrast, Lira 2014 found that controls produced more total words than patients with AD but found no difference between mild and moderate groups.
+\newline
+\par
+\section{Syntax and Morphology (Language Form)}
+Syntax can be defined as the rules that govern how words can be combined to form sentences, whilst Morphology is the system that governs the structure of words and the construction of word forms.
+Multiple studies of language decline in dementia included at least one measure of syntax or syntactic complexity.
+Common constructs included words per clause, grammatical form (measures of an appropriate use of syntactic conjections, tenses, conditionals, subordinate clauses and passive constructions), measures of phrase length and proportions of words in sentences.
+Some researchers have explored the use of formulaic language in those with dementia, the theory being that well practiced phrases are less effortful and therefore place low load on the cognitive abilities of those with AD.
+The general hypothesis motivating these studies is that either working memory limitations or semantic memory limitations in AD affect one's ability to use complex constructions.
+\newline
+\par
+\subsection{N-grams}
 One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
@@
 This is useful as, given a large portion of text or speech, we can predict the probability of a word being close by to a given word.
-\newline
-\par
 A number of researchers have used n-grams as features.
@@
 \par
-\subsection{Measures of Syntactic and Semantic Complexity}
-Another approach to linguistic analysis in this field is the idea of measuring syntactic and semantic complexity.
-According to Emery (2000) \cite{Emery2000} in which she states that Semantic and Syntactic skills deteriorate first in people with MCI and AD.
-If this is true, then psychological measures of semantic and syntactic skills should be able to pick up signs of deterioration and act as markers for possible MCI and AD.
-An example of these measures is the concept of idea density.
-Formally, idea density is defined as the average number of propositions per sentence \cite{Kintsch1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}.
-Type to Token Ratio, Brunet's Index and Honore's Statistic are two more examples of validated measures which are used to measure the lexical diversity of a given piece of text.
-This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2005}.
+\subsection{Mean length of utterance (MLU)}
+Murray found that MLU was not a distinguishing factor among health adults, adults with depression and adults with AD.
+Ripich et al found a decrease in MLU in adults with severe AD over time, and this was supported by findings of Le et al in their studies of authors \cite{Le2012} \newline
+\par 
+\subsection{Proportion of verbs to nouns plus verbs}
+Kave and Levy used a verb index to capture syntactic complexity and found that adults with AD expressed the same amount of verbs to nouns plus verbs as adult controls. \newline
+\par 
+\subsection{Syntactic Complexity - Composite measures of MLU, syntactic errors and verbs}
+Ahmed et al, and Ahmed, Haigh et al found differences in syntactic complexity between adults with MCI and controls, and between MCI and moderate AD stages.
+The differences in syntactic complexity were not significant when individual measures were tested, but were apparent using a composite score consisting of MLU, words in sentences, syntactic errors, nouns with determiners, and verbs with inflections.
 \newline
 \par 
+\subsection{Formulaic Language}
 Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
@@
 \par
-It appears that both frequency based approaches and approaches that use measures of linguistic complexity have both been shown to be effective in various experimental settings.
-The fact that there are two vastly different perspectives to tackling this question, and even within these perspectives there are a range of various methodologies leaves the field with a sense of confusion as to the most effective way to solving the problem.
+\section{Pragmatic Language}
+The pragmatic language domain refers to the social rules for language for the purposes of communication including, using language to achieve goals, using information from the context to achieve these goals and using the interaction between people to initiate, maintain and terminate conversations.
+\subsection{Questions, turn-taking, unsure statements, egocentric comments}
+One study, Ripich et al, examined several pragmatic abilities among patients at different stages of AD.
+The severe AD group asked more questions over time than the other group.
+The authors argued that question-asking was a compensatory strategy, and as a result, adults in late-stage AD may have had insight into their communication problems.
+Potentially, there were a number of flaws of the methodology if their research.
+Firstly, a caregiver was asked to administer the picture description task in order to mirror a more typical communicative interaction.
+Whilst this improved the ecological validity of the study, the caregivers delivery would be varied and inconsistent.
+Secondly, due to the constrained nature of the picture description task, it is unlikely that it would be able to capture the pragmatic skills that are typioical of conversations in everyday life.
 \newline
 \par 
-
-\section{Measures of Lexical Richness}
-\subsection{Type token ratio(TTR)}
-Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
-\begin{equation} \label{x1}
-TTR = numberOfUniqueWords / totalNumberOfWords.
-\end{equation}
-\subsection{Brunet's Index(W)} % Don't forget to cite the original article.
-Brunet's Index (W) differentiates itself for TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
-\begin{equation} \label{x2}
-W = N^{V(-0.165)}
-\end{equation}
-where N is the total length of the utterance being measured and V is equal to the total vocabulary being used by the subject.
-Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers.
+\subsection{Coherence}
+Coherence refers to the appropriate maintenance of topic in discourse and is thematically related to the immediately preceding utterance (local coherence) and by how closely an utterance relates to the general topic at hand (global coherence).
+One study looked at coherence, Chapman et al used picture descriptions of Norman Rockwell prints within a frame analysis, with frames being defined as internalized knowledge structures.
+The authors identified aspects of content, including typical frames of interpretation, atypical, incorrect or no frames, propositions supporting frames and propositions disrupting frames as measures of coherence.
+They examined these variables with early stage AD, old-elderly and normal controls.
+Old-elderly and normal controls produced significantly more typical frames and more frame supporting information that the AD group.
+The authors attributed AD patients' difficulties to memory deficits, attentional deficits, visual perceptual problems, disruption of internalized frame representation, or failure to access frame knowledge.
 \newline
 \par 
-\subsection{Honore's Statistic (R)} % Don't forget to cite the original article.
-Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation: 
-\begin{equation} \label{x3}
-%% Check to see if this is accurate.
-R = (100 \log N) / (1 - V1/V)
-\end{equation}
-where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
+\subsection{Perseveration}
+One study examined verbal preservation in the description of Norman Rockwell prints, dividing the total number of words within perseverations by total number of words in the speech sample.
+The authors also calculated rate of perseveration on two other language tasks: confrontation naming and generative naming.
+Across all tasks, the AD group produced significantly more perseverations than controls; however, on the picture description task alone, there were no significant differences between the two groups.
+The authors theorised that this was because picture description was an easier taks because there was the visual stimulus, similar to the arguement made by Bschor et al.
+\newline
+\par 
+\subsection{Speech intonation / prosody}
+There are a number of studies, that have examined "melodic line", which is a subjective measure of speech prosody defined as "the appropriate use of intonational contour, including alterations in pitch, volume and duration".
+For instance, Forbes-McKay compared melocdic line using a "simple" picuture such as the Cookie Theft picture vs a complex picture.
+The number of pictorial themes differentiated the simple tasks from the complex.
+Results showed no group differences on simple picture tasks (Cookie Theft and Tripping Woman) but there were differences in melodic line for the complex picture tasks.
+Fraser et al examined several acoustic features of speech in patients with AD ussing machine learning methods that captureed both rate and phonation patterns, and found acoustic abnormalities to be a significant factor.
+Konig et al, also used an automated feature analysis examining vocal and temporal aspects of speech among controls, and patients with MCI or AD, and reported a classification of 81%.
+\subsection{Discourse Fluency}
+Verbal fluency is a term used in neuropsychological contexts generally referring to timed, word-generation tasks, while in speech-language pathology contexts, "fluency disorders" are defined as interruptions int he flow of speaking characterized by atypical rate, rhythm and repetitions in sounds, syllables, words and phrases.
+"Fluency", in the literature of discourse of adults with AD, typically refers to the smoothness or flow of spoken language.
+Abnormalities of fluency in this population are typically characterised by filled and unfilled pauses, word repetitions, circumlocutions, and revisions.
+In contrast with fluency disorders (i.e.
+stuttering), abnormailities in the fluency of adults with MCI or AD are rarely manifested at the phonological level.
+\newline
+\par
+The study of "empty speech" by Nicholas et al was one of the first to examine aspects of fluency in the connected speech of persons with AD.
+They found that adults with AD had significantly more repetitions than controls.
+Similarly, Tomoeda et al found more aborted phrases, revisions and ideational repetitions in the AD group than in controls.
+Several other studies support the idea that in AD population there are a greater number of repetitions and revisions than in healthy controls.
+However, there are some studies which contradict these findings (Ahmed et al).
+\newline
+\par 
+\subsection{Speech Monitoring}
+Speech monitoring is related not only to word retrieval deficits and to pragmatic language skill but also to "anosognosia" define as the awareness of one's own deficits.
+McNamara et al investigated word error monitoring skills by comparing uncorrected and repaid errors in adults with AD vs health controls.
+The AD group was equally impaired in error monitoring as the controls.
+Severity of naming deficits correlated negatively to the amount of uncorrected errors.
+The authors suggested that this impairment was related to the executive function dificulties in the clinical groups.
+The authors did not report correlations between error monitoring and executive function test scores, however, which could have strengthened that hypothesis.
+\newline
+\par 
+Another measure of speech monitoring used in the picture description literature is "response to word-finding delays," defined as "whether patients appear unaware of their problem, produce an approximation of the target word or actively search and produce the target word".
+Response to word-finding delays differed significantly between minimal  AD and normal controls (forbes-McKay and Venneri).
+The measure was based on Goodglass and Kaplan's scale for rating discourse on the Boston Diagnostic Aphasia Examination (BDAE) and is composed of clinical judgement of behaviour that is rated on a likert-type scale ranging fdrom 1 to 7 (7 being no abnormality).
 
@@
 
-\subsection{Semantics vs Pragmatics}
-When navigating the English language it is necessary to distinguish between what a sentence says in both semantic and pragmatic terms.
-Semantic meaning refers to the meaning of the words in a sentence local only to the given sentence.
-Another way to put this is, semantics considers the meaning of words without taking into account the context in which these words are spoken.
-Pragmatic meaning refers looks at the same sentence in terms of words and grammar but takes into account the situation or context in which these words are spoken.
-\newline
-\par
 %% IS THERE ANY RESEARCH ON THIS?
@@
 \section{Conclusions and Future Work}
+It appears that both frequency based approaches and approaches that use measures of linguistic complexity have both been shown to be effective in various experimental settings.
+The fact that there are two vastly different perspectives to tackling this question, and even within these perspectives there are a range of various methodologies leaves the field with a sense of confusion as to the most effective way to solving the problem.
+\newline
 \par

### 2018-12-30 Source/LiteratureReview/LiteratureReview.tex
added source

--- 2018-08-28 LiteratureReview/LiteratureReview.tex
@@
+%% TO DO LIST
+% Fix the table on the literature search.
+% Talk about measures of lexical diversity (TTR, Brunets, Honore's)
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \input{settings/packages}
+\usepackage{graphicx}
+\usepackage{float}
 %% Page Settings
@@
 \title{Literature Review}
-\date{2018-08-22}
+\date{2018-09-05}
 \author{Jomar Alcantara}
@@
 \par
-% IN THIS PARAGRAPH - CITE UK REPORT
-Alzheimer's disease and other forms of dementia are affecting a significant proportion of geriatric population in the world today.
+Alzheimer's disease (AD) and other forms of dementia affect a significant proportion of the geriatric population in the world today and is currently the sixth leading cause of death in the US and the leading killer of women in the UK.
 According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
-This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population.
+This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population \cite{Sturrock2016}.
 From a financial perspective, they estimate an annual spend of £4.3 billion of which approximately £85 million is spent solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
 Globally, this picture is a lot bleaker.
-Another report by Alzheimer's Disease International suggests that in 2015 there were 46 million people with a diagnosis of AD and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
-The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.\newline
+Another report by Alzheimer's Disease International suggests that in 2015 there were 46 million people with a diagnosis of dementia globally and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
+\newline
 \par
-Alzheimer's Disease (AD) is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
-However, a definitive diagnosis of Dementia can only produced at post-mortem.
+% In this paragraph, look for the study which actually has these figures.
+Despite this growing problem, at present there are no drugs that improve the prognosis of those suffering with AD.
+All the drugs that are on the market are designed to manage symptoms.
+Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{Alzheimer2016}.
+Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place.
+\newline
+\par 
+Alzheimer's Disease is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
+Whilst a definitive diagnosis of Dementia can only produced at post-mortem, there are a number of clinical indicators from psychological perspective that can indicate that dementia is present.
 From a clinical psychology perspective, those who have dementia have cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits\cite{AmericanPsychiatricAssociation2013}.
@@
 Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent.
-Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time and alter cognitive function until a threshold is reached and clinical symptoms become more obvious.
-\cite{Nestor2006}.
+Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time and alter cognitive function until a threshold is reached and clinical symptoms become more obvious \cite{Nestor2006}.
 The case of Iris Murdoch illustrates this theory well.
 Le et al \cite{Le2012}found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel.
-If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.\newline
+If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
+However, one of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of dementia.\newline
 \par
@@
 The MMSE was not specifically created to screen for dementias and therefore does not interrogate key aspects of cognitive impairment known to be affected in dementia.
-It also has limited value in assessing under educated subjects and a meta-analysis on the effectiveness of the MMSE as a diagnostic tool for dementia to showed that it's accuracy was low (sensitivity between 78.4\% and 85.1\% and specificity between 81.3\% and 87.8\%) \newline
-% Need to cite Should the Mini-Mental State Examination be retired above.
-\par 
-As the MMSE is shown to have low accuracy specifically in the diagnosis of dementia, it becomes necessary for professionals to employ the use of other tools or measures such as the Free Cued Selective Reminding Test (FCSRT) or the Montreal Cognitive Assessment.
+It also has limited value in assessing under-educated subjects and a meta-analysis on the effectiveness of the MMSE as a diagnostic tool for dementia to showed that it's accuracy was low (sensitivity between 78.4\% and 85.1\% and specificity between 81.3\% and 87.8\%).
+As the MMSE is shown to have low accuracy specifically in the diagnosis of dementia, it becomes necessary for professionals to employ the use of other tools or measures such as the Free Cued Selective Reminding Test (FCSRT) or the Montreal Cognitive Assessment (MoCA).
 These tests have the benefits of being much more accurate at diagnosing cognitive impairment and discriminating between dementia and other types of cognitive impairment at the cost time and training of psychological professionals such as clinical psychologists in administering these tests.
-Given the burden on these professionals is likely to increase due to a growing elderly population or in the case of developing countries where the clinician / patient ratio is already unsustainable, it seems useful to find less burdensome ways do aid professionals in diagnosing dementia.
+Given the burden on these professionals is likely to increase due to a growing elderly population or in the case of developing countries where the clinician / patient ratio is already unsustainable, it seems useful to find less burdensome ways to aid professionals in diagnosing dementia.
 \newline
@@
 The two main ways in which diagnosis is performed is through assessment of memory and language.
-Tests of memory are classically among the most accurate ways of diagnosing dementia, however they suffer from the same reliance on clinicians to administer these tests in a clinical setting.
+Tests of memory are classically among the most accurate ways of diagnosing dementia, however these tests suffer from the same reliance on trained clinicians to administer these tests in a clinical setting.
 Language however is a lot easier to collect and can be done in more naturalistic settings.
@@
 Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on measure decline in the use of language in those with AD.
+There are a wide number of factors that are involved in language degeneration in the elderly, and consequently there will be an expected amount of variability between subjects.
+The consensus is that this language degeneration is typically accelerated by the presence of dementia \cite{Berisha2015} and that a potential indicator of dementia is the rate of change in which the decline occurs relative to a fixed point in time rather than a comparison across a cohort of individuals.
 \newline
 \par
-There are a wide number of factors that are involved in language degeneration in the elderly, and consequently there will be an expected amount of variability between subjects.
-The consensus is that this decline is typically accelerated by the presence of dementia \cite{Berisha2015}.
-Given this statement, it seems logical to conclude that one of the key variables that distinguishes language decline in healthy individuals vs those with dementia is the rate of change in which the decline occurs.
-\newline
-\par
-According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with Mild dementia suffer from noticeable word finding difficulty.
+According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with mild dementia suffer from noticeable word finding difficulty.
 They may substitute general terms for more specific terms and may avoid the use of specific names of acquaintances.
@@
 They found that those who would go on to develop AD all had low idea density in early life and they found no AD present in those with high idea density in early life.
-As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are wide and varied and differ as the disease progresses.
+As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are extremely variable and can differ from patient to patient as the disease progresses.
 \newline
@@
 It was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}\newline
+\begin{figure}[H]
+\centering 
+\includegraphics[width=240px, height=150px]{BCTPicture.png}
+\caption{Cookie Theft Picture.\label{white}}
+\end{figure}
 \par 
@@
 \par
-Given the relative strengths of the Narrative description task vs Picture description task, there are few pieces of research research that have used Machine Learning to analyse features from Narrative picture tasks \cite{Fraser2014}.
-This could be down to the availability of data and the absence of any meaningful sets of transcripts of participants performing this task.
+Given the relative strengths of the Narrative description task vs Picture description task, there are few pieces of research  that have used Machine Learning to analyse features from Narrative picture tasks \cite{Fraser2014}.
+This could be due to the availability of data and the absence of any meaningful sets of transcripts of participants performing this task.
 However, this could be an interesting direction to take research in the future to see if features generated from this task could be used to predict MCI or AD.
@@
 \subsection{Interviews}
-Interviews can also be used to elicit language, the idea of employing questions to guide a conversation between speakers.
+Interviews can also be used to elicit language in a more natural way by asking questions to guide a conversation between speakers.
 There are three types of interviews: unstructured, structured and semi-structured.
@@
 Whilst this is the most ecologically valid form of conversation and therefore language generation, it's unstructured nature means that the protocol cannot be consistent and therefore reproduced.
-Semi-structured interviews are therefore preferred other forms of interview as a middle ground.
+Semi-structured interviews are therefore preferred over other forms of interview as a middle ground.
 The semi structured nature of these interviews means that there is some replicability but does not constrain the participant in answering questions.
@@
 Nevertheless, this is the most naturalistic setting for looking at language production and can be used to look at the syntactic and semantic parts of language generation \cite{Sajjadi2012}.
-There have been some attempts to use interviews to to assess language production in AD with promising results \cite{Asgari2017, Guinn2015}.\newline
+There have been some attempts to use interviews to assess language production in AD with promising results \cite{Asgari2017, Guinn2015 .\newline
 \par
@@
 \par
+Single word tasks such as the Boston Naming Test and other such standardized language tests generally target a participants verbal fluency where this is defined as the ability to form and express words in accordance with certain criteria.
+For example, a common test of verbal fluency would be to ask a participant to name as many words in a given category in a limited time frame.
+These categories can be objects such as animals or sports, or can be words beginning with a specific letter.
+There are a number of benefits to using single word tasks.
+From a research methodology perspective, using a standardized test allows researchers to target a very specific process in language generation and isolate factors that impact performance in language well.
+However, this approach does not take into account the symbiotic nature of language processes.
+\newline
+\par 
+Connected language tasks, such as the picture description task and interviews described above are a much more reflective of processes involved in natural language generation.
+They are able to be relatively constrained in the language available to be used, for example in the picture description task, or they can be unconstrained in the form of interviews.
+\newline
+\par
 \subsection{Semantics vs Pragmatics}
@@
 Pragmatic meaning refers looks at the same sentence in terms of words and grammar but takes into account the situation or context in which these words are spoken.
-\newline
+Emery in her literature review looks at all levels of language tasks except for pragmatics, however   \newline
 \par
@@
 If this is true, then psychological measures of semantic and syntactic skills should be able to pick up signs of deterioration and act as markers for possible MCI and AD.
-An example of these measures is the concept of idea density.
+An example of a semantic complexity measure is the concept of idea density.
 Formally, idea density is defined as the average number of propositions per sentence \cite{Kintsch1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}.
-Type to Token Ratio, Brunet's Index and Honore's Statistic are two more examples of validated measures which are used to measure the lexical diversity of a given piece of text.
+Some examples of semantic content measures are Type to Token Ratio, Brunet's Index and Honore's Statistic  which are used to measure the lexical diversity of a given piece of text and/or utterance.
 This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2005}.
@@
 \par
+\begin{figure}[H]
+\centering
+\includegraphics[width=240px, height=150px]{picnic-scene.jpg}
+\caption{Picnic Scene taken from the Western Aphasia Battery (WAB).\label{white}}
+\end{figure}
 Sajjadi et al (2012) examined 10 pictorial themes in picture description the Comprehensive Aphasia Test and found that the group with mild AD produced similar themes than controls.
@@
 Across all tasks, the AD group produced significantly more perseverations than controls; however, on the picture description task alone, there were no significant differences between the two groups.
-The authors theorised that this was because picture description was an easier taks because there was the visual stimulus, similar to the arguement made by Bschor et al.
+The authors theorised that this was because picture description was an easier taks because there was the visual stimulus, similar to the argument made by Bschor et al.
 \newline
@@
 Fraser et al examined several acoustic features of speech in patients with AD ussing machine learning methods that captureed both rate and phonation patterns, and found acoustic abnormalities to be a significant factor.
-Konig et al, also used an automated feature analysis examining vocal and temporal aspects of speech among controls, and patients with MCI or AD, and reported a classification of 81%.
+Konig et al, also used an automated feature analysis examining vocal and temporal aspects of speech among controls, and patients with MCI or AD, and reported a classification of 81\%.
 \subsection{Discourse Fluency}
@@
 However, a lot of more current research has called for the use of machine learning as a way of assisting in the process of diagnosis \cite{Boschi2017}.
+Generally speaking, Natural Language Processing (NLP) systems mimic the semiotic perspective on language.
+That is to say that we can subdivide NLP processes into different components.
+This broadly fits with the literature review conducted by Emery.
+\begin{enumerate}
+	\item Morphological/Lexical: providing the basic language elements or vocabulary, such as words, their roots and inflections.
+	\item Syntactic: for grouping and sequencing elementds within samples of language (usually sentences).
+	\item Semantic: for knowing the meaning of an utterance, usually defined in terms of the 'truth value' of the logical propositions that are thought to be expressed by sentences.
+	\item Pragmatic: for understanding the context and purpose of an utterance.
+\end{enumerate}
+\subsection{Traditional methods of Machine Learning and Natural Language Processing}
 
-%% IS THERE ANY RESEARCH ON THIS?
+\subsection{Deep Learning}
 
-%Current state-of-the-art diagnostic measures of AD are
-%invasive (CSF analysis), expensive (neuroimaging), and
-%time-consuming (neuropsychological assessment). Furthermore,
-%these measures are limited to specialty clinics and
-%thus have limited accessibility as high-throughput, or frontline,
-%screening and diagnostic tools for AD. More importantly,
-%nonspecialists are often inaccurate at identifying
-%early AD and MCI [5]. Thus, there is an increasing need
-%for additional noninvasive and/or cost-effective tools, allowing
-%frontline identification of subjects in the preclinical or
-%early clinical stages of AD who could be suitable for monitoring
-%in specialty clinics and for early treatment. Implementation
-%of effective screening instruments will allow diagnosis
-%earlier in the course of dementia, even at the point when
-%memory function is still essentially within the normal range.
-%This strategy would enable an earlier, and potentially more
-%effective, prevention and treatment of AD with a special
-%focus to preserve cognitive functions
-
+\section{The problems with attempting early diagnosis of dementia and the ethics of actively screening for it}
+It is challenging to characterise Dementia as a disease as it's more syndrome.
+The clinical features that combine to meet the diagnostic criteria for Dementia or it's variants are continuous in nature and also impacted by other variables.
+For example, cognitive performance is affected by previous education and ability to live independently is impacted by a person's physical health.
+Also, the individual who may develop symptoms that indicate the onset of dementia needs the insight to recognise them, or family members will need to do so and also deduce that there is value to admitting there is a problem and seeking help before the deterioration is too great.
+\newline
+\par 
+To detect dementia early requires an understanding of how this presents early in it's progression, but as described above there are a number of potential ways in which individuals with dementia could be impacted but no consistent list of criteria has been produced.
+Even for AD, the most prevalent form of dementia, clinical diagnosis can only be confirmed after death through autopsy.
+In addition, population studies have shown that people over the age of 80 show similar changes in the brain in those that do not have dementia.
+Therefore, it's difficult to say with certainty that any symptoms can definitely cause dementia despite the fact that these symptoms are commonly present in those that have dementia.
+\newline
+\par 
 \section{Conclusions and Future Work}
-It appears that both frequency based approaches and approaches that use measures of linguistic complexity have both been shown to be effective in various experimental settings.
-The fact that there are two vastly different perspectives to tackling this question, and even within these perspectives there are a range of various methodologies leaves the field with a sense of confusion as to the most effective way to solving the problem.
+It appears that there are a number of approaches that have been shown to be effective in various experimental settings.
+The fact that there are a numerous array of different perspectives to tackling this question, and even within these perspectives there are a range of various methodologies leaves the field with a sense of confusion as to the most effective way to solving the problem.
 \newline
+\par
+Current state-of-the-art diagnostic measures of AD are invasive (CSF analysis), expensive (neuroimaging), and time-consuming (neuropsychological assessment).
+Furthermore, these measures are limited to specialty clinics and thus have limited accessibility as high-throughput, or frontline, screening and diagnostic tools for AD.
+More importantly, nonspecialists are often inaccurate at identifying early AD and MCI [5].
+Thus, there is an increasing need for additional noninvasive and/or cost-effective tools, allowing frontline identification of subjects in the preclinical or early clinical stages of AD who could be suitable for monitoring in specialty clinics and for early treatment.
+Implementation of effective screening instruments will allow diagnosis earlier in the course of dementia, even at the point when memory function is still essentially within the normal range.
+This strategy would enable an earlier, and potentially more effective, prevention and treatment of AD with a special focus to preserve cognitive functions \newline
 \par 
@@
 Another key benefit of deep learning is it's ability to automate the process of feature engineering.
-Feature engineering, with this particular problem is an interesting debate as there are currently numerous ways in which to try to generate features.
-\newline
+Feature engineering, with this particular problem is an interesting debate as there are currently numerous ways in which to try to generate features as I have described above.\newline
 \par 

### 2018-12-30 Source/QualifyingReport/QualifyingReport.tex
added source

--- 2018-12-30 Source/LiteratureReview/LiteratureReview.tex
@@
-%% TO DO LIST
-% Fix the table on the literature search.
-% Talk about measures of lexical diversity (TTR, Brunets, Honore's)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Document Class
-\documentclass[a4paper]{article}
+% HEADER BEGINS
+\documentclass{article}
 %% Packages
@@
 \input{settings/macros}
-\title{Literature Review}
-\date{2018-09-05}
+\title{Qualifying Report}
+\date{2018-12-18}
 \author{Jomar Alcantara}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin {document}
+% HEADER ENDS
+% DOCUMENT BEGINS
+\begin{document}
 \maketitle
@@
 \newpage
-\section{Introduction}
+\section{Overview of Problem and potential contribution.}
+Dementia has been identified as one of those fast growing difficulties facing the world.
+A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
+\newline
+\par
+In 2009, the Department of Health in it's National Dementia strategy made 'early diagnosis and support' one of it's key themes as part of this strategy.
+With that in mind, a lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cogntive Impairment (MCI).
+As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages and thus it is important that we utilise technology wherever possible to aid clinicians.
+At the present time diagnosis is typically conducted at memory clinics by trained clinicians.
+I theorise that we might be able to use diagnose patients in their own homes utilising smart-home technology.
+\newline
+\par
+Language analysis is an important part of assessing those suspected of suffering with AD or MCI and this language analysis in part informs an official clinical diagnosis.
+This analysis often takes the form of tasks such as Free Cued Selective Reminding Task and the picture description task which are all quite targeted in the language required of the person being assessed.
+Other studies have analysed more spontaneous samples of language and have identified patterns of speech which decline.
+\newline
+\par
+A step towards the ambient detection of AD and/or MCI in population requires that we look for similar features in language used naturally as part of every day conversation.
+Part of this work therefore is to see whether spontaneous discourse, such a semi-structured interview is a task that has the ability to put pressure both the cognitive and linguistic systems as it requires the ability to understand speech, formulating a coherent answer and then responding in an appropriate way.
+Berisha et al \cite{Berisha2015}, has shown through language analysis in this area that there are marked differences in this process between those who have a diagnosis of AD and healthy controls.
+\newline
+\par
+The potential impact of this research is immense.
+Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can slow the progress of the disease.
+Equally, early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
+\section{Literature Review}
+\subsection{Introduction}
 \par
@@
 \par
-% In this paragraph, look for the study which actually has these figures.
 Despite this growing problem, at present there are no drugs that improve the prognosis of those suffering with AD.
@@
 These tests have the benefits of being much more accurate at diagnosing cognitive impairment and discriminating between dementia and other types of cognitive impairment at the cost time and training of psychological professionals such as clinical psychologists in administering these tests.
-Given the burden on these professionals is likely to increase due to a growing elderly population or in the case of developing countries where the clinician / patient ratio is already unsustainable, it seems useful to find less burdensome ways to aid professionals in diagnosing dementia.
+Given the burden on these professionals is likely to increase due to a growing elderly population or in the case of developing countries where the clinician / patient ratio is already unsustainable, it seems useful to find less burdensome ways to aid professionals in identifying those at risk of developing dementia.
 \newline
 \par
+Many researchers have studied the early detection of AD.
+These studies usually follow two main approaches.
+The analysis of biomarkers and the examination of patients who have demonstrated decreasing cognitive abilities.
+The first approach yields reliable results in the detection of AD in its moderate and advanced states but does not perform well during the early stages of the disease.
+The second approach has gained more attention in recent years, due to the fact that in clinical practice it has shown promise in the early detection of AD.
+In addition, the analysis of the decline of cognitive abilities is comparitively inexpensive and less invasive that the first approach which commonly requires the collection of a sample of cerebro-spinal fluid which is painful for the patient involved.
+/newline
+/par
 The two main ways in which diagnosis is performed is through assessment of memory and language.
@@
 Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on measure decline in the use of language in those with AD.
+There are a number of difficulties to watch out for with this approach.
 There are a wide number of factors that are involved in language degeneration in the elderly, and consequently there will be an expected amount of variability between subjects.
+The administration of such tests may induce nervousness and discomfort which may impact performance, and also repeatedly administering the same language tests to tests for differences over time by be confounded by improved performance at tasks via practice effects.
+Also, generally speaking the language tests that are currently being used do not necessarily describe patients 'real performance' in language production.
+However, there is enough promise in this approach such that it could help further our understanding of the disease, it's progression and the parts of the brain affected in the early stages.
+\newline
+\par  
 The consensus is that this language degeneration is typically accelerated by the presence of dementia \cite{Berisha2015} and that a potential indicator of dementia is the rate of change in which the decline occurs relative to a fixed point in time rather than a comparison across a cohort of individuals.
@@
 
-\subsection{Aims and Methodology}
+\subsubsection{Aims and Methodology}
 \par
@@
 This resulted in 17 journal articles and/or conference papers which form this review.
-
 \begin{table}
@@
 
-\section {Types of Language Assessment}
+\subsection {Types of Language Assessment}
 One of the key debates when looking at how to analyse language is the type of task provided to elicit language production in participants.
@@
 \par
-\subsection{Picture Description Tasks}
+\subsubsection{Picture Description Tasks}
 One of the most commonly used tasks to measure language is the Picture Description task.
@@
 \par
-\subsection{Narrative description task}
+\subsubsection{Narrative description task}
 The story narration task is designed to study a participant's ability to describe and elaborate on a story which is depicted using a series of pictures.
@@
 \par
-\subsection{Interviews}
+\subsubsection{Interviews}
 Interviews can also be used to elicit language in a more natural way by asking questions to guide a conversation between speakers.
@@
 Nevertheless, this is the most naturalistic setting for looking at language production and can be used to look at the syntactic and semantic parts of language generation \cite{Sajjadi2012}.
-There have been some attempts to use interviews to assess language production in AD with promising results \cite{Asgari2017, Guinn2015 .\newline
+There have been some attempts to use interviews to assess language production in AD with promising results \cite{Asgari2017, Guinn2015} .\newline
 \par
-\section{How do we analyse language, issues and debates}
-\subsection{Single Word Language tasks vs Connected Language tasks}
+\subsection{How do we analyse language, issues and debates}
+\subsubsection{Single Word Language tasks vs Connected Language tasks}
 Part of the reason we need to pay attention to how we ask participants to generate data is understanding how we wish to analyse the data afterwards.
@@
 \par
-\subsection{Semantics vs Pragmatics}
+\subsubsection{Semantics vs Pragmatics}
 When navigating the English language it is necessary to distinguish between what a sentence says in both semantic and pragmatic terms.
@@
 \par
-\section{Semantic Content}
+\subsection{Semantic Content}
 Another approach to linguistic analysis in this field is the idea of measuring semantic content and complexity.
@@
 \par
-\subsection{Picture-related content thematic elements}
+\subsubsection{Picture-related content thematic elements}
 Several studies examined the amount of thematic elements expressed that were directly relevant to picture stimulus in picture description tasks.
@@
 Studies using these features found some differences between AD and controls, and some could differentiate between different stages of AD.
-\subsection{General Information Units}
+\subsubsection{General Information Units}
 Some studies used a more general concept of content, defining "information units" as "the smallest non redundant meaningful fact or inference," whether or not the information conveyed was specific to the context in which the conversation happened.
 Giles et al, for example, studied adults with minimal, mild or moderate AD vs controls and found that adults with AD produced fewer overall information units than controls.
-\subsection{Conciseness of information}
+\subsubsection{Conciseness of information}
 Conciseness has been defined as the number of words a speaker uses to express ideas.
@@
 \par
-\subsection{Efficiency}
+\subsubsection{Efficiency}
 Efficiency is the rate at which meaningful information is conveyed over time, calculated by dividing the total number of information units by the duration in seconds of the speech sample.
@@
 Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in dialogue in which a speaker provides responsees do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls..\newline
-\subsection{Lexical richness and diversity}
+\subsubsection{Lexical richness and diversity}
 \subsubsection{Type token ratio(TTR)}
@@
 where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
-\subsection{Quantity - Total number of words}
+\subsubsection{Quantity - Total number of words}
 Several studies report that adults with moderate AD produce fewer words than controls on picture description, however other studies found no differences in total words among groups of controls and patients with MCI or AD.
@@
 \par
-\section{Syntax and Morphology (Language Form)}
+\subsection{Syntax and Morphology (Language Form)}
 Syntax can be defined as the rules that govern how words can be combined to form sentences, whilst Morphology is the system that governs the structure of words and the construction of word forms.
@@
 \par
-\subsection{N-grams}
+\subsubsection{N-grams}
 One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
@@
 \par
-\subsection{Mean length of utterance (MLU)}
+\subsubsection{Mean length of utterance (MLU)}
 Murray found that MLU was not a distinguishing factor among health adults, adults with depression and adults with AD.
@@
 \par
-\subsection{Proportion of verbs to nouns plus verbs}
+\subsubsection{Proportion of verbs to nouns plus verbs}
 Kave and Levy used a verb index to capture syntactic complexity and found that adults with AD expressed the same amount of verbs to nouns plus verbs as adult controls. \newline
 \par
-\subsection{Syntactic Complexity - Composite measures of MLU, syntactic errors and verbs}
+\subsubsection{Syntactic Complexity - Composite measures of MLU, syntactic errors and verbs}
 Ahmed et al, and Ahmed, Haigh et al found differences in syntactic complexity between adults with MCI and controls, and between MCI and moderate AD stages.
@@
 \par
-\subsection{Formulaic Language}
+\subsubsection{Formulaic Language}
 Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
@@
 \par
-\section{Pragmatic Language}
+\subsection{Pragmatic Language}
 The pragmatic language domain refers to the social rules for language for the purposes of communication including, using language to achieve goals, using information from the context to achieve these goals and using the interaction between people to initiate, maintain and terminate conversations.
-\subsection{Questions, turn-taking, unsure statements, egocentric comments}
+\subsubsection{Questions, turn-taking, unsure statements, egocentric comments}
 One study, Ripich et al, examined several pragmatic abilities among patients at different stages of AD.
@@
 \par
-\subsection{Coherence}
+\subsubsection{Coherence}
 Coherence refers to the appropriate maintenance of topic in discourse and is thematically related to the immediately preceding utterance (local coherence) and by how closely an utterance relates to the general topic at hand (global coherence).
@@
 \par
-\subsection{Perseveration}
+\subsubsection{Perseveration}
 One study examined verbal preservation in the description of Norman Rockwell prints, dividing the total number of words within perseverations by total number of words in the speech sample.
@@
 \par
-\subsection{Speech intonation / prosody}
+\subsubsection{Speech intonation / prosody}
 There are a number of studies, that have examined "melodic line", which is a subjective measure of speech prosody defined as "the appropriate use of intonational contour, including alterations in pitch, volume and duration".
@@
 Konig et al, also used an automated feature analysis examining vocal and temporal aspects of speech among controls, and patients with MCI or AD, and reported a classification of 81\%.
-\subsection{Discourse Fluency}
+\subsubsection{Discourse Fluency}
 Verbal fluency is a term used in neuropsychological contexts generally referring to timed, word-generation tasks, while in speech-language pathology contexts, "fluency disorders" are defined as interruptions int he flow of speaking characterized by atypical rate, rhythm and repetitions in sounds, syllables, words and phrases.
@@
 \par
-\subsection{Speech Monitoring}
+\subsubsection{Speech Monitoring}
 Speech monitoring is related not only to word retrieval deficits and to pragmatic language skill but also to "anosognosia" define as the awareness of one's own deficits.
@@
 
-\section{State of literature into Natural Language processing techniques}
+\subsection{Existing Neuropsychological Measures of Cognitive Impairment, and Repeatable Battery of Neuropsychological Tests}
+In terms of standardised cognitive tests, there are two main aims.
+Does the test distinguish accurately between normal aging, MCI and AD (diagnostic utility) and, does the test distinguish between those individuals with MCI who will then go on to develop AD and those individuals with MCI who don't then go on to develop AD (prognostic utility).
+This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their perfomance in terms of both diagnostic and prognostic utility.
+\newline
+\par 
+\subsubsection{RBANS - Repeated Battery for the Assessment of Neuropsychological Status}
+\subsubsection{Patient Health Questionnaire)}
+\subsubsection{WASI}
+\subsubsection{Digit Span}
+\subsubsection{Rey Words}
+\subsubsection{Prospective Memory Recall}
+\subsubsection{Digit symbol coding}
+\subsubsection{Corsi Span}
+\subsubsection{Verbal Fluency}
+%% In this paragraph cite Henry et al, 2004 - taken from Taler and Phillips (2008) - Henry et al, 2004 is a meta analysis
+There are a number of tests that characterise this category of verbal fluency, but generally fall into two categories.
+Letter (Phonemic) fluency involves the generation of as many words as possible which begin with a specified letter.
+Category fluency involves the generation of as many words as possible that fall into a specfied category.
+Both these tasks impose demands on a number of different cognitive processes namely, executive function, verbal retrieval and recall, giving appropriate answers while monitoring previous answers and inhibit inappropriate responses.
+However, these two tasks require different strategies when attempting them.
+Letter fluency relies on search strategies based on lexical representations whereas category fluency requires a search for semantic extensions of a superordinate term, meaning that semantic associations within the lexicon must be intact in order for the task to be carried out successfully.
+\par
+There have been numerous studies which have documented the impact that AD has on verbal fluency tests in both categories.
+A review carried out by Henry et al, found that performance in both letter fluency and category fluency was impaired in those with AD vs controls, but found a larger effect for tests of semantic fluency.
+\subsubsection{Rey Delayed Recall}
+\subsubsection{Naming - Speeded}
+
+\subsubsection{Wisconsin Card Sorting}
+\subsubsection{Complex Figure Task immediate and delayed recall}
+\subsubsection{Cinderella Story - Narrative Description}
+\subsubsection{Sentence Completion}
+\subsubsection{Non-word Repetition}
+\subsubsection{Grooved Peg Board}
+\subsubsection{Visual Search}
+\subsubsection{Spelling}
+\subsubsection{Continuous Picture Naming}
+\subsubsection{Free Cued Selective Reminding Test}
+The Free Cued Selective Reminding Test (FCSRT) was borne out of the premise that by controlling the conditions of learning, a measure of memory is possible that is not confounded by normal age related changes in cognition.
+Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cogntive decline due to a cognitive impairment.
+/newline
+/par
+FCSRT-IR (The instant recall version of the Free Cued Selective Reminding Test) performance has beeen shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the authors to define prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT-IR performance (Sarazin et al, 2007 /newline
+/par
+The sensitivity and specificity of the different test forms in correctly classifying patients were compared to each other using a previously established FCSRT-IR cut off score. /newline
+/par
+
+\subsubsection{ Description (Cookie Theft Picture)}
+
+\subsection{State of literature into Natural Language processing techniques}
 As we can see, diagnosing dementia through analysis has a large background in terms of psychological research.
@@
 \end{enumerate}
-\subsection{Traditional methods of Machine Learning and Natural Language Processing}
+\subsubsection{Traditional methods of Machine Learning and Natural Language Processing}
 
-\subsection{Deep Learning}
+\subsubsection{Deep Learning}
 
-\section{The problems with attempting early diagnosis of dementia and the ethics of actively screening for it}
+\subsection{The problems with attempting early diagnosis of dementia and the ethics of actively screening for it}
 It is challenging to characterise Dementia as a disease as it's more syndrome.
@@
 \par
-\section{Conclusions and Future Work}
+\subsection{Conclusions and Future Work}
 It appears that there are a number of approaches that have been shown to be effective in various experimental settings.
@@
 \par
-\subsection{Future Work}
+\subsubsection{Future Work}
 Future research should also be directed towards developing non-intrusive ways of detecting subtle changes in language based in the home, such that any deterioration that could indicate the presence of MCI or AD and be flagged up early in the progression of any potential cogntive impairment for further review via referral.
@@
 
+\section{Project Proposal}
+\textbf{Full Title of the Research:} Title here \newline
+\textbf{Research Questions and Hypotheses}
+The broad question is can we use Machine Learning to assist in the early diagnosis of dementia and/or mild cognitive impairment through language produced naturally rather than language that is produced as part of a formal cognitive assessment.
+There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia.
+These studies have identified a number of speech and language differences between healthy controls and those with dementia.
+Some of these features have been validated in work using machine learning to classify healthy controls from those with dementia.
+However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'.
+While this is help to learn that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments.
+\newline
+\par
+Machine Learning itself is undergoing a shift in focus at the present time.
+I will describe this shift using the context of my proposed research.
+In traditional machine learning, the transcripts from cognitive assessments are used to generate features such as idea density.
+These features are then used to classify patients.
+A company in Toronto, Canada has taken this approach and currently uses over 400 acoustic and language features to make this classification and to generate a psuedo MMSE score.
+\newline
+\par
+Deep Learning takes a different approach by bypassing the feature generating stage. Deep learning uses neural networks as a fundamental building block.
+
+\subsection{Aims and Rationale}
+The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) 	and without resorting to to time-consuming and expensive psychological tests.
+There is a lot of research into the analysis of language as a bio-marker for MCI and Early Dementia.
+Given 			that sampling a person's language is relatively effortless, my research looks at whether we can find bio-markers of MCI and Early Dementia in natural language.\newline
+Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
+I'm not going to control for memory problems as a 			potential confound, but does this weaken the research?
+How do I defend this?
+\newline
+\begin{itemize}
+	\item Hypothesis.
+I predict that, given a number of samples of language over a given time from, is it possible to detect language decline in humans such that you can differentiate between normal language decline and decline in those with MCI.
+	\item Hypothesis. I predict that my new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures.
+	\item My first hypothesis is whether monitoring language samples and other measures of cognitive decline over the course of the year is enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment.
+	\item My second hypothesis is that those with MCI show a different language profile to controls, if matched similarly for education and other significant factors.
+\end{itemize}	
+\section{Experiment 1 - Pilot Study of Neuropsychological Tests}
+\subsection{Methods}
+I recruited a cohort of 9 volunteers which which to complete the battery of Neuropsychological tests.
+We compiled a series of tests (see literature review for a comprehensive list, explanation and rationale for each test) and completed each version of the test on three consequetive days.
+The results of which were compiled and analysed.
+The hypothesis is that there would be no difference in performance between one battery of tests and another.
+An additional hypothesis is that the language tasks included on each task would be enough to collect an appropriate sample for linguistic analysis, such that we may be able to detect cognitive decline in a clinical population.
+\newline
+\par
+\subsubsection{Participants}
+\subsection{Results}
+\subsection{Discussion}
+\subsection{Conclusion}
+\section{Experiment 2 - Initial Study of the Presidents corpus}
+\subsection{Background}
+\subsection{Methods}
+I took 46 transcripts of Ronald Reagan’s (RR) press conferences from 1981 to 1988 and compared them with 134 press conferences by George H.
+W.
+Bush (GHWB) and 25 press conferences conducted by Donald J.
+Trump (DJT).
+I analyzed transcripts for lexical features shown to change longitudinally with dementia  (for a comprehensive review of these, see the literature review above).
+For this collection of documents, I generated a number of features which looked at a number of different aspects of each document.
+These features encompassed, word level, sentence level and document level features and included a number of features contained in the study by Berisha and Liss with the aim of replicating and extending on their findings.
+These findings were.
+number of unique words, non-specific nouns and fillers and low imageability (LI) verbs.
+Imageability is characterized, according to Berisha and Liss, as the ease with which a term gives rise to a sensory .mental image.
+I compared the trends described in the transcripts of RR and GHWB, but also included DT.
+Berisha and Liss originally made the comparison as it GHWB (GHWB - age at the start of presidency - 64 years, 222 days) was the closest match to RR in terms of age (RR  - age at the start of presidency, 69 yeaars and 349 days).
+However, with the inauguration of Trump, he now is the closest comparible president in terms of age (DJT - age at the start of presidency - 70 years, 220 days).
+It would be interesting to look at a comparison of RR and DJT to see whether the comparisons made by Berisha and Liss hold true with this more appropriate match (in terms of age).
+DJT as with GHWB has no known diagnosis of AD.
+I used the press conference transcripts in the American Presidency Project (APP) archive as a data source for this project.
+The APP is a comprehensive and organized searchable database of presidential documents, including transcripts of speeches, transcripts of news conferences, and other public documents.
+\subsubsection{Pre-processing}
+To generate the files necessary for analysis, I downloaded each transcript and performed the following changes.
+I omitted the prepared statement by the president and any speech by other individuals.
+I started each transcript at the beginning of the first answer to a question.
+I filtered any annotations that were added to the transcript, including any references or clarifications, and any laughter.
+It's worth noting that there appears to be a difference in how 'hesitations' were marked down between each president, for RR hesitations were marked by a single hyphen whereas for GWB hesitations are marked by a double hyphen.
+In order to maintain consistency when parsing through the documents, I have changed both types of hesitation to be marked by a single hyphen.
+I also omitted one word sentences as this data would, from a theoretical perspective, not be relevant for language analysis.
+I did not control for the length of the document, but generated features which would normalise by the length of the document.
+I therefore was able to include all press conferences by both Ronald Reagan and George H.W.
+Bush where there was a question and answer session conducted at least in part by the sitting president (2 press conferences of GWB were ommited due to a lack of a question and answer session).
+\subsection{Results}
+\subsection{Discussion}
+\subsection{Conclusion}
+
+\subsection{Timeline of Future Work and Publication plan}
+\textbf{Year One} \newline
+\textbf{Year Two} \newline
+\textbf{Year Three} \newline
+
+\subsection{Provisional Table of Contents}
+\subsubsection{Chapter 1: Introduction}
+Introduction to the problem of dementia in the context of the wider world including quality of life and financial implications.
+Exploration of dementia as a syndrome rather than a disease, and a look at the different variants of dementia.
+A look at the rationale behind research into the early diagnosis of dementia as well as a brief look at what has been done in the area (wide context, so pharmacological and psychological).
+\newline
+\subsubsection{Chapter 2: Background}
+Literature Review which looks at the background of what has been done in the area of early diagnosis of dementia, from a purely psychological perspective as well as the work done in machine learning in this topic.
+Look at the different attributes that researchers have explored.
+Look at the difference between traditional machine learning methods and deep learning and discuss the pros and cons of each perspective.
+\newline
+\subsection{Chapter 3: Experiments with existing datasets: Presidential Press Conferences, Three Authors - Murdoch, Christie and James, DementiaBank dataset.}
+Aim: To replicate and extend the work by Berisha and Liss.
+To use traditional machine learning techniques to test whether they can differentiate between Reagan, Bush and Trump using similar features.
+To test to see whether Deep Learning can also differentiate between the three.
+\newline
+Dataset: the presidential press conferences given by Presidents Ronald Reagan, George H.W. Bush and Donald Trump. \newline
+Paper: This experiment replicates work done by Berisha and Liss and extends this by adding Donald Trump as an alternative comparison to Ronald Reagan.
+This experiment will look at the features originally recommended by Berisha and Liss, as well as any others that have potential as discussed in the literature review above.
+\newline
+Aim: To replicate the work by Le et al.
+To use traditional machine learning techniques to see if it can differentiate between the three authors using similar features.
+To test to see whether deep learning can also differentiate between the three.
+To add 'pre-trained' layers from Presidential Debates experiment to see if this improves accuracy.
+\newline
+Dataset is the literary novels of Iris Murdoch, Agatha Christie and P.D. James. \newline
+Paper:
+Aim: To replicate and extend the work of Orimaye et al.
+To use traditional machine learning techniques to see if it can categorise people into dementia or healthy categories using the samples given.
+To see whether can use pre-trained layer from presidential debates to improve accuracy.
+It doesn't seem relevant to use pretrained layer from Experiment four, but it might be worth looking at the results of this.
+\newline
+Dataset: The DementiaBank Corpus. \newline
+Paper: To replicate and extend the work of Orimaye et al. To compare traditional machine learning techniques (already completed by Orimaye) and deep learning. \newline
+\subsection{Chapter 4: Validation of new repeatable neuropsychological battery for detecting MCI.}
+Aim: \newline
+Dataset: To be created.
+
+\subsection{Chapter 5: Longditunal Study }
+\subsubsection{Dataset creation}
+The aim is to recruit a set of 50 participants with MCI / Early Dementia (diagnosed or undiagnosed) and a set of 50 controls with a similar age and broadly similar educational background (a known potential confound in assessing cognitive decline).
+The rationale for recruiting this number is that there will be an expected drop out rate and so
+Paper:
+\subsubsection{Ethics}
+Need to discuss confidentiality right at the start of each conversation and to reiterate at every meeting.
+I will also provide an information sheet which details the confidentiality statement.
+The spoken confidentiality statement will be as follows.
+"Anything you say to me today will be confidential unless I feel that there is a risk to you, or someone else.
+If this is the case, I may need to speak to or write to third parties such as your GP.
+Is this ok?" \newline
+\par
+\subsubsection{What if someone reveals suicidal ideation or other risks?}
+As part of the battery of tests for this experiment, there is an opportunity for a participant to reveal that he or she has had suicidal ideation in the past two weeks (Patient Health Questionnaire, Question 9).
+This is something to explore further.
+As I have been trained in psychological risk assessments, it seems appropriate to carry out a further risk assessment.
+Optionally with the tape recorder turned on, or off as requested.
+Any text recorded during the risk assessment will not be included in the language analysis.
+Following this, some information about local services such as Improving Access to Psychological Therapies (IAPT), Samaritans and other appropriate services will be provided.
+A discussion will then be made about whether it is appropriate to notify the GP of the disclosure.
+In the event that there is imminent risk to self or others, a discussion will be made about who I need to contact.
+(Crisis Team and GP).
+In the event that the participant scores 0 on Q9 of the PHQ, no risk assessment will be undertaken.
+\newline
+\subsubsection{What happens at the end of the data collection phase if there is a suspected decline in a participant's cognitive function}
+\subsection{How does language differ between those with AD, MCI and Controls}
+\subsection{Chapter 7: Overall results and Discussion}
+\subsection{Chapter 8: Conclusions and Future work}
+
 \bibliographystyle{unsrt}
-\bibliography{LiteratureReview}
-\end {document}
+\bibliography{QualifyingReport.bib}
+
+\appendix
+\section{Title of Appendix A}
+Text of Appendix A is Here
+
+\section{Title of Appendix B}
+Text of Appendix B is Here
+
+\section{Title of Appendix C}
+Text of Appendix C is Here.
+
+\end{document}
+%DOCUMENT ENDS

### 2018-12-30 Source/Thesis/Thesis.tex
added source

--- 2018-08-28 LiteratureReview/LiteratureReview.tex
@@
+%% TO DO LIST
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% Document Class
-\documentclass[a4paper]{article}
+\documentclass[a4paper]{report}
 %% Packages
@@
 \input{settings/macros}
-\title{Literature Review}
-\date{2018-08-22}
+\title{Thesis}
+\date{2018-09-17}
 \author{Jomar Alcantara}
@@
 \newpage
+% Abstract
+\section{Abstract}
+% Acknowledgements
+\section{Acknowledgements}
 \tableofcontents
 \newpage
+\listoffigures
+\listoftables
+\chapter{Introduction}
+The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to to time-consuming and expensive psychological tests.
+There is a lot of research into the analysis of language as a bio-marker for MCI and Early Dementia.
+Given that sampling a person's language is relatively effortless, my research looks at whether we can find bio-markers of MCI and Early Dementia in natural language.\newline
+Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
+I'm not going to control for memory problems as a potential confound, but does this weaken the research?
+How do I defend this?
+\newline
+\par 
+Introduction to the problem of dementia in the context of the wider world including quality of life and financial implications.
+Exploration of dementia as a syndrome rather than a disease, and a look at the different variants of dementia.
+A look at the rationale behind research into the early diagnosis of dementia as well as a brief look at what has been done in the area (wide context, so pharmacological and psychological).
+\newline
+\chapter{Background}
 \section{Introduction}
 \par
-% IN THIS PARAGRAPH - CITE UK REPORT
-Alzheimer's disease and other forms of dementia are affecting a significant proportion of geriatric population in the world today.
+Alzheimer's disease and other forms of dementia affect a significant proportion of the geriatric population in the world today and is currently the sixth leading cause of death in the US and the leading killer of women in the UK.
 According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
-This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population.
+This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population \cite{Sturrock2016}.
 From a financial perspective, they estimate an annual spend of £4.3 billion of which approximately £85 million is spent solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
 Globally, this picture is a lot bleaker.
-Another report by Alzheimer's Disease International suggests that in 2015 there were 46 million people with a diagnosis of AD and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
-The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.\newline
+Another report by Alzheimer's Disease International suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
+\newline
+\par
+% In this paragraph, look for the study which actually has these figures.
+At present there are no drugs that improve the prognosis of those suffering with AD, all the drugs that are on the market are designed to manage symptoms.
+Whilst there are numerous investigational drugs for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{Alzheimer2016}.
+A possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place.
+Research has therefore been aimed at earlier diagnosis.
+\newline
 \par 
@@
 \section{Semantic Content}
+% Bridge sentence, idea density and type token ratio.
 Another approach to linguistic analysis in this field is the idea of measuring semantic content and complexity.
@@
 However, a lot of more current research has called for the use of machine learning as a way of assisting in the process of diagnosis \cite{Boschi2017}.
+\subsection{Traditional methods of Machine Learning and Natural Language Processing}
+
+\subsection{Deep Learning}
 
@@
 %focus to preserve cognitive functions
-
+\section{The problems with attempting early diagnosis of dementia and the ethics of actively screening for it}
+It is challenging to characterise Dementia as a disease as it's more syndrome.
+The clinical features that combine to meet the diagnostic criteria for Dementia or it's variants are continuous in nature and also impacted by other variables.
+For example, cognitive performance is affected by previous education and ability to live independently is impacted by a person's physical health.
+Also, the individual who may develop symptoms that indicate the onset of dementia needs the insight to recognise them, or family members will need to do so and also deduce that there is value to admitting there is a problem and seeking help before the deterioration is too great.
+\newline
+\par 
+To detect dementia early requires an understanding of how this presents early in it's progression, but as described above there are a number of potential ways in which individuals with dementia could be impacted but no consistent list of criteria has been produced.
+Even for AD, the most prevalent form of dementia, clinical diagnosis can only be confirmed after death through autopsy.
+In addition, population studies have shown that people over the age of 80 show similar changes in the brain in those that do not have dementia.
+Therefore, it's difficult to say with certainty that any symptoms can definitely cause dementia despite the fact that these symptoms are commonly present in those that have dementia.
+\newline
+\par 
 \section{Conclusions and Future Work}
-It appears that both frequency based approaches and approaches that use measures of linguistic complexity have both been shown to be effective in various experimental settings.
-The fact that there are two vastly different perspectives to tackling this question, and even within these perspectives there are a range of various methodologies leaves the field with a sense of confusion as to the most effective way to solving the problem.
+It appears that there are a number of approaches that have been shown to be effective in various experimental settings.
+The fact that there are a numerous array of different perspectives to tackling this question, and even within these perspectives there are a range of various methodologies leaves the field with a sense of confusion as to the most effective way to solving the problem.
 \newline
@@
 \newline
+\chapter{Research Questions and Hypotheses}
+The broad question is can we use Machine Learning to assist in the diagnosis of dementia through natural language.
+There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia.
+These
+
+\chapter{Experiment: Presidential Debates revisited}
+\section{Background}
+Aim: To replicate and extend the work by Berisha and Liss.
+To use traditional machine learning techniques to test whether they can differentiate between Reagan, Bush and Trump using similar features.
+To test to see whether Deep Learning can also differentiate between the three.
+\newline
+\section{Data}
+Dataset: the presidential press conferences given by Presidents Ronald Reagan, George H.W. Bush and Donald Trump. \newline 
+Paper: This experiment replicates work done by Berisha and Liss and extends this by adding Donald Trump as an alternative comparison to Ronald Reagan.
+This experiment will look at the features originally recommended by Berisha and Liss, as well as any others that have potential as discussed in the literature review above.
+\newline
+\chapter{Experiment: Three Authors - Murdoch, Christie and James}
+\section{Background}
+\section{Data}
+Aim: To replicate the work by Le et al.
+To use traditional machine learning techniques to see if it can differentiate between the three authors using similar features.
+To test to see whether deep learning can also differentiate between the three.
+To add 'pre-trained' layers from Presidential Debates experiment to see if this improves accuracy.
+\newline
+Dataset is the literary novels of Iris Murdoch, Agatha Christie and P.D. James. \newline
+Paper: 
+\chapter{Experiment: Using the DementiaBank Corpus}
+Aim: To replicate and extend the work of Orimaye et al.
+To use traditional machine learning techniques to see if it can categorise people into dementia or healthy categories using the samples given.
+To see whether can use pre-trained layer from presidential debates to improve accuracy.
+It doesn't seem relevant to use pretrained layer from Experiment four, but it might be worth looking at the results of this.
+\newline
+Dataset: The DementiaBank Corpus. \newline
+Paper: To replicate and extend the work of Orimaye et al. To compare traditional machine learning techniques (already completed by Orimaye) and deep learning. \newline
+\chapter{Experiment: Longitudinal Data - Current Cohort}
+Aim: \newline
+Dataset: To be created.
+\subsection{Dataset creation}
+The aim is to recruit a set of 50 participants with MCI / Early Dementia (diagnosed or undiagnosed) and a set of 50 controls with a similar age and broadly similar educational background (a known potential confound in assessing cognitive decline).
+The rationale for recruiting this number is that there will be an expected drop out rate and so
+\subsubsection{Patient Health Questionnaire (PHQ-9)}
+\subsubsection{Free Cued Selective Reminding Test (FCSRT)}
+\subsubsection{Mini Mental State Examination(MMSE)}
+\subsubsection{Written Description (Cookie Theft Picture)}
+\subsubsection{Semi-Structured Interview}
+
+Paper:
+
+\subsection{Ethics} 
+Need to discuss confidentiality right at the start of each conversation and to reiterate at every meeting.
+I will also provide an information sheet which details the confidentiality statement.
+The spoken confidentiality statement will be as follows.
+"Anything you say to me today will be confidential unless I feel that there is a risk to you, or someone else.
+If this is the case, I may need to speak to or write to third parties such as your GP.
+Is this ok?" \newline
+\par
+\subsection{What if someone reveals suicidal ideation or other risks?}
+As part of the battery of tests for this experiment, there is an opportunity for a participant to reveal that he or she has had suicidal ideation in the past two weeks (Patient Health Questionnaire, Question 9).
+This is something to explore further.
+As I have been trained in psychological risk assessments, it seems appropriate to carry out a further risk assessment.
+Optionally with the tape recorder turned on, or off as requested.
+Any text recorded during the risk assessment will not be included in the language analysis.
+Following this, some information about local services such as Improving Access to Psychological Therapies (IAPT), Samaritans and other appropriate services will be provided.
+A discussion will then be made about whether it is appropriate to notify the GP of the disclosure.
+In the event that there is imminent risk to self or others, a discussion will be made about who I need to contact.
+(Crisis Team and GP).
+In the event that the participant scores 0 on Q9 of the PHQ, no risk assessment will be undertaken.
+\newline
+\subsection{What happens at the end of the data collection phase if there is a suspected decline in a participant's cognitive function}
+
+\chapter{Overall results and Discussion}
+\chapter{Conclusions and Future work}
 
 \bibliographystyle{unsrt}
-\bibliography{LiteratureReview}
+\bibliography{Bibliography}
+
 \end {document}

### 2019-01-01 Source/QualifyingReport/QualifyingReport.tex
updated QR

@@
 \usepackage{float}
+\usepackage{parskip}
 %% Page Settings
@@
 \title{Qualifying Report}
-\date{2018-12-18}
+\date{2018-12-30}
 \author{Jomar Alcantara}
@@
 \newpage
-\section{Overview of Problem and potential contribution.}
+\section{Overview of Problem}
 Dementia has been identified as one of those fast growing difficulties facing the world.
@@
 The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
-\newline
 \par
-In 2009, the Department of Health in it's National Dementia strategy made 'early diagnosis and support' one of it's key themes as part of this strategy.
-With that in mind, a lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cogntive Impairment (MCI).
-As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages and thus it is important that we utilise technology wherever possible to aid clinicians.
+In 2009, the Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities.
+A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cogntive Impairment (MCI) with research focused two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease.
+As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
 At the present time diagnosis is typically conducted at memory clinics by trained clinicians.
-I theorise that we might be able to use diagnose patients in their own homes utilising smart-home technology.
-\newline
+I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using natural language.
 \par
-Language analysis is an important part of assessing those suspected of suffering with AD or MCI and this language analysis in part informs an official clinical diagnosis.
-This analysis often takes the form of tasks such as Free Cued Selective Reminding Task and the picture description task which are all quite targeted in the language required of the person being assessed.
-Other studies have analysed more spontaneous samples of language and have identified patterns of speech which decline.
-\newline
-\par
-A step towards the ambient detection of AD and/or MCI in population requires that we look for similar features in language used naturally as part of every day conversation.
-Part of this work therefore is to see whether spontaneous discourse, such a semi-structured interview is a task that has the ability to put pressure both the cognitive and linguistic systems as it requires the ability to understand speech, formulating a coherent answer and then responding in an appropriate way.
-Berisha et al \cite{Berisha2015}, has shown through language analysis in this area that there are marked differences in this process between those who have a diagnosis of AD and healthy controls.
-\newline
+There is a large body of research that looks at the decline in language in those with MCI and AD.
+However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD.
+Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
+One of the difficulties in research of this nature is the process of collecting appropriate language samples.
+Part of this work therefore is to see whether spontaneous discourse, such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD.
+There is some evidence to support this.
+Berisha et al \cite{Berisha2015}, has shown through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control.
 \par
 The potential impact of this research is immense.
-Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can slow the progress of the disease.
-Equally, early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
+Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease.
+Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
 \section{Literature Review}
 \subsection{Introduction}
-\par
-Alzheimer's disease (AD) and other forms of dementia affect a significant proportion of the geriatric population in the world today and is currently the sixth leading cause of death in the US and the leading killer of women in the UK.
+AD and other forms of dementia affect a significant proportion of the geriatric population in the world today and is currently the sixth leading cause of death in the US and was named the leading killer of women in the UK.
 According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
@@
 The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
-\newline
 \par
@@
 Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place.
-\newline
 \par
@@
 In addition, these symptoms are often accompanied by emotional difficulties such as depression and irritability and behavioural difficulties.
-\newline
 \par
@@
 If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
-However, one of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of dementia.\newline
+However, one of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of dementia.
 \par
@@
 Given the burden on these professionals is likely to increase due to a growing elderly population or in the case of developing countries where the clinician / patient ratio is already unsustainable, it seems useful to find less burdensome ways to aid professionals in identifying those at risk of developing dementia.
-\newline
 \par
@@
 In addition, the analysis of the decline of cognitive abilities is comparitively inexpensive and less invasive that the first approach which commonly requires the collection of a sample of cerebro-spinal fluid which is painful for the patient involved.
-/newline
 /par
@@
 However, there is enough promise in this approach such that it could help further our understanding of the disease, it's progression and the parts of the brain affected in the early stages.
-\newline
 \par  
 The consensus is that this language degeneration is typically accelerated by the presence of dementia \cite{Berisha2015} and that a potential indicator of dementia is the rate of change in which the decline occurs relative to a fixed point in time rather than a comparison across a cohort of individuals.
-\newline
 \par
@@
 Echolalia (meaningless repetition of another person's spoken words) and automatic speech typically precede mutism.
-With the wide range of deficits someone with AD can suffer, it makes sense to try to categorise these deficits in some way.\newline
+With the wide range of deficits someone with AD can suffer, it makes sense to try to categorise these deficits in some way.
 \par
@@
 As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are extremely variable and can differ from patient to patient as the disease progresses.
-\newline
 \par
@@
 She stated that the language forms we learn last are the first to deteriorate.
-We generally learn language in small simple units initially and build syntax and complexity as we are more comfortable with language.\newline
+We generally learn language in small simple units initially and build syntax and complexity as we are more comfortable with language.
 \par
@@
 The field of machine learning and natural language processing has been suggested as a way to improve the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline \cite{Boschi2017}.
-\newline
-
+% -- END OF INTRODUCTION --
 \subsubsection{Aims and Methodology}
 \par
-The purpose of this review is to seek to understand what techniques for assessing language have been used within the field of cognitive and clinical psychology.
-We then go on to look at what techniques have been developed in the field of machine learning (ML), deep learning (DL) and natural language processing (NLP) that might enable the automated analysis of language easier as well as an obstacles and/or limitations of current technology.
-Finally, we look at some studies which have already looked at the intersection of these two domains.
+The purpose of this review is to look at the current standards of cognitive tests available that can be used as a diagnostic tool for MCI.
+I then go on to explore what techniques for assessing language have been used within the field of cognitive and clinical psychology.
+I then go on to look at what techniques have been developed in the field of machine learning (ML), deep learning (DL) and natural language processing (NLP) that might enable the automated analysis of language easier as well as an obstacles and/or limitations of current technology.
+Finally, I look at some studies which have already looked at the intersection of these two domains.
 \newline
@@
 \end{table}
+% -- END OF Aims and Methodology
+\subsection{Existing Neuropsychological Measures of Cognitive Impairment, and Repeatable Battery of Neuropsychological Tests}
+In terms of standardised cognitive tests, there are two main aims.
+Does the test distinguish accurately between normal aging, MCI and AD (diagnostic utility) and, does the test distinguish between those individuals with MCI who will then go on to develop AD and those individuals with MCI who don't then go on to develop AD (prognostic utility).
+This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their perfomance in terms of both diagnostic and prognostic utility.
+\newline
+\par 
+\subsubsection{RBANS - Repeated Battery for the Assessment of Neuropsychological Status}
+\subsubsection{Patient Health Questionnaire)}
+\subsubsection{WASI}
+\subsubsection{Digit Span}
+\subsubsection{Rey Words}
+\subsubsection{Prospective Memory Recall}
+\subsubsection{Digit symbol coding}
+\subsubsection{Corsi Span}
+\subsubsection{Verbal Fluency}
+%% In this paragraph cite Henry et al, 2004 - taken from Taler and Phillips (2008) - Henry et al, 2004 is a meta analysis
+There are a number of tests that characterise this category of verbal fluency, but generally fall into two categories.
+Letter (Phonemic) fluency involves the generation of as many words as possible which begin with a specified letter.
+Category fluency involves the generation of as many words as possible that fall into a specfied category.
+Both these tasks impose demands on a number of different cognitive processes namely, executive function, verbal retrieval and recall, giving appropriate answers while monitoring previous answers and inhibit inappropriate responses.
+However, these two tasks require different strategies when attempting them.
+Letter fluency relies on search strategies based on lexical representations whereas category fluency requires a search for semantic extensions of a superordinate term, meaning that semantic associations within the lexicon must be intact in order for the task to be carried out successfully.
+\par
+There have been numerous studies which have documented the impact that AD has on verbal fluency tests in both categories.
+A review carried out by Henry et al, found that performance in both letter fluency and category fluency was impaired in those with AD vs controls, but found a larger effect for tests of semantic fluency.
+\subsubsection{Rey Delayed Recall}
+\subsubsection{Naming - Speeded}
+
+\subsubsection{Wisconsin Card Sorting}
+\subsubsection{Complex Figure Task immediate and delayed recall}
+\subsubsection{Cinderella Story - Narrative Description}
+\subsubsection{Sentence Completion}
+\subsubsection{Non-word Repetition}
+\subsubsection{Grooved Peg Board}
+\subsubsection{Visual Search}
+\subsubsection{Spelling}
+\subsubsection{Continuous Picture Naming}
+\subsubsection{Free Cued Selective Reminding Test}
+The Free Cued Selective Reminding Test (FCSRT) was borne out of the premise that by controlling the conditions of learning, a measure of memory is possible that is not confounded by normal age related changes in cognition.
+Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cogntive decline due to a cognitive impairment.
+/newline
+/par
+FCSRT-IR (The instant recall version of the Free Cued Selective Reminding Test) performance has beeen shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the authors to define prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT-IR performance (Sarazin et al, 2007 /newline
+/par
+The sensitivity and specificity of the different test forms in correctly classifying patients were compared to each other using a previously established FCSRT-IR cut off score. /newline
+/par
 
@@
 
-\subsection{Existing Neuropsychological Measures of Cognitive Impairment, and Repeatable Battery of Neuropsychological Tests}
-In terms of standardised cognitive tests, there are two main aims.
-Does the test distinguish accurately between normal aging, MCI and AD (diagnostic utility) and, does the test distinguish between those individuals with MCI who will then go on to develop AD and those individuals with MCI who don't then go on to develop AD (prognostic utility).
-This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their perfomance in terms of both diagnostic and prognostic utility.
-\newline
-\par 
-\subsubsection{RBANS - Repeated Battery for the Assessment of Neuropsychological Status}
-\subsubsection{Patient Health Questionnaire)}
-\subsubsection{WASI}
-\subsubsection{Digit Span}
-\subsubsection{Rey Words}
-\subsubsection{Prospective Memory Recall}
-\subsubsection{Digit symbol coding}
-\subsubsection{Corsi Span}
-\subsubsection{Verbal Fluency}
-%% In this paragraph cite Henry et al, 2004 - taken from Taler and Phillips (2008) - Henry et al, 2004 is a meta analysis
-There are a number of tests that characterise this category of verbal fluency, but generally fall into two categories.
-Letter (Phonemic) fluency involves the generation of as many words as possible which begin with a specified letter.
-Category fluency involves the generation of as many words as possible that fall into a specfied category.
-Both these tasks impose demands on a number of different cognitive processes namely, executive function, verbal retrieval and recall, giving appropriate answers while monitoring previous answers and inhibit inappropriate responses.
-However, these two tasks require different strategies when attempting them.
-Letter fluency relies on search strategies based on lexical representations whereas category fluency requires a search for semantic extensions of a superordinate term, meaning that semantic associations within the lexicon must be intact in order for the task to be carried out successfully.
-\par
-There have been numerous studies which have documented the impact that AD has on verbal fluency tests in both categories.
-A review carried out by Henry et al, found that performance in both letter fluency and category fluency was impaired in those with AD vs controls, but found a larger effect for tests of semantic fluency.
-\subsubsection{Rey Delayed Recall}
-\subsubsection{Naming - Speeded}
-
-\subsubsection{Wisconsin Card Sorting}
-\subsubsection{Complex Figure Task immediate and delayed recall}
-\subsubsection{Cinderella Story - Narrative Description}
-\subsubsection{Sentence Completion}
-\subsubsection{Non-word Repetition}
-\subsubsection{Grooved Peg Board}
-\subsubsection{Visual Search}
-\subsubsection{Spelling}
-\subsubsection{Continuous Picture Naming}
-\subsubsection{Free Cued Selective Reminding Test}
-The Free Cued Selective Reminding Test (FCSRT) was borne out of the premise that by controlling the conditions of learning, a measure of memory is possible that is not confounded by normal age related changes in cognition.
-Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cogntive decline due to a cognitive impairment.
-/newline
-/par
-FCSRT-IR (The instant recall version of the Free Cued Selective Reminding Test) performance has beeen shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the authors to define prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT-IR performance (Sarazin et al, 2007 /newline
-/par
-The sensitivity and specificity of the different test forms in correctly classifying patients were compared to each other using a previously established FCSRT-IR cut off score. /newline
-/par
-
-\subsubsection{ Description (Cookie Theft Picture)}
-
 \subsection{State of literature into Natural Language processing techniques}

### 2019-01-05 Source/QualifyingReport/QualifyingReport.tex
test
Added extra to QR

@@
 \subsection{Introduction}
+Alzheimer's Disease is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
+Whilst a definitive diagnosis of Dementia can only produced at post-mortem, there are a number of clinical indicators from psychological perspective that can indicate that dementia is present.
+From a clinical psychology perspective, those who have dementia have cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits\cite{AmericanPsychiatricAssociation2013}.
+In addition, these symptoms are often accompanied by emotional difficulties such as depression and irritability and behavioural difficulties.
+\par
 AD and other forms of dementia affect a significant proportion of the geriatric population in the world today and is currently the sixth leading cause of death in the US and was named the leading killer of women in the UK.
@@
 Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place.
-\par
-Alzheimer's Disease is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
-Whilst a definitive diagnosis of Dementia can only produced at post-mortem, there are a number of clinical indicators from psychological perspective that can indicate that dementia is present.
-From a clinical psychology perspective, those who have dementia have cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits\cite{AmericanPsychiatricAssociation2013}.
-In addition, these symptoms are often accompanied by emotional difficulties such as depression and irritability and behavioural difficulties.
+It is therefore important to focus AD at it's earliest stages which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.
 \par
@@
 However, one of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of dementia.
+Albert et al have worked to define clinical criteria which professionals can use to diagnose MCI due to AD and differentiate this from age-associated memory impairment and age-associated cognitive decline.
+They note that the diagnosis of MCI requires evidence of intraindividual change and optimally requires evaluation at two or more points.
+\begin{enumerate}
+	\item Concern regarding a change in cognition - A person or an informant should express concern that there is a change in cognitive ability in comparison to previous level of performance.
+	\item Impairment in one or more cognitive domains - There should be evidence of lower performance in one or more cognitive domains beyond what would be expected of a person given their age and education.
+	\item Preservation of independence in functional - Whilst persons with MCI are expected to be able to maintain independence, it is common to experience mild problems in complex functional tasks which they may have been able to perform previously.
+This might mean that they take more time or be less efficient at completing these tasks, or it may be that may make more mistakes.
+	\item Not demented - The deterioration should be mild to the point that there is no significant loss of functioning in social or occupational contexts.
+\end{enumerate}
+In addition to meeting the above criteria, a clinician must rule out other conditions or factors that could account for the decline in cognition with the goal to increase the likelihood that the underlying cause of this decline is AD.
+\par 
+Many researchers have studied the early detection of MCI \ AD.
+These studies usually follow two main approaches.
+The analysis of biomarkers and the examination of patients who have demonstrated decreasing cognitive abilities.
+The first approach yields reliable results in the detection of AD in its moderate and advanced states but does not perform well during the early stages of the disease.
+The second approach has gained more attention in recent years, due to the fact that in clinical practice it has shown promise in the early detection of AD.
+In addition, the analysis of the decline of cognitive abilities is comparitively inexpensive and less invasive that the first approach which commonly requires the collection of a sample of cerebro-spinal fluid which is painful for the patient involved.
 \par
@@
 \par
-Many researchers have studied the early detection of AD.
-These studies usually follow two main approaches.
-The analysis of biomarkers and the examination of patients who have demonstrated decreasing cognitive abilities.
-The first approach yields reliable results in the detection of AD in its moderate and advanced states but does not perform well during the early stages of the disease.
-The second approach has gained more attention in recent years, due to the fact that in clinical practice it has shown promise in the early detection of AD.
-In addition, the analysis of the decline of cognitive abilities is comparitively inexpensive and less invasive that the first approach which commonly requires the collection of a sample of cerebro-spinal fluid which is painful for the patient involved.
-/par
 The two main ways in which diagnosis is performed is through assessment of memory and language.
@@
 However, there is enough promise in this approach such that it could help further our understanding of the disease, it's progression and the parts of the brain affected in the early stages.
-\par  
-The consensus is that this language degeneration is typically accelerated by the presence of dementia \cite{Berisha2015} and that a potential indicator of dementia is the rate of change in which the decline occurs relative to a fixed point in time rather than a comparison across a cohort of individuals.
 \par
@@
 As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are extremely variable and can differ from patient to patient as the disease progresses.
+The consensus among researchers that this language degeneration is typically accelerated by the presence of dementia \cite{Berisha2015} and that a potential indicator of dementia is the rate of change in which the decline occurs relative to a fixed point in time rather than a comparison across a cohort of individuals.
 \par
@@
 We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English.
-This resulted in 17 journal articles and/or conference papers which form this review.
 \begin{table}
@@
 This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their perfomance in terms of both diagnostic and prognostic utility.
-\newline
 \par 
-\subsubsection{RBANS - Repeated Battery for the Assessment of Neuropsychological Status}
+\subsubsection{Repeatable Battery for the Assessment of Neuropsychological Status}
+The Repeatable Battery for the Assessment of Neuropsychological Status (RBANS) was originally developed as an assessment tool for dementia, specfically looking at detecting a characterizing very mild dementia.
+The authors felt that there was a shortfall of appropriate measures that were sensitive enough to milder impairments as well as a number of other shortcomings of existing tests.
+They met a number of design goals for this new battery of tests that addressed these shortcomings.
+The RBANS consists of a number of sub-tests across five distinct domains.
+
+\begin{enumerate}
+	\item Immediate Memory 
+	\item Visuospatial / Constructional
+	\item Language
+	\item Attention
+	\item Delayed Memory 
+\end{enumerate}
+
+However, whilst the authors claim that the RBANS is adequately sensitive in person's with MCI, other research points out that the RBANS has poor sensitivity in detecting MCI.
+Another drawback of this battery is the lack of executive function measures and object naming tasks.
+Research has shown that the RBANS has good test-retest reliability and convergent validity, pa
+
 \subsubsection{Patient Health Questionnaire)}
+The Patient Health Questionnaire
 \subsubsection{WASI}
@@
 \subsubsection{Digit symbol coding}
-\subsubsection{Corsi Span}
+\subsubsection{Corsi Block Tapping Task}
+The original Corisi apparatus consists of a series of nine blocks arranged irregularly on a 9 x 11in board.
+The blocks are tapped by the adminstrator in a randomized sequence of increasing length.
+Immediately after each sequence is demonstrated, the task is to mirror the sequence back to the adminstrator with the sequences increasing in length until a mistake is made.
 \subsubsection{Verbal Fluency}
@@
 \subsubsection{Rey Delayed Recall}
-\subsubsection{Naming - Speeded}
+\subsubsection{Naming Tests}
+Word-finding difficulty is a common symptom of AD and these deficits usually occur during the early stages of the disease progression.
+As such , a test of a patients ability to find words (known as confrontation naming) is a common way to measure cognitive decline.
+One common way to do this is the Boston Naming Test (BNT) which comprises 60 items on a spectrum of very frequent to very infrequent.
+This has been reduced subsequently to two thirty item versions and four fifteen item versions, which correlate significantly with the original sixty item version and the benefit of the shorter versions of the test is that it facilitates testing of individuals with AD who may suffer from fatigue or limited attention span.
 
+Whilst these studies and tests provide evidence that there are word finding difficulties in those with MCI, on their own they do not provide sufficient ability to diagnose MCI or provide an prognosis of disease progression.
 \subsubsection{Wisconsin Card Sorting}
@@
 The sensitivity and specificity of the different test forms in correctly classifying patients were compared to each other using a previously established FCSRT-IR cut off score. /newline
-/par
+\subsubsection{Conclusion}
+I have looked at a number of different cognitive tests and a battery of tests that aim to have high diagnostic and prognostic capabilities in the MCI population.
+However, particularly with the RBANS, there lacks sufficient sensitivity in differentiating those with MCI from healthy elderly individuals such that this tool could be used with a level of confidence in the results.
+In regard to the tests, a common theme is a lack of studies and therefore evidence into the utility of these tests with this particular population.
+However if, as researcher, we aim to investigate this population further then a benchmark battery of tests which is sensitive enough in both diagnostic and/or prognostic utility should be a goal.
 
-\subsection {Types of Language Assessment}
+\subsection{Types of Language Assessment}
 One of the key debates when looking at how to analyse language is the type of task provided to elicit language production in participants.
@@
 
+\subsubsection{How can Machine Learning be applied to this problem}
+
 \subsection{The problems with attempting early diagnosis of dementia and the ethics of actively screening for it}

### 2019-01-05 Source/QualifyingReport/QualifyingReport.tex
edited QR

@@
 \subsubsection{Patient Health Questionnaire)}
-The Patient Health Questionnaire
+The Patient Health Questionnaire is a nine-item questionnaire Likert scale which is designed to measure a 
 \subsubsection{WASI}
 \subsubsection{Digit Span}
-\subsubsection{Rey Words}
-\subsubsection{Prospective Memory Recall}
-\subsubsection{Digit symbol coding}
+\subsubsection{Rey Auditory-Verbal learning test}
+
+\subsubsection{Digit Symbol Substitution Test}
+The digit symbol test involves a key consisting of the numbers 1-9 and a corresponding unique symbol.
+Below this key is a series of numbers from 1-9 in a randomized order and repeated multiple times.
+The participant is asked to fill in the corresponding symbol for each number.
+The task requires that the participant move between the key and the randomized sequence such that  they may retrieve the correct answer from the key, hold this in short-term memory and transcribe the key in the appropriate place.
+
 \subsubsection{Corsi Block Tapping Task}
-The original Corisi apparatus consists of a series of nine blocks arranged irregularly on a 9 x 11in board.
+The original Corsi apparatus consists of a series of nine blocks arranged irregularly on a 9 x 11in board.
 The blocks are tapped by the adminstrator in a randomized sequence of increasing length.
@@
 A review carried out by Henry et al, found that performance in both letter fluency and category fluency was impaired in those with AD vs controls, but found a larger effect for tests of semantic fluency.
-\subsubsection{Rey Delayed Recall}
 \subsubsection{Naming Tests}
@@
 \subsubsection{Wisconsin Card Sorting}
-\subsubsection{Complex Figure Task immediate and delayed recall}
-\subsubsection{Cinderella Story - Narrative Description}
-\subsubsection{Sentence Completion}
-\subsubsection{Non-word Repetition}
-\subsubsection{Grooved Peg Board}
+\subsubsection{Complex Figure Task}
+\subsubsection{Narrative Description Task}
+\subsubsection{Sentence Completion Task}
+\subsubsection{Serial Non-word Repetition}
+\subsubsection{Grooved Pegboard Test}
+The Grooved Pegboard Test (GPT) was originally designed to cover a variety of different psycho-motor functions including hand-eye coordination and motor speed.
+However, some studies have shown a correlation of performance in the GPT and measures of cognitive performance such as the Montreal Cognitive Assessment.
 \subsubsection{Visual Search}
-\subsubsection{Spelling}
+In research, there have been shown to be deficits in both AD and MCI populations.
+Whilst these deficits were not as apparent in MCI populations, there remained a difference that could be used to differentiate healthy controls from those with MCI.
+% Abnormal Visual Search in MCI and Alzheimer's Disease. Tales.A 2005
+
 \subsubsection{Continuous Picture Naming}
@@
 \subsubsection{How can Machine Learning be applied to this problem}
+One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
 

### 2019-01-07 Source/QualifyingReport/QualifyingReport.tex
edited QR
AM work on QR

@@
 
-\subsubsection{Patient Health Questionnaire)}
-The Patient Health Questionnaire is a nine-item questionnaire Likert scale which is designed to measure a 
-\subsubsection{WASI}
-\subsubsection{Digit Span}
+\subsubsection{Digit Span Test}
+The Digit Span Test is predominantly used to measure a person's working memory capacity, specifically the capacity used to store and recall numbers.
+A participant is presented with a series of numbers of fixed length and is asked to recall those numbers in normal or reverse order.
+The series of numbers gets progressively longer until such time as a participant fails three or more times out of eight presentations.
+\par
+% Muangpaisan - Digit span and verbal fluency tests in patients with MCI and normal subjects in Thai community.
+MCI patients had significantly lower digit span score, in both normal and reverse order versions of this test.
+Research has shown that Digits Backwards can, to some degree, predict a diagnosis of MCI.
+They found that Age, Gender and Education have an impact on the performance of the tests.
+\par
+% Emrani - Assessing Working Memory in Mild Cognitive Impairment with Serial Order Recall
+Found no differences between specifically amnestic MCI and controls, but could differentiate between Mixed MCI and the other groups with mixed MCI recalling fewer correct responses than other groups.
+Notably, there was an attenuated recency effect in those with mixed MCI.
+% Kessels - Assessment of Working-memory deficits in patients with MCI and AD using Wechsler's Working Memory Index which includes digit span.
+Kessels identifies that that there are working memory deficits in MCI patients and these worsen with AD patients.
+% Kessels - Assessment of verbal and visuospatial working memory in MCI and AD.
+Kessels identifies that both MCI and AD have impaired performance on all three conditions of the digit span test.
+No differences were found between forward and backward conditions in any of the groups.
+However, available tests may not detect subtle impairments.
 \subsubsection{Rey Auditory-Verbal learning test}
-
+Rey's Auditory Verbal Learning Test (RAVLT+) looks a wide range of neuropsychological processess including short-term auditory-verbal memory, retention of information.
+Participants are given a list of 15 unrelated words, repeated over five different trials and are asked to repeat.
+Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then again after 30 minutes.
+% Moradi - Rey's Auditory Verbal Learning Test scores can be predicted from whole brain MRI in AD.
+Several studies have shown that an impairment in RAVLT score reflect well the underlying pathology caused by AD.
+Thus making the RAVLT an effective early marker to detect AD in persons with memory complaints.
+Moradi investigated to what extent the RAVLT scores are predictable based on MRI data using machine learning approaches, as well as to find out what the most important brain regions are for the estimation of RAVLT scores.
+They found a highly significant cross validated correlation between the estimated and observed RAVLT immediate and RAVLT Percent Forgetting.
+Further, they found that the conversion of MCI subjects to AD in 3-years could be predicted based on either observed or estimated RAVLT scores with an accuract comparible to MRI-based biomarkers.
+% Schoenberg - Test performance and classification statistics for the Rey Auditory Verbal Learning Test in selected clinical samples.
+Study found that RAVLT to best distguish patients suspected of Alzheimer's disease from the psychiatric comparison group.
 \subsubsection{Digit Symbol Substitution Test}
-The digit symbol test involves a key consisting of the numbers 1-9 and a corresponding unique symbol.
+The Digit Symbol Substitution Test (DSST) involves a key consisting of the numbers 1-9 and a corresponding unique symbol.
 Below this key is a series of numbers from 1-9 in a randomized order and repeated multiple times.
@@
 The task requires that the participant move between the key and the randomized sequence such that  they may retrieve the correct answer from the key, hold this in short-term memory and transcribe the key in the appropriate place.
-
+% Rosano - Digit symbol substitution test and future clinical and subclinical disorders of cognition, mobility and mood in older adults.
+Among those with no disorder in cognition, mobility and mood, being in the lowest DSST quartile compared to the highest was associated with nearly twice the odds of developing one or more clinical or subclinical disorders.
+Associations were stronger for incident clinical disorders in cognition.
+Slower psychomotor speed may serve as a biomarker of risk of clinical disorders, mobility and mood.
+While in part attributable to vascular brain disease, other potentially modifiable contributors may be present.
+Further studying the causes of psychomoto slowing with ageing might provide novel insights into age-related brain disorders.
+% Pascoe - The symbol-digit modalities test in MCI: Evidence from Parkinson's Disease Patients.
+Pascoe compared patients with PD with Normal Cognition (PD-N) with those with PD and MCI (PD-MCI) and healthy participants.
+PD-MCI pariticipants achieved significantly lower scores than other groups in the DSST task.
 \subsubsection{Corsi Block Tapping Task}
@@
 Immediately after each sequence is demonstrated, the task is to mirror the sequence back to the adminstrator with the sequences increasing in length until a mistake is made.
+% Perrochon - How specific are cognitive deficits in short-term visuospatial memory in mild cognitive impairment patients?
+The CBT performances were significantly less satisfactory in the multi-domain MCI category than in other groups (Healthy adults, Amnesic MCI and Non-amnesic MCI with executive function impairment).  
 \subsubsection{Verbal Fluency}
@@
 \subsubsection{Wisconsin Card Sorting}
+
 \subsubsection{Complex Figure Task}
+
 \subsubsection{Narrative Description Task}
+
 \subsubsection{Sentence Completion Task}
+
 \subsubsection{Serial Non-word Repetition}
+
 \subsubsection{Grooved Pegboard Test}

### 2019-01-07 Source/QualifyingReport/QualifyingReport.tex
QR PM work

@@
 This has been reduced subsequently to two thirty item versions and four fifteen item versions, which correlate significantly with the original sixty item version and the benefit of the shorter versions of the test is that it facilitates testing of individuals with AD who may suffer from fatigue or limited attention span.
+% Willers - Subclinical naming errors in MCI - A semantic deficit.
+Twenty aMCI patients, twenty AD and 21 controls matched by age, sex and education level were evaluated.
+AD patients obtained significantly lower total scores on the BNT than aMCI patients and controls.
+aMCI patients did not obtain significant differences in total scores but showed significantly higher semantic errors compared to controls.
+Semantic processing is impaired during confrontation naming in aMCI.
+% Vadikolias - Mild Cognitive Impairment: effect of education on the verbal and nonverbal tasks performance decline.
+Higher educational attainment in aMCI subjects were correlated with better performance in verbal and non-verbal tasks during repeated examinations over 1-year.
+Subjects with a lowere level of education performed worse than patients with a high level of education who presented a more stable clinical score.
+The explanation for this is the idea of a 'cognitive reserve' in participants with a higher education such that this provides a buffer that, while not preventing the physiological symptoms of AD, can potentially delay the clinical onset of cognitive symptoms that characterise AD.
+This theory is supported by Snowdon who found a relationship between early life linguistic ability and the density of neurofibrillary tangles in his nun study \cite{Snowdon1996}.
 
 Whilst these studies and tests provide evidence that there are word finding difficulties in those with MCI, on their own they do not provide sufficient ability to diagnose MCI or provide an prognosis of disease progression.
-\subsubsection{Wisconsin Card Sorting}
 
-\subsubsection{Complex Figure Task}
+\subsubsection{Rey-Osterrieth Complex Figure Task}
+% Cite Rey 1941 - L'examen psychologique dans les cas d'encephalopathie traumatique and Osterrieth 1944 - Le test de copie d'une figure complexe.
+The Rey-Osterrieth Complex Figure (ROCF) is a task widely as a test of visuo-spatial skill and visual memory.
+The task, which was originally designed by Rey (1944) and standardised by Osterrieth (1944), asks a participant to copy a complex geometrical figure (known as the immediate copy condition and to recall and reproduce the figure from memory without warning (known as the delayed recall condition).
+In the immediate copy condition, the complexity of the figure requires an integrative cognitive ability.
+The reproduction of such a complex structure involves processes such as planning and organizational strategies that are related to executive functions.
+% Salvadori - Qualitative Evaluation of the Immediate Copy of the Rey-Osterrieth Complex Figure: Comparison Between Vascular and Degenerative MCI patients.
+Patients with vascular MCI had a worse performance in the immediate copy of the ROCF compared to individuals with degenerative MCI, despite their significant impairment in terms of general cognitive status and visual memory.
+Evidence shows that patients with disorders that possibly involve attention and executive functions are characterized by a more disorganised approach when copying the ROCF compared to controls.
+% Teng - Persistence of Neuropsychological Testing Deficits in Mild Cognitive Impairment.
 
-\subsubsection{Narrative Description Task}
+One of the difficulties with the presentation of this task in a repeated battery will be the fact that it turns from an incidental memory task (it is a memory task but the participant is not forewarned that they will be required to memorise the picture) into an intentional memory task (given the previous exposure to the task, it would be expected that a participant would pre-empt the delayed recall portion of the test and spend more time attempting to memorise the details).
 
-\subsubsection{Sentence Completion Task}
+\subsubsection{Hayling Sentence Completion Test}
+Inhibitory deficits are a common in all stages of dementia.
+This is usually tested using Stroop test, however this has the drawback of lacking ecological validity.
+Therefore researchers have moved towards using the Hayling Sentence Completion Test (HCST) which uses skills such as word retreval as well as the ability to inhibit ones responses where appropriate and is therefore a much more ecologically valid task.
+There are two parts to the HCST.
+In the first part, participants have to complete a sentence by providing a word that best fits the given sentence (this is known as the initiation condition).
+In this second part, participants have to complete sentences by inhibiting an impulse to give the word that best first the sentence as in the first part, and producing a semantically unconnected word.
+Performance in both conditions is measured by the time taken for the participant to initiate a response, and in the inhibition condition also by the correctness of the word.
+\par 
+% Martyr - Assessing inhibitory control in early-stage Alzheimer's and Parkinson's disease using the Hayling Sentence Completion Test.
+Martyr et al compared healthy controls with patients with dementia and patients with parkinson's disease.
+They found that a high proportion of Category A errors (producing a word that fits the sentence when instructed otherwise) was a factor in performance loss for participants with dementia.
+Findings suggest that the HSCT may be sensitive to verbal suppression deficits and may provide insight into inhibitory control in participants with dementia.
+Patients with Dementia were significantly slower than controls in the initiation and inhibition conditions vs healthy controls, and slower than patients with parkinsons disease in the initiation but not the inhibition condition.
 
@@
 However if, as researcher, we aim to investigate this population further then a benchmark battery of tests which is sensitive enough in both diagnostic and/or prognostic utility should be a goal.
+\par 
+A criticism of the current literature is the lack of consistency with regard to the experimental groups.
+For example, some studies focus on differentiating between an MCI group, a AD group and a healthy controls group whereas other studies may further subdivide the MCI category according to a number of factors such as Amnesic or Non-Amnesic MCI, Mixed MCI (sometimes called Multi-domain MCI in the literature).
+Given the inconsistency in defining the experimental groups the confusing and often conflicting results that researchers produce is to be expected.
+Future research should look at the standardization of the operational definition of cognitive impairment in MCI may result in more consistent predictions of progression to AD.
 
@@
 This area of research is extremely promising in its early results and the impact of successful research would be life changing for both individuals and the health of the worlds aging population in general.
-\newline
+\par 
+As described above, there is a large gap in the knowledge of how, and it what circumstances does MCI progress to AD as well as how best to assess MCI over time and at this time, to the authors knowledge, there is no consensus in how to approach these two problems.
+Teng et al suggests that work should focus on the MCI population and concentrate on developing a consensus neuropsychological battery that could yield predictable rates of progression to AD.
+This, and with the development of a model of language, sensitive enough to detect subtle deterioration in language use to act as an additional cognitive marker to aid diganosis could potentially move some way to providing this.
+\par 
+% McGuiness - Predicting conversion to dementia in a memory clinic: A standard clinical approach compared with an empirically defined clustering method (latent profile analysis) for mild cognitive impairment subtyping.
 
 \section{Project Proposal}
-\textbf{Full Title of the Research:} Title here \newline
+\textbf{Full Provisional Title of the Research:} Title here \newline
 \textbf{Research Questions and Hypotheses}

### 2019-01-08 Source/QualifyingReport/QualifyingReport.tex
work done overnight on QR

@@
 \usepackage{float}
-\usepackage{parskip}
+%%\usepackage{parskip}
+\usepackage{enumitem}
+\usepackage{fancyhdr}
 %% Page Settings
@@
 %% Own Commands
+\pagestyle{fancy}
+\fancyhf{}
+\fancyhead[CE,CO]{Early Detection of MCI - An Interdisciplinary Approach}
+\fancyfoot[CE,CO]{\leftmark}
+\fancyfoot[LE,RO]{\thepage}
+
+\renewcommand{\headrulewidth}{2pt}
+\renewcommand{\footrulewidth}{1pt}
 \input{settings/macros}
 \title{Qualifying Report}
-\date{2018-12-30}
+\date{2019-01-07}
 \author{Jomar Alcantara}
@@
 \newpage
+\begin{abstract}
+
+\end{abstract}
 \tableofcontents
+\newpage
+\listoftables
+\newpage
+\listoffigures
 \newpage
@@
 Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
+\subsection{What is this subsection called?}
+\textbf{Full Provisional Title of the Research:} Title here \newline
+\textbf{Research Questions and Hypotheses}
+The broad question is can we use Machine Learning to assist in the early diagnosis of dementia and/or mild cognitive impairment through language produced naturally rather than language that is produced as part of a formal cognitive assessment.
+There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia.
+These studies have identified a number of speech and language differences between healthy controls and those with dementia.
+Some of these features have been validated in work using machine learning to classify healthy controls from those with dementia.
+However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'.
+While this is help to learn that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments.
+\newline
+\par
+Machine Learning itself is undergoing a shift in focus at the present time.
+I will describe this shift using the context of my proposed research.
+In traditional machine learning, the transcripts from cognitive assessments are used to generate features such as idea density.
+These features are then used to classify patients.
+A company in Toronto, Canada has taken this approach and currently uses over 400 acoustic and language features to make this classification and to generate a psuedo MMSE score.
+\newline
+\par
+Deep Learning takes a different approach by bypassing the feature generating stage. Deep learning uses neural networks as a fundamental building block.
+
+\subsection{Aims and Rationale}
+The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to to time-consuming and expensive psychological tests.
+There is a lot of research into the analysis of language as a bio-marker for MCI and Early Dementia.
+Given that sampling a person's language is relatively effortless, my research looks at whether we can find bio-markers of MCI and Early Dementia in natural language.\newline
+Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
+I'm not going to control for memory problems as a potential confound, but does this weaken the research?
+How do I defend this?
+\newline
+\begin{itemize}
+	\item Hypothesis.
+I predict that, given a number of samples of language over a given time from, is it possible to detect language decline in humans such that you can differentiate between normal language decline and decline in those with MCI.
+	\item Hypothesis. I predict that my new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures.
+	\item My first hypothesis is whether monitoring language samples and other measures of cognitive decline over the course of the year is enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment.
+	\item My second hypothesis is that those with MCI show a different language profile to controls, if matched similarly for education and other significant factors.
+\end{itemize}
+
+\subsection{Research Challenges}
 \section{Literature Review}
@@
 The measure was based on Goodglass and Kaplan's scale for rating discourse on the Boston Diagnostic Aphasia Examination (BDAE) and is composed of clinical judgement of behaviour that is rated on a likert-type scale ranging fdrom 1 to 7 (7 being no abnormality).
+\par
 
@@
 \section{Project Proposal}
-\textbf{Full Provisional Title of the Research:} Title here \newline
-\textbf{Research Questions and Hypotheses}
-The broad question is can we use Machine Learning to assist in the early diagnosis of dementia and/or mild cognitive impairment through language produced naturally rather than language that is produced as part of a formal cognitive assessment.
-There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia.
-These studies have identified a number of speech and language differences between healthy controls and those with dementia.
-Some of these features have been validated in work using machine learning to classify healthy controls from those with dementia.
-However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'.
-While this is help to learn that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments.
-\newline
-\par
-Machine Learning itself is undergoing a shift in focus at the present time.
-I will describe this shift using the context of my proposed research.
-In traditional machine learning, the transcripts from cognitive assessments are used to generate features such as idea density.
-These features are then used to classify patients.
-A company in Toronto, Canada has taken this approach and currently uses over 400 acoustic and language features to make this classification and to generate a psuedo MMSE score.
-\newline
-\par
-Deep Learning takes a different approach by bypassing the feature generating stage. Deep learning uses neural networks as a fundamental building block.
 
-\subsection{Aims and Rationale}
-The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to to time-consuming and expensive psychological tests.
-There is a lot of research into the analysis of language as a bio-marker for MCI and Early Dementia.
-Given that sampling a person's language is relatively effortless, my research looks at whether we can find bio-markers of MCI and Early Dementia in natural language.\newline
-Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
-I'm not going to control for memory problems as a potential confound, but does this weaken the research?
-How do I defend this?
-\newline
-\begin{itemize}
-	\item Hypothesis.
-I predict that, given a number of samples of language over a given time from, is it possible to detect language decline in humans such that you can differentiate between normal language decline and decline in those with MCI.
-	\item Hypothesis. I predict that my new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures.
-	\item My first hypothesis is whether monitoring language samples and other measures of cognitive decline over the course of the year is enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment.
-	\item My second hypothesis is that those with MCI show a different language profile to controls, if matched similarly for education and other significant factors.
-\end{itemize}	
-\section{Experiment 1 - Pilot Study of Neuropsychological Tests}
-\subsection{Methods}
+\subsection{Experiment 1 - Pilot Study of Neuropsychological Tests}
+\subsubsection{Methods}
 I recruited a cohort of 9 volunteers which which to complete the battery of Neuropsychological tests.
@@
 \subsubsection{Participants}
-\subsection{Results}
-\subsection{Discussion}
-\subsection{Conclusion}
-\section{Experiment 2 - Initial Study of the Presidents corpus}
-\subsection{Background}
-\subsection{Methods}
+\subsubsection{Results}
+\subsubsection{Discussion}
+\subsubsection{Conclusion}
+\subsection{Experiment 2 - Initial Study of the Presidents corpus}
+\subsubsection{Background}
+\subsubsection{Methods}
 I took 46 transcripts of Ronald Reagan’s (RR) press conferences from 1981 to 1988 and compared them with 134 press conferences by George H.
@@
 Bush where there was a question and answer session conducted at least in part by the sitting president (2 press conferences of GWB were ommited due to a lack of a question and answer session).
-\subsection{Results}
-\subsection{Discussion}
-\subsection{Conclusion}
-
-\subsection{Timeline of Future Work and Publication plan}
-\textbf{Year One} \newline
-\textbf{Year Two} \newline
-\textbf{Year Three} \newline
-
-\subsection{Provisional Table of Contents}
-\subsubsection{Chapter 1: Introduction}
-Introduction to the problem of dementia in the context of the wider world including quality of life and financial implications.
-Exploration of dementia as a syndrome rather than a disease, and a look at the different variants of dementia.
-A look at the rationale behind research into the early diagnosis of dementia as well as a brief look at what has been done in the area (wide context, so pharmacological and psychological).
-\newline
-\subsubsection{Chapter 2: Background}
-Literature Review which looks at the background of what has been done in the area of early diagnosis of dementia, from a purely psychological perspective as well as the work done in machine learning in this topic.
-Look at the different attributes that researchers have explored.
-Look at the difference between traditional machine learning methods and deep learning and discuss the pros and cons of each perspective.
-\newline
+\subsubsection{Results}
+\subsubsection{Discussion}
+\subsubsection{Conclusion}
 \subsection{Chapter 3: Experiments with existing datasets: Presidential Press Conferences, Three Authors - Murdoch, Christie and James, DementiaBank dataset.}
@@
 
-\subsection{Chapter 5: Longditunal Study }
+\subsection{Chapter 5: Longditunal Study}
 \subsubsection{Dataset creation}
@@
 \newline
-\subsubsection{What happens at the end of the data collection phase if there is a suspected decline in a participant's cognitive function}
-\subsection{How does language differ between those with AD, MCI and Controls}
-\subsection{Chapter 7: Overall results and Discussion}
-\subsection{Chapter 8: Conclusions and Future work}
+
+\section{Future Work}
+\subsection{Introduction}
+
+\subsection{Project Plan}
+
+\subsection{Risk Assessment}
+
+\subsection{Ethical Considerations}
+
+\subsection{Thesis - Provisional Table of Contents}
+In this section I provide a provisional table of contents based on the work I have either completed or proposed above.
+\begin{enumerate}
+	\item Introduction
+	\begin{enumerate}
+		\item Overview
+		\item Background and Context
+		\item Aims and Objectives
+		\item Challenges and Achievements
+	\end{enumerate}
+	\item Literature Review
+	\begin{enumerate}
+		\item Introduction
+		\item Assessing Mild Cognitive Impairment using cognitive tests.
+		\item Assessing Mild Cognitive Impairment using Language.
+		\item Natural Language Processing and Machine Learning
+		\item Conclusion
+	\end{enumerate}
+	\item Further Exploration of the Presidents Corpus
+	\begin{enumerate}
+		\item Introduction
+		\item Methodology
+		\item Results
+		\item Discussion
+		\item Conclusion
+	\end{enumerate}
+	\item Validation of a Repeatable Neuropsychological Battery of Tests
+	\begin{enumerate}
+		\item Introduction
+		\item Methodology
+		\item Pilot Study Results
+		\item Longitundinal Study results
+		\item Discussion
+		\item Conclusion
+	\end{enumerate}
+	\item Longitudnal Study
+	\begin{enumerate}
+		\item Introduction
+		\item Methodology
+		\item Experiment 1 Results
+		\item Experiment 2 Results
+		\item Discussion
+		\item Conclusion
+	\end{enumerate}
+	\item General Discussion
+	\begin{enumerate}
+		\item Discussion
+		\item Future Work
+		\item Conclusion
+	\end{enumerate}
+\end{enumerate}
+
+\section{Conclusion}
 

### 2019-01-08 Source/QualifyingReport/QualifyingReport.tex
morning work on QR 08Jan

@@
 \usepackage{float}
-%%\usepackage{parskip}
 \usepackage{enumitem}
@@
 Patients with Dementia were significantly slower than controls in the initiation and inhibition conditions vs healthy controls, and slower than patients with parkinsons disease in the initiation but not the inhibition condition.
-
-\subsubsection{Serial Non-word Repetition}
-
 \subsubsection{Grooved Pegboard Test}
@@
 However, some studies have shown a correlation of performance in the GPT and measures of cognitive performance such as the Montreal Cognitive Assessment.
+% Bezdicek - Grooved Pegboard Predicates More of Cognitive Than Motor Involvement in PD.
+Bezdicek found that the GPT predicted performance on the MoCA and concluded that in addition to being a measure of motor skills, there were results that showed that the GPT could also provide information about a participants cognitive skills.
+Whilst this was specifically using a population of patients with Parkinson's disease, given the MoCA and GPT have shown degree of correlation, there remains an opportunity for research that explores the use of the GPT with MCI and Early AD populations.
+% Darweesh - Simple Test of Manual Dexterity can identify persons at high risk for neurodegenerative diseases in the community.
+Darweesh et al, used the Purdue Pegboard Test to assess manual dexterity in a healthy older adult population and followed their participants for between eight and twelve years until an indication of the onset of a neurodegenerative disease.
+In this time 227 (4\%) of their participants went on to develop a diagnosis of dementia.
+They noted that higher PPT scores were associated with lower risk of incident neurodegnerative disease and noted significant associations of PPT scores with all forms of dementia and this potentially highlights a deterioration of motor function in the pre-clinical  phase of dementia.
 \subsubsection{Visual Search}
+The visual search task requires participants to be watching a computerised display in which a target appears either by itself or surrounded by other elements which are used as a distraction.
+These other elements can be vary in similarity to the target.
+This is inherently a measure of attention shifting and processing of various similarity and as one would expect, the visual search time is increased in the presence of distractors.
+This effect is magnified in participants who have a cognitive impairment.
+% Abnormal Visual Search in MCI and Alzheimer's Disease. Tales.A 2005
 In research, there have been shown to be deficits in both AD and MCI populations.
 Whilst these deficits were not as apparent in MCI populations, there remained a difference that could be used to differentiate healthy controls from those with MCI.
-% Abnormal Visual Search in MCI and Alzheimer's Disease. Tales.A 2005
-
-\subsubsection{Continuous Picture Naming}
+% Tales - Visual Search in MCI: A Longitudinal Study
+In a comparison of visual search performance between those with aMCI and healthy controls, Tales noted significantly poorer performance in the aMCI group.
+However, she also noted a good deal of heterogeneity in the aMCI group which illustrates that whilst the aMCI group have essentially the same condition, presentations within this population can differ markedly.
+The results from this trial also illustrates the use of a non-memory task as a means of diagnosing dementia as patients who went on to be diagnosed with dementia whilst this study was in progress had significantly poorer visual search performance.
 \subsubsection{Free Cued Selective Reminding Test}
 The Free Cued Selective Reminding Test (FCSRT) was borne out of the premise that by controlling the conditions of learning, a measure of memory is possible that is not confounded by normal age related changes in cognition.
-Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cogntive decline due to a cognitive impairment.
-/newline
+Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cognitive decline due to a cognitive impairment.
 /par
-FCSRT-IR (The instant recall version of the Free Cued Selective Reminding Test) performance has beeen shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the authors to define prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT-IR performance (Sarazin et al, 2007 /newline
+FCSRT-IR (The instant recall version of the Free Cued Selective Reminding Test) performance has beeen shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the authors to define prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT-IR performance (Sarazin et al, 2007
 /par
-The sensitivity and specificity of the different test forms in correctly classifying patients were compared to each other using a previously established FCSRT-IR cut off score. /newline
+The sensitivity and specificity of the different test forms in correctly classifying patients were compared to each other using a previously established FCSRT-IR cut off score. 
 \subsubsection{Conclusion}
@@
 In the literature researchers have primarily focused on Picture Description tasks but have also suggested other ways in which we might collect data.
-\newline
 \par
@@
 The picture itself depicts a familiar domestic scene and would not require participants to use any vocabulary beyond that learned in childhood.
-It was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}\newline
+It was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}
 \begin{figure}[H]
 \centering
-\includegraphics[width=240px, height=150px]{BCTPicture.png}
+\includegraphics[width=240px, height=150px]{images\BCTPicture.png}
 \caption{Cookie Theft Picture.\label{white}}
@@
 In explaining the differences, it is worth noting that these researchers were using differing variables and this could explain their different perspectives.
-\newline
 \par
-In terms of Machine Learning research in this area, a number of researchers have used transcripts based on picture description tasks \cite{Zimmerer2016, Orimaye2017, Mueller2018a, Fraser2015} and have successfully extracted linguistic features that could differentiate between AD and controls.\newline
+In terms of Machine Learning research in this area, a number of researchers have used transcripts based on picture description tasks \cite{Zimmerer2016, Orimaye2017, Mueller2018a, Fraser2015} and have successfully extracted linguistic features that could differentiate between AD and controls.
 \par
@@
 As noted above, Ash \cite{Ash2012} felt that this task was interchangable with the Picture Description task and because this task requires elaboration rather than simple description, is a sturdier test of lexical and semantic abilities as well as syntactic complexity.
-\cite{DeLira2011} \newline
+\cite{DeLira2011}
 \par
@@
 However, this could be an interesting direction to take research in the future to see if features generated from this task could be used to predict MCI or AD.
-\newline
 \par
@@
 The semi structured nature of these interviews means that there is some replicability but does not constrain the participant in answering questions.
-\newline
 \par
@@
 Nevertheless, this is the most naturalistic setting for looking at language production and can be used to look at the syntactic and semantic parts of language generation \cite{Sajjadi2012}.
-There have been some attempts to use interviews to assess language production in AD with promising results \cite{Asgari2017, Guinn2015} .\newline
+There have been some attempts to use interviews to assess language production in AD with promising results \cite{Asgari2017, Guinn2015}.
 \par
@@
 There are other less common methods of analysing language but these are beyond the scope of this review.
-\newline
 \par
@@
 However, this approach does not take into account the symbiotic nature of language processes.
-\newline
 \par
@@
 They are able to be relatively constrained in the language available to be used, for example in the picture description task, or they can be unconstrained in the form of interviews.
-\newline
 \par
@@
 Pragmatic meaning refers looks at the same sentence in terms of words and grammar but takes into account the situation or context in which these words are spoken.
-Emery in her literature review looks at all levels of language tasks except for pragmatics, however   \newline
+Emery in her literature review looks at all levels of language tasks except for pragmatics, however ---FINISH THIS!
 \par
@@
 This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2005}.
-\newline
 \par
@@
 He found that patients with AD expressed significantly fewer content elements than controls.
-\newline
 \par
@@
 This study was replicated by Vuorinen et al (2001).
-\newline
 \par
@@
 The authors found no difference in the number of semantic elements produced but did not that the group with moderate to moderately severe AD took more time and more syllables to communicate these elements.
-\newline
 \par
@@
 \centering
-\includegraphics[width=240px, height=150px]{picnic-scene.jpg}
+\includegraphics[width=240px, height=150px]{\images\picnic-scene.jpg}
 \caption{Picnic Scene taken from the Western Aphasia Battery (WAB).\label{white}}
@@
 They identified 12 elements of the cookie theft picture and found that whilst each AD group differed significantly from the others and from controls, the measures did not distinguish between MCI and normal controls.
-\newline
 \par
@@
 Riley et al extended these findings by concluding that early-life idea density was associated with lower brain weight, higher degree of cerebral atrophy and increased neurofibrillary pathology in later life.
-\newline
 \par
@@
 In their study they found that AD patients produced more of these behaviours than did controls.
-\newline
 \par
@@
 The authors suggested that discourse information measures may help disentangle the similarities in symptoms of early AD versus depression in older adults.
-Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in dialogue in which a speaker provides responsees do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls..\newline
+Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in dialogue in which a speaker provides responsees do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls.
 \subsubsection{Lexical richness and diversity}
@@
 In contrast, Lira 2014 found that controls produced more total words than patients with AD but found no difference between mild and moderate groups.
-\newline
 \par
@@
 The general hypothesis motivating these studies is that either working memory limitations or semantic memory limitations in AD affect one's ability to use complex constructions.
-\newline
 \par
@@
 Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
-\newline
 \par
@@
 They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
-\newline
 \par
@@
 Murray found that MLU was not a distinguishing factor among health adults, adults with depression and adults with AD.
-Ripich et al found a decrease in MLU in adults with severe AD over time, and this was supported by findings of Le et al in their studies of authors \cite{Le2012} \newline
+Ripich et al found a decrease in MLU in adults with severe AD over time, and this was supported by findings of Le et al in their studies of authors \cite{Le2012}
 \par
 \subsubsection{Proportion of verbs to nouns plus verbs}
-Kave and Levy used a verb index to capture syntactic complexity and found that adults with AD expressed the same amount of verbs to nouns plus verbs as adult controls. \newline
+Kave and Levy used a verb index to capture syntactic complexity and found that adults with AD expressed the same amount of verbs to nouns plus verbs as adult controls.
 \par
@@
 The differences in syntactic complexity were not significant when individual measures were tested, but were apparent using a composite score consisting of MLU, words in sentences, syntactic errors, nouns with determiners, and verbs with inflections.
-\newline
 \par
@@
 He noticed a significant difference in the use of formulaic sentences between AD and Healthy Controls.
-\newline
 \par
@@
 Secondly, due to the constrained nature of the picture description task, it is unlikely that it would be able to capture the pragmatic skills that are typioical of conversations in everyday life.
-\newline
 \par
@@
 The authors attributed AD patients' difficulties to memory deficits, attentional deficits, visual perceptual problems, disruption of internalized frame representation, or failure to access frame knowledge.
-\newline
 \par
 \subsubsection{Perseveration}
+% EXPLAIN PERSEVERATIONS
 One study examined verbal preservation in the description of Norman Rockwell prints, dividing the total number of words within perseverations by total number of words in the speech sample.
@@
 The authors theorised that this was because picture description was an easier taks because there was the visual stimulus, similar to the argument made by Bschor et al.
-\newline
 \par
@@
 stuttering), abnormailities in the fluency of adults with MCI or AD are rarely manifested at the phonological level.
-\newline
 \par
@@
 However, there are some studies which contradict these findings (Ahmed et al).
-\newline
 \par
@@
 The authors did not report correlations between error monitoring and executive function test scores, however, which could have strengthened that hypothesis.
-\newline
 \par
@@
 \par
+\subsection{Conclusions}
 
@@
 
-\subsubsection{How can Machine Learning be applied to this problem}
+\subsection{Discussion}
 One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
-
-\subsection{The problems with attempting early diagnosis of dementia and the ethics of actively screening for it}
 It is challenging to characterise Dementia as a disease as it's more syndrome.
@@
 Also, the individual who may develop symptoms that indicate the onset of dementia needs the insight to recognise them, or family members will need to do so and also deduce that there is value to admitting there is a problem and seeking help before the deterioration is too great.
-\newline
 \par
@@
 Therefore, it's difficult to say with certainty that any symptoms can definitely cause dementia despite the fact that these symptoms are commonly present in those that have dementia.
-\newline
 \par
@@
 The fact that there are a numerous array of different perspectives to tackling this question, and even within these perspectives there are a range of various methodologies leaves the field with a sense of confusion as to the most effective way to solving the problem.
-\newline
 \par
@@
 Implementation of effective screening instruments will allow diagnosis earlier in the course of dementia, even at the point when memory function is still essentially within the normal range.
-This strategy would enable an earlier, and potentially more effective, prevention and treatment of AD with a special focus to preserve cognitive functions \newline
+This strategy would enable an earlier, and potentially more effective, prevention and treatment of AD with a special focus to preserve cognitive functions.
 \par
@@
 A significant amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
-\newline
 \par
@@
 Further, despite the quality of datasets being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which we could claim are limited in scope.
-\newline
 \par
@@
 Another key benefit of deep learning is it's ability to automate the process of feature engineering.
-Feature engineering, with this particular problem is an interesting debate as there are currently numerous ways in which to try to generate features as I have described above.\newline
+Feature engineering, with this particular problem is an interesting debate as there are currently numerous ways in which to try to generate features as I have described above.
 \par
@@
 
-\section{Project Proposal}
+\section{Initial Study of the Presidents corpus}
 
-\subsection{Experiment 1 - Pilot Study of Neuropsychological Tests}
-\subsubsection{Methods}
-I recruited a cohort of 9 volunteers which which to complete the battery of Neuropsychological tests.
-We compiled a series of tests (see literature review for a comprehensive list, explanation and rationale for each test) and completed each version of the test on three consequetive days.
-The results of which were compiled and analysed.
-The hypothesis is that there would be no difference in performance between one battery of tests and another.
-An additional hypothesis is that the language tasks included on each task would be enough to collect an appropriate sample for linguistic analysis, such that we may be able to detect cognitive decline in a clinical population.
-\newline
-\par
-\subsubsection{Participants}
-\subsubsection{Results}
-\subsubsection{Discussion}
-\subsubsection{Conclusion}
-\subsection{Experiment 2 - Initial Study of the Presidents corpus}
-\subsubsection{Background}
-\subsubsection{Methods}
+\subsection{Background}
+\subsection{Methods}
 I took 46 transcripts of Ronald Reagan’s (RR) press conferences from 1981 to 1988 and compared them with 134 press conferences by George H.
@@
 \subsubsection{Conclusion}
+
+\section{Future Work}
+\subsection{Introduction}
+
+\subsection{Project Plan}
+I recruited a cohort of 9 volunteers which which to complete the battery of Neuropsychological tests.
+We compiled a series of tests (see literature review for a comprehensive list, explanation and rationale for each test) and completed each version of the test on three consequetive days.
+The results of which were compiled and analysed.
+The hypothesis is that there would be no difference in performance between one battery of tests and another.
+An additional hypothesis is that the language tasks included on each task would be enough to collect an appropriate sample for linguistic analysis, such that we may be able to detect cognitive decline in a clinical population.
+\newline
+
 \subsection{Chapter 3: Experiments with existing datasets: Presidential Press Conferences, Three Authors - Murdoch, Christie and James, DementiaBank dataset.}
@@
 Paper: To replicate and extend the work of Orimaye et al. To compare traditional machine learning techniques (already completed by Orimaye) and deep learning. \newline
-\subsection{Chapter 4: Validation of new repeatable neuropsychological battery for detecting MCI.}
-Aim: \newline
-Dataset: To be created.
 
@@
 Paper:
-\subsubsection{Ethics}
+
+\subsection{Risk Assessment}
+
+\subsection{Ethical Considerations}
 Need to discuss confidentiality right at the start of each conversation and to reiterate at every meeting.
@@
 \newline
-
-\section{Future Work}
-\subsection{Introduction}
-
-\subsection{Project Plan}
-
-\subsection{Risk Assessment}
-
-\subsection{Ethical Considerations}
 

### 2019-01-08 Source/QualifyingReport/QualifyingReport.tex
moved images to separate folder
PM work on QR including updating the bibliography

@@
 % HEADER BEGINS
-\documentclass{article}
+\documentclass[12pt, twoside, a4paper]{article}
 %% Packages
@@
 \fancyhf{}
-\fancyhead[CE,CO]{Early Detection of MCI - An Interdisciplinary Approach}
-\fancyfoot[CE,CO]{\leftmark}
+\fancyhead[CE,CO]{\leftmark}
+\fancyfoot[CE,CO]{Early Detection of MCI that progresses to dementia - An Interdisciplinary Approach}
 \fancyfoot[LE,RO]{\thepage}
@@
 \par
-In 2009, the Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities.
-A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cogntive Impairment (MCI) with research focused two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease.
-As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
-At the present time diagnosis is typically conducted at memory clinics by trained clinicians.
-I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using natural language.
+In 2009, the Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{Health2009}.
+A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cogntive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease\cite{Taler2008}.
+As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages \cite{Prince2015} and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
+At the present time diagnosis is typically conducted at memory clinics by trained clinicians\cite{Boschi2017}.
+I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
 \par
-There is a large body of research that looks at the decline in language in those with MCI and AD.
-However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD.
+There is a large body of research that looks at the decline in language in those with MCI and AD\cite{Taler2008, Boschi2017}.
+However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD\cite{Taler2008, Boschi2017}.
 Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
@@
 Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
-\subsection{What is this subsection called?}
+\par
+This report is structured in the following way.
+The context of the project, the aims and hypotheses of the research and the challenges the research presents are detailed in the remainder of this chapter.
+A literature review of current methods of cognitive assessment of MCI, an review of how language is said to be impacted in the MCI and Early AD population and a review of NLP and ML techniques that have the potential to be applied to this domain is presented in Chapter 2.
+The findings from a replication and extension of previous research using an existing dataset is presented in Chapter 3.
+I detail my future work plan including distinct tasks to be completed over the next two years in Chapter 4.
+Finally, my conclusions are summarised in Chapter 5.
+\subsection{Background}
 \textbf{Full Provisional Title of the Research:} Title here \newline
@@
 There are a number of studies, that have examined "melodic line", which is a subjective measure of speech prosody defined as "the appropriate use of intonational contour, including alterations in pitch, volume and duration".
-For instance, Forbes-McKay compared melocdic line using a "simple" picuture such as the Cookie Theft picture vs a complex picture.
+For instance, Forbes-McKay compared melodic line using a "simple" picuture such as the Cookie Theft picture vs a complex picture.
 The number of pictorial themes differentiated the simple tasks from the complex.
 Results showed no group differences on simple picture tasks (Cookie Theft and Tripping Woman) but there were differences in melodic line for the complex picture tasks.
-Fraser et al examined several acoustic features of speech in patients with AD ussing machine learning methods that captureed both rate and phonation patterns, and found acoustic abnormalities to be a significant factor.
+Fraser et al examined several acoustic features of speech in patients with AD ussing machine learning methods that captured both rate and phonation patterns, and found acoustic abnormalities to be a significant factor.
 Konig et al, also used an automated feature analysis examining vocal and temporal aspects of speech among controls, and patients with MCI or AD, and reported a classification of 81\%.
@@
 In contrast with fluency disorders (i.e.
-stuttering), abnormailities in the fluency of adults with MCI or AD are rarely manifested at the phonological level.
+stuttering), abnormalities in the fluency of adults with MCI or AD are rarely manifested at the phonological level.
 \par
@@
 \section{Initial Study of the Presidents corpus}
+\subsection{Background}
 
-\subsection{Background}
 \subsection{Methods}
@@
 Bush where there was a question and answer session conducted at least in part by the sitting president (2 press conferences of GWB were ommited due to a lack of a question and answer session).
-\subsubsection{Results}
-\subsubsection{Discussion}
-\subsubsection{Conclusion}
+\subsection{Results}
+\subsection{Discussion}
+\subsection{Conclusion}
 
@@
 
-\subsection{Project Plan}
+\subsection{Additional Research Aims and Objectives}
+
+\subsection{Additional Research Challenges}
+
+\subsection{Project Plan and Task Definition}
+\subsubsection{Task 2 - Pilot Study for the Neuropsychological Battery of Tests}
 I recruited a cohort of 9 volunteers which which to complete the battery of Neuropsychological tests.
@@
 \newline
+\subsubsection{Task 3 - Exploration of language with the DementiaBank dataset}
 
-\subsection{Chapter 3: Experiments with existing datasets: Presidential Press Conferences, Three Authors - Murdoch, Christie and James, DementiaBank dataset.}
 Aim: To replicate and extend the work by Berisha and Liss.
@@
 
-\subsection{Chapter 5: Longditunal Study}
-\subsubsection{Dataset creation}
+\subsubsection{Task 4 - Longditunal Study}
+
 The aim is to recruit a set of 50 participants with MCI / Early Dementia (diagnosed or undiagnosed) and a set of 50 controls with a similar age and broadly similar educational background (a known potential confound in assessing cognitive decline).
@@
 
+\subsubsection{Task 5 - Analysis of performance and validation of proposed Repeatable Neuropsychological Battery of Tests}
+
+\subsubsection{Task 6 - Thesis Writing}
+
 \subsection{Risk Assessment}
+In this section, I look at the potential difficulties that may produce a delay to the proposed research, define any strategies that have already been put in place to account for this as well as define any other potential solutions.
+\subsubsection{Delay in data collection}
+
+\subsubsection{Rejected Publications}
+
+\subsubsection{Rejected NHS Ethics}
 
 \subsection{Ethical Considerations}
+In this section I discuss the ethical considerations related to this project and any adjustments that I have made to my research because of these considerations.
+I am aware of the code of human  research ethics published by the British Psychological Society \cite{Society2012} and my research has been designed in a way that fits with these guidelines.
+\subsubsection{Risk}
+I will be working with a middle-aged to elderly population who have identified in some way that there is a difficulty with cognition.
+However, I do not anticipate working with vulnerable groups, those lacking capacity or individuals in a dependant or unequal relationship.
+Should any participants I am working with lose capacity during the course of the study, I will remove that participants details and data from the study.
+\par
+\subsubsection{
+
 Need to discuss confidentiality right at the start of each conversation and to reiterate at every meeting.

### 2019-01-10 Source/QualifyingReport/QualifyingReport.tex
QR work done

@@
 % HEADER BEGINS
-\documentclass[12pt, twoside, a4paper]{article}
+\documentclass{book}
 %% Packages
@@
 \fancyfoot[CE,CO]{Early Detection of MCI that progresses to dementia - An Interdisciplinary Approach}
-\fancyfoot[LE,RO]{\thepage}
+\fancyhead[LE,RO]{\thepage}
 
@@
 \newpage
-\begin{abstract}
-
-\end{abstract}
 \tableofcontents
@@
 \newpage
-\section{Overview of Problem}
+\chapter{Overview of Problem}
 Dementia has been identified as one of those fast growing difficulties facing the world.
@@
 \subsection{Research Challenges}
-\section{Literature Review}
+\chapter{Literature Review}
 \subsection{Introduction}
@@
 \centering
-\includegraphics[width=240px, height=150px]{images\BCTPicture.png}
+\includegraphics[width=240px, height=150px]{images/BCTPicture.png}
 \caption{Cookie Theft Picture.\label{white}}
@@
 \centering
-\includegraphics[width=240px, height=150px]{\images\picnic-scene.jpg}
+\includegraphics[width=240px, height=150px]{images/picnic-scene.jpg}
 \caption{Picnic Scene taken from the Western Aphasia Battery (WAB).\label{white}}
@@
 
-\section{Initial Study of the Presidents corpus}
+\chapter{Initial Study of the Presidents corpus}
 \subsection{Background}
@@
 Bush where there was a question and answer session conducted at least in part by the sitting president (2 press conferences of GWB were ommited due to a lack of a question and answer session).
+\subsubsection{Feature Selection}
 \subsection{Results}
@@
 
-\section{Future Work}
+\chapter{Future Work}
 \subsection{Introduction}
@@
 \subsubsection{Task 6 - Thesis Writing}
+The writing of the thesis, including additional over-running data analysis is planned for the last six months of the project (Oct '20 to Mar '21).
 
@@
 \subsubsection{Delay in data collection}
-
+When working with a clinical population, the collection of data can be subject to delays.
+This can be for a number of reasons including but not limited to sickness (patient or researcher), unavailability of participant for another reason or unavailability of venue to collect data.
+The collection of data could also be subject to delays due to ethical approval for the research being delayed.
+For individual participants unavailability, flexibility has been built into the data collection schedule to allow for an extra month (capacity for one additional cohort) of data collection per collection cycle.
 \subsubsection{Rejected Publications}
-
-\subsubsection{Rejected NHS Ethics}
-
+For each task defined above, with the exception of task 2 and task 6 I have identified suitable conferences at which I can present my work.
+With the target conferences identified every effort will be made to meet the submission deadlines for these conferences.
+If submissions are rejected at these conferences, I will identify alternative conferences to present at.
+\subsubsection{Rejected NHS Ethics and/or University Ethics}
+Every effort will be made to submit a complete ethics application through the NHS.
+However, in the event that there is a delay in the ethical approval process then data collection will proceed with a priority placed on the control group.
+Future cohorts will then have proportionally more clinical participants until the research meets it's target numbers.
 \subsection{Ethical Considerations}
@@
 \par
-\subsubsection{
+\subsubsection{Informed Consent}
 
@@
 
-\section{Conclusion}
+\chapter{Conclusion}
 

### 2019-01-14 Source/QualifyingReport/QualifyingReport.tex
QR edits

@@
 % HEADER BEGINS
-\documentclass{book}
+\documentclass[10pt, letterpaper, twoside, openany]{book}
 %% Packages
@@
 \usepackage{fancyhdr}
+\usepackage{rotating}
 %% Page Settings
@@
 \par
-In 2009, the Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{Health2009}.
+In 2009, the Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
 A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cogntive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease\cite{Taler2008}.
@@
 Finally, my conclusions are summarised in Chapter 5.
-\subsection{Background}
-\textbf{Full Provisional Title of the Research:} Title here \newline
-\textbf{Research Questions and Hypotheses}
+\section{Aims and Rationale}
+The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to to time-consuming and expensive psychological tests.
+There is a lot of research into the analysis of language as a bio-marker for MCI and Early Dementia.
+Given that sampling a person's language is relatively effortless, my research looks at whether we can find bio-markers of MCI and Early Dementia in natural language.\newline
+Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
+I'm not going to control for memory problems as a potential confound, but does this weaken the research?
+How do I defend this?
+\newline
 The broad question is can we use Machine Learning to assist in the early diagnosis of dementia and/or mild cognitive impairment through language produced naturally rather than language that is produced as part of a formal cognitive assessment.
@@
 \newline
-\par
-Deep Learning takes a different approach by bypassing the feature generating stage. Deep learning uses neural networks as a fundamental building block.
 
-\subsection{Aims and Rationale}
-The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to to time-consuming and expensive psychological tests.
-There is a lot of research into the analysis of language as a bio-marker for MCI and Early Dementia.
-Given that sampling a person's language is relatively effortless, my research looks at whether we can find bio-markers of MCI and Early Dementia in natural language.\newline
-Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
-I'm not going to control for memory problems as a potential confound, but does this weaken the research?
-How do I defend this?
-\newline
-\begin{itemize}
-	\item Hypothesis.
-I predict that, given a number of samples of language over a given time from, is it possible to detect language decline in humans such that you can differentiate between normal language decline and decline in those with MCI.
-	\item Hypothesis. I predict that my new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures.
-	\item My first hypothesis is whether monitoring language samples and other measures of cognitive decline over the course of the year is enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment.
-	\item My second hypothesis is that those with MCI show a different language profile to controls, if matched similarly for education and other significant factors.
-\end{itemize}
-
-\subsection{Research Challenges}
 \chapter{Literature Review}
-\subsection{Introduction}
+\section{Introduction}
 Alzheimer's Disease is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
 Whilst a definitive diagnosis of Dementia can only produced at post-mortem, there are a number of clinical indicators from psychological perspective that can indicate that dementia is present.
-From a clinical psychology perspective, those who have dementia have cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits\cite{AmericanPsychiatricAssociation2013}.
+From a clinical psychology perspective, those who have dementia have cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits\cite{McKhann2011}.
 In addition, these symptoms are often accompanied by emotional difficulties such as depression and irritability and behavioural difficulties.
@@
 According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
-This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population \cite{Sturrock2016}.
+This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population \cite{AlzheimersSociety2014}.
 From a financial perspective, they estimate an annual spend of £4.3 billion of which approximately £85 million is spent solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
@@
 All the drugs that are on the market are designed to manage symptoms.
-Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{Alzheimer2016}.
+Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{AlzheimersSociety2014}.
 Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place.
@@
 The case of Iris Murdoch illustrates this theory well.
-Le et al \cite{Le2012}found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel.
+Le et al \cite{Le2011}found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel.
 If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
 However, one of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of dementia.
-Albert et al have worked to define clinical criteria which professionals can use to diagnose MCI due to AD and differentiate this from age-associated memory impairment and age-associated cognitive decline.
-They note that the diagnosis of MCI requires evidence of intraindividual change and optimally requires evaluation at two or more points.
+Albert and his team have worked to define clinical criteria which professionals can use to diagnose MCI due to AD and differentiate this from age-associated memory impairment and age-associated cognitive decline.
+They note that the diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}.
 \begin{enumerate}
@@
 The second approach has gained more attention in recent years, due to the fact that in clinical practice it has shown promise in the early detection of AD.
-In addition, the analysis of the decline of cognitive abilities is comparitively inexpensive and less invasive that the first approach which commonly requires the collection of a sample of cerebro-spinal fluid which is painful for the patient involved.
+In addition, the analysis of the decline of cognitive abilities is comparatively inexpensive and less invasive that the first approach which commonly requires the collection of a sample of cerebro-spinal fluid which is painful for the patient involved.
 \par
@@
 It also has limited value in assessing under-educated subjects and a meta-analysis on the effectiveness of the MMSE as a diagnostic tool for dementia to showed that it's accuracy was low (sensitivity between 78.4\% and 85.1\% and specificity between 81.3\% and 87.8\%).
-As the MMSE is shown to have low accuracy specifically in the diagnosis of dementia, it becomes necessary for professionals to employ the use of other tools or measures such as the Free Cued Selective Reminding Test (FCSRT) or the Montreal Cognitive Assessment (MoCA).
+As the MMSE is shown to have low accuracy specifically in the diagnosis of dementia, it becomes necessary for professionals to employ the use of other tools or measures such as the Free Cued Selective Reminding Test (FCSRT) \cite{Grober2010} or the Montreal Cognitive Assessment (MoCA) \cite{Davis2015}.
 These tests have the benefits of being much more accurate at diagnosing cognitive impairment and discriminating between dementia and other types of cognitive impairment at the cost time and training of psychological professionals such as clinical psychologists in administering these tests.
@@
 % -- END OF INTRODUCTION --
-\subsubsection{Aims and Methodology}
+\subsection{Aims and Methodology}
 \par
@@
 We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English.
-\begin{table}
+\begin{sidewaystable}
+	\begin{center}
 	\begin{tabular}{ | c | c | c | p{1cm} |}
@@
 	\end{tabular}
+	\end{center}
 	\caption{\label{tab:table-name}Search Terms and Number of Results.}
-\end{table}
+\end{sidewaystable}
 % -- END OF Aims and Methodology
@@
 \subsubsection{Repeatable Battery for the Assessment of Neuropsychological Status}
-The Repeatable Battery for the Assessment of Neuropsychological Status (RBANS) was originally developed as an assessment tool for dementia, specfically looking at detecting a characterizing very mild dementia.
+The Repeatable Battery for the Assessment of Neuropsychological Status (RBANS) was originally developed as an assessment tool for dementia, specfically looking at detecting a characterizing very mild dementia \cite{Randolph1998}.
 The authors felt that there was a shortfall of appropriate measures that were sensitive enough to milder impairments as well as a number of other shortcomings of existing tests.
@@
 Another drawback of this battery is the lack of executive function measures and object naming tasks.
-Research has shown that the RBANS has good test-retest reliability and convergent validity, pa
+Research has shown that the RBANS has good test-retest reliability and convergent validity.
+However, Duff et al warned that caution should be exercised when using the RBANS in a MCI population as it has lower sensitivity in this population \cite{Duff2010}.
 
@@
 \par
-% Emrani - Assessing Working Memory in Mild Cognitive Impairment with Serial Order Recall
-Found no differences between specifically amnestic MCI and controls, but could differentiate between Mixed MCI and the other groups with mixed MCI recalling fewer correct responses than other groups.
+Emrani et al found no differences between specifically amnestic MCI and controls, but could differentiate between Mixed MCI and the other groups with mixed MCI recalling fewer correct responses than other groups.
 Notably, there was an attenuated recency effect in those with mixed MCI.
-% Kessels - Assessment of Working-memory deficits in patients with MCI and AD using Wechsler's Working Memory Index which includes digit span.
+\cite{Emrani2018}.
 Kessels identifies that that there are working memory deficits in MCI patients and these worsen with AD patients.
-% Kessels - Assessment of verbal and visuospatial working memory in MCI and AD.
-Kessels identifies that both MCI and AD have impaired performance on all three conditions of the digit span test.
+\cite{Kessels2011} He also identified that both MCI and AD have impaired performance on all three conditions of the digit span test.
 No differences were found between forward and backward conditions in any of the groups.
-However, available tests may not detect subtle impairments.
+However, available tests may not detect subtle impairments \cite{Kessels2015}.
 \subsubsection{Rey Auditory-Verbal learning test}
@@
 Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then again after 30 minutes.
-% Moradi - Rey's Auditory Verbal Learning Test scores can be predicted from whole brain MRI in AD.
+\par
 Several studies have shown that an impairment in RAVLT score reflect well the underlying pathology caused by AD.
@@
 They found a highly significant cross validated correlation between the estimated and observed RAVLT immediate and RAVLT Percent Forgetting.
-Further, they found that the conversion of MCI subjects to AD in 3-years could be predicted based on either observed or estimated RAVLT scores with an accuract comparible to MRI-based biomarkers.
+Further, they found that the conversion of MCI subjects to AD in 3-years could be predicted based on either observed or estimated RAVLT scores with an accuracy comparible to MRI-based biomarkers \cite{Moradi2017}.
 % Schoenberg - Test performance and classification statistics for the Rey Auditory Verbal Learning Test in selected clinical samples.
@@
 A number of researchers have used n-grams as features.
-One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class.
-Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls.
-Their main finding supported n-grams as the most significant predictor.
-One of the criticisms is the use of picture description tasks and n-grams.
-Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
 \par
@@
 
-\subsection{State of literature into Natural Language processing techniques}
+\subsection{State of literature into Machine Learning and Natural Language processing techniques}
 As we can see, diagnosing dementia through analysis has a large background in terms of psychological research.
@@
 That is to say that we can subdivide NLP processes into different components.
-This broadly fits with the literature review conducted by Emery.
+This broadly fits with the literature review conducted by Emery \cite{Emery2000}.
 \begin{enumerate}
@@
 \end{enumerate}
-\subsubsection{Traditional methods of Machine Learning and Natural Language Processing}
+\subsubsection{Traditional methods of Machine Learning}
+% Thomas - Automatic Detection and Rating of AD through the Lexical Analysis of Spontaneous Speech
+Thomas et al, detail several lexical approaches to the problem of detecting and rating DAT.
+The approaches rely on character n-gram techniques but also explore correlation of usage of frequency of different types of speech.
+Their results act as a proof in concept of the utility of using a pure computational approach to the diagnosis of dementia using spontaneous speech.
+Results - 95\% accuracy dementia vs controls.
+70\% accuracy classifying dementia into two categories.
+50\% accuracy classifying dementia into four categories.
+They suggest further exploration of characteristics to classify.
+% Bucks et al - Analysis of spontaneous, conversational speech in dementia of the Alzheimer's Type: Evaluation of an objective technique for analyzing lexical performance.
+AD patients had higher mean P rate, A rate and V rate scores but lower N rate scores vs normal older controls \cite{Bucks2000}.
+One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class.
+Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls.
+Their main finding supported n-grams as the most significant predictor.
+One of the criticisms is the use of picture description tasks and n-grams.
+Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
+\subsubsection{The case for Deep Learning}
 
-\subsubsection{Deep Learning}
+\subsection{Conclusions}
 
@@
 \chapter{Initial Study of the Presidents corpus}
-\subsection{Background}
+\section{Background}
 
-\subsection{Methods}
+\section{Methods}
 I took 46 transcripts of Ronald Reagan’s (RR) press conferences from 1981 to 1988 and compared them with 134 press conferences by George H.
@@
 The APP is a comprehensive and organized searchable database of presidential documents, including transcripts of speeches, transcripts of news conferences, and other public documents.
-\subsubsection{Pre-processing}
+\subsection{Pre-processing}
 To generate the files necessary for analysis, I downloaded each transcript and performed the following changes.
@@
 Bush where there was a question and answer session conducted at least in part by the sitting president (2 press conferences of GWB were ommited due to a lack of a question and answer session).
-\subsubsection{Feature Selection}
-\subsection{Results}
-\subsection{Discussion}
-\subsection{Conclusion}
+\subsection{Feature Selection}
+\section{Results}
+\section{Discussion}
+\section{Conclusion}
 
 \chapter{Future Work}
-\subsection{Introduction}
+\section{Introduction}
+This section looks at my research aims and objectives given the literature review presented above.
+This section will then go on to explore any challenges to my research.
+I then present the work I have planned for the next two years, divided into distinct tasks with their own aims and objectives.
+As part of this I will identify target conferences where I could present my work.
+Finally, I look at potential risks and ethical concerns that may occur as part of my research.
+\section{Research Aims and Objectives}
+\par
+The broad aim of my research is to explore whether the application of Natural Language Processing and Machine Learning techniques to spontaneous speech samples can improve our ability to diagnose Mild Cognitive Impairment in patients that will directly lead to Alzheimer's Disease.
+An additional outcome of my research is the development of a repeatable battery of neuropsychological tests that specifically focus on being sensitive to Mild Cognitive Impairment.
+\begin{itemize}
+	\item \textbf{Hypothesis} Is monitoring language samples and other measures of cognitive decline over the course of the year enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment.
+	\item \textbf{Hypothesis} Those with MCI will show a different language profile to controls if matched similarly for age, gender, education and other significant factors.
+	\item \textbf{Hypothesis} The new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures.
+	\item \textbf{Hypothesis} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, those with Alzheimer's Disease and decline in those with MCI.
+\end{itemize}
 
-\subsection{Additional Research Aims and Objectives}
+\section{Research Challenges}
+\subsection{Data Collection}
+\subsection{Commonality of MCI}
 
-\subsection{Additional Research Challenges}
+\section{Project Plan and Task Definition}
+\subsection{Task 1 - Replication and Extension of language analysis in the Presidents Speeches Dataset}
 
-\subsection{Project Plan and Task Definition}
-\subsubsection{Task 2 - Pilot Study for the Neuropsychological Battery of Tests}
-I recruited a cohort of 9 volunteers which which to complete the battery of Neuropsychological tests.
+\subsection{Task 2 - Pilot Study for the Neuropsychological Battery of Tests}
+I recruited a cohort of 4 volunteers which which to complete the battery of Neuropsychological tests.
 We compiled a series of tests (see literature review for a comprehensive list, explanation and rationale for each test) and completed each version of the test on three consequetive days.
@@
 \newline
-\subsubsection{Task 3 - Exploration of language with the DementiaBank dataset}
+\subsection{Task 3 - Exploration of language with the DementiaBank dataset}
 
@@
 
-\subsubsection{Task 4 - Longditunal Study}
+\subsection{Task 4 - Longditunal Study}
 
@@
 
-\subsubsection{Task 5 - Analysis of performance and validation of proposed Repeatable Neuropsychological Battery of Tests}
+\subsubsection{Participants}
+We are looking at a number of people with subjective MCI and a diagnosis of MCI.
 
-\subsubsection{Task 6 - Thesis Writing}
+\subsection{Task 5 - Analysis of performance and validation of proposed Repeatable Neuropsychological Battery of Tests}
+
+\subsection{Task 6 - Thesis Writing}
 The writing of the thesis, including additional over-running data analysis is planned for the last six months of the project (Oct '20 to Mar '21).
 
-\subsection{Risk Assessment}
+\section{Skills Development}
+
+\section{Risk Assessment}
 In this section, I look at the potential difficulties that may produce a delay to the proposed research, define any strategies that have already been put in place to account for this as well as define any other potential solutions.
-\subsubsection{Delay in data collection}
+\subsection{Delay in data collection}
 When working with a clinical population, the collection of data can be subject to delays.
@@
 For individual participants unavailability, flexibility has been built into the data collection schedule to allow for an extra month (capacity for one additional cohort) of data collection per collection cycle.
-\subsubsection{Rejected Publications}
+\subsection{Rejected Publications}
 For each task defined above, with the exception of task 2 and task 6 I have identified suitable conferences at which I can present my work.
@@
 If submissions are rejected at these conferences, I will identify alternative conferences to present at.
-\subsubsection{Rejected NHS Ethics and/or University Ethics}
+\subsection{Rejected NHS Ethics and/or University Ethics}
 Every effort will be made to submit a complete ethics application through the NHS.
@@
 Future cohorts will then have proportionally more clinical participants until the research meets it's target numbers.
-\subsection{Ethical Considerations}
+
+\section{Ethical Considerations}
 In this section I discuss the ethical considerations related to this project and any adjustments that I have made to my research because of these considerations.
 I am aware of the code of human  research ethics published by the British Psychological Society \cite{Society2012} and my research has been designed in a way that fits with these guidelines.
-\subsubsection{Risk}
+\subsection{Risk}
 I will be working with a middle-aged to elderly population who have identified in some way that there is a difficulty with cognition.
@@
 Should any participants I am working with lose capacity during the course of the study, I will remove that participants details and data from the study.
-\par
-\subsubsection{Informed Consent}
+\subsection{Informed Consent}
+Consent will be sought overtly at the recruitment and initial data collection point by way of an information sheet and associated consent form.
+At other data collection points, I will have information sheets available if participants request them and if such a request is made then I will seek consent again at this stage.
+As part of this process, I will be supplying my participants with my contact details should they have any question or for the purposes of withdrawing consent.
+These consent forms will be stored in a secure, locked cabinet at the university.
+\subsection{Confidentiality}
+Given the nature of the study and the possibility of disclosures of suicidal ideation, it is important that confidentiality is discussed at each data collection point.
+I will also provide an information sheet which details the confidentiality statement and also the situations in which I may need to break confidentiality.
+In addition, I will make a verbal confidentiality statement at each collection point and seek acknowledgement of their understanding and permission to continue.
+The spoken confidentiality statement will be as follows.
+\begin{center}
+	\textit{"Anything you say to me today will be confidential unless I feel that there is a risk to you, or to someone else.
+If this is the case, I may need to speak to or write to third parties such as your GP.
+I will endeavor to discuss this communications with yourself where it is possible to do so.
+Is this ok?"} \newline
+\end{center}
 
-Need to discuss confidentiality right at the start of each conversation and to reiterate at every meeting.
-I will also provide an information sheet which details the confidentiality statement.
-The spoken confidentiality statement will be as follows.
-"Anything you say to me today will be confidential unless I feel that there is a risk to you, or someone else.
-If this is the case, I may need to speak to or write to third parties such as your GP.
-Is this ok?" \newline
-\par
-\subsubsection{What if someone reveals suicidal ideation or other risks?}
+\subsection{Disclosure of suicidal ideation or other risks}
 As part of the battery of tests for this experiment, there is an opportunity for a participant to reveal that he or she has had suicidal ideation in the past two weeks (Patient Health Questionnaire, Question 9).
@@
 
-\subsection{Thesis - Provisional Table of Contents}
+\section{Thesis - Provisional Table of Contents}
 In this section I provide a provisional table of contents based on the work I have either completed or proposed above.
@@
 \bibliographystyle{unsrt}
-\bibliography{QualifyingReport.bib}
+\bibliography{QualifyingReport}
 

### 2019-01-15 Source/QualifyingReport/QualifyingReport.tex
work done on QR

@@
 \listoftables
-\newpage
 \listoffigures
@@
 While this is help to learn that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments.
-\newline
 \par
@@
 A company in Toronto, Canada has taken this approach and currently uses over 400 acoustic and language features to make this classification and to generate a psuedo MMSE score.
-\newline
 
@@
 \par
+\subsubsection{Type token ratio(TTR)}
+Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
+\begin{equation} \label{x1}
+TTR = numberOfUniqueWords / totalNumberOfWords.
+\end{equation}
+\subsubsection{Brunet's Index(W)} % Don't forget to cite the original article.
+Brunet's Index (W) differentiates itself for TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
+\begin{equation} \label{x2}
+W = N^{V(-0.165)}
+\end{equation}
+where N is the total length of the utterance being measured and V is equal to the total vocabulary being used by the subject.
+Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers.
+\newline
+\par
+\subsubsection{Honore's Statistic (R)} % Don't forget to cite the original article.
+Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation:
+\begin{equation} \label{x3}
+%% Check to see if this is accurate.
+R = (100 \log N) / (1 - V1/V)
+\end{equation}
+where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
 \subsubsection{Picture-related content thematic elements}
@@
 Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in dialogue in which a speaker provides responsees do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls.
-\subsubsection{Lexical richness and diversity}
-\subsubsection{Type token ratio(TTR)}
-Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
-\begin{equation} \label{x1}
-TTR = numberOfUniqueWords / totalNumberOfWords.
-\end{equation}
-\subsubsection{Brunet's Index(W)} % Don't forget to cite the original article.
-Brunet's Index (W) differentiates itself for TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
-\begin{equation} \label{x2}
-W = N^{V(-0.165)}
-\end{equation}
-where N is the total length of the utterance being measured and V is equal to the total vocabulary being used by the subject.
-Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers.
-\newline
-\par
-\subsubsection{Honore's Statistic (R)} % Don't forget to cite the original article.
-Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation:
-\begin{equation} \label{x3}
-%% Check to see if this is accurate.
-R = (100 \log N) / (1 - V1/V)
-\end{equation}
-where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
 \subsubsection{Quantity - Total number of words}
@@
 Teng et al suggests that work should focus on the MCI population and concentrate on developing a consensus neuropsychological battery that could yield predictable rates of progression to AD.
-This, and with the development of a model of language, sensitive enough to detect subtle deterioration in language use to act as an additional cognitive marker to aid diganosis could potentially move some way to providing this.
+This, and with the development of a model of language, sensitive enough to detect subtle deterioration in language use to act as an additional cognitive marker to aid diagnosis could potentially move some way to providing this.
 \par 
@@
 \section{Background}
-
+There has been a significant research in the area of language deterioration as a means of detecting Alzheimer's Disease.
+This usually takes the form analysis of speech recorded as part of a cognitive assessment such as the Picture Description Task \cite{Orimaye2008}.
+Given that language samples are the relatively easy to collect, research has moved towards analysis of spontaneous speech.
+An good example of this type of research is the study conducted by Berisha and Liss which looked at the differences in language use between two US presidents, Ronald Reagan (who would go on to receive a diagnosis of Dementia) and George H.W.
+Bush who acted as a matched control based on Age \citep{Berisha2015}.
+They found several differences in language use which they felt acted as indicators of Reagan's difficulties with language due to dementia.
+These significant differences were in the number of unique words used per speech, the use of non-specific nouns and fillers and low-imageability verbs \citep{Berisha2015}.
+/par   
+This study replicates work done by Berisha and Liss and extends this by adding Donald Trump as an alternative, more appropriate comparison to Ronald Reagan as he is much closer in age than George H.W.
+Bush.
+This experiment will look at the features originally identified by Berisha and Liss, as well as any others that have potential as discussed in the literature review above.
 \section{Methods}
@@
 The broad aim of my research is to explore whether the application of Natural Language Processing and Machine Learning techniques to spontaneous speech samples can improve our ability to diagnose Mild Cognitive Impairment in patients that will directly lead to Alzheimer's Disease.
-An additional outcome of my research is the development of a repeatable battery of neuropsychological tests that specifically focus on being sensitive to Mild Cognitive Impairment.
+Another aim of my research is to provide some clarity as to which language features are most important to measure when looking for evidence of MCI and/or AD.
+Finally, an additional outcome of my research is the development of a repeatable battery of neuropsychological tests that specifically focus on being sensitive to Mild Cognitive Impairment.
 \begin{itemize}
@@
 \subsection{Data Collection}
+
 \subsection{Commonality of MCI}
 
-\section{Project Plan and Task Definition}
-\subsection{Task 1 - Replication and Extension of language analysis in the Presidents Speeches Dataset}
+\subsection{Sparsity of Language}
 
+\section{Project Plan and Task Definitions}
+In this section, I describe each task in detail along with a projected timeline depicted in a Gantt chart.
 \subsection{Task 2 - Pilot Study for the Neuropsychological Battery of Tests}
-I recruited a cohort of 4 volunteers which which to complete the battery of Neuropsychological tests.
-We compiled a series of tests (see literature review for a comprehensive list, explanation and rationale for each test) and completed each version of the test on three consequetive days.
-The results of which were compiled and analysed.
-The hypothesis is that there would be no difference in performance between one battery of tests and another.
-An additional hypothesis is that the language tasks included on each task would be enough to collect an appropriate sample for linguistic analysis, such that we may be able to detect cognitive decline in a clinical population.
-\newline
-\subsection{Task 3 - Exploration of language with the DementiaBank dataset}
-
-Aim: To replicate and extend the work by Berisha and Liss.
-To use traditional machine learning techniques to test whether they can differentiate between Reagan, Bush and Trump using similar features.
-To test to see whether Deep Learning can also differentiate between the three.
-\newline
-Dataset: the presidential press conferences given by Presidents Ronald Reagan, George H.W. Bush and Donald Trump. \newline
-Paper: This experiment replicates work done by Berisha and Liss and extends this by adding Donald Trump as an alternative comparison to Ronald Reagan.
-This experiment will look at the features originally recommended by Berisha and Liss, as well as any others that have potential as discussed in the literature review above.
-\newline
-Aim: To replicate the work by Le et al.
-To use traditional machine learning techniques to see if it can differentiate between the three authors using similar features.
-To test to see whether deep learning can also differentiate between the three.
-To add 'pre-trained' layers from Presidential Debates experiment to see if this improves accuracy.
-\newline
-Dataset is the literary novels of Iris Murdoch, Agatha Christie and P.D. James. \newline
-Paper:
-Aim: To replicate and extend the work of Orimaye et al.
-To use traditional machine learning techniques to see if it can categorise people into dementia or healthy categories using the samples given.
-To see whether can use pre-trained layer from presidential debates to improve accuracy.
-It doesn't seem relevant to use pretrained layer from Experiment four, but it might be worth looking at the results of this.
-\newline
-Dataset: The DementiaBank Corpus. \newline
-Paper: To replicate and extend the work of Orimaye et al. To compare traditional machine learning techniques (already completed by Orimaye) and deep learning. \newline
-
+The aim of this study is complete a trial run of the administration of a new battery of Neuropsychological Battery of Tests.
+We aim to find out whether the tests in question generate enough language so that features such as those generated in Task 1 are meaningful and can be used as language markers that can then be tracked over time.
+We also aim to look at whether there are significant practice effects between three versions of these tests with a gap of seven days between each test.
+\par 
+We compiled a series of tests which look a number of cognitive functions which have previously demonstrated a level of effectiveness at discriminating an MCI population from a healthy population (see literature review for a comprehensive list, explanation and rationale for each test).
+In addition to these tests, we have incorporated two language tasks in which a participants ability to generate spontaneous speech that we can then use to further analyse.
+As a planned future study is longitudinal, it is important that this battery of tests is repeatable but does not have significant practice effects that may interfere with our ability to track cognitive decline.
+\par 
+We have recruited a cohort of 4 volunteers who will complete the battery of Neuropsychological tests on three separate days with a gap of at least seven days between each administration.
+The results of which will compiled and analysed.
+Our hypothesis is that there would be no difference in performance between one battery of tests and another.
+An additional hypothesis is that the language tasks included on each test would be enough to collect an appropriate sample for linguistic analysis such that we may be able to detect cognitive decline in a clinical population.
+\par 
+While the study in itself is not valid for submission to a conference or journal, it is hoped that the results of the study will allow subsequent studies using the same or adjusted battery of tests to be published.
+\subsection{Task 3 - Exploration of language with the DementiaBank dataset and Three Authors dataset.}
+The aim of this task is to replicate and extend the work of Fraser and Orimaye who both looked at the DementiaBank dataset and set out to use traditional machine learning techniques to see whether they could categorise people into dementia or healthy categories using the samples given.
+We also aim to replicate and extend the work of Le et al who used electronic versions of the novels of three authors.
+Both datasets are longitudinal and have been shown to track language decline over time in different ways.
+\par 
+The DementiaBank dataset was a product of a longitudinal study which was run by the University of Pittsburgh School of Medicine.
+The dataset contains transcripts of interviews conducted with participants with MCI / AD and other forms of dementia.
+These interviews were conducted in English and used the cookie-theft picture language task (see Literature Review for a full description).
+The full data set consists of 314 patients classified as having dementia and 242 patients who are classified as healthy and these patients are tested on a minimum of two separate occasions.
+\par 
+The Three Authors dataset includes the novels of P.D.
+James, Agatha Christie and Iris Murdoch.
+It contains twenty of Murdoch's novels which were published between the ages of 35 and 76, sixteen of Christie's novels written between the ages of 28 and 82 and fifteen of P.D.
+James novels written between the ages of 42 and 82.
+These three authors are notable as Murdoch had a confirmed diagnosis of Dementia after her death, Christie had a suspected diagnosis of Dementia and P.D.
+James was used as a control.
+\par 
+These two existing datasets in addition to the President's Press Conference corpus described above will provide three sources of validation of features that we generate as well as give us clues as how to best process the data we receive from our longitudinal study.
+\par 
+The final output of this task will be a paper which looks at the importance of various word-level and sentence-level features in the deterioration of language for people with AD.
+The target conference for this work is the UK Alzheimer's Research conference in March 2020 with an expected Abstract submission date of December 2019 and the Student Research Workshop of the ACL in July 2020 with an expected submission date of April 2020.
 \subsection{Task 4 - Longditunal Study}
-
-The aim is to recruit a set of 50 participants with MCI / Early Dementia (diagnosed or undiagnosed) and a set of 50 controls with a similar age and broadly similar educational background (a known potential confound in assessing cognitive decline).
-The rationale for recruiting this number is that there will be an expected drop out rate and so
-Paper:
-
-\subsubsection{Participants}
-We are looking at a number of people with subjective MCI and a diagnosis of MCI.
-
-\subsection{Task 5 - Analysis of performance and validation of proposed Repeatable Neuropsychological Battery of Tests}
+Previously we have looked at historical datasets, the aim of this task is see whether a formalised collection of specific data, namely our language tasks are enough to be able to be able to detect cognitive decline in participants at different stages of the MCI / Alzheimer's disease progression.
+To this end, we plan on administering our neuropsychological battery of tests from Task 2 to each participant three times (six monthly intervals) over the course of the year.
+\par 
+We therefore plan to recruit a set of 80 participants divided evenly into four groups.
+These groups are Healthy Older Controls (HOC), Early Alzheimers (EA), Subjective Mild Cognitive Impairment (SMCI) and Diagnosed Mild Cognitive Impairment (DMCI).
+We expect to see typical age related decline in our HOC group and more dramatic decline in our EA group.
+We are most interested to see if there is any measurable language decline in the SMCI or DMCI experimental groups.
+\par 
+Whilst we have divided participants into experimental groups we aim to investigate each participant individually.
+Because there is a large amount of heterogeneity in this population, we will initially be analysing each participants results in isolation with the participant acting as their own control.
+However, we will also be exploring the use of traditional machine learning and/or deep learning across cohorts to investigate whether each cohort has their own distinct 'language profile'.
+\par 
+The final output of this task will be a paper discussing the results of our longitudinal study from an language analysis perspective.
+The target conference for this work will be the UK Alzheimer's Research conference in March 2021 with an expected submission date of December 2020, the Alzheimer's Association International Conference in July 2021 with an expected submission date of January 2021 and the Association of Computational Linguistics Conference (ACL) in July 2021 with an expected submission date of March 2021.
+\subsection{Task 5 - Analysis of performance and validation of the proposed Repeatable Neuropsychological Battery of Tests}  
+In this task, I will be looking at the results of the neuropsychological battery of tests without the additional language tasks to determine whether we can detect cognitive decline. 
+\par 
+We expect to see typical age related decline in our HOC group and more severe decline in our EA group.
+Once again, we will be interested in whether there is any measurable decline in our SMCI or DMCI experimental groups with our battery of tests.
+\par 
+The final output of this task will be a paper discussing the results of the battery of tests, including how the battery does in terms of diagnostic and prognostic utility.
+The target conference for this work is to be decided.
 
@@
 The writing of the thesis, including additional over-running data analysis is planned for the last six months of the project (Oct '20 to Mar '21).
-
-\section{Skills Development}
 
@@
 With the target conferences identified every effort will be made to meet the submission deadlines for these conferences.
-If submissions are rejected at these conferences, I will identify alternative conferences to present at.
+If submissions are rejected at these conferences, I have identified alternative conferences to present at.
 \subsection{Rejected NHS Ethics and/or University Ethics}
@@
 Future cohorts will then have proportionally more clinical participants until the research meets it's target numbers.
+In the event that NHS ethics is not granted in time for me to write up, participants will be sought from other areas such as the Alzheimer's Research pool of volunteers and venues where suitable candidates may be present such as meetings of the University of the Third age.
 
@@
 I will endeavor to discuss this communications with yourself where it is possible to do so.
-Is this ok?"} \newline
+Is this ok?"}
 \end{center}
-
 \subsection{Disclosure of suicidal ideation or other risks}
@@
 In the event that the participant scores 0 on Q9 of the PHQ, no risk assessment will be undertaken.
-\newline
 

### 2019-01-16 Source/QualifyingReport/QualifyingReport.tex
QR edits 14-1-19

@@
 \par
-The potential impact of this research is immense.
+The potential impact of this research in this area is immense.
 Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease.
@@
 This report is structured in the following way.
-The context of the project, the aims and hypotheses of the research and the challenges the research presents are detailed in the remainder of this chapter.
 A literature review of current methods of cognitive assessment of MCI, an review of how language is said to be impacted in the MCI and Early AD population and a review of NLP and ML techniques that have the potential to be applied to this domain is presented in Chapter 2.
 The findings from a replication and extension of previous research using an existing dataset is presented in Chapter 3.
-I detail my future work plan including distinct tasks to be completed over the next two years in Chapter 4.
+I detail the aims of my research and my proposed future work plan including distinct tasks to be completed over the next two years in Chapter 4.
 Finally, my conclusions are summarised in Chapter 5.
-\section{Aims and Rationale}
-The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to to time-consuming and expensive psychological tests.
-There is a lot of research into the analysis of language as a bio-marker for MCI and Early Dementia.
-Given that sampling a person's language is relatively effortless, my research looks at whether we can find bio-markers of MCI and Early Dementia in natural language.\newline
-Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
-I'm not going to control for memory problems as a potential confound, but does this weaken the research?
-How do I defend this?
-\newline
-The broad question is can we use Machine Learning to assist in the early diagnosis of dementia and/or mild cognitive impairment through language produced naturally rather than language that is produced as part of a formal cognitive assessment.
-There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia.
-These studies have identified a number of speech and language differences between healthy controls and those with dementia.
-Some of these features have been validated in work using machine learning to classify healthy controls from those with dementia.
-However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'.
-While this is help to learn that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments.
-\par
-Machine Learning itself is undergoing a shift in focus at the present time.
-I will describe this shift using the context of my proposed research.
-In traditional machine learning, the transcripts from cognitive assessments are used to generate features such as idea density.
-These features are then used to classify patients.
-A company in Toronto, Canada has taken this approach and currently uses over 400 acoustic and language features to make this classification and to generate a psuedo MMSE score.
 
@@
 \section{Introduction}
-Alzheimer's Disease is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
-Whilst a definitive diagnosis of Dementia can only produced at post-mortem, there are a number of clinical indicators from psychological perspective that can indicate that dementia is present.
-From a clinical psychology perspective, those who have dementia have cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits\cite{McKhann2011}.
-In addition, these symptoms are often accompanied by emotional difficulties such as depression and irritability and behavioural difficulties.
+Alzheimer's Disease (AD) is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
+Whilst a definitive diagnosis of Dementia can only produced at post-mortem, there are a number of clinical indicators from psychological perspective that can indicate dementia is present.
+From a clinical psychology perspective, those who have dementia demonstrate cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language and visuospatial deficits\cite{McKhann2011}.
+In addition, these symptoms are often accompanied by emotional problems such as depression and behavioural difficulties.
 \par
@@
 Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time and alter cognitive function until a threshold is reached and clinical symptoms become more obvious \cite{Nestor2006}.
-The case of Iris Murdoch illustrates this theory well.
-Le et al \cite{Le2011}found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel.
+The case of Iris Murdoch, who had a confirmed diagnosis of dementia, illustrates this theory well.
+Le et al \cite{Le2011}found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel when, it is theorized, the symptoms of AD manifested themselves more significantly.
 If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
@@
 The MMSE was not specifically created to screen for dementias and therefore does not interrogate key aspects of cognitive impairment known to be affected in dementia.
-It also has limited value in assessing under-educated subjects and a meta-analysis on the effectiveness of the MMSE as a diagnostic tool for dementia to showed that it's accuracy was low (sensitivity between 78.4\% and 85.1\% and specificity between 81.3\% and 87.8\%).
+It also has limited value in assessing under-educated subjects and a meta-analysis on the effectiveness of the MMSE as a diagnostic tool for dementia showed that it's accuracy was low (sensitivity between 78.4\% and 85.1\% and specificity between 81.3\% and 87.8\%).
 As the MMSE is shown to have low accuracy specifically in the diagnosis of dementia, it becomes necessary for professionals to employ the use of other tools or measures such as the Free Cued Selective Reminding Test (FCSRT) \cite{Grober2010} or the Montreal Cognitive Assessment (MoCA) \cite{Davis2015}.
@@
 There are a wide number of factors that are involved in language degeneration in the elderly, and consequently there will be an expected amount of variability between subjects.
-The administration of such tests may induce nervousness and discomfort which may impact performance, and also repeatedly administering the same language tests to tests for differences over time by be confounded by improved performance at tasks via practice effects.
-Also, generally speaking the language tests that are currently being used do not necessarily describe patients 'real performance' in language production.
+The administration of such tests may induce nervousness and discomfort which may impact performance, and also repeatedly administering the same language tests for differences over time may be confounded by improved performance at tasks via practice effects.
 However, there is enough promise in this approach such that it could help further our understanding of the disease, it's progression and the parts of the brain affected in the early stages.
@@
 She found that people with AD generally had intact Phonology and Morphology but more impaired Syntax and Semantics.
-She stated that the language forms we learn last are the first to deteriorate.
-We generally learn language in small simple units initially and build syntax and complexity as we are more comfortable with language.
+She asserted that the language forms we learn last are the first to deteriorate as we generally learn language in small simple units initially and build syntax and complexity as we are more comfortable with language.
 \par
@@
 % -- END OF Aims and Methodology
-\subsection{Existing Neuropsychological Measures of Cognitive Impairment, and Repeatable Battery of Neuropsychological Tests}
+\section{Existing Neuropsychological Measures of Cognitive Impairment, and Repeatable Battery of Neuropsychological Tests}
 In terms of standardised cognitive tests, there are two main aims.
@@
 \par 
-\subsubsection{Repeatable Battery for the Assessment of Neuropsychological Status}
+\subsection{Repeatable Battery for the Assessment of Neuropsychological Status}
 The Repeatable Battery for the Assessment of Neuropsychological Status (RBANS) was originally developed as an assessment tool for dementia, specfically looking at detecting a characterizing very mild dementia \cite{Randolph1998}.
@@
 An good example of this type of research is the study conducted by Berisha and Liss which looked at the differences in language use between two US presidents, Ronald Reagan (who would go on to receive a diagnosis of Dementia) and George H.W.
-Bush who acted as a matched control based on Age \citep{Berisha2015}.
+Bush who acted as a matched control based on Age \cite{Berisha2015}.
 They found several differences in language use which they felt acted as indicators of Reagan's difficulties with language due to dementia.
-These significant differences were in the number of unique words used per speech, the use of non-specific nouns and fillers and low-imageability verbs \citep{Berisha2015}.
+These significant differences were in the number of unique words used per speech, the use of non-specific nouns and fillers and low-imageability verbs \cite{Berisha2015}.
 /par   
@@
 \par
+\section{Research Aims}
+The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to time-consuming and expensive psychological tests.
+There is a lot of research into the analysis of language as a marker for MCI and Early Dementia \cite{Taler2008, Boschi2017} and there have been calls for research that explores the use of Natural Language Processing and Machine Learning techniques to aid in the analysis of language \cite{Boschi2017}.
+It is this area which is the focus of my research.
+\par 
+Given that sampling a person's language is relatively effortless, my proposed research will look at whether we can find markers of MCI in natural language.
+I focus particularly on MCI as this is often a precursor to AD and therefore detection of MCI in an elderly population would allow I plan to explore the use Natural Language Processing and Machine Learning techniques to assist in this process, and focus on MCI specifically.
+early diagnosis of dementia and/or mild cognitive impairment through language produced naturally rather than language that is produced as part of a formal cognitive assessment.
+There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia.
+These studies have identified a number of speech and language differences between healthy controls and those with dementia.
+Some of these features have been validated in work using machine learning to classify healthy controls from those with dementia.
+However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'.
+While this is help to learn that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments.
+\par
+Machine Learning itself is undergoing a shift in focus at the present time.
+I will describe this shift using the context of my proposed research.
+In traditional machine learning, the transcripts from cognitive assessments are used to generate features such as idea density.
+These features are then used to classify patients.
+A company in Toronto, Canada has taken this approach and currently uses over 400 acoustic and language features to make this classification and to generate a psuedo MMSE score.
 The broad aim of my research is to explore whether the application of Natural Language Processing and Machine Learning techniques to spontaneous speech samples can improve our ability to diagnose Mild Cognitive Impairment in patients that will directly lead to Alzheimer's Disease.
@@
 \subsection{Sparsity of Language}
+
+Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
+I'm not going to control for memory problems as a potential confound, but does this weaken the research?
+How do I defend this?
+\newline
 

### 2019-01-16 Source/QualifyingReport/QualifyingReport.tex
QR adjustments for Brie

@@
 \section{Existing Neuropsychological Measures of Cognitive Impairment, and Repeatable Battery of Neuropsychological Tests}
-In terms of standardised cognitive tests, there are two main aims.
-Does the test distinguish accurately between normal aging, MCI and AD (diagnostic utility) and, does the test distinguish between those individuals with MCI who will then go on to develop AD and those individuals with MCI who don't then go on to develop AD (prognostic utility).
-This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their perfomance in terms of both diagnostic and prognostic utility.
+In terms of standardised cognitive tests there are two main aims.
+Does the test distinguish accurately between normal aging, MCI and AD (diagnostic utility) and does the test distinguish between those individuals with MCI who will then go on to develop AD and those individuals with MCI who don't then go on to develop AD (prognostic utility).
+This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their performance in terms of both diagnostic and prognostic utility.
 \par 
 \subsection{Repeatable Battery for the Assessment of Neuropsychological Status}
-The Repeatable Battery for the Assessment of Neuropsychological Status (RBANS) was originally developed as an assessment tool for dementia, specfically looking at detecting a characterizing very mild dementia \cite{Randolph1998}.
+The Repeatable Battery for the Assessment of Neuropsychological Status (RBANS) was originally developed as an assessment tool for dementia, specifically looking at detecting a characterizing very mild dementia \cite{Randolph1998}.
 The authors felt that there was a shortfall of appropriate measures that were sensitive enough to milder impairments as well as a number of other shortcomings of existing tests.
@@
 
-\subsubsection{Digit Span Test}
-The Digit Span Test is predominantly used to measure a person's working memory capacity, specifically the capacity used to store and recall numbers.
+\subsection{Digit Span Test}
+The Digit Span Test (DST) is predominantly used to measure a person's working memory capacity, specifically the capacity used to store and recall numbers.
 A participant is presented with a series of numbers of fixed length and is asked to recall those numbers in normal or reverse order.
@@
 % Muangpaisan - Digit span and verbal fluency tests in patients with MCI and normal subjects in Thai community.
-MCI patients had significantly lower digit span score, in both normal and reverse order versions of this test.
-Research has shown that Digits Backwards can, to some degree, predict a diagnosis of MCI.
-They found that Age, Gender and Education have an impact on the performance of the tests.
-\par
-Emrani et al found no differences between specifically amnestic MCI and controls, but could differentiate between Mixed MCI and the other groups with mixed MCI recalling fewer correct responses than other groups.
+Research has consistently shown that MCI patients have a significantly lower digit span score in both normal and reverse order versions of this test and a study by Muangpaisan has shown that the reverse order version of this test can, to some degree, predict a diagnosis of MCI.
+Muangpaisan also found that Age, Gender and Education have an impact on the performance of the tests.
+Emrani et al found that the DST revealed no differences between specifically amnestic MCI and controls, but could differentiate between Mixed MCI and the other groups with mixed MCI recalling fewer correct responses than other groups.
 Notably, there was an attenuated recency effect in those with mixed MCI.
@@
 However, available tests may not detect subtle impairments \cite{Kessels2015}.
-\subsubsection{Rey Auditory-Verbal learning test}
+\subsection{Rey Auditory-Verbal learning test}
 Rey's Auditory Verbal Learning Test (RAVLT+) looks a wide range of neuropsychological processess including short-term auditory-verbal memory, retention of information.
@@
 Further, they found that the conversion of MCI subjects to AD in 3-years could be predicted based on either observed or estimated RAVLT scores with an accuracy comparible to MRI-based biomarkers \cite{Moradi2017}.
+Another study by Schoenberg found that RAVLT to best distinguish patients suspected of Alzheimer's disease from the psychiatric comparison group.
 % Schoenberg - Test performance and classification statistics for the Rey Auditory Verbal Learning Test in selected clinical samples.
-Study found that RAVLT to best distguish patients suspected of Alzheimer's disease from the psychiatric comparison group.
-\subsubsection{Digit Symbol Substitution Test}
+
+\subsection{Digit Symbol Substitution Test}
 The Digit Symbol Substitution Test (DSST) involves a key consisting of the numbers 1-9 and a corresponding unique symbol.
@@
 While in part attributable to vascular brain disease, other potentially modifiable contributors may be present.
-Further studying the causes of psychomoto slowing with ageing might provide novel insights into age-related brain disorders.
-% Pascoe - The symbol-digit modalities test in MCI: Evidence from Parkinson's Disease Patients.
+Further studying the causes of psychomotor slowing with ageing might provide novel insights into age-related brain disorders.
 Pascoe compared patients with PD with Normal Cognition (PD-N) with those with PD and MCI (PD-MCI) and healthy participants.
 PD-MCI pariticipants achieved significantly lower scores than other groups in the DSST task.
-\subsubsection{Corsi Block Tapping Task}
+% Pascoe - The symbol-digit modalities test in MCI: Evidence from Parkinson's Disease Patients.
+
+\subsection{**Corsi Block Tapping Task}
 The original Corsi apparatus consists of a series of nine blocks arranged irregularly on a 9 x 11in board.
@@
 The CBT performances were significantly less satisfactory in the multi-domain MCI category than in other groups (Healthy adults, Amnesic MCI and Non-amnesic MCI with executive function impairment).  
-\subsubsection{Verbal Fluency}
+\subsection{Verbal Fluency}
 %% In this paragraph cite Henry et al, 2004 - taken from Taler and Phillips (2008) - Henry et al, 2004 is a meta analysis
@@
 A review carried out by Henry et al, found that performance in both letter fluency and category fluency was impaired in those with AD vs controls, but found a larger effect for tests of semantic fluency.
-\subsubsection{Naming Tests}
+\subsection{Naming Tests}
 Word-finding difficulty is a common symptom of AD and these deficits usually occur during the early stages of the disease progression.
@@
 
-\subsubsection{Rey-Osterrieth Complex Figure Task}
+\subsection{Rey-Osterrieth Complex Figure Task}
 % Cite Rey 1941 - L'examen psychologique dans les cas d'encephalopathie traumatique and Osterrieth 1944 - Le test de copie d'une figure complexe.
@@
 
-\subsubsection{Hayling Sentence Completion Test}
+\subsection{Hayling Sentence Completion Test}
 Inhibitory deficits are a common in all stages of dementia.
@@
 Patients with Dementia were significantly slower than controls in the initiation and inhibition conditions vs healthy controls, and slower than patients with parkinsons disease in the initiation but not the inhibition condition.
-\subsubsection{Grooved Pegboard Test}
+\subsection{Grooved Pegboard Test}
 The Grooved Pegboard Test (GPT) was originally designed to cover a variety of different psycho-motor functions including hand-eye coordination and motor speed.
@@
 They noted that higher PPT scores were associated with lower risk of incident neurodegnerative disease and noted significant associations of PPT scores with all forms of dementia and this potentially highlights a deterioration of motor function in the pre-clinical  phase of dementia.
-\subsubsection{Visual Search}
+\subsection{Visual Search}
 The visual search task requires participants to be watching a computerised display in which a target appears either by itself or surrounded by other elements which are used as a distraction.
@@
 The results from this trial also illustrates the use of a non-memory task as a means of diagnosing dementia as patients who went on to be diagnosed with dementia whilst this study was in progress had significantly poorer visual search performance.
-\subsubsection{Free Cued Selective Reminding Test}
+\subsection{Free Cued Selective Reminding Test}
 The Free Cued Selective Reminding Test (FCSRT) was borne out of the premise that by controlling the conditions of learning, a measure of memory is possible that is not confounded by normal age related changes in cognition.
@@
 
-\subsection{Types of Language Assessment}
+\section{Types of Language Assessment}
 One of the key debates when looking at how to analyse language is the type of task provided to elicit language production in participants.
@@
 \par
-\subsubsection{Picture Description Tasks}
+\subsection{Picture Description Tasks}
 One of the most commonly used tasks to measure language is the Picture Description task.
@@
 \par
-\subsubsection{Narrative description task}
+\subsection{Narrative description task}
 The story narration task is designed to study a participant's ability to describe and elaborate on a story which is depicted using a series of pictures.
@@
 \par
-\subsubsection{Interviews}
+\subsection{Interviews}
 Interviews can also be used to elicit language in a more natural way by asking questions to guide a conversation between speakers.
@@
 \par
-\subsection{How do we analyse language, issues and debates}
-\subsubsection{Single Word Language tasks vs Connected Language tasks}
+\section{How do we analyse language, issues and debates}
+\subsection{Single Word Language tasks vs Connected Language tasks}
 Part of the reason we need to pay attention to how we ask participants to generate data is understanding how we wish to analyse the data afterwards.
@@
 where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
-\subsubsection{Picture-related content thematic elements}
+\subsection{Picture-related content thematic elements}
 Several studies examined the amount of thematic elements expressed that were directly relevant to picture stimulus in picture description tasks.
@@
 Studies using these features found some differences between AD and controls, and some could differentiate between different stages of AD.
-\subsubsection{General Information Units}
+\subsection{General Information Units}
 Some studies used a more general concept of content, defining "information units" as "the smallest non redundant meaningful fact or inference," whether or not the information conveyed was specific to the context in which the conversation happened.
 Giles et al, for example, studied adults with minimal, mild or moderate AD vs controls and found that adults with AD produced fewer overall information units than controls.
-\subsubsection{Conciseness of information}
+\subsection{Conciseness of information}
 Conciseness has been defined as the number of words a speaker uses to express ideas.
@@
 \par
-\subsubsection{Efficiency}
+\subsection{Efficiency}
 Efficiency is the rate at which meaningful information is conveyed over time, calculated by dividing the total number of information units by the duration in seconds of the speech sample.
@@
 Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in dialogue in which a speaker provides responsees do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls.
-\subsubsection{Quantity - Total number of words}
+\subsection{Quantity - Total number of words}
 Several studies report that adults with moderate AD produce fewer words than controls on picture description, however other studies found no differences in total words among groups of controls and patients with MCI or AD.
@@
 \par
-\subsection{Syntax and Morphology (Language Form)}
+\section{Syntax and Morphology (Language Form)}
 Syntax can be defined as the rules that govern how words can be combined to form sentences, whilst Morphology is the system that governs the structure of words and the construction of word forms.
@@
 \par
-\subsubsection{Mean length of utterance (MLU)}
+\subsection{Mean length of utterance (MLU)}
 Murray found that MLU was not a distinguishing factor among health adults, adults with depression and adults with AD.
@@
 \par
-\subsubsection{Proportion of verbs to nouns plus verbs}
+\subsection{Proportion of verbs to nouns plus verbs}
 Kave and Levy used a verb index to capture syntactic complexity and found that adults with AD expressed the same amount of verbs to nouns plus verbs as adult controls.
 \par
-\subsubsection{Syntactic Complexity - Composite measures of MLU, syntactic errors and verbs}
+\subsection{Syntactic Complexity - Composite measures of MLU, syntactic errors and verbs}
 Ahmed et al, and Ahmed, Haigh et al found differences in syntactic complexity between adults with MCI and controls, and between MCI and moderate AD stages.
@@
 \par
-\subsubsection{Formulaic Language}
+\subsection{Formulaic Language}
 Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
@@
 \par
-\subsection{Pragmatic Language}
+\section{Pragmatic Language}
 The pragmatic language domain refers to the social rules for language for the purposes of communication including, using language to achieve goals, using information from the context to achieve these goals and using the interaction between people to initiate, maintain and terminate conversations.
-\subsubsection{Questions, turn-taking, unsure statements, egocentric comments}
+\subsection{Questions, turn-taking, unsure statements, egocentric comments}
 One study, Ripich et al, examined several pragmatic abilities among patients at different stages of AD.
@@
 \par
-\subsubsection{Coherence}
+\subsection{Coherence}
 Coherence refers to the appropriate maintenance of topic in discourse and is thematically related to the immediately preceding utterance (local coherence) and by how closely an utterance relates to the general topic at hand (global coherence).
@@
 \par
-\subsubsection{Perseveration}
+\subsection{**Perseveration}
 % EXPLAIN PERSEVERATIONS
@@
 \par
-\subsubsection{Speech intonation / prosody}
+\subsection{Speech intonation / prosody}
 There are a number of studies, that have examined "melodic line", which is a subjective measure of speech prosody defined as "the appropriate use of intonational contour, including alterations in pitch, volume and duration".
@@
 Konig et al, also used an automated feature analysis examining vocal and temporal aspects of speech among controls, and patients with MCI or AD, and reported a classification of 81\%.
-\subsubsection{Discourse Fluency}
+\subsection{Discourse Fluency}
 Verbal fluency is a term used in neuropsychological contexts generally referring to timed, word-generation tasks, while in speech-language pathology contexts, "fluency disorders" are defined as interruptions int he flow of speaking characterized by atypical rate, rhythm and repetitions in sounds, syllables, words and phrases.
@@
 \par
-\subsubsection{Speech Monitoring}
+\subsection{Speech Monitoring}
 Speech monitoring is related not only to word retrieval deficits and to pragmatic language skill but also to "anosognosia" define as the awareness of one's own deficits.
@@
 \subsection{Conclusions}
+This section has looked at a number of different attributes which researchers have cited as being important in the detection of MCI and AD.
+Whilst in some cases there is a broad consensus about a given feature in the vast majority of cases there is some uncertainty.
+Some of the difficulty lies in different approaches used particularly around the criteria for experimental groups which means that it is difficult to be able to compare studies, like for like, as the populations of the participants being studied vary in subtle or obvious ways.
+It is important for any studies moving forward to use comparable standards when it comes to their experimental groups, such that the picture can be made clearer.
 
-\subsection{State of literature into Machine Learning and Natural Language processing techniques}
+\section{**State of literature into Machine Learning and Natural Language processing techniques}
 As we can see, diagnosing dementia through analysis has a large background in terms of psychological research.
@@
 	\item Morphological/Lexical: providing the basic language elements or vocabulary, such as words, their roots and inflections.
-	\item Syntactic: for grouping and sequencing elementds within samples of language (usually sentences).
+	\item Syntactic: for grouping and sequencing elements within samples of language (usually sentences).
 	\item Semantic: for knowing the meaning of an utterance, usually defined in terms of the 'truth value' of the logical propositions that are thought to be expressed by sentences.
@@
 
+One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
+
 \subsection{Conclusions}
@@
 \subsection{Discussion}
-One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
 It is challenging to characterise Dementia as a disease as it's more syndrome.
@@
 \section{Research Aims and Objectives}
-\par
-\section{Research Aims}
-The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to time-consuming and expensive psychological tests.
-There is a lot of research into the analysis of language as a marker for MCI and Early Dementia \cite{Taler2008, Boschi2017} and there have been calls for research that explores the use of Natural Language Processing and Machine Learning techniques to aid in the analysis of language \cite{Boschi2017}.
+The aim of my research is to ultimately find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to time-consuming and expensive psychological tests.
+There is a lot of research into the analysis of language as a marker for MCI \cite{Taler2008, Boschi2017} and there have been calls for research that explores the use of Natural Language Processing and Machine Learning techniques to aid in the analysis of language \cite{Boschi2017}.
 It is this area which is the focus of my research.
 \par
+I have deliberately selected the MCI population as examining language decline at this stage has the potential to further our understanding of how MCI progresses into dementia.
 Given that sampling a person's language is relatively effortless, my proposed research will look at whether we can find markers of MCI in natural language.
-I focus particularly on MCI as this is often a precursor to AD and therefore detection of MCI in an elderly population would allow I plan to explore the use Natural Language Processing and Machine Learning techniques to assist in this process, and focus on MCI specifically.
-early diagnosis of dementia and/or mild cognitive impairment through language produced naturally rather than language that is produced as part of a formal cognitive assessment.
+I plan to explore the use Natural Language Processing and Machine Learning techniques to assist in this process.
 There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia.
-These studies have identified a number of speech and language differences between healthy controls and those with dementia.
-Some of these features have been validated in work using machine learning to classify healthy controls from those with dementia.
+As documented above, these studies have identified a number of speech and language differences between healthy controls and those with dementia.
+Some of these features have been incorporated in work using machine learning to classify healthy controls from those with dementia.
 However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'.
-While this is help to learn that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments.
+While this is helpful to learn that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments.
+Therefore we would like to see whether natural language, outside the framework of a deliberate language generating task, has the same ability to discriminate.
 \par
@@
 I will describe this shift using the context of my proposed research.
-In traditional machine learning, the transcripts from cognitive assessments are used to generate features such as idea density.
+In traditional machine learning, the transcripts from cognitive assessments are used to generate abstract features such as idea density and more concrete features such as mean length of sentence or number of unique words.
 These features are then used to classify patients.
-A company in Toronto, Canada has taken this approach and currently uses over 400 acoustic and language features to make this classification and to generate a psuedo MMSE score.
-The broad aim of my research is to explore whether the application of Natural Language Processing and Machine Learning techniques to spontaneous speech samples can improve our ability to diagnose Mild Cognitive Impairment in patients that will directly lead to Alzheimer's Disease.
-Another aim of my research is to provide some clarity as to which language features are most important to measure when looking for evidence of MCI and/or AD.
+As we have seen above, the picture with regards to what features deem important in this area is somewhat muddy.
+Machine Learning has a number of tools at it's disposal to aid this.
+The first is to statistically determine, given a model, which features contribute most highly to a given prediction.
+\par 
+The broad objectives of my research are to explore whether the application of Natural Language Processing and Machine Learning techniques to spontaneous speech samples can improve our ability to diagnose Mild Cognitive Impairment in patients that will directly lead to Alzheimer's Disease.
+Another objective of my research is to provide some clarity as to which language features are most important to measure when looking for evidence of MCI and/or AD.
 Finally, an additional outcome of my research is the development of a repeatable battery of neuropsychological tests that specifically focus on being sensitive to Mild Cognitive Impairment.
@@
 
-\section{Research Challenges}
+\section{**Research Challenges}
 \subsection{Data Collection}
@@
 
-Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
+\subsection{Co-morbidity of Language and Memory difficulties in MCI}Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
 I'm not going to control for memory problems as a potential confound, but does this weaken the research?

### 2019-01-16 Source/QualifyingReport/QualifyingReport.tex
added picture

@@
 \par
-\subsection{**Perseveration}
-% EXPLAIN PERSEVERATIONS
+\subsection{*Perseveration}
+Perseveration can be broadly defined as the repetition of a response regardless of the absence or cessation of stimulus which could have generated an appropriate response in the first place. 
 One study examined verbal preservation in the description of Norman Rockwell prints, dividing the total number of words within perseverations by total number of words in the speech sample.
@@
 \end{enumerate}
-\subsubsection{Traditional methods of Machine Learning}
+\begin{figure}[H]
+\centering
+\includegraphics[width=240px, height=150px]{images/workflow.jpg}
+\caption{Cookie Theft Picture.\label{white}}
+\end{figure}
+\subsection{Traditional methods of Machine Learning}
 % Thomas - Automatic Detection and Rating of AD through the Lexical Analysis of Spontaneous Speech
@@
 Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
-\subsubsection{The case for Deep Learning}
+\subsection{The case for Deep Learning}
+Deep Learning is an
+
+One of the main benefits of deep learning is that it does not rely on pre defined features being fed to the model, instead generating it's own features.
+This move away from relying on features solves one of the difficulties with
 

### 2019-01-17 Source/QualifyingReport/QualifyingReport.tex
submitted draft

@@
 \tableofcontents
-\newpage
 \listoftables
@@
 
-\subsection{**Corsi Block Tapping Task}
-The original Corsi apparatus consists of a series of nine blocks arranged irregularly on a 9 x 11in board.
-The blocks are tapped by the adminstrator in a randomized sequence of increasing length.
-Immediately after each sequence is demonstrated, the task is to mirror the sequence back to the adminstrator with the sequences increasing in length until a mistake is made.
-% Perrochon - How specific are cognitive deficits in short-term visuospatial memory in mild cognitive impairment patients?
-The CBT performances were significantly less satisfactory in the multi-domain MCI category than in other groups (Healthy adults, Amnesic MCI and Non-amnesic MCI with executive function impairment).  
 \subsection{Verbal Fluency}
@@
 This has been reduced subsequently to two thirty item versions and four fifteen item versions, which correlate significantly with the original sixty item version and the benefit of the shorter versions of the test is that it facilitates testing of individuals with AD who may suffer from fatigue or limited attention span.
+\par 
 % Willers - Subclinical naming errors in MCI - A semantic deficit.
-Twenty aMCI patients, twenty AD and 21 controls matched by age, sex and education level were evaluated.
-AD patients obtained significantly lower total scores on the BNT than aMCI patients and controls.
+Willers et al studied Twenty aMCI patients, twenty AD and 21 controls matched by age, sex and education level.
+They found AD patients obtained significantly lower total scores on the BNT than aMCI patients and controls.
 aMCI patients did not obtain significant differences in total scores but showed significantly higher semantic errors compared to controls.
 Semantic processing is impaired during confrontation naming in aMCI.
+\par 
 % Vadikolias - Mild Cognitive Impairment: effect of education on the verbal and nonverbal tasks performance decline.
-Higher educational attainment in aMCI subjects were correlated with better performance in verbal and non-verbal tasks during repeated examinations over 1-year.
-Subjects with a lowere level of education performed worse than patients with a high level of education who presented a more stable clinical score.
+Vadikolias investigated the impact of education on naming tests and found that higher educational attainment in aMCI subjects were correlated with better performance in verbal and non-verbal tasks during repeated examinations over 1-year.
+Subjects with a lower level of education performed worse than patients with a high level of education who presented a more stable clinical score.
 The explanation for this is the idea of a 'cognitive reserve' in participants with a higher education such that this provides a buffer that, while not preventing the physiological symptoms of AD, can potentially delay the clinical onset of cognitive symptoms that characterise AD.
 This theory is supported by Snowdon who found a relationship between early life linguistic ability and the density of neurofibrillary tangles in his nun study \cite{Snowdon1996}.
-
 Whilst these studies and tests provide evidence that there are word finding difficulties in those with MCI, on their own they do not provide sufficient ability to diagnose MCI or provide an prognosis of disease progression.
@@
 % Salvadori - Qualitative Evaluation of the Immediate Copy of the Rey-Osterrieth Complex Figure: Comparison Between Vascular and Degenerative MCI patients.
-Patients with vascular MCI had a worse performance in the immediate copy of the ROCF compared to individuals with degenerative MCI, despite their significant impairment in terms of general cognitive status and visual memory.
+Salvadori found that patients with vascular MCI had a worse performance in the immediate copy of the ROCF compared to individuals with degenerative MCI, despite their significant impairment in terms of general cognitive status and visual memory.
 Evidence shows that patients with disorders that possibly involve attention and executive functions are characterized by a more disorganised approach when copying the ROCF compared to controls.
+One of the difficulties with the presentation of this task in a repeated battery will be the fact that it turns from an incidental memory task (it is a memory task but the participant is not forewarned that they will be required to memorise the picture) into an intentional memory task (given the previous exposure to the task, it would be expected that a participant would pre-empt the delayed recall portion of the test and spend more time attempting to memorise the details).
 % Teng - Persistence of Neuropsychological Testing Deficits in Mild Cognitive Impairment.
-
-One of the difficulties with the presentation of this task in a repeated battery will be the fact that it turns from an incidental memory task (it is a memory task but the participant is not forewarned that they will be required to memorise the picture) into an intentional memory task (given the previous exposure to the task, it would be expected that a participant would pre-empt the delayed recall portion of the test and spend more time attempting to memorise the details).
 
@@
 % Abnormal Visual Search in MCI and Alzheimer's Disease. Tales.A 2005
-In research, there have been shown to be deficits in both AD and MCI populations.
-Whilst these deficits were not as apparent in MCI populations, there remained a difference that could be used to differentiate healthy controls from those with MCI.
-% Tales - Visual Search in MCI: A Longitudinal Study
+Research has shown there to be deficits in both AD and MCI populations in this task.
+Whilst these deficits were not as apparent in MCI populations, there remained a significant enough difference that could be used to differentiate healthy controls from those with MCI.
 In a comparison of visual search performance between those with aMCI and healthy controls, Tales noted significantly poorer performance in the aMCI group.
@@
 Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cognitive decline due to a cognitive impairment.
-/par
-FCSRT-IR (The instant recall version of the Free Cued Selective Reminding Test) performance has beeen shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the authors to define prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT-IR performance (Sarazin et al, 2007
-/par
+FCSRT-IR (The instant recall version of the Free Cued Selective Reminding Test) performance has beeen shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the authors to define prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT-IR performance (Sarazin et al, 2007.
 The sensitivity and specificity of the different test forms in correctly classifying patients were compared to each other using a previously established FCSRT-IR cut off score.
@@
 I have looked at a number of different cognitive tests and a battery of tests that aim to have high diagnostic and prognostic capabilities in the MCI population.
-However, particularly with the RBANS, there lacks sufficient sensitivity in differentiating those with MCI from healthy elderly individuals such that this tool could be used with a level of confidence in the results.
-In regard to the tests, a common theme is a lack of studies and therefore evidence into the utility of these tests with this particular population.
+However, particularly with the RBANS, there lacks sufficient sensitivity in differentiating those with MCI from healthy elderly individuals such that this tool could be used with a level of confidence in a clinical setting.
+In regard to the tests, a common theme is a lack of studies and therefore evidence into the utility of these tests with an MCI population.
 However if, as researcher, we aim to investigate this population further then a benchmark battery of tests which is sensitive enough in both diagnostic and/or prognostic utility should be a goal.
 \par 
-A criticism of the current literature is the lack of consistency with regard to the experimental groups.
+Another criticism of the current literature is the lack of consistency with regard to the experimental groups.
 For example, some studies focus on differentiating between an MCI group, a AD group and a healthy controls group whereas other studies may further subdivide the MCI category according to a number of factors such as Amnesic or Non-Amnesic MCI, Mixed MCI (sometimes called Multi-domain MCI in the literature).
@@
 Formally, idea density is defined as the average number of propositions per sentence \cite{Kintsch1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}.
-Some examples of semantic content measures are Type to Token Ratio, Brunet's Index and Honore's Statistic  which are used to measure the lexical diversity of a given piece of text and/or utterance.
+Some examples of semantic content measures are Type to Token Ratio, Brunet's Index and Honore's Statistic (described below) which are used to measure the lexical diversity of a given piece of text and/or utterance.
 This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2005}.
@@
 \par
-\subsection{*Perseveration}
+\subsection{Perseveration}
 Perseveration can be broadly defined as the repetition of a response regardless of the absence or cessation of stimulus which could have generated an appropriate response in the first place.
+A typical example could be the idea of conversation moving from introductions to a more general conversation, someone who has difficulties with perseveration will struggle with the move between social contexts and repeat language that would be more used in an introduction context.
+\par 
 One study examined verbal preservation in the description of Norman Rockwell prints, dividing the total number of words within perseverations by total number of words in the speech sample.
@@
 Across all tasks, the AD group produced significantly more perseverations than controls; however, on the picture description task alone, there were no significant differences between the two groups.
-The authors theorised that this was because picture description was an easier taks because there was the visual stimulus, similar to the argument made by Bschor et al.
+The authors theorised that this was because picture description was an easier task because there was the visual stimulus, similar to the argument made by Bschor et al \cite{Bschor2001}.
 \par
@@
 Generally speaking, Natural Language Processing (NLP) systems mimic the semiotic perspective on language.
-That is to say that we can subdivide NLP processes into different components.
-This broadly fits with the literature review conducted by Emery \cite{Emery2000}.
+That is to say that we can subdivide NLP processes into different components (see Figure 2.3) and model broadly fits with the literature review conducted by Emery \cite{Emery2000}.
 \begin{enumerate}
@@
 \includegraphics[width=240px, height=150px]{images/workflow.jpg}
-\caption{Cookie Theft Picture.\label{white}}
+\caption{How Natural Language Processing tasks are subdivided.\label{white}}
 \end{figure}
@@
 \subsection{The case for Deep Learning}
-Deep Learning is an
+% Deep Learning is an NOT
 
-One of the main benefits of deep learning is that it does not rely on pre defined features being fed to the model, instead generating it's own features.
+% One of the main benefits of deep learning is that it does not rely on pre defined features being fed to the model, instead generating it's own features.
 This move away from relying on features solves one of the difficulties with
 
-One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
+% One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
 
@@
 \subsection{Discussion}
-It is challenging to characterise Dementia as a disease as it's more syndrome.
-The clinical features that combine to meet the diagnostic criteria for Dementia or it's variants are continuous in nature and also impacted by other variables.
-For example, cognitive performance is affected by previous education and ability to live independently is impacted by a person's physical health.
-Also, the individual who may develop symptoms that indicate the onset of dementia needs the insight to recognise them, or family members will need to do so and also deduce that there is value to admitting there is a problem and seeking help before the deterioration is too great.
-\par
-To detect dementia early requires an understanding of how this presents early in it's progression, but as described above there are a number of potential ways in which individuals with dementia could be impacted but no consistent list of criteria has been produced.
-Even for AD, the most prevalent form of dementia, clinical diagnosis can only be confirmed after death through autopsy.
-In addition, population studies have shown that people over the age of 80 show similar changes in the brain in those that do not have dementia.
-Therefore, it's difficult to say with certainty that any symptoms can definitely cause dementia despite the fact that these symptoms are commonly present in those that have dementia.
-\par
-\subsection{Conclusions and Future Work}
-It appears that there are a number of approaches that have been shown to be effective in various experimental settings.
-The fact that there are a numerous array of different perspectives to tackling this question, and even within these perspectives there are a range of various methodologies leaves the field with a sense of confusion as to the most effective way to solving the problem.
-\par
 Current state-of-the-art diagnostic measures of AD are invasive (CSF analysis), expensive (neuroimaging), and time-consuming (neuropsychological assessment).
-Furthermore, these measures are limited to specialty clinics and thus have limited accessibility as high-throughput, or frontline, screening and diagnostic tools for AD.
-More importantly, nonspecialists are often inaccurate at identifying early AD and MCI [5].
-Thus, there is an increasing need for additional noninvasive and/or cost-effective tools, allowing frontline identification of subjects in the preclinical or early clinical stages of AD who could be suitable for monitoring in specialty clinics and for early treatment.
+Furthermore, these measures are limited to speciality clinics and thus have limited accessibility as frontline screening and diagnostic tools for AD.
+More importantly, nonspecialists are often inaccurate at identifying early AD and MCI.
+Thus, there is an increasing need for additional noninvasive and/or cost-effective tools, allowing effective frontline identification of subjects in the preclinical or early clinical stages of AD who could be suitable for monitoring in speciality clinics and for early treatment.
 Implementation of effective screening instruments will allow diagnosis earlier in the course of dementia, even at the point when memory function is still essentially within the normal range.
@@
 \par 
-Identification of language impairment is key idea in the diagnosis of AD and early diagnosis can alter the prognosis and change the management of this degenerative disorder as well as provide new opportunities to study the progression of the disease.
-Given the burden on the diagnosis of AD on professionals there has been a call to use technology to potentially ease this burden \cite{Boschi2017}.
-It has already been shown that analysis of speech and language has shown markers that pre-date the official diagnosis of dementia \cite{Snowdon1996, Berisha2015}.
-A significant amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
+However the literature has identified a number of challenges when approaching this problem.
+Firstly, the clinical features that combine to meet the diagnostic criteria for Dementia or it's variants are continuous in nature and heterogenous between patients and are also impacted by other variables.
+For example as shown above cognitive performance is affected in part by a patients educational attainment and a patients ability to live independently is impacted by a their physical health as much as their cognitive health.
+The challenge therefore is to find features that are minimally impacted by other factors, or that can be controlled for by a experimental design such as a matched pairs design to control for educational attainment.
 \par 
+Another challenge lies with the recruitment of suitable individuals who may notice a decline in cognition to the point where we might classify them as having MCI, but these individuals deduce that there is little to no value to admitting there is a problem and seeking help whilst their symptoms are 'manageable'.
+\par
+Finally there is a large amount of variability in the presentations of those with MCI and early dementia, and this is compounded by an similar amount of variability in the criteria researchers have used for experimental groups and the approaches researchers have used to tackle this problem.
+This had led to a confused literature.
+Recently a call for research that has consistent inclusion / exclusion criteria has been made along with some proposed definitions of MCI and it's subgroups \cite{Petersen2014}.
+Researchers have identified the analysis of language impairment as an area of promise to explore in the diagnosis of MCI and early AD and recent developments in natural language processing and machine learning techniques have the potential to assist in this research.
+Indeed, given the increased burden on the diagnosis of MCI and AD on professionals there has been a call to use technology to potentially ease this burden \cite{Boschi2017}.
+A small but growing amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
+
+\subsubsection{Future Work}
 One of the areas for research to study is a careful examination of the features that are being used to measure language deterioration.
@@
 \par
-\subsubsection{Future Work}
-Future research should also be directed towards developing non-intrusive ways of detecting subtle changes in language based in the home, such that any deterioration that could indicate the presence of MCI or AD and be flagged up early in the progression of any potential cogntive impairment for further review via referral.
-Machine learning approaches seem to be the most logical approach for achieving this aim.
-Further, despite the quality of datasets being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which we could claim are limited in scope.
+Teng et al suggests that work should focus on the MCI population and concentrate on developing a consensus neuropsychological battery that could yield predictable rates of progression to AD.
+This, in conjunction with the development of a model of language, sensitive enough to detect subtle deterioration in language use to act as an additional cognitive marker to aid diagnosis could potentially move some way to providing this.
 \par 
-The recent resurgence in the use of neural networks and deep learning is because of it's significantly improved performance on many problems and it's ability to scale from small to large datasets.
-Another key benefit of deep learning is it's ability to automate the process of feature engineering.
-Feature engineering, with this particular problem is an interesting debate as there are currently numerous ways in which to try to generate features as I have described above.
+Future research should also be directed towards developing non-intrusive ways of detecting subtle changes in natural language such that any perceived deterioration that could indicate the presence of MCI or AD could be flagged up early.
+Machine learning approaches seem to be the most logical approach for achieving this aim as language could be collected in non-intrusive ways and passed to a machine learning algorithm for preliminary classification.
+Further, despite the excellent quality of datasets being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which we could claim are limited in scope.
+Finally, the recent resurgence in the use of neural networks and deep learning could provide the answer to the confused literature in terms of features.
+A key benefit of deep learning is it's ability to automate the process of feature engineering.
+So there is an opportunity to explore the use of deep learning, to not only develop new features but also
 \par
 This area of research is extremely promising in its early results and the impact of successful research would be life changing for both individuals and the health of the worlds aging population in general.
-\par 
-As described above, there is a large gap in the knowledge of how, and it what circumstances does MCI progress to AD as well as how best to assess MCI over time and at this time, to the authors knowledge, there is no consensus in how to approach these two problems.
-Teng et al suggests that work should focus on the MCI population and concentrate on developing a consensus neuropsychological battery that could yield predictable rates of progression to AD.
-This, and with the development of a model of language, sensitive enough to detect subtle deterioration in language use to act as an additional cognitive marker to aid diagnosis could potentially move some way to providing this.
-\par 
-% McGuiness - Predicting conversion to dementia in a memory clinic: A standard clinical approach compared with an empirically defined clustering method (latent profile analysis) for mild cognitive impairment subtyping.
 
@@
 While this is helpful to learn that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments.
-Therefore we would like to see whether natural language, outside the framework of a deliberate language generating task, has the same ability to discriminate.
+Therefore we would like to see whether natural language generated outside the framework of a deliberate language generating task, has the same ability to discriminate.
 \par
@@
 The first is to statistically determine, given a model, which features contribute most highly to a given prediction.
+The second is to use a deep learning model which has the ability to automate the process of feature engineering.
 \par 

### 2019-01-20 Source/QualifyingReport/QualifyingReport.tex
QR edits 17-1-19

@@
 As part of this I will identify target conferences where I could present my work.
-Finally, I look at potential risks and ethical concerns that may occur as part of my research.
+Finally, I look at potential risks and ethical concerns that could conceivably occur.
 \section{Research Aims and Objectives}
-The aim of my research is to ultimately find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to time-consuming and expensive psychological tests.
+The aim of my research is to ultimately to identify less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to time-consuming and expensive psychological tests.
 There is a lot of research into the analysis of language as a marker for MCI \cite{Taler2008, Boschi2017} and there have been calls for research that explores the use of Natural Language Processing and Machine Learning techniques to aid in the analysis of language \cite{Boschi2017}.
-It is this area which is the focus of my research.
+This will be the focus of my research.
 \par
@@
 However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'.
-While this is helpful to learn that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments.
-Therefore we would like to see whether natural language generated outside the framework of a deliberate language generating task, has the same ability to discriminate.
+While this is helpful to establish that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments.
+Therefore we would like to clarify whether natural language generated outside the framework of a deliberate language generating task, has the same ability to discriminate.
 \par
@@
 
-\subsection{Co-morbidity of Language and Memory difficulties in MCI}Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
+\subsection{Co-morbidity of Language and Memory difficulties in MCI}Concerns: Language and Memory are naturally intertwined and it would be difficult to test one without some reliance on the other.
 I'm not going to control for memory problems as a potential confound, but does this weaken the research?
@@
 \par 
-We compiled a series of tests which look a number of cognitive functions which have previously demonstrated a level of effectiveness at discriminating an MCI population from a healthy population (see literature review for a comprehensive list, explanation and rationale for each test).
-In addition to these tests, we have incorporated two language tasks in which a participants ability to generate spontaneous speech that we can then use to further analyse.
+We compiled a series of tests which look at a number of cognitive functions which have previously demonstrated a level of effectiveness in discriminating an MCI population from a healthy population (see literature review for a comprehensive list, explanation and rationale for each test).
+In addition to these tests, we have incorporated two language tasks in which a participant's ability to generate spontaneous speech that we can use to further analyse.
 As a planned future study is longitudinal, it is important that this battery of tests is repeatable but does not have significant practice effects that may interfere with our ability to track cognitive decline.
@@
 \subsection{Task 3 - Exploration of language with the DementiaBank dataset and Three Authors dataset.}
-The aim of this task is to replicate and extend the work of Fraser and Orimaye who both looked at the DementiaBank dataset and set out to use traditional machine learning techniques to see whether they could categorise people into dementia or healthy categories using the samples given.
+The aim of this task is to replicate and extend the work of Fraser and Orimaye who both used the DementiaBank dataset and set out to use traditional machine learning techniques to see whether they could categorize people into dementia or healthy categories using the samples given.
 We also aim to replicate and extend the work of Le et al who used electronic versions of the novels of three authors.
@@
 Whilst we have divided participants into experimental groups we aim to investigate each participant individually.
-Because there is a large amount of heterogeneity in this population, we will initially be analysing each participants results in isolation with the participant acting as their own control.
+Because there is a large amount of heterogeneity in this population, we will initially be analysing each participant's results in isolation with the participant acting as their own control.
 However, we will also be exploring the use of traditional machine learning and/or deep learning across cohorts to investigate whether each cohort has their own distinct 'language profile'.
 \par 
-The final output of this task will be a paper discussing the results of our longitudinal study from an language analysis perspective.
+The final output of this task will be a paper discussing the results of our longitudinal study from a language analysis perspective.
 The target conference for this work will be the UK Alzheimer's Research conference in March 2021 with an expected submission date of December 2020, the Alzheimer's Association International Conference in July 2021 with an expected submission date of January 2021 and the Association of Computational Linguistics Conference (ACL) in July 2021 with an expected submission date of March 2021.

### 2019-01-25 Source/QualifyingReport/QualifyingReport.tex
25-01-2019 am edits

@@
 \title{Qualifying Report}
-\date{2019-01-07}
+\date{2019-01-21}
 \author{Jomar Alcantara}
@@
 	\item Immediate Memory 
+	\begin{enumerate}
+		\item{List Learning}
+		\item{Story Memory}
+	\end{enumerate}
 	\item Visuospatial / Constructional
+	\begin{enumerate}
+		\item{Figure Copy}
+		\item{Line Orientation}
+	\end{enumerate}
 	\item Language
+	\begin{enumerate}
+		\item{Picture Naming}
+		\item{Semantic Fluency}
+	\end{enumerate}
 	\item Attention
+	\begin{enumerate}
+		\item{Digit Span}
+		\item{Digit Coding}
+	\end{enumerate}
 	\item Delayed Memory
+	\begin{enumerate}
+		\item{List Recall}
+		\item{List Recognitions}
+		\item{Story Memory}
+		\item{Figure Recall}
+	\end{enumerate}
 \end{enumerate}
@@
 The sensitivity and specificity of the different test forms in correctly classifying patients were compared to each other using a previously established FCSRT-IR cut off score.
-\subsubsection{Conclusion}
+\subsubsection{Conclusions}
 I have looked at a number of different cognitive tests and a battery of tests that aim to have high diagnostic and prognostic capabilities in the MCI population.
@@
 An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Kaplan2010}.
-In this task participants are asked to describe a picture presented to them in as much detail as possible.
-The picture itself depicts a familiar domestic scene and would not require participants to use any vocabulary beyond that learned in childhood.
-It was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}
+The Cookie Theft picture (pictured below) depicts a scene of a home typical of the period of time when it was created and would generally not require participants to use any complicated vocabulary to describe.
+In this task participants are asked to describe the picture presented to them in as much detail as possible.
+This task was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}
 \begin{figure}[H]
@@
 \par
-The picture description task does a fine job of eliciting descriptive language, however because of the limited content has limited use.
-The task in itself just a descriptive task, and therefore elicits a certain type of language.
+The picture description task does a fine job of eliciting descriptive language however because of the specific content the language produce could be considered quite limited.
 There is some disagreement as to the benefits of this using this methodology.
@@
 \par
-In terms of Machine Learning research in this area, a number of researchers have used transcripts based on picture description tasks \cite{Zimmerer2016, Orimaye2017, Mueller2018a, Fraser2015} and have successfully extracted linguistic features that could differentiate between AD and controls.
-\par
 \subsection{Narrative description task}
 The story narration task is designed to study a participant's ability to describe and elaborate on a story which is depicted using a series of pictures.
-The stories depicted are usually based on children's books or famous stories such as Cinderella.
-\cite{Fraser2014} This task requires ordering the story in a structured and coherent framework.
-It also requires comprehension and understanding of the stories characters and the events depicted, as well as an awareness of a character's goals and internal responses to given events.
-This task is particularly useful, as the procedure reduces the demands on memory and is therefore able rule out memory as a confounding variable for any results observed.
-As noted above, Ash \cite{Ash2012} felt that this task was interchangable with the Picture Description task and because this task requires elaboration rather than simple description, is a sturdier test of lexical and semantic abilities as well as syntactic complexity.
-\cite{DeLira2011}
+The stories depicted are usually based on children's books or famous stories with the Cinderella being the one most typically used \cite{Fraser2014}.
+This task requires ordering the story in a structured and coherent framework.
+It also requires comprehension and understanding of the stories characters and the events depicted, as well as an awareness of a character's actions, motivations and internal reactions to given events.
+This task is particularly useful as the procedure reduces the demands on memory, due to the participant being able to access the picture book during the description and is therefore able rule out memory as a confounding variable for any results observed.
+As noted above, Ash \cite{Ash2012} felt that this task was interchangable with the Picture Description task.
+However, other research felt that this was a studier test of lexical and semantic abilities as well as syntactic complexity because this task requires interpretation and elaboration in additional to a simple description \cite{DeLira2011}.
 \par
@@
 \par
-The analysis of interviews can be difficult to analyse as both the content can vary even between participants.
+The analysis of interviews can be difficult to analyse as both the content can vary even between participants, although it can be argued that content should not affect the type of language being generated unless it is narrow topic or the participant is constrained in how they answer a given question.
 It is also difficult to measure as there are no pre-defined task goals in comparison to the other two methods.
@@
 \par
+\subsection{Conclusions}
+One can view the different types of tasks above as a continuum where picture tasks represent a much more controllable task with a lot of supporting research but which generates a much more constrained set of language that is atypical of normal speech in terms of the cognitive functions used.
+
 \section{How do we analyse language, issues and debates}
@@
 \par
-Single word tasks such as the Boston Naming Test and other such standardized language tests generally target a participants verbal fluency where this is defined as the ability to form and express words in accordance with certain criteria.
-For example, a common test of verbal fluency would be to ask a participant to name as many words in a given category in a limited time frame.
-These categories can be objects such as animals or sports, or can be words beginning with a specific letter.
+Single word tasks such as the Boston Naming Test and other such standardized language tests generally target a participants verbal fluency where this is defined as the ability to form and express words in accordance with certain criteria (see above for a full description of verbal fluency as task).
 There are a number of benefits to using single word tasks.
@@
 They are able to be relatively constrained in the language available to be used, for example in the picture description task, or they can be unconstrained in the form of interviews.
-\par
-\subsubsection{Semantics vs Pragmatics}
+% ADD MORE!
+\par
+\subsection{Semantics vs Pragmatics}
 When navigating the English language it is necessary to distinguish between what a sentence says in both semantic and pragmatic terms.
@@
 Pragmatic meaning refers looks at the same sentence in terms of words and grammar but takes into account the situation or context in which these words are spoken.
-Emery in her literature review looks at all levels of language tasks except for pragmatics, however ---FINISH THIS!
+Emery in her literature review looks at all levels of language tasks except for pragmatics, however this is an area that should not be ignored.
 \par
@@
 
-\section{**State of literature into Machine Learning and Natural Language processing techniques}
+\section{State of literature into Machine Learning and Natural Language processing techniques}
 As we can see, diagnosing dementia through analysis has a large background in terms of psychological research.
-However, a lot of more current research has called for the use of machine learning as a way of assisting in the process of diagnosis \cite{Boschi2017}.
+However, increasingly current research has called for the use of machine learning as a way of assisting in the process of diagnosis \cite{Boschi2017}.
+This section will look at what areas of machine learning could potentially be used as tools to help in this domain as well as any research that has applied machine learning to this problem.
+\subsection{Natural Language Processing}
+Natural Language Processing as a topic can be defined as the intersection between Machine Learning and Linguistics and looks at
+
 Generally speaking, Natural Language Processing (NLP) systems mimic the semiotic perspective on language.
@@
 \end{figure}
+
+\subsubsection{Natural Language Toolkit}
+% CITE Bird, Steven, Edward Loper and Ewan Klein (2009). Natural Language Processing with Python.  O'Reilly Media Inc.
+The Natural Language Toolkit (NLTK) was originally designed as part of a computational linguistics course
+
+\begin{center}
+	\begin{tabular}{ | c | c | c | p{1cm} |}
+		\hline
+		Language processing task & Functionality \\ \hline
+		Accessing corpora & standardized interfaces to corpora and lexicons \\ \hline
+		String processing & tokenizers, sentence tokenizers, stemmers \\ \hline
+		Collocation discovery & t-test, chi-squared, point-wise mutual information \\ \hline
+		Part-of-Speech tagging & n-gram, backoff, Brill, HMM, TnT \\ \hline
+		Machine Learning & decision tree, maximum entropy, naive Bayes, EM, k-means \\ \hline
+		Chunking & regular expression, n-gram, named-entity \\ \hline
+		Parsing & chart, feature-based, unification, probabilistic, dependency \\ \hline
+		Semantic interpretation & lambda calculus, first-order logic, model checking \\ \hline
+		Evaluation metrics & precision, recall, agreement coefficients \\ \hline
+		Probability and estimation & frequency distributions, smoothed probability distributions \\ \hline
+		Applications & Graphical concordance, parsers, WordNet browser, chatbots \\ \hline
+		Linguistic Framework & manipulate data in SIL Toolbox format \\ \hline
+	\end{tabular}
+\end{center}
+caption{\label{tab:table-name}Search Terms and Number of Results.}
+
+\subsubsection{The Stanford Parser}
+
+\subsubsection{Linguistic Inquiry and Word Count}
+
 \subsection{Traditional methods of Machine Learning}
+
+In terms of Machine Learning research in this area, a number of researchers have used transcripts based on picture description tasks \cite{Zimmerer2016, Orimaye2017, Mueller2018a, Fraser2015} and have successfully extracted linguistic features that could differentiate between AD and controls.
+\par
 % Thomas - Automatic Detection and Rating of AD through the Lexical Analysis of Spontaneous Speech
@@
 \subsection{Conclusions}
-
-\subsection{Discussion}
+We can see that both Natural Language Processing and Machine Learning techniques have a lot to offer, Indeed there has been a lot of research which have used pre-existing datasets to explore this area with promising results.
+To date, there has not been much in the way of research in which these techniques are applied to newly created samples of language.
+
+\section{Discussion}
 Current state-of-the-art diagnostic measures of AD are invasive (CSF analysis), expensive (neuroimaging), and time-consuming (neuropsychological assessment).
@@
 
-\subsubsection{Future Work}
+\subsection{Future Work}
 One of the areas for research to study is a careful examination of the features that are being used to measure language deterioration.
@@
 	\item \textbf{Hypothesis} The new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures.
-	\item \textbf{Hypothesis} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, those with Alzheimer's Disease and decline in those with MCI.
+	\item \textbf{Hypothesis} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, those with Alzheimers Disease and decline in those with MCI.
 \end{itemize}
-
-\section{**Research Challenges}
-\subsection{Data Collection}
-
-\subsection{Commonality of MCI}
-
-\subsection{Sparsity of Language}
-
-\subsection{Co-morbidity of Language and Memory difficulties in MCI}Concerns: Language and Memory are naturally intertwined and it would be difficult to test one without some reliance on the other.
-I'm not going to control for memory problems as a potential confound, but does this weaken the research?
-How do I defend this?
-\newline
 
@@
 The writing of the thesis, including additional over-running data analysis is planned for the last six months of the project (Oct '20 to Mar '21).
+
+\section{Research Challenges}
+\subsection{Data Collection}
+There are a few challenges that I have identified with the process of data collection.
+As with all longitudinal studies, the data collected is only valid if there are at least two sets of data points for a given participant.
+Given the population I am exploring, it is expected that there is some level of drop out rate over the course of the year and this has been accounted for.
+Nevertheless, a higher than average drop out rate would weaken the strength of the study, especially if those who dropped out came from the experimental group.
+
+\subsection{Commonality of MCI}
+% Roberts - Higher risk of progression to dementia in mild cognitive impairment cases who revert to normal.
+Whilst research has consistently shown that those with a diagnosis of MCI tend to have a significantly higher risk of progression to dementia (CITE Roberts).
+The progression at any given time is low and variable.
+% Farias - Progression of Mild Cognitive Impairment to Dementia in Clinic- vs Community Based Cohorts
+Annual conversion rates vary across longitudinal studies conducted in this population, but researchers have shown that there is a higher conversion rate in clinical samples of this population than in community samples.
+In my proposed longitudinal study, I aim to collect a clinical MCI population, but depending on the availability of such samples I may have to collect from the community.
+If this is the case then, the prevalence of those with MCI who have language deficits within my participants may be low.
+
+\subsection{Sparsity of Language}
+I have tried to estimate the amount of language I will collect by conducting a pilot study on healthy controls.
+However, until the first sets of data collection with the MCI population, I will not be able to tell if this is a problem.
+I fully expect there to be little difference in the amount of language collected in the MCI population, but this remains a possibility which I will need to adjust for at the 2nd and 3rd data collection phases.
 

### 2019-01-25 Source/QualifyingReport/QualifyingReport.tex
fixed tables

@@
 We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English.
-\begin{sidewaystable}
+\begin{table}
 	\begin{center}
-	\begin{tabular}{ | c | c | c | p{1cm} |}
+	\begin{tabular}{ | p{4cm} | p{1cm} | p{6cm} | p{1cm} |}
 		\hline
-		Database & Number of Results & Search Terms  \\ \hline
-		ProQuest(PsychArticles) & 1484 Results & Language AND Decline AND Dementia \\ \hline
-		ProQuest(PsychArticles) & 486 Results  & Language AND Decline AND Dementia AND Speech \\ \hline
-		ProQuest(PsychArticles) & 159 Results & Machine Learning AND Dementia AND Language \\ \hline
-		Web of Science & 1207 Results  & Language AND Decline AND Dementia   \\ \hline
-		Web of Science & 151 Results  & Language AND Decline AND Dementia AND Speech  \\ \hline
-		Web of Science & 34 Results & Machine Learning AND Dementia AND Language \\ \hline
-		Scopus & 791 Results & Language AND Decline AND Dementia  \\ \hline
-		Scopus & 91 Results & Language AND Decline AND Dementia AND Speech   \\ \hline
-		Scopus & 29 Results & Machine Learning AND Dementia AND Langauge \\ \hline
-		Scopus & 1292 Results & "Language Deficits" AND Dementia \\ \hline
+		Database & Results & Search Terms  \\ \hline
+		ProQuest(PsychArticles) & 1484 & Language AND Decline AND Dementia \\ \hline
+		ProQuest(PsychArticles) & 486  & Language AND Decline AND Dementia AND Speech \\ \hline
+		ProQuest(PsychArticles) & 159 & Machine Learning AND Dementia AND Language \\ \hline
+		Web of Science & 1207  & Language AND Decline AND Dementia   \\ \hline
+		Web of Science & 151  & Language AND Decline AND Dementia AND Speech  \\ \hline
+		Web of Science & 34 & Machine Learning AND Dementia AND Language \\ \hline
+		Scopus & 791 & Language AND Decline AND Dementia  \\ \hline
+		Scopus & 91 & Language AND Decline AND Dementia AND Speech   \\ \hline
+		Scopus & 29 & Machine Learning AND Dementia AND Langauge \\ \hline
+		Scopus & 1292 & "Language Deficits" AND Dementia \\ \hline
 	\end{tabular}
@@
 	\caption{\label{tab:table-name}Search Terms and Number of Results.}
-\end{sidewaystable}
+\end{table}
 % -- END OF Aims and Methodology
@@
 
-\begin{center}
-	\begin{tabular}{ | c | c | c | p{1cm} |}
+\begin{table}
+	\begin{tabular}{ | p{4cm} | p{7cm} | p{1cm} |}
 		\hline
-		Language processing task & Functionality \\ \hline
+		\textbf{Language processing task} & \textbf{Functionality} \\ \hline
 		Accessing corpora & standardized interfaces to corpora and lexicons \\ \hline
@@
 	\end{tabular}
-\end{center}
-caption{\label{tab:table-name}Search Terms and Number of Results.}
+		\caption{\label{tab:table-name}NLTK tasks and functionality}
+\end{table}
 

### 2019-01-25 Source/QualifyingReport/QualifyingReport.tex
25-01-2019 lunch edits

@@
 % CITE Bird, Steven, Edward Loper and Ewan Klein (2009). Natural Language Processing with Python.  O'Reilly Media Inc.
-The Natural Language Toolkit (NLTK) was originally designed as part of a computational linguistics course
+The Natural Language Toolkit (NLTK) was originally designed as part of a computational linguistics course but has evolved to become an open-source framework which includes methods and modules for a wide array of Natural Language tasks (see Table 2.2 for a full list).
 
@@
 
-\subsubsection{The Stanford Parser}
+\subsubsection{CoreNLP}
+The stanford parser is
+
+Stanford CoreNLP provides a set of human language technology tools.
+It can give the base forms of words, their parts of speech, whether they are names of companies, people, etc., normalize dates, times, and numeric quantities, mark up the structure of sentences in terms of phrases and syntactic dependencies, indicate which noun phrases refer to the same entities, indicate sentiment, extract particular or open-class relations between entity mentions, get the quotes people said, etc.
+
+Stanford CoreNLP’s goal is to make it very easy to apply a bunch of linguistic analysis tools to a piece of text.
+A tool pipeline can be run on a piece of plain text with just two lines of code.
+CoreNLP is designed to be highly flexible and extensible.
+With a single option you can change which tools should be enabled and disabled.
+Stanford CoreNLP integrates many of Stanford’s NLP tools, including the part-of-speech (POS) tagger, the named entity recognizer (NER), the parser, the coreference resolution system, sentiment analysis, bootstrapped pattern learning, and the open information extraction tools.
+Moreover, an annotator pipeline can include additional custom or third-party annotators.
+CoreNLP’s analyses provide the foundational building blocks for higher-level and domain-specific text understanding applications.
 
 \subsubsection{Linguistic Inquiry and Word Count}
+The premise behind the Linguistic Inquiry and Word Count (LIWC) is the words we speak inform an observer of the state of mind of an individual.
+To this end, the creators have constructed a framework which uses a dictionary of words which are pre-categorised.
+% Al-Mosaiwi - In an Absolute State: Elevated Use of Absolutist Words Is a Marker Specific to Anxiety, Depression, and Suicidal Ideation 
+A typical example of this might be someone with depression who may typically be more inwardly focused and therefore talking in the first person is common. 
+% Sonnenschein - Linguistic analysis of patients with mood and anxiety disorders during cognitive behavioral therapy
+The 2015 version of the LIWC has over 90 different categories and some supporting research suggests that it is effective at identifying themes and associating these themes with mental health difficulties such as depression and anxiety.
+\par
+A framework like this could be used as another tool to explore themes within a persons speech.
 

### 2019-01-27 Source/QualifyingReport/QualifyingReport.tex
references pt1

@@
 Participants are given a list of 15 unrelated words, repeated over five different trials and are asked to repeat.
-Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then again after 30 minutes.
+Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then again after 30 minutes \cite{}.
 \par
@@
 The Rey-Osterrieth Complex Figure (ROCF) is a task widely as a test of visuo-spatial skill and visual memory.
-The task, which was originally designed by Rey (1944) and standardised by Osterrieth (1944), asks a participant to copy a complex geometrical figure (known as the immediate copy condition and to recall and reproduce the figure from memory without warning (known as the delayed recall condition).
+The task, which was originally designed by Rey (1944) and standardised by Osterrieth (1944) \cite{Rey1941}, asks a participant to copy a complex geometrical figure (known as the immediate copy condition and to recall and reproduce the figure from memory without warning (known as the delayed recall condition).
 In the immediate copy condition, the complexity of the figure requires an integrative cognitive ability.

### 2019-01-27 Source/QualifyingReport/QualifyingReport.tex
references almost complete

@@
 Participants are given a list of 15 unrelated words, repeated over five different trials and are asked to repeat.
-Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then again after 30 minutes \cite{}.
+Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then again after 30 minutes \cite{Rey1941}.
 \par
@@
 Further, they found that the conversion of MCI subjects to AD in 3-years could be predicted based on either observed or estimated RAVLT scores with an accuracy comparible to MRI-based biomarkers \cite{Moradi2017}.
-Another study by Schoenberg found that RAVLT to best distinguish patients suspected of Alzheimer's disease from the psychiatric comparison group.
-% Schoenberg - Test performance and classification statistics for the Rey Auditory Verbal Learning Test in selected clinical samples.
-
+Another study by Schoenberg found that RAVLT to best distinguish patients suspected of Alzheimer's disease from the psychiatric comparison group \cite{Schoenberg2006}.
 \subsection{Digit Symbol Substitution Test}
@@
 The task requires that the participant move between the key and the randomized sequence such that  they may retrieve the correct answer from the key, hold this in short-term memory and transcribe the key in the appropriate place.
-% Rosano - Digit symbol substitution test and future clinical and subclinical disorders of cognition, mobility and mood in older adults.
+\par 
 Among those with no disorder in cognition, mobility and mood, being in the lowest DSST quartile compared to the highest was associated with nearly twice the odds of developing one or more clinical or subclinical disorders.
@@
 Slower psychomotor speed may serve as a biomarker of risk of clinical disorders, mobility and mood.
-While in part attributable to vascular brain disease, other potentially modifiable contributors may be present.
+While in part attributable to vascular brain disease, other potentially modifiable contributors may be present \cite{Rosano2016}.
 Further studying the causes of psychomotor slowing with ageing might provide novel insights into age-related brain disorders.
 Pascoe compared patients with PD with Normal Cognition (PD-N) with those with PD and MCI (PD-MCI) and healthy participants.
-PD-MCI pariticipants achieved significantly lower scores than other groups in the DSST task.
-% Pascoe - The symbol-digit modalities test in MCI: Evidence from Parkinson's Disease Patients.
-
+PD-MCI pariticipants achieved significantly lower scores than other groups in the DSST task \cite{Pascoe2018}.
 \subsection{Verbal Fluency}
-%% In this paragraph cite Henry et al, 2004 - taken from Taler and Phillips (2008) - Henry et al, 2004 is a meta analysis
 There are a number of tests that characterise this category of verbal fluency, but generally fall into two categories.
@@
 There have been numerous studies which have documented the impact that AD has on verbal fluency tests in both categories.
-A review carried out by Henry et al, found that performance in both letter fluency and category fluency was impaired in those with AD vs controls, but found a larger effect for tests of semantic fluency.
+A review carried out by Henry et al \cite{Henry2004}, found that performance in both letter fluency and category fluency was impaired in those with AD vs controls, but found a larger effect for tests of semantic fluency.
 \subsection{Naming Tests}
@@
 \par 
-% Willers - Subclinical naming errors in MCI - A semantic deficit.
-Willers et al studied Twenty aMCI patients, twenty AD and 21 controls matched by age, sex and education level.
+Willers et al \cite{Willers2008} studied Twenty aMCI patients, twenty AD and 21 controls matched by age, sex and education level.
 They found AD patients obtained significantly lower total scores on the BNT than aMCI patients and controls.
@@
 \par 
-% Vadikolias - Mild Cognitive Impairment: effect of education on the verbal and nonverbal tasks performance decline.
-Vadikolias investigated the impact of education on naming tests and found that higher educational attainment in aMCI subjects were correlated with better performance in verbal and non-verbal tasks during repeated examinations over 1-year.
+Vadikolias investigated the impact of education on naming tests \cite{Vadikolias2012} and found that higher educational attainment in aMCI subjects were correlated with better performance in verbal and non-verbal tasks during repeated examinations over 1-year.
 Subjects with a lower level of education performed worse than patients with a high level of education who presented a more stable clinical score.
@@
 Whilst these studies and tests provide evidence that there are word finding difficulties in those with MCI, on their own they do not provide sufficient ability to diagnose MCI or provide an prognosis of disease progression.
-
 \subsection{Rey-Osterrieth Complex Figure Task}
-% Cite Rey 1941 - L'examen psychologique dans les cas d'encephalopathie traumatique and Osterrieth 1944 - Le test de copie d'une figure complexe.
 The Rey-Osterrieth Complex Figure (ROCF) is a task widely as a test of visuo-spatial skill and visual memory.
@@
 The reproduction of such a complex structure involves processes such as planning and organizational strategies that are related to executive functions.
-% Salvadori - Qualitative Evaluation of the Immediate Copy of the Rey-Osterrieth Complex Figure: Comparison Between Vascular and Degenerative MCI patients.
-Salvadori found that patients with vascular MCI had a worse performance in the immediate copy of the ROCF compared to individuals with degenerative MCI, despite their significant impairment in terms of general cognitive status and visual memory.
+\par
+Salvadori found that patients with vascular MCI had a worse performance in the immediate copy of the ROCF compared to individuals with degenerative MCI, despite their significant impairment in terms of general cognitive status and visual memory \cite{Salvadori2018}.
 Evidence shows that patients with disorders that possibly involve attention and executive functions are characterized by a more disorganised approach when copying the ROCF compared to controls.
-One of the difficulties with the presentation of this task in a repeated battery will be the fact that it turns from an incidental memory task (it is a memory task but the participant is not forewarned that they will be required to memorise the picture) into an intentional memory task (given the previous exposure to the task, it would be expected that a participant would pre-empt the delayed recall portion of the test and spend more time attempting to memorise the details).
-% Teng - Persistence of Neuropsychological Testing Deficits in Mild Cognitive Impairment.
-
+One of the difficulties with the presentation of this task in a repeated battery will be the fact that it turns from an incidental memory task (it is a memory task but the participant is not forewarned that they will be required to memorise the picture) into an intentional memory task (given the previous exposure to the task, it would be expected that a participant would pre-empt the delayed recall portion of the test and spend more time attempting to memorise the details) \cite{Teng2009}.
 \subsection{Hayling Sentence Completion Test}
@@
 \par 
-% Martyr - Assessing inhibitory control in early-stage Alzheimer's and Parkinson's disease using the Hayling Sentence Completion Test.
-Martyr et al compared healthy controls with patients with dementia and patients with parkinson's disease.
+Martyr et al compared healthy controls with patients with dementia and patients with parkinson's disease \cite{Martyr2017}.
 They found that a high proportion of Category A errors (producing a word that fits the sentence when instructed otherwise) was a factor in performance loss for participants with dementia.
@@
 However, some studies have shown a correlation of performance in the GPT and measures of cognitive performance such as the Montreal Cognitive Assessment.
-% Bezdicek - Grooved Pegboard Predicates More of Cognitive Than Motor Involvement in PD.
-Bezdicek found that the GPT predicted performance on the MoCA and concluded that in addition to being a measure of motor skills, there were results that showed that the GPT could also provide information about a participants cognitive skills.
+\par 
+Bezdicek found that the GPT predicted performance on the MoCA and concluded that in addition to being a measure of motor skills, there were results that showed that the GPT could also provide information about a participants cognitive skills \cite{Bezdicek2014}.
 Whilst this was specifically using a population of patients with Parkinson's disease, given the MoCA and GPT have shown degree of correlation, there remains an opportunity for research that explores the use of the GPT with MCI and Early AD populations.
-% Darweesh - Simple Test of Manual Dexterity can identify persons at high risk for neurodegenerative diseases in the community.
-Darweesh et al, used the Purdue Pegboard Test to assess manual dexterity in a healthy older adult population and followed their participants for between eight and twelve years until an indication of the onset of a neurodegenerative disease.
+\par 
+Darweesh et al, used the Purdue Pegboard Test to assess manual dexterity in a healthy older adult population and followed their participants for between eight and twelve years until an indication of the onset of a neurodegenerative disease \cite{Darweesh2017}.
 In this time 227 (4\%) of their participants went on to develop a diagnosis of dementia.
@@
 This effect is magnified in participants who have a cognitive impairment.
-% Abnormal Visual Search in MCI and Alzheimer's Disease. Tales.A 2005
-Research has shown there to be deficits in both AD and MCI populations in this task.
+\par
+Research has shown there to be deficits in both AD and MCI populations in this task \cite{Tales2005}.
 Whilst these deficits were not as apparent in MCI populations, there remained a significant enough difference that could be used to differentiate healthy controls from those with MCI.
@@
 Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cognitive decline due to a cognitive impairment.
-FCSRT-IR (The instant recall version of the Free Cued Selective Reminding Test) performance has beeen shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the authors to define prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT-IR performance (Sarazin et al, 2007.
+FCSRT-IR (The instant recall version of the Free Cued Selective Reminding Test) performance has beeen shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the authors to define prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT-IR performance \cite{Sarazin2007}.
 The sensitivity and specificity of the different test forms in correctly classifying patients were compared to each other using a previously established FCSRT-IR cut off score.
@@
 One of the most commonly used tasks to measure language is the Picture Description task.
-An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Kaplan2010}.
+An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Borod1980}.
 The Cookie Theft picture (pictured below) depicts a scene of a home typical of the period of time when it was created and would generally not require participants to use any complicated vocabulary to describe.
@@
 Some examples of semantic content measures are Type to Token Ratio, Brunet's Index and Honore's Statistic (described below) which are used to measure the lexical diversity of a given piece of text and/or utterance.
-This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2005}.
+This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2013}.
 \par
@@
 \par
-\subsubsection{Honore's Statistic (R)} % Don't forget to cite the original article.
+\subsubsection{Honore's Statistic (R)}
 Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation:
@@
 \par
-Hier, Hagenlocker and Shindler assessed content using a similar list of thematic elements.
+Hier, Hagenlocker and Shindler assessed content using a similar list of thematic elements \cite{Hier1985}.
 They divided their participants into early-stage and late-stage AD, as well as including a control group.
 The late-stage AD group produced significantly fewer relevant observations than the early stage group, and the AD group combined produced fewer relevant observations than controls.
-This study was replicated by Vuorinen et al (2001).
-\par
-Smith, Chenery and Murdoch (1989) applied Hier's methodology for constructing pictorial 'themes' with the Picnic Scene from the Western Aphasia Battery (WAB) with a control and patients with moderate to moderately severe AD.
+This study was replicated by Lukatela et al \cite{Lukatela1998}.
+\par
+Smith, Chenery and Murdoch applied Hier's methodology for constructing pictorial 'themes' with the Picnic Scene from the Western Aphasia Battery (WAB) with a control and patients with moderate to moderately severe AD \cite{Smith1989}.
 The authors found no difference in the number of semantic elements produced but did not that the group with moderate to moderately severe AD took more time and more syllables to communicate these elements.
@@
 \end{figure}
-Sajjadi et al (2012) examined 10 pictorial themes in picture description the Comprehensive Aphasia Test and found that the group with mild AD produced similar themes than controls.
+Sajjadi et al examined 10 pictorial themes in picture description the Comprehensive Aphasia Test and found that the group with mild AD produced similar themes than controls \cite{Sajjadi2012}.
 Bschor et al.
@@
 Some studies used a more general concept of content, defining "information units" as "the smallest non redundant meaningful fact or inference," whether or not the information conveyed was specific to the context in which the conversation happened.
-Giles et al, for example, studied adults with minimal, mild or moderate AD vs controls and found that adults with AD produced fewer overall information units than controls.
+Giles et al for example studied adults with minimal, mild or moderate AD vs controls and found that adults with AD produced fewer overall information units than controls \cite{Giles1996}.
 \subsection{Conciseness of information}
@@
 Conciseness is calculated by dividing the number of ideas expressed by the total number of words in a measure commonly referred to as idea density but also known as lexical index, information content and information unit conciseness index.
-Snowdon et al examined written discourse from the Nun study and found that low idea density in early life was associated with reduced cognitive performance in later life.
+Snowdon examined written discourse from the Nun study and found that low idea density in early life was associated with reduced cognitive performance in later life \cite{Snowdon1996}.
 Riley et al extended these findings by concluding that early-life idea density was associated with lower brain weight, higher degree of cerebral atrophy and increased neurofibrillary pathology in later life.
@@
 Murray found that MLU was not a distinguishing factor among health adults, adults with depression and adults with AD.
-Ripich et al found a decrease in MLU in adults with severe AD over time, and this was supported by findings of Le et al in their studies of authors \cite{Le2012}
+Ripich et al found a decrease in MLU in adults with severe AD over time, and this was supported by findings of Le et al in their studies of authors \cite{Ripich1988, Le2011}
 \par
 \subsection{Proportion of verbs to nouns plus verbs}
-Kave and Levy used a verb index to capture syntactic complexity and found that adults with AD expressed the same amount of verbs to nouns plus verbs as adult controls.
+Kave and Levy used a verb index to capture syntactic complexity and found that adults with AD expressed the same amount of verbs to nouns plus verbs as adult controls \cite{Kave2018}.
 \par
 \subsection{Syntactic Complexity - Composite measures of MLU, syntactic errors and verbs}
-Ahmed et al, and Ahmed, Haigh et al found differences in syntactic complexity between adults with MCI and controls, and between MCI and moderate AD stages.
+A number of studies which explored syntactic complexity found differences between adults with MCI and controls, and between MCI and moderate AD stages \cite{Rentoumi2014, Ahmed2013}.
 The differences in syntactic complexity were not significant when individual measures were tested, but were apparent using a composite score consisting of MLU, words in sentences, syntactic errors, nouns with determiners, and verbs with inflections.
@@
 \subsection{Questions, turn-taking, unsure statements, egocentric comments}
-One study, Ripich et al, examined several pragmatic abilities among patients at different stages of AD.
+Ripich et al examined several pragmatic abilities among patients at different stages of AD \cite{Ripich1988}.
 The severe AD group asked more questions over time than the other group.
@@
 Whilst this improved the ecological validity of the study, the caregivers delivery would be varied and inconsistent.
-Secondly, due to the constrained nature of the picture description task, it is unlikely that it would be able to capture the pragmatic skills that are typioical of conversations in everyday life.
+Secondly, due to the constrained nature of the picture description task, it is unlikely that it would be able to capture the pragmatic skills that are typical of conversations in everyday life.
 \par
@@
 Coherence refers to the appropriate maintenance of topic in discourse and is thematically related to the immediately preceding utterance (local coherence) and by how closely an utterance relates to the general topic at hand (global coherence).
-One study looked at coherence, Chapman et al used picture descriptions of Norman Rockwell prints within a frame analysis, with frames being defined as internalized knowledge structures.
+One study looked at coherence, Chapman et al used picture descriptions of Norman Rockwell prints within a frame analysis, with frames being defined as internalized knowledge structures \cite{Chapman1998}.
 The authors identified aspects of content, including typical frames of interpretation, atypical, incorrect or no frames, propositions supporting frames and propositions disrupting frames as measures of coherence.
@@
 There are a number of studies, that have examined "melodic line", which is a subjective measure of speech prosody defined as "the appropriate use of intonational contour, including alterations in pitch, volume and duration".
-For instance, Forbes-McKay compared melodic line using a "simple" picuture such as the Cookie Theft picture vs a complex picture.
-The number of pictorial themes differentiated the simple tasks from the complex.
+For instance, Forbes-McKay compared melodic line using a "simple" picture such as the Cookie Theft picture vs a complex picture.
+The number of pictorial themes differentiated the simple tasks from the complex \cite{Forbes-Mckay2005}.
 Results showed no group differences on simple picture tasks (Cookie Theft and Tripping Woman) but there were differences in melodic line for the complex picture tasks.
-Fraser et al examined several acoustic features of speech in patients with AD ussing machine learning methods that captured both rate and phonation patterns, and found acoustic abnormalities to be a significant factor.
-Konig et al, also used an automated feature analysis examining vocal and temporal aspects of speech among controls, and patients with MCI or AD, and reported a classification of 81\%.
+Fraser et al examined several acoustic features of speech in patients with AD using machine learning methods that captured both rate and phonation patterns, and found acoustic abnormalities to be a significant factor.
+Konig et al, also used an automated feature analysis examining vocal and temporal aspects of speech among controls, and patients with MCI or AD, and reported a classification of 81\% \cite{Konig2018}.
 \subsection{Discourse Fluency}
@@
 \par
-The study of "empty speech" by Nicholas et al was one of the first to examine aspects of fluency in the connected speech of persons with AD.
+The study of "empty speech" by Nicholas et al was one of the first to examine aspects of fluency in the connected speech of persons with AD \cite{Nicholas1985}.
 They found that adults with AD had significantly more repetitions than controls.
-Similarly, Tomoeda et al found more aborted phrases, revisions and ideational repetitions in the AD group than in controls.
+Similarly, Bayles and Tomoeda found more aborted phrases, revisions and ideational repetitions in the AD group than in controls \cite{Bayles1983}.
 Several other studies support the idea that in AD population there are a greater number of repetitions and revisions than in healthy controls.
-However, there are some studies which contradict these findings (Ahmed et al).
+However, there are some studies which contradict these findings\cite{Ahmed2013}.
 \par
@@
 Severity of naming deficits correlated negatively to the amount of uncorrected errors.
-The authors suggested that this impairment was related to the executive function dificulties in the clinical groups.
+The authors suggested that this impairment was related to the executive function difficulties in the clinical groups.
 The authors did not report correlations between error monitoring and executive function test scores, however, which could have strengthened that hypothesis.
@@
 Another measure of speech monitoring used in the picture description literature is "response to word-finding delays," defined as "whether patients appear unaware of their problem, produce an approximation of the target word or actively search and produce the target word".
-Response to word-finding delays differed significantly between minimal  AD and normal controls (forbes-McKay and Venneri).
+Response to word-finding delays differed significantly between minimal  AD and normal controls \cite{Forbes-Mckay2005}.
 The measure was based on Goodglass and Kaplan's scale for rating discourse on the Boston Diagnostic Aphasia Examination (BDAE) and is composed of clinical judgement of behaviour that is rated on a likert-type scale ranging fdrom 1 to 7 (7 being no abnormality).
@@
 % CITE Bird, Steven, Edward Loper and Ewan Klein (2009). Natural Language Processing with Python.  O'Reilly Media Inc.
-The Natural Language Toolkit (NLTK) was originally designed as part of a computational linguistics course but has evolved to become an open-source framework which includes methods and modules for a wide array of Natural Language tasks (see Table 2.2 for a full list).
+The Natural Language Toolkit (NLTK) was originally designed as part of a computational linguistics course but has evolved to become an open-source framework which includes methods and modules for a wide array of Natural Language tasks (see Table 2.2 for a full list) \cite{Bird2009}.
 
@@
 \par
-Teng et al suggests that work should focus on the MCI population and concentrate on developing a consensus neuropsychological battery that could yield predictable rates of progression to AD.
+Teng et al suggests that work should focus on the MCI population and concentrate on developing a consensus neuropsychological battery that could yield predictable rates of progression to AD \cite{Teng2009}.
 This, in conjunction with the development of a model of language, sensitive enough to detect subtle deterioration in language use to act as an additional cognitive marker to aid diagnosis could potentially move some way to providing this.
@@
 \begin{itemize}
-	\item \textbf{Hypothesis} Is monitoring language samples and other measures of cognitive decline over the course of the year enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment.
-	\item \textbf{Hypothesis} Those with MCI will show a different language profile to controls if matched similarly for age, gender, education and other significant factors.
-	\item \textbf{Hypothesis} The new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures.
-	\item \textbf{Hypothesis} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, those with Alzheimers Disease and decline in those with MCI.
+	\item \textbf{Hypothesis One} Is monitoring language samples and other measures of cognitive decline over the course of the year enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment.
+	\item \textbf{Hypothesis Two} Those with MCI will show a different language profile to controls if matched similarly for age, gender, education and other significant factors.
+	\item \textbf{Hypothesis Three} The new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures.
+	\item \textbf{Hypothesis Four} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, those with Alzheimers Disease and decline in those with MCI.
 \end{itemize}

### 2019-01-28 Source/QualifyingReport/QualifyingReport.tex
28-01-2019 edits

@@
 In 2009, the Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
-A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cogntive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease\cite{Taler2008}.
+A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease\cite{Taler2008}.
 As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages \cite{Prince2015} and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
@@
 Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
-One of the difficulties in research of this nature is the process of collecting appropriate language samples.
-Part of this work therefore is to see whether spontaneous discourse, such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD.
+Another area of focus for research of this nature is the process of collecting appropriate language samples.
+Whilst collecting samples of language is comparatively unintrusive, researchers recognise that these samples require a rich sample of language that potentially cannot be generated by tasks such as the picture description task.
+Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD.
 There is some evidence to support this.
@@
 Alzheimer's Disease (AD) is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
-Whilst a definitive diagnosis of Dementia can only produced at post-mortem, there are a number of clinical indicators from psychological perspective that can indicate dementia is present.
+Whilst a definitive diagnosis of Dementia can only be produced at post-mortem, there are a number of clinical indicators from psychological perspective that can indicate dementia is present.
 From a clinical psychology perspective, those who have dementia demonstrate cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language and visuospatial deficits\cite{McKhann2011}.
@@
 \end{enumerate}
-In addition to meeting the above criteria, a clinician must rule out other conditions or factors that could account for the decline in cognition with the goal to increase the likelihood that the underlying cause of this decline is AD.
-\par 
-Many researchers have studied the early detection of MCI \ AD.
+In addition to meeting the above criteria, a clinician must rule out other conditions or factors that could account for the decline in cognition with the goal to increase the likelihood that the underlying cause of this decline is dementia.
+\par 
+Many researchers have studied the early detection of MCI.
 These studies usually follow two main approaches.
@@
 \par
-One of the most common ways in which clinicians traditionally make an early diagnosis of cognitive impairment is through the use of the Mini Mental State Examination (MMSE) \cite{Folstein1975}.
+One of the most common ways in which clinicians traditionally make an early diagnosis of dementia is through the use of the Mini Mental State Examination (MMSE) \cite{Folstein1975}.
 The MMSE is a brief questionnaire consisting of eleven questions which tests cognitive aspects of mental function and requires only 5-10 minutes to administer \cite{Folstein1975}.
@@
 It also has limited value in assessing under-educated subjects and a meta-analysis on the effectiveness of the MMSE as a diagnostic tool for dementia showed that it's accuracy was low (sensitivity between 78.4\% and 85.1\% and specificity between 81.3\% and 87.8\%).
-As the MMSE is shown to have low accuracy specifically in the diagnosis of dementia, it becomes necessary for professionals to employ the use of other tools or measures such as the Free Cued Selective Reminding Test (FCSRT) \cite{Grober2010} or the Montreal Cognitive Assessment (MoCA) \cite{Davis2015}.
+As the MMSE has been shown to have low accuracy specifically in the diagnosis of dementia, it becomes necessary for professionals to employ the use of other tools or measures such as the Free Cued Selective Reminding Test (FCSRT) \cite{Grober2010} or the Montreal Cognitive Assessment (MoCA) \cite{Davis2015}.
+\par 
 These tests have the benefits of being much more accurate at diagnosing cognitive impairment and discriminating between dementia and other types of cognitive impairment at the cost time and training of psychological professionals such as clinical psychologists in administering these tests.
-Given the burden on these professionals is likely to increase due to a growing elderly population or in the case of developing countries where the clinician / patient ratio is already unsustainable, it seems useful to find less burdensome ways to aid professionals in identifying those at risk of developing dementia.
+However, the utility of diagnosing dementia at the point where clinical intervention is warranted is limited because at this stage both psychological and pharmacological interventions have been shown to not be effective \cite{Prince2015}.
+In order to further our understanding of the progression of dementia it is important to detect the signs of dementia before they are clinically apparent.
 \par
 The two main ways in which diagnosis is performed is through assessment of memory and language.
-Tests of memory are classically among the most accurate ways of diagnosing dementia, however these tests suffer from the same reliance on trained clinicians to administer these tests in a clinical setting.
+Tests of memory are classically among the most accurate ways of diagnosing dementia, however these tests suffer from the same reliance on clinicians to administer these tests in a clinical setting.
 Language however is a lot easier to collect and can be done in more naturalistic settings.
@@
 \par
-One of the most famous pieces of research on the topic of language decline in dementia was by Berisha and Liss (2015) \cite{Berisha2015} who examined speeches and public interviews of former US president Ronald Reagan.
+One of the most famous pieces of research on the topic of language decline in dementia was by Berisha and Liss who examined speeches and public interviews of former US president Ronald Reagan \cite{Berisha2015}.
 They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease.
 It turned out later to be the case that he had Alzheimer's Disease.
-Another classical study by Snowdon et al (1996) \cite{Snowdon1996} looked at whether linguistic ability in early life was associated with cognitive function and AD in later life.
+Another classical study by Snowdon et al looked at whether linguistic ability in early life was associated with cognitive function and AD in later life \cite{Snowdon1996} .
 They found that idea density (defined as the number of expressed propositions divided by the number of words) was a key predictor in predicting whether nuns would go on to develop AD in later life.
@@
 She asserted that the language forms we learn last are the first to deteriorate as we generally learn language in small simple units initially and build syntax and complexity as we are more comfortable with language.
+However it is important to note that different variants of dementia show different deficits in terms of language productions.
+Regardless, dividing language in this way is useful as it allows us to detect deteriorations in different parts of language usage and therefore may provide a way of discriminating between different forms of dementia.
 \par
 It is clear from both the clinical diagnostic criteria and supporting research that language is impacted in those with AD.
-However, one of the costs of analysing language is that there is a huge burden on trained practitioners, be it clinical psychologists, audio transcribers and text encoders to facilitate the process of collecting data and analysis.
+However, whilst there is a move towards research aimed at looking more specifically at MCI we currently lack the measures that are sensitive enough to detect MCI.
+Given that we know language is affected before a clinical diagnosis of dementia is usually made \cite{Berisha2015, Snowdon1996, Le2011}, it makes sense to explore whether language on it's own can provide markers that may indicate a cognitive impairment that could progress to dementia.
 The field of machine learning and natural language processing has been suggested as a way to improve the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline \cite{Boschi2017}.
@@
 \subsection{Rey Auditory-Verbal learning test}
-Rey's Auditory Verbal Learning Test (RAVLT+) looks a wide range of neuropsychological processess including short-term auditory-verbal memory, retention of information.
+Rey's Auditory Verbal Learning Test (RAVLT+) looks a wide range of neuropsychological processess including short-term auditory-verbal memory, learning and retention of information.
 Participants are given a list of 15 unrelated words, repeated over five different trials and are asked to repeat.
-Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then again after 30 minutes \cite{Rey1941}.
+Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then repeat it again after 30 minutes \cite{Rey1941}.
 \par
@@
 Thus making the RAVLT an effective early marker to detect AD in persons with memory complaints.
-Moradi investigated to what extent the RAVLT scores are predictable based on MRI data using machine learning approaches, as well as to find out what the most important brain regions are for the estimation of RAVLT scores.
+Moradi investigated to what extent the RAVLT scores are predictable based on MRI data using machine learning approaches, as well as to find out what the most important brain regions are for the estimation of RAVLT scores \cite{Moradi2017}.
 They found a highly significant cross validated correlation between the estimated and observed RAVLT immediate and RAVLT Percent Forgetting.
@@
 \subsection{Rey-Osterrieth Complex Figure Task}
+% CITE - Osterrieth - Filetest de copied’une figure complex: contribution al’etude de la perception et de la memoire [The test of copying a complex figure: a contribution to the study of perception and memory]
+% CITE - Taylor 1969 1979 - Variants
+% CITE - Strauss - A comparison of the Rey and Taylor figures
 The Rey-Osterrieth Complex Figure (ROCF) is a task widely as a test of visuo-spatial skill and visual memory.
@@
 \par
-Research has shown there to be deficits in both AD and MCI populations in this task \cite{Tales2005}.
-Whilst these deficits were not as apparent in MCI populations, there remained a significant enough difference that could be used to differentiate healthy controls from those with MCI.
+Research has shown there to be deficits in both AD and MCI populations in this task \cite{Tales2005}, and those with MCI exhibited less severe deficits compabut that these deficits were not as apparent in MCI populations, there remained a significant enough difference that could be used to differentiate healthy controls from those with MCI.
 In a comparison of visual search performance between those with aMCI and healthy controls, Tales noted significantly poorer performance in the aMCI group.
 However, she also noted a good deal of heterogeneity in the aMCI group which illustrates that whilst the aMCI group have essentially the same condition, presentations within this population can differ markedly.
-The results from this trial also illustrates the use of a non-memory task as a means of diagnosing dementia as patients who went on to be diagnosed with dementia whilst this study was in progress had significantly poorer visual search performance.
+The results from this study also illustrates the use of a non-memory task as a means of diagnosing dementia as patients who went on to be diagnosed with dementia whilst this study was in progress had significantly poorer visual search performance.
 \subsection{Free Cued Selective Reminding Test}
+% Describe the test
 The Free Cued Selective Reminding Test (FCSRT) was borne out of the premise that by controlling the conditions of learning, a measure of memory is possible that is not confounded by normal age related changes in cognition.
 Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cognitive decline due to a cognitive impairment.
-FCSRT-IR (The instant recall version of the Free Cued Selective Reminding Test) performance has beeen shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the authors to define prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT-IR performance \cite{Sarazin2007}.
-The sensitivity and specificity of the different test forms in correctly classifying patients were compared to each other using a previously established FCSRT-IR cut off score.
+The FCSRT starts with a learning phase in which participants are asked to look at a card containing four pictures (e.g., grapes) for an item that belongs to a named category(e.g., fruit).
+After these four items are identified, immediate cued recall of these four items is tested and this is repeated for a total of 16 items.
+Following this learning phase is the recall phase in which participants are asked to recall all the items identified without cues and any items which are not recalled are then cued.
+There are three outcome measures for this test, free recall (the number of items recalled without cues), total recall (the number of items recalled with or without cues) and cue efficiency which is identified as follows \cite{Grober2010}.
+\begin{equation} \label{x1}
+Cue efficiency = (totalRecall-freeRecall)/48-freeRecall, range 0.0-1.0)
+\end{equation}
+Grober found that patients with impaired free recall with four times more likely to develop dementia than those with intact free recall \cite{Grober2010}.
+This test has also been shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the researchers to used the test to categorise prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT score \cite{Sarazin2007}.
 \subsubsection{Conclusions}
@@
 Future research should look at the standardization of the operational definition of cognitive impairment in MCI may result in more consistent predictions of progression to AD.
+\par 
+Finally, it can be said that in testing for MCI and comparing these results with a cohort of those with a similar diagnosis has limited utility.
+This is because there is huge variability in the presentation of MCI within participants that even if controlled for with a matched pairs design for age, gender and education will produce inconsistent results.
+There is an argument therefore for the use of longitudinal studies, where the comparisons on the performance of these tasks are with the participants themselves.
+There are not many studies which use a longitudinal approach, although those that do show promising results.
 
@@
 \par
-Single word tasks such as the Boston Naming Test and other such standardized language tests generally target a participants verbal fluency where this is defined as the ability to form and express words in accordance with certain criteria (see above for a full description of verbal fluency as task).
+Single word tasks such as the Boston Naming Test and other such standardized language tests generally target a participants word production where this is defined as the ability to form and express words in accordance with certain criteria (see above for a full description of verbal fluency as task).
 There are a number of benefits to using single word tasks.
@@
 Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
-\begin{equation} \label{x1}
+\begin{equation} \label{x2}
 TTR = numberOfUniqueWords / totalNumberOfWords.
@@
 Brunet's Index (W) differentiates itself for TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
-\begin{equation} \label{x2}
+\begin{equation} \label{x3}
 W = N^{V(-0.165)}
@@
 Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation:
-\begin{equation} \label{x3}
+\begin{equation} \label{x4}
 %% Check to see if this is accurate.
@@
 Studies using these features found some differences between AD and controls, and some could differentiate between different stages of AD.
-\subsection{General Information Units}
+\subsection{General Information Units or Content Information Units}
 Some studies used a more general concept of content, defining "information units" as "the smallest non redundant meaningful fact or inference," whether or not the information conveyed was specific to the context in which the conversation happened.
@@
 \subsection{Conciseness of information}
+% CITE RILEY!
 Conciseness has been defined as the number of words a speaker uses to express ideas.
@@
 \par
-Ahmed, de Jager et al examined idea density with patients who had confirmed AD post mortem.
+Ahmed, de Jager et al examined idea density with patients who had confirmed AD post mortem \cite{Ahmed2013}.
 They found that those with AD produced fewer total semantic units than controls but there was no significant difference between the groups with regards to idea density.
@@
 \subsection{Efficiency}
+% CITE MURRAY!
 Efficiency is the rate at which meaningful information is conveyed over time, calculated by dividing the total number of information units by the duration in seconds of the speech sample.
-Smith et al, 1989 found that 18 adults with moderately severe AD produced fewer content units per minute on average than controls, and attributed this different to increased circumlocutions and repetitions.
-Murray 2010, used an analogous measure which he called "performance deviations per minute", in which fillers, irrelevant words, revisions or false stars, vague or non-specific vocabulary and inaccurate output (e.g.
+Smith et al, 1989 found that 18 adults with moderately severe AD produced fewer content units per minute on average than controls, and attributed this different to increased circumlocutions and repetitions \cite{Smith1989}.
+Murray used an analogous measure which he called "performance deviations per minute", in which fillers, irrelevant words, revisions or false stars, vague or non-specific vocabulary and inaccurate output (e.g.
 paraphasias) were divided by the total number of minutes in the speech sample; this measure was lower for those with AD than those with depression, and also healthy controls.
 The authors suggested that discourse information measures may help disentangle the similarities in symptoms of early AD versus depression in older adults.
-Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in dialogue in which a speaker provides responsees do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls.
+Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in dialogue in which a speaker provides responses do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls.
 \subsection{Quantity - Total number of words}
@@
 Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
-They found that there were four factors which informed the classification of participants as either healthy or AD.
+They found that there were four factors which they identified as important in the classification of participants as either healthy or AD.
 These four factors were semantic impairment, acoustic abnormality, syntactic impairment and information impairment and were based on existing measures of semantic and syntactic complexity.
@@
 Ripich et al examined several pragmatic abilities among patients at different stages of AD \cite{Ripich1988}.
-The severe AD group asked more questions over time than the other group.
-The authors argued that question-asking was a compensatory strategy, and as a result, adults in late-stage AD may have had insight into their communication problems.
-Potentially, there were a number of flaws of the methodology if their research.
+The severe AD group asked more questions over time than the other group which infers problems with language processing.
+The authors argued that question-asking was a compensatory strategy and as a result, the proposed that adults in late-stage AD may have had insight into their communication problems.
+\par 
+There were a number of flaws of the methodology if their research.
 Firstly, a caregiver was asked to administer the picture description task in order to mirror a more typical communicative interaction.
@@
 Secondly, due to the constrained nature of the picture description task, it is unlikely that it would be able to capture the pragmatic skills that are typical of conversations in everyday life.
-\par
 \subsection{Coherence}
@@
 \par
-\subsection{Speech intonation / prosody}
-There are a number of studies, that have examined "melodic line", which is a subjective measure of speech prosody defined as "the appropriate use of intonational contour, including alterations in pitch, volume and duration".
-For instance, Forbes-McKay compared melodic line using a "simple" picture such as the Cookie Theft picture vs a complex picture.
-The number of pictorial themes differentiated the simple tasks from the complex \cite{Forbes-Mckay2005}.
-Results showed no group differences on simple picture tasks (Cookie Theft and Tripping Woman) but there were differences in melodic line for the complex picture tasks.
-Fraser et al examined several acoustic features of speech in patients with AD using machine learning methods that captured both rate and phonation patterns, and found acoustic abnormalities to be a significant factor.
-Konig et al, also used an automated feature analysis examining vocal and temporal aspects of speech among controls, and patients with MCI or AD, and reported a classification of 81\% \cite{Konig2018}.
-\subsection{Discourse Fluency}
-Verbal fluency is a term used in neuropsychological contexts generally referring to timed, word-generation tasks, while in speech-language pathology contexts, "fluency disorders" are defined as interruptions int he flow of speaking characterized by atypical rate, rhythm and repetitions in sounds, syllables, words and phrases.
+\subsection{Empty Speech}
+Verbal fluency is a term used in neuropsychological contexts generally referring to timed, word-generation tasks, while in speech-language pathology contexts, "fluency disorders" are defined as interruptions in the flow of speaking characterized by atypical rate, rhythm and repetitions in sounds, syllables, words and phrases.
 "Fluency", in the literature of discourse of adults with AD, typically refers to the smoothness or flow of spoken language.
 Abnormalities of fluency in this population are typically characterised by filled and unfilled pauses, word repetitions, circumlocutions, and revisions.
-In contrast with fluency disorders (i.e.
-stuttering), abnormalities in the fluency of adults with MCI or AD are rarely manifested at the phonological level.
 \par
@@
 However, there are some studies which contradict these findings\cite{Ahmed2013}.
-\par
-\subsection{Speech Monitoring}
-Speech monitoring is related not only to word retrieval deficits and to pragmatic language skill but also to "anosognosia" define as the awareness of one's own deficits.
-McNamara et al investigated word error monitoring skills by comparing uncorrected and repaid errors in adults with AD vs health controls.
-The AD group was equally impaired in error monitoring as the controls.
-Severity of naming deficits correlated negatively to the amount of uncorrected errors.
-The authors suggested that this impairment was related to the executive function difficulties in the clinical groups.
-The authors did not report correlations between error monitoring and executive function test scores, however, which could have strengthened that hypothesis.
-\par
-Another measure of speech monitoring used in the picture description literature is "response to word-finding delays," defined as "whether patients appear unaware of their problem, produce an approximation of the target word or actively search and produce the target word".
-Response to word-finding delays differed significantly between minimal  AD and normal controls \cite{Forbes-Mckay2005}.
-The measure was based on Goodglass and Kaplan's scale for rating discourse on the Boston Diagnostic Aphasia Examination (BDAE) and is composed of clinical judgement of behaviour that is rated on a likert-type scale ranging fdrom 1 to 7 (7 being no abnormality).
-\par
 \subsection{Conclusions}
@@
 % Thomas - Automatic Detection and Rating of AD through the Lexical Analysis of Spontaneous Speech
-Thomas et al, detail several lexical approaches to the problem of detecting and rating DAT.
+Thomas et al, detail several lexical approaches to the problem of detecting and rating AD.
 The approaches rely on character n-gram techniques but also explore correlation of usage of frequency of different types of speech.
@@
 This usually takes the form analysis of speech recorded as part of a cognitive assessment such as the Picture Description Task \cite{Orimaye2008}.
-Given that language samples are the relatively easy to collect, research has moved towards analysis of spontaneous speech.
+Given that language samples are relatively easy to collect, research has moved towards analysis of spontaneous speech.
 An good example of this type of research is the study conducted by Berisha and Liss which looked at the differences in language use between two US presidents, Ronald Reagan (who would go on to receive a diagnosis of Dementia) and George H.W.
@@
 I filtered any annotations that were added to the transcript, including any references or clarifications, and any laughter.
-It's worth noting that there appears to be a difference in how 'hesitations' were marked down between each president, for RR hesitations were marked by a single hyphen whereas for GWB hesitations are marked by a double hyphen.
+It's worth noting that there appears to be a difference in how 'hesitations' were marked down between each president, for RR hesitations were marked by a single hyphen whereas for GHWB hesitations are marked by a double hyphen.
 In order to maintain consistency when parsing through the documents, I have changed both types of hesitation to be marked by a single hyphen.
@@
 I therefore was able to include all press conferences by both Ronald Reagan and George H.W.
-Bush where there was a question and answer session conducted at least in part by the sitting president (2 press conferences of GWB were ommited due to a lack of a question and answer session).
+Bush where there was a question and answer session conducted at least in part by the sitting president (2 press conferences of GHWB were ommited due to a lack of a question and answer session).
 \subsection{Feature Selection}
@@
 \section{Research Aims and Objectives}
-The aim of my research is to ultimately to identify less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to time-consuming and expensive psychological tests.
+The aim of my research is to ultimately to identify more accurate and efficient ways of detecting MCI and early dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRIs and EEGs) by utilising spontaneous speech.
 There is a lot of research into the analysis of language as a marker for MCI \cite{Taler2008, Boschi2017} and there have been calls for research that explores the use of Natural Language Processing and Machine Learning techniques to aid in the analysis of language \cite{Boschi2017}.
@@
 Given that sampling a person's language is relatively effortless, my proposed research will look at whether we can find markers of MCI in natural language.
-I plan to explore the use Natural Language Processing and Machine Learning techniques to assist in this process.
+I plan to explore the use of Natural Language Processing and Machine Learning techniques to assist in this process.
 There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia.
@@
 However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'.
-While this is helpful to establish that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments.
+While this is helpful to establish that machine learning can be used to differentiate between healthy controls and those with dementia, there has been little work looking at those in the MCI phase although it is apparent that language is still affected \cite{Berisha2015, Snowdon1996, Le2011}.
 Therefore we would like to clarify whether natural language generated outside the framework of a deliberate language generating task, has the same ability to discriminate.
@@
 \begin{itemize}
-	\item \textbf{Hypothesis One} Is monitoring language samples and other measures of cognitive decline over the course of the year enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment.
+	\item \textbf{Hypothesis One} Monitoring language samples and other measures of cognitive decline over the course of the year will be enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment
 	\item \textbf{Hypothesis Two} Those with MCI will show a different language profile to controls if matched similarly for age, gender, education and other significant factors.
-	\item \textbf{Hypothesis Three} The new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures.
+	\item \textbf{Hypothesis Three} The new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures such as the RBANS.
 	\item \textbf{Hypothesis Four} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, those with Alzheimers Disease and decline in those with MCI.
+	\item \textbf{Hypothesis Five} Using individuals as their own control will be enough to identify trends in the decline of their use of language.
 \end{itemize}
@@
 \subsection{Task 2 - Pilot Study for the Neuropsychological Battery of Tests}
-The aim of this study is complete a trial run of the administration of a new battery of Neuropsychological Battery of Tests.
+The aim of this study is complete a trial run of the administration of a new battery of neuropsychological battery of tests.
 We aim to find out whether the tests in question generate enough language so that features such as those generated in Task 1 are meaningful and can be used as language markers that can then be tracked over time.
@@
 \par 
-We have recruited a cohort of 4 volunteers who will complete the battery of Neuropsychological tests on three separate days with a gap of at least seven days between each administration.
+We have recruited a cohort of 4 volunteers who will complete the battery of neuropsychological tests on three separate days with a gap of at least seven days between each administration.
 The results of which will compiled and analysed.
@@
 \par 
-While the study in itself is not valid for submission to a conference or journal, it is hoped that the results of the study will allow subsequent studies using the same or adjusted battery of tests to be published.
+While the pilot study in itself is not valid for submission to a conference or journal, it is hoped that the results of the study will allow subsequent studies using the same or adjusted battery of tests to be published.
 \subsection{Task 3 - Exploration of language with the DementiaBank dataset and Three Authors dataset.}
@@
 \par 
-We expect to see typical age related decline in our HOC group and more severe decline in our EA group.
+We expect to see no decline in our HOC group within this timeframe and more severe decline in our EA group.
 Once again, we will be interested in whether there is any measurable decline in our SMCI or DMCI experimental groups with our battery of tests.

### 2019-01-29 Source/QualifyingReport/QualifyingReport.tex
QR

@@
 In 2009, the Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
-A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease\cite{Taler2008}.
+A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
 As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages \cite{Prince2015} and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
-At the present time diagnosis is typically conducted at memory clinics by trained clinicians\cite{Boschi2017}.
+At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017}.
 I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
 \par
-There is a large body of research that looks at the decline in language in those with MCI and AD\cite{Taler2008, Boschi2017}.
-However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD\cite{Taler2008, Boschi2017}.
+There is a large body of research that looks at the decline in language in those with MCI and AD \cite{Taler2008, Boschi2017}.
+However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD \cite{Taler2008, Boschi2017}.
 Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
@@
 Whilst a definitive diagnosis of Dementia can only be produced at post-mortem, there are a number of clinical indicators from psychological perspective that can indicate dementia is present.
-From a clinical psychology perspective, those who have dementia demonstrate cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language and visuospatial deficits\cite{McKhann2011}.
+From a clinical psychology perspective, those who have dementia demonstrate cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}.
 In addition, these symptoms are often accompanied by emotional problems such as depression and behavioural difficulties.
@@
 The case of Iris Murdoch, who had a confirmed diagnosis of dementia, illustrates this theory well.
-Le et al \cite{Le2011}found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel when, it is theorized, the symptoms of AD manifested themselves more significantly.
+Le et al \cite{Le2011} found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel when, it is theorized, the symptoms of AD manifested themselves more significantly.
 If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
-However, one of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of dementia.
+One of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of dementia.
 Albert and his team have worked to define clinical criteria which professionals can use to diagnose MCI due to AD and differentiate this from age-associated memory impairment and age-associated cognitive decline.
@@
 
-However, whilst the authors claim that the RBANS is adequately sensitive in person's with MCI, other research points out that the RBANS has poor sensitivity in detecting MCI.
+However, whilst the authors claim that the RBANS is adequately sensitive in person's with MCI \cite{Randolph1998} and there is some research to support this assertion \cite{Karantzoulis2013}, other research points out that the RBANS has poor sensitivity in detecting MCI \cite{Duff2010}.
 Another drawback of this battery is the lack of executive function measures and object naming tasks.
@@
 % Muangpaisan - Digit span and verbal fluency tests in patients with MCI and normal subjects in Thai community.
-Research has consistently shown that MCI patients have a significantly lower digit span score in both normal and reverse order versions of this test and a study by Muangpaisan has shown that the reverse order version of this test can, to some degree, predict a diagnosis of MCI.
+Research has consistently shown that MCI patients have a significantly lower digit span score in both normal and reverse order versions of this test and a study by Muangpaisan has shown that the reverse order version of this test can, to some degree, predict a diagnosis of MCI \cite{Muangpaisan2008}.
 Muangpaisan also found that Age, Gender and Education have an impact on the performance of the tests.
@@
 Participants are given a list of 15 unrelated words, repeated over five different trials and are asked to repeat.
-Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then repeat it again after 30 minutes \cite{Rey1941}.
+Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then repeat it again after 30 minutes \cite{Schmidt1996}.
 \par
@@
 The reproduction of such a complex structure involves processes such as planning and organizational strategies that are related to executive functions.
+In order to make this task repeatable, Taylor et al designed a comparable set of figures which have proven to be of equivalent difficulty \cite{Taylor1969, Strauss1990}.
 \par

### 2019-01-29 Source/QualifyingReport/QualifyingReport.tex
more edits

@@
 \section{How do we analyse language, issues and debates}
+There are a wide variety of approaches that we can take in the analysis of language and there is no real consensus on the best approach to this particular problem.
+This section looks at the different ways in which researchers have looked at the problem as well as discussing some of the areas of contention.
 \subsection{Single Word Language tasks vs Connected Language tasks}
@@
 From a research methodology perspective, using a standardized test allows researchers to target a very specific process in language generation and isolate factors that impact performance in language well.
-However, this approach does not take into account the symbiotic nature of language processes.
+However, this approach does not take into account other cognitive processes that are used in lengthier speech tasks.
+More specifically, single word language tasks do not look at the interaction between language, executive functions and reasoning abilities.
+They also do not require the logical and efficient organisation of ideas.
 \par
@@
 An example of a semantic complexity measure is the concept of idea density.
-Formally, idea density is defined as the average number of propositions per sentence \cite{Kintsch1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}.
+Formally, idea density is defined as the average number of propositions per sentence \cite{Keenan1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}.
 Some examples of semantic content measures are Type to Token Ratio, Brunet's Index and Honore's Statistic (described below) which are used to measure the lexical diversity of a given piece of text and/or utterance.
@@
 \end{equation}
-\subsubsection{Brunet's Index(W)} % Don't forget to cite the original article.
-Brunet's Index (W) differentiates itself for TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
-\begin{equation} \label{x3}
-W = N^{V(-0.165)}
-\end{equation}
-where N is the total length of the utterance being measured and V is equal to the total vocabulary being used by the subject.
-Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers.
-\newline
-\par
-\subsubsection{Honore's Statistic (R)}
-Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation:
-\begin{equation} \label{x4}
-%% Check to see if this is accurate.
-R = (100 \log N) / (1 - V1/V)
-\end{equation}
-where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
 \subsection{Picture-related content thematic elements}
@@
 (2001) examined Cookie Theft picture descriptions at four stages of AD.
-They identified 12 elements of the cookie theft picture and found that whilst each AD group differed significantly from the others and from controls, the measures did not distinguish between MCI and normal controls.
+They identified 12 elements of the cookie theft picture and found that whilst each AD group differed significantly from the others and from controls, the measures did not distinguish between MCI and normal controls \cite{Bschor2001}.
 \par
@@
 Snowdon examined written discourse from the Nun study and found that low idea density in early life was associated with reduced cognitive performance in later life \cite{Snowdon1996}.
-Riley et al extended these findings by concluding that early-life idea density was associated with lower brain weight, higher degree of cerebral atrophy and increased neurofibrillary pathology in later life.
+Riley et al extended these findings by concluding that early-life idea density was associated with lower brain weight, higher degree of cerebral atrophy and increased neurofibrillary pathology in later life \cite{Riley2005}.
 \par
@@
 Several studies report that adults with moderate AD produce fewer words than controls on picture description, however other studies found no differences in total words among groups of controls and patients with MCI or AD.
-Murray and Nicholas et al investigated normal controls, patients with AD and older adults with depression and found no group differences in total words.
-In contrast, Lira 2014 found that controls produced more total words than patients with AD but found no difference between mild and moderate groups.
+Murray and Nicholas et al investigated normal controls, patients with AD and older adults with depression and found no group differences in total words \cite{Nicholas1985}.
+In contrast, Lira 2014 found that controls produced more total words than patients with AD but found no difference between mild and moderate groups \cite{Lira2014}.
 \par
@@
 Syntax can be defined as the rules that govern how words can be combined to form sentences, whilst Morphology is the system that governs the structure of words and the construction of word forms.
-Multiple studies of language decline in dementia included at least one measure of syntax or syntactic complexity.
+Multiple studies of language decline in dementia included at least one measure of syntax or syntactic complexity \cite{Orimaye2017}.
 Common constructs included words per clause, grammatical form (measures of an appropriate use of syntactic conjections, tenses, conditionals, subordinate clauses and passive constructions), measures of phrase length and proportions of words in sentences.
@@
 The premise behind the Linguistic Inquiry and Word Count (LIWC) is the words we speak inform an observer of the state of mind of an individual.
-To this end, the creators have constructed a framework which uses a dictionary of words which are pre-categorised.
+To this end, the creators have constructed a framework which uses a dictionary of words which are pre-categorised \cite{Pennebaker2015}.
 % Al-Mosaiwi - In an Absolute State: Elevated Use of Absolutist Words Is a Marker Specific to Anxiety, Depression, and Suicidal Ideation 
@@
 There has been a significant research in the area of language deterioration as a means of detecting Alzheimer's Disease.
-This usually takes the form analysis of speech recorded as part of a cognitive assessment such as the Picture Description Task \cite{Orimaye2008}.
+This usually takes the form analysis of speech recorded as part of a cognitive assessment such as the Picture Description Task \cite{Orimaye2014,Fraser2015}.
 Given that language samples are relatively easy to collect, research has moved towards analysis of spontaneous speech.
@@
 These significant differences were in the number of unique words used per speech, the use of non-specific nouns and fillers and low-imageability verbs \cite{Berisha2015}.
-/par   
+\par   
 This study replicates work done by Berisha and Liss and extends this by adding Donald Trump as an alternative, more appropriate comparison to Ronald Reagan as he is much closer in age than George H.W.
@@
 \section{Results}
+\begin{table}
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{3cm} | p{3cm} | p{3cm} |}
+		\hline
+		& RR & GHWB & DJT \\ \hline
+		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
+		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
+		LI Verbs & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
+
+	\end{tabular}
+	\caption{\label{tab:table-name}NLTK tasks and functionality}
+	\end{center} 
+\end{table}
 \section{Discussion}
@@
 In this section I discuss the ethical considerations related to this project and any adjustments that I have made to my research because of these considerations.
-I am aware of the code of human  research ethics published by the British Psychological Society \cite{Society2012} and my research has been designed in a way that fits with these guidelines.
+I am aware of the code of human  research ethics published by the British Psychological Society \cite{BritishPsychologicalSociety2014} and my research has been designed in a way that fits with these guidelines.
 \subsection{Risk}

### 2019-01-30 Source/QualifyingReport/QualifyingReport.tex
stats and QR update

@@
 W.
-Bush (GHWB) and 25 press conferences conducted by Donald J.
+Bush (GHWB) and 29 press conferences conducted by Donald J.
 Trump (DJT).
@@
 Imageability is characterized, according to Berisha and Liss, as the ease with which a term gives rise to a sensory .mental image.
-I compared the trends described in the transcripts of RR and GHWB, but also included DT.
+I compared the trends described in the transcripts of RR and GHWB, but also included DJT.
 Berisha and Liss originally made the comparison as it GHWB (GHWB - age at the start of presidency - 64 years, 222 days) was the closest match to RR in terms of age (RR  - age at the start of presidency, 69 yeaars and 349 days).
@@
 \subsection{Feature Selection}
+We calculating the following features for each transcript in turn using the NLTK (see section 2 for a description) \cite{Bird2009} and Python. 
+\par 
+\subsubsection{Measures of lexical variation}
+% CITE Richards 1986 Type/Token Ratios: What do they really tell us?
+We constructed two features of lexical variation.
+Firstly we looked at the number of unique words.
+To do this we were able to split each transcript into individual words and changed them to lowercase using NLTK and were then count the number of unique words that appeared in each transcript.
+We also used the TTR formula (see section 2 for a description) for a feature that measures lexical diversity independent of sample size \cite{}.
+\subsubsection{Fillers, Non-Specific Nouns and LI Verbs}
+For these features, we counted the number of occurrences for different categories (see table for list of categories tracked and the words counted.
+
+\begin{table}
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{6cm} |}
+		\hline
+		Category & Words \\ \hline
+		Fillers & \textit{"well", "so", "basically", "actually", "literally", "um", "ah"} \\ \hline
+		Non Specific Nouns & \textit{"something", "anything", "thing", "everything"} \\ \hline
+		LI Verbs & \textit{"be", "come", "do", "get", "give", go", "know", "look", "make", "see", "tell", "think", "want"} \\ \hline
+
+	\end{tabular}
+	\caption{\label{tab:table-name}Categories and Words Counted}
+	\end{center} 
+\end{table}
+
+\subsubsection{Usage of parts of speech}
+This section involves using a Part of Speech tagger (PoS) which analyses a sentence and assigns a 'tag' to each word based on the function the word has in a sentence.
+At a basic level this can be divided into the eight defined parts of speech: 'nouns', 'pronouns', verbs', 'adjectives', 'adverbs', 'conjunctions', 'prepositions' and 'interjections' but can be further subcategorised.
+We used the PoS tagger built into NLTK to tag each transcript in turn and used these the counts from each of these eight categories in our analysis.
+In addition to frequency counts we also normalised these features by dividing the frequency count by the number of words in the document to take into account transcript length.
+
 \section{Results}
+\begin{table}
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{3cm} | p{3cm} | p{3cm} |}
+		\hline
+		& RR & GHWB & DJT \\ \hline
+		Total Words & 3423.91 (416.42) & 2607.72 (1210.38) & 1848.65 (1549.38) \\ \hline
+		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
+		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
+		LI Verbs & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
+	\end{tabular}
+	\caption{\label{tab:table-name}Means and Standard Deviations of important features}
+	\end{center} 
+\end{table}
+
 \begin{table}
@@
 	\end{tabular}
-	\caption{\label{tab:table-name}NLTK tasks and functionality}
+	\caption{\label{tab:table-name}RR T-tests vs GWB and DJT}
 	\end{center} 
 \end{table}
+
+\begin{table}
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{3cm} | p{3cm} | p{3cm} |}
+		\hline
+		& RR & GHWB & DJT \\ \hline
+		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
+		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
+		LI Verbs & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
+		Nouns Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
+		Verbs Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
+		Adjectives Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
+		Adverbs Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
+		Pronouns Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
+		Conjunctions Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
+		Prepositions Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
+		Interjections Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
+
+	\end{tabular}
+	\caption{\label{tab:table-name}Pearson Correlations for Features}
+	\end{center} 
+\end{table}
+
 \section{Discussion}
+% Cite Garrard 2005 -The effects of very early Alzheimer's disease on the characteristics of writing by a renowned author.
+
+Our research supports the findings of Berisha and Liss with regards to unique words, non-specific nouns and low-imageability words\cite{Berisha2015}.
+\par 
+A criticism of Berisha and Liss's work is the problems they had with normalising the transcripts in terms of length.
+This was also a problem in the work of Garrard et al.
+Whilst it is important to control for outliers, there are other ways in which we can control for length of sample.
+In this study, we explored features such as TTR, and also normalised versions of the frequency statistics we used, and also dividing the frequency counts by 100 which and these showed similar significant results.
+
+ In addition, when adding DJT as an additional control to compare to and despite a comparatively smaller sample per press conference held, their findings also hold true. 
+\par 
+It's worth noting that DJT's conferences are, with the exception of one press conference, held jointly with other people and therefore the amount of speech produced per conference is a lot less.
+This finding is, in itself notable as it illustrates the problems with amounts of speech.
+This problem could potentially be solved with
+\par 
+There are limitations of this research.
+Whilst in terms of age, DJT is certainly more suitable as a control to match with RR, in some ways they held very different styles of press conferences in that RR preferred to do solo press conferences and DJT has shown a preference for doing joint press conferences which have an impact on the amount of language produced.
+Another limitation is the way in which I normalised the features.
+\par 
+With further work, we will not have the luxury of having multiple samples and so it would be worth exploring how the quality of these predictions might lessen when faced with considerably fewer samples.
 \section{Conclusion}
+The results of this work show that we can track a person's use of language through time in a number of ways, and that
 

### 2019-01-30 Source/QualifyingReport/QualifyingReport.tex
edits to QR

@@
 \par
-% Muangpaisan - Digit span and verbal fluency tests in patients with MCI and normal subjects in Thai community.
 Research has consistently shown that MCI patients have a significantly lower digit span score in both normal and reverse order versions of this test and a study by Muangpaisan has shown that the reverse order version of this test can, to some degree, predict a diagnosis of MCI \cite{Muangpaisan2008}.
@@
 Pascoe compared patients with PD with Normal Cognition (PD-N) with those with PD and MCI (PD-MCI) and healthy participants.
-PD-MCI pariticipants achieved significantly lower scores than other groups in the DSST task \cite{Pascoe2018}.
+PD-MCI participants achieved significantly lower scores than other groups in the DSST task \cite{Pascoe2018}.
 \subsection{Verbal Fluency}
@@
 % CITE - Osterrieth - Filetest de copied’une figure complex: contribution al’etude de la perception et de la memoire [The test of copying a complex figure: a contribution to the study of perception and memory]
-% CITE - Taylor 1969 1979 - Variants
-% CITE - Strauss - A comparison of the Rey and Taylor figures
 The Rey-Osterrieth Complex Figure (ROCF) is a task widely as a test of visuo-spatial skill and visual memory.
@@
 This is usually tested using Stroop test, however this has the drawback of lacking ecological validity.
-Therefore researchers have moved towards using the Hayling Sentence Completion Test (HCST) which uses skills such as word retreval as well as the ability to inhibit ones responses where appropriate and is therefore a much more ecologically valid task.
+Therefore researchers have moved towards using the Hayling Sentence Completion Test (HCST) which uses skills such as word retrieval as well as the ability to inhibit ones responses where appropriate and is therefore a much more ecologically valid task.
 There are two parts to the HCST.
@@
 \par 
-Martyr et al compared healthy controls with patients with dementia and patients with parkinson's disease \cite{Martyr2017}.
+Martyr et al compared healthy controls with patients with dementia and patients with Parkinson's disease \cite{Martyr2017}.
 They found that a high proportion of Category A errors (producing a word that fits the sentence when instructed otherwise) was a factor in performance loss for participants with dementia.
 Findings suggest that the HSCT may be sensitive to verbal suppression deficits and may provide insight into inhibitory control in participants with dementia.
-Patients with Dementia were significantly slower than controls in the initiation and inhibition conditions vs healthy controls, and slower than patients with parkinsons disease in the initiation but not the inhibition condition.
+Patients with Dementia were significantly slower than controls in the initiation and inhibition conditions vs healthy controls, and slower than patients with Parkinson's disease in the initiation but not the inhibition condition.
 \subsection{Grooved Pegboard Test}
@@
 \par 
-Darweesh et al, used the Purdue Pegboard Test to assess manual dexterity in a healthy older adult population and followed their participants for between eight and twelve years until an indication of the onset of a neurodegenerative disease \cite{Darweesh2017}.
+Darweesh et al, used the Purdue Pegboard Test (PPT) to assess manual dexterity in a healthy older adult population and followed their participants for between eight and twelve years until an indication of the onset of a neurodegenerative disease \cite{Darweesh2017}.
 In this time 227 (4\%) of their participants went on to develop a diagnosis of dementia.
@@
 \subsection{Free Cued Selective Reminding Test}
-% Describe the test
 The Free Cued Selective Reminding Test (FCSRT) was borne out of the premise that by controlling the conditions of learning, a measure of memory is possible that is not confounded by normal age related changes in cognition.
@@
 \subsection{Conciseness of information}
-% CITE RILEY!
 Conciseness has been defined as the number of words a speaker uses to express ideas.
@@
 \subsubsection{Natural Language Toolkit}
-% CITE Bird, Steven, Edward Loper and Ewan Klein (2009). Natural Language Processing with Python.  O'Reilly Media Inc.
 The Natural Language Toolkit (NLTK) was originally designed as part of a computational linguistics course but has evolved to become an open-source framework which includes methods and modules for a wide array of Natural Language tasks (see Table 2.2 for a full list) \cite{Bird2009}.
@@
 \subsubsection{Fillers, Non-Specific Nouns and LI Verbs}
+% CITE Bird et al 2000 The Rise and Fall of Frequency and Imageability Noun and Verb Production in Semantic Dementia
 For these features, we counted the number of occurrences for different categories (see table for list of categories tracked and the words counted.
@@
 		& RR & GHWB & DJT \\ \hline
-		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
-		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
-		LI Verbs & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
-		Nouns Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
-		Verbs Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
-		Adjectives Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
-		Adverbs Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
-		Pronouns Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
-		Conjunctions Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
-		Prepositions Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
-		Interjections Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
-
+		Unique Words & -0.58 & -0.25 & 0.16 \\ \hline
+		Non Specific Nouns & 0.25 & -0.08 & -0.03 \\ \hline
+		LI Verbs & -0.23 & -0.20 & 0.02 \\ \hline
+		Nouns Normalised & -0.692 & -0.03 & 0.02 \\ \hline
+		Verbs Normalised & 0.32 & 0.24 & -0.03 \\ \hline
+		Adjectives Normalised & -0.38 & 0.08 & -0.34 \\ \hline
+		Adverbs Normalised & 0.38 & 0.02 & -0.36 \\ \hline
+		Pronouns Normalised & 0.63 & 0.13 & 0.12 \\ \hline
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize (\,) * denotes p < 0.05} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize [\,] ** denotes p < 0.01} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize [\,] *** denotes p < 0.001} \\
 	\end{tabular}
@@
 
-\appendix
-\section{Title of Appendix A}
-Text of Appendix A is Here
-
-\section{Title of Appendix B}
-Text of Appendix B is Here
-
-\section{Title of Appendix C}
-Text of Appendix C is Here.
-
 \end{document}

### 2019-01-31 Source/QualifyingReport/QualifyingReport.tex

@@
 We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English.
-\begin{table}
+\begin{table}[H]
 	\begin{center}
@@
 \subsection{Rey Auditory-Verbal learning test}
-Rey's Auditory Verbal Learning Test (RAVLT+) looks a wide range of neuropsychological processess including short-term auditory-verbal memory, learning and retention of information.
+Rey's Auditory Verbal Learning Test (RAVLT+) looks a wide range of neuropsychological processes including short-term auditory-verbal memory, learning and retention of information.
 Participants are given a list of 15 unrelated words, repeated over five different trials and are asked to repeat.
@@
 One common way to do this is the Boston Naming Test (BNT) which comprises 60 items on a spectrum of very frequent to very infrequent.
-This has been reduced subsequently to two thirty item versions and four fifteen item versions, which correlate significantly with the original sixty item version and the benefit of the shorter versions of the test is that it facilitates testing of individuals with AD who may suffer from fatigue or limited attention span.
+This has been reduced subsequently to two thirty item versions and four fifteen item versions, which correlate significantly with the original sixty item version and the benefit of the shorter versions of the test is that it facilitates testing of individuals with AD who may suffer from fatigue or limited attention span \cite{Williams1989}.
 \par 
@@
 \subsection{Rey-Osterrieth Complex Figure Task}
-% CITE - Osterrieth - Filetest de copied’une figure complex: contribution al’etude de la perception et de la memoire [The test of copying a complex figure: a contribution to the study of perception and memory]
 The Rey-Osterrieth Complex Figure (ROCF) is a task widely as a test of visuo-spatial skill and visual memory.
-The task, which was originally designed by Rey (1944) and standardised by Osterrieth (1944) \cite{Rey1941}, asks a participant to copy a complex geometrical figure (known as the immediate copy condition and to recall and reproduce the figure from memory without warning (known as the delayed recall condition).
+The task, which was originally designed by Rey (1944) and standardised by Osterrieth (1944) \cite{Rey1941, Osterrieth1944}, asks a participant to copy a complex geometrical figure (known as the immediate copy condition and to recall and reproduce the figure from memory without warning (known as the delayed recall condition).
 In the immediate copy condition, the complexity of the figure requires an integrative cognitive ability.
@@
 Evidence shows that patients with disorders that possibly involve attention and executive functions are characterized by a more disorganised approach when copying the ROCF compared to controls.
-One of the difficulties with the presentation of this task in a repeated battery will be the fact that it turns from an incidental memory task (it is a memory task but the participant is not forewarned that they will be required to memorise the picture) into an intentional memory task (given the previous exposure to the task, it would be expected that a participant would pre-empt the delayed recall portion of the test and spend more time attempting to memorise the details) \cite{Teng2009}.
+One of the difficulties with the presentation of this task in a repeated battery will be the fact that it turns from an incidental memory task (it is a memory task but the participant is not forewarned that they will be required to memorise the picture) into an intentional memory task (given the previous exposure to the task, it would be expected that a participant would preempt the delayed recall portion of the test and spend more time attempting to memorise the details) \cite{Teng2009}.
 \subsection{Hayling Sentence Completion Test}
@@
 This test has also been shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the researchers to used the test to categorise prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT score \cite{Sarazin2007}.
-\subsubsection{Conclusions}
+\subsection{Conclusions}
 I have looked at a number of different cognitive tests and a battery of tests that aim to have high diagnostic and prognostic capabilities in the MCI population.
@@
 \par
-Connected language tasks, such as the picture description task and interviews described above are a much more reflective of processes involved in natural language generation.
+Connected language tasks, such as the picture description task and interviews described above are much more reflective of the processes involved in natural language generation.
 They are able to be relatively constrained in the language available to be used, for example in the picture description task, or they can be unconstrained in the form of interviews.
-% ADD MORE!
+But regardless of how they are framed, they are potentially much more useful in terms looking at all the processes involved in language generation with the drawback that you will not be able to isolate specific parts of the language generation process.
+A review by Boschi et al concludes that analysis of connected speech is potentially useful in guiding clinicians to identify language disorders \cite{Boschi2017} and also highlights the role of NLP and Machine learning in assisting in this endeavour.
 \par
@@
 Emery in her literature review looks at all levels of language tasks except for pragmatics, however this is an area that should not be ignored.
+Whilst the study of language from a pragmatic perspective is much more complex, it is perhaps one of the most vital areas to study because of the number of different cognitive processes involved.
 \par
@@
 Formally, idea density is defined as the average number of propositions per sentence \cite{Keenan1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}.
-Some examples of semantic content measures are Type to Token Ratio, Brunet's Index and Honore's Statistic (described below) which are used to measure the lexical diversity of a given piece of text and/or utterance.
+An example of semantic content measures is Type to Token Ratio (described below) which is used to measure the lexical diversity of a given piece of text and/or utterance.
 This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2013}.
@@
 % Al-Mosaiwi - In an Absolute State: Elevated Use of Absolutist Words Is a Marker Specific to Anxiety, Depression, and Suicidal Ideation 
-A typical example of this might be someone with depression who may typically be more inwardly focused and therefore talking in the first person is common. 
-% Sonnenschein - Linguistic analysis of patients with mood and anxiety disorders during cognitive behavioral therapy
-The 2015 version of the LIWC has over 90 different categories and some supporting research suggests that it is effective at identifying themes and associating these themes with mental health difficulties such as depression and anxiety.
+A typical example of this might be someone with depression who may typically be more inwardly focused and therefore talking in the first person is common\cite{Al-Mosaiwi2018}.
+
+The 2015 version of the LIWC has over 90 different categories and some supporting research suggests that it is effective at identifying themes and associating these themes with mental health difficulties such as depression and anxiety \cite{Sonnenschein2018}.
 \par
@@
 \par
-% Thomas - Automatic Detection and Rating of AD through the Lexical Analysis of Spontaneous Speech
 Thomas et al, detail several lexical approaches to the problem of detecting and rating AD.
@@
 50\% accuracy classifying dementia into four categories.
-They suggest further exploration of characteristics to classify.
-% Bucks et al - Analysis of spontaneous, conversational speech in dementia of the Alzheimer's Type: Evaluation of an objective technique for analyzing lexical performance.
+They suggest further exploration of characteristics to classify \cite{Thomas2005}.
+\par 
 AD patients had higher mean P rate, A rate and V rate scores but lower N rate scores vs normal older controls \cite{Bucks2000}.
@@
 
-% One of the main benefits of deep learning is that it does not rely on pre defined features being fed to the model, instead generating it's own features.
+One of the main benefits of deep learning is that it does not rely on pre defined features being fed to the model, instead generating it's own features.
 This move away from relying on features solves one of the difficulties with
@@
 
-\begin{table}
+\begin{table}[H]
 	\begin{center}
@@
 \section{Results}
-\begin{table}
+One of the most important thing to note is the wide variety of samples between the three presidents and also the varying timescales.
+RR participated in 46 press conferences over eight years (an average of 5.75 a year) which is the fewest number of press conferences given by an American president during their term of office.
+GHWB participated in 136 press conferences over four years (an average of  34 a year) and DJT participated in 29 press conferences to date (an average of 19.3 per year).
+Equally, there are variances in the average number of words.
+RR produced an average of 3424 words per conference compared to 2608 by GHWB (unpaired t = 4.434, p\textless 0.001) and DJT at 1849 words (unpaired t = 6.524, p\textless 0.001).
+\begin{table}[H]
 	\begin{center}
-	\begin{tabular}{ | p{3cm} | p{3cm} | p{3cm} | p{3cm} |}
+	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
 		\hline
@@
 		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
-		LI Verbs & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
+		LI Verbs & 124.22 (17.89)& 103.45 (51.87) & 84.48 (75.78) \\ \hline
 	\end{tabular}
@@
 
-\begin{table}
+In terms of unique words, we found that RR used significantly more unique words, non-specific nouns and low imageability verbs than GHWB and DJT (see Table 3.3).
+Some of these differences are due to the length of the sample, particularly in the case of DJT where his average sample is almost half the sample of RR.
+It could also be said that this could be down to differences in linguistic abilities or speaking style \cite{Berisha2015, Le2011}.
+However, we can certainly see that as controls GWHB and DJT are comparative in relation to non-specific nouns and LI verbs.
+
+\begin{table}[H]
 	\begin{center}
-	\begin{tabular}{ | p{3cm} | p{3cm} | p{3cm} | p{3cm} |}
+	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
 		\hline
-		& RR & GHWB & DJT \\ \hline
-		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
-		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
-		LI Verbs & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
-
+		& RR v GHWB & RR v DJT & GHWB v DJT \\ \hline
+		Total Words & 4.434*** & 6.524*** & 2.899** \\ \hline
+		Unique Words & 6.832*** & 11.403*** & 4.137*** \\ \hline
+		Non Specific Nouns & 7.877*** & 3.426** & -0.574 \\ \hline
+		LI Verbs & 2.656** & 3.420*** & 1.628 \\ \hline
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
 	\end{tabular}
@@
 
-\begin{table}
+We then looked at the data from a longitudinal perspective as we are interested seeing whether we can track various language variables and their progress over time.
+We ran a number of Pearsons correlations with transcript index number as a time reference and the dependant variables (Table 3.4).
+For our controls, we found them to be stable for the most part with the main highlights being a decrease in Adverb usage for DJT (R=-0.36, p=0.049) and a steady but not severe decline in a number of variables for GWHB, namely total word count, unique words, low imageability words and verb usage.
+\par 
+For RR, his decline is more marked and more widespread through his language use.
+We noticed an significant increase in adverb (R=0.41, p=0.004) and pronoun usage (R=0.65, p\textless0.001), as well as a slight usage increase in Non-specific nouns(R=0.30, p=0.03).
+There was a highly significant decrease in number of unique words (R=-0.56, p\textless0.001) and noun usage (R=-0.70, p\textless0.001).
+Also very significant decrease in adjective usage (R=-0.40, p=0.005) and a significant decrease in total word count (R=-0.31, p=0.03).
+
+\begin{table}[H]
 	\begin{center}
-	\begin{tabular}{ | p{3cm} | p{3cm} | p{3cm} | p{3cm} |}
+	\begin{tabular}{ | l | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
 		\hline
 		& RR & GHWB & DJT \\ \hline
-		Unique Words & -0.58 & -0.25 & 0.16 \\ \hline
-		Non Specific Nouns & 0.25 & -0.08 & -0.03 \\ \hline
-		LI Verbs & -0.23 & -0.20 & 0.02 \\ \hline
-		Nouns Normalised & -0.692 & -0.03 & 0.02 \\ \hline
-		Verbs Normalised & 0.32 & 0.24 & -0.03 \\ \hline
-		Adjectives Normalised & -0.38 & 0.08 & -0.34 \\ \hline
-		Adverbs Normalised & 0.38 & 0.02 & -0.36 \\ \hline
-		Pronouns Normalised & 0.63 & 0.13 & 0.12 \\ \hline
-		\multicolumn{4}{@{}p{1.5in}}{\footnotesize (\,) * denotes p < 0.05} \\
-    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize [\,] ** denotes p < 0.01} \\
-    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize [\,] *** denotes p < 0.001} \\
+		Word Count & \textbf{-0.31*} & \textbf{-0.21*} & 0.08 \\ \hline 
+		Unique Words & \textbf{-0.56***} & \textbf{-0.25**} & 0.16 \\ \hline
+		Non Specific Nouns & \textbf{0.30*} & -0.08 & -0.03 \\ \hline
+		LI Verbs & -0.19 & \textbf{-0.20**} & 0.02 \\ \hline
+		Nouns Normalised & \textbf{-0.70***} & -0.03 & 0.14 \\ \hline
+		Verbs Normalised & \textbf{0.36**} & \textbf{0.24***} & -0.03 \\ \hline
+		Adjectives Normalised & \textbf{-0.40**} & 0.08 & -0.34 \\ \hline
+		Adverbs Normalised & \textbf{0.41***} & 0.02 & \textbf{-0.36*} \\ \hline
+		Pronouns Normalised & \textbf{0.65***} & 0.13 & 0.07 \\ \hline
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
 	\end{tabular}
@@
 \section{Discussion}
-% Cite Garrard 2005 -The effects of very early Alzheimer's disease on the characteristics of writing by a renowned author.
-
-Our research supports the findings of Berisha and Liss with regards to unique words, non-specific nouns and low-imageability words\cite{Berisha2015}.
-\par 
+President Reagan received his diagnosis of AD in August 1994 but using transcripts of speeches he made in his two terms as President (January 1981 - January 1989) we have be able to identify certain changes in his use of language that we might ascribe to the onset of MCI and early AD.
+Despite differences in our methodology, our research supports the findings of Berisha and Liss in that we both find a significant decrease in unique words over time and an increase in non-specifc noun usage.
+Compared to our controls (GWHB andDJT), we find some slight trends with GWHB but no such trends with DJT in his speech albeit his samples of speech span a shorter amount of time.
+\begin{figure}[H]
+	\centering
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/RRUniqueWords.png}
+		\caption{Ronald Reagan - Unique Words over time}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/GWHBUniqueWords.png}
+		\caption{George H.W. Bush - Unique Words over time}
+	\end{minipage}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/RRNSNouns.png}
+		\caption{Ronald Reagan - Non-specifc Nouns over time}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/DJTNSNouns.png}
+		\caption{Donald J. Trump - Non-specific Nouns over time}
+	\end{minipage}
+\end{figure}
+
 A criticism of Berisha and Liss's work is the problems they had with normalising the transcripts in terms of length.
-This was also a problem in the work of Garrard et al.
+This was also a problem in the work of Garrard et al \cite{Garrard2005, Le2011}.
 Whilst it is important to control for outliers, there are other ways in which we can control for length of sample.
-In this study, we explored features such as TTR, and also normalised versions of the frequency statistics we used, and also dividing the frequency counts by 100 which and these showed similar significant results.
-
- In addition, when adding DJT as an additional control to compare to and despite a comparatively smaller sample per press conference held, their findings also hold true. 
-\par 
-It's worth noting that DJT's conferences are, with the exception of one press conference, held jointly with other people and therefore the amount of speech produced per conference is a lot less.
-This finding is, in itself notable as it illustrates the problems with amounts of speech.
-This problem could potentially be solved with
-\par 
+\par 
+Interestingly, when we normalised the various types of words used by the presidents we found some interesting patterns that further differentiated RR from the controls.
+Whilst Non-specific nouns increased over time, we found that noun usage in general significantly decreased and pronouns increased similarly significantly.
+The increase in pronoun for those with early AD has been identified in literature, although there are only a few studies that explore this \cite{Almor1999, Wendelstein2015}.
+Wendlestein et al propose that the increased used of pronouns is an expression of an impaired ability to adapt language to the listener's needs \cite{Wendelstein2015}.
+Almor et al attributed this reliance on pronouns due to a impaired working memory \cite{Almor1999}.
+\par 
+The decrease in overall noun usage has also been identified as a feature.
+Jarrold et al found that AD patients would use more pronouns, verbs and fewer nouns than controls \cite{Jarrold2014}.
+Wendlestein in their investigations into noun usage found that decreased later on in AD progression and was unaffected in the pre-clinical stages of AD \cite{Wendelstein2014}.
+Our results are supported by existing literature and this potentially means that language analysis in the way we have structured it may have diagnostic or prognostic properties.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=240px, height=150px]{images/RRNounsNormalised.png}
+	\caption{Ronald Reagan - Nouns Normalised over time}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=240px, height=150px]{images/RRPronouns.png}
+	\caption{Ronald Reagan - Pronouns Normalised over time}
+\end{figure}
+
 There are limitations of this research.
 Whilst in terms of age, DJT is certainly more suitable as a control to match with RR, in some ways they held very different styles of press conferences in that RR preferred to do solo press conferences and DJT has shown a preference for doing joint press conferences which have an impact on the amount of language produced.
-Another limitation is the way in which I normalised the features.
-\par 
-With further work, we will not have the luxury of having multiple samples and so it would be worth exploring how the quality of these predictions might lessen when faced with considerably fewer samples.
-\section{Conclusion}
-The results of this work show that we can track a person's use of language through time in a number of ways, and that
+This finding is, in itself notable as it illustrates the problems with small amounts of speech.
+\par
+The problem of finding an appropriate control is a common one in this domain.
+Given that those with MCI and early dementia have such variable presentations, it might prove of limited value in matched pairs design.
+\par 
+With further work, we will not have the luxury of having multiple samples and so it would be worth exploring how the quality of these predictions might lessen when faced with considerably fewer samples and over a smaller time period.
+In addition, this replication and extension has demonstrated the potential utility of using longitudinal data as a means of comparing language use of a person at two or more time periods and using this information as a diagnostic aid for MCI.
+\par 
+The results of this work show that we can track a person's use of language through time in a number of ways, and th
 
@@
 Nevertheless, a higher than average drop out rate would weaken the strength of the study, especially if those who dropped out came from the experimental group.
-
 \subsection{Commonality of MCI}
-% Roberts - Higher risk of progression to dementia in mild cognitive impairment cases who revert to normal.
-Whilst research has consistently shown that those with a diagnosis of MCI tend to have a significantly higher risk of progression to dementia (CITE Roberts).
+Whilst research has consistently shown that those with a diagnosis of MCI tend to have a significantly higher risk of progression to dementia \cite{Roberts2014}.
 The progression at any given time is low and variable.
-% Farias - Progression of Mild Cognitive Impairment to Dementia in Clinic- vs Community Based Cohorts
-Annual conversion rates vary across longitudinal studies conducted in this population, but researchers have shown that there is a higher conversion rate in clinical samples of this population than in community samples.
+Annual conversion rates vary across longitudinal studies conducted in this population, but researchers have shown that there is a higher conversion rate in clinical samples of this population than in community samples \cite{Farias2009}.
 In my proposed longitudinal study, I aim to collect a clinical MCI population, but depending on the availability of such samples I may have to collect from the community.
 If this is the case then, the prevalence of those with MCI who have language deficits within my participants may be low.
-
 \subsection{Sparsity of Language}
@@
 I will be working with a middle-aged to elderly population who have identified in some way that there is a difficulty with cognition.
-However, I do not anticipate working with vulnerable groups, those lacking capacity or individuals in a dependant or unequal relationship.
+However, I do not anticipate working with vulnerable groups, those lacking capacity or individuals in a dependent or unequal relationship.
 Should any participants I am working with lose capacity during the course of the study, I will remove that participants details and data from the study.

### 2019-01-31 Source/QualifyingReport/QualifyingReport.tex

@@
 \begin{equation} \label{x1}
-Cue efficiency = (totalRecall-freeRecall)/48-freeRecall, range 0.0-1.0)
+CueEfficiency = (totalRecall-freeRecall)/48-freeRecall, range 0.0-1.0)
 \end{equation}
@@
 \section{State of literature into Machine Learning and Natural Language processing techniques}
-As we can see, diagnosing dementia through analysis has a large background in terms of psychological research.
-However, increasingly current research has called for the use of machine learning as a way of assisting in the process of diagnosis \cite{Boschi2017}.
-This section will look at what areas of machine learning could potentially be used as tools to help in this domain as well as any research that has applied machine learning to this problem.
+Diagnosing dementia through language analysis has a large background in terms of psychological research.
+Increasingly current research has called for the use of machine learning as a way of assisting in the process of diagnosis \cite{Orimaye2017, Boschi2017}.
+This section will look at what areas of natural language processing (NLP) and machine learning (ML) could potentially be used as tools to help in this domain as well as any research that has applied machine learning to this problem.
 \subsection{Natural Language Processing}
-Natural Language Processing as a topic can be defined as the intersection between Machine Learning and Linguistics and looks at
-
+Natural Language Processing as a topic can be defined as the intersection between Machine Learning and Linguistics and looks at a number of language tasks, particularly important in this domain is how to enable a computer to process and analyse large amounts of language data.
+Some other tasks include speech recognition, natural language understanding and natural language generation.
 Generally speaking, Natural Language Processing (NLP) systems mimic the semiotic perspective on language.
-That is to say that we can subdivide NLP processes into different components (see Figure 2.3) and model broadly fits with the literature review conducted by Emery \cite{Emery2000}.
+That is to say that we can subdivide NLP processes into different components (see Figure 2.3).
 \begin{enumerate}
@@
 \end{figure}
-
+In order to facilitate these and other tasks, a number of frameworks have been developed that automate some of the more common processing tasks. A brief review of these frameworks follows. 
 \subsubsection{Natural Language Toolkit}
-The Natural Language Toolkit (NLTK) was originally designed as part of a computational linguistics course but has evolved to become an open-source framework which includes methods and modules for a wide array of Natural Language tasks (see Table 2.2 for a full list) \cite{Bird2009}.
-
+The Natural Language Toolkit (NLTK) was originally designed as part of a computational linguistics course at the University of Pennsylvania but has evolved to become an open-source framework which includes methods and modules for a wide array of Natural Language tasks (see Table 2.2 for a full list) \cite{Bird2009}.
 \begin{table}
@@
 \end{table}
+The NLTK has been successfully used in both teaching and research purposes and as part of this provides access to large array of databases and corpora which have potential uses in common tasks.
 
 \subsubsection{CoreNLP}
-The stanford parser is
-
 Stanford CoreNLP provides a set of human language technology tools.
 It can give the base forms of words, their parts of speech, whether they are names of companies, people, etc., normalize dates, times, and numeric quantities, mark up the structure of sentences in terms of phrases and syntactic dependencies, indicate which noun phrases refer to the same entities, indicate sentiment, extract particular or open-class relations between entity mentions, get the quotes people said, etc.
-
+\par
+\begin{figure}[H]
+\centering
+\includegraphics[width=350px, height=150px]{images/CoreNLP.jpg}
+\caption{How Natural Language Processing tasks are subdivided.\label{white}}
+\end{figure}
 Stanford CoreNLP’s goal is to make it very easy to apply a bunch of linguistic analysis tools to a piece of text.
@@
 To this end, the creators have constructed a framework which uses a dictionary of words which are pre-categorised \cite{Pennebaker2015}.
-% Al-Mosaiwi - In an Absolute State: Elevated Use of Absolutist Words Is a Marker Specific to Anxiety, Depression, and Suicidal Ideation 
 A typical example of this might be someone with depression who may typically be more inwardly focused and therefore talking in the first person is common\cite{Al-Mosaiwi2018}.
-
 The 2015 version of the LIWC has over 90 different categories and some supporting research suggests that it is effective at identifying themes and associating these themes with mental health difficulties such as depression and anxiety \cite{Sonnenschein2018}.
@@
 \begin{itemize}
-	\item \textbf{Hypothesis One} Monitoring language samples and other measures of cognitive decline over the course of the year will be enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment
-	\item \textbf{Hypothesis Two} Those with MCI will show a different language profile to controls if matched similarly for age, gender, education and other significant factors.
-	\item \textbf{Hypothesis Three} The new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures such as the RBANS.
-	\item \textbf{Hypothesis Four} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, those with Alzheimers Disease and decline in those with MCI.
-	\item \textbf{Hypothesis Five} Using individuals as their own control will be enough to identify trends in the decline of their use of language.
+	\item \textbf{H1} Monitoring language samples and other measures of cognitive decline over the course of the year will be enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment
+	\item \textbf{H2} Those with MCI will show a different language profile to controls if matched similarly for age, gender, education and other significant factors.
+	\item \textbf{H3} The new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures such as the RBANS.
+	\item \textbf{H4} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, those with Alzheimers Disease and decline in those with MCI.
+	\item \textbf{H5} Using individuals as their own control will be enough to identify trends in the decline of their use of language.
 \end{itemize}

### 2019-01-31 Source/QualifyingReport/QualifyingReport.tex
added GanttChart

@@
 Stanford CoreNLP provides a set of human language technology tools.
-It can give the base forms of words, their parts of speech, whether they are names of companies, people, etc., normalize dates, times, and numeric quantities, mark up the structure of sentences in terms of phrases and syntactic dependencies, indicate which noun phrases refer to the same entities, indicate sentiment, extract particular or open-class relations between entity mentions, get the quotes people said, etc.
+This comes as an installable package that requires JAVA and can be interacted with through the command line or by an accompanying API.
+Whilst NLTK is much more oriented to parsing text and generating features, CoreNLP does this and adds more functionality such as a Named Entity Recogniser and sentiment analysis functions.
+At it's centre is the Stanford parser, which parses text and can give the base forms of words, their parts of speech, whether they are names of companies, people, etc., normalize dates, times, and numeric quantities, mark up the structure of sentences in terms of phrases and syntactic dependencies, indicate which noun phrases refer to the same entities, indicate sentiment, extract particular or open-class relations between entity mentions, get the quotes people said, etc.
 \par
@@
 \centering
-\includegraphics[width=350px, height=150px]{images/CoreNLP.jpg}
-\caption{How Natural Language Processing tasks are subdivided.\label{white}}
+\includegraphics[width=350px, height=200px]{images/CoreNLP.jpg}
+\caption{Depiction of how Core NLP marks up a sentence}
 \end{figure}
-Stanford CoreNLP’s goal is to make it very easy to apply a bunch of linguistic analysis tools to a piece of text.
-A tool pipeline can be run on a piece of plain text with just two lines of code.
-CoreNLP is designed to be highly flexible and extensible.
-With a single option you can change which tools should be enabled and disabled.
-Stanford CoreNLP integrates many of Stanford’s NLP tools, including the part-of-speech (POS) tagger, the named entity recognizer (NER), the parser, the coreference resolution system, sentiment analysis, bootstrapped pattern learning, and the open information extraction tools.
-Moreover, an annotator pipeline can include additional custom or third-party annotators.
-CoreNLP’s analyses provide the foundational building blocks for higher-level and domain-specific text understanding applications.
+It's intended for CoreNLP to provide tools and frameworks that will enable higher level language analysis whilst remaining domain agnostic.
 
 \subsubsection{Linguistic Inquiry and Word Count}
-The premise behind the Linguistic Inquiry and Word Count (LIWC) is the words we speak inform an observer of the state of mind of an individual.
-To this end, the creators have constructed a framework which uses a dictionary of words which are pre-categorised \cite{Pennebaker2015}.
+The Linguistic Inquiry and Word Count (LIWC) is a piece of software that counts words and assigns these counts to categories \cite{Pennebaker2015}.
+The rationale behind this is that the words we speak inform an observer of the state of mind of an individual.
 A typical example of this might be someone with depression who may typically be more inwardly focused and therefore talking in the first person is common\cite{Al-Mosaiwi2018}.
-The 2015 version of the LIWC has over 90 different categories and some supporting research suggests that it is effective at identifying themes and associating these themes with mental health difficulties such as depression and anxiety \cite{Sonnenschein2018}.
-\par
-A framework like this could be used as another tool to explore themes within a persons speech.
+The 2015 version of the LIWC has over 90 different categories \cite{Pennebaker2015} and some supporting research suggests that it is effective at identifying themes and associating these themes with mental health difficulties such as depression and anxiety \cite{Sonnenschein2018}.
+\par
+Whilst not necessarily a tool that can parse data and look at trends from a linguistic perspective, a framework like this could be used as another tool to explore themes within a persons speech from a more qualitative angle.
+To the authors knowledge, there is no research that looks at speech changes from this perspective.
 
 \subsection{Traditional methods of Machine Learning}
-
 In terms of Machine Learning research in this area, a number of researchers have used transcripts based on picture description tasks \cite{Zimmerer2016, Orimaye2017, Mueller2018a, Fraser2015} and have successfully extracted linguistic features that could differentiate between AD and controls.
 \par
-Thomas et al, detail several lexical approaches to the problem of detecting and rating AD.
-The approaches rely on character n-gram techniques but also explore correlation of usage of frequency of different types of speech.
+One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the most frequent class.
+They detail several lexical approaches to the problem of detecting and rating AD.
+The approaches they looked at relied primarily on character n-gram techniques but also explore correlation of usage of frequency of different types of speech.
 Their results act as a proof in concept of the utility of using a pure computational approach to the diagnosis of dementia using spontaneous speech.
-Results - 95\% accuracy dementia vs controls.
-70\% accuracy classifying dementia into two categories.
-50\% accuracy classifying dementia into four categories.
+The were able to obtain 95\% accuracy dementia vs controls, 70\% accuracy classifying dementia into two categories and 50\% accuracy classifying dementia into four categories.
 They suggest further exploration of characteristics to classify \cite{Thomas2005}.
 \par 
-AD patients had higher mean P rate, A rate and V rate scores but lower N rate scores vs normal older controls \cite{Bucks2000}.
-One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class.
+% Cite CRUM Population-based norms for the Mini-Mental Status Examination by age and educational level
+Bucks et al \cite{Bucks2000} took a more traditional approach using frequency rates per 100 words of various word types such as nouns, verbs, adjectives and pronouns and some measures of lexical richness to attempt to discriminate between healthy older participants and participants with probable AD.
+They found that AD patients had higher mean pronoun rate, adjective rate and verb rate but lower noun rate vs normal older controls \cite{Bucks2000}.
+It's important to note that the clinical group averaged a score on the MMSE of 15 (sd=6.8) which puts the clinical group in the 'moderately impaired range'.
+At this stage, the two groups should be reasonably easy to differentiate in terms of language and other cognitive abilities.
+\par 
+Fraser et al \cite{Fraser2015} built upon the work by Bucks by adding an acoustic analysis element to her analysis using the DementiaBank corpus, as well as significantly expanding the list of linguistic features they used.
+They used features generated from the Stanford parser (Now known as Core NLP) and computed the frequency of the occurrence of different parts of speech which they then normalised by the total number of words in each utterance.
+They also computed ratios between these parts of speech such as pronouns to nouns.
+They also included measures of syntactic complexity such as mean length of sentences and measures of lexical richness such as TTR.
+Their analysis showed was able to achieve 81\% accuracy in distinguishing individuals with AD from controls based on short samples of speech.
+They found that there are four factors which they felt were important in discriminating between these two groups: semantic impairment, acoustic abnormality, syntactic impairment and information impairment.
+\par 
+Zimmerer et al \cite{Zimmerer2016} took a slightly different approach in that they developed the Frequency in Language Analysis Tool (FLAT) which looked at the degree of formulaicity in an individual sample.
+The rationale behind this is that people with impaired cognitive function are able to depend on formulaic sentences for language generation to ease the burden of language generation on cognitive symptoms.
+They were able find that those with probable AD produced more formulaic language than controls.
+\par 
 Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls.
@@
 Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
+\par
+
 \subsection{The case for Deep Learning}
-% Deep Learning is an NOT
-
-One of the main benefits of deep learning is that it does not rely on pre defined features being fed to the model, instead generating it's own features.
-This move away from relying on features solves one of the difficulties with
-
-% One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
+One of the criticisms of traditional learning models is it's reliance on features that are generated for them.
+As we can see from the brief look at some of the research in this field, researchers have taken a number of different approaches with success.
+However, there does not seem to be a consensus.
+As machine learning in general moves away from traditional machine learning models to deep learning techniques, it poses the question can deep learning assist with this problem?
+\par 
+One of the main benefits of deep learning is that it does not rely on pre defined features being fed to the model.
+Instead, deep learning models take raw data in with some amount of preprocessing and generates it's own features.
+This move away from relying on features solves one of the difficulties we have with the current psychological literature, namely that there are some disagreement on the which
+\par 
+One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
+This is the case with all machine learning techniques, but more so with deep learning.
+There have been ways to 'create' more data such as data augmentation.
 
@@
 We can see that both Natural Language Processing and Machine Learning techniques have a lot to offer, Indeed there has been a lot of research which have used pre-existing datasets to explore this area with promising results.
-To date, there has not been much in the way of research in which these techniques are applied to newly created samples of language.
+One of the difficulties with the current research is the approach of trying to discriminate between those AD and healthy controls.
+This is not necessarily a problem in the real world as it is trivially easy to do for trained clinicians.
+A more interesting, but potentially harder to problem to solve is to discriminate between those with MCI and healthy controls, and more importantly to track their decline over time.
+Further, to date the authors are unaware of any research in which these techniques are applied to newly created samples of language.
 
@@
 
+% GANTT CHART
+\begin{sidewaysfigure}[ht]
+    \includegraphics[width=\textwidth]{/images/GanttChart.png}
+    \caption{Proposed Timetable.}
+\end{sidewaysfigure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=240px, height=150px]{images/RRPronouns.png}
+	\caption{Ronald Reagan - Pronouns Normalised over time}
+\end{figure}
+
 \section{Research Challenges}

### 2019-02-05 Source/QualifyingReport/QualifyingReport.tex

@@
 \title{Qualifying Report}
-\date{2019-01-21}
+\date{2019-01-31}
 \author{Jomar Alcantara}
@@
 \includegraphics[width=240px, height=150px]{images/BCTPicture.png}
-\caption{Cookie Theft Picture.\label{white}}
+\caption{Cookie Theft Picture - From Kaplan and Goodglass (1983)}
 \end{figure}
@@
 \end{equation}
-\subsection{Picture-related content thematic elements}
-Several studies examined the amount of thematic elements expressed that were directly relevant to picture stimulus in picture description tasks.
-The studies used a variety of phrases to denote these thematic elements, these included 'pictorial themes', 'relevant observations and 'semantic units' but are notionally similar.
-The only difference was the number of thematic elements that 'scored' for the studies in question.
+\subsection{Thematic and Content elements in relation to the Picture description task}
+A number of studies looked at the accuracy of picture descriptions using counts of thematic and content elements within a picture.
+Whilst called diffferent names such as 'pictorial themes', 'relevant observations and 'semantic units', they all represented the same idea.
+The only difference between the studies was the number of thematic elements that 'scored' correctly.
 Nicholas et al identified eight thematic elements of the Cookie Theft picture and used the number of elements as an outcome measure in different groups.
@@
 \includegraphics[width=240px, height=150px]{images/picnic-scene.jpg}
-\caption{Picnic Scene taken from the Western Aphasia Battery (WAB).\label{white}}
+\caption{Picnic Scene taken from the Western Aphasia Battery (WAB).}
 \end{figure}
@@
 (2001) examined Cookie Theft picture descriptions at four stages of AD.
-They identified 12 elements of the cookie theft picture and found that whilst each AD group differed significantly from the others and from controls, the measures did not distinguish between MCI and normal controls \cite{Bschor2001}.
+They found that whilst each AD group differed significantly from the others and also from controls, the measures were not able to distinguish between MCI and normal controls \cite{Bschor2001}.
 \par
@@
 \subsection{General Information Units or Content Information Units}
-Some studies used a more general concept of content, defining "information units" as "the smallest non redundant meaningful fact or inference," whether or not the information conveyed was specific to the context in which the conversation happened.
+Some studies used a more general concept of content, using terms such as "general information units" or "content information units" and this could be defined as "the smallest non redundant meaningful fact or inference,"  and was counted whether or not the information conveyed was specific to the context in which the conversation happened.
 Giles et al for example studied adults with minimal, mild or moderate AD vs controls and found that adults with AD produced fewer overall information units than controls \cite{Giles1996}.
-\subsection{Conciseness of information}
+\subsection{Conciseness of information}1`
 Conciseness has been defined as the number of words a speaker uses to express ideas.
-The theory is that people with AD will need more words to convey ideas because of word-finding related behaviours such as circumlocutions and repetitions.
-Conciseness is calculated by dividing the number of ideas expressed by the total number of words in a measure commonly referred to as idea density but also known as lexical index, information content and information unit conciseness index.
+The theory is that people with AD would need more words to convey ideas because of difficulties with word-finding and compensatory  behaviours such as circumlocutions and repetitions.
+Conciseness has previously been calculated by dividing the number of ideas expressed by the total number of words in a measure commonly referred to as idea density but also known as lexical index, information content and information unit conciseness index.
 Snowdon examined written discourse from the Nun study and found that low idea density in early life was associated with reduced cognitive performance in later life \cite{Snowdon1996}.
@@
 They found that those with AD produced fewer total semantic units than controls but there was no significant difference between the groups with regards to idea density.
-The study of "empty speech" by Nicholas et al examined conciseness with measures thought to contribute to the "nonspecifity" of discourse in AD, such as empty phrases (defined as common idioms contributing no relevant information), deictic terms (e.g.
+The study of "empty speech" by Nicholas et al examined conciseness and specifically looked at empty phrases (defined as common utterances which contribute no relevant information), deictic terms (e.g.
 "this", "that" without referents), indefinite terms (e.g.
 "thing" or "stuff"), pronouns without proper noun antecedants, and repetitions.
-In their study they found that AD patients produced more of these behaviours than did controls.
+In their study they found that AD patients produced more of these than did controls.
 \par
@@
 % CITE MURRAY!
-Efficiency is the rate at which meaningful information is conveyed over time, calculated by dividing the total number of information units by the duration in seconds of the speech sample.
-Smith et al, 1989 found that 18 adults with moderately severe AD produced fewer content units per minute on average than controls, and attributed this different to increased circumlocutions and repetitions \cite{Smith1989}.
-Murray used an analogous measure which he called "performance deviations per minute", in which fillers, irrelevant words, revisions or false stars, vague or non-specific vocabulary and inaccurate output (e.g.
-paraphasias) were divided by the total number of minutes in the speech sample; this measure was lower for those with AD than those with depression, and also healthy controls.
+Efficiency is the rate at which meaningful information is conveyed over time,  and can be calculated by dividing the total number of information units by the duration in seconds of the speech sample.
+Smith et al, 1989 found that 18 adults AD produced fewer content units over time on average than controls, he attributed these differences to increased circumlocutions and repetitions in the AD group \cite{Smith1989}.
+Murray used an similar measure in which fillers, irrelevant words, revisions or false stars, vague or non-specific vocabulary and inaccurate output were group together as 'performance deviations" and were divided by the total number of minutes in the sample.
+This measure was lower for those with AD than those with depression, and also healthy controls.
 The authors suggested that discourse information measures may help disentangle the similarities in symptoms of early AD versus depression in older adults.
 Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in dialogue in which a speaker provides responses do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls.
-\subsection{Quantity - Total number of words}
+\subsection{Total number of words}
 Several studies report that adults with moderate AD produce fewer words than controls on picture description, however other studies found no differences in total words among groups of controls and patients with MCI or AD.
 Murray and Nicholas et al investigated normal controls, patients with AD and older adults with depression and found no group differences in total words \cite{Nicholas1985}.
-In contrast, Lira 2014 found that controls produced more total words than patients with AD but found no difference between mild and moderate groups \cite{Lira2014}.
+In contrast, Lira et al found that controls produced more total words than patients with AD but found no difference between mild and moderate groups \cite{Lira2014}.
 \par
@@
 \par
-\subsection{Mean length of utterance (MLU)}
-Murray found that MLU was not a distinguishing factor among health adults, adults with depression and adults with AD.
-Ripich et al found a decrease in MLU in adults with severe AD over time, and this was supported by findings of Le et al in their studies of authors \cite{Ripich1988, Le2011}
-\par
-\subsection{Proportion of verbs to nouns plus verbs}
-Kave and Levy used a verb index to capture syntactic complexity and found that adults with AD expressed the same amount of verbs to nouns plus verbs as adult controls \cite{Kave2018}.
-\par
-\subsection{Syntactic Complexity - Composite measures of MLU, syntactic errors and verbs}
-A number of studies which explored syntactic complexity found differences between adults with MCI and controls, and between MCI and moderate AD stages \cite{Rentoumi2014, Ahmed2013}.
-The differences in syntactic complexity were not significant when individual measures were tested, but were apparent using a composite score consisting of MLU, words in sentences, syntactic errors, nouns with determiners, and verbs with inflections.
-\par
 \subsection{Formulaic Language}
@@
 The pragmatic language domain refers to the social rules for language for the purposes of communication including, using language to achieve goals, using information from the context to achieve these goals and using the interaction between people to initiate, maintain and terminate conversations.
-\subsection{Questions, turn-taking, unsure statements, egocentric comments}
-Ripich et al examined several pragmatic abilities among patients at different stages of AD \cite{Ripich1988}.
-The severe AD group asked more questions over time than the other group which infers problems with language processing.
-The authors argued that question-asking was a compensatory strategy and as a result, the proposed that adults in late-stage AD may have had insight into their communication problems.
-\par 
-There were a number of flaws of the methodology if their research.
-Firstly, a caregiver was asked to administer the picture description task in order to mirror a more typical communicative interaction.
-Whilst this improved the ecological validity of the study, the caregivers delivery would be varied and inconsistent.
-Secondly, due to the constrained nature of the picture description task, it is unlikely that it would be able to capture the pragmatic skills that are typical of conversations in everyday life.
 \subsection{Coherence}
-Coherence refers to the appropriate maintenance of topic in discourse and is thematically related to the immediately preceding utterance (local coherence) and by how closely an utterance relates to the general topic at hand (global coherence).
-One study looked at coherence, Chapman et al used picture descriptions of Norman Rockwell prints within a frame analysis, with frames being defined as internalized knowledge structures \cite{Chapman1998}.
-The authors identified aspects of content, including typical frames of interpretation, atypical, incorrect or no frames, propositions supporting frames and propositions disrupting frames as measures of coherence.
+Coherence, in lay terms, can be defined as the ability to maintain awareness of the topic at hand.
+It can be separated into local coherence which is related to themes of the immediately preceding utterance and global coherence which looks at how closely an utterance is related to the topic currently being discussed.
+Chapman et al used picture descriptions of Norman Rockwell prints within a frame analysis, with frames being defined as the context in which the picture is viewed \cite{Chapman1998}.
+The authors identified aspects of content, including whether the frames of interpretation being offered were typical, atypical, incorrect or had no frame.
+They also looked propositions supporting frames and propositions disrupting frames as measures of coherence.
 They examined these variables with early stage AD, old-elderly and normal controls.
-Old-elderly and normal controls produced significantly more typical frames and more frame supporting information that the AD group.
+Healthy older adults and normal controls produced significantly more typical frames and more frame supporting information than the AD group.
 The authors attributed AD patients' difficulties to memory deficits, attentional deficits, visual perceptual problems, disruption of internalized frame representation, or failure to access frame knowledge.
@@
 \par 
-One study examined verbal preservation in the description of Norman Rockwell prints, dividing the total number of words within perseverations by total number of words in the speech sample.
+One study examined verbal preservation in the description of Norman Rockwell prints \cite{Bayles198}.
+The presented participants with a number of similar pictures that had significantly different contexts and asked participants to describe these prints.
+They divided the total number of words within perseverations by total number of words in the speech sample.
 The authors also calculated rate of perseveration on two other language tasks: confrontation naming and generative naming.
-Across all tasks, the AD group produced significantly more perseverations than controls; however, on the picture description task alone, there were no significant differences between the two groups.
-The authors theorised that this was because picture description was an easier task because there was the visual stimulus, similar to the argument made by Bschor et al \cite{Bschor2001}.
+In all tasks, the AD group produced significantly more perseverations than controls but there were no significant differences between the two groups in the picture task in isolation.
+The authors felt that this was because picture description was an easier task, as it was a visual task in contrast to the other tasks which tapped into other cognitive processes.
 \par
@@
 \centering
-\includegraphics[width=350px, height=200px]{images/CoreNLP.jpg}
+\includegraphics[width=350px, height=200px]{images/CoreNLP.png}
 \caption{Depiction of how Core NLP marks up a sentence}
@@
 Firstly, the clinical features that combine to meet the diagnostic criteria for Dementia or it's variants are continuous in nature and heterogenous between patients and are also impacted by other variables.
-For example as shown above cognitive performance is affected in part by a patients educational attainment and a patients ability to live independently is impacted by a their physical health as much as their cognitive health.
+For example as shown above, cognitive performance is affected in part by a patients educational attainment and a patients ability to live independently is impacted by a their physical health as much as their cognitive health.
 The challenge therefore is to find features that are minimally impacted by other factors, or that can be controlled for by a experimental design such as a matched pairs design to control for educational attainment.
@@
 For example, Zimmerer (2016) \cite{Zimmerer2016} describes connectivity in such as a way that correlates directly with what Mueller (2018) \cite{Mueller2018a} calls Fluency.
-Whilst these are very nuanced measures which differ slightly in the form they take, the sheer range of measures and features being produced make it difficult to conceptualise as part of the larger problem.
+Whilst these are very nuanced measures which differ slightly in the form they take, the sheer range of measures and features being produced make it difficult to organise and explore what is truly going on in those with MCI and early AD..
 Some work needs to be done in producing a consistent list of measures that are validated using existing datasets and can be used for future research moving forward.
@@
 Machine learning approaches seem to be the most logical approach for achieving this aim as language could be collected in non-intrusive ways and passed to a machine learning algorithm for preliminary classification.
-Further, despite the excellent quality of datasets being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which we could claim are limited in scope.
+Despite the excellent quality of datasets, for example the DementiaBank dataset, being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which some researchers could claim are limited in scope.
 Finally, the recent resurgence in the use of neural networks and deep learning could provide the answer to the confused literature in terms of features.
 A key benefit of deep learning is it's ability to automate the process of feature engineering.
-So there is an opportunity to explore the use of deep learning, to not only develop new features but also
+So there is an opportunity to explore the use of deep learning, to not only develop new features but also validate existing features independently.
 \par
@@
 \subsection{Feature Selection}
-We calculating the following features for each transcript in turn using the NLTK (see section 2 for a description) \cite{Bird2009} and Python. 
+We calculated the following features for each transcript in turn using the NLTK (see section 2 for a description) \cite{Bird2009} and Python. 
 \par 
 \subsubsection{Measures of lexical variation}
-% CITE Richards 1986 Type/Token Ratios: What do they really tell us?
 We constructed two features of lexical variation.
@@
 To do this we were able to split each transcript into individual words and changed them to lowercase using NLTK and were then count the number of unique words that appeared in each transcript.
-We also used the TTR formula (see section 2 for a description) for a feature that measures lexical diversity independent of sample size \cite{}.
+We also used the TTR formula (see section 2 for a description) for a feature that measures lexical diversity independent of sample size \cite{Richards1987}.
 \subsubsection{Fillers, Non-Specific Nouns and LI Verbs}
-% CITE Bird et al 2000 The Rise and Fall of Frequency and Imageability Noun and Verb Production in Semantic Dementia
-For these features, we counted the number of occurrences for different categories (see table for list of categories tracked and the words counted.
+For these features, we counted the number of occurrences for different categories (see table for list of categories tracked and the words counted).
+The features were used by Berisha and Liss in their research \cite{Berisha2015} and were taken from work done by Bird et al \cite{Bird2000}.
 
@@
 		& RR v GHWB & RR v DJT & GHWB v DJT \\ \hline
-		Total Words & 4.434*** & 6.524*** & 2.899** \\ \hline
-		Unique Words & 6.832*** & 11.403*** & 4.137*** \\ \hline
-		Non Specific Nouns & 7.877*** & 3.426** & -0.574 \\ \hline
-		LI Verbs & 2.656** & 3.420*** & 1.628 \\ \hline
+		Total Words & \textbf{4.434***} & \textbf{6.524***} & \textbf{2.899**} \\ \hline
+		Unique Words & \textbf{6.832***} & \textbf{11.403***} & \textbf{4.137***} \\ \hline
+		Non Specific Nouns & \textbf{7.877***} & \textbf{3.426**} & -0.574 \\ \hline
+		LI Verbs & \textbf{2.656**} & \textbf{3.420***} & 1.628 \\ \hline
 		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
@@
 Whilst in terms of age, DJT is certainly more suitable as a control to match with RR, in some ways they held very different styles of press conferences in that RR preferred to do solo press conferences and DJT has shown a preference for doing joint press conferences which have an impact on the amount of language produced.
-This finding is, in itself notable as it illustrates the problems with small amounts of speech.
-\par
-The problem of finding an appropriate control is a common one in this domain.
+This artifact of the data is in itself notable as it illustrates the problems we may have with smaller amounts of speech.
+Also, the problem of finding an appropriate control is a common one in this domain.
 Given that those with MCI and early dementia have such variable presentations, it might prove of limited value in matched pairs design.
 \par 
-With further work, we will not have the luxury of having multiple samples and so it would be worth exploring how the quality of these predictions might lessen when faced with considerably fewer samples and over a smaller time period.
-In addition, this replication and extension has demonstrated the potential utility of using longitudinal data as a means of comparing language use of a person at two or more time periods and using this information as a diagnostic aid for MCI.
-\par 
-The results of this work show that we can track a person's use of language through time in a number of ways, and th
+With further work, it is not feasible to the vast array of samples over a timeframe, as we have had with the president corpus and so it would be worth exploring how the quality of these predictions might lessen when faced with considerably fewer samples and over a smaller time period.
+It would also be worth extending this research further to encompass more of the linguistic features Fraser used in her work \cite{Fraser2015} to see if there are any further insights to be gained.
+In addition, this replication and extension has demonstrated the potential utility of using longitudinal data as a means of comparing language use of a person at two or more time periods and using this information as a diagnostic aid for MCI and therefore more work would be helpful from a longitudinal perspective to see if this approach may be valid in moving towards a solution for this particular problem.
+\par 
+The results of this work show that we can track a person's use of language through time in a number of ways, and it is possible for an individual to be his or her own control.
+This is important as it means the heterogenous nature of the MCI population does not impact results as much as if we were comparing those with MCI to controls.
+Equally, it would be helpful to have controls to ascertain what would be usual to expect in the decline of language in a healthy older adult.
 
@@
 \begin{itemize}
-	\item \textbf{H1} Monitoring language samples and other measures of cognitive decline over the course of the year will be enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment
-	\item \textbf{H2} Those with MCI will show a different language profile to controls if matched similarly for age, gender, education and other significant factors.
-	\item \textbf{H3} The new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures such as the RBANS.
-	\item \textbf{H4} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, those with Alzheimers Disease and decline in those with MCI.
-	\item \textbf{H5} Using individuals as their own control will be enough to identify trends in the decline of their use of language.
+	\item \textbf{(H1)} Monitoring language samples and other measures of cognitive decline over the course of the year will be enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment
+	\item \textbf{(H2)} Those with MCI will show a different language profile to controls if matched similarly for age, gender, education and other significant factors.
+	\item \textbf{(H3)} The new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures such as the RBANS.
+	\item \textbf{(H4)} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, those with Alzheimers Disease and decline in those with MCI.
+	\item \textbf{(H5)} Using individuals as their own control will be enough to identify trends in the decline of their use of language.
 \end{itemize}
@@
 % GANTT CHART
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=350px, height=200px]{images/TaskList.png}
+	\caption{Task List}
+\end{figure}
+
 \begin{sidewaysfigure}[ht]
-    \includegraphics[width=\textwidth]{/images/GanttChart.png}
+	\includegraphics[width=\textwidth]{images/GanttChart.png}
 	\caption{Proposed Timetable.}
 \end{sidewaysfigure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=240px, height=150px]{images/RRPronouns.png}
-	\caption{Ronald Reagan - Pronouns Normalised over time}
-\end{figure}
 
@@
 \chapter{Conclusion}
+With the older population of the world continuing to grow year upon year, Dementia in all it's forms is also going to continue to grow.
+Indeed it is already one of the most common causes of death among the elderly \cite{Prince2015}.
+Whilst a lot of research gone into Alzheimer's disease and other forms of dementia, this research thus far has not yielded positive results in terms of our understanding of the disease or of a potential cure.
+Focus therefore should be on the early, pre-clinical stages of dementia to increase our understanding of the disease and to research potential early interventions in those who are likely to go on to develop dementia.
+\par 
+We looked at areas of research in terms of what is available for testing for MCI at the moment and what language markers have been identified as potentially significant.
+We also reviewed the contributions that Natural Language Processing and Machine Learning techniques could make to the field as well as a brief look at some of the key pieces of research that have used techniques.
+The picture is confusing as there has not been much in the way of consistency in researchers results in either of these areas for those with MCI.
+We propose that the problem is due to the heterogeneity of the MCI population and that simple differentiation of MCI or Healthy may not be possible.
+\par 
+We then looked at a potential longitudinal approach, originally done by Berisha and Liss \cite{Berisha2015}.
+We not only replicated a lot of their findings, despite slightly differing methodologies, but we found other language factors which change with the progression of MCI in our results.
+This is significant as it proves that we can use an individual as his or her own control in the measure of language deterioration and this goes some way to solving the problem of heterogeneity in this population.
+\par
+We then go on to propose more research looking at other available longitudinal data (Task 3), as well as proposing a methodology for collecting language data for analysis (Task 2, 5).
+Finally, we propose a longitudinal study (Task 4) which aims to use our newly created methodology for collecting language data in healthy controls, those with probable AD, those subjective cognitive impairment and those with a diagnosis of MCI to see whether we can detect signs of MCI through language analysis.
+\par
+If this work is successful, we have the opportunity to increase our understanding of the very early stages of AD as well as potentially having a new approach to the diagnosis of MCI and AD in the future.
 

### 2019-02-05 Source/Thesis/Thesis.tex
QR merge with Thesis

@@
 \listoftables
+\chapter{Overview of Problem}
+Dementia has been identified as one of those fast growing difficulties facing the world.
+A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
+\par
+In 2009, the Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
+A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
+As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages \cite{Prince2015} and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
+At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017}.
+I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
+\par
+There is a large body of research that looks at the decline in language in those with MCI and AD \cite{Taler2008, Boschi2017}.
+However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD \cite{Taler2008, Boschi2017}.
+Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
+Another area of focus for research of this nature is the process of collecting appropriate language samples.
+Whilst collecting samples of language is comparatively unintrusive, researchers recognise that these samples require a rich sample of language that potentially cannot be generated by tasks such as the picture description task.
+Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD.
+There is some evidence to support this.
+Berisha et al \cite{Berisha2015}, has shown through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control.
+\par
+The potential impact of this research in this area is immense.
+Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease.
+Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
+\par
+This report is structured in the following way.
+A literature review of current methods of cognitive assessment of MCI, an review of how language is said to be impacted in the MCI and Early AD population and a review of NLP and ML techniques that have the potential to be applied to this domain is presented in Chapter 2.
+The findings from a replication and extension of previous research using an existing dataset is presented in Chapter 3.
+I detail the aims of my research and my proposed future work plan including distinct tasks to be completed over the next two years in Chapter 4.
+Finally, my conclusions are summarised in Chapter 5.
+
 \chapter{Introduction}
@@
 \chapter{Background}
+\chapter{Literature Review}
 \section{Introduction}
+Alzheimer's Disease (AD) is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
+Whilst a definitive diagnosis of Dementia can only be produced at post-mortem, there are a number of clinical indicators from psychological perspective that can indicate dementia is present.
+From a clinical psychology perspective, those who have dementia demonstrate cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}.
+In addition, these symptoms are often accompanied by emotional problems such as depression and behavioural difficulties.
 \par
-Alzheimer's disease and other forms of dementia affect a significant proportion of the geriatric population in the world today and is currently the sixth leading cause of death in the US and the leading killer of women in the UK.
+AD and other forms of dementia affect a significant proportion of the geriatric population in the world today and is currently the sixth leading cause of death in the US and was named the leading killer of women in the UK.
 According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
-This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population \cite{Sturrock2016}.
+This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population \cite{AlzheimersSociety2014}.
 From a financial perspective, they estimate an annual spend of £4.3 billion of which approximately £85 million is spent solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
 Globally, this picture is a lot bleaker.
-Another report by Alzheimer's Disease International suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+Another report by Alzheimer's Disease International suggests that in 2015 there were 46 million people with a diagnosis of dementia globally and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
 The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
-\newline
 \par
-% In this paragraph, look for the study which actually has these figures.
-At present there are no drugs that improve the prognosis of those suffering with AD, all the drugs that are on the market are designed to manage symptoms.
-Whilst there are numerous investigational drugs for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{Alzheimer2016}.
-A possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place.
-Research has therefore been aimed at earlier diagnosis.
-\newline
-\par 
-Alzheimer's Disease (AD) is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
-However, a definitive diagnosis of Dementia can only produced at post-mortem.
-From a clinical psychology perspective, those who have dementia have cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits\cite{AmericanPsychiatricAssociation2013}.
-In addition, these symptoms are often accompanied by emotional difficulties such as depression and irritability and behavioural difficulties.
-\newline
+Despite this growing problem, at present there are no drugs that improve the prognosis of those suffering with AD.
+All the drugs that are on the market are designed to manage symptoms.
+Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{AlzheimersSociety2014}.
+Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place.
+It is therefore important to focus AD at it's earliest stages which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.
 \par
 Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent.
-Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time and alter cognitive function until a threshold is reached and clinical symptoms become more obvious.
-\cite{Nestor2006}.
-The case of Iris Murdoch illustrates this theory well.
-Le et al \cite{Le2012}found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel.
-If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.\newline
+Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time and alter cognitive function until a threshold is reached and clinical symptoms become more obvious \cite{Nestor2006}.
+The case of Iris Murdoch, who had a confirmed diagnosis of dementia, illustrates this theory well.
+Le et al \cite{Le2011} found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel when, it is theorized, the symptoms of AD manifested themselves more significantly.
+If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
+One of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of dementia.
+Albert and his team have worked to define clinical criteria which professionals can use to diagnose MCI due to AD and differentiate this from age-associated memory impairment and age-associated cognitive decline.
+They note that the diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}.
+\begin{enumerate}
+	\item Concern regarding a change in cognition - A person or an informant should express concern that there is a change in cognitive ability in comparison to previous level of performance.
+	\item Impairment in one or more cognitive domains - There should be evidence of lower performance in one or more cognitive domains beyond what would be expected of a person given their age and education.
+	\item Preservation of independence in functional - Whilst persons with MCI are expected to be able to maintain independence, it is common to experience mild problems in complex functional tasks which they may have been able to perform previously.
+This might mean that they take more time or be less efficient at completing these tasks, or it may be that may make more mistakes.
+	\item Not demented - The deterioration should be mild to the point that there is no significant loss of functioning in social or occupational contexts.
+\end{enumerate}
+In addition to meeting the above criteria, a clinician must rule out other conditions or factors that could account for the decline in cognition with the goal to increase the likelihood that the underlying cause of this decline is dementia.
 \par 
-One of the most common ways in which clinicians traditionally make an early diagnosis of cognitive impairment is through the use of the Mini Mental State Examination (MMSE) \cite{Folstein1975}.
+Many researchers have studied the early detection of MCI.
+These studies usually follow two main approaches.
+The analysis of biomarkers and the examination of patients who have demonstrated decreasing cognitive abilities.
+The first approach yields reliable results in the detection of AD in its moderate and advanced states but does not perform well during the early stages of the disease.
+The second approach has gained more attention in recent years, due to the fact that in clinical practice it has shown promise in the early detection of AD.
+In addition, the analysis of the decline of cognitive abilities is comparatively inexpensive and less invasive that the first approach which commonly requires the collection of a sample of cerebro-spinal fluid which is painful for the patient involved.
+\par
+One of the most common ways in which clinicians traditionally make an early diagnosis of dementia is through the use of the Mini Mental State Examination (MMSE) \cite{Folstein1975}.
 The MMSE is a brief questionnaire consisting of eleven questions which tests cognitive aspects of mental function and requires only 5-10 minutes to administer \cite{Folstein1975}.
@@
 The MMSE was not specifically created to screen for dementias and therefore does not interrogate key aspects of cognitive impairment known to be affected in dementia.
-It also has limited value in assessing under educated subjects and a meta-analysis on the effectiveness of the MMSE as a diagnostic tool for dementia to showed that it's accuracy was low (sensitivity between 78.4\% and 85.1\% and specificity between 81.3\% and 87.8\%) \newline
-% Need to cite Should the Mini-Mental State Examination be retired above.
+It also has limited value in assessing under-educated subjects and a meta-analysis on the effectiveness of the MMSE as a diagnostic tool for dementia showed that it's accuracy was low (sensitivity between 78.4\% and 85.1\% and specificity between 81.3\% and 87.8\%).
+As the MMSE has been shown to have low accuracy specifically in the diagnosis of dementia, it becomes necessary for professionals to employ the use of other tools or measures such as the Free Cued Selective Reminding Test (FCSRT) \cite{Grober2010} or the Montreal Cognitive Assessment (MoCA) \cite{Davis2015}.
 \par 
-As the MMSE is shown to have low accuracy specifically in the diagnosis of dementia, it becomes necessary for professionals to employ the use of other tools or measures such as the Free Cued Selective Reminding Test (FCSRT) or the Montreal Cognitive Assessment.
 These tests have the benefits of being much more accurate at diagnosing cognitive impairment and discriminating between dementia and other types of cognitive impairment at the cost time and training of psychological professionals such as clinical psychologists in administering these tests.
-Given the burden on these professionals is likely to increase due to a growing elderly population or in the case of developing countries where the clinician / patient ratio is already unsustainable, it seems useful to find less burdensome ways do aid professionals in diagnosing dementia.
-\newline
+However, the utility of diagnosing dementia at the point where clinical intervention is warranted is limited because at this stage both psychological and pharmacological interventions have been shown to not be effective \cite{Prince2015}.
+In order to further our understanding of the progression of dementia it is important to detect the signs of dementia before they are clinically apparent.
 \par
 The two main ways in which diagnosis is performed is through assessment of memory and language.
-Tests of memory are classically among the most accurate ways of diagnosing dementia, however they suffer from the same reliance on clinicians to administer these tests in a clinical setting.
+Tests of memory are classically among the most accurate ways of diagnosing dementia, however these tests suffer from the same reliance on clinicians to administer these tests in a clinical setting.
 Language however is a lot easier to collect and can be done in more naturalistic settings.
@@
 Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on measure decline in the use of language in those with AD.
-\newline
+There are a number of difficulties to watch out for with this approach.
+There are a wide number of factors that are involved in language degeneration in the elderly, and consequently there will be an expected amount of variability between subjects.
+The administration of such tests may induce nervousness and discomfort which may impact performance, and also repeatedly administering the same language tests for differences over time may be confounded by improved performance at tasks via practice effects.
+However, there is enough promise in this approach such that it could help further our understanding of the disease, it's progression and the parts of the brain affected in the early stages.
 \par
-There are a wide number of factors that are involved in language degeneration in the elderly, and consequently there will be an expected amount of variability between subjects.
-The consensus is that this decline is typically accelerated by the presence of dementia \cite{Berisha2015}.
-Given this statement, it seems logical to conclude that one of the key variables that distinguishes language decline in healthy individuals vs those with dementia is the rate of change in which the decline occurs.
-\newline
-\par
-According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with Mild dementia suffer from noticeable word finding difficulty.
+According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with mild dementia suffer from noticeable word finding difficulty.
 They may substitute general terms for more specific terms and may avoid the use of specific names of acquaintances.
@@
 Echolalia (meaningless repetition of another person's spoken words) and automatic speech typically precede mutism.
-With the wide range of deficits someone with AD can suffer, it makes sense to try to categorise these deficits in some way.\newline
+With the wide range of deficits someone with AD can suffer, it makes sense to try to categorise these deficits in some way.
 \par
-One of the most famous pieces of research on the topic of language decline in dementia was by Berisha and Liss (2015) \cite{Berisha2015} who examined speeches and public interviews of former US president Ronald Reagan.
+One of the most famous pieces of research on the topic of language decline in dementia was by Berisha and Liss who examined speeches and public interviews of former US president Ronald Reagan \cite{Berisha2015}.
 They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease.
 It turned out later to be the case that he had Alzheimer's Disease.
-Another classical study by Snowdon et al (1996) \cite{Snowdon1996} looked at whether linguistic ability in early life was associated with cognitive function and AD in later life.
+Another classical study by Snowdon et al looked at whether linguistic ability in early life was associated with cognitive function and AD in later life \cite{Snowdon1996} .
 They found that idea density (defined as the number of expressed propositions divided by the number of words) was a key predictor in predicting whether nuns would go on to develop AD in later life.
 They found that those who would go on to develop AD all had low idea density in early life and they found no AD present in those with high idea density in early life.
-As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are wide and varied and differ as the disease progresses.
-\newline
+As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are extremely variable and can differ from patient to patient as the disease progresses.
+The consensus among researchers that this language degeneration is typically accelerated by the presence of dementia \cite{Berisha2015} and that a potential indicator of dementia is the rate of change in which the decline occurs relative to a fixed point in time rather than a comparison across a cohort of individuals.
 \par
@@
 She found that people with AD generally had intact Phonology and Morphology but more impaired Syntax and Semantics.
-She stated that the language forms we learn last are the first to deteriorate.
-We generally learn language in small simple units initially and build syntax and complexity as we are more comfortable with language.\newline
+She asserted that the language forms we learn last are the first to deteriorate as we generally learn language in small simple units initially and build syntax and complexity as we are more comfortable with language.
+However it is important to note that different variants of dementia show different deficits in terms of language productions.
+Regardless, dividing language in this way is useful as it allows us to detect deteriorations in different parts of language usage and therefore may provide a way of discriminating between different forms of dementia.
 \par
 It is clear from both the clinical diagnostic criteria and supporting research that language is impacted in those with AD.
-However, one of the costs of analysing language is that there is a huge burden on trained practitioners, be it clinical psychologists, audio transcribers and text encoders to facilitate the process of collecting data and analysis.
+However, whilst there is a move towards research aimed at looking more specifically at MCI we currently lack the measures that are sensitive enough to detect MCI.
+Given that we know language is affected before a clinical diagnosis of dementia is usually made \cite{Berisha2015, Snowdon1996, Le2011}, it makes sense to explore whether language on it's own can provide markers that may indicate a cognitive impairment that could progress to dementia.
 The field of machine learning and natural language processing has been suggested as a way to improve the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline \cite{Boschi2017}.
-\newline
-
+% -- END OF INTRODUCTION --
 \subsection{Aims and Methodology}
 \par
-The purpose of this review is to seek to understand what techniques for assessing language have been used within the field of cognitive and clinical psychology.
-We then go on to look at what techniques have been developed in the field of machine learning (ML), deep learning (DL) and natural language processing (NLP) that might enable the automated analysis of language easier as well as an obstacles and/or limitations of current technology.
-Finally, we look at some studies which have already looked at the intersection of these two domains.
+The purpose of this review is to look at the current standards of cognitive tests available that can be used as a diagnostic tool for MCI.
+I then go on to explore what techniques for assessing language have been used within the field of cognitive and clinical psychology.
+I then go on to look at what techniques have been developed in the field of machine learning (ML), deep learning (DL) and natural language processing (NLP) that might enable the automated analysis of language easier as well as an obstacles and/or limitations of current technology.
+Finally, I look at some studies which have already looked at the intersection of these two domains.
 \newline
@@
 We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English.
-This resulted in 17 journal articles and/or conference papers which form this review.
-
-\begin{table}
-	\begin{tabular}{ | c | c | c | p{1cm} |}
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{4cm} | p{1cm} | p{6cm} | p{1cm} |}
 		\hline
-		Database & Number of Results & Search Terms  \\ \hline
-		ProQuest(PsychArticles) & 1484 Results & Language AND Decline AND Dementia \\ \hline
-		ProQuest(PsychArticles) & 486 Results  & Language AND Decline AND Dementia AND Speech \\ \hline
-		ProQuest(PsychArticles) & 159 Results & Machine Learning AND Dementia AND Language \\ \hline
-		Web of Science & 1207 Results  & Language AND Decline AND Dementia   \\ \hline
-		Web of Science & 151 Results  & Language AND Decline AND Dementia AND Speech  \\ \hline
-		Web of Science & 34 Results & Machine Learning AND Dementia AND Language \\ \hline
-		Scopus & 791 Results & Language AND Decline AND Dementia  \\ \hline
-		Scopus & 91 Results & Language AND Decline AND Dementia AND Speech   \\ \hline
-		Scopus & 29 Results & Machine Learning AND Dementia AND Langauge \\ \hline
-		Scopus & 1292 Results & "Language Deficits" AND Dementia \\ \hline
+		Database & Results & Search Terms  \\ \hline
+		ProQuest(PsychArticles) & 1484 & Language AND Decline AND Dementia \\ \hline
+		ProQuest(PsychArticles) & 486  & Language AND Decline AND Dementia AND Speech \\ \hline
+		ProQuest(PsychArticles) & 159 & Machine Learning AND Dementia AND Language \\ \hline
+		Web of Science & 1207  & Language AND Decline AND Dementia   \\ \hline
+		Web of Science & 151  & Language AND Decline AND Dementia AND Speech  \\ \hline
+		Web of Science & 34 & Machine Learning AND Dementia AND Language \\ \hline
+		Scopus & 791 & Language AND Decline AND Dementia  \\ \hline
+		Scopus & 91 & Language AND Decline AND Dementia AND Speech   \\ \hline
+		Scopus & 29 & Machine Learning AND Dementia AND Langauge \\ \hline
+		Scopus & 1292 & "Language Deficits" AND Dementia \\ \hline
 	\end{tabular}
+	\end{center}
 	\caption{\label{tab:table-name}Search Terms and Number of Results.}
 \end{table}
+% -- END OF Aims and Methodology
+\section{Existing Neuropsychological Measures of Cognitive Impairment, and Repeatable Battery of Neuropsychological Tests}
+In terms of standardised cognitive tests there are two main aims.
+Does the test distinguish accurately between normal aging, MCI and AD (diagnostic utility) and does the test distinguish between those individuals with MCI who will then go on to develop AD and those individuals with MCI who don't then go on to develop AD (prognostic utility).
+This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their performance in terms of both diagnostic and prognostic utility.
+\par 
+\subsection{Repeatable Battery for the Assessment of Neuropsychological Status}
+The Repeatable Battery for the Assessment of Neuropsychological Status (RBANS) was originally developed as an assessment tool for dementia, specifically looking at detecting a characterizing very mild dementia \cite{Randolph1998}.
+The authors felt that there was a shortfall of appropriate measures that were sensitive enough to milder impairments as well as a number of other shortcomings of existing tests.
+They met a number of design goals for this new battery of tests that addressed these shortcomings.
+The RBANS consists of a number of sub-tests across five distinct domains.
 
-\section {Types of Language Assessment}
+\begin{enumerate}
+	\item Immediate Memory 
+	\begin{enumerate}
+		\item{List Learning}
+		\item{Story Memory}
+	\end{enumerate}
+	\item Visuospatial / Constructional
+	\begin{enumerate}
+		\item{Figure Copy}
+		\item{Line Orientation}
+	\end{enumerate}
+	\item Language
+	\begin{enumerate}
+		\item{Picture Naming}
+		\item{Semantic Fluency}
+	\end{enumerate}
+	\item Attention
+	\begin{enumerate}
+		\item{Digit Span}
+		\item{Digit Coding}
+	\end{enumerate}
+	\item Delayed Memory
+	\begin{enumerate}
+		\item{List Recall}
+		\item{List Recognitions}
+		\item{Story Memory}
+		\item{Figure Recall}
+	\end{enumerate}
+\end{enumerate}
+
+However, whilst the authors claim that the RBANS is adequately sensitive in person's with MCI \cite{Randolph1998} and there is some research to support this assertion \cite{Karantzoulis2013}, other research points out that the RBANS has poor sensitivity in detecting MCI \cite{Duff2010}.
+Another drawback of this battery is the lack of executive function measures and object naming tasks.
+Research has shown that the RBANS has good test-retest reliability and convergent validity.
+However, Duff et al warned that caution should be exercised when using the RBANS in a MCI population as it has lower sensitivity in this population \cite{Duff2010}.
+
+\subsection{Digit Span Test}
+The Digit Span Test (DST) is predominantly used to measure a person's working memory capacity, specifically the capacity used to store and recall numbers.
+A participant is presented with a series of numbers of fixed length and is asked to recall those numbers in normal or reverse order.
+The series of numbers gets progressively longer until such time as a participant fails three or more times out of eight presentations.
+\par
+Research has consistently shown that MCI patients have a significantly lower digit span score in both normal and reverse order versions of this test and a study by Muangpaisan has shown that the reverse order version of this test can, to some degree, predict a diagnosis of MCI \cite{Muangpaisan2008}.
+Muangpaisan also found that Age, Gender and Education have an impact on the performance of the tests.
+Emrani et al found that the DST revealed no differences between specifically amnestic MCI and controls, but could differentiate between Mixed MCI and the other groups with mixed MCI recalling fewer correct responses than other groups.
+Notably, there was an attenuated recency effect in those with mixed MCI.
+\cite{Emrani2018}.
+Kessels identifies that that there are working memory deficits in MCI patients and these worsen with AD patients.
+\cite{Kessels2011} He also identified that both MCI and AD have impaired performance on all three conditions of the digit span test.
+No differences were found between forward and backward conditions in any of the groups.
+However, available tests may not detect subtle impairments \cite{Kessels2015}.
+\subsection{Rey Auditory-Verbal learning test}
+Rey's Auditory Verbal Learning Test (RAVLT+) looks a wide range of neuropsychological processes including short-term auditory-verbal memory, learning and retention of information.
+Participants are given a list of 15 unrelated words, repeated over five different trials and are asked to repeat.
+Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then repeat it again after 30 minutes \cite{Schmidt1996}.
+\par
+Several studies have shown that an impairment in RAVLT score reflect well the underlying pathology caused by AD.
+Thus making the RAVLT an effective early marker to detect AD in persons with memory complaints.
+Moradi investigated to what extent the RAVLT scores are predictable based on MRI data using machine learning approaches, as well as to find out what the most important brain regions are for the estimation of RAVLT scores \cite{Moradi2017}.
+They found a highly significant cross validated correlation between the estimated and observed RAVLT immediate and RAVLT Percent Forgetting.
+Further, they found that the conversion of MCI subjects to AD in 3-years could be predicted based on either observed or estimated RAVLT scores with an accuracy comparible to MRI-based biomarkers \cite{Moradi2017}.
+Another study by Schoenberg found that RAVLT to best distinguish patients suspected of Alzheimer's disease from the psychiatric comparison group \cite{Schoenberg2006}.
+\subsection{Digit Symbol Substitution Test}
+The Digit Symbol Substitution Test (DSST) involves a key consisting of the numbers 1-9 and a corresponding unique symbol.
+Below this key is a series of numbers from 1-9 in a randomized order and repeated multiple times.
+The participant is asked to fill in the corresponding symbol for each number.
+The task requires that the participant move between the key and the randomized sequence such that  they may retrieve the correct answer from the key, hold this in short-term memory and transcribe the key in the appropriate place.
+\par 
+Among those with no disorder in cognition, mobility and mood, being in the lowest DSST quartile compared to the highest was associated with nearly twice the odds of developing one or more clinical or subclinical disorders.
+Associations were stronger for incident clinical disorders in cognition.
+Slower psychomotor speed may serve as a biomarker of risk of clinical disorders, mobility and mood.
+While in part attributable to vascular brain disease, other potentially modifiable contributors may be present \cite{Rosano2016}.
+Further studying the causes of psychomotor slowing with ageing might provide novel insights into age-related brain disorders.
+Pascoe compared patients with PD with Normal Cognition (PD-N) with those with PD and MCI (PD-MCI) and healthy participants.
+PD-MCI participants achieved significantly lower scores than other groups in the DSST task \cite{Pascoe2018}.
+\subsection{Verbal Fluency}
+There are a number of tests that characterise this category of verbal fluency, but generally fall into two categories.
+Letter (Phonemic) fluency involves the generation of as many words as possible which begin with a specified letter.
+Category fluency involves the generation of as many words as possible that fall into a specfied category.
+Both these tasks impose demands on a number of different cognitive processes namely, executive function, verbal retrieval and recall, giving appropriate answers while monitoring previous answers and inhibit inappropriate responses.
+However, these two tasks require different strategies when attempting them.
+Letter fluency relies on search strategies based on lexical representations whereas category fluency requires a search for semantic extensions of a superordinate term, meaning that semantic associations within the lexicon must be intact in order for the task to be carried out successfully.
+\par  
+There have been numerous studies which have documented the impact that AD has on verbal fluency tests in both categories.
+A review carried out by Henry et al \cite{Henry2004}, found that performance in both letter fluency and category fluency was impaired in those with AD vs controls, but found a larger effect for tests of semantic fluency.
+\subsection{Naming Tests}
+Word-finding difficulty is a common symptom of AD and these deficits usually occur during the early stages of the disease progression.
+As such , a test of a patients ability to find words (known as confrontation naming) is a common way to measure cognitive decline.
+One common way to do this is the Boston Naming Test (BNT) which comprises 60 items on a spectrum of very frequent to very infrequent.
+This has been reduced subsequently to two thirty item versions and four fifteen item versions, which correlate significantly with the original sixty item version and the benefit of the shorter versions of the test is that it facilitates testing of individuals with AD who may suffer from fatigue or limited attention span \cite{Williams1989}.
+\par 
+Willers et al \cite{Willers2008} studied Twenty aMCI patients, twenty AD and 21 controls matched by age, sex and education level.
+They found AD patients obtained significantly lower total scores on the BNT than aMCI patients and controls.
+aMCI patients did not obtain significant differences in total scores but showed significantly higher semantic errors compared to controls.
+Semantic processing is impaired during confrontation naming in aMCI.
+\par 
+Vadikolias investigated the impact of education on naming tests \cite{Vadikolias2012} and found that higher educational attainment in aMCI subjects were correlated with better performance in verbal and non-verbal tasks during repeated examinations over 1-year.
+Subjects with a lower level of education performed worse than patients with a high level of education who presented a more stable clinical score.
+The explanation for this is the idea of a 'cognitive reserve' in participants with a higher education such that this provides a buffer that, while not preventing the physiological symptoms of AD, can potentially delay the clinical onset of cognitive symptoms that characterise AD.
+This theory is supported by Snowdon who found a relationship between early life linguistic ability and the density of neurofibrillary tangles in his nun study \cite{Snowdon1996}.
+Whilst these studies and tests provide evidence that there are word finding difficulties in those with MCI, on their own they do not provide sufficient ability to diagnose MCI or provide an prognosis of disease progression.
+\subsection{Rey-Osterrieth Complex Figure Task}
+The Rey-Osterrieth Complex Figure (ROCF) is a task widely as a test of visuo-spatial skill and visual memory.
+The task, which was originally designed by Rey (1944) and standardised by Osterrieth (1944) \cite{Rey1941, Osterrieth1944}, asks a participant to copy a complex geometrical figure (known as the immediate copy condition and to recall and reproduce the figure from memory without warning (known as the delayed recall condition).
+In the immediate copy condition, the complexity of the figure requires an integrative cognitive ability.
+The reproduction of such a complex structure involves processes such as planning and organizational strategies that are related to executive functions.
+In order to make this task repeatable, Taylor et al designed a comparable set of figures which have proven to be of equivalent difficulty \cite{Taylor1969, Strauss1990}.
+\par
+Salvadori found that patients with vascular MCI had a worse performance in the immediate copy of the ROCF compared to individuals with degenerative MCI, despite their significant impairment in terms of general cognitive status and visual memory \cite{Salvadori2018}.
+Evidence shows that patients with disorders that possibly involve attention and executive functions are characterized by a more disorganised approach when copying the ROCF compared to controls.
+One of the difficulties with the presentation of this task in a repeated battery will be the fact that it turns from an incidental memory task (it is a memory task but the participant is not forewarned that they will be required to memorise the picture) into an intentional memory task (given the previous exposure to the task, it would be expected that a participant would preempt the delayed recall portion of the test and spend more time attempting to memorise the details) \cite{Teng2009}.
+\subsection{Hayling Sentence Completion Test}
+Inhibitory deficits are a common in all stages of dementia.
+This is usually tested using Stroop test, however this has the drawback of lacking ecological validity.
+Therefore researchers have moved towards using the Hayling Sentence Completion Test (HCST) which uses skills such as word retrieval as well as the ability to inhibit ones responses where appropriate and is therefore a much more ecologically valid task.
+There are two parts to the HCST.
+In the first part, participants have to complete a sentence by providing a word that best fits the given sentence (this is known as the initiation condition).
+In this second part, participants have to complete sentences by inhibiting an impulse to give the word that best first the sentence as in the first part, and producing a semantically unconnected word.
+Performance in both conditions is measured by the time taken for the participant to initiate a response, and in the inhibition condition also by the correctness of the word.
+\par 
+Martyr et al compared healthy controls with patients with dementia and patients with Parkinson's disease \cite{Martyr2017}.
+They found that a high proportion of Category A errors (producing a word that fits the sentence when instructed otherwise) was a factor in performance loss for participants with dementia.
+Findings suggest that the HSCT may be sensitive to verbal suppression deficits and may provide insight into inhibitory control in participants with dementia.
+Patients with Dementia were significantly slower than controls in the initiation and inhibition conditions vs healthy controls, and slower than patients with Parkinson's disease in the initiation but not the inhibition condition.
+\subsection{Grooved Pegboard Test}
+The Grooved Pegboard Test (GPT) was originally designed to cover a variety of different psycho-motor functions including hand-eye coordination and motor speed.
+However, some studies have shown a correlation of performance in the GPT and measures of cognitive performance such as the Montreal Cognitive Assessment.
+\par 
+Bezdicek found that the GPT predicted performance on the MoCA and concluded that in addition to being a measure of motor skills, there were results that showed that the GPT could also provide information about a participants cognitive skills \cite{Bezdicek2014}.
+Whilst this was specifically using a population of patients with Parkinson's disease, given the MoCA and GPT have shown degree of correlation, there remains an opportunity for research that explores the use of the GPT with MCI and Early AD populations.
+\par 
+Darweesh et al, used the Purdue Pegboard Test (PPT) to assess manual dexterity in a healthy older adult population and followed their participants for between eight and twelve years until an indication of the onset of a neurodegenerative disease \cite{Darweesh2017}.
+In this time 227 (4\%) of their participants went on to develop a diagnosis of dementia.
+They noted that higher PPT scores were associated with lower risk of incident neurodegnerative disease and noted significant associations of PPT scores with all forms of dementia and this potentially highlights a deterioration of motor function in the pre-clinical phase of dementia.
+\subsection{Visual Search}
+The visual search task requires participants to be watching a computerised display in which a target appears either by itself or surrounded by other elements which are used as a distraction.
+These other elements can be vary in similarity to the target.
+This is inherently a measure of attention shifting and processing of various similarity and as one would expect, the visual search time is increased in the presence of distractors.
+This effect is magnified in participants who have a cognitive impairment.
+\par
+Research has shown there to be deficits in both AD and MCI populations in this task \cite{Tales2005}, and those with MCI exhibited less severe deficits compabut that these deficits were not as apparent in MCI populations, there remained a significant enough difference that could be used to differentiate healthy controls from those with MCI.
+In a comparison of visual search performance between those with aMCI and healthy controls, Tales noted significantly poorer performance in the aMCI group.
+However, she also noted a good deal of heterogeneity in the aMCI group which illustrates that whilst the aMCI group have essentially the same condition, presentations within this population can differ markedly.
+The results from this study also illustrates the use of a non-memory task as a means of diagnosing dementia as patients who went on to be diagnosed with dementia whilst this study was in progress had significantly poorer visual search performance.
+\subsection{Free Cued Selective Reminding Test}
+The Free Cued Selective Reminding Test (FCSRT) was borne out of the premise that by controlling the conditions of learning, a measure of memory is possible that is not confounded by normal age related changes in cognition.
+Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cognitive decline due to a cognitive impairment.
+The FCSRT starts with a learning phase in which participants are asked to look at a card containing four pictures (e.g., grapes) for an item that belongs to a named category(e.g., fruit).
+After these four items are identified, immediate cued recall of these four items is tested and this is repeated for a total of 16 items.
+Following this learning phase is the recall phase in which participants are asked to recall all the items identified without cues and any items which are not recalled are then cued.
+There are three outcome measures for this test, free recall (the number of items recalled without cues), total recall (the number of items recalled with or without cues) and cue efficiency which is identified as follows \cite{Grober2010}.
+\begin{equation} \label{x1}
+CueEfficiency = (totalRecall-freeRecall)/48-freeRecall, range 0.0-1.0)
+\end{equation}
+Grober found that patients with impaired free recall with four times more likely to develop dementia than those with intact free recall \cite{Grober2010}.
+This test has also been shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the researchers to used the test to categorise prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT score \cite{Sarazin2007}.
+\subsection{Conclusions}
+I have looked at a number of different cognitive tests and a battery of tests that aim to have high diagnostic and prognostic capabilities in the MCI population.
+However, particularly with the RBANS, there lacks sufficient sensitivity in differentiating those with MCI from healthy elderly individuals such that this tool could be used with a level of confidence in a clinical setting.
+In regard to the tests, a common theme is a lack of studies and therefore evidence into the utility of these tests with an MCI population.
+However if, as researcher, we aim to investigate this population further then a benchmark battery of tests which is sensitive enough in both diagnostic and/or prognostic utility should be a goal.
+\par 
+Another criticism of the current literature is the lack of consistency with regard to the experimental groups.
+For example, some studies focus on differentiating between an MCI group, a AD group and a healthy controls group whereas other studies may further subdivide the MCI category according to a number of factors such as Amnesic or Non-Amnesic MCI, Mixed MCI (sometimes called Multi-domain MCI in the literature).
+Given the inconsistency in defining the experimental groups the confusing and often conflicting results that researchers produce is to be expected.
+Future research should look at the standardization of the operational definition of cognitive impairment in MCI may result in more consistent predictions of progression to AD.
+\par 
+Finally, it can be said that in testing for MCI and comparing these results with a cohort of those with a similar diagnosis has limited utility.
+This is because there is huge variability in the presentation of MCI within participants that even if controlled for with a matched pairs design for age, gender and education will produce inconsistent results.
+There is an argument therefore for the use of longitudinal studies, where the comparisons on the performance of these tasks are with the participants themselves.
+There are not many studies which use a longitudinal approach, although those that do show promising results.
+
+\section{Types of Language Assessment}
 One of the key debates when looking at how to analyse language is the type of task provided to elicit language production in participants.
 In the literature researchers have primarily focused on Picture Description tasks but have also suggested other ways in which we might collect data.
-\newline
 \par
@@
 One of the most commonly used tasks to measure language is the Picture Description task.
-An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Kaplan2010}.
-In this task participants are asked to describe a picture presented to them in as much detail as possible.
-The picture itself depicts a familiar domestic scene and would not require participants to use any vocabulary beyond that learned in childhood.
-It was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}\newline
+An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Borod1980}.
+The Cookie Theft picture (pictured below) depicts a scene of a home typical of the period of time when it was created and would generally not require participants to use any complicated vocabulary to describe.
+In this task participants are asked to describe the picture presented to them in as much detail as possible.
+This task was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}
+\begin{figure}[H]
+\centering
+\includegraphics[width=240px, height=150px]{images/BCTPicture.png}
+\caption{Cookie Theft Picture - From Kaplan and Goodglass (1983)}
+\end{figure}
 \par
-The picture description task does a fine job of eliciting descriptive language, however because of the limited content has limited use.
-The task in itself just a descriptive task, and therefore elicits a certain type of language.
+The picture description task does a fine job of eliciting descriptive language however because of the specific content the language produce could be considered quite limited.
 There is some disagreement as to the benefits of this using this methodology.
@@
 In explaining the differences, it is worth noting that these researchers were using differing variables and this could explain their different perspectives.
-\newline
-\par
-In terms of Machine Learning research in this area, a number of researchers have used transcripts based on picture description tasks \cite{Zimmerer2016, Orimaye2017, Mueller2018a, Fraser2015} and have successfully extracted linguistic features that could differentiate between AD and controls.\newline
 \par
@@
 The story narration task is designed to study a participant's ability to describe and elaborate on a story which is depicted using a series of pictures.
-The stories depicted are usually based on children's books or famous stories such as Cinderella.
-\cite{Fraser2014} This task requires ordering the story in a structured and coherent framework.
-It also requires comprehension and understanding of the stories characters and the events depicted, as well as an awareness of a character's goals and internal responses to given events.
-This task is particularly useful, as the procedure reduces the demands on memory and is therefore able rule out memory as a confounding variable for any results observed.
-As noted above, Ash \cite{Ash2012} felt that this task was interchangable with the Picture Description task and because this task requires elaboration rather than simple description, is a sturdier test of lexical and semantic abilities as well as syntactic complexity.
-\cite{DeLira2011} \newline
+The stories depicted are usually based on children's books or famous stories with the Cinderella being the one most typically used \cite{Fraser2014}.
+This task requires ordering the story in a structured and coherent framework.
+It also requires comprehension and understanding of the stories characters and the events depicted, as well as an awareness of a character's actions, motivations and internal reactions to given events.
+This task is particularly useful as the procedure reduces the demands on memory, due to the participant being able to access the picture book during the description and is therefore able rule out memory as a confounding variable for any results observed.
+As noted above, Ash \cite{Ash2012} felt that this task was interchangable with the Picture Description task.
+However, other research felt that this was a studier test of lexical and semantic abilities as well as syntactic complexity because this task requires interpretation and elaboration in additional to a simple description \cite{DeLira2011}.
 \par
-Given the relative strengths of the Narrative description task vs Picture description task, there are few pieces of research research that have used Machine Learning to analyse features from Narrative picture tasks \cite{Fraser2014}.
-This could be down to the availability of data and the absence of any meaningful sets of transcripts of participants performing this task.
+Given the relative strengths of the Narrative description task vs Picture description task, there are few pieces of research that have used Machine Learning to analyse features from Narrative picture tasks \cite{Fraser2014}.
+This could be due to the availability of data and the absence of any meaningful sets of transcripts of participants performing this task.
 However, this could be an interesting direction to take research in the future to see if features generated from this task could be used to predict MCI or AD.
-\newline
 \par
 \subsection{Interviews}
-Interviews can also be used to elicit language, the idea of employing questions to guide a conversation between speakers.
+Interviews can also be used to elicit language in a more natural way by asking questions to guide a conversation between speakers.
 There are three types of interviews: unstructured, structured and semi-structured.
@@
 Whilst this is the most ecologically valid form of conversation and therefore language generation, it's unstructured nature means that the protocol cannot be consistent and therefore reproduced.
-Semi-structured interviews are therefore preferred other forms of interview as a middle ground.
+Semi-structured interviews are therefore preferred over other forms of interview as a middle ground.
 The semi structured nature of these interviews means that there is some replicability but does not constrain the participant in answering questions.
-\newline
 \par
-The analysis of interviews can be difficult to analyse as both the content can vary even between participants.
+The analysis of interviews can be difficult to analyse as both the content can vary even between participants, although it can be argued that content should not affect the type of language being generated unless it is narrow topic or the participant is constrained in how they answer a given question.
 It is also difficult to measure as there are no pre-defined task goals in comparison to the other two methods.
 Nevertheless, this is the most naturalistic setting for looking at language production and can be used to look at the syntactic and semantic parts of language generation \cite{Sajjadi2012}.
-There have been some attempts to use interviews to to assess language production in AD with promising results \cite{Asgari2017, Guinn2015}.\newline
+There have been some attempts to use interviews to assess language production in AD with promising results \cite{Asgari2017, Guinn2015}.
 \par
+\subsection{Conclusions}
+One can view the different types of tasks above as a continuum where picture tasks represent a much more controllable task with a lot of supporting research but which generates a much more constrained set of language that is atypical of normal speech in terms of the cognitive functions used.
+
 \section{How do we analyse language, issues and debates}
+There are a wide variety of approaches that we can take in the analysis of language and there is no real consensus on the best approach to this particular problem.
+This section looks at the different ways in which researchers have looked at the problem as well as discussing some of the areas of contention.
 \subsection{Single Word Language tasks vs Connected Language tasks}
@@
 There are other less common methods of analysing language but these are beyond the scope of this review.
-\newline
+\par
+Single word tasks such as the Boston Naming Test and other such standardized language tests generally target a participants word production where this is defined as the ability to form and express words in accordance with certain criteria (see above for a full description of verbal fluency as task).
+There are a number of benefits to using single word tasks.
+From a research methodology perspective, using a standardized test allows researchers to target a very specific process in language generation and isolate factors that impact performance in language well.
+However, this approach does not take into account other cognitive processes that are used in lengthier speech tasks.
+More specifically, single word language tasks do not look at the interaction between language, executive functions and reasoning abilities.
+They also do not require the logical and efficient organisation of ideas.
+\par
+Connected language tasks, such as the picture description task and interviews described above are much more reflective of the processes involved in natural language generation.
+They are able to be relatively constrained in the language available to be used, for example in the picture description task, or they can be unconstrained in the form of interviews.
+But regardless of how they are framed, they are potentially much more useful in terms looking at all the processes involved in language generation with the drawback that you will not be able to isolate specific parts of the language generation process.
+A review by Boschi et al concludes that analysis of connected speech is potentially useful in guiding clinicians to identify language disorders \cite{Boschi2017} and also highlights the role of NLP and Machine learning in assisting in this endeavour.
 \par
@@
 Pragmatic meaning refers looks at the same sentence in terms of words and grammar but takes into account the situation or context in which these words are spoken.
-\newline
+Emery in her literature review looks at all levels of language tasks except for pragmatics, however this is an area that should not be ignored.
+Whilst the study of language from a pragmatic perspective is much more complex, it is perhaps one of the most vital areas to study because of the number of different cognitive processes involved.
 \par
-\section{Semantic Content}
-% Bridge sentence, idea density and type token ratio.
+\subsection{Semantic Content}
 Another approach to linguistic analysis in this field is the idea of measuring semantic content and complexity.
@@
 If this is true, then psychological measures of semantic and syntactic skills should be able to pick up signs of deterioration and act as markers for possible MCI and AD.
-An example of these measures is the concept of idea density.
-Formally, idea density is defined as the average number of propositions per sentence \cite{Kintsch1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}.
-Type to Token Ratio, Brunet's Index and Honore's Statistic are two more examples of validated measures which are used to measure the lexical diversity of a given piece of text.
-This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2005}.
-\newline
+An example of a semantic complexity measure is the concept of idea density.
+Formally, idea density is defined as the average number of propositions per sentence \cite{Keenan1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}.
+An example of semantic content measures is Type to Token Ratio (described below) which is used to measure the lexical diversity of a given piece of text and/or utterance.
+This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2013}.
 \par
-\subsection{Picture-related content thematic elements}
-Several studies examined the amount of thematic elements expressed that were directly relevant to picture stimulus in picture description tasks.
-The studies used a variety of phrases to denote these thematic elements, these included 'pictorial themes', 'relevant observations and 'semantic units' but are notionally similar.
-The only difference was the number of thematic elements that 'scored' for the studies in question.
+\subsubsection{Type token ratio(TTR)}
+Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
+\begin{equation} \label{x2}
+TTR = numberOfUniqueWords / totalNumberOfWords.
+\end{equation}
+\subsection{Thematic and Content elements in relation to the Picture description task}
+A number of studies looked at the accuracy of picture descriptions using counts of thematic and content elements within a picture.
+Whilst called diffferent names such as 'pictorial themes', 'relevant observations and 'semantic units', they all represented the same idea.
+The only difference between the studies was the number of thematic elements that 'scored' correctly.
 Nicholas et al identified eight thematic elements of the Cookie Theft picture and used the number of elements as an outcome measure in different groups.
 He found that patients with AD expressed significantly fewer content elements than controls.
-\newline
 \par
-Hier, Hagenlocker and Shindler assessed content using a similar list of thematic elements.
+Hier, Hagenlocker and Shindler assessed content using a similar list of thematic elements \cite{Hier1985}.
 They divided their participants into early-stage and late-stage AD, as well as including a control group.
 The late-stage AD group produced significantly fewer relevant observations than the early stage group, and the AD group combined produced fewer relevant observations than controls.
-This study was replicated by Vuorinen et al (2001).
-\newline
+This study was replicated by Lukatela et al \cite{Lukatela1998}.
 \par
-Smith, Chenery and Murdoch (1989) applied Hier's methodology for constructing pictorial 'themes' with the Picnic Scene from the Western Aphasia Battery (WAB) with a control and patients with moderate to moderately severe AD.
+Smith, Chenery and Murdoch applied Hier's methodology for constructing pictorial 'themes' with the Picnic Scene from the Western Aphasia Battery (WAB) with a control and patients with moderate to moderately severe AD \cite{Smith1989}.
 The authors found no difference in the number of semantic elements produced but did not that the group with moderate to moderately severe AD took more time and more syllables to communicate these elements.
-\newline
 \par
-Sajjadi et al (2012) examined 10 pictorial themes in picture description the Comprehensive Aphasia Test and found that the group with mild AD produced similar themes than controls.
+\begin{figure}[H]
+\centering
+\includegraphics[width=240px, height=150px]{images/picnic-scene.jpg}
+\caption{Picnic Scene taken from the Western Aphasia Battery (WAB).}
+\end{figure}
+Sajjadi et al examined 10 pictorial themes in picture description the Comprehensive Aphasia Test and found that the group with mild AD produced similar themes than controls \cite{Sajjadi2012}.
 Bschor et al.
 (2001) examined Cookie Theft picture descriptions at four stages of AD.
-They identified 12 elements of the cookie theft picture and found that whilst each AD group differed significantly from the others and from controls, the measures did not distinguish between MCI and normal controls.
-\newline
+They found that whilst each AD group differed significantly from the others and also from controls, the measures were not able to distinguish between MCI and normal controls \cite{Bschor2001}.
 \par
@@
 Studies using these features found some differences between AD and controls, and some could differentiate between different stages of AD.
-\subsection{General Information Units}
-Some studies used a more general concept of content, defining "information units" as "the smallest non redundant meaningful fact or inference," whether or not the information conveyed was specific to the context in which the conversation happened.
-Giles et al, for example, studied adults with minimal, mild or moderate AD vs controls and found that adults with AD produced fewer overall information units than controls.
-\subsection{Conciseness of information}
+\subsection{General Information Units or Content Information Units}
+Some studies used a more general concept of content, using terms such as "general information units" or "content information units" and this could be defined as "the smallest non redundant meaningful fact or inference,"  and was counted whether or not the information conveyed was specific to the context in which the conversation happened.
+Giles et al for example studied adults with minimal, mild or moderate AD vs controls and found that adults with AD produced fewer overall information units than controls \cite{Giles1996}.
+\subsection{Conciseness of information}1`
 Conciseness has been defined as the number of words a speaker uses to express ideas.
-The theory is that people with AD will need more words to convey ideas because of word-finding related behaviours such as circumlocutions and repetitions.
-Conciseness is calculated by dividing the number of ideas expressed by the total number of words in a measure commonly referred to as idea density but also known as lexical index, information content and information unit conciseness index.
-Snowdon et al examined written discourse from the Nun study and found that low idea density in early life was associated with reduced cognitive performance in later life.
-Riley et al extended these findings by concluding that early-life idea density was associated with lower brain weight, higher degree of cerebral atrophy and increased neurofibrillary pathology in later life.
-\newline
+The theory is that people with AD would need more words to convey ideas because of difficulties with word-finding and compensatory  behaviours such as circumlocutions and repetitions.
+Conciseness has previously been calculated by dividing the number of ideas expressed by the total number of words in a measure commonly referred to as idea density but also known as lexical index, information content and information unit conciseness index.
+Snowdon examined written discourse from the Nun study and found that low idea density in early life was associated with reduced cognitive performance in later life \cite{Snowdon1996}.
+Riley et al extended these findings by concluding that early-life idea density was associated with lower brain weight, higher degree of cerebral atrophy and increased neurofibrillary pathology in later life \cite{Riley2005}.
 \par
-Ahmed, de Jager et al examined idea density with patients who had confirmed AD post mortem.
+Ahmed, de Jager et al examined idea density with patients who had confirmed AD post mortem \cite{Ahmed2013}.
 They found that those with AD produced fewer total semantic units than controls but there was no significant difference between the groups with regards to idea density.
-The study of "empty speech" by Nicholas et al examined conciseness with measures thought to contribute to the "nonspecifity" of discourse in AD, such as empty phrases (defined as common idioms contributing no relevant information), deictic terms (e.g.
+The study of "empty speech" by Nicholas et al examined conciseness and specifically looked at empty phrases (defined as common utterances which contribute no relevant information), deictic terms (e.g.
 "this", "that" without referents), indefinite terms (e.g.
 "thing" or "stuff"), pronouns without proper noun antecedants, and repetitions.
-In their study they found that AD patients produced more of these behaviours than did controls.
-\newline
+In their study they found that AD patients produced more of these than did controls.
 \par
 \subsection{Efficiency}
-Efficiency is the rate at which meaningful information is conveyed over time, calculated by dividing the total number of information units by the duration in seconds of the speech sample.
-Smith et al, 1989 found that 18 adults with moderately severe AD produced fewer content units per minute on average than controls, and attributed this different to increased circumlocutions and repetitions.
-Murray 2010, used an analogous measure which he called "performance deviations per minute", in which fillers, irrelevant words, revisions or false stars, vague or non-specific vocabulary and inaccurate output (e.g.
-paraphasias) were divided by the total number of minutes in the speech sample; this measure was lower for those with AD than those with depression, and also healthy controls.
+% CITE MURRAY!
+Efficiency is the rate at which meaningful information is conveyed over time,  and can be calculated by dividing the total number of information units by the duration in seconds of the speech sample.
+Smith et al, 1989 found that 18 adults AD produced fewer content units over time on average than controls, he attributed these differences to increased circumlocutions and repetitions in the AD group \cite{Smith1989}.
+Murray used an similar measure in which fillers, irrelevant words, revisions or false stars, vague or non-specific vocabulary and inaccurate output were group together as 'performance deviations" and were divided by the total number of minutes in the sample.
+This measure was lower for those with AD than those with depression, and also healthy controls.
 The authors suggested that discourse information measures may help disentangle the similarities in symptoms of early AD versus depression in older adults.
-Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in dialogue in which a speaker provides responsees do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls..\newline
-\subsection{Lexical richness and diversity}
-\subsubsection{Type token ratio(TTR)}
-Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
-\begin{equation} \label{x1}
-TTR = numberOfUniqueWords / totalNumberOfWords.
-\end{equation}
-\subsubsection{Brunet's Index(W)} % Don't forget to cite the original article.
-Brunet's Index (W) differentiates itself for TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
-\begin{equation} \label{x2}
-W = N^{V(-0.165)}
-\end{equation}
-where N is the total length of the utterance being measured and V is equal to the total vocabulary being used by the subject.
-Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers.
-\newline
-\par 
-\subsubsection{Honore's Statistic (R)} % Don't forget to cite the original article.
-Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation: 
-\begin{equation} \label{x3}
-%% Check to see if this is accurate.
-R = (100 \log N) / (1 - V1/V)
-\end{equation}
-where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
-\subsection{Quantity - Total number of words}
+Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in dialogue in which a speaker provides responses do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls.
+\subsection{Total number of words}
 Several studies report that adults with moderate AD produce fewer words than controls on picture description, however other studies found no differences in total words among groups of controls and patients with MCI or AD.
-Murray and Nicholas et al investigated normal controls, patients with AD and older adults with depression and found no group differences in total words.
-In contrast, Lira 2014 found that controls produced more total words than patients with AD but found no difference between mild and moderate groups.
-\newline
+Murray and Nicholas et al investigated normal controls, patients with AD and older adults with depression and found no group differences in total words \cite{Nicholas1985}.
+In contrast, Lira et al found that controls produced more total words than patients with AD but found no difference between mild and moderate groups \cite{Lira2014}.
 \par
@@
 The general hypothesis motivating these studies is that either working memory limitations or semantic memory limitations in AD affect one's ability to use complex constructions.
-\newline
 \par
-\subsection{N-grams}
+\subsubsection{N-grams}
 One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
@@
 A number of researchers have used n-grams as features.
-One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class.
+\par
+Asgari, Kaye and Dodge (2017) \cite{Asgari2017} used another form of word frequency measurement.
+Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words \cite{Pennebaker2015}.
+They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
+\par
+\subsection{Formulaic Language}
+Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
+They found that there were four factors which they identified as important in the classification of participants as either healthy or AD.
+These four factors were semantic impairment, acoustic abnormality, syntactic impairment and information impairment and were based on existing measures of semantic and syntactic complexity.
+Zimmerer (2016) \cite{Zimmerer2016} looked at whether language was more formulaic in those suffering from AD.
+He proposed that those who suffer from AD rely on formulatic sentences, for example 'Noun-Verb-Noun', and this is done to reduce language complexity.
+He noticed a significant difference in the use of formulaic sentences between AD and Healthy Controls.
+\par
+\section{Pragmatic Language}
+The pragmatic language domain refers to the social rules for language for the purposes of communication including, using language to achieve goals, using information from the context to achieve these goals and using the interaction between people to initiate, maintain and terminate conversations.
+\subsection{Coherence}
+Coherence, in lay terms, can be defined as the ability to maintain awareness of the topic at hand.
+It can be separated into local coherence which is related to themes of the immediately preceding utterance and global coherence which looks at how closely an utterance is related to the topic currently being discussed.
+Chapman et al used picture descriptions of Norman Rockwell prints within a frame analysis, with frames being defined as the context in which the picture is viewed \cite{Chapman1998}.
+The authors identified aspects of content, including whether the frames of interpretation being offered were typical, atypical, incorrect or had no frame.
+They also looked propositions supporting frames and propositions disrupting frames as measures of coherence.
+They examined these variables with early stage AD, old-elderly and normal controls.
+Healthy older adults and normal controls produced significantly more typical frames and more frame supporting information than the AD group.
+The authors attributed AD patients' difficulties to memory deficits, attentional deficits, visual perceptual problems, disruption of internalized frame representation, or failure to access frame knowledge.
+\par
+\subsection{Perseveration}
+Perseveration can be broadly defined as the repetition of a response regardless of the absence or cessation of stimulus which could have generated an appropriate response in the first place.
+A typical example could be the idea of conversation moving from introductions to a more general conversation, someone who has difficulties with perseveration will struggle with the move between social contexts and repeat language that would be more used in an introduction context.
+\par 
+One study examined verbal preservation in the description of Norman Rockwell prints \cite{Bayles198}.
+The presented participants with a number of similar pictures that had significantly different contexts and asked participants to describe these prints.
+They divided the total number of words within perseverations by total number of words in the speech sample.
+The authors also calculated rate of perseveration on two other language tasks: confrontation naming and generative naming.
+In all tasks, the AD group produced significantly more perseverations than controls but there were no significant differences between the two groups in the picture task in isolation.
+The authors felt that this was because picture description was an easier task, as it was a visual task in contrast to the other tasks which tapped into other cognitive processes.
+\par
+\subsection{Empty Speech}
+Verbal fluency is a term used in neuropsychological contexts generally referring to timed, word-generation tasks, while in speech-language pathology contexts, "fluency disorders" are defined as interruptions in the flow of speaking characterized by atypical rate, rhythm and repetitions in sounds, syllables, words and phrases.
+"Fluency", in the literature of discourse of adults with AD, typically refers to the smoothness or flow of spoken language.
+Abnormalities of fluency in this population are typically characterised by filled and unfilled pauses, word repetitions, circumlocutions, and revisions.
+\par
+The study of "empty speech" by Nicholas et al was one of the first to examine aspects of fluency in the connected speech of persons with AD \cite{Nicholas1985}.
+They found that adults with AD had significantly more repetitions than controls.
+Similarly, Bayles and Tomoeda found more aborted phrases, revisions and ideational repetitions in the AD group than in controls \cite{Bayles1983}.
+Several other studies support the idea that in AD population there are a greater number of repetitions and revisions than in healthy controls.
+However, there are some studies which contradict these findings\cite{Ahmed2013}.
+\subsection{Conclusions}
+This section has looked at a number of different attributes which researchers have cited as being important in the detection of MCI and AD.
+Whilst in some cases there is a broad consensus about a given feature in the vast majority of cases there is some uncertainty.
+Some of the difficulty lies in different approaches used particularly around the criteria for experimental groups which means that it is difficult to be able to compare studies, like for like, as the populations of the participants being studied vary in subtle or obvious ways.
+It is important for any studies moving forward to use comparable standards when it comes to their experimental groups, such that the picture can be made clearer.
+
+\section{State of literature into Machine Learning and Natural Language processing techniques}
+Diagnosing dementia through language analysis has a large background in terms of psychological research.
+Increasingly current research has called for the use of machine learning as a way of assisting in the process of diagnosis \cite{Orimaye2017, Boschi2017}.
+This section will look at what areas of natural language processing (NLP) and machine learning (ML) could potentially be used as tools to help in this domain as well as any research that has applied machine learning to this problem.
+\subsection{Natural Language Processing}
+Natural Language Processing as a topic can be defined as the intersection between Machine Learning and Linguistics and looks at a number of language tasks, particularly important in this domain is how to enable a computer to process and analyse large amounts of language data.
+Some other tasks include speech recognition, natural language understanding and natural language generation.
+Generally speaking, Natural Language Processing (NLP) systems mimic the semiotic perspective on language.
+That is to say that we can subdivide NLP processes into different components (see Figure 2.3).
+\begin{enumerate}
+	\item Morphological/Lexical: providing the basic language elements or vocabulary, such as words, their roots and inflections.
+	\item Syntactic: for grouping and sequencing elements within samples of language (usually sentences).
+	\item Semantic: for knowing the meaning of an utterance, usually defined in terms of the 'truth value' of the logical propositions that are thought to be expressed by sentences.
+	\item Pragmatic: for understanding the context and purpose of an utterance.
+\end{enumerate}
+\begin{figure}[H]
+\centering
+\includegraphics[width=240px, height=150px]{images/workflow.jpg}
+\caption{How Natural Language Processing tasks are subdivided.\label{white}}
+\end{figure}
+In order to facilitate these and other tasks, a number of frameworks have been developed that automate some of the more common processing tasks. A brief review of these frameworks follows. 
+\subsubsection{Natural Language Toolkit}
+The Natural Language Toolkit (NLTK) was originally designed as part of a computational linguistics course at the University of Pennsylvania but has evolved to become an open-source framework which includes methods and modules for a wide array of Natural Language tasks (see Table 2.2 for a full list) \cite{Bird2009}.
+\begin{table}
+	\begin{tabular}{ | p{4cm} | p{7cm} | p{1cm} |}
+		\hline
+		\textbf{Language processing task} & \textbf{Functionality} \\ \hline
+		Accessing corpora & standardized interfaces to corpora and lexicons \\ \hline
+		String processing & tokenizers, sentence tokenizers, stemmers \\ \hline
+		Collocation discovery & t-test, chi-squared, point-wise mutual information \\ \hline
+		Part-of-Speech tagging & n-gram, backoff, Brill, HMM, TnT \\ \hline
+		Machine Learning & decision tree, maximum entropy, naive Bayes, EM, k-means \\ \hline
+		Chunking & regular expression, n-gram, named-entity \\ \hline
+		Parsing & chart, feature-based, unification, probabilistic, dependency \\ \hline
+		Semantic interpretation & lambda calculus, first-order logic, model checking \\ \hline
+		Evaluation metrics & precision, recall, agreement coefficients \\ \hline
+		Probability and estimation & frequency distributions, smoothed probability distributions \\ \hline
+		Applications & Graphical concordance, parsers, WordNet browser, chatbots \\ \hline
+		Linguistic Framework & manipulate data in SIL Toolbox format \\ \hline
+	\end{tabular}
+		\caption{\label{tab:table-name}NLTK tasks and functionality}
+\end{table}
+The NLTK has been successfully used in both teaching and research purposes and as part of this provides access to large array of databases and corpora which have potential uses in common tasks.
+
+\subsubsection{CoreNLP}
+Stanford CoreNLP provides a set of human language technology tools.
+This comes as an installable package that requires JAVA and can be interacted with through the command line or by an accompanying API.
+Whilst NLTK is much more oriented to parsing text and generating features, CoreNLP does this and adds more functionality such as a Named Entity Recogniser and sentiment analysis functions.
+At it's centre is the Stanford parser, which parses text and can give the base forms of words, their parts of speech, whether they are names of companies, people, etc., normalize dates, times, and numeric quantities, mark up the structure of sentences in terms of phrases and syntactic dependencies, indicate which noun phrases refer to the same entities, indicate sentiment, extract particular or open-class relations between entity mentions, get the quotes people said, etc.
+\par
+\begin{figure}[H]
+\centering
+\includegraphics[width=350px, height=200px]{images/CoreNLP.png}
+\caption{Depiction of how Core NLP marks up a sentence}
+\end{figure}
+It's intended for CoreNLP to provide tools and frameworks that will enable higher level language analysis whilst remaining domain agnostic.
+
+\subsubsection{Linguistic Inquiry and Word Count}
+The Linguistic Inquiry and Word Count (LIWC) is a piece of software that counts words and assigns these counts to categories \cite{Pennebaker2015}.
+The rationale behind this is that the words we speak inform an observer of the state of mind of an individual.
+A typical example of this might be someone with depression who may typically be more inwardly focused and therefore talking in the first person is common\cite{Al-Mosaiwi2018}.
+The 2015 version of the LIWC has over 90 different categories \cite{Pennebaker2015} and some supporting research suggests that it is effective at identifying themes and associating these themes with mental health difficulties such as depression and anxiety \cite{Sonnenschein2018}.
+\par
+Whilst not necessarily a tool that can parse data and look at trends from a linguistic perspective, a framework like this could be used as another tool to explore themes within a persons speech from a more qualitative angle.
+To the authors knowledge, there is no research that looks at speech changes from this perspective.
+
+\subsection{Traditional methods of Machine Learning}
+In terms of Machine Learning research in this area, a number of researchers have used transcripts based on picture description tasks \cite{Zimmerer2016, Orimaye2017, Mueller2018a, Fraser2015} and have successfully extracted linguistic features that could differentiate between AD and controls.
+\par
+One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the most frequent class.
+They detail several lexical approaches to the problem of detecting and rating AD.
+The approaches they looked at relied primarily on character n-gram techniques but also explore correlation of usage of frequency of different types of speech.
+Their results act as a proof in concept of the utility of using a pure computational approach to the diagnosis of dementia using spontaneous speech.
+The were able to obtain 95\% accuracy dementia vs controls, 70\% accuracy classifying dementia into two categories and 50\% accuracy classifying dementia into four categories.
+They suggest further exploration of characteristics to classify \cite{Thomas2005}.
+\par 
+% Cite CRUM Population-based norms for the Mini-Mental Status Examination by age and educational level
+Bucks et al \cite{Bucks2000} took a more traditional approach using frequency rates per 100 words of various word types such as nouns, verbs, adjectives and pronouns and some measures of lexical richness to attempt to discriminate between healthy older participants and participants with probable AD.
+They found that AD patients had higher mean pronoun rate, adjective rate and verb rate but lower noun rate vs normal older controls \cite{Bucks2000}.
+It's important to note that the clinical group averaged a score on the MMSE of 15 (sd=6.8) which puts the clinical group in the 'moderately impaired range'.
+At this stage, the two groups should be reasonably easy to differentiate in terms of language and other cognitive abilities.
+\par 
+Fraser et al \cite{Fraser2015} built upon the work by Bucks by adding an acoustic analysis element to her analysis using the DementiaBank corpus, as well as significantly expanding the list of linguistic features they used.
+They used features generated from the Stanford parser (Now known as Core NLP) and computed the frequency of the occurrence of different parts of speech which they then normalised by the total number of words in each utterance.
+They also computed ratios between these parts of speech such as pronouns to nouns.
+They also included measures of syntactic complexity such as mean length of sentences and measures of lexical richness such as TTR.
+Their analysis showed was able to achieve 81\% accuracy in distinguishing individuals with AD from controls based on short samples of speech.
+They found that there are four factors which they felt were important in discriminating between these two groups: semantic impairment, acoustic abnormality, syntactic impairment and information impairment.
+\par 
+Zimmerer et al \cite{Zimmerer2016} took a slightly different approach in that they developed the Frequency in Language Analysis Tool (FLAT) which looked at the degree of formulaicity in an individual sample.
+The rationale behind this is that people with impaired cognitive function are able to depend on formulaic sentences for language generation to ease the burden of language generation on cognitive symptoms.
+They were able find that those with probable AD produced more formulaic language than controls.
+\par 
 Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls.
@@
 Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
-\newline
 \par
-Asgari, Kaye and Dodge (2017) \cite{Asgari2017} used another form of word frequency measurement.
-Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words \cite{Pennebaker2015}.
-They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
-\newline
+
+\subsection{The case for Deep Learning}
+One of the criticisms of traditional learning models is it's reliance on features that are generated for them.
+As we can see from the brief look at some of the research in this field, researchers have taken a number of different approaches with success.
+However, there does not seem to be a consensus.
+As machine learning in general moves away from traditional machine learning models to deep learning techniques, it poses the question can deep learning assist with this problem?
 \par 
-\subsection{Mean length of utterance (MLU)}
-Murray found that MLU was not a distinguishing factor among health adults, adults with depression and adults with AD.
-Ripich et al found a decrease in MLU in adults with severe AD over time, and this was supported by findings of Le et al in their studies of authors \cite{Le2012} \newline
+One of the main benefits of deep learning is that it does not rely on pre defined features being fed to the model.
+Instead, deep learning models take raw data in with some amount of preprocessing and generates it's own features.
+This move away from relying on features solves one of the difficulties we have with the current psychological literature, namely that there are some disagreement on the which
 \par 
-\subsection{Proportion of verbs to nouns plus verbs}
-Kave and Levy used a verb index to capture syntactic complexity and found that adults with AD expressed the same amount of verbs to nouns plus verbs as adult controls. \newline
+One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
+This is the case with all machine learning techniques, but more so with deep learning.
+There have been ways to 'create' more data such as data augmentation.
+
+\subsection{Conclusions}
+We can see that both Natural Language Processing and Machine Learning techniques have a lot to offer, Indeed there has been a lot of research which have used pre-existing datasets to explore this area with promising results.
+One of the difficulties with the current research is the approach of trying to discriminate between those AD and healthy controls.
+This is not necessarily a problem in the real world as it is trivially easy to do for trained clinicians.
+A more interesting, but potentially harder to problem to solve is to discriminate between those with MCI and healthy controls, and more importantly to track their decline over time.
+Further, to date the authors are unaware of any research in which these techniques are applied to newly created samples of language.
+
+\section{Discussion}
+Current state-of-the-art diagnostic measures of AD are invasive (CSF analysis), expensive (neuroimaging), and time-consuming (neuropsychological assessment).
+Furthermore, these measures are limited to speciality clinics and thus have limited accessibility as frontline screening and diagnostic tools for AD.
+More importantly, nonspecialists are often inaccurate at identifying early AD and MCI.
+Thus, there is an increasing need for additional noninvasive and/or cost-effective tools, allowing effective frontline identification of subjects in the preclinical or early clinical stages of AD who could be suitable for monitoring in speciality clinics and for early treatment.
+Implementation of effective screening instruments will allow diagnosis earlier in the course of dementia, even at the point when memory function is still essentially within the normal range.
+This strategy would enable an earlier, and potentially more effective, prevention and treatment of AD with a special focus to preserve cognitive functions.
 \par 
-\subsection{Syntactic Complexity - Composite measures of MLU, syntactic errors and verbs}
-Ahmed et al, and Ahmed, Haigh et al found differences in syntactic complexity between adults with MCI and controls, and between MCI and moderate AD stages.
-The differences in syntactic complexity were not significant when individual measures were tested, but were apparent using a composite score consisting of MLU, words in sentences, syntactic errors, nouns with determiners, and verbs with inflections.
-\newline
+However the literature has identified a number of challenges when approaching this problem.
+Firstly, the clinical features that combine to meet the diagnostic criteria for Dementia or it's variants are continuous in nature and heterogenous between patients and are also impacted by other variables.
+For example as shown above, cognitive performance is affected in part by a patients educational attainment and a patients ability to live independently is impacted by a their physical health as much as their cognitive health.
+The challenge therefore is to find features that are minimally impacted by other factors, or that can be controlled for by a experimental design such as a matched pairs design to control for educational attainment.
 \par 
-\subsection{Formulaic Language}
-Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
-They found that there were four factors which informed the classification of participants as either healthy or AD.
-These four factors were semantic impairment, acoustic abnormality, syntactic impairment and information impairment and were based on existing measures of semantic and syntactic complexity.
-Zimmerer (2016) \cite{Zimmerer2016} looked at whether language was more formulaic in those suffering from AD.
-He proposed that those who suffer from AD rely on formulatic sentences, for example 'Noun-Verb-Noun', and this is done to reduce language complexity.
-He noticed a significant difference in the use of formulaic sentences between AD and Healthy Controls.
-\newline
+Another challenge lies with the recruitment of suitable individuals who may notice a decline in cognition to the point where we might classify them as having MCI, but these individuals deduce that there is little to no value to admitting there is a problem and seeking help whilst their symptoms are 'manageable'.
 \par
-\section{Pragmatic Language}
-The pragmatic language domain refers to the social rules for language for the purposes of communication including, using language to achieve goals, using information from the context to achieve these goals and using the interaction between people to initiate, maintain and terminate conversations.
-\subsection{Questions, turn-taking, unsure statements, egocentric comments}
-One study, Ripich et al, examined several pragmatic abilities among patients at different stages of AD.
-The severe AD group asked more questions over time than the other group.
-The authors argued that question-asking was a compensatory strategy, and as a result, adults in late-stage AD may have had insight into their communication problems.
-Potentially, there were a number of flaws of the methodology if their research.
-Firstly, a caregiver was asked to administer the picture description task in order to mirror a more typical communicative interaction.
-Whilst this improved the ecological validity of the study, the caregivers delivery would be varied and inconsistent.
-Secondly, due to the constrained nature of the picture description task, it is unlikely that it would be able to capture the pragmatic skills that are typioical of conversations in everyday life.
-\newline
-\par 
-\subsection{Coherence}
-Coherence refers to the appropriate maintenance of topic in discourse and is thematically related to the immediately preceding utterance (local coherence) and by how closely an utterance relates to the general topic at hand (global coherence).
-One study looked at coherence, Chapman et al used picture descriptions of Norman Rockwell prints within a frame analysis, with frames being defined as internalized knowledge structures.
-The authors identified aspects of content, including typical frames of interpretation, atypical, incorrect or no frames, propositions supporting frames and propositions disrupting frames as measures of coherence.
-They examined these variables with early stage AD, old-elderly and normal controls.
-Old-elderly and normal controls produced significantly more typical frames and more frame supporting information that the AD group.
-The authors attributed AD patients' difficulties to memory deficits, attentional deficits, visual perceptual problems, disruption of internalized frame representation, or failure to access frame knowledge.
-\newline
-\par 
-\subsection{Perseveration}
-One study examined verbal preservation in the description of Norman Rockwell prints, dividing the total number of words within perseverations by total number of words in the speech sample.
-The authors also calculated rate of perseveration on two other language tasks: confrontation naming and generative naming.
-Across all tasks, the AD group produced significantly more perseverations than controls; however, on the picture description task alone, there were no significant differences between the two groups.
-The authors theorised that this was because picture description was an easier taks because there was the visual stimulus, similar to the argument made by Bschor et al.
-\newline
-\par 
-\subsection{Speech intonation / prosody}
-There are a number of studies, that have examined "melodic line", which is a subjective measure of speech prosody defined as "the appropriate use of intonational contour, including alterations in pitch, volume and duration".
-For instance, Forbes-McKay compared melocdic line using a "simple" picuture such as the Cookie Theft picture vs a complex picture.
-The number of pictorial themes differentiated the simple tasks from the complex.
-Results showed no group differences on simple picture tasks (Cookie Theft and Tripping Woman) but there were differences in melodic line for the complex picture tasks.
-Fraser et al examined several acoustic features of speech in patients with AD ussing machine learning methods that captureed both rate and phonation patterns, and found acoustic abnormalities to be a significant factor.
-Konig et al, also used an automated feature analysis examining vocal and temporal aspects of speech among controls, and patients with MCI or AD, and reported a classification of 81\%.
-\subsection{Discourse Fluency}
-Verbal fluency is a term used in neuropsychological contexts generally referring to timed, word-generation tasks, while in speech-language pathology contexts, "fluency disorders" are defined as interruptions int he flow of speaking characterized by atypical rate, rhythm and repetitions in sounds, syllables, words and phrases.
-"Fluency", in the literature of discourse of adults with AD, typically refers to the smoothness or flow of spoken language.
-Abnormalities of fluency in this population are typically characterised by filled and unfilled pauses, word repetitions, circumlocutions, and revisions.
-In contrast with fluency disorders (i.e.
-stuttering), abnormailities in the fluency of adults with MCI or AD are rarely manifested at the phonological level.
-\newline
-\par
-The study of "empty speech" by Nicholas et al was one of the first to examine aspects of fluency in the connected speech of persons with AD.
-They found that adults with AD had significantly more repetitions than controls.
-Similarly, Tomoeda et al found more aborted phrases, revisions and ideational repetitions in the AD group than in controls.
-Several other studies support the idea that in AD population there are a greater number of repetitions and revisions than in healthy controls.
-However, there are some studies which contradict these findings (Ahmed et al).
-\newline
-\par 
-\subsection{Speech Monitoring}
-Speech monitoring is related not only to word retrieval deficits and to pragmatic language skill but also to "anosognosia" define as the awareness of one's own deficits.
-McNamara et al investigated word error monitoring skills by comparing uncorrected and repaid errors in adults with AD vs health controls.
-The AD group was equally impaired in error monitoring as the controls.
-Severity of naming deficits correlated negatively to the amount of uncorrected errors.
-The authors suggested that this impairment was related to the executive function dificulties in the clinical groups.
-The authors did not report correlations between error monitoring and executive function test scores, however, which could have strengthened that hypothesis.
-\newline
-\par 
-Another measure of speech monitoring used in the picture description literature is "response to word-finding delays," defined as "whether patients appear unaware of their problem, produce an approximation of the target word or actively search and produce the target word".
-Response to word-finding delays differed significantly between minimal  AD and normal controls (forbes-McKay and Venneri).
-The measure was based on Goodglass and Kaplan's scale for rating discourse on the Boston Diagnostic Aphasia Examination (BDAE) and is composed of clinical judgement of behaviour that is rated on a likert-type scale ranging fdrom 1 to 7 (7 being no abnormality).
+Finally there is a large amount of variability in the presentations of those with MCI and early dementia, and this is compounded by an similar amount of variability in the criteria researchers have used for experimental groups and the approaches researchers have used to tackle this problem.
+This had led to a confused literature.
+Recently a call for research that has consistent inclusion / exclusion criteria has been made along with some proposed definitions of MCI and it's subgroups \cite{Petersen2014}.
+Researchers have identified the analysis of language impairment as an area of promise to explore in the diagnosis of MCI and early AD and recent developments in natural language processing and machine learning techniques have the potential to assist in this research.
+Indeed, given the increased burden on the diagnosis of MCI and AD on professionals there has been a call to use technology to potentially ease this burden \cite{Boschi2017}.
+A small but growing amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
 
-\section{State of literature into Natural Language processing techniques}
-As we can see, diagnosing dementia through analysis has a large background in terms of psychological research.
-However, a lot of more current research has called for the use of machine learning as a way of assisting in the process of diagnosis \cite{Boschi2017}.
-\subsection{Traditional methods of Machine Learning and Natural Language Processing}
-
-\subsection{Deep Learning}
-
-%% IS THERE ANY RESEARCH ON THIS?
-
-%Current state-of-the-art diagnostic measures of AD are
-%invasive (CSF analysis), expensive (neuroimaging), and
-%time-consuming (neuropsychological assessment). Furthermore,
-%these measures are limited to specialty clinics and
-%thus have limited accessibility as high-throughput, or frontline,
-%screening and diagnostic tools for AD. More importantly,
-%nonspecialists are often inaccurate at identifying
-%early AD and MCI [5]. Thus, there is an increasing need
-%for additional noninvasive and/or cost-effective tools, allowing
-%frontline identification of subjects in the preclinical or
-%early clinical stages of AD who could be suitable for monitoring
-%in specialty clinics and for early treatment. Implementation
-%of effective screening instruments will allow diagnosis
-%earlier in the course of dementia, even at the point when
-%memory function is still essentially within the normal range.
-%This strategy would enable an earlier, and potentially more
-%effective, prevention and treatment of AD with a special
-%focus to preserve cognitive functions
-\section{The problems with attempting early diagnosis of dementia and the ethics of actively screening for it}
-It is challenging to characterise Dementia as a disease as it's more syndrome.
-The clinical features that combine to meet the diagnostic criteria for Dementia or it's variants are continuous in nature and also impacted by other variables.
-For example, cognitive performance is affected by previous education and ability to live independently is impacted by a person's physical health.
-Also, the individual who may develop symptoms that indicate the onset of dementia needs the insight to recognise them, or family members will need to do so and also deduce that there is value to admitting there is a problem and seeking help before the deterioration is too great.
-\newline
-\par 
-To detect dementia early requires an understanding of how this presents early in it's progression, but as described above there are a number of potential ways in which individuals with dementia could be impacted but no consistent list of criteria has been produced.
-Even for AD, the most prevalent form of dementia, clinical diagnosis can only be confirmed after death through autopsy.
-In addition, population studies have shown that people over the age of 80 show similar changes in the brain in those that do not have dementia.
-Therefore, it's difficult to say with certainty that any symptoms can definitely cause dementia despite the fact that these symptoms are commonly present in those that have dementia.
-\newline
-\par 
-\section{Conclusions and Future Work}
-It appears that there are a number of approaches that have been shown to be effective in various experimental settings.
-The fact that there are a numerous array of different perspectives to tackling this question, and even within these perspectives there are a range of various methodologies leaves the field with a sense of confusion as to the most effective way to solving the problem.
-\newline
-\par
-Identification of language impairment is key idea in the diagnosis of AD and early diagnosis can alter the prognosis and change the management of this degenerative disorder as well as provide new opportunities to study the progression of the disease.
-Given the burden on the diagnosis of AD on professionals there has been a call to use technology to potentially ease this burden \cite{Boschi2017}.
-It has already been shown that analysis of speech and language has shown markers that pre-date the official diagnosis of dementia \cite{Snowdon1996, Berisha2015}.
-A significant amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
-\newline
-\par
+\subsection{Future Work}
 One of the areas for research to study is a careful examination of the features that are being used to measure language deterioration.
 For example, Zimmerer (2016) \cite{Zimmerer2016} describes connectivity in such as a way that correlates directly with what Mueller (2018) \cite{Mueller2018a} calls Fluency.
-Whilst these are very nuanced measures which differ slightly in the form they take, the sheer range of measures and features being produced make it difficult to conceptualise as part of the larger problem.
+Whilst these are very nuanced measures which differ slightly in the form they take, the sheer range of measures and features being produced make it difficult to organise and explore what is truly going on in those with MCI and early AD..
 Some work needs to be done in producing a consistent list of measures that are validated using existing datasets and can be used for future research moving forward.
@@
 \par
-\subsection{Future Work}
-Future research should also be directed towards developing non-intrusive ways of detecting subtle changes in language based in the home, such that any deterioration that could indicate the presence of MCI or AD and be flagged up early in the progression of any potential cogntive impairment for further review via referral.
-Machine learning approaches seem to be the most logical approach for achieving this aim.
-Further, despite the quality of datasets being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which we could claim are limited in scope.
-\newline
+Teng et al suggests that work should focus on the MCI population and concentrate on developing a consensus neuropsychological battery that could yield predictable rates of progression to AD \cite{Teng2009}.
+This, in conjunction with the development of a model of language, sensitive enough to detect subtle deterioration in language use to act as an additional cognitive marker to aid diagnosis could potentially move some way to providing this.
 \par 
-The recent resurgence in the use of neural networks and deep learning is because of it's significantly improved performance on many problems and it's ability to scale from small to large datasets.
-Another key benefit of deep learning is it's ability to automate the process of feature engineering.
-Feature engineering, with this particular problem is an interesting debate as there are currently numerous ways in which to try to generate features as I have described above.\newline
+Future research should also be directed towards developing non-intrusive ways of detecting subtle changes in natural language such that any perceived deterioration that could indicate the presence of MCI or AD could be flagged up early.
+Machine learning approaches seem to be the most logical approach for achieving this aim as language could be collected in non-intrusive ways and passed to a machine learning algorithm for preliminary classification.
+Despite the excellent quality of datasets, for example the DementiaBank dataset, being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which some researchers could claim are limited in scope.
+Finally, the recent resurgence in the use of neural networks and deep learning could provide the answer to the confused literature in terms of features.
+A key benefit of deep learning is it's ability to automate the process of feature engineering.
+So there is an opportunity to explore the use of deep learning, to not only develop new features but also validate existing features independently.
 \par
 This area of research is extremely promising in its early results and the impact of successful research would be life changing for both individuals and the health of the worlds aging population in general.
-\newline
+\chapter{Initial Study of the Presidents corpus}
+\section{Background}
+There has been a significant research in the area of language deterioration as a means of detecting Alzheimer's Disease.
+This usually takes the form analysis of speech recorded as part of a cognitive assessment such as the Picture Description Task \cite{Orimaye2014,Fraser2015}.
+Given that language samples are relatively easy to collect, research has moved towards analysis of spontaneous speech.
+An good example of this type of research is the study conducted by Berisha and Liss which looked at the differences in language use between two US presidents, Ronald Reagan (who would go on to receive a diagnosis of Dementia) and George H.W.
+Bush who acted as a matched control based on Age \cite{Berisha2015}.
+They found several differences in language use which they felt acted as indicators of Reagan's difficulties with language due to dementia.
+These significant differences were in the number of unique words used per speech, the use of non-specific nouns and fillers and low-imageability verbs \cite{Berisha2015}.
+\par   
+This study replicates work done by Berisha and Liss and extends this by adding Donald Trump as an alternative, more appropriate comparison to Ronald Reagan as he is much closer in age than George H.W.
+Bush.
+This experiment will look at the features originally identified by Berisha and Liss, as well as any others that have potential as discussed in the literature review above.
+\section{Methods}
+I took 46 transcripts of Ronald Reagan’s (RR) press conferences from 1981 to 1988 and compared them with 134 press conferences by George H.
+W.
+Bush (GHWB) and 29 press conferences conducted by Donald J.
+Trump (DJT).
+I analyzed transcripts for lexical features shown to change longitudinally with dementia  (for a comprehensive review of these, see the literature review above).
+For this collection of documents, I generated a number of features which looked at a number of different aspects of each document.
+These features encompassed, word level, sentence level and document level features and included a number of features contained in the study by Berisha and Liss with the aim of replicating and extending on their findings.
+These findings were.
+number of unique words, non-specific nouns and fillers and low imageability (LI) verbs.
+Imageability is characterized, according to Berisha and Liss, as the ease with which a term gives rise to a sensory .mental image.
+I compared the trends described in the transcripts of RR and GHWB, but also included DJT.
+Berisha and Liss originally made the comparison as it GHWB (GHWB - age at the start of presidency - 64 years, 222 days) was the closest match to RR in terms of age (RR  - age at the start of presidency, 69 yeaars and 349 days).
+However, with the inauguration of Trump, he now is the closest comparible president in terms of age (DJT - age at the start of presidency - 70 years, 220 days).
+It would be interesting to look at a comparison of RR and DJT to see whether the comparisons made by Berisha and Liss hold true with this more appropriate match (in terms of age).
+DJT as with GHWB has no known diagnosis of AD.
+I used the press conference transcripts in the American Presidency Project (APP) archive as a data source for this project.
+The APP is a comprehensive and organized searchable database of presidential documents, including transcripts of speeches, transcripts of news conferences, and other public documents.
+\subsection{Pre-processing}
+To generate the files necessary for analysis, I downloaded each transcript and performed the following changes.
+I omitted the prepared statement by the president and any speech by other individuals.
+I started each transcript at the beginning of the first answer to a question.
+I filtered any annotations that were added to the transcript, including any references or clarifications, and any laughter.
+It's worth noting that there appears to be a difference in how 'hesitations' were marked down between each president, for RR hesitations were marked by a single hyphen whereas for GHWB hesitations are marked by a double hyphen.
+In order to maintain consistency when parsing through the documents, I have changed both types of hesitation to be marked by a single hyphen.
+I also omitted one word sentences as this data would, from a theoretical perspective, not be relevant for language analysis.
+I did not control for the length of the document, but generated features which would normalise by the length of the document.
+I therefore was able to include all press conferences by both Ronald Reagan and George H.W.
+Bush where there was a question and answer session conducted at least in part by the sitting president (2 press conferences of GHWB were ommited due to a lack of a question and answer session).
+\subsection{Feature Selection}
+We calculated the following features for each transcript in turn using the NLTK (see section 2 for a description) \cite{Bird2009} and Python. 
+\par 
+\subsubsection{Measures of lexical variation}
+We constructed two features of lexical variation.
+Firstly we looked at the number of unique words.
+To do this we were able to split each transcript into individual words and changed them to lowercase using NLTK and were then count the number of unique words that appeared in each transcript.
+We also used the TTR formula (see section 2 for a description) for a feature that measures lexical diversity independent of sample size \cite{Richards1987}.
+\subsubsection{Fillers, Non-Specific Nouns and LI Verbs}
+For these features, we counted the number of occurrences for different categories (see table for list of categories tracked and the words counted).
+The features were used by Berisha and Liss in their research \cite{Berisha2015} and were taken from work done by Bird et al \cite{Bird2000}.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{6cm} |}
+		\hline
+		Category & Words \\ \hline
+		Fillers & \textit{"well", "so", "basically", "actually", "literally", "um", "ah"} \\ \hline
+		Non Specific Nouns & \textit{"something", "anything", "thing", "everything"} \\ \hline
+		LI Verbs & \textit{"be", "come", "do", "get", "give", go", "know", "look", "make", "see", "tell", "think", "want"} \\ \hline
+
+	\end{tabular}
+	\caption{\label{tab:table-name}Categories and Words Counted}
+	\end{center} 
+\end{table}
+
+\subsubsection{Usage of parts of speech}
+This section involves using a Part of Speech tagger (PoS) which analyses a sentence and assigns a 'tag' to each word based on the function the word has in a sentence.
+At a basic level this can be divided into the eight defined parts of speech: 'nouns', 'pronouns', verbs', 'adjectives', 'adverbs', 'conjunctions', 'prepositions' and 'interjections' but can be further subcategorised.
+We used the PoS tagger built into NLTK to tag each transcript in turn and used these the counts from each of these eight categories in our analysis.
+In addition to frequency counts we also normalised these features by dividing the frequency count by the number of words in the document to take into account transcript length.
+
+\section{Results}
+One of the most important thing to note is the wide variety of samples between the three presidents and also the varying timescales.
+RR participated in 46 press conferences over eight years (an average of 5.75 a year) which is the fewest number of press conferences given by an American president during their term of office.
+GHWB participated in 136 press conferences over four years (an average of  34 a year) and DJT participated in 29 press conferences to date (an average of 19.3 per year).
+Equally, there are variances in the average number of words.
+RR produced an average of 3424 words per conference compared to 2608 by GHWB (unpaired t = 4.434, p\textless 0.001) and DJT at 1849 words (unpaired t = 6.524, p\textless 0.001).
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR & GHWB & DJT \\ \hline
+		Total Words & 3423.91 (416.42) & 2607.72 (1210.38) & 1848.65 (1549.38) \\ \hline
+		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
+		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
+		LI Verbs & 124.22 (17.89)& 103.45 (51.87) & 84.48 (75.78) \\ \hline
+	\end{tabular}
+	\caption{\label{tab:table-name}Means and Standard Deviations of important features}
+	\end{center} 
+\end{table}
+
+In terms of unique words, we found that RR used significantly more unique words, non-specific nouns and low imageability verbs than GHWB and DJT (see Table 3.3).
+Some of these differences are due to the length of the sample, particularly in the case of DJT where his average sample is almost half the sample of RR.
+It could also be said that this could be down to differences in linguistic abilities or speaking style \cite{Berisha2015, Le2011}.
+However, we can certainly see that as controls GWHB and DJT are comparative in relation to non-specific nouns and LI verbs.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR v GHWB & RR v DJT & GHWB v DJT \\ \hline
+		Total Words & \textbf{4.434***} & \textbf{6.524***} & \textbf{2.899**} \\ \hline
+		Unique Words & \textbf{6.832***} & \textbf{11.403***} & \textbf{4.137***} \\ \hline
+		Non Specific Nouns & \textbf{7.877***} & \textbf{3.426**} & -0.574 \\ \hline
+		LI Verbs & \textbf{2.656**} & \textbf{3.420***} & 1.628 \\ \hline
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+	\end{tabular}
+	\caption{\label{tab:table-name}RR T-tests vs GWB and DJT}
+	\end{center} 
+\end{table}
+
+We then looked at the data from a longitudinal perspective as we are interested seeing whether we can track various language variables and their progress over time.
+We ran a number of Pearsons correlations with transcript index number as a time reference and the dependant variables (Table 3.4).
+For our controls, we found them to be stable for the most part with the main highlights being a decrease in Adverb usage for DJT (R=-0.36, p=0.049) and a steady but not severe decline in a number of variables for GWHB, namely total word count, unique words, low imageability words and verb usage.
+\par 
+For RR, his decline is more marked and more widespread through his language use.
+We noticed an significant increase in adverb (R=0.41, p=0.004) and pronoun usage (R=0.65, p\textless0.001), as well as a slight usage increase in Non-specific nouns(R=0.30, p=0.03).
+There was a highly significant decrease in number of unique words (R=-0.56, p\textless0.001) and noun usage (R=-0.70, p\textless0.001).
+Also very significant decrease in adjective usage (R=-0.40, p=0.005) and a significant decrease in total word count (R=-0.31, p=0.03).
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | l | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR & GHWB & DJT \\ \hline
+		Word Count & \textbf{-0.31*} & \textbf{-0.21*} & 0.08 \\ \hline 
+		Unique Words & \textbf{-0.56***} & \textbf{-0.25**} & 0.16 \\ \hline
+		Non Specific Nouns & \textbf{0.30*} & -0.08 & -0.03 \\ \hline
+		LI Verbs & -0.19 & \textbf{-0.20**} & 0.02 \\ \hline
+		Nouns Normalised & \textbf{-0.70***} & -0.03 & 0.14 \\ \hline
+		Verbs Normalised & \textbf{0.36**} & \textbf{0.24***} & -0.03 \\ \hline
+		Adjectives Normalised & \textbf{-0.40**} & 0.08 & -0.34 \\ \hline
+		Adverbs Normalised & \textbf{0.41***} & 0.02 & \textbf{-0.36*} \\ \hline
+		Pronouns Normalised & \textbf{0.65***} & 0.13 & 0.07 \\ \hline
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+	\end{tabular}
+	\caption{\label{tab:table-name}Pearson Correlations for Features}
+	\end{center} 
+\end{table}
+
+\section{Discussion}
+President Reagan received his diagnosis of AD in August 1994 but using transcripts of speeches he made in his two terms as President (January 1981 - January 1989) we have be able to identify certain changes in his use of language that we might ascribe to the onset of MCI and early AD.
+Despite differences in our methodology, our research supports the findings of Berisha and Liss in that we both find a significant decrease in unique words over time and an increase in non-specifc noun usage.
+Compared to our controls (GWHB andDJT), we find some slight trends with GWHB but no such trends with DJT in his speech albeit his samples of speech span a shorter amount of time.
+\begin{figure}[H]
+	\centering
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/RRUniqueWords.png}
+		\caption{Ronald Reagan - Unique Words over time}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/GWHBUniqueWords.png}
+		\caption{George H.W. Bush - Unique Words over time}
+	\end{minipage}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/RRNSNouns.png}
+		\caption{Ronald Reagan - Non-specifc Nouns over time}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/DJTNSNouns.png}
+		\caption{Donald J. Trump - Non-specific Nouns over time}
+	\end{minipage}
+\end{figure}
+
+A criticism of Berisha and Liss's work is the problems they had with normalising the transcripts in terms of length.
+This was also a problem in the work of Garrard et al \cite{Garrard2005, Le2011}.
+Whilst it is important to control for outliers, there are other ways in which we can control for length of sample.
+\par 
+Interestingly, when we normalised the various types of words used by the presidents we found some interesting patterns that further differentiated RR from the controls.
+Whilst Non-specific nouns increased over time, we found that noun usage in general significantly decreased and pronouns increased similarly significantly.
+The increase in pronoun for those with early AD has been identified in literature, although there are only a few studies that explore this \cite{Almor1999, Wendelstein2015}.
+Wendlestein et al propose that the increased used of pronouns is an expression of an impaired ability to adapt language to the listener's needs \cite{Wendelstein2015}.
+Almor et al attributed this reliance on pronouns due to a impaired working memory \cite{Almor1999}.
+\par 
+The decrease in overall noun usage has also been identified as a feature.
+Jarrold et al found that AD patients would use more pronouns, verbs and fewer nouns than controls \cite{Jarrold2014}.
+Wendlestein in their investigations into noun usage found that decreased later on in AD progression and was unaffected in the pre-clinical stages of AD \cite{Wendelstein2014}.
+Our results are supported by existing literature and this potentially means that language analysis in the way we have structured it may have diagnostic or prognostic properties.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=240px, height=150px]{images/RRNounsNormalised.png}
+	\caption{Ronald Reagan - Nouns Normalised over time}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=240px, height=150px]{images/RRPronouns.png}
+	\caption{Ronald Reagan - Pronouns Normalised over time}
+\end{figure}
+
+There are limitations of this research.
+Whilst in terms of age, DJT is certainly more suitable as a control to match with RR, in some ways they held very different styles of press conferences in that RR preferred to do solo press conferences and DJT has shown a preference for doing joint press conferences which have an impact on the amount of language produced.
+This artifact of the data is in itself notable as it illustrates the problems we may have with smaller amounts of speech.
+Also, the problem of finding an appropriate control is a common one in this domain.
+Given that those with MCI and early dementia have such variable presentations, it might prove of limited value in matched pairs design.
+\par 
+With further work, it is not feasible to the vast array of samples over a timeframe, as we have had with the president corpus and so it would be worth exploring how the quality of these predictions might lessen when faced with considerably fewer samples and over a smaller time period.
+It would also be worth extending this research further to encompass more of the linguistic features Fraser used in her work \cite{Fraser2015} to see if there are any further insights to be gained.
+In addition, this replication and extension has demonstrated the potential utility of using longitudinal data as a means of comparing language use of a person at two or more time periods and using this information as a diagnostic aid for MCI and therefore more work would be helpful from a longitudinal perspective to see if this approach may be valid in moving towards a solution for this particular problem.
+\par 
+The results of this work show that we can track a person's use of language through time in a number of ways, and it is possible for an individual to be his or her own control.
+This is important as it means the heterogenous nature of the MCI population does not impact results as much as if we were comparing those with MCI to controls.
+Equally, it would be helpful to have controls to ascertain what would be usual to expect in the decline of language in a healthy older adult.
+\chapter{Future Work}
+\section{Introduction}
+This section looks at my research aims and objectives given the literature review presented above.
+This section will then go on to explore any challenges to my research.
+I then present the work I have planned for the next two years, divided into distinct tasks with their own aims and objectives.
+As part of this I will identify target conferences where I could present my work.
+Finally, I look at potential risks and ethical concerns that could conceivably occur.
+\section{Research Aims and Objectives}
+The aim of my research is to ultimately to identify more accurate and efficient ways of detecting MCI and early dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRIs and EEGs) by utilising spontaneous speech.
+There is a lot of research into the analysis of language as a marker for MCI \cite{Taler2008, Boschi2017} and there have been calls for research that explores the use of Natural Language Processing and Machine Learning techniques to aid in the analysis of language \cite{Boschi2017}.
+This will be the focus of my research.
+\par
+I have deliberately selected the MCI population as examining language decline at this stage has the potential to further our understanding of how MCI progresses into dementia.
+Given that sampling a person's language is relatively effortless, my proposed research will look at whether we can find markers of MCI in natural language.
+I plan to explore the use of Natural Language Processing and Machine Learning techniques to assist in this process.
+There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia.
+As documented above, these studies have identified a number of speech and language differences between healthy controls and those with dementia.
+Some of these features have been incorporated in work using machine learning to classify healthy controls from those with dementia.
+However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'.
+While this is helpful to establish that machine learning can be used to differentiate between healthy controls and those with dementia, there has been little work looking at those in the MCI phase although it is apparent that language is still affected \cite{Berisha2015, Snowdon1996, Le2011}.
+Therefore we would like to clarify whether natural language generated outside the framework of a deliberate language generating task, has the same ability to discriminate.
+\par
+Machine Learning itself is undergoing a shift in focus at the present time.
+I will describe this shift using the context of my proposed research.
+In traditional machine learning, the transcripts from cognitive assessments are used to generate abstract features such as idea density and more concrete features such as mean length of sentence or number of unique words.
+These features are then used to classify patients.
+As we have seen above, the picture with regards to what features deem important in this area is somewhat muddy.
+Machine Learning has a number of tools at it's disposal to aid this.
+The first is to statistically determine, given a model, which features contribute most highly to a given prediction.
+The second is to use a deep learning model which has the ability to automate the process of feature engineering.
+\par 
+The broad objectives of my research are to explore whether the application of Natural Language Processing and Machine Learning techniques to spontaneous speech samples can improve our ability to diagnose Mild Cognitive Impairment in patients that will directly lead to Alzheimer's Disease.
+Another objective of my research is to provide some clarity as to which language features are most important to measure when looking for evidence of MCI and/or AD.
+Finally, an additional outcome of my research is the development of a repeatable battery of neuropsychological tests that specifically focus on being sensitive to Mild Cognitive Impairment.
+\begin{itemize}
+	\item \textbf{(H1)} Monitoring language samples and other measures of cognitive decline over the course of the year will be enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment
+	\item \textbf{(H2)} Those with MCI will show a different language profile to controls if matched similarly for age, gender, education and other significant factors.
+	\item \textbf{(H3)} The new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures such as the RBANS.
+	\item \textbf{(H4)} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, those with Alzheimers Disease and decline in those with MCI.
+	\item \textbf{(H5)} Using individuals as their own control will be enough to identify trends in the decline of their use of language.
+\end{itemize}
+
+\section{Project Plan and Task Definitions}
+In this section, I describe each task in detail along with a projected timeline depicted in a Gantt chart.
+\subsection{Task 2 - Pilot Study for the Neuropsychological Battery of Tests}
+The aim of this study is complete a trial run of the administration of a new battery of neuropsychological battery of tests.
+We aim to find out whether the tests in question generate enough language so that features such as those generated in Task 1 are meaningful and can be used as language markers that can then be tracked over time.
+We also aim to look at whether there are significant practice effects between three versions of these tests with a gap of seven days between each test.
+\par 
+We compiled a series of tests which look at a number of cognitive functions which have previously demonstrated a level of effectiveness in discriminating an MCI population from a healthy population (see literature review for a comprehensive list, explanation and rationale for each test).
+In addition to these tests, we have incorporated two language tasks in which a participant's ability to generate spontaneous speech that we can use to further analyse.
+As a planned future study is longitudinal, it is important that this battery of tests is repeatable but does not have significant practice effects that may interfere with our ability to track cognitive decline.
+\par 
+We have recruited a cohort of 4 volunteers who will complete the battery of neuropsychological tests on three separate days with a gap of at least seven days between each administration.
+The results of which will compiled and analysed.
+Our hypothesis is that there would be no difference in performance between one battery of tests and another.
+An additional hypothesis is that the language tasks included on each test would be enough to collect an appropriate sample for linguistic analysis such that we may be able to detect cognitive decline in a clinical population.
+\par 
+While the pilot study in itself is not valid for submission to a conference or journal, it is hoped that the results of the study will allow subsequent studies using the same or adjusted battery of tests to be published.
+\subsection{Task 3 - Exploration of language with the DementiaBank dataset and Three Authors dataset.}
+The aim of this task is to replicate and extend the work of Fraser and Orimaye who both used the DementiaBank dataset and set out to use traditional machine learning techniques to see whether they could categorize people into dementia or healthy categories using the samples given.
+We also aim to replicate and extend the work of Le et al who used electronic versions of the novels of three authors.
+Both datasets are longitudinal and have been shown to track language decline over time in different ways.
+\par 
+The DementiaBank dataset was a product of a longitudinal study which was run by the University of Pittsburgh School of Medicine.
+The dataset contains transcripts of interviews conducted with participants with MCI / AD and other forms of dementia.
+These interviews were conducted in English and used the cookie-theft picture language task (see Literature Review for a full description).
+The full data set consists of 314 patients classified as having dementia and 242 patients who are classified as healthy and these patients are tested on a minimum of two separate occasions.
+\par 
+The Three Authors dataset includes the novels of P.D.
+James, Agatha Christie and Iris Murdoch.
+It contains twenty of Murdoch's novels which were published between the ages of 35 and 76, sixteen of Christie's novels written between the ages of 28 and 82 and fifteen of P.D.
+James novels written between the ages of 42 and 82.
+These three authors are notable as Murdoch had a confirmed diagnosis of Dementia after her death, Christie had a suspected diagnosis of Dementia and P.D.
+James was used as a control.
+\par 
+These two existing datasets in addition to the President's Press Conference corpus described above will provide three sources of validation of features that we generate as well as give us clues as how to best process the data we receive from our longitudinal study.
+\par 
+The final output of this task will be a paper which looks at the importance of various word-level and sentence-level features in the deterioration of language for people with AD.
+The target conference for this work is the UK Alzheimer's Research conference in March 2020 with an expected Abstract submission date of December 2019 and the Student Research Workshop of the ACL in July 2020 with an expected submission date of April 2020.
+\subsection{Task 4 - Longditunal Study}
+Previously we have looked at historical datasets, the aim of this task is see whether a formalised collection of specific data, namely our language tasks are enough to be able to be able to detect cognitive decline in participants at different stages of the MCI / Alzheimer's disease progression.
+To this end, we plan on administering our neuropsychological battery of tests from Task 2 to each participant three times (six monthly intervals) over the course of the year.
+\par 
+We therefore plan to recruit a set of 80 participants divided evenly into four groups.
+These groups are Healthy Older Controls (HOC), Early Alzheimers (EA), Subjective Mild Cognitive Impairment (SMCI) and Diagnosed Mild Cognitive Impairment (DMCI).
+We expect to see typical age related decline in our HOC group and more dramatic decline in our EA group.
+We are most interested to see if there is any measurable language decline in the SMCI or DMCI experimental groups.
+\par 
+Whilst we have divided participants into experimental groups we aim to investigate each participant individually.
+Because there is a large amount of heterogeneity in this population, we will initially be analysing each participant's results in isolation with the participant acting as their own control.
+However, we will also be exploring the use of traditional machine learning and/or deep learning across cohorts to investigate whether each cohort has their own distinct 'language profile'.
+\par 
+The final output of this task will be a paper discussing the results of our longitudinal study from a language analysis perspective.
+The target conference for this work will be the UK Alzheimer's Research conference in March 2021 with an expected submission date of December 2020, the Alzheimer's Association International Conference in July 2021 with an expected submission date of January 2021 and the Association of Computational Linguistics Conference (ACL) in July 2021 with an expected submission date of March 2021.
+\subsection{Task 5 - Analysis of performance and validation of the proposed Repeatable Neuropsychological Battery of Tests}  
+In this task, I will be looking at the results of the neuropsychological battery of tests without the additional language tasks to determine whether we can detect cognitive decline. 
+\par 
+We expect to see no decline in our HOC group within this timeframe and more severe decline in our EA group.
+Once again, we will be interested in whether there is any measurable decline in our SMCI or DMCI experimental groups with our battery of tests.
+\par 
+The final output of this task will be a paper discussing the results of the battery of tests, including how the battery does in terms of diagnostic and prognostic utility.
+The target conference for this work is to be decided.
+
+\subsection{Task 6 - Thesis Writing}
+The writing of the thesis, including additional over-running data analysis is planned for the last six months of the project (Oct '20 to Mar '21).
+
+% GANTT CHART
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=350px, height=200px]{images/TaskList.png}
+	\caption{Task List}
+\end{figure}
+
+\begin{sidewaysfigure}[ht]
+	\includegraphics[width=\textwidth]{images/GanttChart.png}
+	\caption{Proposed Timetable.}
+\end{sidewaysfigure}
+
+\section{Research Challenges}
+\subsection{Data Collection}
+There are a few challenges that I have identified with the process of data collection.
+As with all longitudinal studies, the data collected is only valid if there are at least two sets of data points for a given participant.
+Given the population I am exploring, it is expected that there is some level of drop out rate over the course of the year and this has been accounted for.
+Nevertheless, a higher than average drop out rate would weaken the strength of the study, especially if those who dropped out came from the experimental group.
+\subsection{Commonality of MCI}
+Whilst research has consistently shown that those with a diagnosis of MCI tend to have a significantly higher risk of progression to dementia \cite{Roberts2014}.
+The progression at any given time is low and variable.
+Annual conversion rates vary across longitudinal studies conducted in this population, but researchers have shown that there is a higher conversion rate in clinical samples of this population than in community samples \cite{Farias2009}.
+In my proposed longitudinal study, I aim to collect a clinical MCI population, but depending on the availability of such samples I may have to collect from the community.
+If this is the case then, the prevalence of those with MCI who have language deficits within my participants may be low.
+\subsection{Sparsity of Language}
+I have tried to estimate the amount of language I will collect by conducting a pilot study on healthy controls.
+However, until the first sets of data collection with the MCI population, I will not be able to tell if this is a problem.
+I fully expect there to be little difference in the amount of language collected in the MCI population, but this remains a possibility which I will need to adjust for at the 2nd and 3rd data collection phases.
+
+\section{Risk Assessment}
+In this section, I look at the potential difficulties that may produce a delay to the proposed research, define any strategies that have already been put in place to account for this as well as define any other potential solutions.
+\subsection{Delay in data collection}
+When working with a clinical population, the collection of data can be subject to delays.
+This can be for a number of reasons including but not limited to sickness (patient or researcher), unavailability of participant for another reason or unavailability of venue to collect data.
+The collection of data could also be subject to delays due to ethical approval for the research being delayed.
+For individual participants unavailability, flexibility has been built into the data collection schedule to allow for an extra month (capacity for one additional cohort) of data collection per collection cycle.
+\subsection{Rejected Publications}
+For each task defined above, with the exception of task 2 and task 6 I have identified suitable conferences at which I can present my work.
+With the target conferences identified every effort will be made to meet the submission deadlines for these conferences.
+If submissions are rejected at these conferences, I have identified alternative conferences to present at.
+\subsection{Rejected NHS Ethics and/or University Ethics}
+Every effort will be made to submit a complete ethics application through the NHS.
+However, in the event that there is a delay in the ethical approval process then data collection will proceed with a priority placed on the control group.
+Future cohorts will then have proportionally more clinical participants until the research meets it's target numbers.
+In the event that NHS ethics is not granted in time for me to write up, participants will be sought from other areas such as the Alzheimer's Research pool of volunteers and venues where suitable candidates may be present such as meetings of the University of the Third age.
+
+\section{Ethical Considerations}
+In this section I discuss the ethical considerations related to this project and any adjustments that I have made to my research because of these considerations.
+I am aware of the code of human  research ethics published by the British Psychological Society \cite{BritishPsychologicalSociety2014} and my research has been designed in a way that fits with these guidelines.
+\subsection{Risk}
+I will be working with a middle-aged to elderly population who have identified in some way that there is a difficulty with cognition.
+However, I do not anticipate working with vulnerable groups, those lacking capacity or individuals in a dependent or unequal relationship.
+Should any participants I am working with lose capacity during the course of the study, I will remove that participants details and data from the study.
+\subsection{Informed Consent}
+Consent will be sought overtly at the recruitment and initial data collection point by way of an information sheet and associated consent form.
+At other data collection points, I will have information sheets available if participants request them and if such a request is made then I will seek consent again at this stage.
+As part of this process, I will be supplying my participants with my contact details should they have any question or for the purposes of withdrawing consent.
+These consent forms will be stored in a secure, locked cabinet at the university.
+\subsection{Confidentiality}
+Given the nature of the study and the possibility of disclosures of suicidal ideation, it is important that confidentiality is discussed at each data collection point.
+I will also provide an information sheet which details the confidentiality statement and also the situations in which I may need to break confidentiality.
+In addition, I will make a verbal confidentiality statement at each collection point and seek acknowledgement of their understanding and permission to continue.
+The spoken confidentiality statement will be as follows.
+\begin{center}
+	\textit{"Anything you say to me today will be confidential unless I feel that there is a risk to you, or to someone else.
+If this is the case, I may need to speak to or write to third parties such as your GP.
+I will endeavor to discuss this communications with yourself where it is possible to do so.
+Is this ok?"}
+\end{center}
+\subsection{Disclosure of suicidal ideation or other risks}
+As part of the battery of tests for this experiment, there is an opportunity for a participant to reveal that he or she has had suicidal ideation in the past two weeks (Patient Health Questionnaire, Question 9).
+This is something to explore further.
+As I have been trained in psychological risk assessments, it seems appropriate to carry out a further risk assessment.
+Optionally with the tape recorder turned on, or off as requested.
+Any text recorded during the risk assessment will not be included in the language analysis.
+Following this, some information about local services such as Improving Access to Psychological Therapies (IAPT), Samaritans and other appropriate services will be provided.
+A discussion will then be made about whether it is appropriate to notify the GP of the disclosure.
+In the event that there is imminent risk to self or others, a discussion will be made about who I need to contact.
+(Crisis Team and GP).
+In the event that the participant scores 0 on Q9 of the PHQ, no risk assessment will be undertaken.
+
+\section{Thesis - Provisional Table of Contents}
+In this section I provide a provisional table of contents based on the work I have either completed or proposed above.
+\begin{enumerate}
+	\item Introduction
+	\begin{enumerate}
+		\item Overview
+		\item Background and Context
+		\item Aims and Objectives
+		\item Challenges and Achievements
+	\end{enumerate}
+	\item Literature Review
+	\begin{enumerate}
+		\item Introduction
+		\item Assessing Mild Cognitive Impairment using cognitive tests.
+		\item Assessing Mild Cognitive Impairment using Language.
+		\item Natural Language Processing and Machine Learning
+		\item Conclusion
+	\end{enumerate}
+	\item Further Exploration of the Presidents Corpus
+	\begin{enumerate}
+		\item Introduction
+		\item Methodology
+		\item Results
+		\item Discussion
+		\item Conclusion
+	\end{enumerate}
+	\item Validation of a Repeatable Neuropsychological Battery of Tests
+	\begin{enumerate}
+		\item Introduction
+		\item Methodology
+		\item Pilot Study Results
+		\item Longitundinal Study results
+		\item Discussion
+		\item Conclusion
+	\end{enumerate}
+	\item Longitudnal Study
+	\begin{enumerate}
+		\item Introduction
+		\item Methodology
+		\item Experiment 1 Results
+		\item Experiment 2 Results
+		\item Discussion
+		\item Conclusion
+	\end{enumerate}
+	\item General Discussion
+	\begin{enumerate}
+		\item Discussion
+		\item Future Work
+		\item Conclusion
+	\end{enumerate}
+\end{enumerate}
+
+\chapter{Conclusion}
+With the older population of the world continuing to grow year upon year, Dementia in all it's forms is also going to continue to grow.
+Indeed it is already one of the most common causes of death among the elderly \cite{Prince2015}.
+Whilst a lot of research gone into Alzheimer's disease and other forms of dementia, this research thus far has not yielded positive results in terms of our understanding of the disease or of a potential cure.
+Focus therefore should be on the early, pre-clinical stages of dementia to increase our understanding of the disease and to research potential early interventions in those who are likely to go on to develop dementia.
+\par 
+We looked at areas of research in terms of what is available for testing for MCI at the moment and what language markers have been identified as potentially significant.
+We also reviewed the contributions that Natural Language Processing and Machine Learning techniques could make to the field as well as a brief look at some of the key pieces of research that have used techniques.
+The picture is confusing as there has not been much in the way of consistency in researchers results in either of these areas for those with MCI.
+We propose that the problem is due to the heterogeneity of the MCI population and that simple differentiation of MCI or Healthy may not be possible.
+\par 
+We then looked at a potential longitudinal approach, originally done by Berisha and Liss \cite{Berisha2015}.
+We not only replicated a lot of their findings, despite slightly differing methodologies, but we found other language factors which change with the progression of MCI in our results.
+This is significant as it proves that we can use an individual as his or her own control in the measure of language deterioration and this goes some way to solving the problem of heterogeneity in this population.
+\par
+We then go on to propose more research looking at other available longitudinal data (Task 3), as well as proposing a methodology for collecting language data for analysis (Task 2, 5).
+Finally, we propose a longitudinal study (Task 4) which aims to use our newly created methodology for collecting language data in healthy controls, those with probable AD, those subjective cognitive impairment and those with a diagnosis of MCI to see whether we can detect signs of MCI through language analysis.
+\par
+If this work is successful, we have the opportunity to increase our understanding of the very early stages of AD as well as potentially having a new approach to the diagnosis of MCI and AD in the future.
+
 \chapter{Research Questions and Hypotheses}

### 2019-02-05 Jupyter Notebooks/lex_processing.py
Tidy and moved coding to separate repo

@@ -1,74 +1,0 @@
-import math
-import nltk, re, pprint, string
-from nltk import word_tokenize, sent_tokenize
-import datetime
-import pandas as pd
-import numpy as np
-import matplotlib
-import matplotlib.pyplot as plt
- 
-
-def get_julian_datetime(date):
- 
-    # Ensure correct format
-    if not isinstance(date, datetime.datetime):
-        raise TypeError('Invalid type for parameter "date" - expecting datetime')
-    elif date.year < 1801 or date.year > 2099:
-        raise ValueError('Datetime must be between year 1801 and 2099')
-
-    # Perform the calculation
-    julian_datetime = 367 * date.year - int((7 * (date.year + int((date.month + 9) / 12.0))) / 4.0) + int(
-        (275 * date.month) / 9.0) + date.day + 1721013.5 + (
-                          date.hour + date.minute / 60.0 + date.second / math.pow(60,
-                                                                                  2)) / 24.0 - 0.5 * math.copysign(
-        1, 100 * date.year + date.month - 190002.5) + 0.5
-
-    return julian_datetime
-
-def wordCount(text):
-    wordDict = {}
-    for sentence in text:
-        for entry in sentence:
-            keyTitle = entry[1]
-            if keyTitle not in wordDict:
-                wordDict[keyTitle] = 1
-            else:
-                count = wordDict[keyTitle]
-                count = count + 1
-                wordDict[keyTitle] = count
-    return wordDict
-
-def meanLengthSentence(text):
-    total_length = 0
-    for sent in text:
-        total_length = total_length+len(sent)
-    return total_length / len(text)
-
-def preprocess(text):
-    sentences = sent_tokenize(text)
-    sentences = [word_tokenize(sent) for sent in sentences]
-    sentences = [nltk.pos_tag(sent) for sent in sentences]
-    return sentences
-
-def percentage(count, total):
-    return 100*count / total
-
-def lexical_diversity(text):
-    return len(set(text)) / len(text)
-
-def strip_punctuation(text):
-    table = str.maketrans(dict.fromkeys(':,.?'))
-    s = text.translate(table)
-    return s
-
-def BrunetsIndex (TextLength, UniqueWords):
-    w = TextLength**(UniqueWords**-.165)
-    return w
-
-def plotter (xs, ys):
-    xs = np.asarray(xs)
[…]

### 2019-02-07 Source/Presidents Corpus Paper/paper1.tex
added paper1

--- 2019-01-31 Source/QualifyingReport/QualifyingReport.tex
@@
 % HEADER BEGINS
-\documentclass[10pt, letterpaper, twoside, openany]{book}
+\documentclass[10pt]{article}
 %% Packages
@@
 \fancyhf{}
-\fancyhead[CE,CO]{\leftmark}
-\fancyfoot[CE,CO]{Early Detection of MCI that progresses to dementia - An Interdisciplinary Approach}
-\fancyhead[LE,RO]{\thepage}
+\fancyhead[CE,CO]{Early Detection of MCI that progresses to dementia - An Interdisciplinary Approach}
+\fancyfoot[LE,RO]{\thepage}
 
@@
 \begin{document}
-\maketitle
-\newpage
-\tableofcontents
-\listoftables
-\listoffigures
-\newpage
-\chapter{Overview of Problem}
-Dementia has been identified as one of those fast growing difficulties facing the world.
+\section{Abstract}
+\section{Background}
+Dementia has been identified as one of the fastest growing difficulties facing the world today.
 A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
@@
 \par
-There is a large body of research that looks at the decline in language in those with MCI and AD \cite{Taler2008, Boschi2017}.
-However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD \cite{Taler2008, Boschi2017}.
+There is a large body of research that looks at the decline in language of those with MCI and AD \cite{Taler2008, Boschi2017}.
+However there is conflicting evidence in these studies about which declining language factors are associated with MCI and AD \cite{Taler2008, Boschi2017}.
 Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
-Another area of focus for research of this nature is the process of collecting appropriate language samples.
-Whilst collecting samples of language is comparatively unintrusive, researchers recognise that these samples require a rich sample of language that potentially cannot be generated by tasks such as the picture description task.
-Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD.
-There is some evidence to support this.
-Berisha et al \cite{Berisha2015}, has shown through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control.
+There is some evidence to support this, Berisha et al \cite{Berisha2015}, has shown through a longitudinal language analysis of spontaneous speech that there are marked differences in different parts of speech in this process between those who would go on to have a diagnosis of AD and a healthy control.
 \par
@@
 Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
-\par
-This report is structured in the following way.
-A literature review of current methods of cognitive assessment of MCI, an review of how language is said to be impacted in the MCI and Early AD population and a review of NLP and ML techniques that have the potential to be applied to this domain is presented in Chapter 2.
-The findings from a replication and extension of previous research using an existing dataset is presented in Chapter 3.
-I detail the aims of my research and my proposed future work plan including distinct tasks to be completed over the next two years in Chapter 4.
-Finally, my conclusions are summarised in Chapter 5.
 
-\chapter{Literature Review}
-\section{Introduction}
-Alzheimer's Disease (AD) is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
-Whilst a definitive diagnosis of Dementia can only be produced at post-mortem, there are a number of clinical indicators from psychological perspective that can indicate dementia is present.
-From a clinical psychology perspective, those who have dementia demonstrate cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}.
-In addition, these symptoms are often accompanied by emotional problems such as depression and behavioural difficulties.
-\par
-AD and other forms of dementia affect a significant proportion of the geriatric population in the world today and is currently the sixth leading cause of death in the US and was named the leading killer of women in the UK.
-According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
-This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population \cite{AlzheimersSociety2014}.
-From a financial perspective, they estimate an annual spend of £4.3 billion of which approximately £85 million is spent solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
-Globally, this picture is a lot bleaker.
-Another report by Alzheimer's Disease International suggests that in 2015 there were 46 million people with a diagnosis of dementia globally and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
-The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
-\par
-Despite this growing problem, at present there are no drugs that improve the prognosis of those suffering with AD.
-All the drugs that are on the market are designed to manage symptoms.
-Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{AlzheimersSociety2014}.
-Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place.
-It is therefore important to focus AD at it's earliest stages which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.
-\par
-Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent.
-Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time and alter cognitive function until a threshold is reached and clinical symptoms become more obvious \cite{Nestor2006}.
-The case of Iris Murdoch, who had a confirmed diagnosis of dementia, illustrates this theory well.
-Le et al \cite{Le2011} found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel when, it is theorized, the symptoms of AD manifested themselves more significantly.
-If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
-One of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of dementia.
-Albert and his team have worked to define clinical criteria which professionals can use to diagnose MCI due to AD and differentiate this from age-associated memory impairment and age-associated cognitive decline.
-They note that the diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}.
-\begin{enumerate}
-	\item Concern regarding a change in cognition - A person or an informant should express concern that there is a change in cognitive ability in comparison to previous level of performance.
-	\item Impairment in one or more cognitive domains - There should be evidence of lower performance in one or more cognitive domains beyond what would be expected of a person given their age and education.
-	\item Preservation of independence in functional - Whilst persons with MCI are expected to be able to maintain independence, it is common to experience mild problems in complex functional tasks which they may have been able to perform previously.
-This might mean that they take more time or be less efficient at completing these tasks, or it may be that may make more mistakes.
-	\item Not demented - The deterioration should be mild to the point that there is no significant loss of functioning in social or occupational contexts.
-\end{enumerate}
-In addition to meeting the above criteria, a clinician must rule out other conditions or factors that could account for the decline in cognition with the goal to increase the likelihood that the underlying cause of this decline is dementia.
-\par 
-Many researchers have studied the early detection of MCI.
-These studies usually follow two main approaches.
-The analysis of biomarkers and the examination of patients who have demonstrated decreasing cognitive abilities.
-The first approach yields reliable results in the detection of AD in its moderate and advanced states but does not perform well during the early stages of the disease.
-The second approach has gained more attention in recent years, due to the fact that in clinical practice it has shown promise in the early detection of AD.
-In addition, the analysis of the decline of cognitive abilities is comparatively inexpensive and less invasive that the first approach which commonly requires the collection of a sample of cerebro-spinal fluid which is painful for the patient involved.
-\par
-One of the most common ways in which clinicians traditionally make an early diagnosis of dementia is through the use of the Mini Mental State Examination (MMSE) \cite{Folstein1975}.
-The MMSE is a brief questionnaire consisting of eleven questions which tests cognitive aspects of mental function and requires only 5-10 minutes to administer \cite{Folstein1975}.
-The MMSE is chosen due to it's effectiveness at assessing a person's cognitive mental state at a specific point in time, as well as being as sensitive to changes as a more detailed and complex assessment such as the Wechler Adult Intelligence Scale \cite{Folstein1975}.
-Whilst the MMSE is useful as a brief screening tool it has it's limitations.
-The MMSE was not specifically created to screen for dementias and therefore does not interrogate key aspects of cognitive impairment known to be affected in dementia.
-It also has limited value in assessing under-educated subjects and a meta-analysis on the effectiveness of the MMSE as a diagnostic tool for dementia showed that it's accuracy was low (sensitivity between 78.4\% and 85.1\% and specificity between 81.3\% and 87.8\%).
-As the MMSE has been shown to have low accuracy specifically in the diagnosis of dementia, it becomes necessary for professionals to employ the use of other tools or measures such as the Free Cued Selective Reminding Test (FCSRT) \cite{Grober2010} or the Montreal Cognitive Assessment (MoCA) \cite{Davis2015}.
-\par 
-These tests have the benefits of being much more accurate at diagnosing cognitive impairment and discriminating between dementia and other types of cognitive impairment at the cost time and training of psychological professionals such as clinical psychologists in administering these tests.
-However, the utility of diagnosing dementia at the point where clinical intervention is warranted is limited because at this stage both psychological and pharmacological interventions have been shown to not be effective \cite{Prince2015}.
-In order to further our understanding of the progression of dementia it is important to detect the signs of dementia before they are clinically apparent.
-\par
-The two main ways in which diagnosis is performed is through assessment of memory and language.
-Tests of memory are classically among the most accurate ways of diagnosing dementia, however these tests suffer from the same reliance on clinicians to administer these tests in a clinical setting.
-Language however is a lot easier to collect and can be done in more naturalistic settings.
-As with memory, these tests can be done over time and would be able to chart a patients language degeneration over time.
-Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on measure decline in the use of language in those with AD.
-There are a number of difficulties to watch out for with this approach.
-There are a wide number of factors that are involved in language degeneration in the elderly, and consequently there will be an expected amount of variability between subjects.
-The administration of such tests may induce nervousness and discomfort which may impact performance, and also repeatedly administering the same language tests for differences over time may be confounded by improved performance at tasks via practice effects.
-However, there is enough promise in this approach such that it could help further our understanding of the disease, it's progression and the parts of the brain affected in the early stages.
-\par
-According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with mild dementia suffer from noticeable word finding difficulty.
-They may substitute general terms for more specific terms and may avoid the use of specific names of acquaintances.
-There may be grammatical errors involving subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc.
-Those who have progressed from Mild to Major depression also have difficulties with expressive or receptive language.
-They will often use general-use phrases such as "that thing" and "you know what I mean" and prefers general pronouns rather than names.
-With severe impairment, sufferers may not even recall names of closer friends and family.
-Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur.
-Echolalia (meaningless repetition of another person's spoken words) and automatic speech typically precede mutism.
-With the wide range of deficits someone with AD can suffer, it makes sense to try to categorise these deficits in some way.
-\par
-One of the most famous pieces of research on the topic of language decline in dementia was by Berisha and Liss who examined speeches and public interviews of former US president Ronald Reagan \cite{Berisha2015}.
-They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease.
-It turned out later to be the case that he had Alzheimer's Disease.
-Another classical study by Snowdon et al looked at whether linguistic ability in early life was associated with cognitive function and AD in later life \cite{Snowdon1996} .
-They found that idea density (defined as the number of expressed propositions divided by the number of words) was a key predictor in predicting whether nuns would go on to develop AD in later life.
-They found that those who would go on to develop AD all had low idea density in early life and they found no AD present in those with high idea density in early life.
-As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are extremely variable and can differ from patient to patient as the disease progresses.
-The consensus among researchers that this language degeneration is typically accelerated by the presence of dementia \cite{Berisha2015} and that a potential indicator of dementia is the rate of change in which the decline occurs relative to a fixed point in time rather than a comparison across a cohort of individuals.
-\par
-Emery \cite{Emery2000} completed a literature review looking at all the potential language deficits that could exist in those with AD and / or MCI.
-She divided these deficits into four levels of language: Phonology, Morphology, Syntax and Semantics.
-She proposed that language and the processes involved in language are hierarchical in nature and that language moves from simple units of construction (Phonology and Morphology), and build layers of complexity and sophistication (Syntax and Semantics).
-She found that people with AD generally had intact Phonology and Morphology but more impaired Syntax and Semantics.
-She asserted that the language forms we learn last are the first to deteriorate as we generally learn language in small simple units initially and build syntax and complexity as we are more comfortable with language.
-However it is important to note that different variants of dementia show different deficits in terms of language productions.
-Regardless, dividing language in this way is useful as it allows us to detect deteriorations in different parts of language usage and therefore may provide a way of discriminating between different forms of dementia.
-\par
-It is clear from both the clinical diagnostic criteria and supporting research that language is impacted in those with AD.
-However, whilst there is a move towards research aimed at looking more specifically at MCI we currently lack the measures that are sensitive enough to detect MCI.
-Given that we know language is affected before a clinical diagnosis of dementia is usually made \cite{Berisha2015, Snowdon1996, Le2011}, it makes sense to explore whether language on it's own can provide markers that may indicate a cognitive impairment that could progress to dementia.
-The field of machine learning and natural language processing has been suggested as a way to improve the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline \cite{Boschi2017}.
-% -- END OF INTRODUCTION --
-\subsection{Aims and Methodology}
-\par
-The purpose of this review is to look at the current standards of cognitive tests available that can be used as a diagnostic tool for MCI.
-I then go on to explore what techniques for assessing language have been used within the field of cognitive and clinical psychology.
-I then go on to look at what techniques have been developed in the field of machine learning (ML), deep learning (DL) and natural language processing (NLP) that might enable the automated analysis of language easier as well as an obstacles and/or limitations of current technology.
-Finally, I look at some studies which have already looked at the intersection of these two domains.
-\newline
-\par
-A search of the literature was conducted using ProQuest (PsychArticles), SCOPUS, Web of Science.
-The following results were found (Table 1).
-All papers were then reviewed for relevance by reading the abstract and full text where appropriate and a shortlist was compiled.
-An additional search through references of shortlisted papers was also conducted and any papers who upon further review appeared relevant were added to the shortlist.
-Papers were included where researchers used machine learning to classify participants as MCI, AD or Healthy using language.
-We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English.
-\begin{table}[H]
-	\begin{center}
-	\begin{tabular}{ | p{4cm} | p{1cm} | p{6cm} | p{1cm} |}
-		\hline
-		Database & Results & Search Terms  \\ \hline
-		ProQuest(PsychArticles) & 1484 & Language AND Decline AND Dementia \\ \hline
-		ProQuest(PsychArticles) & 486  & Language AND Decline AND Dementia AND Speech \\ \hline
-		ProQuest(PsychArticles) & 159 & Machine Learning AND Dementia AND Language \\ \hline
-		Web of Science & 1207  & Language AND Decline AND Dementia   \\ \hline
-		Web of Science & 151  & Language AND Decline AND Dementia AND Speech  \\ \hline
-		Web of Science & 34 & Machine Learning AND Dementia AND Language \\ \hline
-		Scopus & 791 & Language AND Decline AND Dementia  \\ \hline
-		Scopus & 91 & Language AND Decline AND Dementia AND Speech   \\ \hline
-		Scopus & 29 & Machine Learning AND Dementia AND Langauge \\ \hline
-		Scopus & 1292 & "Language Deficits" AND Dementia \\ \hline
-	\end{tabular}
-	\end{center}
-	\caption{\label{tab:table-name}Search Terms and Number of Results.}
-\end{table}
-% -- END OF Aims and Methodology
-\section{Existing Neuropsychological Measures of Cognitive Impairment, and Repeatable Battery of Neuropsychological Tests}
-In terms of standardised cognitive tests there are two main aims.
-Does the test distinguish accurately between normal aging, MCI and AD (diagnostic utility) and does the test distinguish between those individuals with MCI who will then go on to develop AD and those individuals with MCI who don't then go on to develop AD (prognostic utility).
-This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their performance in terms of both diagnostic and prognostic utility.
-\par 
-\subsection{Repeatable Battery for the Assessment of Neuropsychological Status}
-The Repeatable Battery for the Assessment of Neuropsychological Status (RBANS) was originally developed as an assessment tool for dementia, specifically looking at detecting a characterizing very mild dementia \cite{Randolph1998}.
-The authors felt that there was a shortfall of appropriate measures that were sensitive enough to milder impairments as well as a number of other shortcomings of existing tests.
-They met a number of design goals for this new battery of tests that addressed these shortcomings.
-The RBANS consists of a number of sub-tests across five distinct domains.
-
-\begin{enumerate}
-	\item Immediate Memory 
-	\begin{enumerate}
-		\item{List Learning}
-		\item{Story Memory}
-	\end{enumerate}
-	\item Visuospatial / Constructional
-	\begin{enumerate}
-		\item{Figure Copy}
-		\item{Line Orientation}
-	\end{enumerate}
-	\item Language
-	\begin{enumerate}
-		\item{Picture Naming}
-		\item{Semantic Fluency}
-	\end{enumerate}
-	\item Attention
-	\begin{enumerate}
-		\item{Digit Span}
-		\item{Digit Coding}
-	\end{enumerate}
-	\item Delayed Memory
-	\begin{enumerate}
-		\item{List Recall}
-		\item{List Recognitions}
-		\item{Story Memory}
-		\item{Figure Recall}
-	\end{enumerate}
-\end{enumerate}
-
-However, whilst the authors claim that the RBANS is adequately sensitive in person's with MCI \cite{Randolph1998} and there is some research to support this assertion \cite{Karantzoulis2013}, other research points out that the RBANS has poor sensitivity in detecting MCI \cite{Duff2010}.
-Another drawback of this battery is the lack of executive function measures and object naming tasks.
-Research has shown that the RBANS has good test-retest reliability and convergent validity.
-However, Duff et al warned that caution should be exercised when using the RBANS in a MCI population as it has lower sensitivity in this population \cite{Duff2010}.
-
-\subsection{Digit Span Test}
-The Digit Span Test (DST) is predominantly used to measure a person's working memory capacity, specifically the capacity used to store and recall numbers.
-A participant is presented with a series of numbers of fixed length and is asked to recall those numbers in normal or reverse order.
-The series of numbers gets progressively longer until such time as a participant fails three or more times out of eight presentations.
-\par
-Research has consistently shown that MCI patients have a significantly lower digit span score in both normal and reverse order versions of this test and a study by Muangpaisan has shown that the reverse order version of this test can, to some degree, predict a diagnosis of MCI \cite{Muangpaisan2008}.
-Muangpaisan also found that Age, Gender and Education have an impact on the performance of the tests.
-Emrani et al found that the DST revealed no differences between specifically amnestic MCI and controls, but could differentiate between Mixed MCI and the other groups with mixed MCI recalling fewer correct responses than other groups.
-Notably, there was an attenuated recency effect in those with mixed MCI.
-\cite{Emrani2018}.
-Kessels identifies that that there are working memory deficits in MCI patients and these worsen with AD patients.
-\cite{Kessels2011} He also identified that both MCI and AD have impaired performance on all three conditions of the digit span test.
-No differences were found between forward and backward conditions in any of the groups.
-However, available tests may not detect subtle impairments \cite{Kessels2015}.
-\subsection{Rey Auditory-Verbal learning test}
-Rey's Auditory Verbal Learning Test (RAVLT+) looks a wide range of neuropsychological processes including short-term auditory-verbal memory, learning and retention of information.
-Participants are given a list of 15 unrelated words, repeated over five different trials and are asked to repeat.
-Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then repeat it again after 30 minutes \cite{Schmidt1996}.
-\par
-Several studies have shown that an impairment in RAVLT score reflect well the underlying pathology caused by AD.
-Thus making the RAVLT an effective early marker to detect AD in persons with memory complaints.
-Moradi investigated to what extent the RAVLT scores are predictable based on MRI data using machine learning approaches, as well as to find out what the most important brain regions are for the estimation of RAVLT scores \cite{Moradi2017}.
-They found a highly significant cross validated correlation between the estimated and observed RAVLT immediate and RAVLT Percent Forgetting.
-Further, they found that the conversion of MCI subjects to AD in 3-years could be predicted based on either observed or estimated RAVLT scores with an accuracy comparible to MRI-based biomarkers \cite{Moradi2017}.
-Another study by Schoenberg found that RAVLT to best distinguish patients suspected of Alzheimer's disease from the psychiatric comparison group \cite{Schoenberg2006}.
-\subsection{Digit Symbol Substitution Test}
-The Digit Symbol Substitution Test (DSST) involves a key consisting of the numbers 1-9 and a corresponding unique symbol.
-Below this key is a series of numbers from 1-9 in a randomized order and repeated multiple times.
-The participant is asked to fill in the corresponding symbol for each number.
-The task requires that the participant move between the key and the randomized sequence such that  they may retrieve the correct answer from the key, hold this in short-term memory and transcribe the key in the appropriate place.
-\par 
-Among those with no disorder in cognition, mobility and mood, being in the lowest DSST quartile compared to the highest was associated with nearly twice the odds of developing one or more clinical or subclinical disorders.
-Associations were stronger for incident clinical disorders in cognition.
-Slower psychomotor speed may serve as a biomarker of risk of clinical disorders, mobility and mood.
-While in part attributable to vascular brain disease, other potentially modifiable contributors may be present \cite{Rosano2016}.
-Further studying the causes of psychomotor slowing with ageing might provide novel insights into age-related brain disorders.
-Pascoe compared patients with PD with Normal Cognition (PD-N) with those with PD and MCI (PD-MCI) and healthy participants.
-PD-MCI participants achieved significantly lower scores than other groups in the DSST task \cite{Pascoe2018}.
-\subsection{Verbal Fluency}
-There are a number of tests that characterise this category of verbal fluency, but generally fall into two categories.
-Letter (Phonemic) fluency involves the generation of as many words as possible which begin with a specified letter.
-Category fluency involves the generation of as many words as possible that fall into a specfied category.
-Both these tasks impose demands on a number of different cognitive processes namely, executive function, verbal retrieval and recall, giving appropriate answers while monitoring previous answers and inhibit inappropriate responses.
-However, these two tasks require different strategies when attempting them.
-Letter fluency relies on search strategies based on lexical representations whereas category fluency requires a search for semantic extensions of a superordinate term, meaning that semantic associations within the lexicon must be intact in order for the task to be carried out successfully.
-\par  
-There have been numerous studies which have documented the impact that AD has on verbal fluency tests in both categories.
-A review carried out by Henry et al \cite{Henry2004}, found that performance in both letter fluency and category fluency was impaired in those with AD vs controls, but found a larger effect for tests of semantic fluency.
-\subsection{Naming Tests}
-Word-finding difficulty is a common symptom of AD and these deficits usually occur during the early stages of the disease progression.
-As such , a test of a patients ability to find words (known as confrontation naming) is a common way to measure cognitive decline.
-One common way to do this is the Boston Naming Test (BNT) which comprises 60 items on a spectrum of very frequent to very infrequent.
-This has been reduced subsequently to two thirty item versions and four fifteen item versions, which correlate significantly with the original sixty item version and the benefit of the shorter versions of the test is that it facilitates testing of individuals with AD who may suffer from fatigue or limited attention span \cite{Williams1989}.
-\par 
-Willers et al \cite{Willers2008} studied Twenty aMCI patients, twenty AD and 21 controls matched by age, sex and education level.
-They found AD patients obtained significantly lower total scores on the BNT than aMCI patients and controls.
-aMCI patients did not obtain significant differences in total scores but showed significantly higher semantic errors compared to controls.
-Semantic processing is impaired during confrontation naming in aMCI.
-\par 
-Vadikolias investigated the impact of education on naming tests \cite{Vadikolias2012} and found that higher educational attainment in aMCI subjects were correlated with better performance in verbal and non-verbal tasks during repeated examinations over 1-year.
-Subjects with a lower level of education performed worse than patients with a high level of education who presented a more stable clinical score.
-The explanation for this is the idea of a 'cognitive reserve' in participants with a higher education such that this provides a buffer that, while not preventing the physiological symptoms of AD, can potentially delay the clinical onset of cognitive symptoms that characterise AD.
-This theory is supported by Snowdon who found a relationship between early life linguistic ability and the density of neurofibrillary tangles in his nun study \cite{Snowdon1996}.
-Whilst these studies and tests provide evidence that there are word finding difficulties in those with MCI, on their own they do not provide sufficient ability to diagnose MCI or provide an prognosis of disease progression.
-\subsection{Rey-Osterrieth Complex Figure Task}
-The Rey-Osterrieth Complex Figure (ROCF) is a task widely as a test of visuo-spatial skill and visual memory.
-The task, which was originally designed by Rey (1944) and standardised by Osterrieth (1944) \cite{Rey1941, Osterrieth1944}, asks a participant to copy a complex geometrical figure (known as the immediate copy condition and to recall and reproduce the figure from memory without warning (known as the delayed recall condition).
-In the immediate copy condition, the complexity of the figure requires an integrative cognitive ability.
-The reproduction of such a complex structure involves processes such as planning and organizational strategies that are related to executive functions.
-In order to make this task repeatable, Taylor et al designed a comparable set of figures which have proven to be of equivalent difficulty \cite{Taylor1969, Strauss1990}.
-\par
-Salvadori found that patients with vascular MCI had a worse performance in the immediate copy of the ROCF compared to individuals with degenerative MCI, despite their significant impairment in terms of general cognitive status and visual memory \cite{Salvadori2018}.
-Evidence shows that patients with disorders that possibly involve attention and executive functions are characterized by a more disorganised approach when copying the ROCF compared to controls.
-One of the difficulties with the presentation of this task in a repeated battery will be the fact that it turns from an incidental memory task (it is a memory task but the participant is not forewarned that they will be required to memorise the picture) into an intentional memory task (given the previous exposure to the task, it would be expected that a participant would preempt the delayed recall portion of the test and spend more time attempting to memorise the details) \cite{Teng2009}.
-\subsection{Hayling Sentence Completion Test}
-Inhibitory deficits are a common in all stages of dementia.
-This is usually tested using Stroop test, however this has the drawback of lacking ecological validity.
-Therefore researchers have moved towards using the Hayling Sentence Completion Test (HCST) which uses skills such as word retrieval as well as the ability to inhibit ones responses where appropriate and is therefore a much more ecologically valid task.
-There are two parts to the HCST.
-In the first part, participants have to complete a sentence by providing a word that best fits the given sentence (this is known as the initiation condition).
-In this second part, participants have to complete sentences by inhibiting an impulse to give the word that best first the sentence as in the first part, and producing a semantically unconnected word.
-Performance in both conditions is measured by the time taken for the participant to initiate a response, and in the inhibition condition also by the correctness of the word.
-\par 
-Martyr et al compared healthy controls with patients with dementia and patients with Parkinson's disease \cite{Martyr2017}.
-They found that a high proportion of Category A errors (producing a word that fits the sentence when instructed otherwise) was a factor in performance loss for participants with dementia.
-Findings suggest that the HSCT may be sensitive to verbal suppression deficits and may provide insight into inhibitory control in participants with dementia.
-Patients with Dementia were significantly slower than controls in the initiation and inhibition conditions vs healthy controls, and slower than patients with Parkinson's disease in the initiation but not the inhibition condition.
-\subsection{Grooved Pegboard Test}
-The Grooved Pegboard Test (GPT) was originally designed to cover a variety of different psycho-motor functions including hand-eye coordination and motor speed.
-However, some studies have shown a correlation of performance in the GPT and measures of cognitive performance such as the Montreal Cognitive Assessment.
-\par 
-Bezdicek found that the GPT predicted performance on the MoCA and concluded that in addition to being a measure of motor skills, there were results that showed that the GPT could also provide information about a participants cognitive skills \cite{Bezdicek2014}.
-Whilst this was specifically using a population of patients with Parkinson's disease, given the MoCA and GPT have shown degree of correlation, there remains an opportunity for research that explores the use of the GPT with MCI and Early AD populations.
-\par 
-Darweesh et al, used the Purdue Pegboard Test (PPT) to assess manual dexterity in a healthy older adult population and followed their participants for between eight and twelve years until an indication of the onset of a neurodegenerative disease \cite{Darweesh2017}.
-In this time 227 (4\%) of their participants went on to develop a diagnosis of dementia.
-They noted that higher PPT scores were associated with lower risk of incident neurodegnerative disease and noted significant associations of PPT scores with all forms of dementia and this potentially highlights a deterioration of motor function in the pre-clinical phase of dementia.
-\subsection{Visual Search}
-The visual search task requires participants to be watching a computerised display in which a target appears either by itself or surrounded by other elements which are used as a distraction.
-These other elements can be vary in similarity to the target.
-This is inherently a measure of attention shifting and processing of various similarity and as one would expect, the visual search time is increased in the presence of distractors.
-This effect is magnified in participants who have a cognitive impairment.
-\par
-Research has shown there to be deficits in both AD and MCI populations in this task \cite{Tales2005}, and those with MCI exhibited less severe deficits compabut that these deficits were not as apparent in MCI populations, there remained a significant enough difference that could be used to differentiate healthy controls from those with MCI.
-In a comparison of visual search performance between those with aMCI and healthy controls, Tales noted significantly poorer performance in the aMCI group.
-However, she also noted a good deal of heterogeneity in the aMCI group which illustrates that whilst the aMCI group have essentially the same condition, presentations within this population can differ markedly.
-The results from this study also illustrates the use of a non-memory task as a means of diagnosing dementia as patients who went on to be diagnosed with dementia whilst this study was in progress had significantly poorer visual search performance.
-\subsection{Free Cued Selective Reminding Test}
-The Free Cued Selective Reminding Test (FCSRT) was borne out of the premise that by controlling the conditions of learning, a measure of memory is possible that is not confounded by normal age related changes in cognition.
-Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cognitive decline due to a cognitive impairment.
-The FCSRT starts with a learning phase in which participants are asked to look at a card containing four pictures (e.g., grapes) for an item that belongs to a named category(e.g., fruit).
-After these four items are identified, immediate cued recall of these four items is tested and this is repeated for a total of 16 items.
-Following this learning phase is the recall phase in which participants are asked to recall all the items identified without cues and any items which are not recalled are then cued.
-There are three outcome measures for this test, free recall (the number of items recalled without cues), total recall (the number of items recalled with or without cues) and cue efficiency which is identified as follows \cite{Grober2010}.
-\begin{equation} \label{x1}
-CueEfficiency = (totalRecall-freeRecall)/48-freeRecall, range 0.0-1.0)
-\end{equation}
-Grober found that patients with impaired free recall with four times more likely to develop dementia than those with intact free recall \cite{Grober2010}.
-This test has also been shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the researchers to used the test to categorise prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT score \cite{Sarazin2007}.
-\subsection{Conclusions}
-I have looked at a number of different cognitive tests and a battery of tests that aim to have high diagnostic and prognostic capabilities in the MCI population.
-However, particularly with the RBANS, there lacks sufficient sensitivity in differentiating those with MCI from healthy elderly individuals such that this tool could be used with a level of confidence in a clinical setting.
-In regard to the tests, a common theme is a lack of studies and therefore evidence into the utility of these tests with an MCI population.
-However if, as researcher, we aim to investigate this population further then a benchmark battery of tests which is sensitive enough in both diagnostic and/or prognostic utility should be a goal.
-\par 
-Another criticism of the current literature is the lack of consistency with regard to the experimental groups.
-For example, some studies focus on differentiating between an MCI group, a AD group and a healthy controls group whereas other studies may further subdivide the MCI category according to a number of factors such as Amnesic or Non-Amnesic MCI, Mixed MCI (sometimes called Multi-domain MCI in the literature).
-Given the inconsistency in defining the experimental groups the confusing and often conflicting results that researchers produce is to be expected.
-Future research should look at the standardization of the operational definition of cognitive impairment in MCI may result in more consistent predictions of progression to AD.
-\par 
-Finally, it can be said that in testing for MCI and comparing these results with a cohort of those with a similar diagnosis has limited utility.
-This is because there is huge variability in the presentation of MCI within participants that even if controlled for with a matched pairs design for age, gender and education will produce inconsistent results.
-There is an argument therefore for the use of longitudinal studies, where the comparisons on the performance of these tasks are with the participants themselves.
-There are not many studies which use a longitudinal approach, although those that do show promising results.
-
-\section{Types of Language Assessment}
-One of the key debates when looking at how to analyse language is the type of task provided to elicit language production in participants.
-In the literature researchers have primarily focused on Picture Description tasks but have also suggested other ways in which we might collect data.
-\par
-\subsection{Picture Description Tasks}
-One of the most commonly used tasks to measure language is the Picture Description task.
-An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Borod1980}.
-The Cookie Theft picture (pictured below) depicts a scene of a home typical of the period of time when it was created and would generally not require participants to use any complicated vocabulary to describe.
-In this task participants are asked to describe the picture presented to them in as much detail as possible.
-This task was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}
-\begin{figure}[H]
-\centering
-\includegraphics[width=240px, height=150px]{images/BCTPicture.png}
-\caption{Cookie Theft Picture.\label{white}}
-\end{figure}
-\par
-The picture description task does a fine job of eliciting descriptive language however because of the specific content the language produce could be considered quite limited.
-There is some disagreement as to the benefits of this using this methodology.
-This task is reported as being useful to lexico-semantic disorders \cite{Boschi2017, Sajjadi2012} as the language being generated is primarly nouns and deixis (words to identify items and words to put those items into context).
-However, Ash \cite{Ash2012}felt that there was no difference in using this task vs Story Narration (described below).
-In explaining the differences, it is worth noting that these researchers were using differing variables and this could explain their different perspectives.
-\par
-\subsection{Narrative description task}
-The story narration task is designed to study a participant's ability to describe and elaborate on a story which is depicted using a series of pictures.
-The stories depicted are usually based on children's books or famous stories with the Cinderella being the one most typically used \cite{Fraser2014}.
-This task requires ordering the story in a structured and coherent framework.
-It also requires comprehension and understanding of the stories characters and the events depicted, as well as an awareness of a character's actions, motivations and internal reactions to given events.
-This task is particularly useful as the procedure reduces the demands on memory, due to the participant being able to access the picture book during the description and is therefore able rule out memory as a confounding variable for any results observed.
-As noted above, Ash \cite{Ash2012} felt that this task was interchangable with the Picture Description task.
-However, other research felt that this was a studier test of lexical and semantic abilities as well as syntactic complexity because this task requires interpretation and elaboration in additional to a simple description \cite{DeLira2011}.
-\par
-Given the relative strengths of the Narrative description task vs Picture description task, there are few pieces of research that have used Machine Learning to analyse features from Narrative picture tasks \cite{Fraser2014}.
-This could be due to the availability of data and the absence of any meaningful sets of transcripts of participants performing this task.
-However, this could be an interesting direction to take research in the future to see if features generated from this task could be used to predict MCI or AD.
-\par
-\subsection{Interviews}
-Interviews can also be used to elicit language in a more natural way by asking questions to guide a conversation between speakers.
-There are three types of interviews: unstructured, structured and semi-structured.
-Structured interviews tend to produce very limited speech and therefore has never been used in this area \cite{Boschi2017}.
-Unstructured interviews are open ended and generally do not conform to any particular pattern.
-They use generic themes such as family or hobbies to guide the conversation.
-Whilst this is the most ecologically valid form of conversation and therefore language generation, it's unstructured nature means that the protocol cannot be consistent and therefore reproduced.
-Semi-structured interviews are therefore preferred over other forms of interview as a middle ground.
-The semi structured nature of these interviews means that there is some replicability but does not constrain the participant in answering questions.
-\par
-The analysis of interviews can be difficult to analyse as both the content can vary even between participants, although it can be argued that content should not affect the type of language being generated unless it is narrow topic or the participant is constrained in how they answer a given question.
-It is also difficult to measure as there are no pre-defined task goals in comparison to the other two methods.
-Nevertheless, this is the most naturalistic setting for looking at language production and can be used to look at the syntactic and semantic parts of language generation \cite{Sajjadi2012}.
-There have been some attempts to use interviews to assess language production in AD with promising results \cite{Asgari2017, Guinn2015}.
-\par
-\subsection{Conclusions}
-One can view the different types of tasks above as a continuum where picture tasks represent a much more controllable task with a lot of supporting research but which generates a much more constrained set of language that is atypical of normal speech in terms of the cognitive functions used.
-
-\section{How do we analyse language, issues and debates}
-There are a wide variety of approaches that we can take in the analysis of language and there is no real consensus on the best approach to this particular problem.
-This section looks at the different ways in which researchers have looked at the problem as well as discussing some of the areas of contention.
-\subsection{Single Word Language tasks vs Connected Language tasks}
-Part of the reason we need to pay attention to how we ask participants to generate data is understanding how we wish to analyse the data afterwards.
-As discussed above, the different methodologies to collect data generate different types of language.
-There are two main approaches which we have looked at to analyse language, using frequencies of words and combinations of words and measures of syntax and semantics.
-There are other less common methods of analysing language but these are beyond the scope of this review.
-\par
-Single word tasks such as the Boston Naming Test and other such standardized language tests generally target a participants word production where this is defined as the ability to form and express words in accordance with certain criteria (see above for a full description of verbal fluency as task).
-There are a number of benefits to using single word tasks.
-From a research methodology perspective, using a standardized test allows researchers to target a very specific process in language generation and isolate factors that impact performance in language well.
-However, this approach does not take into account other cognitive processes that are used in lengthier speech tasks.
-More specifically, single word language tasks do not look at the interaction between language, executive functions and reasoning abilities.
-They also do not require the logical and efficient organisation of ideas.
-\par
-Connected language tasks, such as the picture description task and interviews described above are much more reflective of the processes involved in natural language generation.
-They are able to be relatively constrained in the language available to be used, for example in the picture description task, or they can be unconstrained in the form of interviews.
-But regardless of how they are framed, they are potentially much more useful in terms looking at all the processes involved in language generation with the drawback that you will not be able to isolate specific parts of the language generation process.
-A review by Boschi et al concludes that analysis of connected speech is potentially useful in guiding clinicians to identify language disorders \cite{Boschi2017} and also highlights the role of NLP and Machine learning in assisting in this endeavour.
-\par
-\subsection{Semantics vs Pragmatics}
-When navigating the English language it is necessary to distinguish between what a sentence says in both semantic and pragmatic terms.
-Semantic meaning refers to the meaning of the words in a sentence local only to the given sentence.
-Another way to put this is, semantics considers the meaning of words without taking into account the context in which these words are spoken.
-Pragmatic meaning refers looks at the same sentence in terms of words and grammar but takes into account the situation or context in which these words are spoken.
-Emery in her literature review looks at all levels of language tasks except for pragmatics, however this is an area that should not be ignored.
-Whilst the study of language from a pragmatic perspective is much more complex, it is perhaps one of the most vital areas to study because of the number of different cognitive processes involved.
-\par
-\subsection{Semantic Content}
-Another approach to linguistic analysis in this field is the idea of measuring semantic content and complexity.
-According to Emery (2000) \cite{Emery2000} in which she states that Semantic and Syntactic skills deteriorate first in people with MCI and AD.
-If this is true, then psychological measures of semantic and syntactic skills should be able to pick up signs of deterioration and act as markers for possible MCI and AD.
-An example of a semantic complexity measure is the concept of idea density.
-Formally, idea density is defined as the average number of propositions per sentence \cite{Keenan1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}.
-An example of semantic content measures is Type to Token Ratio (described below) which is used to measure the lexical diversity of a given piece of text and/or utterance.
-This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2013}.
-\par
-\subsubsection{Type token ratio(TTR)}
-Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
-\begin{equation} \label{x2}
-TTR = numberOfUniqueWords / totalNumberOfWords.
-\end{equation}
-\subsection{Picture-related content thematic elements}
-Several studies examined the amount of thematic elements expressed that were directly relevant to picture stimulus in picture description tasks.
-The studies used a variety of phrases to denote these thematic elements, these included 'pictorial themes', 'relevant observations and 'semantic units' but are notionally similar.
-The only difference was the number of thematic elements that 'scored' for the studies in question.
-Nicholas et al identified eight thematic elements of the Cookie Theft picture and used the number of elements as an outcome measure in different groups.
-He found that patients with AD expressed significantly fewer content elements than controls.
-\par
-Hier, Hagenlocker and Shindler assessed content using a similar list of thematic elements \cite{Hier1985}.
-They divided their participants into early-stage and late-stage AD, as well as including a control group.
-The late-stage AD group produced significantly fewer relevant observations than the early stage group, and the AD group combined produced fewer relevant observations than controls.
-This study was replicated by Lukatela et al \cite{Lukatela1998}.
-\par
-Smith, Chenery and Murdoch applied Hier's methodology for constructing pictorial 'themes' with the Picnic Scene from the Western Aphasia Battery (WAB) with a control and patients with moderate to moderately severe AD \cite{Smith1989}.
-The authors found no difference in the number of semantic elements produced but did not that the group with moderate to moderately severe AD took more time and more syllables to communicate these elements.
-\par
-\begin{figure}[H]
-\centering
-\includegraphics[width=240px, height=150px]{images/picnic-scene.jpg}
-\caption{Picnic Scene taken from the Western Aphasia Battery (WAB).\label{white}}
-\end{figure}
-Sajjadi et al examined 10 pictorial themes in picture description the Comprehensive Aphasia Test and found that the group with mild AD produced similar themes than controls \cite{Sajjadi2012}.
-Bschor et al.
-(2001) examined Cookie Theft picture descriptions at four stages of AD.
-They identified 12 elements of the cookie theft picture and found that whilst each AD group differed significantly from the others and from controls, the measures did not distinguish between MCI and normal controls \cite{Bschor2001}.
-\par
-Finally, a number of studies used composite measures which contained thematic elements and other unspecified information units resulting in a list of 23 possible information units of the Cookie Theft picture.
-The authors felt that this provided a wider, more liberal range of relevant content and thus subtler differences could be noted.
-Studies using these features found some differences between AD and controls, and some could differentiate between different stages of AD.
-\subsection{General Information Units or Content Information Units}
-Some studies used a more general concept of content, defining "information units" as "the smallest non redundant meaningful fact or inference," whether or not the information conveyed was specific to the context in which the conversation happened.
-Giles et al for example studied adults with minimal, mild or moderate AD vs controls and found that adults with AD produced fewer overall information units than controls \cite{Giles1996}.
-\subsection{Conciseness of information}
-Conciseness has been defined as the number of words a speaker uses to express ideas.
-The theory is that people with AD will need more words to convey ideas because of word-finding related behaviours such as circumlocutions and repetitions.
-Conciseness is calculated by dividing the number of ideas expressed by the total number of words in a measure commonly referred to as idea density but also known as lexical index, information content and information unit conciseness index.
-Snowdon examined written discourse from the Nun study and found that low idea density in early life was associated with reduced cognitive performance in later life \cite{Snowdon1996}.
-Riley et al extended these findings by concluding that early-life idea density was associated with lower brain weight, higher degree of cerebral atrophy and increased neurofibrillary pathology in later life \cite{Riley2005}.
-\par
-Ahmed, de Jager et al examined idea density with patients who had confirmed AD post mortem \cite{Ahmed2013}.
-They found that those with AD produced fewer total semantic units than controls but there was no significant difference between the groups with regards to idea density.
-The study of "empty speech" by Nicholas et al examined conciseness with measures thought to contribute to the "nonspecifity" of discourse in AD, such as empty phrases (defined as common idioms contributing no relevant information), deictic terms (e.g.
-"this", "that" without referents), indefinite terms (e.g.
-"thing" or "stuff"), pronouns without proper noun antecedants, and repetitions.
-In their study they found that AD patients produced more of these behaviours than did controls.
-\par
-\subsection{Efficiency}
-% CITE MURRAY!
-Efficiency is the rate at which meaningful information is conveyed over time, calculated by dividing the total number of information units by the duration in seconds of the speech sample.
-Smith et al, 1989 found that 18 adults with moderately severe AD produced fewer content units per minute on average than controls, and attributed this different to increased circumlocutions and repetitions \cite{Smith1989}.
-Murray used an analogous measure which he called "performance deviations per minute", in which fillers, irrelevant words, revisions or false stars, vague or non-specific vocabulary and inaccurate output (e.g.
-paraphasias) were divided by the total number of minutes in the speech sample; this measure was lower for those with AD than those with depression, and also healthy controls.
-The authors suggested that discourse information measures may help disentangle the similarities in symptoms of early AD versus depression in older adults.
-Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in dialogue in which a speaker provides responses do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls.
-\subsection{Quantity - Total number of words}
-Several studies report that adults with moderate AD produce fewer words than controls on picture description, however other studies found no differences in total words among groups of controls and patients with MCI or AD.
-Murray and Nicholas et al investigated normal controls, patients with AD and older adults with depression and found no group differences in total words \cite{Nicholas1985}.
-In contrast, Lira 2014 found that controls produced more total words than patients with AD but found no difference between mild and moderate groups \cite{Lira2014}.
-\par
-\section{Syntax and Morphology (Language Form)}
-Syntax can be defined as the rules that govern how words can be combined to form sentences, whilst Morphology is the system that governs the structure of words and the construction of word forms.
-Multiple studies of language decline in dementia included at least one measure of syntax or syntactic complexity \cite{Orimaye2017}.
-Common constructs included words per clause, grammatical form (measures of an appropriate use of syntactic conjections, tenses, conditionals, subordinate clauses and passive constructions), measures of phrase length and proportions of words in sentences.
-Some researchers have explored the use of formulaic language in those with dementia, the theory being that well practiced phrases are less effortful and therefore place low load on the cognitive abilities of those with AD.
-The general hypothesis motivating these studies is that either working memory limitations or semantic memory limitations in AD affect one's ability to use complex constructions.
-\par
-\subsubsection{N-grams}
-One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
-An n-gram is a contiguous sequence of n items from a given sample of text or speech.
-The items can be phonemes, syllables, letters, words or base pairs according to the application.
-For example, given the sequence of words "to be or not to be", this extract is said to contain six 1-gram sequences (to, be, or, not, to, be), five 2-gram sequences (to be, be or, or not, not to, to be), four 3-gram sequences(to be or, be or not, or not to, not to be) and so on.
-This is useful as, given a large portion of text or speech, we can predict the probability of a word being close by to a given word.
-A number of researchers have used n-grams as features.
-\par
-Asgari, Kaye and Dodge (2017) \cite{Asgari2017} used another form of word frequency measurement.
-Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words \cite{Pennebaker2015}.
-They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
-\par
-\subsection{Mean length of utterance (MLU)}
-Murray found that MLU was not a distinguishing factor among health adults, adults with depression and adults with AD.
-Ripich et al found a decrease in MLU in adults with severe AD over time, and this was supported by findings of Le et al in their studies of authors \cite{Ripich1988, Le2011}
-\par
-\subsection{Proportion of verbs to nouns plus verbs}
-Kave and Levy used a verb index to capture syntactic complexity and found that adults with AD expressed the same amount of verbs to nouns plus verbs as adult controls \cite{Kave2018}.
-\par
-\subsection{Syntactic Complexity - Composite measures of MLU, syntactic errors and verbs}
-A number of studies which explored syntactic complexity found differences between adults with MCI and controls, and between MCI and moderate AD stages \cite{Rentoumi2014, Ahmed2013}.
-The differences in syntactic complexity were not significant when individual measures were tested, but were apparent using a composite score consisting of MLU, words in sentences, syntactic errors, nouns with determiners, and verbs with inflections.
-\par
-\subsection{Formulaic Language}
-Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
-They found that there were four factors which they identified as important in the classification of participants as either healthy or AD.
-These four factors were semantic impairment, acoustic abnormality, syntactic impairment and information impairment and were based on existing measures of semantic and syntactic complexity.
-Zimmerer (2016) \cite{Zimmerer2016} looked at whether language was more formulaic in those suffering from AD.
-He proposed that those who suffer from AD rely on formulatic sentences, for example 'Noun-Verb-Noun', and this is done to reduce language complexity.
-He noticed a significant difference in the use of formulaic sentences between AD and Healthy Controls.
-\par
-\section{Pragmatic Language}
-The pragmatic language domain refers to the social rules for language for the purposes of communication including, using language to achieve goals, using information from the context to achieve these goals and using the interaction between people to initiate, maintain and terminate conversations.
-\subsection{Questions, turn-taking, unsure statements, egocentric comments}
-Ripich et al examined several pragmatic abilities among patients at different stages of AD \cite{Ripich1988}.
-The severe AD group asked more questions over time than the other group which infers problems with language processing.
-The authors argued that question-asking was a compensatory strategy and as a result, the proposed that adults in late-stage AD may have had insight into their communication problems.
-\par 
-There were a number of flaws of the methodology if their research.
-Firstly, a caregiver was asked to administer the picture description task in order to mirror a more typical communicative interaction.
-Whilst this improved the ecological validity of the study, the caregivers delivery would be varied and inconsistent.
-Secondly, due to the constrained nature of the picture description task, it is unlikely that it would be able to capture the pragmatic skills that are typical of conversations in everyday life.
-\subsection{Coherence}
-Coherence refers to the appropriate maintenance of topic in discourse and is thematically related to the immediately preceding utterance (local coherence) and by how closely an utterance relates to the general topic at hand (global coherence).
-One study looked at coherence, Chapman et al used picture descriptions of Norman Rockwell prints within a frame analysis, with frames being defined as internalized knowledge structures \cite{Chapman1998}.
-The authors identified aspects of content, including typical frames of interpretation, atypical, incorrect or no frames, propositions supporting frames and propositions disrupting frames as measures of coherence.
-They examined these variables with early stage AD, old-elderly and normal controls.
-Old-elderly and normal controls produced significantly more typical frames and more frame supporting information that the AD group.
-The authors attributed AD patients' difficulties to memory deficits, attentional deficits, visual perceptual problems, disruption of internalized frame representation, or failure to access frame knowledge.
-\par
-\subsection{Perseveration}
-Perseveration can be broadly defined as the repetition of a response regardless of the absence or cessation of stimulus which could have generated an appropriate response in the first place.
-A typical example could be the idea of conversation moving from introductions to a more general conversation, someone who has difficulties with perseveration will struggle with the move between social contexts and repeat language that would be more used in an introduction context.
-\par 
-One study examined verbal preservation in the description of Norman Rockwell prints, dividing the total number of words within perseverations by total number of words in the speech sample.
-The authors also calculated rate of perseveration on two other language tasks: confrontation naming and generative naming.
-Across all tasks, the AD group produced significantly more perseverations than controls; however, on the picture description task alone, there were no significant differences between the two groups.
-The authors theorised that this was because picture description was an easier task because there was the visual stimulus, similar to the argument made by Bschor et al \cite{Bschor2001}.
-\par
-\subsection{Empty Speech}
-Verbal fluency is a term used in neuropsychological contexts generally referring to timed, word-generation tasks, while in speech-language pathology contexts, "fluency disorders" are defined as interruptions in the flow of speaking characterized by atypical rate, rhythm and repetitions in sounds, syllables, words and phrases.
-"Fluency", in the literature of discourse of adults with AD, typically refers to the smoothness or flow of spoken language.
-Abnormalities of fluency in this population are typically characterised by filled and unfilled pauses, word repetitions, circumlocutions, and revisions.
-\par
-The study of "empty speech" by Nicholas et al was one of the first to examine aspects of fluency in the connected speech of persons with AD \cite{Nicholas1985}.
-They found that adults with AD had significantly more repetitions than controls.
-Similarly, Bayles and Tomoeda found more aborted phrases, revisions and ideational repetitions in the AD group than in controls \cite{Bayles1983}.
-Several other studies support the idea that in AD population there are a greater number of repetitions and revisions than in healthy controls.
-However, there are some studies which contradict these findings\cite{Ahmed2013}.
-\subsection{Conclusions}
-This section has looked at a number of different attributes which researchers have cited as being important in the detection of MCI and AD.
-Whilst in some cases there is a broad consensus about a given feature in the vast majority of cases there is some uncertainty.
-Some of the difficulty lies in different approaches used particularly around the criteria for experimental groups which means that it is difficult to be able to compare studies, like for like, as the populations of the participants being studied vary in subtle or obvious ways.
-It is important for any studies moving forward to use comparable standards when it comes to their experimental groups, such that the picture can be made clearer.
-
-\section{State of literature into Machine Learning and Natural Language processing techniques}
-Diagnosing dementia through language analysis has a large background in terms of psychological research.
-Increasingly current research has called for the use of machine learning as a way of assisting in the process of diagnosis \cite{Orimaye2017, Boschi2017}.
-This section will look at what areas of natural language processing (NLP) and machine learning (ML) could potentially be used as tools to help in this domain as well as any research that has applied machine learning to this problem.
-\subsection{Natural Language Processing}
-Natural Language Processing as a topic can be defined as the intersection between Machine Learning and Linguistics and looks at a number of language tasks, particularly important in this domain is how to enable a computer to process and analyse large amounts of language data.
-Some other tasks include speech recognition, natural language understanding and natural language generation.
-Generally speaking, Natural Language Processing (NLP) systems mimic the semiotic perspective on language.
-That is to say that we can subdivide NLP processes into different components (see Figure 2.3).
-\begin{enumerate}
-	\item Morphological/Lexical: providing the basic language elements or vocabulary, such as words, their roots and inflections.
-	\item Syntactic: for grouping and sequencing elements within samples of language (usually sentences).
-	\item Semantic: for knowing the meaning of an utterance, usually defined in terms of the 'truth value' of the logical propositions that are thought to be expressed by sentences.
-	\item Pragmatic: for understanding the context and purpose of an utterance.
-\end{enumerate}
-\begin{figure}[H]
-\centering
-\includegraphics[width=240px, height=150px]{images/workflow.jpg}
-\caption{How Natural Language Processing tasks are subdivided.\label{white}}
-\end{figure}
-In order to facilitate these and other tasks, a number of frameworks have been developed that automate some of the more common processing tasks. A brief review of these frameworks follows. 
-\subsubsection{Natural Language Toolkit}
-The Natural Language Toolkit (NLTK) was originally designed as part of a computational linguistics course at the University of Pennsylvania but has evolved to become an open-source framework which includes methods and modules for a wide array of Natural Language tasks (see Table 2.2 for a full list) \cite{Bird2009}.
-\begin{table}
-	\begin{tabular}{ | p{4cm} | p{7cm} | p{1cm} |}
-		\hline
-		\textbf{Language processing task} & \textbf{Functionality} \\ \hline
-		Accessing corpora & standardized interfaces to corpora and lexicons \\ \hline
-		String processing & tokenizers, sentence tokenizers, stemmers \\ \hline
-		Collocation discovery & t-test, chi-squared, point-wise mutual information \\ \hline
-		Part-of-Speech tagging & n-gram, backoff, Brill, HMM, TnT \\ \hline
-		Machine Learning & decision tree, maximum entropy, naive Bayes, EM, k-means \\ \hline
-		Chunking & regular expression, n-gram, named-entity \\ \hline
-		Parsing & chart, feature-based, unification, probabilistic, dependency \\ \hline
-		Semantic interpretation & lambda calculus, first-order logic, model checking \\ \hline
-		Evaluation metrics & precision, recall, agreement coefficients \\ \hline
-		Probability and estimation & frequency distributions, smoothed probability distributions \\ \hline
-		Applications & Graphical concordance, parsers, WordNet browser, chatbots \\ \hline
-		Linguistic Framework & manipulate data in SIL Toolbox format \\ \hline
-	\end{tabular}
-		\caption{\label{tab:table-name}NLTK tasks and functionality}
-\end{table}
-The NLTK has been successfully used in both teaching and research purposes and as part of this provides access to large array of databases and corpora which have potential uses in common tasks.
-
-\subsubsection{CoreNLP}
-Stanford CoreNLP provides a set of human language technology tools.
-It can give the base forms of words, their parts of speech, whether they are names of companies, people, etc., normalize dates, times, and numeric quantities, mark up the structure of sentences in terms of phrases and syntactic dependencies, indicate which noun phrases refer to the same entities, indicate sentiment, extract particular or open-class relations between entity mentions, get the quotes people said, etc.
-\par
-\begin{figure}[H]
-\centering
-\includegraphics[width=350px, height=150px]{images/CoreNLP.jpg}
-\caption{How Natural Language Processing tasks are subdivided.\label{white}}
-\end{figure}
-Stanford CoreNLP’s goal is to make it very easy to apply a bunch of linguistic analysis tools to a piece of text.
-A tool pipeline can be run on a piece of plain text with just two lines of code.
-CoreNLP is designed to be highly flexible and extensible.
-With a single option you can change which tools should be enabled and disabled.
-Stanford CoreNLP integrates many of Stanford’s NLP tools, including the part-of-speech (POS) tagger, the named entity recognizer (NER), the parser, the coreference resolution system, sentiment analysis, bootstrapped pattern learning, and the open information extraction tools.
-Moreover, an annotator pipeline can include additional custom or third-party annotators.
-CoreNLP’s analyses provide the foundational building blocks for higher-level and domain-specific text understanding applications.
-
-\subsubsection{Linguistic Inquiry and Word Count}
-The premise behind the Linguistic Inquiry and Word Count (LIWC) is the words we speak inform an observer of the state of mind of an individual.
-To this end, the creators have constructed a framework which uses a dictionary of words which are pre-categorised \cite{Pennebaker2015}.
-A typical example of this might be someone with depression who may typically be more inwardly focused and therefore talking in the first person is common\cite{Al-Mosaiwi2018}.
-The 2015 version of the LIWC has over 90 different categories and some supporting research suggests that it is effective at identifying themes and associating these themes with mental health difficulties such as depression and anxiety \cite{Sonnenschein2018}.
-\par
-A framework like this could be used as another tool to explore themes within a persons speech.
-
-\subsection{Traditional methods of Machine Learning}
-
-In terms of Machine Learning research in this area, a number of researchers have used transcripts based on picture description tasks \cite{Zimmerer2016, Orimaye2017, Mueller2018a, Fraser2015} and have successfully extracted linguistic features that could differentiate between AD and controls.
-\par
-Thomas et al, detail several lexical approaches to the problem of detecting and rating AD.
-The approaches rely on character n-gram techniques but also explore correlation of usage of frequency of different types of speech.
-Their results act as a proof in concept of the utility of using a pure computational approach to the diagnosis of dementia using spontaneous speech.
-Results - 95\% accuracy dementia vs controls.
-70\% accuracy classifying dementia into two categories.
-50\% accuracy classifying dementia into four categories.
-They suggest further exploration of characteristics to classify \cite{Thomas2005}.
-\par 
-AD patients had higher mean P rate, A rate and V rate scores but lower N rate scores vs normal older controls \cite{Bucks2000}.
-One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class.
-Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls.
-Their main finding supported n-grams as the most significant predictor.
-One of the criticisms is the use of picture description tasks and n-grams.
-Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
-\subsection{The case for Deep Learning}
-% Deep Learning is an NOT
-
-One of the main benefits of deep learning is that it does not rely on pre defined features being fed to the model, instead generating it's own features.
-This move away from relying on features solves one of the difficulties with
-
-% One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
-
-\subsection{Conclusions}
-We can see that both Natural Language Processing and Machine Learning techniques have a lot to offer, Indeed there has been a lot of research which have used pre-existing datasets to explore this area with promising results.
-To date, there has not been much in the way of research in which these techniques are applied to newly created samples of language.
-
-\section{Discussion}
-Current state-of-the-art diagnostic measures of AD are invasive (CSF analysis), expensive (neuroimaging), and time-consuming (neuropsychological assessment).
-Furthermore, these measures are limited to speciality clinics and thus have limited accessibility as frontline screening and diagnostic tools for AD.
-More importantly, nonspecialists are often inaccurate at identifying early AD and MCI.
-Thus, there is an increasing need for additional noninvasive and/or cost-effective tools, allowing effective frontline identification of subjects in the preclinical or early clinical stages of AD who could be suitable for monitoring in speciality clinics and for early treatment.
-Implementation of effective screening instruments will allow diagnosis earlier in the course of dementia, even at the point when memory function is still essentially within the normal range.
-This strategy would enable an earlier, and potentially more effective, prevention and treatment of AD with a special focus to preserve cognitive functions.
-\par 
-However the literature has identified a number of challenges when approaching this problem.
-Firstly, the clinical features that combine to meet the diagnostic criteria for Dementia or it's variants are continuous in nature and heterogenous between patients and are also impacted by other variables.
-For example as shown above cognitive performance is affected in part by a patients educational attainment and a patients ability to live independently is impacted by a their physical health as much as their cognitive health.
-The challenge therefore is to find features that are minimally impacted by other factors, or that can be controlled for by a experimental design such as a matched pairs design to control for educational attainment.
-\par 
-Another challenge lies with the recruitment of suitable individuals who may notice a decline in cognition to the point where we might classify them as having MCI, but these individuals deduce that there is little to no value to admitting there is a problem and seeking help whilst their symptoms are 'manageable'.
-\par
-Finally there is a large amount of variability in the presentations of those with MCI and early dementia, and this is compounded by an similar amount of variability in the criteria researchers have used for experimental groups and the approaches researchers have used to tackle this problem.
-This had led to a confused literature.
-Recently a call for research that has consistent inclusion / exclusion criteria has been made along with some proposed definitions of MCI and it's subgroups \cite{Petersen2014}.
-Researchers have identified the analysis of language impairment as an area of promise to explore in the diagnosis of MCI and early AD and recent developments in natural language processing and machine learning techniques have the potential to assist in this research.
-Indeed, given the increased burden on the diagnosis of MCI and AD on professionals there has been a call to use technology to potentially ease this burden \cite{Boschi2017}.
-A small but growing amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
-
-\subsection{Future Work}
-One of the areas for research to study is a careful examination of the features that are being used to measure language deterioration.
-For example, Zimmerer (2016) \cite{Zimmerer2016} describes connectivity in such as a way that correlates directly with what Mueller (2018) \cite{Mueller2018a} calls Fluency.
-Whilst these are very nuanced measures which differ slightly in the form they take, the sheer range of measures and features being produced make it difficult to conceptualise as part of the larger problem.
-Some work needs to be done in producing a consistent list of measures that are validated using existing datasets and can be used for future research moving forward.
-\newline
-\par
-Teng et al suggests that work should focus on the MCI population and concentrate on developing a consensus neuropsychological battery that could yield predictable rates of progression to AD \cite{Teng2009}.
-This, in conjunction with the development of a model of language, sensitive enough to detect subtle deterioration in language use to act as an additional cognitive marker to aid diagnosis could potentially move some way to providing this.
-\par 
-Future research should also be directed towards developing non-intrusive ways of detecting subtle changes in natural language such that any perceived deterioration that could indicate the presence of MCI or AD could be flagged up early.
-Machine learning approaches seem to be the most logical approach for achieving this aim as language could be collected in non-intrusive ways and passed to a machine learning algorithm for preliminary classification.
-Further, despite the excellent quality of datasets being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which we could claim are limited in scope.
-Finally, the recent resurgence in the use of neural networks and deep learning could provide the answer to the confused literature in terms of features.
-A key benefit of deep learning is it's ability to automate the process of feature engineering.
-So there is an opportunity to explore the use of deep learning, to not only develop new features but also
-\par
-This area of research is extremely promising in its early results and the impact of successful research would be life changing for both individuals and the health of the worlds aging population in general.
-
-\chapter{Initial Study of the Presidents corpus}
-\section{Background}
 There has been a significant research in the area of language deterioration as a means of detecting Alzheimer's Disease.
@@
 This experiment will look at the features originally identified by Berisha and Liss, as well as any others that have potential as discussed in the literature review above.
-\section{Methods}
+\section{Methodology}
 I took 46 transcripts of Ronald Reagan’s (RR) press conferences from 1981 to 1988 and compared them with 134 press conferences by George H.
@@
 		LI Verbs & \textit{"be", "come", "do", "get", "give", go", "know", "look", "make", "see", "tell", "think", "want"} \\ \hline
-
 	\end{tabular}
@@
 Whilst in terms of age, DJT is certainly more suitable as a control to match with RR, in some ways they held very different styles of press conferences in that RR preferred to do solo press conferences and DJT has shown a preference for doing joint press conferences which have an impact on the amount of language produced.
-This finding is, in itself notable as it illustrates the problems with small amounts of speech.
-\par
-The problem of finding an appropriate control is a common one in this domain.
+This artifact of the data is in itself notable as it illustrates the problems we may have with smaller amounts of speech.
+Also, the problem of finding an appropriate control is a common one in this domain.
 Given that those with MCI and early dementia have such variable presentations, it might prove of limited value in matched pairs design.
 \par 
-With further work, we will not have the luxury of having multiple samples and so it would be worth exploring how the quality of these predictions might lessen when faced with considerably fewer samples and over a smaller time period.
-In addition, this replication and extension has demonstrated the potential utility of using longitudinal data as a means of comparing language use of a person at two or more time periods and using this information as a diagnostic aid for MCI.
+With further work, it is not feasible to the vast array of samples over a timeframe, as we have had with the president corpus and so it would be worth exploring how the quality of these predictions might lessen when faced with considerably fewer samples and over a smaller time period.
+It would also be worth extending this research further to encompass more of the linguistic features Fraser used in her work \cite{Fraser2015} to see if there are any further insights to be gained.
+In addition, this replication and extension has demonstrated the potential utility of using longitudinal data as a means of comparing language use of a person at two or more time periods and using this information as a diagnostic aid for MCI and therefore more work would be helpful from a longitudinal perspective to see if this approach may be valid in moving towards a solution for this particular problem.
 \par 
-The results of this work show that we can track a person's use of language through time in a number of ways, and th
-
-\chapter{Future Work}
-\section{Introduction}
-This section looks at my research aims and objectives given the literature review presented above.
-This section will then go on to explore any challenges to my research.
-I then present the work I have planned for the next two years, divided into distinct tasks with their own aims and objectives.
-As part of this I will identify target conferences where I could present my work.
-Finally, I look at potential risks and ethical concerns that could conceivably occur.
-\section{Research Aims and Objectives}
-The aim of my research is to ultimately to identify more accurate and efficient ways of detecting MCI and early dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRIs and EEGs) by utilising spontaneous speech.
-There is a lot of research into the analysis of language as a marker for MCI \cite{Taler2008, Boschi2017} and there have been calls for research that explores the use of Natural Language Processing and Machine Learning techniques to aid in the analysis of language \cite{Boschi2017}.
-This will be the focus of my research.
-\par
-I have deliberately selected the MCI population as examining language decline at this stage has the potential to further our understanding of how MCI progresses into dementia.
-Given that sampling a person's language is relatively effortless, my proposed research will look at whether we can find markers of MCI in natural language.
-I plan to explore the use of Natural Language Processing and Machine Learning techniques to assist in this process.
-There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia.
-As documented above, these studies have identified a number of speech and language differences between healthy controls and those with dementia.
-Some of these features have been incorporated in work using machine learning to classify healthy controls from those with dementia.
-However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'.
-While this is helpful to establish that machine learning can be used to differentiate between healthy controls and those with dementia, there has been little work looking at those in the MCI phase although it is apparent that language is still affected \cite{Berisha2015, Snowdon1996, Le2011}.
-Therefore we would like to clarify whether natural language generated outside the framework of a deliberate language generating task, has the same ability to discriminate.
-\par
-Machine Learning itself is undergoing a shift in focus at the present time.
-I will describe this shift using the context of my proposed research.
-In traditional machine learning, the transcripts from cognitive assessments are used to generate abstract features such as idea density and more concrete features such as mean length of sentence or number of unique words.
-These features are then used to classify patients.
-As we have seen above, the picture with regards to what features deem important in this area is somewhat muddy.
-Machine Learning has a number of tools at it's disposal to aid this.
-The first is to statistically determine, given a model, which features contribute most highly to a given prediction.
-The second is to use a deep learning model which has the ability to automate the process of feature engineering.
-\par 
-The broad objectives of my research are to explore whether the application of Natural Language Processing and Machine Learning techniques to spontaneous speech samples can improve our ability to diagnose Mild Cognitive Impairment in patients that will directly lead to Alzheimer's Disease.
-Another objective of my research is to provide some clarity as to which language features are most important to measure when looking for evidence of MCI and/or AD.
-Finally, an additional outcome of my research is the development of a repeatable battery of neuropsychological tests that specifically focus on being sensitive to Mild Cognitive Impairment.
-\begin{itemize}
-	\item \textbf{H1} Monitoring language samples and other measures of cognitive decline over the course of the year will be enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment
-	\item \textbf{H2} Those with MCI will show a different language profile to controls if matched similarly for age, gender, education and other significant factors.
-	\item \textbf{H3} The new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures such as the RBANS.
-	\item \textbf{H4} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, those with Alzheimers Disease and decline in those with MCI.
-	\item \textbf{H5} Using individuals as their own control will be enough to identify trends in the decline of their use of language.
-\end{itemize}
-
-\section{Project Plan and Task Definitions}
-In this section, I describe each task in detail along with a projected timeline depicted in a Gantt chart.
-\subsection{Task 2 - Pilot Study for the Neuropsychological Battery of Tests}
-The aim of this study is complete a trial run of the administration of a new battery of neuropsychological battery of tests.
-We aim to find out whether the tests in question generate enough language so that features such as those generated in Task 1 are meaningful and can be used as language markers that can then be tracked over time.
-We also aim to look at whether there are significant practice effects between three versions of these tests with a gap of seven days between each test.
-\par 
-We compiled a series of tests which look at a number of cognitive functions which have previously demonstrated a level of effectiveness in discriminating an MCI population from a healthy population (see literature review for a comprehensive list, explanation and rationale for each test).
-In addition to these tests, we have incorporated two language tasks in which a participant's ability to generate spontaneous speech that we can use to further analyse.
-As a planned future study is longitudinal, it is important that this battery of tests is repeatable but does not have significant practice effects that may interfere with our ability to track cognitive decline.
-\par 
-We have recruited a cohort of 4 volunteers who will complete the battery of neuropsychological tests on three separate days with a gap of at least seven days between each administration.
-The results of which will compiled and analysed.
-Our hypothesis is that there would be no difference in performance between one battery of tests and another.
-An additional hypothesis is that the language tasks included on each test would be enough to collect an appropriate sample for linguistic analysis such that we may be able to detect cognitive decline in a clinical population.
-\par 
-While the pilot study in itself is not valid for submission to a conference or journal, it is hoped that the results of the study will allow subsequent studies using the same or adjusted battery of tests to be published.
-\subsection{Task 3 - Exploration of language with the DementiaBank dataset and Three Authors dataset.}
-The aim of this task is to replicate and extend the work of Fraser and Orimaye who both used the DementiaBank dataset and set out to use traditional machine learning techniques to see whether they could categorize people into dementia or healthy categories using the samples given.
-We also aim to replicate and extend the work of Le et al who used electronic versions of the novels of three authors.
-Both datasets are longitudinal and have been shown to track language decline over time in different ways.
-\par 
-The DementiaBank dataset was a product of a longitudinal study which was run by the University of Pittsburgh School of Medicine.
-The dataset contains transcripts of interviews conducted with participants with MCI / AD and other forms of dementia.
-These interviews were conducted in English and used the cookie-theft picture language task (see Literature Review for a full description).
-The full data set consists of 314 patients classified as having dementia and 242 patients who are classified as healthy and these patients are tested on a minimum of two separate occasions.
-\par 
-The Three Authors dataset includes the novels of P.D.
-James, Agatha Christie and Iris Murdoch.
-It contains twenty of Murdoch's novels which were published between the ages of 35 and 76, sixteen of Christie's novels written between the ages of 28 and 82 and fifteen of P.D.
-James novels written between the ages of 42 and 82.
-These three authors are notable as Murdoch had a confirmed diagnosis of Dementia after her death, Christie had a suspected diagnosis of Dementia and P.D.
-James was used as a control.
-\par 
-These two existing datasets in addition to the President's Press Conference corpus described above will provide three sources of validation of features that we generate as well as give us clues as how to best process the data we receive from our longitudinal study.
-\par 
-The final output of this task will be a paper which looks at the importance of various word-level and sentence-level features in the deterioration of language for people with AD.
-The target conference for this work is the UK Alzheimer's Research conference in March 2020 with an expected Abstract submission date of December 2019 and the Student Research Workshop of the ACL in July 2020 with an expected submission date of April 2020.
-\subsection{Task 4 - Longditunal Study}
-Previously we have looked at historical datasets, the aim of this task is see whether a formalised collection of specific data, namely our language tasks are enough to be able to be able to detect cognitive decline in participants at different stages of the MCI / Alzheimer's disease progression.
-To this end, we plan on administering our neuropsychological battery of tests from Task 2 to each participant three times (six monthly intervals) over the course of the year.
-\par 
-We therefore plan to recruit a set of 80 participants divided evenly into four groups.
-These groups are Healthy Older Controls (HOC), Early Alzheimers (EA), Subjective Mild Cognitive Impairment (SMCI) and Diagnosed Mild Cognitive Impairment (DMCI).
-We expect to see typical age related decline in our HOC group and more dramatic decline in our EA group.
-We are most interested to see if there is any measurable language decline in the SMCI or DMCI experimental groups.
-\par 
-Whilst we have divided participants into experimental groups we aim to investigate each participant individually.
-Because there is a large amount of heterogeneity in this population, we will initially be analysing each participant's results in isolation with the participant acting as their own control.
-However, we will also be exploring the use of traditional machine learning and/or deep learning across cohorts to investigate whether each cohort has their own distinct 'language profile'.
-\par 
-The final output of this task will be a paper discussing the results of our longitudinal study from a language analysis perspective.
-The target conference for this work will be the UK Alzheimer's Research conference in March 2021 with an expected submission date of December 2020, the Alzheimer's Association International Conference in July 2021 with an expected submission date of January 2021 and the Association of Computational Linguistics Conference (ACL) in July 2021 with an expected submission date of March 2021.
-\subsection{Task 5 - Analysis of performance and validation of the proposed Repeatable Neuropsychological Battery of Tests}  
-In this task, I will be looking at the results of the neuropsychological battery of tests without the additional language tasks to determine whether we can detect cognitive decline. 
-\par 
-We expect to see no decline in our HOC group within this timeframe and more severe decline in our EA group.
-Once again, we will be interested in whether there is any measurable decline in our SMCI or DMCI experimental groups with our battery of tests.
-\par 
-The final output of this task will be a paper discussing the results of the battery of tests, including how the battery does in terms of diagnostic and prognostic utility.
-The target conference for this work is to be decided.
-
-\subsection{Task 6 - Thesis Writing}
-The writing of the thesis, including additional over-running data analysis is planned for the last six months of the project (Oct '20 to Mar '21).
-
-\section{Research Challenges}
-\subsection{Data Collection}
-There are a few challenges that I have identified with the process of data collection.
-As with all longitudinal studies, the data collected is only valid if there are at least two sets of data points for a given participant.
-Given the population I am exploring, it is expected that there is some level of drop out rate over the course of the year and this has been accounted for.
-Nevertheless, a higher than average drop out rate would weaken the strength of the study, especially if those who dropped out came from the experimental group.
-\subsection{Commonality of MCI}
-Whilst research has consistently shown that those with a diagnosis of MCI tend to have a significantly higher risk of progression to dementia \cite{Roberts2014}.
-The progression at any given time is low and variable.
-Annual conversion rates vary across longitudinal studies conducted in this population, but researchers have shown that there is a higher conversion rate in clinical samples of this population than in community samples \cite{Farias2009}.
-In my proposed longitudinal study, I aim to collect a clinical MCI population, but depending on the availability of such samples I may have to collect from the community.
-If this is the case then, the prevalence of those with MCI who have language deficits within my participants may be low.
-\subsection{Sparsity of Language}
-I have tried to estimate the amount of language I will collect by conducting a pilot study on healthy controls.
-However, until the first sets of data collection with the MCI population, I will not be able to tell if this is a problem.
-I fully expect there to be little difference in the amount of language collected in the MCI population, but this remains a possibility which I will need to adjust for at the 2nd and 3rd data collection phases.
-
-\section{Risk Assessment}
-In this section, I look at the potential difficulties that may produce a delay to the proposed research, define any strategies that have already been put in place to account for this as well as define any other potential solutions.
-\subsection{Delay in data collection}
-When working with a clinical population, the collection of data can be subject to delays.
-This can be for a number of reasons including but not limited to sickness (patient or researcher), unavailability of participant for another reason or unavailability of venue to collect data.
-The collection of data could also be subject to delays due to ethical approval for the research being delayed.
-For individual participants unavailability, flexibility has been built into the data collection schedule to allow for an extra month (capacity for one additional cohort) of data collection per collection cycle.
-\subsection{Rejected Publications}
-For each task defined above, with the exception of task 2 and task 6 I have identified suitable conferences at which I can present my work.
-With the target conferences identified every effort will be made to meet the submission deadlines for these conferences.
-If submissions are rejected at these conferences, I have identified alternative conferences to present at.
-\subsection{Rejected NHS Ethics and/or University Ethics}
-Every effort will be made to submit a complete ethics application through the NHS.
-However, in the event that there is a delay in the ethical approval process then data collection will proceed with a priority placed on the control group.
-Future cohorts will then have proportionally more clinical participants until the research meets it's target numbers.
-In the event that NHS ethics is not granted in time for me to write up, participants will be sought from other areas such as the Alzheimer's Research pool of volunteers and venues where suitable candidates may be present such as meetings of the University of the Third age.
-
-\section{Ethical Considerations}
-In this section I discuss the ethical considerations related to this project and any adjustments that I have made to my research because of these considerations.
-I am aware of the code of human  research ethics published by the British Psychological Society \cite{BritishPsychologicalSociety2014} and my research has been designed in a way that fits with these guidelines.
-\subsection{Risk}
-I will be working with a middle-aged to elderly population who have identified in some way that there is a difficulty with cognition.
-However, I do not anticipate working with vulnerable groups, those lacking capacity or individuals in a dependent or unequal relationship.
-Should any participants I am working with lose capacity during the course of the study, I will remove that participants details and data from the study.
-\subsection{Informed Consent}
-Consent will be sought overtly at the recruitment and initial data collection point by way of an information sheet and associated consent form.
-At other data collection points, I will have information sheets available if participants request them and if such a request is made then I will seek consent again at this stage.
-As part of this process, I will be supplying my participants with my contact details should they have any question or for the purposes of withdrawing consent.
-These consent forms will be stored in a secure, locked cabinet at the university.
-\subsection{Confidentiality}
-Given the nature of the study and the possibility of disclosures of suicidal ideation, it is important that confidentiality is discussed at each data collection point.
-I will also provide an information sheet which details the confidentiality statement and also the situations in which I may need to break confidentiality.
-In addition, I will make a verbal confidentiality statement at each collection point and seek acknowledgement of their understanding and permission to continue.
-The spoken confidentiality statement will be as follows.
-\begin{center}
-	\textit{"Anything you say to me today will be confidential unless I feel that there is a risk to you, or to someone else.
-If this is the case, I may need to speak to or write to third parties such as your GP.
-I will endeavor to discuss this communications with yourself where it is possible to do so.
-Is this ok?"}
-\end{center}
-\subsection{Disclosure of suicidal ideation or other risks}
-As part of the battery of tests for this experiment, there is an opportunity for a participant to reveal that he or she has had suicidal ideation in the past two weeks (Patient Health Questionnaire, Question 9).
-This is something to explore further.
-As I have been trained in psychological risk assessments, it seems appropriate to carry out a further risk assessment.
-Optionally with the tape recorder turned on, or off as requested.
-Any text recorded during the risk assessment will not be included in the language analysis.
-Following this, some information about local services such as Improving Access to Psychological Therapies (IAPT), Samaritans and other appropriate services will be provided.
-A discussion will then be made about whether it is appropriate to notify the GP of the disclosure.
-In the event that there is imminent risk to self or others, a discussion will be made about who I need to contact.
-(Crisis Team and GP).
-In the event that the participant scores 0 on Q9 of the PHQ, no risk assessment will be undertaken.
-
-\section{Thesis - Provisional Table of Contents}
-In this section I provide a provisional table of contents based on the work I have either completed or proposed above.
-\begin{enumerate}
-	\item Introduction
-	\begin{enumerate}
-		\item Overview
-		\item Background and Context
-		\item Aims and Objectives
-		\item Challenges and Achievements
-	\end{enumerate}
-	\item Literature Review
-	\begin{enumerate}
-		\item Introduction
-		\item Assessing Mild Cognitive Impairment using cognitive tests.
-		\item Assessing Mild Cognitive Impairment using Language.
-		\item Natural Language Processing and Machine Learning
-		\item Conclusion
-	\end{enumerate}
-	\item Further Exploration of the Presidents Corpus
-	\begin{enumerate}
-		\item Introduction
-		\item Methodology
-		\item Results
-		\item Discussion
-		\item Conclusion
-	\end{enumerate}
-	\item Validation of a Repeatable Neuropsychological Battery of Tests
-	\begin{enumerate}
-		\item Introduction
-		\item Methodology
-		\item Pilot Study Results
-		\item Longitundinal Study results
-		\item Discussion
-		\item Conclusion
-	\end{enumerate}
-	\item Longitudnal Study
-	\begin{enumerate}
-		\item Introduction
-		\item Methodology
-		\item Experiment 1 Results
-		\item Experiment 2 Results
-		\item Discussion
-		\item Conclusion
-	\end{enumerate}
-	\item General Discussion
-	\begin{enumerate}
-		\item Discussion
-		\item Future Work
-		\item Conclusion
-	\end{enumerate}
-\end{enumerate}
-
-\chapter{Conclusion}
+The results of this work show that we can track a person's use of language through time in a number of ways, and it is possible for an individual to be his or her own control.
+This is important as it means the heterogenous nature of the MCI population does not impact results as much as if we were comparing those with MCI to controls.
+Equally, it would be helpful to have controls to ascertain what would be usual to expect in the decline of language in a healthy older adult.
 

### 2019-02-07 Source/Presidents Corpus Paper/paper1.tex
Paper 1 - Rough Draft

@@
 \usepackage{rotating}
+
+\usepackage{arxiv}
+
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{lipsum}
+
 %% Page Settings
@@
 %% Own Commands
-\pagestyle{fancy}
-\fancyhf{}
-\fancyhead[CE,CO]{Early Detection of MCI that progresses to dementia - An Interdisciplinary Approach}
-\fancyfoot[LE,RO]{\thepage}
+\input{settings/macros}
+\title{Longitudinal Language Analysis in the Diagnosis of Mild Cognitive Impairment: A Case Study}
 
-\renewcommand{\headrulewidth}{2pt}
-\renewcommand{\footrulewidth}{1pt}
-\input{settings/macros}
-\title{Qualifying Report}
-\date{2019-01-31}
-\author{Jomar Alcantara}
+\author{
+  Jomar C. Alcantara \\
+  Department of Computer Science \\
+  School of Engineering and Applied Sciences\\
+  Aston University\\
+  Birmingham, UK \\
+  \texttt{alcantaj@aston.ac.uk} \\
+  \And
+  Peter Sawyer \\
+  Department of Computer Science \\
+  School of Engineering and Applied Sciences \\
+  Aston University\\
+  Birmingham, UK \\
+  \texttt{p.sawyer@aston.ac.uk} \\
+  \And
+  Cristina Romani \\
+  Department of Psychology \\
+  School of Life and Health Science\\
+  Aston University\\
+  Birmingham, UK \\
+  \texttt{c.romani@aston.ac.uk} \\
+}
 % HEADER ENDS
@@
 \begin{document}
+\maketitle
 \section{Abstract}
@@
 Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
-
+\par 
 There has been a significant research in the area of language deterioration as a means of detecting Alzheimer's Disease.
@@
 
-\bibliographystyle{unsrt}
-\bibliography{QualifyingReport}
+\begin{thebibliography}{10}
+
+\bibitem{Prince2015}
+Martin Prince, Anders Wimo, Guerchet M, Ali GC, Wu~YT, and Prina M.
+\newblock {World Alzheimer Report 2015 The Global Impact of Dementia An
+  analysis of prevalence, incidence, cost and trends}.
+\newblock {\em Alzheimer's Disease International}, 2015.
+
+\bibitem{England2009}
+{Department of Health}.
+\newblock {Living well with dementia : Dementia Strategy (England)}.
+\newblock Technical report, Department of Health, 2009.
+
+\bibitem{Taler2008}
+Vanessa Taler and Natalie~A. Phillips.
+\newblock {Language performance in Alzheimer's disease and mild cognitive
+  impairment: A comparative review}.
+\newblock {\em Journal of Clinical and Experimental Neuropsychology},
+  30(5):501--556, 2008.
+
+\bibitem{Boschi2017}
+Veronica Boschi, Eleonora Catrical{\`{a}}, Monica Consonni, Cristiano Chesi,
+  Andrea Moro, and Stefano~F. Cappa.
+\newblock {Connected speech in neurodegenerative language disorders: A review}.
+\newblock {\em Frontiers in Psychology}, 8(MAR), 2017.
+
+\bibitem{Berisha2015}
+Visar Berisha, Shuai Wang, Amy LaCross, and Julie Liss.
+\newblock {Tracking Discourse Complexity Preceding Alzheimer's Disease
+  Diagnosis: A Case Study Comparing the Press Conferences of Presidents Ronald
+  Reagan and George Herbert Walker Bush}.
+\newblock {\em Journal of Alzheimer's Disease}, 45(3):959--963, 2015.
+
+\bibitem{Orimaye2014}
+Sylvester~Olubolu Orimaye, Jojo Sze-Meng Wong, and Karen~Jennifer Golden.
+\newblock {Learning Predictive Linguistic Features for Alzheimer's Disease and
+  related Dementias using Verbal Utterances}.
+\newblock {\em Proceedings of the Workshop on Computational Linguistics and
+  Clinical Psychology: From Linguistic Signal to Clinical Reality}, pages
+  78--87, 2014.
+
+\bibitem{Fraser2015}
+Kathleen~C. Fraser, Jed~A. Meltzer, and Frank Rudzicz.
+\newblock {Linguistic Features Identify Alzheimer's Disease in Narrative
+  Speech}.
+\newblock {\em Journal of Alzheimer's Disease}, 49(2):407--422, oct 2015.
+
+\bibitem{Bird2009}
+Steven Bird, Ewan Klein, and Edward Loper.
+\newblock {\em {Language Processing and Python}}.
+\newblock O'Reilly Media Inc, 2009.
+
+\bibitem{Richards1987}
+Brian Richards.
+\newblock {Type/Token Ratios: what do they really tell us?}
+\newblock {\em Journal of Child Language}, 14(02):201, jun 1987.
+
+\bibitem{Bird2000}
+Helen Bird, Matthew~A. {Lambon Ralph}, Karalyn Patterson, and John~R. Hodges.
+\newblock {The rise and fall of frequency and imageability: Noun and verb
+  production in semantic dementia}.
+\newblock {\em Brain and Language}, 73(1):17--49, 2000.
+
+\bibitem{Le2011}
+Xuan Le, Ian Lancashire, Graeme Hirst, and Regina Jokel.
+\newblock {Longitudinal detection of dementia through lexical and syntactic
+  changes in writing: A case study of three British novelists}.
+\newblock {\em Literary and Linguistic Computing}, 26(4):435--461, 2011.
+
+\bibitem{Garrard2005}
+Peter Garrard, Lisa~M. Maloney, John~R. Hodges, and Karalyn Patterson.
+\newblock {The effects of very early Alzheimer's disease on the characteristics
+  of writing by a renowned author}.
+\newblock {\em Brain}, 2005.
+
+\bibitem{Almor1999}
+Amit Almor, Daniel Kempler, Maryellen~C MacDonald, Elaine~S Andersen,
+  Lorraine~K Tyler, Lee Willis, Lori Altmann, Mariela Gil, Laila Lalami, Karen
+  Mar-blestone, Sarah Schuster, and Karen Stevens.
+\newblock {Why Do Alzheimer Patients Have Difficulty with Pronouns? Working
+  Memory, Semantics, and Reference in Comprehension and Production in
+  Alzheimer's Disease}.
+\newblock {\em Brain and Language}, 67:202--227, 1999.
+
+\bibitem{Wendelstein2015}
+Britta Wendelstein, J{\"{o}}rn Stegmeier, Claudia Frankenberg, Ekkehard Felder,
+  and Johannes Schr{\"{o}}der.
+\newblock {Changes in the use of pronouns in spoken language in the course of
+  preclinical to early stages of Alzheimer's disease}.
+\newblock {\em Alzheimer's {\&} Dementia}, 11(7):P784--P785, jul 2015.
+
+\bibitem{Jarrold2014}
+William Jarrold, Bart Peintner, David Wilkins, Dimitra Vergryi, Colleen Richey,
+  Maria~Luisa Gorno-Tempini, and Jennifer Ogar.
+\newblock {Aided diagnosis of dementia type through computer-based analysis of
+  spontaneous speech}.
+\newblock {\em Proceedings of the Workshop on Computational Linguistics and
+  Clinical Psychology: From Linguistic Signal to Clinical Reality}, pages
+  27--37, 2014.
+
+\bibitem{Wendelstein2014}
+Britta Wendelstein, Ekkehard Felder, J{\"{o}}rn Stegmeier, and Johannes
+  Schr{\"{o}}der.
+\newblock {Noun or Pronoun?: Using Parts of Speech in ALzheimer's Disease}.
+\newblock {\em Alzheimer's {\&} Dementia}, 10(4):P434, jul 2014.
+
+\end{thebibliography}
 

### 2019-02-07 Source/Presidents Corpus Paper/paper1.tex
Paper1 update

@@
 \maketitle
-\section{Abstract}
+\begin{abstract}
+Dementia is a neurological disorder that has been consistently growing in numbers over time.
+The analysis of language has been highlighted as a potential method for diagnosing Alzheimer's disease at its early or pre-clinical stages.
+This paper builds on the work of Berisha and Liss who looked at extracting language features from the non-scripted portions of press conferences from two US presidents, President Ronald Reagan, who would later go on to receive a diagnosis of Alzheimer's Disease and President George Herbert Walker Bush who received no such diagnosis.
+In this paper, we add the transcripts of President Donald John Trump who in terms of age is a much more appropriate match and we extract a larger variety of language features which have been shown to be associated with language decline due to Dementia.
+Once the features were extracted we completed a regression analyses.
+We found that President Reagan showed a significant decrease in unique words over time and his usage of nouns and adjectives.
+We also found that there was an increase in his usage of non-specific nouns, adverbs and pronouns.
+We found no significant trends in the transcripts of Presidents Bush and Trump.
+This study shows that language deterioration can be tracked over time in those with Mild Cognitive Impairment and is generally differentiable from healthy controls.
+The results also potentially demonstrate the utility of longitudinal language analysis as a method of diganosing Mild Cognitive Impairment in the future.
+\end{abstract}
+\keywords{Mild Cognitive Impairment \and Alzheimer's Disease \and Natural Language Processing \and Longitudinal \and Early Diagnosis \and Language}
+
 \section{Background}
@@
 \par
-In 2009, the Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
-A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
-As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages \cite{Prince2015} and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
-At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017}.
-I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
+In 2009, the UK's Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
+A lot of work has gone into this with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
+Recently, researchers have highlighted that technology, more specifically the domains of Natural Language Processing (NLP) and Machine Learning (ML), is ideally placed to aid in this endeavour \cite{Boschi2017}.
+It is theorized that we may be able to use these technologies to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML) \cite{Orimaye2014, Boschi2017}.
 \par
-There is a large body of research that looks at the decline in language of those with MCI and AD \cite{Taler2008, Boschi2017}.
+There is a large body of research that looks at the decline in language of those with MCI and AD \cite{Taler2008, Boschi2017, Emery2000}.
 However there is conflicting evidence in these studies about which declining language factors are associated with MCI and AD \cite{Taler2008, Boschi2017}.
-Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
+Some work therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
 There is some evidence to support this, Berisha et al \cite{Berisha2015}, has shown through a longitudinal language analysis of spontaneous speech that there are marked differences in different parts of speech in this process between those who would go on to have a diagnosis of AD and a healthy control.
@@
 
+\bibliographystyle{unsrt}
 \begin{thebibliography}{10}
@@
 
-\bibitem{Berisha2015}
-Visar Berisha, Shuai Wang, Amy LaCross, and Julie Liss.
-\newblock {Tracking Discourse Complexity Preceding Alzheimer's Disease
-  Diagnosis: A Case Study Comparing the Press Conferences of Presidents Ronald
-  Reagan and George Herbert Walker Bush}.
-\newblock {\em Journal of Alzheimer's Disease}, 45(3):959--963, 2015.
-
 \bibitem{Orimaye2014}
@@
   78--87, 2014.
+
+\bibitem{Emery2000}
+Olga~B. Emery.
+\newblock {Language Impairment in Dementia of the Alzheimer's Type: A
+  Hierarchical Decline?}
+\newblock {\em International Journal of Psychiatry in Medicine},
+  30(2):145--164, jul 2000.
+
+\bibitem{Berisha2015}
+Visar Berisha, Shuai Wang, Amy LaCross, and Julie Liss.
+\newblock {Tracking Discourse Complexity Preceding Alzheimer's Disease
+  Diagnosis: A Case Study Comparing the Press Conferences of Presidents Ronald
+  Reagan and George Herbert Walker Bush}.
+\newblock {\em Journal of Alzheimer's Disease}, 45(3):959--963, 2015.
 
@@
   Schr{\"{o}}der.
-\newblock {Noun or Pronoun?: Using Parts of Speech in ALzheimer's Disease}.
+\newblock {NOUN OR PRONOUN?: USING PARTS OF SPEECH IN ALZHEIMER'S DISEASE}.
 \newblock {\em Alzheimer's {\&} Dementia}, 10(4):P434, jul 2014.

### 2019-02-09 Source/Presidents Corpus Paper/paper1.tex
changes to paper

@@
 
-\usepackage{arxiv}
+\usepackage{arxiv}					% ARXIV Template
 
@@
 \usepackage{microtype}      		% microtypography
-\usepackage{lipsum}
 
@@
   \texttt{c.romani@aston.ac.uk} \\
+  \And
+  George Vogiatzis \\
+  Department of Computer Science \\
+  School of Engineering and Applied Sciences \\
+  Aston University \\
+  Birmingham, UK \\
+  \texttt{g.vogiatzis@aston.ac.uk} \\
 }
@@
 
+\twocolumn
 \section{Background}
@@
 These significant differences were in the number of unique words used per speech, the use of non-specific nouns and fillers and low-imageability verbs \cite{Berisha2015}.
+\par
+
+The contributions of this paper are.......
+
 \par 
@@
 	\begin{center}
-	\begin{tabular}{ | p{3cm} | p{6cm} |}
+	\begin{tabular}{ | p{3cm} | p{4cm} |}
 		\hline
@@
 	\begin{center}
-	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+	\begin{tabular}{ | p{2cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
 		\hline
@@
 	\begin{center}
-	\begin{tabular}{ | l | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+	\begin{tabular}{ | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
 		\hline
@@
 Compared to our controls (GWHB andDJT), we find some slight trends with GWHB but no such trends with DJT in his speech albeit his samples of speech span a shorter amount of time.
+
+\onecolumn
 \begin{figure}[H]
@@
 
+\twocolumn
 A criticism of Berisha and Liss's work is the problems they had with normalising the transcripts in terms of length.

### 2019-03-05 Source/Presidents Corpus Paper/paper1.tex
changes to paper1

--- 2019-02-07 Source/Presidents Corpus Paper/paper1.tex
@@
 \end{abstract}
-\keywords{Mild Cognitive Impairment \and Alzheimer's Disease \and Natural Language Processing \and Longitudinal \and Early Diagnosis \and Language}
+\keywords{Cognitive Dysfunction \and Alzheimer Disease \and Natural Language Processing \and Early Diagnosis \and Speech}
 
-\section{Background}
+\section{Introduction}
 Dementia has been identified as one of the fastest growing difficulties facing the world today.
@@
 \section{Methodology}
-I took 46 transcripts of Ronald Reagan’s (RR) press conferences from 1981 to 1988 and compared them with 134 press conferences by George H.
+We took 46 transcripts of Ronald Reagan’s (RR) press conferences from 1981 to 1988 and compared them with 134 press conferences by George H.
 W.
@@
 Trump (DJT).
-I analyzed transcripts for lexical features shown to change longitudinally with dementia  (for a comprehensive review of these, see the literature review above).
-For this collection of documents, I generated a number of features which looked at a number of different aspects of each document.
+We analyzed transcripts for lexical features shown to change longitudinally with dementia.
+For this collection of documents, We generated a number of features which looked at a number of different aspects of each document.
 These features encompassed, word level, sentence level and document level features and included a number of features contained in the study by Berisha and Liss with the aim of replicating and extending on their findings.
-These findings were.
-number of unique words, non-specific nouns and fillers and low imageability (LI) verbs.
-Imageability is characterized, according to Berisha and Liss, as the ease with which a term gives rise to a sensory .mental image.
-I compared the trends described in the transcripts of RR and GHWB, but also included DJT.
+These features were: number of unique words, non-specific nouns and fillers and low imageability (LI) verbs.
+Imageability is characterized, according to Berisha and Liss, as the ease with which a term gives rise to a sensory mental image.
+We compared the trends described in the transcripts of RR and GHWB, but also included DJT.
 Berisha and Liss originally made the comparison as it GHWB (GHWB - age at the start of presidency - 64 years, 222 days) was the closest match to RR in terms of age (RR  - age at the start of presidency, 69 yeaars and 349 days).
-However, with the inauguration of Trump, he now is the closest comparible president in terms of age (DJT - age at the start of presidency - 70 years, 220 days).
+With the inauguration of Trump, he now is the closest comparible president in terms of age (DJT - age at the start of presidency - 70 years, 220 days).
 It would be interesting to look at a comparison of RR and DJT to see whether the comparisons made by Berisha and Liss hold true with this more appropriate match (in terms of age).
 DJT as with GHWB has no known diagnosis of AD.
-I used the press conference transcripts in the American Presidency Project (APP) archive as a data source for this project.
+We used the press conference transcripts in the American Presidency Project (APP) archive as a data source for this project.
 The APP is a comprehensive and organized searchable database of presidential documents, including transcripts of speeches, transcripts of news conferences, and other public documents.
 \subsection{Pre-processing}
-To generate the files necessary for analysis, I downloaded each transcript and performed the following changes.
-I omitted the prepared statement by the president and any speech by other individuals.
-I started each transcript at the beginning of the first answer to a question.
-I filtered any annotations that were added to the transcript, including any references or clarifications, and any laughter.
-It's worth noting that there appears to be a difference in how 'hesitations' were marked down between each president, for RR hesitations were marked by a single hyphen whereas for GHWB hesitations are marked by a double hyphen.
-In order to maintain consistency when parsing through the documents, I have changed both types of hesitation to be marked by a single hyphen.
-I also omitted one word sentences as this data would, from a theoretical perspective, not be relevant for language analysis.
-I did not control for the length of the document, but generated features which would normalise by the length of the document.
-I therefore was able to include all press conferences by both Ronald Reagan and George H.W.
+To generate the files necessary for analysis, We downloaded each transcript and performed the following changes.
+We omitted the prepared statement by the president and any speech by other individuals.
+We started each transcript at the beginning of the first answer to a question.
+We omitted any annotations that were added to the transcript, including any references or clarifications, and any laughter.
+It's worth noting that there appears to be a difference in how 'hesitations' were marked down between each president, for RR hesitations were marked by a single hyphen whereas for GHWB and DJT hesitations are marked by a double hyphen.
+In order to maintain consistency when parsing through the documents, we changed both types of hesitation to be marked by a single hyphen.
+We also omitted one word sentences as this data would, from a theoretical perspective, not be relevant for language analysis.
+We did not control for the length of the document, but generated features which would subsequently be normalised by the length of the document.
+We therefore were able to include all press conferences by both Ronald Reagan and George H.W.
 Bush where there was a question and answer session conducted at least in part by the sitting president (2 press conferences of GHWB were ommited due to a lack of a question and answer session).

### 2019-03-06 Source/Thesis/Thesis.tex
PhD Folder Tidy
changes to Thesis
Changes made to Thesis

@@
 \input{settings/packages}
+\usepackage{graphicx}
+\usepackage{float}
+\usepackage{enumitem}
+\usepackage{fancyhdr}
+\usepackage{rotating}
 %% Page Settings
@@
 %% Own Commands
+\pagestyle{fancy}
+\fancyhf{}
+\fancyhead[CE,CO]{\leftmark}
+\fancyfoot[CE,CO]{Early Detection of MCI that progresses to dementia - An Interdisciplinary Approach}
+\fancyhead[LE,RO]{\thepage}
+
+\renewcommand{\headrulewidth}{2pt}
+\renewcommand{\footrulewidth}{1pt}
 \input{settings/macros}
 \title{Thesis}
-\date{2018-09-17}
+\date{2019-02-05}
 \author{Jomar Alcantara}
@@
 \listoftables
-\chapter{Overview of Problem}
+\chapter{Introduction}
 Dementia has been identified as one of those fast growing difficulties facing the world.
@@
 Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
-\par
-This report is structured in the following way.
-A literature review of current methods of cognitive assessment of MCI, an review of how language is said to be impacted in the MCI and Early AD population and a review of NLP and ML techniques that have the potential to be applied to this domain is presented in Chapter 2.
-The findings from a replication and extension of previous research using an existing dataset is presented in Chapter 3.
-I detail the aims of my research and my proposed future work plan including distinct tasks to be completed over the next two years in Chapter 4.
-Finally, my conclusions are summarised in Chapter 5.
-
-\chapter{Introduction}
+
 The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to to time-consuming and expensive psychological tests.
@@
 \newline
-\chapter{Background}
-\chapter{Literature Review}
-\section{Introduction}
+\chapter{Background and related work}
 Alzheimer's Disease (AD) is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
@@
 It is important for any studies moving forward to use comparable standards when it comes to their experimental groups, such that the picture can be made clearer.
+
+\section{Mild Cognitive Impairment}
+% Mild cognitive impairment as a diagnostic entity. Petersen (2004)
+Research has noted that there is an observable transitional period between normal ageing and the point where it is possible to diagnose probable AD.
+This transitional period has been described in a range of different terms including Mild Cognitive Impairment (MCI), prodromal dementia and Mild Neurocognitive Disorder (mNCD).
+Traditionally, a diagnosis of probable dementia can only be given once cognitive deterioration has reached a certain point, for example - when a patient scores between 20-24 on the MMSE, this has been given the term Mild Alzheimer's Disease.
+\par 
+In some ways this transitional period is difficult to pinpoint.
+At the start of this transitional period, this may manifest in extremely subtle ways such that it might not be differentiable from cognitive decline associated with aging.
+Equally there is an overlap at the end of this transitional period and the very early stages of clinically diagnosable probably AD.
+\par
+The American Psychiatric Association (APA) recently published the DSM-5 and added the term Mild Neurocognitive Disorder
+
+% Mild Cognitive Impairment. Petersen (2016)
+
+\subsection{Criteria for Mild Cognitive Impairment due to Alzheimer Disease}
+
+\subsection{}
+
+\subsection{Rates of conversion from MCI to Dementia}
+The American Academy of Neurology (AAN) concluded that MCI is a useful marker to identify since the condition significantly increases the risk of progression to Dementia.
+
+% A Longitudinal Follow-Up of 550 Mild Cognitive Impairment Patients: Evidence for Large Conversion to Dementia Rates and Detection of Major Risk Factors - Espinosa et al (2013)
+The annual conversion of amnestic MCI to AD reported in different studies ranges from 10\% to 40\%.
+
+\section{Online collection of Data}
 

### 2019-03-14 Source/Thesis/Response.tex
added personspec and response

\documentclass[11pt]{article}
\begin{document}
\textbf{Response to the Qualification Report and Viva Voce feedback}
\par 
\textbf{Your feedback}
The student had a very good understanding of the literature and could summarise it well. It was a bit lacking in the machine learning area but he has the capacity for improving that and is aware of what he needs to do. His research background was good but there needed to be more focus on the research questions. These were not explored well in the QR but were much clearer in the viva. He now has some very good ideas to take forward that should lead to a Ph.D.
\par 
The main advice we would give is to be much more careful about how and what data you collect. Getting the participants to do it themselves on a daily basis by, for example, a speech diary that they record and submit, is much more practical than a battery of tests in a clinic. Plus you do not want to be asking about clinical diagnoses using, for example, PHQ9 that measures depression because this raises a lot of ethics issues that will make it harder for NHS ethics approval. Keep the data collection simple, especially as the focus of the Ph.D is about changes in language and not about the cognitive tests themselves. We do not think the latter are necessary because the focus of the Ph.D is on whether you can detect decline by differences in language use over time and not about determining the clinical underpinnings of those differences.

\section{Research Aims and Questions}
The broad objectives of my research are to explore whether the application of Natural Language Processing and Machine Learning techniques to spontaneous speech samples can improve our ability to diagnose Mild Cognitive Impairment in patients that will directly lead to Alzheimer's Disease. Another objective of my research is to provide some clarity as to which language features are most important to measure when looking for evidence of MCI and/or AD. 
\begin{itemize}
	\item \textbf{(H1)} Monitoring language samples and other measures of cognitive decline over the course of the year will be enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment
	\item \textbf{(H2)} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, these with Alzheimers Disease and decline in those with MCI.
\end{itemize}

\section{Data Collection Methodology}

\section{Project Aims}

\end{document}

### 2019-03-22 Archive/DhaneshaPersonSpec/PersonSpec.tex
reorganised files

\documentclass[11pt]{article}
\begin{document}
\textbf{MAIN DUTIES AND RESPONSIBILITIES}
\begin{enumerate}
	\item To support and contribute to the research activity of the project
	\item To collect data from project participants
	\item To recruit project participants from Join Dementia Research databases and other sources as guided and to identify needs and priorities
	\item To work independently according to a plan agreed with project staff
	\item To manage data for statistical analysis including data entry and data cleaning
	\item To ensure the secure and proper maintenance of relevant project files, research materials and data
	\item To assist in the preparation of interim and final reports for the project lead
	\item To be willing to travel as this project may involve working at other sites
	\item To assist the project team to ensure that all deadlines are met and that the study progresses according to the agreed timetable
	\item To seek and utilize supervision from the Project Lead as appropriate
	\item To carry out administrative duties required for co-ordination of the project
	\item To attend such meetings as necessary for satisfactory completion of the duties of the post
	\item Through liaison with clinical, administrative and managerial staff, to ensure that that ethical standards relating to confidentiality and the protection of information are adhered to
	\item Sensitively deal with emotive and distressing topics through project
	\item Sensitively and effectively work with patient groups to explore a range of emotive and highly contentious subjects
	\item Regularly undertake travel to a variety of sites using public transport
	\item Use Visual Display Unit equipment more or less continuously on almost all days
	\item To prioritise own workload, seeking guidance from and accepting the advice of the Project Lead where appropriate.  
	\item To be involved in the discussion of any queries associated with the project with the participants and professionals involved, in consultation with the project lead
	\item To conduct all other duties as agreed by the project lead 
\end{enumerate}

\end{document}

### 2019-03-22 ryanCalc.py

@@ -1,58 +1,0 @@
-def add(x, y):
-    return x + y
-
-def subtract(x, y):
-    return x - y
-
-def multiply(x, y):
-    return x * y
-
-def divide(x, y):
-    return x / y
-
-def modulus(x, y):
-    return x % y
-
-def getNumbers():
-    num1 = input('Enter the first number : ')
-    num2 = input('Enter the second number : ')
-    return(num1, num2)
-
-def Menu(choice):
-    if choice == 1:
-        num1, num2 = getNumbers()
-        num1 = int(num1)
-        num2 = int(num2)
-        print(num1, "+", num2, add(num1, num2))
-    elif choice == 2:
-        num1, num2 = getNumbers()
-        num1 = int(num1)
-        num2 = int(num2)
-        print(num1, "-", num2, subtract(num1, num2))
-    elif choice == 3:
-        num1, num2 = getNumbers()
-        num1 = int(num1)
-        num2 = int(num2)
-        print(num1, "*", num2, multiply(num1, num2))
-    elif choice == 4:
-        num1, num2 = getNumbers()
-        num1 = int(num1)
-        num2 = int(num2)
-        print(num1, "/", num2, divide(num1, num2))
-    elif choice == 5:
-        num1, num2 = getNumbers()
-        num1 = int(num1)
-        num2 = int(num2)
-        print(num1, "%", num2, modulus(num1, num2))
-    else:
-        print('Not a legal operation')
-    
-def MenuChoice ():
-    print('Select Operation')
-    print('1: Add')
-    print('2: Subtract')
-    print('3: Multiply')
-    print('4: Divide')
-    choice = input('Please enter a selection: ')
-    int(choice)
-    return choice

### 2019-03-25 Source/Thesis/Response.tex
adds to Response

@@
 \documentclass[11pt]{article}
+
+\usepackage{setspace}
+\onehalfspacing
+
+\usepackage{etoolbox}
+\AtBeginEnvironment{quote}{\singlespacing\small\itshape}
+
 \begin{document}
-\textbf{Response to the Qualification Report and Viva Voce feedback}
+\title{Response to the Qualification Report and Viva Voce feedback}
+\maketitle
 \par 
-\textbf{Your feedback}
+\begin{quote}
 The student had a very good understanding of the literature and could summarise it well.
@@
 \par 
+\bigskip
 The main advice we would give is to be much more careful about how and what data you collect.
@@
 We do not think the latter are necessary because the focus of the Ph.D is on whether you can detect decline by differences in language use over time and not about determining the clinical underpinnings of those differences.
+\end{quote}
 
 \section{Research Aims and Questions}
-The broad objectives of my research are to explore whether the application of Natural Language Processing and Machine Learning techniques to spontaneous speech samples can improve our ability to diagnose Mild Cognitive Impairment in patients that will directly lead to Alzheimer's Disease.
-Another objective of my research is to provide some clarity as to which language features are most important to measure when looking for evidence of MCI and/or AD.
+In my The revised objectives of my research are to explore whether the application of Natural Language Processing and Machine Learning techniques to spontaneous speech samples can improve our ability to diagnose Mild Cognitive Impairment in patients that will directly lead to Alzheimer's Disease and to provide some clarity as to which language features are most important to measure when looking for evidence of MCI and/or AD.
 \begin{itemize}
-	\item \textbf{(H1)} Monitoring language samples and other measures of cognitive decline over the course of the year will be enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment
-	\item \textbf{(H2)} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, these with Alzheimers Disease and decline in those with MCI.
+	\item \textbf{(Hypothesis One)} Monitoring language samples over the course of the year will be enough to detect cognitive decline in those diagnosed with Mild Cognitive Impairment and Early Alzheimer's Disease at specific points in time.
+	\item \textbf{(Hypothesis Two)} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, these with Alzheimers Disease and decline in those with MCI.
+	\item \textbf{(Hypothesis Three)} T
 \end{itemize}
 
-\section{Data Collection Methodology}
+\section{Data Collection and Analysis Methodology}
+One of the main concerns I have had was about the amount of data collection that needed to be done.
+Given the low rate of progression from MCI to AD year on year, there is a non-zero possibility that I could have a cohort where no-one progresses to AD.
+Equally, working with a clinical population would invariably require NHS Ethics.
+Since this problem is essentially a classification problem, in theory we can look at things from a different perspective.
 
-\section{Project Aims}
+\subsection{Why not stick to picture tasks or narrative storytelling tasks?}
+Semi-structured speech data in commonly available data sets such as the DementiaBank corpus or the Wisconsin Registry for Alzheimer's Prevention (WRAP) focus on picture description tasks.
+These samples are usually collected once an individual is referred to a clinician for assessment.
+
+\subsection{One Class Classification}
+The prevalence of Dementia in the UK's population is roughly 1 in 14 or 7\%.
+The prevalence of MCI has only briefly been explored, however some research by a team at the Univesity of Leeds indicate that the prevalence of MCI is between 3.2 - 24\%, and found that the prevalence of MCI in people who have self-reported memory problems is almost double (18\%) the amount of do not self report.
+Whilst this means that it is entirely possible to recruit a cohort with MCI and AD, it is comparatively costly in terms of time and resources.
+Given that, from a statistical perspective, this is a highly imbalanced population we can look at the problem as a kind of anomaly detection.
+\par 
+This has some precedent in the field of . In 
+\par 
+One-class classifiers analyze only examples of the majority class (in this case - Healthy Older Adults) to learn a classification boundary that excludes outlier examples.
+When dealing with heavily imbalanced datasets with text classifcation as an objective, the one-class classifier approach has been demonstrated to outperform conventional two-class classification.
+
+\section{Proposed Methodology}
+\subsection{Stage 1 - Exploratory Data Analysis of DementiaBank dataset}
+
+\subsection{Stage 2 - Collection of Longitudinal Data}
+Given the aim is to collect a large sample of language from a healthy population.
+I now estimate the numbers of samples to be as follows.
+100 Healthy Older Adults, 20 MCI, 20 Early AD.
+The 20 MCI and 20 Early AD will be categorised based on a clinical diagnosis given by a mental health professional.
+In addition, from the Healthy Older Adults, I will ascertain a participants level of concern about their cognitive ability with a simple 2 item questionnaire.
+The intention, if there are enough people who express a level of concern, is to conduct an of the majority class with and without those who demonstrate concern.
+\par 
+\begin{itemize}
+	\item Question 1. How concerned are you about your cognitive health? Rate on a scale of 0 to 10, where 0 is not concerned at all and 10 is extremely concerned.
+	\item Question 2. If you answered 1 or more on Question 1, please explain what your concerns are.
+\end{itemize}
+\par
+
+\subsection{Stage 3 - Analysis of Longitudinal Data}
+Once the data has been collected.
+The analysis of the data will be begin from the perspective of treating the healthy older adults as the Majority class, and all others as 'outliers'.
+The first stage of this is to see whether using a One-class classifier is sensitive enough to detect those with a diagnosis of MCI and AD without differentiating between the two.
+
+This is important from the perspective of identifying people
+
+\section{Proposed Extension}
+\subsection{Analysis of written blogs}
 

### 2019-03-27 Source/Thesis/Response.tex
update to response

@@
 \subsection{Stage 1 - Exploratory Data Analysis of DementiaBank dataset}
+The aim of this stage is to explore language features and acoustic features to an existing dataset.
+This allows the development of an automated pipeline using a more accurate speech-to-language application (IBM Watson).
+This also allows us the opportunity to explore the use of the one-class classifier to an existing data set as a proof in concept.
 
@@
 In addition, from the Healthy Older Adults, I will ascertain a participants level of concern about their cognitive ability with a simple 2 item questionnaire.
-The intention, if there are enough people who express a level of concern, is to conduct an of the majority class with and without those who demonstrate concern.
+The intention, if there are enough people who express a level of concern, is to conduct an analysis of the majority class with and without those who demonstrate concern.
 \par 
@@
 \subsection{Stage 3 - Analysis of Longitudinal Data}
-Once the data has been collected.
+Once the data has been collected, we can test the automated pipeline on new data which allows the validation of this protocol.
 The analysis of the data will be begin from the perspective of treating the healthy older adults as the Majority class, and all others as 'outliers'.
@@
 
-\section{Proposed Extension}
-\subsection{Analysis of written blogs}
-
+\section{Revised Chapters of Thesis}
+\begin{itemize}
+	\item Chapter 1 - Introduction
+	\item Chapter 2 - Background and Related Work
+	\item Chapter 3 - Feature Engineering and Classification of DementiaBank Transcripts and Audio.
+	\item Chapter 4 - Comparison of different speech to language applications.
+	\item Chapter 5 - Classification of language generated from Longitudinal Study.
+	\item Chapter 6 - Techniques in feature engineering - What features matter? Comparing language generated from picture description task to language generated from our task.
+	\item Chapter 7 - Conclusions and Future Work
+\end{itemize}
 \end{document}

### 2019-03-27 Source/Thesis/Thesis.tex
changes to Thesis.tex
changes 27th March

@@
-%% TO DO LIST
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% Document Class
-\documentclass[a4paper]{report}
+\documentclass[a4paper]{book}
 %% Packages
@@
 \title{Thesis}
-\date{2019-02-05}
+\date{2019-03-27}
 \author{Jomar Alcantara}
@@
 % Abstract
-\section{Abstract}
+\section*{Abstract}
 % Acknowledgements
-\section{Acknowledgements}
+\section*{Acknowledgements}
 \tableofcontents
@@
 Giles et al for example studied adults with minimal, mild or moderate AD vs controls and found that adults with AD produced fewer overall information units than controls \cite{Giles1996}.
-\subsection{Conciseness of information}1`
+\subsection{Conciseness of information}
 Conciseness has been defined as the number of words a speaker uses to express ideas.
@@
 
+\subsection{One Class Classification}
+The prevalence of Dementia in the UK's population is roughly 1 in 14 or 7\%.
+The prevalence of MCI has only briefly been explored, however some research by a team at the Univesity of Leeds indicate that the prevalence of MCI is between 3.2 - 24\%, and found that the prevalence of MCI in people who have self-reported memory problems is almost double (18\%) the amount of do not self report.
+Whilst this means that it is entirely possible to recruit a cohort with MCI and AD, it is comparatively costly in terms of time and resources.
+Given that, from a statistical perspective, this is a highly imbalanced population we can look at the problem as a kind of anomaly detection.
+\par 
+This has some precedent in the field of . In 
+\par 
+One-class classifiers analyze only examples of the majority class (in this case - Healthy Older Adults) to learn a classification boundary that excludes outlier examples.
+When dealing with heavily imbalanced datasets with text classifcation as an objective, the one-class classifier approach has been demonstrated to outperform conventional two-class classification.
+
 \section{Discussion}
@@
 A small but growing amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
-
-\subsection{Future Work}
+\par 
 One of the areas for research to study is a careful examination of the features that are being used to measure language deterioration.
@@
 This area of research is extremely promising in its early results and the impact of successful research would be life changing for both individuals and the health of the worlds aging population in general.
-\chapter{Initial Study of the Presidents corpus}
-\section{Background}
-There has been a significant research in the area of language deterioration as a means of detecting Alzheimer's Disease.
-This usually takes the form analysis of speech recorded as part of a cognitive assessment such as the Picture Description Task \cite{Orimaye2014,Fraser2015}.
-Given that language samples are relatively easy to collect, research has moved towards analysis of spontaneous speech.
-An good example of this type of research is the study conducted by Berisha and Liss which looked at the differences in language use between two US presidents, Ronald Reagan (who would go on to receive a diagnosis of Dementia) and George H.W.
-Bush who acted as a matched control based on Age \cite{Berisha2015}.
-They found several differences in language use which they felt acted as indicators of Reagan's difficulties with language due to dementia.
-These significant differences were in the number of unique words used per speech, the use of non-specific nouns and fillers and low-imageability verbs \cite{Berisha2015}.
-\par   
-This study replicates work done by Berisha and Liss and extends this by adding Donald Trump as an alternative, more appropriate comparison to Ronald Reagan as he is much closer in age than George H.W.
-Bush.
-This experiment will look at the features originally identified by Berisha and Liss, as well as any others that have potential as discussed in the literature review above.
-\section{Methods}
-I took 46 transcripts of Ronald Reagan’s (RR) press conferences from 1981 to 1988 and compared them with 134 press conferences by George H.
-W.
-Bush (GHWB) and 29 press conferences conducted by Donald J.
-Trump (DJT).
-I analyzed transcripts for lexical features shown to change longitudinally with dementia  (for a comprehensive review of these, see the literature review above).
-For this collection of documents, I generated a number of features which looked at a number of different aspects of each document.
-These features encompassed, word level, sentence level and document level features and included a number of features contained in the study by Berisha and Liss with the aim of replicating and extending on their findings.
-These findings were.
-number of unique words, non-specific nouns and fillers and low imageability (LI) verbs.
-Imageability is characterized, according to Berisha and Liss, as the ease with which a term gives rise to a sensory .mental image.
-I compared the trends described in the transcripts of RR and GHWB, but also included DJT.
-Berisha and Liss originally made the comparison as it GHWB (GHWB - age at the start of presidency - 64 years, 222 days) was the closest match to RR in terms of age (RR  - age at the start of presidency, 69 yeaars and 349 days).
-However, with the inauguration of Trump, he now is the closest comparible president in terms of age (DJT - age at the start of presidency - 70 years, 220 days).
-It would be interesting to look at a comparison of RR and DJT to see whether the comparisons made by Berisha and Liss hold true with this more appropriate match (in terms of age).
-DJT as with GHWB has no known diagnosis of AD.
-I used the press conference transcripts in the American Presidency Project (APP) archive as a data source for this project.
-The APP is a comprehensive and organized searchable database of presidential documents, including transcripts of speeches, transcripts of news conferences, and other public documents.
-\subsection{Pre-processing}
-To generate the files necessary for analysis, I downloaded each transcript and performed the following changes.
-I omitted the prepared statement by the president and any speech by other individuals.
-I started each transcript at the beginning of the first answer to a question.
-I filtered any annotations that were added to the transcript, including any references or clarifications, and any laughter.
-It's worth noting that there appears to be a difference in how 'hesitations' were marked down between each president, for RR hesitations were marked by a single hyphen whereas for GHWB hesitations are marked by a double hyphen.
-In order to maintain consistency when parsing through the documents, I have changed both types of hesitation to be marked by a single hyphen.
-I also omitted one word sentences as this data would, from a theoretical perspective, not be relevant for language analysis.
-I did not control for the length of the document, but generated features which would normalise by the length of the document.
-I therefore was able to include all press conferences by both Ronald Reagan and George H.W.
-Bush where there was a question and answer session conducted at least in part by the sitting president (2 press conferences of GHWB were ommited due to a lack of a question and answer session).
-\subsection{Feature Selection}
-We calculated the following features for each transcript in turn using the NLTK (see section 2 for a description) \cite{Bird2009} and Python. 
-\par 
-\subsubsection{Measures of lexical variation}
-We constructed two features of lexical variation.
-Firstly we looked at the number of unique words.
-To do this we were able to split each transcript into individual words and changed them to lowercase using NLTK and were then count the number of unique words that appeared in each transcript.
-We also used the TTR formula (see section 2 for a description) for a feature that measures lexical diversity independent of sample size \cite{Richards1987}.
-\subsubsection{Fillers, Non-Specific Nouns and LI Verbs}
-For these features, we counted the number of occurrences for different categories (see table for list of categories tracked and the words counted).
-The features were used by Berisha and Liss in their research \cite{Berisha2015} and were taken from work done by Bird et al \cite{Bird2000}.
-
-\begin{table}[H]
-	\begin{center}
-	\begin{tabular}{ | p{3cm} | p{6cm} |}
-		\hline
-		Category & Words \\ \hline
-		Fillers & \textit{"well", "so", "basically", "actually", "literally", "um", "ah"} \\ \hline
-		Non Specific Nouns & \textit{"something", "anything", "thing", "everything"} \\ \hline
-		LI Verbs & \textit{"be", "come", "do", "get", "give", go", "know", "look", "make", "see", "tell", "think", "want"} \\ \hline
-
-	\end{tabular}
-	\caption{\label{tab:table-name}Categories and Words Counted}
-	\end{center} 
-\end{table}
-
-\subsubsection{Usage of parts of speech}
-This section involves using a Part of Speech tagger (PoS) which analyses a sentence and assigns a 'tag' to each word based on the function the word has in a sentence.
-At a basic level this can be divided into the eight defined parts of speech: 'nouns', 'pronouns', verbs', 'adjectives', 'adverbs', 'conjunctions', 'prepositions' and 'interjections' but can be further subcategorised.
-We used the PoS tagger built into NLTK to tag each transcript in turn and used these the counts from each of these eight categories in our analysis.
-In addition to frequency counts we also normalised these features by dividing the frequency count by the number of words in the document to take into account transcript length.
-
-\section{Results}
-One of the most important thing to note is the wide variety of samples between the three presidents and also the varying timescales.
-RR participated in 46 press conferences over eight years (an average of 5.75 a year) which is the fewest number of press conferences given by an American president during their term of office.
-GHWB participated in 136 press conferences over four years (an average of  34 a year) and DJT participated in 29 press conferences to date (an average of 19.3 per year).
-Equally, there are variances in the average number of words.
-RR produced an average of 3424 words per conference compared to 2608 by GHWB (unpaired t = 4.434, p\textless 0.001) and DJT at 1849 words (unpaired t = 6.524, p\textless 0.001).
-\begin{table}[H]
-	\begin{center}
-	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
-		\hline
-		& RR & GHWB & DJT \\ \hline
-		Total Words & 3423.91 (416.42) & 2607.72 (1210.38) & 1848.65 (1549.38) \\ \hline
-		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
-		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
-		LI Verbs & 124.22 (17.89)& 103.45 (51.87) & 84.48 (75.78) \\ \hline
-	\end{tabular}
-	\caption{\label{tab:table-name}Means and Standard Deviations of important features}
-	\end{center} 
-\end{table}
-
-In terms of unique words, we found that RR used significantly more unique words, non-specific nouns and low imageability verbs than GHWB and DJT (see Table 3.3).
-Some of these differences are due to the length of the sample, particularly in the case of DJT where his average sample is almost half the sample of RR.
-It could also be said that this could be down to differences in linguistic abilities or speaking style \cite{Berisha2015, Le2011}.
-However, we can certainly see that as controls GWHB and DJT are comparative in relation to non-specific nouns and LI verbs.
-
-\begin{table}[H]
-	\begin{center}
-	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
-		\hline
-		& RR v GHWB & RR v DJT & GHWB v DJT \\ \hline
-		Total Words & \textbf{4.434***} & \textbf{6.524***} & \textbf{2.899**} \\ \hline
-		Unique Words & \textbf{6.832***} & \textbf{11.403***} & \textbf{4.137***} \\ \hline
-		Non Specific Nouns & \textbf{7.877***} & \textbf{3.426**} & -0.574 \\ \hline
-		LI Verbs & \textbf{2.656**} & \textbf{3.420***} & 1.628 \\ \hline
-		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
-		\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
-		\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
-	\end{tabular}
-	\caption{\label{tab:table-name}RR T-tests vs GWB and DJT}
-	\end{center} 
-\end{table}
-
-We then looked at the data from a longitudinal perspective as we are interested seeing whether we can track various language variables and their progress over time.
-We ran a number of Pearsons correlations with transcript index number as a time reference and the dependant variables (Table 3.4).
-For our controls, we found them to be stable for the most part with the main highlights being a decrease in Adverb usage for DJT (R=-0.36, p=0.049) and a steady but not severe decline in a number of variables for GWHB, namely total word count, unique words, low imageability words and verb usage.
-\par 
-For RR, his decline is more marked and more widespread through his language use.
-We noticed an significant increase in adverb (R=0.41, p=0.004) and pronoun usage (R=0.65, p\textless0.001), as well as a slight usage increase in Non-specific nouns(R=0.30, p=0.03).
-There was a highly significant decrease in number of unique words (R=-0.56, p\textless0.001) and noun usage (R=-0.70, p\textless0.001).
-Also very significant decrease in adjective usage (R=-0.40, p=0.005) and a significant decrease in total word count (R=-0.31, p=0.03).
-
-\begin{table}[H]
-	\begin{center}
-	\begin{tabular}{ | l | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
-		\hline
-		& RR & GHWB & DJT \\ \hline
-		Word Count & \textbf{-0.31*} & \textbf{-0.21*} & 0.08 \\ \hline 
-		Unique Words & \textbf{-0.56***} & \textbf{-0.25**} & 0.16 \\ \hline
-		Non Specific Nouns & \textbf{0.30*} & -0.08 & -0.03 \\ \hline
-		LI Verbs & -0.19 & \textbf{-0.20**} & 0.02 \\ \hline
-		Nouns Normalised & \textbf{-0.70***} & -0.03 & 0.14 \\ \hline
-		Verbs Normalised & \textbf{0.36**} & \textbf{0.24***} & -0.03 \\ \hline
-		Adjectives Normalised & \textbf{-0.40**} & 0.08 & -0.34 \\ \hline
-		Adverbs Normalised & \textbf{0.41***} & 0.02 & \textbf{-0.36*} \\ \hline
-		Pronouns Normalised & \textbf{0.65***} & 0.13 & 0.07 \\ \hline
-		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
-    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
-    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
-	\end{tabular}
-	\caption{\label{tab:table-name}Pearson Correlations for Features}
-	\end{center} 
-\end{table}
-
-\section{Discussion}
-President Reagan received his diagnosis of AD in August 1994 but using transcripts of speeches he made in his two terms as President (January 1981 - January 1989) we have be able to identify certain changes in his use of language that we might ascribe to the onset of MCI and early AD.
-Despite differences in our methodology, our research supports the findings of Berisha and Liss in that we both find a significant decrease in unique words over time and an increase in non-specifc noun usage.
-Compared to our controls (GWHB andDJT), we find some slight trends with GWHB but no such trends with DJT in his speech albeit his samples of speech span a shorter amount of time.
-\begin{figure}[H]
-	\centering
-	\begin{minipage}[b]{0.4\textwidth}
-		\includegraphics[width=190px, height=140px]{images/RRUniqueWords.png}
-		\caption{Ronald Reagan - Unique Words over time}
-	\end{minipage}
-	\hfill
-	\begin{minipage}[b]{0.4\textwidth}
-		\includegraphics[width=190px, height=140px]{images/GWHBUniqueWords.png}
-		\caption{George H.W. Bush - Unique Words over time}
-	\end{minipage}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\begin{minipage}[b]{0.4\textwidth}
-		\includegraphics[width=190px, height=140px]{images/RRNSNouns.png}
-		\caption{Ronald Reagan - Non-specifc Nouns over time}
-	\end{minipage}
-	\hfill
-	\begin{minipage}[b]{0.4\textwidth}
-		\includegraphics[width=190px, height=140px]{images/DJTNSNouns.png}
-		\caption{Donald J. Trump - Non-specific Nouns over time}
-	\end{minipage}
-\end{figure}
-
-A criticism of Berisha and Liss's work is the problems they had with normalising the transcripts in terms of length.
-This was also a problem in the work of Garrard et al \cite{Garrard2005, Le2011}.
-Whilst it is important to control for outliers, there are other ways in which we can control for length of sample.
-\par 
-Interestingly, when we normalised the various types of words used by the presidents we found some interesting patterns that further differentiated RR from the controls.
-Whilst Non-specific nouns increased over time, we found that noun usage in general significantly decreased and pronouns increased similarly significantly.
-The increase in pronoun for those with early AD has been identified in literature, although there are only a few studies that explore this \cite{Almor1999, Wendelstein2015}.
-Wendlestein et al propose that the increased used of pronouns is an expression of an impaired ability to adapt language to the listener's needs \cite{Wendelstein2015}.
-Almor et al attributed this reliance on pronouns due to a impaired working memory \cite{Almor1999}.
-\par 
-The decrease in overall noun usage has also been identified as a feature.
-Jarrold et al found that AD patients would use more pronouns, verbs and fewer nouns than controls \cite{Jarrold2014}.
-Wendlestein in their investigations into noun usage found that decreased later on in AD progression and was unaffected in the pre-clinical stages of AD \cite{Wendelstein2014}.
-Our results are supported by existing literature and this potentially means that language analysis in the way we have structured it may have diagnostic or prognostic properties.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=240px, height=150px]{images/RRNounsNormalised.png}
-	\caption{Ronald Reagan - Nouns Normalised over time}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=240px, height=150px]{images/RRPronouns.png}
-	\caption{Ronald Reagan - Pronouns Normalised over time}
-\end{figure}
-
-There are limitations of this research.
-Whilst in terms of age, DJT is certainly more suitable as a control to match with RR, in some ways they held very different styles of press conferences in that RR preferred to do solo press conferences and DJT has shown a preference for doing joint press conferences which have an impact on the amount of language produced.
-This artifact of the data is in itself notable as it illustrates the problems we may have with smaller amounts of speech.
-Also, the problem of finding an appropriate control is a common one in this domain.
-Given that those with MCI and early dementia have such variable presentations, it might prove of limited value in matched pairs design.
-\par 
-With further work, it is not feasible to the vast array of samples over a timeframe, as we have had with the president corpus and so it would be worth exploring how the quality of these predictions might lessen when faced with considerably fewer samples and over a smaller time period.
-It would also be worth extending this research further to encompass more of the linguistic features Fraser used in her work \cite{Fraser2015} to see if there are any further insights to be gained.
-In addition, this replication and extension has demonstrated the potential utility of using longitudinal data as a means of comparing language use of a person at two or more time periods and using this information as a diagnostic aid for MCI and therefore more work would be helpful from a longitudinal perspective to see if this approach may be valid in moving towards a solution for this particular problem.
-\par 
-The results of this work show that we can track a person's use of language through time in a number of ways, and it is possible for an individual to be his or her own control.
-This is important as it means the heterogenous nature of the MCI population does not impact results as much as if we were comparing those with MCI to controls.
-Equally, it would be helpful to have controls to ascertain what would be usual to expect in the decline of language in a healthy older adult.
-\chapter{Future Work}
-\section{Introduction}
-This section looks at my research aims and objectives given the literature review presented above.
-This section will then go on to explore any challenges to my research.
-I then present the work I have planned for the next two years, divided into distinct tasks with their own aims and objectives.
-As part of this I will identify target conferences where I could present my work.
-Finally, I look at potential risks and ethical concerns that could conceivably occur.
-\section{Research Aims and Objectives}
-The aim of my research is to ultimately to identify more accurate and efficient ways of detecting MCI and early dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRIs and EEGs) by utilising spontaneous speech.
-There is a lot of research into the analysis of language as a marker for MCI \cite{Taler2008, Boschi2017} and there have been calls for research that explores the use of Natural Language Processing and Machine Learning techniques to aid in the analysis of language \cite{Boschi2017}.
-This will be the focus of my research.
-\par
-I have deliberately selected the MCI population as examining language decline at this stage has the potential to further our understanding of how MCI progresses into dementia.
-Given that sampling a person's language is relatively effortless, my proposed research will look at whether we can find markers of MCI in natural language.
-I plan to explore the use of Natural Language Processing and Machine Learning techniques to assist in this process.
-There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia.
-As documented above, these studies have identified a number of speech and language differences between healthy controls and those with dementia.
-Some of these features have been incorporated in work using machine learning to classify healthy controls from those with dementia.
-However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'.
-While this is helpful to establish that machine learning can be used to differentiate between healthy controls and those with dementia, there has been little work looking at those in the MCI phase although it is apparent that language is still affected \cite{Berisha2015, Snowdon1996, Le2011}.
-Therefore we would like to clarify whether natural language generated outside the framework of a deliberate language generating task, has the same ability to discriminate.
-\par
-Machine Learning itself is undergoing a shift in focus at the present time.
-I will describe this shift using the context of my proposed research.
-In traditional machine learning, the transcripts from cognitive assessments are used to generate abstract features such as idea density and more concrete features such as mean length of sentence or number of unique words.
-These features are then used to classify patients.
-As we have seen above, the picture with regards to what features deem important in this area is somewhat muddy.
-Machine Learning has a number of tools at it's disposal to aid this.
-The first is to statistically determine, given a model, which features contribute most highly to a given prediction.
-The second is to use a deep learning model which has the ability to automate the process of feature engineering.
-\par 
-The broad objectives of my research are to explore whether the application of Natural Language Processing and Machine Learning techniques to spontaneous speech samples can improve our ability to diagnose Mild Cognitive Impairment in patients that will directly lead to Alzheimer's Disease.
-Another objective of my research is to provide some clarity as to which language features are most important to measure when looking for evidence of MCI and/or AD.
-Finally, an additional outcome of my research is the development of a repeatable battery of neuropsychological tests that specifically focus on being sensitive to Mild Cognitive Impairment.
-\begin{itemize}
-	\item \textbf{(H1)} Monitoring language samples and other measures of cognitive decline over the course of the year will be enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment
-	\item \textbf{(H2)} Those with MCI will show a different language profile to controls if matched similarly for age, gender, education and other significant factors.
-	\item \textbf{(H3)} The new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures such as the RBANS.
-	\item \textbf{(H4)} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, those with Alzheimers Disease and decline in those with MCI.
-	\item \textbf{(H5)} Using individuals as their own control will be enough to identify trends in the decline of their use of language.
-\end{itemize}
-
-\section{Project Plan and Task Definitions}
-In this section, I describe each task in detail along with a projected timeline depicted in a Gantt chart.
-\subsection{Task 2 - Pilot Study for the Neuropsychological Battery of Tests}
-The aim of this study is complete a trial run of the administration of a new battery of neuropsychological battery of tests.
-We aim to find out whether the tests in question generate enough language so that features such as those generated in Task 1 are meaningful and can be used as language markers that can then be tracked over time.
-We also aim to look at whether there are significant practice effects between three versions of these tests with a gap of seven days between each test.
-\par 
-We compiled a series of tests which look at a number of cognitive functions which have previously demonstrated a level of effectiveness in discriminating an MCI population from a healthy population (see literature review for a comprehensive list, explanation and rationale for each test).
-In addition to these tests, we have incorporated two language tasks in which a participant's ability to generate spontaneous speech that we can use to further analyse.
-As a planned future study is longitudinal, it is important that this battery of tests is repeatable but does not have significant practice effects that may interfere with our ability to track cognitive decline.
-\par 
-We have recruited a cohort of 4 volunteers who will complete the battery of neuropsychological tests on three separate days with a gap of at least seven days between each administration.
-The results of which will compiled and analysed.
-Our hypothesis is that there would be no difference in performance between one battery of tests and another.
-An additional hypothesis is that the language tasks included on each test would be enough to collect an appropriate sample for linguistic analysis such that we may be able to detect cognitive decline in a clinical population.
-\par 
-While the pilot study in itself is not valid for submission to a conference or journal, it is hoped that the results of the study will allow subsequent studies using the same or adjusted battery of tests to be published.
-\subsection{Task 3 - Exploration of language with the DementiaBank dataset and Three Authors dataset.}
-The aim of this task is to replicate and extend the work of Fraser and Orimaye who both used the DementiaBank dataset and set out to use traditional machine learning techniques to see whether they could categorize people into dementia or healthy categories using the samples given.
-We also aim to replicate and extend the work of Le et al who used electronic versions of the novels of three authors.
-Both datasets are longitudinal and have been shown to track language decline over time in different ways.
-\par 
-The DementiaBank dataset was a product of a longitudinal study which was run by the University of Pittsburgh School of Medicine.
-The dataset contains transcripts of interviews conducted with participants with MCI / AD and other forms of dementia.
-These interviews were conducted in English and used the cookie-theft picture language task (see Literature Review for a full description).
-The full data set consists of 314 patients classified as having dementia and 242 patients who are classified as healthy and these patients are tested on a minimum of two separate occasions.
-\par 
-The Three Authors dataset includes the novels of P.D.
-James, Agatha Christie and Iris Murdoch.
-It contains twenty of Murdoch's novels which were published between the ages of 35 and 76, sixteen of Christie's novels written between the ages of 28 and 82 and fifteen of P.D.
-James novels written between the ages of 42 and 82.
-These three authors are notable as Murdoch had a confirmed diagnosis of Dementia after her death, Christie had a suspected diagnosis of Dementia and P.D.
-James was used as a control.
-\par 
-These two existing datasets in addition to the President's Press Conference corpus described above will provide three sources of validation of features that we generate as well as give us clues as how to best process the data we receive from our longitudinal study.
-\par 
-The final output of this task will be a paper which looks at the importance of various word-level and sentence-level features in the deterioration of language for people with AD.
-The target conference for this work is the UK Alzheimer's Research conference in March 2020 with an expected Abstract submission date of December 2019 and the Student Research Workshop of the ACL in July 2020 with an expected submission date of April 2020.
-\subsection{Task 4 - Longditunal Study}
-Previously we have looked at historical datasets, the aim of this task is see whether a formalised collection of specific data, namely our language tasks are enough to be able to be able to detect cognitive decline in participants at different stages of the MCI / Alzheimer's disease progression.
-To this end, we plan on administering our neuropsychological battery of tests from Task 2 to each participant three times (six monthly intervals) over the course of the year.
-\par 
-We therefore plan to recruit a set of 80 participants divided evenly into four groups.
-These groups are Healthy Older Controls (HOC), Early Alzheimers (EA), Subjective Mild Cognitive Impairment (SMCI) and Diagnosed Mild Cognitive Impairment (DMCI).
-We expect to see typical age related decline in our HOC group and more dramatic decline in our EA group.
-We are most interested to see if there is any measurable language decline in the SMCI or DMCI experimental groups.
-\par 
-Whilst we have divided participants into experimental groups we aim to investigate each participant individually.
-Because there is a large amount of heterogeneity in this population, we will initially be analysing each participant's results in isolation with the participant acting as their own control.
-However, we will also be exploring the use of traditional machine learning and/or deep learning across cohorts to investigate whether each cohort has their own distinct 'language profile'.
-\par 
-The final output of this task will be a paper discussing the results of our longitudinal study from a language analysis perspective.
-The target conference for this work will be the UK Alzheimer's Research conference in March 2021 with an expected submission date of December 2020, the Alzheimer's Association International Conference in July 2021 with an expected submission date of January 2021 and the Association of Computational Linguistics Conference (ACL) in July 2021 with an expected submission date of March 2021.
-\subsection{Task 5 - Analysis of performance and validation of the proposed Repeatable Neuropsychological Battery of Tests}  
-In this task, I will be looking at the results of the neuropsychological battery of tests without the additional language tasks to determine whether we can detect cognitive decline. 
-\par 
-We expect to see no decline in our HOC group within this timeframe and more severe decline in our EA group.
-Once again, we will be interested in whether there is any measurable decline in our SMCI or DMCI experimental groups with our battery of tests.
-\par 
-The final output of this task will be a paper discussing the results of the battery of tests, including how the battery does in terms of diagnostic and prognostic utility.
-The target conference for this work is to be decided.
-
-\subsection{Task 6 - Thesis Writing}
-The writing of the thesis, including additional over-running data analysis is planned for the last six months of the project (Oct '20 to Mar '21).
-
-% GANTT CHART
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=350px, height=200px]{images/TaskList.png}
-	\caption{Task List}
-\end{figure}
-
-\begin{sidewaysfigure}[ht]
-	\includegraphics[width=\textwidth]{images/GanttChart.png}
-	\caption{Proposed Timetable.}
-\end{sidewaysfigure}
-
-\section{Research Challenges}
-\subsection{Data Collection}
-There are a few challenges that I have identified with the process of data collection.
-As with all longitudinal studies, the data collected is only valid if there are at least two sets of data points for a given participant.
-Given the population I am exploring, it is expected that there is some level of drop out rate over the course of the year and this has been accounted for.
-Nevertheless, a higher than average drop out rate would weaken the strength of the study, especially if those who dropped out came from the experimental group.
-\subsection{Commonality of MCI}
-Whilst research has consistently shown that those with a diagnosis of MCI tend to have a significantly higher risk of progression to dementia \cite{Roberts2014}.
-The progression at any given time is low and variable.
-Annual conversion rates vary across longitudinal studies conducted in this population, but researchers have shown that there is a higher conversion rate in clinical samples of this population than in community samples \cite{Farias2009}.
-In my proposed longitudinal study, I aim to collect a clinical MCI population, but depending on the availability of such samples I may have to collect from the community.
-If this is the case then, the prevalence of those with MCI who have language deficits within my participants may be low.
-\subsection{Sparsity of Language}
-I have tried to estimate the amount of language I will collect by conducting a pilot study on healthy controls.
-However, until the first sets of data collection with the MCI population, I will not be able to tell if this is a problem.
-I fully expect there to be little difference in the amount of language collected in the MCI population, but this remains a possibility which I will need to adjust for at the 2nd and 3rd data collection phases.
-
-\section{Risk Assessment}
-In this section, I look at the potential difficulties that may produce a delay to the proposed research, define any strategies that have already been put in place to account for this as well as define any other potential solutions.
-\subsection{Delay in data collection}
-When working with a clinical population, the collection of data can be subject to delays.
-This can be for a number of reasons including but not limited to sickness (patient or researcher), unavailability of participant for another reason or unavailability of venue to collect data.
-The collection of data could also be subject to delays due to ethical approval for the research being delayed.
-For individual participants unavailability, flexibility has been built into the data collection schedule to allow for an extra month (capacity for one additional cohort) of data collection per collection cycle.
-\subsection{Rejected Publications}
-For each task defined above, with the exception of task 2 and task 6 I have identified suitable conferences at which I can present my work.
-With the target conferences identified every effort will be made to meet the submission deadlines for these conferences.
-If submissions are rejected at these conferences, I have identified alternative conferences to present at.
-\subsection{Rejected NHS Ethics and/or University Ethics}
-Every effort will be made to submit a complete ethics application through the NHS.
-However, in the event that there is a delay in the ethical approval process then data collection will proceed with a priority placed on the control group.
-Future cohorts will then have proportionally more clinical participants until the research meets it's target numbers.
-In the event that NHS ethics is not granted in time for me to write up, participants will be sought from other areas such as the Alzheimer's Research pool of volunteers and venues where suitable candidates may be present such as meetings of the University of the Third age.
-
-\section{Ethical Considerations}
-In this section I discuss the ethical considerations related to this project and any adjustments that I have made to my research because of these considerations.
-I am aware of the code of human  research ethics published by the British Psychological Society \cite{BritishPsychologicalSociety2014} and my research has been designed in a way that fits with these guidelines.
-\subsection{Risk}
-I will be working with a middle-aged to elderly population who have identified in some way that there is a difficulty with cognition.
-However, I do not anticipate working with vulnerable groups, those lacking capacity or individuals in a dependent or unequal relationship.
-Should any participants I am working with lose capacity during the course of the study, I will remove that participants details and data from the study.
-\subsection{Informed Consent}
-Consent will be sought overtly at the recruitment and initial data collection point by way of an information sheet and associated consent form.
-At other data collection points, I will have information sheets available if participants request them and if such a request is made then I will seek consent again at this stage.
-As part of this process, I will be supplying my participants with my contact details should they have any question or for the purposes of withdrawing consent.
-These consent forms will be stored in a secure, locked cabinet at the university.
-\subsection{Confidentiality}
-Given the nature of the study and the possibility of disclosures of suicidal ideation, it is important that confidentiality is discussed at each data collection point.
-I will also provide an information sheet which details the confidentiality statement and also the situations in which I may need to break confidentiality.
-In addition, I will make a verbal confidentiality statement at each collection point and seek acknowledgement of their understanding and permission to continue.
-The spoken confidentiality statement will be as follows.
-\begin{center}
-	\textit{"Anything you say to me today will be confidential unless I feel that there is a risk to you, or to someone else.
-If this is the case, I may need to speak to or write to third parties such as your GP.
-I will endeavor to discuss this communications with yourself where it is possible to do so.
-Is this ok?"}
-\end{center}
-\subsection{Disclosure of suicidal ideation or other risks}
-As part of the battery of tests for this experiment, there is an opportunity for a participant to reveal that he or she has had suicidal ideation in the past two weeks (Patient Health Questionnaire, Question 9).
-This is something to explore further.
-As I have been trained in psychological risk assessments, it seems appropriate to carry out a further risk assessment.
-Optionally with the tape recorder turned on, or off as requested.
-Any text recorded during the risk assessment will not be included in the language analysis.
-Following this, some information about local services such as Improving Access to Psychological Therapies (IAPT), Samaritans and other appropriate services will be provided.
-A discussion will then be made about whether it is appropriate to notify the GP of the disclosure.
-In the event that there is imminent risk to self or others, a discussion will be made about who I need to contact.
-(Crisis Team and GP).
-In the event that the participant scores 0 on Q9 of the PHQ, no risk assessment will be undertaken.
-
-\section{Thesis - Provisional Table of Contents}
-In this section I provide a provisional table of contents based on the work I have either completed or proposed above.
-\begin{enumerate}
-	\item Introduction
-	\begin{enumerate}
-		\item Overview
-		\item Background and Context
-		\item Aims and Objectives
-		\item Challenges and Achievements
-	\end{enumerate}
-	\item Literature Review
-	\begin{enumerate}
-		\item Introduction
-		\item Assessing Mild Cognitive Impairment using cognitive tests.
-		\item Assessing Mild Cognitive Impairment using Language.
-		\item Natural Language Processing and Machine Learning
-		\item Conclusion
-	\end{enumerate}
-	\item Further Exploration of the Presidents Corpus
-	\begin{enumerate}
-		\item Introduction
-		\item Methodology
-		\item Results
-		\item Discussion
-		\item Conclusion
-	\end{enumerate}
-	\item Validation of a Repeatable Neuropsychological Battery of Tests
-	\begin{enumerate}
-		\item Introduction
-		\item Methodology
-		\item Pilot Study Results
-		\item Longitundinal Study results
-		\item Discussion
-		\item Conclusion
-	\end{enumerate}
-	\item Longitudnal Study
-	\begin{enumerate}
-		\item Introduction
-		\item Methodology
-		\item Experiment 1 Results
-		\item Experiment 2 Results
-		\item Discussion
-		\item Conclusion
-	\end{enumerate}
-	\item General Discussion
-	\begin{enumerate}
-		\item Discussion
-		\item Future Work
-		\item Conclusion
-	\end{enumerate}
-\end{enumerate}
-
-\chapter{Conclusion}
+
+\chapter{Feature Engineering and Classification of DementiaBank Transcripts and Audio}
+
+\chapter{Comparison of different speech to language applications}
+
+\chapter{Development of an application}
+
+\chapter{Techniques in feature engineering - What features matter? Comparing language generated from picture description task to language generated from our task}
+
+\chapter{Conclusions and Future Work}
 With the older population of the world continuing to grow year upon year, Dementia in all it's forms is also going to continue to grow.
@@
 
-\chapter{Research Questions and Hypotheses}
-The broad question is can we use Machine Learning to assist in the diagnosis of dementia through natural language.
-There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia.
-These
-
-\chapter{Experiment: Presidential Debates revisited}
-\section{Background}
-Aim: To replicate and extend the work by Berisha and Liss.
-To use traditional machine learning techniques to test whether they can differentiate between Reagan, Bush and Trump using similar features.
-To test to see whether Deep Learning can also differentiate between the three.
-\newline
-\section{Data}
-Dataset: the presidential press conferences given by Presidents Ronald Reagan, George H.W. Bush and Donald Trump. \newline 
-Paper: This experiment replicates work done by Berisha and Liss and extends this by adding Donald Trump as an alternative comparison to Ronald Reagan.
-This experiment will look at the features originally recommended by Berisha and Liss, as well as any others that have potential as discussed in the literature review above.
-\newline
-\chapter{Experiment: Three Authors - Murdoch, Christie and James}
-\section{Background}
-\section{Data}
-Aim: To replicate the work by Le et al.
-To use traditional machine learning techniques to see if it can differentiate between the three authors using similar features.
-To test to see whether deep learning can also differentiate between the three.
-To add 'pre-trained' layers from Presidential Debates experiment to see if this improves accuracy.
-\newline
-Dataset is the literary novels of Iris Murdoch, Agatha Christie and P.D. James. \newline
-Paper: 
-\chapter{Experiment: Using the DementiaBank Corpus}
-Aim: To replicate and extend the work of Orimaye et al.
-To use traditional machine learning techniques to see if it can categorise people into dementia or healthy categories using the samples given.
-To see whether can use pre-trained layer from presidential debates to improve accuracy.
-It doesn't seem relevant to use pretrained layer from Experiment four, but it might be worth looking at the results of this.
-\newline
-Dataset: The DementiaBank Corpus. \newline
-Paper: To replicate and extend the work of Orimaye et al. To compare traditional machine learning techniques (already completed by Orimaye) and deep learning. \newline
-\chapter{Experiment: Longitudinal Data - Current Cohort}
-Aim: \newline
-Dataset: To be created.
-\subsection{Dataset creation}
-The aim is to recruit a set of 50 participants with MCI / Early Dementia (diagnosed or undiagnosed) and a set of 50 controls with a similar age and broadly similar educational background (a known potential confound in assessing cognitive decline).
-The rationale for recruiting this number is that there will be an expected drop out rate and so
-\subsubsection{Patient Health Questionnaire (PHQ-9)}
-\subsubsection{Free Cued Selective Reminding Test (FCSRT)}
-\subsubsection{Mini Mental State Examination(MMSE)}
-\subsubsection{Written Description (Cookie Theft Picture)}
-\subsubsection{Semi-Structured Interview}
-
-Paper:
-
-\subsection{Ethics} 
-Need to discuss confidentiality right at the start of each conversation and to reiterate at every meeting.
-I will also provide an information sheet which details the confidentiality statement.
-The spoken confidentiality statement will be as follows.
-"Anything you say to me today will be confidential unless I feel that there is a risk to you, or someone else.
-If this is the case, I may need to speak to or write to third parties such as your GP.
-Is this ok?" \newline
-\par
-\subsection{What if someone reveals suicidal ideation or other risks?}
-As part of the battery of tests for this experiment, there is an opportunity for a participant to reveal that he or she has had suicidal ideation in the past two weeks (Patient Health Questionnaire, Question 9).
-This is something to explore further.
-As I have been trained in psychological risk assessments, it seems appropriate to carry out a further risk assessment.
-Optionally with the tape recorder turned on, or off as requested.
-Any text recorded during the risk assessment will not be included in the language analysis.
-Following this, some information about local services such as Improving Access to Psychological Therapies (IAPT), Samaritans and other appropriate services will be provided.
-A discussion will then be made about whether it is appropriate to notify the GP of the disclosure.
-In the event that there is imminent risk to self or others, a discussion will be made about who I need to contact.
-(Crisis Team and GP).
-In the event that the participant scores 0 on Q9 of the PHQ, no risk assessment will be undertaken.
-\newline
-\subsection{What happens at the end of the data collection phase if there is a suspected decline in a participant's cognitive function}
-
-\chapter{Overall results and Discussion}
-\chapter{Conclusions and Future work}
-
 \bibliographystyle{unsrt}

### 2019-04-02 Source/Systematic Review/Systematic Review.tex
added Systematic Review

\title{Systematic Review}
\author{
	Jomar Alcantara \\
	Department of Computer Science\\
	School of Engineering and Applied Sciences\\
	Aston University
	\and
	Peter Sawyer \\
	Department of Computer Science\\
	School of Engineering and Applied Sciences\\
	Aston University
	\and
	George Vogiatzis\\
	Department of Computer Science\\
	School of Engineering and Applied Sciences\\
	Aston University}

\date{\today}

\documentclass[12pt]{article}

\begin{document}
\maketitle

\begin{abstract}
This is the paper's abstract \ldots
\end{abstract}

\section{Introduction}
This is time for all good men to come to the aid of their party!

\paragraph{Outline}
The remainder of this article is organized as follows.
Section~\ref{previous work} gives account of previous work.
Our new and exciting results are described in Section~\ref{results}.
Finally, Section~\ref{conclusions} gives the conclusions.

\section{Previous work}\label{previous work}
A much longer \LaTeXe{} example was written by Gil~\cite{Gil:02}.

\section{Results}\label{results}
In this section we describe the results.

\section{Conclusions}\label{conclusions}
We worked hard, and achieved very little.

\bibliographystyle{abbrv}
\bibliography{main}

\end{document}
This is never printed

### 2019-04-04 Source/Systematic Review/Systematic Review - Protocol.tex
added Systematic Review

\documentclass[12pt]{article}
\begin{document}
% https://github.com/anadallora/SNAC/blob/master/protocol%20v4.pdf
\section{Introduction}
\section{Research Protocol}
\subsection{Research Question}
\subsection{Search Process}
\subsection{Primary Study Selection}
\subsection{Study Quality Assessment}
\subsection{Data Extraction}
\subsection{Data Synthesis and Reporting}
\end{document}

### 2019-04-04 Source/Systematic Review/Systematic Review.tex
added Systematic Review

@@
 
+\usepackage[utf8]{inputenc}
+\usepackage{tikz}
+\usetikzlibrary{shapes.geometric, arrows}
+
+\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
+\tikzstyle{io} = [trapezium, trapezium left angle=70, trapezium right angle=110, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=blue!30]
+\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, text width=3cm, draw=black, fill=orange!30]
+\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]
+\tikzstyle{arrow} = [thick,->,>=stealth]
+
 \begin{document}
@@
 \section{Introduction}
-This is time for all good men to come to the aid of their party!
+Dementia has been identified as one of those fast growing difficulties facing the world.
+A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
+\par
+In 2009, the Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
+A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
+As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages \cite{Prince2015} and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
+At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017}.
+I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
+\par
+There is a large body of research that looks at the decline in language in those with MCI and AD \cite{Taler2008, Boschi2017}.
+However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD \cite{Taler2008, Boschi2017}.
+Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
+Another area of focus for research of this nature is the process of collecting appropriate language samples.
+Whilst collecting samples of language is comparatively unintrusive, researchers recognise that these samples require a rich sample of language that potentially cannot be generated by tasks such as the picture description task.
+Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD.
+There is some evidence to support this.
+Berisha et al \cite{Berisha2015}, has shown through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control.
+\par
+The potential impact of this research in this area is immense.
+Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease.
+Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
 
-\paragraph{Outline}
+The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to to time-consuming and expensive psychological tests.
+There is a lot of research into the analysis of language as a bio-marker for MCI and Early Dementia.
+Given that sampling a person's language is relatively effortless, my research looks at whether we can find bio-markers of MCI and Early Dementia in natural language.\newline
+Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
+I'm not going to control for memory problems as a potential confound, but does this weaken the research?
+How do I defend this?
+\newline
+\par 
+Introduction to the problem of dementia in the context of the wider world including quality of life and financial implications.
+Exploration of dementia as a syndrome rather than a disease, and a look at the different variants of dementia.
+A look at the rationale behind research into the early diagnosis of dementia as well as a brief look at what has been done in the area (wide context, so pharmacological and psychological).
+\newline
+
+The question to be addressed in this systematic review is how has the field of machine learning and natural language processing addresses language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease.
+
 The remainder of this article is organized as follows.
-Section~\ref{previous work} gives account of previous work.
+Section~\ref{methodology} gives an account of the process of this Systematic Review.
 Our new and exciting results are described in Section~\ref{results}.
+We discuss the results and implications in Section~\ref{discussion}.
 Finally, Section~\ref{conclusions} gives the conclusions.
 
-\section{Previous work}\label{previous work}
-A much longer \LaTeXe{} example was written by Gil~\cite{Gil:02}.
+\section{Methodology}\label{methodology}
+
+A systematic literature review (SLR) describes a process which aims to identify, evaluate and interpret the research and literature in a given area.
+They are designed to provide a complete and exhaustive summary of the current evidence relevant to an identified research question.
+SLR's conduct a thorough search of all literature following a pre-defined protocol that specifies focused research questions, identifies criteria for the selection of studies and assessment of their quality, and forms to execute the data extraction and synthesis of results.
+\par 
+Common motivations for conducting a SLR are: 
+\begin{enumerate}
+	\item to summary all the evidence about a topic.
+	\item find gaps in the research. 
+	\item to provide a ground for a fundament to new research.
+	\item and to examine how the current research supports a hypothesis. 
+\end{enumerate}
+
+Performing a SLR comprises the following steps: 
+\begin{enumerate}
+	\item (i) identify the need for performing the SLR.
+	\item (ii) formulate research questions.
+	\item (iii) execute a comprehensive search and selection of primary studies.
+	\item (iv) assess the quality and extract data from the studies.
+	\item (v) interpret the results.
+	\item (vi) report the SLR.
+\end{enumerate}
+
+The main research question this SLR aims to address is: “How has the field of machine learning and natural language processing addresses language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease?”.
+This main question was decomposed further into five research questions:
+\begin{enumerate}
+	\item RQ1: Which ML and MS techniques are being used in the dementia and comorbidities research?
+	\item RQ2: What data characteristics (variables, determinants and indicators) are being considered when applying the ML or/and MS techniques (physiological, demographic/social, genetics, lifestyle etc)?
+	\item RQ3: What are the goals of the studies that employ ML or MS techniques for prognosis of dementia and comorbidities?
+	\item RQ4: How is data censoring being handled in the studies?
+	\item RQ5: Do the studies focus on individuals or populations?
+\end{enumerate}
+
+The present paper builds upon these questions and additionally presents the results of the other two additional research questions.
+Further, the key terms related to comorbidities were included in the search string to ensure that relevant studies about ML or MS for the prognosis of a disease, where dementia is considered a comorbidity to that disease would also be retrieved from the database searches, even when the term dementia was not mentioned in the paper’s title or abstract.
+
+\subsection{Search strategy}
+To address the research questions, a search string was defined using the PICO approach, which decomposes the main question into four parts: population, intervention, comparison and outcome.
+The comparison component was discarded because the SLR was mainly concerned with a characterization.
+For each of the remaining components, keywords were derived and their rationale can be represented as follows:
+\begin{enumerate}
+	\item Population: Studies that present research on dementia and comorbidities.
+Dementia’s keywords were selected from the “Systematized Nomenclature of Medicine–Clinical Terms” and selected by A4.
+Comorbidities’ keywords were extracted from the Marengoni et al.
+SLR in this topic.
+	\item Intervention: ML or MS techniques.
+The ML keywords were selected from the branch “Machine Learning Approaches” of the “2012 ACM Computing Classification System”.
+The MS keywords were selected by A2.
+	\item Outcome: Prognosis on dementia and comorbidies. The prognosis keywords were provided
+by A4.
+\end{enumerate}
+
+The automated searches were performed in the Pubmed, Web of Science and Scopus databases.
+Table 1 shows the search string used for the Pubmed automated search, but note that this search string was adapted to each of the other databases’ search context.
+
+\begin{table}
+	\begin{tabular}{ p{12cm} }
+	\hline
+	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+	Searched - 4th April 2019 - Generated 1257 Articles \\
+	\hline
+	\end{tabular}
+	\caption[Table caption text]{Search Terms for Web of Science database}
+	\label{table:name}
+\end{table}
+
+\begin{table}
+	\begin{tabular}{ p{12cm} }
+	\hline
+	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+	Searched - 4th April 2019 - Generated 1257 Articles \\
+	\hline
+	\end{tabular}
+	\caption[Table caption text]{Search Terms for Scopus database}
+	\label{table:name}
+\end{table}
+
+\begin{table}
+	\begin{tabular}{ p{12cm} }
+	\hline
+	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+	Searched - 4th April 2019 - Generated 1257 Articles \\
+	\hline
+	\end{tabular}
+	\caption[Table caption text]{Search Terms for ProQuest database}
+	\label{table:name}
+\end{table}
+
+\begin{table}
+	\begin{tabular}{ p{12cm} }
+	\hline
+	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+	Searched - 4th April 2019 - Generated 1257 Articles \\
+	\hline
+	\end{tabular}
+	\caption[Table caption text]{Search Terms for IEEE Xplore database}
+	\label{table:name}
+\end{table}
+
+\subsection{Study selection}
+The first step of the study selection was the execution of an evaluation round with 100 random papers from the 593 results returned from the automated searches.
+These had their title and abstracts assessed by A1, A2 and A3, according to inclusion and exclusion criteria defined previously in the protocol (see Table 2).
+This step was mainly concerned in maintaining the consistency of the selection between the participants throughout the SLR.
+The remaining 493 results had their title and abstracts assessed by A1 and A2, according to the inclusion and exclusion criteria.
+After the evaluations, 37 papers were selected.
+Then a one-iteration backward snowballing was carried out looking for possible additional studies.
+The 1199 new identified studies were assessed analogously as the previous ones, resulting in 41 new selected papers.
+Throughout the whole selection process, A3 and A4 acted in conflict resolution in the case where A1 and A2 couldn’t reach an agreement.
+\par 
+In total, 78 papers were selected to be fully read and assessed regarding its eligibility.
+The ones that successfully passed the established criteria previously defined in the protocol, had their relevant data extracted.
+In order to minimize the chance of selecting studies with bias evidence, a quality assessment questionnaire was used.
+This questionnaire was adapted from Kitchenham’s guidelines and can be found in the SLR protocol.
+If the grading attributed to a paper fell below 8 points (out of a total of 12), it would be rejected for quality reasons.
+The 8-point threshold was
+decided in the research group discussions involving all the authors.
+In this phase, a paper could also be rejected due to inclusion and exclusion criteria because the selection process adopted an inclusive approach.
+This means that during the reading of the titles and abstracts, in the case where the information provided was incomplete or too general it was selected to be fully read in the posterior phase.
+A common example is the case when the data analysis technique specified in the abstract was merely “classification”, so it was not possible to know if any machine learning occurred.
+\par
+As in the study selection, a quality assessment evaluation round was performed beforehand to ensure consistency in the evaluations.
+A1, A2 and A5 participated in this task.
+In total, 37 studies composed the final set of included primary studies and had their relevant data extracted, 7 papers were rejected due quality reasons, and 34 papers were rejected due to failing the inclusion and exclusion criteria.
+One reason for the high number of the latter was the decision to exclude the papers that used solely statistical methods as data analysis tech- niques to build the prognostic models.
+The selected studies were also assessed for the risk of cumulative evidence bias.
+This was done by checking, in the case of the same research group with different studies in the final set of included primary studies, if it was justified having both studies (i.e different samples).
+
+\subsection{Data collection}
+For the data collection, a base extraction form was defined in the protocol, but later in the study it was evolved based on the research group discussions.
+Table 3 lists and defines the collected variables.
+
+\begin{table}
+	\begin{tabular}{ p{6cm} | p{6cm} }
+	\hline
+	Inclusion Criteria & Exclusion Criteria \\ \hline
+	Be a primary study in English; AND address research on dementia and comorbidities; AND address at least one ML or MS technique; AND address a prognosis related to dementia and comorbidities. &
+	Be a secondary or tertiary study; OR be written in another language other than English; OR do not address a research on dementia and comorbidities; OR do not address at least one ML or MS technique; OR do not address a prognosis related to dementia and comorbidities.
+\\
+	\hline 
+	\end{tabular}
+	\caption[Table caption text]{Inclusion and Exclusion Criteria}
+	\label{table:name}
+\end{table}
+
+\begin{table}
+	\begin{tabular}{ p{6cm} | p{6cm} }
+	\hline
+	Variable & Definition \\ \hline
+	Conditions Studied & For which dementia disorder is the study deriving a prognosis. \\ \hline
+	Database used in the study & Name and origin of the data source used to derive the prognosis of the studied dementia. \\ \hline
+	Dataset Categories & Classes in which the data units were divided into. \\ \hline
+	Handling of Censored Data & Description of the way in which censored data was handled. \\ \hline
+	Follow-up period & Period of time, which the data units were followed. \\ \hline
+	Data Analysis Techniques & ML or MS techniques that were used to build the prognostic models. \\ \hline 
+	Model Variables & The variables used in building the prognostic models. \\ \hline
+	Aim of the Study & The goal of the built prognostic models. \\ \hline 
+	Focus of the Study & If the built prognostic models aim its predictions on an individual or population level. \\
+	\hline 
+	\end{tabular}
+
+\end{table}
+
+In addition to these variables other basic data about the studies was collected, these were: title, authors, journal/source, year and type of publication.
+No summary measures were used.
+Summary tables were used for the synthesis of results and no additional analyses were carried out
 
@@
 In this section we describe the results.
+
+\begin{figure}
+
+\begin{tikzpicture}[node distance=2cm]
+
+\node (start) [startstop] {Start};
+\node (in1) [io, below of=start] {Input};
+\node (pro1) [process, below of=in1] {Process 1};
+\node (dec1) [decision, below of=pro1, yshift=-0.5cm] {Decision 1};
+\node (pro2a) [process, below of=dec1, yshift=-0.5cm] {Process 2a text text text text text text text text text text};
+\node (pro2b) [process, right of=dec1, xshift=2cm] {Process 2b};
+\node (out1) [io, below of=pro2a] {Output};
+\node (stop) [startstop, below of=out1] {Stop};
+
+\draw [arrow] (start) -- (in1);
+\draw [arrow] (in1) -- (pro1);
+\draw [arrow] (pro1) -- (dec1);
+\draw [arrow] (dec1) -- node[anchor=east] {yes} (pro2a);
+\draw [arrow] (dec1) -- node[anchor=south] {no} (pro2b);
+\draw [arrow] (pro2b) |- (pro1);
+\draw [arrow] (pro2a) -- (out1);
+\draw [arrow] (out1) -- (stop);
+\end{tikzpicture}
+\caption[Figure caption text]{PRISMA flow chart.}
+\label{figure:name}
+\end{figure}
+\subsection{Language Analysis of Speakers with Dementia of the Alzheimer's Type - Guinn and Habash (2012)}
+
+\subsection{Features and Machine Learning Classification of Connected Speech Samples from patients with Autopsy Proven Alzheimer's Disease with and without additional Vascular Pathology - Rentoumi, Raoufian, Ahmed, de Jager, Garrard (2014)}
+
+\subsection{Learning Predictive Linguistic Features for Alzheimer's Disease and related Dementias using Verbal Utterances - Orimaye, Wong and Golden (2014)}
+
+\subsection{Learning Lingustic Biomarkers for Predicting Mild Cognitive Impairment using Compound Skip-grams - Orimaye, Wong, Wong and Tai (2015)}
+
+\subsection{Diagnosing People with Dementia using Automatic Conversation Analysis - Mirheidari, Reuber, Blackburn, Walker (2016)}
+
+\subsection{Formulaic Language in People with Probably Alzheimer's Disease: A Frequency-Based Approach - Zimmerer, Wibrow and Varley (2016)}
+
+\subsection{Linguistic Features Identify Alzheimer's Disease in Narrative Speech - Fraser, Melzer and Rudzicz (2016}
+
+\subsection{Declines in Connected Language Are Associated with Very Early Mild Cognitive Impairment: Results from the Wisconsin Registry for Alzheimer's Prevention - Mueller, Koscik, Hermann, Johnson and Turkstra (2018)}
+
+\subsection{Deep language space neural network for classifying mild cognitive impairment and Alzheimer-type dementia - Orimaye, Wong and Wong (2018)}
+In this paper, Orimaye et al use deep-deep neural networks language models (D2NNLM) to learn linguistic changes that distinguish the language of patients with MCI and AD-type dementia from the healthy controls using higher order n-grams.
+An ordinary DNNLM uses lower order n-gram N-dimensional sparse vectors as descrete feature representations to train the neural network with multiple hidden layers.
+
+\subsubsection{Deep Neural Network Language Models}
+
+\section{Discussion}\label{discussion}
 
@@
 \end{document}
-This is never printed

### 2019-04-18 Source/Thesis/Response.tex
Response.tex updated

--- 2019-03-14 Source/Thesis/Response.tex
@@
 We do not think the latter are necessary because the focus of the Ph.D is on whether you can detect decline by differences in language use over time and not about determining the clinical underpinnings of those differences.
+\end{quote}
 
-\section{Research Aims and Questions}
-The broad objectives of my research are to explore whether the application of Natural Language Processing and Machine Learning techniques to spontaneous speech samples can improve our ability to diagnose Mild Cognitive Impairment in patients that will directly lead to Alzheimer's Disease.
-Another objective of my research is to provide some clarity as to which language features are most important to measure when looking for evidence of MCI and/or AD.
-\begin{itemize}
-	\item \textbf{(H1)} Monitoring language samples and other measures of cognitive decline over the course of the year will be enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment
-	\item \textbf{(H2)} Given a number of samples of language over a given time frame, it is possible to detect language decline in participants such that you can differentiate between normal language decline, these with Alzheimers Disease and decline in those with MCI.
-\end{itemize}
+\subsection*{Aims and Objectives}
+The overall objective of my research is to explore whether machine learning and natural language processing can aid in the detection of Mild Cognitive Impairment and Early Alzheimer's through language alone.
+With this in mind, my research has the following research questions.
 
-\section{Data Collection Methodology}
+\begin{enumerate}
+	\item What research has been conducted in this field?
+	\item How can we use technology to best enable the collection of language samples?
+	\item Can we develop a complete pipeline of that processes data efficiently?
+	\item What can machine learning tell us about how language affects us?
+\end{enumerate}
 
-\section{Project Aims}
+\section*{Proposed Methodology}
+\subsection*{Stage 1 - Systematic Review of the NLP and Machine Learning research to this problem}
+The aim of this stage is to explore, in depth, the literature in this area.
+As this is a new field, and NLP and Machine Learning advances comparitively quickly, bias is given to more recent research.
+I aim to convert my existing literature review into a systematic review which is positioned from the Computer Science and Machine Learning perspective.
+Estimated completion, August 2019.
+This will form chapter 3 of my thesis.
+This will also produce a publication.
 
+\subsection*{Stage 2 - Delphi Methodology and developing consensus on how best to collect language samples using technology}
+The aim of this stage is to develop, in conjunction with experts, a convergence of ideas on how best to collect language samples using technology.
+The aim is to provide input from clinicians, technology experts and those using the technology within the homes to find the best protocol, and ideas for how to collect data with the focus on ease of use and non intrusiveness.
+The aim is to recruit no more than 20 participants with a rough guide of 5 clinicians, 5 technology experts and 10 users.
+Estimated completion, December 2019.
+This will form chapter 4 of my thesis.
+
+\subsection*{Stage 3 - Development of a pipeline that processes language data accurately}
+The aim of this stage is to build a data pipeline, and test the different technologies such as automated speech recognition software that might have an impact on the quality of the features generated.
+This is based on research done by Fraser (2015, 2019) who uses off the shelf products (Nuance - Dragon NaturallySpeaking) and open-source technology (Kaldi) as a means of converting speech to text.
+I would like to include tests of state of the art speech recognition from Google speech-to-text, IBM Watson and Amazon transcribe.
+This will also address the challenges unique to this population such as how to deal with changes in topic, hesitations, repetitions and so on.
+This will form chapter 5 of my thesis.
+The aim is to complete this by December 2019.
+
+\subsection*{Stage 4 - Analysis of DementiaBank dataset, an exploration of existing and novel ways of classification of those with MCI and Early AD.}
+The aim of this stage is to explore and identify language features that appear to be more important in the identification and classification of those with MCI and early AD.
+This stage will build upon the systematic review (stage 1) which, amongst other objectives, aims to identify language features other researchers have identified as useful in this problem.
+This stage will replicate and validate previous research results and introduce One Class Classification as a new way of classifying those with MCI and AD.
+This will form chapter 6 of my thesis.
+This should produce a publication.
+The aim is to complete this by March 2020.
+
+\subsubsection*{One Class Classification}
+The prevalence of Dementia in the UK's population is roughly 1 in 14 or 7\%.
+The prevalence of MCI has only briefly been explored, however some research by a team at the Univesity of Leeds indicate that the prevalence of MCI is between 3.2 - 24\%, and found that the prevalence of MCI in people who have self-reported memory problems is almost double (18\%) the amount of do not self report.
+Whilst this means that it is entirely possible to recruit a cohort with MCI and AD, it is comparatively costly in terms of time and resources.
+Given that, from a statistical perspective, this is a highly imbalanced population we can look at the problem as a kind of anomaly detection.
+This has some precedent in the field of Medicine, where examples of the 'positive' class can be costly or difficult to obtain.
+\par 
+One-class classifiers analyze only examples of the majority class (in this case - Healthy Older Adults) to learn a classification boundary that excludes outlier examples.
+When dealing with heavily imbalanced datasets with text classifcation as an objective, the one-class classifier approach has been demonstrated to outperform conventional two-class classification.
+T
+
+\subsection*{Stage 5 - Pilot study of the methodology developed}
+The aim of this stage is to pilot the methodology developed in stage 2.
+Over a period of six months, 10 participants engage in collection of data and we analyse the results from both the participants perspective and the perspective of the researcher.
+From the participants perspective, we look at whether the process is comfortable, unintrusive and easy to use for the participant.
+From the researchers perspective, we look at the quality of the audio and whether there are any problems in the data pipeline.
+We also look at the quantity of language samples and whether the amount of language generated is enough for analysis.
+This will form chapter 7 of my thesis.
+This should produce a publication.
+The aim is to complete this by September 2020.
 \end{document}

### 2019-04-19 Source/Systematic Review/Systematic Review.tex
updated systematic review

@@
 
+\begin{table}
+	\begin{tabular}{ p{3cm} | p{3cm} | p{3cm} | p{3cm} }
+	\hline
+	Scopus & Web of Science & PubMed & IEEE Xplore \\ \hline
+	1002 & 991 & 376 & 230 \\ 
+	\hline
+	\end{tabular}
+	\caption[Table caption text]{Totals from each database}
+	\label{table:name}
+\end{table}
+
 \subsection{Study selection}
-The first step of the study selection was the execution of an evaluation round with 100 random papers from the 593 results returned from the automated searches.
-These had their title and abstracts assessed by A1, A2 and A3, according to inclusion and exclusion criteria defined previously in the protocol (see Table 2).
-This step was mainly concerned in maintaining the consistency of the selection between the participants throughout the SLR.
-The remaining 493 results had their title and abstracts assessed by A1 and A2, according to the inclusion and exclusion criteria.
+A total of 2599 papers were identified through the searches conducted above.
+Papers that were duplicated were removed and Papers were then evaluated by the JA based on the inclusion and exclusion criteria defined previously (see Table 6).
+Where JA could not reach a decision, PS and GV were consulted and majority vote on inclusion was conducted.
+Each paper was initially presented with just the title and the abstract.
+
+\begin{table}
+	\begin{tabular}{ p{6cm} | p{6cm} }
+	\hline
+	Stage & Papers remaining\\ \hline
+	Start & 2599 \\ \hline
+	Duplicates & Non-papers removed & 2599 \\hline 
+	\hline
+	\end{tabular}
+	\caption[Table caption text]{Totals at each stage}
+	\label{table:name}
+\end{table}
+
 After the evaluations, 37 papers were selected.
@@
 
+A search of the literature was conducted using ProQuest (PsychArticles), SCOPUS, Web of Science.
+The following results were found (Table 1).
+All papers were then reviewed for relevance by reading the abstract and full text where appropriate and a shortlist was compiled.
+An additional search through references of shortlisted papers was also conducted and any papers who upon further review appeared relevant were added to the shortlist.
+Papers were included where researchers used machine learning to classify participants as MCI, AD or Healthy using language.
+We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English.
+
 \subsection{Data collection}
@@
 	Inclusion Criteria & Exclusion Criteria \\ \hline
-	Be a primary study in English; AND address research on dementia and comorbidities; AND address at least one ML or MS technique; AND address a prognosis related to dementia and comorbidities. &
+	Be a primary study in English; AND address research on dementia and comorbidities; AND address at least one ML or MS technique; AND address a prognosis related to dementia and comorbidities; AND use cognition or language decline as a factor for analysis.
+&
 	Be a secondary or tertiary study; OR be written in another language other than English; OR do not address a research on dementia and comorbidities; OR do not address at least one ML or MS technique; OR do not address a prognosis related to dementia and comorbidities.
@@
 
-\begin{tikzpicture}[node distance=2cm]
+\begin{tikzpicture}[sibling distance = 10em,
+	every node/.style = {shape = rectangle, rounded corners,
+	draw, align = center, 
+	top color = white, bottom color=blue!20}]]
 
-\node (start) [startstop] {Start};
-\node (in1) [io, below of=start] {Input};
-\node (pro1) [process, below of=in1] {Process 1};
-\node (dec1) [decision, below of=pro1, yshift=-0.5cm] {Decision 1};
-\node (pro2a) [process, below of=dec1, yshift=-0.5cm] {Process 2a text text text text text text text text text text};
-\node (pro2b) [process, right of=dec1, xshift=2cm] {Process 2b};
-\node (out1) [io, below of=pro2a] {Output};
-\node (stop) [startstop, below of=out1] {Stop};
-
-\draw [arrow] (start) -- (in1);
-\draw [arrow] (in1) -- (pro1);
-\draw [arrow] (pro1) -- (dec1);
-\draw [arrow] (dec1) -- node[anchor=east] {yes} (pro2a);
-\draw [arrow] (dec1) -- node[anchor=south] {no} (pro2b);
-\draw [arrow] (pro2b) |- (pro1);
-\draw [arrow] (pro2a) -- (out1);
-\draw [arrow] (out1) -- (stop);
 \end{tikzpicture}

### 2019-05-08 Source/Thesis/Response.tex
edited Response.pdf
added more searches

@@
 \onehalfspacing
+
+\usepackage{graphicx}
+\usepackage{float}
+\usepackage{pdflscape}
 
@@
 \subsection*{Stage 1 - Systematic Review of the NLP and Machine Learning research to this problem}
-The aim of this stage is to explore, in depth, the literature in this area.
-As this is a new field, and NLP and Machine Learning advances comparitively quickly, bias is given to more recent research.
+The aim of this stage is to explore, in depth, literature in this area.
+As this is a new field, and NLP and Machine Learning advances comparatively quickly, bias is given to more recent research.
 I aim to convert my existing literature review into a systematic review which is positioned from the Computer Science and Machine Learning perspective.
@@
 This will form chapter 3 of my thesis.
-This will also produce a publication.
+This will also produce a publication tentatively titled - Automated language analysis and it's application in the early diagnosis of Mild Cognitive Impairment and Alzheimer's Disease - A systematic review.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=400px, height=150px]{images/TasksAndDates.png}
+\caption{Tasks and Dates}
+\end{figure}
 
@@
 Estimated completion, December 2019.
-This will form chapter 4 of my thesis.
+This will form chapter 4 of my thesis and may produce a publication depending on the methodology used.
 
@@
 \par 
-One-class classifiers analyze only examples of the majority class (in this case - Healthy Older Adults) to learn a classification boundary that excludes outlier examples.
-When dealing with heavily imbalanced datasets with text classifcation as an objective, the one-class classifier approach has been demonstrated to outperform conventional two-class classification.
-T
+One-class classifiers analyse only examples of the majority class (in this case - Healthy Older Adults) to learn a classification boundary that excludes outlier examples.
+When dealing with heavily imbalanced datasets with text classification as an objective, the one-class classifier approach has been demonstrated to outperform conventional two-class classification.
 
@@
 Over a period of six months, 10 participants engage in collection of data and we analyse the results from both the participants perspective and the perspective of the researcher.
-From the participants perspective, we look at whether the process is comfortable, unintrusive and easy to use for the participant.
+From the participants perspective, we look at whether the process is comfortable and easy to use for the participant.
 From the researchers perspective, we look at the quality of the audio and whether there are any problems in the data pipeline.
@@
 The aim is to complete this by September 2020.
+
+\begin{landscape}
+\begin{figure}[H]
+\centering
+\includegraphics[width=650px, height=150px]{images/GanttChartUpdated.png}
+\caption{Gantt Chart}
+\end{figure}
+\end{landscape}
+
+\section{Use of additional resources}
+It has been suggested that I potentially have the resource of a placement student from the Psychology department to help with my research.
+If this is still the case then I propose to get help with the following.
+
+\begin{enumerate}
+	\item Focus Groups - Conduct focus groups with various stakeholders as part of stage 2 (the adapted Delphi Methodology).
+	\item Qualitative Analysis - Assist in the Qualitative Analysis of the transcripts / answers to the questions asked of the various stakeholders. 
+	\item Recruitment - Recruitment into stage 5 (Pilot Study)
+	\item Quantitative Analysis - Assist in the Analysis of the results of stage 5 (Pilot Study)
+	\item Test Administration - Administration of any neuropsychological tests during stage 5 (Pilot Study)
+	\item General research support - Responding to queries from stakeholders in stage 2 and pilot study participants in stage 5.
+\end{enumerate}
+
 \end{document}

### 2019-05-08 Source/Systematic Review/Systematic Review - Protocol.tex
added more searches

@@
 \documentclass[12pt]{article}
+\usepackage{graphicx}
+\usepackage{float}
 \begin{document}
+
 % https://github.com/anadallora/SNAC/blob/master/protocol%20v4.pdf
 \section{Introduction}
+Dementia has been identified as one of those fast growing difficulties facing the world.
+A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
+\par
+In 2009, the Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
+A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
+As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages \cite{Prince2015} and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
+At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017}.
+I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
+\par
+There is a large body of research that looks at the decline in language in those with MCI and AD \cite{Taler2008, Boschi2017}.
+However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD \cite{Taler2008, Boschi2017}.
+Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
+Another area of focus for research of this nature is the process of collecting appropriate language samples.
+Whilst collecting samples of language is comparatively unintrusive, researchers recognise that these samples require a rich sample of language that potentially cannot be generated by tasks such as the picture description task.
+Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD.
+There is some evidence to support this.
+Berisha et al \cite{Berisha2015}, has shown through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control.
+\par
+The potential impact of this research in this area is immense.
+Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease.
+Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
+
+The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to to time-consuming and expensive psychological tests.
+There is a lot of research into the analysis of language as a bio-marker for MCI and Early Dementia.
+Given that sampling a person's language is relatively effortless, my research looks at whether we can find bio-markers of MCI and Early Dementia in natural language.\newline
+Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
+I'm not going to control for memory problems as a potential confound, but does this weaken the research?
+How do I defend this?
+\newline
+\par 
+Introduction to the problem of dementia in the context of the wider world including quality of life and financial implications.
+Exploration of dementia as a syndrome rather than a disease, and a look at the different variants of dementia.
+A look at the rationale behind research into the early diagnosis of dementia as well as a brief look at what has been done in the area (wide context, so pharmacological and psychological).
+\newline
+
+The question to be addressed in this systematic review is how has the field of machine learning and natural language processing addresses language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease.
 \section{Research Protocol}
 \subsection{Research Question}
+The main research question this SLR aims to address is: “How has the field of machine learning and natural language processing addresses language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease?”.
+This main question was decomposed further into five research questions:
+\begin{enumerate}
+	\item RQ1: Which ML and MS techniques are being used in the dementia and comorbidities research?
+	\item RQ2: What data characteristics (variables, determinants and indicators) are being considered when applying the ML or/and MS techniques (physiological, demographic/social, genetics, lifestyle etc)?
+	\item RQ3: What are the goals of the studies that employ ML or MS techniques for prognosis of dementia and comorbidities?
+	\item RQ4: How is data censoring being handled in the studies?
+	\item RQ5: Do the studies focus on individuals or populations?
+\end{enumerate}
 \subsection{Search Process}
+
+To address the research questions, a search string was defined using the PICO approach, which decomposes the main question into four parts: population, intervention, comparison and outcome.
+The comparison component was discarded because the SLR was mainly concerned with a characterization.
+For each of the remaining components, keywords were derived and their rationale can be represented as follows:
+\begin{enumerate}
+	\item Population: Studies that present research on dementia and comorbidities.
+Dementia’s keywords were selected from the “Systematized Nomenclature of Medicine–Clinical Terms” and selected by A4.
+Comorbidities’ keywords were extracted from the Marengoni et al.
+SLR in this topic.
+	\item Intervention: ML or MS techniques.
+The ML keywords were selected from the branch “Machine Learning Approaches” of the “2012 ACM Computing Classification System”.
+The MS keywords were selected by A2.
+	\item Outcome: Prognosis on dementia and comorbidies. The prognosis keywords were provided
+by A4.
+\end{enumerate}
+
+The automated searches were performed in the Pubmed, Web of Science and Scopus databases.
+Table 1 shows the search string used for the Pubmed automated search, but note that this search string was adapted to each of the other databases’ search context.
+
+\begin{table}
+	\begin{tabular}{ p{12cm} }
+	\hline
+	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+	Searched - 4th April 2019 - Generated 1257 Articles \\
+	\hline
+	\end{tabular}
+	\caption[Table caption text]{Search Terms for Web of Science database}
+	\label{table:name}
+\end{table}
+
+\begin{table}
+	\begin{tabular}{ p{12cm} }
+	\hline
+	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+	Searched - 4th April 2019 - Generated 1257 Articles \\
+	\hline
+	\end{tabular}
+	\caption[Table caption text]{Search Terms for Scopus database}
+	\label{table:name}
+\end{table}
+
+\begin{table}
+	\begin{tabular}{ p{12cm} }
+	\hline
+	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+	Searched - 4th April 2019 - Generated 1257 Articles \\
+	\hline
+	\end{tabular}
+	\caption[Table caption text]{Search Terms for ProQuest database}
+	\label{table:name}
+\end{table}
+
+\begin{table}
+	\begin{tabular}{ p{12cm} }
+	\hline
+	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+	Searched - 4th April 2019 - Generated 1257 Articles \\
+	\hline
+	\end{tabular}
+	\caption[Table caption text]{Search Terms for IEEE Xplore database}
+	\label{table:name}
+\end{table}
+
+\begin{table}
+	\begin{tabular}{ p{3cm} | p{3cm} | p{3cm} | p{3cm} }
+	\hline
+	Scopus & Web of Science & PubMed & IEEE Xplore \\ \hline
+	1002 & 991 & 376 & 230 \\ 
+	\hline
+	\end{tabular}
+	\caption[Table caption text]{Totals from each database}
+	\label{table:name}
+\end{table}
+
 \subsection{Primary Study Selection}
+Firstly, the results from the automated searches from the three selected sources (Scopus, Pubmed and Web of Science) will have their duplicates removed, and this set of possible primary studies will be assessed by the evaluation of the titles and abstracts according to the defined inclusion and exclusion criteria.
+The result from this is our first selected subset (Selected Subset 1).
+The next phase is to look at the reference lists of papers in Selected Subset 2 and identify any papers missed in the automated searches.
+This is our second selected subset (Selected Subset 2).
+\par
+Each paper in these subsets will be fully read and assessed for it's quality, using the process documented further down.
+Any papers that meet the established criteria will be included as a primary study for this systematic review.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=300px, height=150px]{images/PRISMA - Chart.png}
+\caption{Selection Chart}
+\end{figure}
 \subsection{Study Quality Assessment}

### 2019-05-09 Source/Thesis/Response.tex
updated 09May

@@
 
-\section{Use of additional resources}
+\section*{Use of additional resources}
 It has been suggested that I potentially have the resource of a placement student from the Psychology department to help with my research.

### 2019-05-09 Source/Thesis/Response.tex
updated Systematic Review protocol

@@
 \subsubsection*{One Class Classification}
-The prevalence of Dementia in the UK's population is roughly 1 in 14 or 7\%.
+The prevalence of Dementia in the UK's population in the over 65 population is roughly 1 in 14 or 7\%.
 The prevalence of MCI has only briefly been explored, however some research by a team at the Univesity of Leeds indicate that the prevalence of MCI is between 3.2 - 24\%, and found that the prevalence of MCI in people who have self-reported memory problems is almost double (18\%) the amount of do not self report.
@@
 When dealing with heavily imbalanced datasets with text classification as an objective, the one-class classifier approach has been demonstrated to outperform conventional two-class classification.
+The goal would be to compare this method of classification, with other more traditional classification methods previously applied to this problem.
 

### 2019-05-09 Source/Systematic Review/Systematic Review - Protocol.tex
updated Systematic Review protocol

@@
 \begin{document}
-
+\tableofcontents
 % https://github.com/anadallora/SNAC/blob/master/protocol%20v4.pdf
@@
 \begin{table}
+	\begin{tabular}{ p{3cm} | p{3cm} | p{3cm} | p{3cm} }
+	\hline
+	Scopus & Web of Science & PubMed & IEEE Xplore \\ \hline
+	1002 & 991 & 376 & 230 \\ 
+	\hline
+	\end{tabular}
+	\caption[Table caption text]{Totals from each database}
+	\label{table:name}
+\end{table}
+
+\subsection{Primary Study Selection}
+Firstly, the results from the automated searches from the three selected sources (Scopus, Pubmed and Web of Science) will have their duplicates removed, and this set of possible primary studies will be assessed by the evaluation of the titles and abstracts according to the defined inclusion and exclusion criteria.
+The result from this is our first selected subset (Selected Subset 1).
+The next phase is to look at the reference lists of papers in Selected Subset 2 and identify any papers missed in the automated searches.
+This is our second selected subset (Selected Subset 2).
+\par
+Each paper in these subsets will be fully read and assessed for it's quality, using the process documented further down.
+Any papers that meet the established criteria will be included as a primary study for this systematic review.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=300px, height=150px]{images/PRISMA - Chart.png}
+\caption{Selection Chart}
+\end{figure}
+\subsection{Study Quality Assessment}
+The papers selected will be full read and will then pass through an assessment of quality. The assessment will answer the following questions. 
+\begin{table}
+	\begin{tabular}{ p{6cm} | p{6cm} }
+	\hline
+	Question & Rating Scale \\ \hline
+	Are the aims of the study clearly stated? & Yes=1; Partly=0.5; No=0 \\ \hline
+	Does the study describe clearly the population being studied? & Yes=1; Partly=0.5; No=0 \\ \hline
+	Was the sample size justified? & Yes=1; Partly=0.5; No=0 \\ \hline
+	Is the sample representative of the population to which the results will generalize? & Yes=1; Partly=0.5; No=0 \\ \hline
+	Were the limitations of the study reported either during the explanation of the study design or during the discussion of the study results? & Yes=1; Partly=0.5; No=0 \\ \hline
+	Were the findings clearly reported? & Yes=1; Partly=0.5; No=0 \\ \hline
+	\hline
+	Are the measures used in the study clearly defined? & Yes=1; Partly=0.5; No=0 \\ \hline
+	Are the measuses used in the study valid? & Yes=1; Partly=0.5; No=0 \\ \hline
+	Are the measures used in the study clearly described? & Yes=1; Partly=0.5; No=0 \\ \hline
+	\hline
+	Is/are the technique(s) being employed clearly described? & Yes=1; Partly=0.5; No=0 \\ \hline
+	In the case of more than one technique, was the statistical significance assessed? & Yes=1; Partly=0.5; No=0 \\ \hline
+	Was a senstivity analysis carried out to assess if the results were due to certain inputs? & Yes=1; Partly=0.5; No=0 \\ \hline
+	\end{tabular}
+	\caption[Table caption text]{Paper quality assessment}
+	\label{table:name}
+\end{table}
+
+\subsection{Data Extraction}
+\subsection{Data Synthesis and Reporting}
+
+\appendix
+\section{\\Appendix A - Change Log}
+
+\section{\\Appendix B - Searches}
+\begin{table}
 	\begin{tabular}{ p{12cm} }
@@
 \end{table}
-
-\begin{table}
-	\begin{tabular}{ p{3cm} | p{3cm} | p{3cm} | p{3cm} }
-	\hline
-	Scopus & Web of Science & PubMed & IEEE Xplore \\ \hline
-	1002 & 991 & 376 & 230 \\ 
-	\hline
-	\end{tabular}
-	\caption[Table caption text]{Totals from each database}
-	\label{table:name}
-\end{table}
-
-\subsection{Primary Study Selection}
-Firstly, the results from the automated searches from the three selected sources (Scopus, Pubmed and Web of Science) will have their duplicates removed, and this set of possible primary studies will be assessed by the evaluation of the titles and abstracts according to the defined inclusion and exclusion criteria.
-The result from this is our first selected subset (Selected Subset 1).
-The next phase is to look at the reference lists of papers in Selected Subset 2 and identify any papers missed in the automated searches.
-This is our second selected subset (Selected Subset 2).
-\par
-Each paper in these subsets will be fully read and assessed for it's quality, using the process documented further down.
-Any papers that meet the established criteria will be included as a primary study for this systematic review.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=300px, height=150px]{images/PRISMA - Chart.png}
-\caption{Selection Chart}
-\end{figure}
-\subsection{Study Quality Assessment}
-\subsection{Data Extraction}
-\subsection{Data Synthesis and Reporting}
 \end{document}

### 2019-05-17 Source/Systematic Review/Systematic Review.tex
changes to systematic review

@@
 
-\begin{table}
-	\begin{tabular}{ p{6cm} | p{6cm} }
-	\hline
-	Stage & Papers remaining\\ \hline
-	Start & 2599 \\ \hline
-	Duplicates & Non-papers removed & 2599 \\hline 
-	\hline
-	\end{tabular}
-	\caption[Table caption text]{Totals at each stage}
-	\label{table:name}
-\end{table}
-
-After the evaluations, 37 papers were selected.
-Then a one-iteration backward snowballing was carried out looking for possible additional studies.
+After these evaluations, 37 papers were selected.
+Then a one-iteration backward snowballing process was carried out looking for possible additional studies.
 The 1199 new identified studies were assessed analogously as the previous ones, resulting in 41 new selected papers.
@@
 
-\begin{figure}
-
-\begin{tikzpicture}[sibling distance = 10em,
-	every node/.style = {shape = rectangle, rounded corners,
-	draw, align = center, 
-	top color = white, bottom color=blue!20}]]
-
-\end{tikzpicture}
-\caption[Figure caption text]{PRISMA flow chart.}
-\label{figure:name}
-\end{figure}
 \subsection{Language Analysis of Speakers with Dementia of the Alzheimer's Type - Guinn and Habash (2012)}

### 2019-05-28 Source/Thesis/Thesis.tex
updated Thesis

@@
-%% Document Class
-\documentclass[a4paper]{book}
-%% Packages
-\input{settings/packages}
-\usepackage{graphicx}
-\usepackage{float}
-\usepackage{enumitem}
-\usepackage{fancyhdr}
-\usepackage{rotating}
-%% Page Settings
-\input{settings/page}
-%% Own Commands
-\pagestyle{fancy}
-\fancyhf{}
-\fancyhead[CE,CO]{\leftmark}
-\fancyfoot[CE,CO]{Early Detection of MCI that progresses to dementia - An Interdisciplinary Approach}
-\fancyhead[LE,RO]{\thepage}
+\title{Early Detection of MCI that progresses to dementia - An Interdisciplinary Approach}
+\author{
+        Jomar Alcantara \\
+        Department of Computer Science \\
+        School of Engineering and Applied Sciences \\
+        Aston University\\
+        Birmingham, United Kingdom
+}
+\date{\today}
 
-\renewcommand{\headrulewidth}{2pt}
-\renewcommand{\footrulewidth}{1pt}
-\input{settings/macros}
-\title{Thesis}
-\date{2019-03-27}
-\author{Jomar Alcantara}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin {document}
+\documentclass[twopages, a4paper]{book}
+
+\begin{document}
 \maketitle
@@
 \section*{Abstract}
+\newpage
 % Acknowledgements
 \section*{Acknowledgements}
+\newpage
 \tableofcontents
@@
 \chapter{Introduction}
-Dementia has been identified as one of those fast growing difficulties facing the world.
+Dementia has been identified as one of the one of the biggest global health and social care challenges facing the world tday.
 A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
@@
 \par
-In 2009, the Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
-A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
-As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages \cite{Prince2015} and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
+In 2009, the UK's Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
+A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two distinct areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
+As our aging population increases, the potential burden on health care and social services will increase and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
 At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017}.
@@
 \par
-There is a large body of research that looks at the decline in language in those with MCI and AD \cite{Taler2008, Boschi2017}.
-However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD \cite{Taler2008, Boschi2017}.
-Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
-Another area of focus for research of this nature is the process of collecting appropriate language samples.
-Whilst collecting samples of language is comparatively unintrusive, researchers recognise that these samples require a rich sample of language that potentially cannot be generated by tasks such as the picture description task.
-Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD.
-There is some evidence to support this.
-Berisha et al \cite{Berisha2015}, has shown through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control.
-\par
-The potential impact of this research in this area is immense.
-Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease.
-Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
+\chapter{Background and Related Work}
 
-The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to to time-consuming and expensive psychological tests.
-There is a lot of research into the analysis of language as a bio-marker for MCI and Early Dementia.
-Given that sampling a person's language is relatively effortless, my research looks at whether we can find bio-markers of MCI and Early Dementia in natural language.\newline
-Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
-I'm not going to control for memory problems as a potential confound, but does this weaken the research?
-How do I defend this?
-\newline
-\par 
-Introduction to the problem of dementia in the context of the wider world including quality of life and financial implications.
-Exploration of dementia as a syndrome rather than a disease, and a look at the different variants of dementia.
-A look at the rationale behind research into the early diagnosis of dementia as well as a brief look at what has been done in the area (wide context, so pharmacological and psychological).
-\newline
-\chapter{Background and related work}
-Alzheimer's Disease (AD) is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
-Whilst a definitive diagnosis of Dementia can only be produced at post-mortem, there are a number of clinical indicators from psychological perspective that can indicate dementia is present.
-From a clinical psychology perspective, those who have dementia demonstrate cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}.
-In addition, these symptoms are often accompanied by emotional problems such as depression and behavioural difficulties.
-\par
-AD and other forms of dementia affect a significant proportion of the geriatric population in the world today and is currently the sixth leading cause of death in the US and was named the leading killer of women in the UK.
-According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
-This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population \cite{AlzheimersSociety2014}.
-From a financial perspective, they estimate an annual spend of £4.3 billion of which approximately £85 million is spent solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
-Globally, this picture is a lot bleaker.
-Another report by Alzheimer's Disease International suggests that in 2015 there were 46 million people with a diagnosis of dementia globally and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
-The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
-\par
-Despite this growing problem, at present there are no drugs that improve the prognosis of those suffering with AD.
-All the drugs that are on the market are designed to manage symptoms.
-Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{AlzheimersSociety2014}.
-Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place.
-It is therefore important to focus AD at it's earliest stages which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.
-\par
-Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent.
-Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time and alter cognitive function until a threshold is reached and clinical symptoms become more obvious \cite{Nestor2006}.
-The case of Iris Murdoch, who had a confirmed diagnosis of dementia, illustrates this theory well.
-Le et al \cite{Le2011} found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel when, it is theorized, the symptoms of AD manifested themselves more significantly.
-If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
-One of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of dementia.
-Albert and his team have worked to define clinical criteria which professionals can use to diagnose MCI due to AD and differentiate this from age-associated memory impairment and age-associated cognitive decline.
-They note that the diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}.
-\begin{enumerate}
-	\item Concern regarding a change in cognition - A person or an informant should express concern that there is a change in cognitive ability in comparison to previous level of performance.
-	\item Impairment in one or more cognitive domains - There should be evidence of lower performance in one or more cognitive domains beyond what would be expected of a person given their age and education.
-	\item Preservation of independence in functional - Whilst persons with MCI are expected to be able to maintain independence, it is common to experience mild problems in complex functional tasks which they may have been able to perform previously.
-This might mean that they take more time or be less efficient at completing these tasks, or it may be that may make more mistakes.
-	\item Not demented - The deterioration should be mild to the point that there is no significant loss of functioning in social or occupational contexts.
-\end{enumerate}
-In addition to meeting the above criteria, a clinician must rule out other conditions or factors that could account for the decline in cognition with the goal to increase the likelihood that the underlying cause of this decline is dementia.
-\par 
-Many researchers have studied the early detection of MCI.
-These studies usually follow two main approaches.
-The analysis of biomarkers and the examination of patients who have demonstrated decreasing cognitive abilities.
-The first approach yields reliable results in the detection of AD in its moderate and advanced states but does not perform well during the early stages of the disease.
-The second approach has gained more attention in recent years, due to the fact that in clinical practice it has shown promise in the early detection of AD.
-In addition, the analysis of the decline of cognitive abilities is comparatively inexpensive and less invasive that the first approach which commonly requires the collection of a sample of cerebro-spinal fluid which is painful for the patient involved.
-\par
-One of the most common ways in which clinicians traditionally make an early diagnosis of dementia is through the use of the Mini Mental State Examination (MMSE) \cite{Folstein1975}.
-The MMSE is a brief questionnaire consisting of eleven questions which tests cognitive aspects of mental function and requires only 5-10 minutes to administer \cite{Folstein1975}.
-The MMSE is chosen due to it's effectiveness at assessing a person's cognitive mental state at a specific point in time, as well as being as sensitive to changes as a more detailed and complex assessment such as the Wechler Adult Intelligence Scale \cite{Folstein1975}.
-Whilst the MMSE is useful as a brief screening tool it has it's limitations.
-The MMSE was not specifically created to screen for dementias and therefore does not interrogate key aspects of cognitive impairment known to be affected in dementia.
-It also has limited value in assessing under-educated subjects and a meta-analysis on the effectiveness of the MMSE as a diagnostic tool for dementia showed that it's accuracy was low (sensitivity between 78.4\% and 85.1\% and specificity between 81.3\% and 87.8\%).
-As the MMSE has been shown to have low accuracy specifically in the diagnosis of dementia, it becomes necessary for professionals to employ the use of other tools or measures such as the Free Cued Selective Reminding Test (FCSRT) \cite{Grober2010} or the Montreal Cognitive Assessment (MoCA) \cite{Davis2015}.
-\par 
-These tests have the benefits of being much more accurate at diagnosing cognitive impairment and discriminating between dementia and other types of cognitive impairment at the cost time and training of psychological professionals such as clinical psychologists in administering these tests.
-However, the utility of diagnosing dementia at the point where clinical intervention is warranted is limited because at this stage both psychological and pharmacological interventions have been shown to not be effective \cite{Prince2015}.
-In order to further our understanding of the progression of dementia it is important to detect the signs of dementia before they are clinically apparent.
-\par
-The two main ways in which diagnosis is performed is through assessment of memory and language.
-Tests of memory are classically among the most accurate ways of diagnosing dementia, however these tests suffer from the same reliance on clinicians to administer these tests in a clinical setting.
-Language however is a lot easier to collect and can be done in more naturalistic settings.
-As with memory, these tests can be done over time and would be able to chart a patients language degeneration over time.
-Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on measure decline in the use of language in those with AD.
-There are a number of difficulties to watch out for with this approach.
-There are a wide number of factors that are involved in language degeneration in the elderly, and consequently there will be an expected amount of variability between subjects.
-The administration of such tests may induce nervousness and discomfort which may impact performance, and also repeatedly administering the same language tests for differences over time may be confounded by improved performance at tasks via practice effects.
-However, there is enough promise in this approach such that it could help further our understanding of the disease, it's progression and the parts of the brain affected in the early stages.
-\par
-According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with mild dementia suffer from noticeable word finding difficulty.
-They may substitute general terms for more specific terms and may avoid the use of specific names of acquaintances.
-There may be grammatical errors involving subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc.
-Those who have progressed from Mild to Major depression also have difficulties with expressive or receptive language.
-They will often use general-use phrases such as "that thing" and "you know what I mean" and prefers general pronouns rather than names.
-With severe impairment, sufferers may not even recall names of closer friends and family.
-Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur.
-Echolalia (meaningless repetition of another person's spoken words) and automatic speech typically precede mutism.
-With the wide range of deficits someone with AD can suffer, it makes sense to try to categorise these deficits in some way.
-\par
-One of the most famous pieces of research on the topic of language decline in dementia was by Berisha and Liss who examined speeches and public interviews of former US president Ronald Reagan \cite{Berisha2015}.
-They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease.
-It turned out later to be the case that he had Alzheimer's Disease.
-Another classical study by Snowdon et al looked at whether linguistic ability in early life was associated with cognitive function and AD in later life \cite{Snowdon1996} .
-They found that idea density (defined as the number of expressed propositions divided by the number of words) was a key predictor in predicting whether nuns would go on to develop AD in later life.
-They found that those who would go on to develop AD all had low idea density in early life and they found no AD present in those with high idea density in early life.
-As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are extremely variable and can differ from patient to patient as the disease progresses.
-The consensus among researchers that this language degeneration is typically accelerated by the presence of dementia \cite{Berisha2015} and that a potential indicator of dementia is the rate of change in which the decline occurs relative to a fixed point in time rather than a comparison across a cohort of individuals.
-\par
-Emery \cite{Emery2000} completed a literature review looking at all the potential language deficits that could exist in those with AD and / or MCI.
-She divided these deficits into four levels of language: Phonology, Morphology, Syntax and Semantics.
-She proposed that language and the processes involved in language are hierarchical in nature and that language moves from simple units of construction (Phonology and Morphology), and build layers of complexity and sophistication (Syntax and Semantics).
-She found that people with AD generally had intact Phonology and Morphology but more impaired Syntax and Semantics.
-She asserted that the language forms we learn last are the first to deteriorate as we generally learn language in small simple units initially and build syntax and complexity as we are more comfortable with language.
-However it is important to note that different variants of dementia show different deficits in terms of language productions.
-Regardless, dividing language in this way is useful as it allows us to detect deteriorations in different parts of language usage and therefore may provide a way of discriminating between different forms of dementia.
-\par
-It is clear from both the clinical diagnostic criteria and supporting research that language is impacted in those with AD.
-However, whilst there is a move towards research aimed at looking more specifically at MCI we currently lack the measures that are sensitive enough to detect MCI.
-Given that we know language is affected before a clinical diagnosis of dementia is usually made \cite{Berisha2015, Snowdon1996, Le2011}, it makes sense to explore whether language on it's own can provide markers that may indicate a cognitive impairment that could progress to dementia.
-The field of machine learning and natural language processing has been suggested as a way to improve the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline \cite{Boschi2017}.
-% -- END OF INTRODUCTION --
-\subsection{Aims and Methodology}
-\par
-The purpose of this review is to look at the current standards of cognitive tests available that can be used as a diagnostic tool for MCI.
-I then go on to explore what techniques for assessing language have been used within the field of cognitive and clinical psychology.
-I then go on to look at what techniques have been developed in the field of machine learning (ML), deep learning (DL) and natural language processing (NLP) that might enable the automated analysis of language easier as well as an obstacles and/or limitations of current technology.
-Finally, I look at some studies which have already looked at the intersection of these two domains.
-\newline
-\par
-A search of the literature was conducted using ProQuest (PsychArticles), SCOPUS, Web of Science.
-The following results were found (Table 1).
-All papers were then reviewed for relevance by reading the abstract and full text where appropriate and a shortlist was compiled.
-An additional search through references of shortlisted papers was also conducted and any papers who upon further review appeared relevant were added to the shortlist.
-Papers were included where researchers used machine learning to classify participants as MCI, AD or Healthy using language.
-We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English.
-\begin{table}[H]
-	\begin{center}
-	\begin{tabular}{ | p{4cm} | p{1cm} | p{6cm} | p{1cm} |}
-		\hline
-		Database & Results & Search Terms  \\ \hline
-		ProQuest(PsychArticles) & 1484 & Language AND Decline AND Dementia \\ \hline
-		ProQuest(PsychArticles) & 486  & Language AND Decline AND Dementia AND Speech \\ \hline
-		ProQuest(PsychArticles) & 159 & Machine Learning AND Dementia AND Language \\ \hline
-		Web of Science & 1207  & Language AND Decline AND Dementia   \\ \hline
-		Web of Science & 151  & Language AND Decline AND Dementia AND Speech  \\ \hline
-		Web of Science & 34 & Machine Learning AND Dementia AND Language \\ \hline
-		Scopus & 791 & Language AND Decline AND Dementia  \\ \hline
-		Scopus & 91 & Language AND Decline AND Dementia AND Speech   \\ \hline
-		Scopus & 29 & Machine Learning AND Dementia AND Langauge \\ \hline
-		Scopus & 1292 & "Language Deficits" AND Dementia \\ \hline
-	\end{tabular}
-	\end{center}
-	\caption{\label{tab:table-name}Search Terms and Number of Results.}
-\end{table}
-% -- END OF Aims and Methodology
-\section{Existing Neuropsychological Measures of Cognitive Impairment, and Repeatable Battery of Neuropsychological Tests}
-In terms of standardised cognitive tests there are two main aims.
-Does the test distinguish accurately between normal aging, MCI and AD (diagnostic utility) and does the test distinguish between those individuals with MCI who will then go on to develop AD and those individuals with MCI who don't then go on to develop AD (prognostic utility).
-This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their performance in terms of both diagnostic and prognostic utility.
-\par 
-\subsection{Repeatable Battery for the Assessment of Neuropsychological Status}
-The Repeatable Battery for the Assessment of Neuropsychological Status (RBANS) was originally developed as an assessment tool for dementia, specifically looking at detecting a characterizing very mild dementia \cite{Randolph1998}.
-The authors felt that there was a shortfall of appropriate measures that were sensitive enough to milder impairments as well as a number of other shortcomings of existing tests.
-They met a number of design goals for this new battery of tests that addressed these shortcomings.
-The RBANS consists of a number of sub-tests across five distinct domains.
+\chapter{Systematic Review}
 
-\begin{enumerate}
-	\item Immediate Memory 
-	\begin{enumerate}
-		\item{List Learning}
-		\item{Story Memory}
-	\end{enumerate}
-	\item Visuospatial / Constructional
-	\begin{enumerate}
-		\item{Figure Copy}
-		\item{Line Orientation}
-	\end{enumerate}
-	\item Language
-	\begin{enumerate}
-		\item{Picture Naming}
-		\item{Semantic Fluency}
-	\end{enumerate}
-	\item Attention
-	\begin{enumerate}
-		\item{Digit Span}
-		\item{Digit Coding}
-	\end{enumerate}
-	\item Delayed Memory
-	\begin{enumerate}
-		\item{List Recall}
-		\item{List Recognitions}
-		\item{Story Memory}
-		\item{Figure Recall}
-	\end{enumerate}
-\end{enumerate}
+\chapter{Language Data Collection - Developing a consensus}
 
-However, whilst the authors claim that the RBANS is adequately sensitive in person's with MCI \cite{Randolph1998} and there is some research to support this assertion \cite{Karantzoulis2013}, other research points out that the RBANS has poor sensitivity in detecting MCI \cite{Duff2010}.
-Another drawback of this battery is the lack of executive function measures and object naming tasks.
-Research has shown that the RBANS has good test-retest reliability and convergent validity.
-However, Duff et al warned that caution should be exercised when using the RBANS in a MCI population as it has lower sensitivity in this population \cite{Duff2010}.
+\chapter{Pipeline Development}
 
-\subsection{Digit Span Test}
-The Digit Span Test (DST) is predominantly used to measure a person's working memory capacity, specifically the capacity used to store and recall numbers.
-A participant is presented with a series of numbers of fixed length and is asked to recall those numbers in normal or reverse order.
-The series of numbers gets progressively longer until such time as a participant fails three or more times out of eight presentations.
-\par
-Research has consistently shown that MCI patients have a significantly lower digit span score in both normal and reverse order versions of this test and a study by Muangpaisan has shown that the reverse order version of this test can, to some degree, predict a diagnosis of MCI \cite{Muangpaisan2008}.
-Muangpaisan also found that Age, Gender and Education have an impact on the performance of the tests.
-Emrani et al found that the DST revealed no differences between specifically amnestic MCI and controls, but could differentiate between Mixed MCI and the other groups with mixed MCI recalling fewer correct responses than other groups.
-Notably, there was an attenuated recency effect in those with mixed MCI.
-\cite{Emrani2018}.
-Kessels identifies that that there are working memory deficits in MCI patients and these worsen with AD patients.
-\cite{Kessels2011} He also identified that both MCI and AD have impaired performance on all three conditions of the digit span test.
-No differences were found between forward and backward conditions in any of the groups.
-However, available tests may not detect subtle impairments \cite{Kessels2015}.
-\subsection{Rey Auditory-Verbal learning test}
-Rey's Auditory Verbal Learning Test (RAVLT+) looks a wide range of neuropsychological processes including short-term auditory-verbal memory, learning and retention of information.
-Participants are given a list of 15 unrelated words, repeated over five different trials and are asked to repeat.
-Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then repeat it again after 30 minutes \cite{Schmidt1996}.
-\par
-Several studies have shown that an impairment in RAVLT score reflect well the underlying pathology caused by AD.
-Thus making the RAVLT an effective early marker to detect AD in persons with memory complaints.
-Moradi investigated to what extent the RAVLT scores are predictable based on MRI data using machine learning approaches, as well as to find out what the most important brain regions are for the estimation of RAVLT scores \cite{Moradi2017}.
-They found a highly significant cross validated correlation between the estimated and observed RAVLT immediate and RAVLT Percent Forgetting.
-Further, they found that the conversion of MCI subjects to AD in 3-years could be predicted based on either observed or estimated RAVLT scores with an accuracy comparible to MRI-based biomarkers \cite{Moradi2017}.
-Another study by Schoenberg found that RAVLT to best distinguish patients suspected of Alzheimer's disease from the psychiatric comparison group \cite{Schoenberg2006}.
-\subsection{Digit Symbol Substitution Test}
-The Digit Symbol Substitution Test (DSST) involves a key consisting of the numbers 1-9 and a corresponding unique symbol.
-Below this key is a series of numbers from 1-9 in a randomized order and repeated multiple times.
-The participant is asked to fill in the corresponding symbol for each number.
-The task requires that the participant move between the key and the randomized sequence such that  they may retrieve the correct answer from the key, hold this in short-term memory and transcribe the key in the appropriate place.
-\par 
-Among those with no disorder in cognition, mobility and mood, being in the lowest DSST quartile compared to the highest was associated with nearly twice the odds of developing one or more clinical or subclinical disorders.
-Associations were stronger for incident clinical disorders in cognition.
-Slower psychomotor speed may serve as a biomarker of risk of clinical disorders, mobility and mood.
-While in part attributable to vascular brain disease, other potentially modifiable contributors may be present \cite{Rosano2016}.
-Further studying the causes of psychomotor slowing with ageing might provide novel insights into age-related brain disorders.
-Pascoe compared patients with PD with Normal Cognition (PD-N) with those with PD and MCI (PD-MCI) and healthy participants.
-PD-MCI participants achieved significantly lower scores than other groups in the DSST task \cite{Pascoe2018}.
-\subsection{Verbal Fluency}
-There are a number of tests that characterise this category of verbal fluency, but generally fall into two categories.
-Letter (Phonemic) fluency involves the generation of as many words as possible which begin with a specified letter.
-Category fluency involves the generation of as many words as possible that fall into a specfied category.
-Both these tasks impose demands on a number of different cognitive processes namely, executive function, verbal retrieval and recall, giving appropriate answers while monitoring previous answers and inhibit inappropriate responses.
-However, these two tasks require different strategies when attempting them.
-Letter fluency relies on search strategies based on lexical representations whereas category fluency requires a search for semantic extensions of a superordinate term, meaning that semantic associations within the lexicon must be intact in order for the task to be carried out successfully.
-\par  
-There have been numerous studies which have documented the impact that AD has on verbal fluency tests in both categories.
-A review carried out by Henry et al \cite{Henry2004}, found that performance in both letter fluency and category fluency was impaired in those with AD vs controls, but found a larger effect for tests of semantic fluency.
-\subsection{Naming Tests}
-Word-finding difficulty is a common symptom of AD and these deficits usually occur during the early stages of the disease progression.
-As such , a test of a patients ability to find words (known as confrontation naming) is a common way to measure cognitive decline.
-One common way to do this is the Boston Naming Test (BNT) which comprises 60 items on a spectrum of very frequent to very infrequent.
-This has been reduced subsequently to two thirty item versions and four fifteen item versions, which correlate significantly with the original sixty item version and the benefit of the shorter versions of the test is that it facilitates testing of individuals with AD who may suffer from fatigue or limited attention span \cite{Williams1989}.
-\par 
-Willers et al \cite{Willers2008} studied Twenty aMCI patients, twenty AD and 21 controls matched by age, sex and education level.
-They found AD patients obtained significantly lower total scores on the BNT than aMCI patients and controls.
-aMCI patients did not obtain significant differences in total scores but showed significantly higher semantic errors compared to controls.
-Semantic processing is impaired during confrontation naming in aMCI.
-\par 
-Vadikolias investigated the impact of education on naming tests \cite{Vadikolias2012} and found that higher educational attainment in aMCI subjects were correlated with better performance in verbal and non-verbal tasks during repeated examinations over 1-year.
-Subjects with a lower level of education performed worse than patients with a high level of education who presented a more stable clinical score.
-The explanation for this is the idea of a 'cognitive reserve' in participants with a higher education such that this provides a buffer that, while not preventing the physiological symptoms of AD, can potentially delay the clinical onset of cognitive symptoms that characterise AD.
-This theory is supported by Snowdon who found a relationship between early life linguistic ability and the density of neurofibrillary tangles in his nun study \cite{Snowdon1996}.
-Whilst these studies and tests provide evidence that there are word finding difficulties in those with MCI, on their own they do not provide sufficient ability to diagnose MCI or provide an prognosis of disease progression.
-\subsection{Rey-Osterrieth Complex Figure Task}
-The Rey-Osterrieth Complex Figure (ROCF) is a task widely as a test of visuo-spatial skill and visual memory.
-The task, which was originally designed by Rey (1944) and standardised by Osterrieth (1944) \cite{Rey1941, Osterrieth1944}, asks a participant to copy a complex geometrical figure (known as the immediate copy condition and to recall and reproduce the figure from memory without warning (known as the delayed recall condition).
-In the immediate copy condition, the complexity of the figure requires an integrative cognitive ability.
-The reproduction of such a complex structure involves processes such as planning and organizational strategies that are related to executive functions.
-In order to make this task repeatable, Taylor et al designed a comparable set of figures which have proven to be of equivalent difficulty \cite{Taylor1969, Strauss1990}.
-\par
-Salvadori found that patients with vascular MCI had a worse performance in the immediate copy of the ROCF compared to individuals with degenerative MCI, despite their significant impairment in terms of general cognitive status and visual memory \cite{Salvadori2018}.
-Evidence shows that patients with disorders that possibly involve attention and executive functions are characterized by a more disorganised approach when copying the ROCF compared to controls.
-One of the difficulties with the presentation of this task in a repeated battery will be the fact that it turns from an incidental memory task (it is a memory task but the participant is not forewarned that they will be required to memorise the picture) into an intentional memory task (given the previous exposure to the task, it would be expected that a participant would preempt the delayed recall portion of the test and spend more time attempting to memorise the details) \cite{Teng2009}.
-\subsection{Hayling Sentence Completion Test}
-Inhibitory deficits are a common in all stages of dementia.
-This is usually tested using Stroop test, however this has the drawback of lacking ecological validity.
-Therefore researchers have moved towards using the Hayling Sentence Completion Test (HCST) which uses skills such as word retrieval as well as the ability to inhibit ones responses where appropriate and is therefore a much more ecologically valid task.
-There are two parts to the HCST.
-In the first part, participants have to complete a sentence by providing a word that best fits the given sentence (this is known as the initiation condition).
-In this second part, participants have to complete sentences by inhibiting an impulse to give the word that best first the sentence as in the first part, and producing a semantically unconnected word.
-Performance in both conditions is measured by the time taken for the participant to initiate a response, and in the inhibition condition also by the correctness of the word.
-\par 
-Martyr et al compared healthy controls with patients with dementia and patients with Parkinson's disease \cite{Martyr2017}.
-They found that a high proportion of Category A errors (producing a word that fits the sentence when instructed otherwise) was a factor in performance loss for participants with dementia.
-Findings suggest that the HSCT may be sensitive to verbal suppression deficits and may provide insight into inhibitory control in participants with dementia.
-Patients with Dementia were significantly slower than controls in the initiation and inhibition conditions vs healthy controls, and slower than patients with Parkinson's disease in the initiation but not the inhibition condition.
-\subsection{Grooved Pegboard Test}
-The Grooved Pegboard Test (GPT) was originally designed to cover a variety of different psycho-motor functions including hand-eye coordination and motor speed.
-However, some studies have shown a correlation of performance in the GPT and measures of cognitive performance such as the Montreal Cognitive Assessment.
-\par 
-Bezdicek found that the GPT predicted performance on the MoCA and concluded that in addition to being a measure of motor skills, there were results that showed that the GPT could also provide information about a participants cognitive skills \cite{Bezdicek2014}.
-Whilst this was specifically using a population of patients with Parkinson's disease, given the MoCA and GPT have shown degree of correlation, there remains an opportunity for research that explores the use of the GPT with MCI and Early AD populations.
-\par 
-Darweesh et al, used the Purdue Pegboard Test (PPT) to assess manual dexterity in a healthy older adult population and followed their participants for between eight and twelve years until an indication of the onset of a neurodegenerative disease \cite{Darweesh2017}.
-In this time 227 (4\%) of their participants went on to develop a diagnosis of dementia.
-They noted that higher PPT scores were associated with lower risk of incident neurodegnerative disease and noted significant associations of PPT scores with all forms of dementia and this potentially highlights a deterioration of motor function in the pre-clinical phase of dementia.
-\subsection{Visual Search}
-The visual search task requires participants to be watching a computerised display in which a target appears either by itself or surrounded by other elements which are used as a distraction.
-These other elements can be vary in similarity to the target.
-This is inherently a measure of attention shifting and processing of various similarity and as one would expect, the visual search time is increased in the presence of distractors.
-This effect is magnified in participants who have a cognitive impairment.
-\par
-Research has shown there to be deficits in both AD and MCI populations in this task \cite{Tales2005}, and those with MCI exhibited less severe deficits compabut that these deficits were not as apparent in MCI populations, there remained a significant enough difference that could be used to differentiate healthy controls from those with MCI.
-In a comparison of visual search performance between those with aMCI and healthy controls, Tales noted significantly poorer performance in the aMCI group.
-However, she also noted a good deal of heterogeneity in the aMCI group which illustrates that whilst the aMCI group have essentially the same condition, presentations within this population can differ markedly.
-The results from this study also illustrates the use of a non-memory task as a means of diagnosing dementia as patients who went on to be diagnosed with dementia whilst this study was in progress had significantly poorer visual search performance.
-\subsection{Free Cued Selective Reminding Test}
-The Free Cued Selective Reminding Test (FCSRT) was borne out of the premise that by controlling the conditions of learning, a measure of memory is possible that is not confounded by normal age related changes in cognition.
-Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cognitive decline due to a cognitive impairment.
-The FCSRT starts with a learning phase in which participants are asked to look at a card containing four pictures (e.g., grapes) for an item that belongs to a named category(e.g., fruit).
-After these four items are identified, immediate cued recall of these four items is tested and this is repeated for a total of 16 items.
-Following this learning phase is the recall phase in which participants are asked to recall all the items identified without cues and any items which are not recalled are then cued.
-There are three outcome measures for this test, free recall (the number of items recalled without cues), total recall (the number of items recalled with or without cues) and cue efficiency which is identified as follows \cite{Grober2010}.
-\begin{equation} \label{x1}
-CueEfficiency = (totalRecall-freeRecall)/48-freeRecall, range 0.0-1.0)
-\end{equation}
-Grober found that patients with impaired free recall with four times more likely to develop dementia than those with intact free recall \cite{Grober2010}.
-This test has also been shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the researchers to used the test to categorise prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT score \cite{Sarazin2007}.
-\subsection{Conclusions}
-I have looked at a number of different cognitive tests and a battery of tests that aim to have high diagnostic and prognostic capabilities in the MCI population.
-However, particularly with the RBANS, there lacks sufficient sensitivity in differentiating those with MCI from healthy elderly individuals such that this tool could be used with a level of confidence in a clinical setting.
-In regard to the tests, a common theme is a lack of studies and therefore evidence into the utility of these tests with an MCI population.
-However if, as researcher, we aim to investigate this population further then a benchmark battery of tests which is sensitive enough in both diagnostic and/or prognostic utility should be a goal.
-\par 
-Another criticism of the current literature is the lack of consistency with regard to the experimental groups.
-For example, some studies focus on differentiating between an MCI group, a AD group and a healthy controls group whereas other studies may further subdivide the MCI category according to a number of factors such as Amnesic or Non-Amnesic MCI, Mixed MCI (sometimes called Multi-domain MCI in the literature).
-Given the inconsistency in defining the experimental groups the confusing and often conflicting results that researchers produce is to be expected.
-Future research should look at the standardization of the operational definition of cognitive impairment in MCI may result in more consistent predictions of progression to AD.
-\par 
-Finally, it can be said that in testing for MCI and comparing these results with a cohort of those with a similar diagnosis has limited utility.
-This is because there is huge variability in the presentation of MCI within participants that even if controlled for with a matched pairs design for age, gender and education will produce inconsistent results.
-There is an argument therefore for the use of longitudinal studies, where the comparisons on the performance of these tasks are with the participants themselves.
-There are not many studies which use a longitudinal approach, although those that do show promising results.
+\chapter{Analysis of DementiaBank dataset}
 
-\section{Types of Language Assessment}
-One of the key debates when looking at how to analyse language is the type of task provided to elicit language production in participants.
-In the literature researchers have primarily focused on Picture Description tasks but have also suggested other ways in which we might collect data.
-\par
-\subsection{Picture Description Tasks}
-One of the most commonly used tasks to measure language is the Picture Description task.
-An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Borod1980}.
-The Cookie Theft picture (pictured below) depicts a scene of a home typical of the period of time when it was created and would generally not require participants to use any complicated vocabulary to describe.
-In this task participants are asked to describe the picture presented to them in as much detail as possible.
-This task was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}
-\begin{figure}[H]
-\centering
-\includegraphics[width=240px, height=150px]{images/BCTPicture.png}
-\caption{Cookie Theft Picture - From Kaplan and Goodglass (1983)}
-\end{figure}
-\par
-The picture description task does a fine job of eliciting descriptive language however because of the specific content the language produce could be considered quite limited.
-There is some disagreement as to the benefits of this using this methodology.
-This task is reported as being useful to lexico-semantic disorders \cite{Boschi2017, Sajjadi2012} as the language being generated is primarly nouns and deixis (words to identify items and words to put those items into context).
-However, Ash \cite{Ash2012}felt that there was no difference in using this task vs Story Narration (described below).
-In explaining the differences, it is worth noting that these researchers were using differing variables and this could explain their different perspectives.
-\par
-\subsection{Narrative description task}
-The story narration task is designed to study a participant's ability to describe and elaborate on a story which is depicted using a series of pictures.
-The stories depicted are usually based on children's books or famous stories with the Cinderella being the one most typically used \cite{Fraser2014}.
-This task requires ordering the story in a structured and coherent framework.
-It also requires comprehension and understanding of the stories characters and the events depicted, as well as an awareness of a character's actions, motivations and internal reactions to given events.
-This task is particularly useful as the procedure reduces the demands on memory, due to the participant being able to access the picture book during the description and is therefore able rule out memory as a confounding variable for any results observed.
-As noted above, Ash \cite{Ash2012} felt that this task was interchangable with the Picture Description task.
-However, other research felt that this was a studier test of lexical and semantic abilities as well as syntactic complexity because this task requires interpretation and elaboration in additional to a simple description \cite{DeLira2011}.
-\par
-Given the relative strengths of the Narrative description task vs Picture description task, there are few pieces of research that have used Machine Learning to analyse features from Narrative picture tasks \cite{Fraser2014}.
-This could be due to the availability of data and the absence of any meaningful sets of transcripts of participants performing this task.
-However, this could be an interesting direction to take research in the future to see if features generated from this task could be used to predict MCI or AD.
-\par
-\subsection{Interviews}
-Interviews can also be used to elicit language in a more natural way by asking questions to guide a conversation between speakers.
-There are three types of interviews: unstructured, structured and semi-structured.
-Structured interviews tend to produce very limited speech and therefore has never been used in this area \cite{Boschi2017}.
-Unstructured interviews are open ended and generally do not conform to any particular pattern.
-They use generic themes such as family or hobbies to guide the conversation.
-Whilst this is the most ecologically valid form of conversation and therefore language generation, it's unstructured nature means that the protocol cannot be consistent and therefore reproduced.
-Semi-structured interviews are therefore preferred over other forms of interview as a middle ground.
-The semi structured nature of these interviews means that there is some replicability but does not constrain the participant in answering questions.
-\par
-The analysis of interviews can be difficult to analyse as both the content can vary even between participants, although it can be argued that content should not affect the type of language being generated unless it is narrow topic or the participant is constrained in how they answer a given question.
-It is also difficult to measure as there are no pre-defined task goals in comparison to the other two methods.
-Nevertheless, this is the most naturalistic setting for looking at language production and can be used to look at the syntactic and semantic parts of language generation \cite{Sajjadi2012}.
-There have been some attempts to use interviews to assess language production in AD with promising results \cite{Asgari2017, Guinn2015}.
-\par
-\subsection{Conclusions}
-One can view the different types of tasks above as a continuum where picture tasks represent a much more controllable task with a lot of supporting research but which generates a much more constrained set of language that is atypical of normal speech in terms of the cognitive functions used.
+\chapter{Pilot Study}
 
-\section{How do we analyse language, issues and debates}
-There are a wide variety of approaches that we can take in the analysis of language and there is no real consensus on the best approach to this particular problem.
-This section looks at the different ways in which researchers have looked at the problem as well as discussing some of the areas of contention.
-\subsection{Single Word Language tasks vs Connected Language tasks}
-Part of the reason we need to pay attention to how we ask participants to generate data is understanding how we wish to analyse the data afterwards.
-As discussed above, the different methodologies to collect data generate different types of language.
-There are two main approaches which we have looked at to analyse language, using frequencies of words and combinations of words and measures of syntax and semantics.
-There are other less common methods of analysing language but these are beyond the scope of this review.
-\par
-Single word tasks such as the Boston Naming Test and other such standardized language tests generally target a participants word production where this is defined as the ability to form and express words in accordance with certain criteria (see above for a full description of verbal fluency as task).
-There are a number of benefits to using single word tasks.
-From a research methodology perspective, using a standardized test allows researchers to target a very specific process in language generation and isolate factors that impact performance in language well.
-However, this approach does not take into account other cognitive processes that are used in lengthier speech tasks.
-More specifically, single word language tasks do not look at the interaction between language, executive functions and reasoning abilities.
-They also do not require the logical and efficient organisation of ideas.
-\par
-Connected language tasks, such as the picture description task and interviews described above are much more reflective of the processes involved in natural language generation.
-They are able to be relatively constrained in the language available to be used, for example in the picture description task, or they can be unconstrained in the form of interviews.
-But regardless of how they are framed, they are potentially much more useful in terms looking at all the processes involved in language generation with the drawback that you will not be able to isolate specific parts of the language generation process.
-A review by Boschi et al concludes that analysis of connected speech is potentially useful in guiding clinicians to identify language disorders \cite{Boschi2017} and also highlights the role of NLP and Machine learning in assisting in this endeavour.
-\par
-\subsection{Semantics vs Pragmatics}
-When navigating the English language it is necessary to distinguish between what a sentence says in both semantic and pragmatic terms.
-Semantic meaning refers to the meaning of the words in a sentence local only to the given sentence.
-Another way to put this is, semantics considers the meaning of words without taking into account the context in which these words are spoken.
-Pragmatic meaning refers looks at the same sentence in terms of words and grammar but takes into account the situation or context in which these words are spoken.
-Emery in her literature review looks at all levels of language tasks except for pragmatics, however this is an area that should not be ignored.
-Whilst the study of language from a pragmatic perspective is much more complex, it is perhaps one of the most vital areas to study because of the number of different cognitive processes involved.
-\par
-\subsection{Semantic Content}
-Another approach to linguistic analysis in this field is the idea of measuring semantic content and complexity.
-According to Emery (2000) \cite{Emery2000} in which she states that Semantic and Syntactic skills deteriorate first in people with MCI and AD.
-If this is true, then psychological measures of semantic and syntactic skills should be able to pick up signs of deterioration and act as markers for possible MCI and AD.
-An example of a semantic complexity measure is the concept of idea density.
-Formally, idea density is defined as the average number of propositions per sentence \cite{Keenan1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}.
-An example of semantic content measures is Type to Token Ratio (described below) which is used to measure the lexical diversity of a given piece of text and/or utterance.
-This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2013}.
-\par
-\subsubsection{Type token ratio(TTR)}
-Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
-\begin{equation} \label{x2}
-TTR = numberOfUniqueWords / totalNumberOfWords.
-\end{equation}
-\subsection{Thematic and Content elements in relation to the Picture description task}
-A number of studies looked at the accuracy of picture descriptions using counts of thematic and content elements within a picture.
-Whilst called diffferent names such as 'pictorial themes', 'relevant observations and 'semantic units', they all represented the same idea.
-The only difference between the studies was the number of thematic elements that 'scored' correctly.
-Nicholas et al identified eight thematic elements of the Cookie Theft picture and used the number of elements as an outcome measure in different groups.
-He found that patients with AD expressed significantly fewer content elements than controls.
-\par
-Hier, Hagenlocker and Shindler assessed content using a similar list of thematic elements \cite{Hier1985}.
-They divided their participants into early-stage and late-stage AD, as well as including a control group.
-The late-stage AD group produced significantly fewer relevant observations than the early stage group, and the AD group combined produced fewer relevant observations than controls.
-This study was replicated by Lukatela et al \cite{Lukatela1998}.
-\par
-Smith, Chenery and Murdoch applied Hier's methodology for constructing pictorial 'themes' with the Picnic Scene from the Western Aphasia Battery (WAB) with a control and patients with moderate to moderately severe AD \cite{Smith1989}.
-The authors found no difference in the number of semantic elements produced but did not that the group with moderate to moderately severe AD took more time and more syllables to communicate these elements.
-\par
-\begin{figure}[H]
-\centering
-\includegraphics[width=240px, height=150px]{images/picnic-scene.jpg}
-\caption{Picnic Scene taken from the Western Aphasia Battery (WAB).}
-\end{figure}
-Sajjadi et al examined 10 pictorial themes in picture description the Comprehensive Aphasia Test and found that the group with mild AD produced similar themes than controls \cite{Sajjadi2012}.
-Bschor et al.
-(2001) examined Cookie Theft picture descriptions at four stages of AD.
-They found that whilst each AD group differed significantly from the others and also from controls, the measures were not able to distinguish between MCI and normal controls \cite{Bschor2001}.
-\par
-Finally, a number of studies used composite measures which contained thematic elements and other unspecified information units resulting in a list of 23 possible information units of the Cookie Theft picture.
-The authors felt that this provided a wider, more liberal range of relevant content and thus subtler differences could be noted.
-Studies using these features found some differences between AD and controls, and some could differentiate between different stages of AD.
-\subsection{General Information Units or Content Information Units}
-Some studies used a more general concept of content, using terms such as "general information units" or "content information units" and this could be defined as "the smallest non redundant meaningful fact or inference,"  and was counted whether or not the information conveyed was specific to the context in which the conversation happened.
-Giles et al for example studied adults with minimal, mild or moderate AD vs controls and found that adults with AD produced fewer overall information units than controls \cite{Giles1996}.
-\subsection{Conciseness of information}
-Conciseness has been defined as the number of words a speaker uses to express ideas.
-The theory is that people with AD would need more words to convey ideas because of difficulties with word-finding and compensatory  behaviours such as circumlocutions and repetitions.
-Conciseness has previously been calculated by dividing the number of ideas expressed by the total number of words in a measure commonly referred to as idea density but also known as lexical index, information content and information unit conciseness index.
-Snowdon examined written discourse from the Nun study and found that low idea density in early life was associated with reduced cognitive performance in later life \cite{Snowdon1996}.
-Riley et al extended these findings by concluding that early-life idea density was associated with lower brain weight, higher degree of cerebral atrophy and increased neurofibrillary pathology in later life \cite{Riley2005}.
-\par
-Ahmed, de Jager et al examined idea density with patients who had confirmed AD post mortem \cite{Ahmed2013}.
-They found that those with AD produced fewer total semantic units than controls but there was no significant difference between the groups with regards to idea density.
-The study of "empty speech" by Nicholas et al examined conciseness and specifically looked at empty phrases (defined as common utterances which contribute no relevant information), deictic terms (e.g.
-"this", "that" without referents), indefinite terms (e.g.
-"thing" or "stuff"), pronouns without proper noun antecedants, and repetitions.
-In their study they found that AD patients produced more of these than did controls.
-\par
-\subsection{Efficiency}
-% CITE MURRAY!
-Efficiency is the rate at which meaningful information is conveyed over time,  and can be calculated by dividing the total number of information units by the duration in seconds of the speech sample.
-Smith et al, 1989 found that 18 adults AD produced fewer content units over time on average than controls, he attributed these differences to increased circumlocutions and repetitions in the AD group \cite{Smith1989}.
-Murray used an similar measure in which fillers, irrelevant words, revisions or false stars, vague or non-specific vocabulary and inaccurate output were group together as 'performance deviations" and were divided by the total number of minutes in the sample.
-This measure was lower for those with AD than those with depression, and also healthy controls.
-The authors suggested that discourse information measures may help disentangle the similarities in symptoms of early AD versus depression in older adults.
-Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in dialogue in which a speaker provides responses do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls.
-\subsection{Total number of words}
-Several studies report that adults with moderate AD produce fewer words than controls on picture description, however other studies found no differences in total words among groups of controls and patients with MCI or AD.
-Murray and Nicholas et al investigated normal controls, patients with AD and older adults with depression and found no group differences in total words \cite{Nicholas1985}.
-In contrast, Lira et al found that controls produced more total words than patients with AD but found no difference between mild and moderate groups \cite{Lira2014}.
-\par
-\section{Syntax and Morphology (Language Form)}
-Syntax can be defined as the rules that govern how words can be combined to form sentences, whilst Morphology is the system that governs the structure of words and the construction of word forms.
-Multiple studies of language decline in dementia included at least one measure of syntax or syntactic complexity \cite{Orimaye2017}.
-Common constructs included words per clause, grammatical form (measures of an appropriate use of syntactic conjections, tenses, conditionals, subordinate clauses and passive constructions), measures of phrase length and proportions of words in sentences.
-Some researchers have explored the use of formulaic language in those with dementia, the theory being that well practiced phrases are less effortful and therefore place low load on the cognitive abilities of those with AD.
-The general hypothesis motivating these studies is that either working memory limitations or semantic memory limitations in AD affect one's ability to use complex constructions.
-\par
-\subsubsection{N-grams}
-One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
-An n-gram is a contiguous sequence of n items from a given sample of text or speech.
-The items can be phonemes, syllables, letters, words or base pairs according to the application.
-For example, given the sequence of words "to be or not to be", this extract is said to contain six 1-gram sequences (to, be, or, not, to, be), five 2-gram sequences (to be, be or, or not, not to, to be), four 3-gram sequences(to be or, be or not, or not to, not to be) and so on.
-This is useful as, given a large portion of text or speech, we can predict the probability of a word being close by to a given word.
-A number of researchers have used n-grams as features.
-\par
-Asgari, Kaye and Dodge (2017) \cite{Asgari2017} used another form of word frequency measurement.
-Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words \cite{Pennebaker2015}.
-They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
-\par
-\subsection{Formulaic Language}
-Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
-They found that there were four factors which they identified as important in the classification of participants as either healthy or AD.
-These four factors were semantic impairment, acoustic abnormality, syntactic impairment and information impairment and were based on existing measures of semantic and syntactic complexity.
-Zimmerer (2016) \cite{Zimmerer2016} looked at whether language was more formulaic in those suffering from AD.
-He proposed that those who suffer from AD rely on formulatic sentences, for example 'Noun-Verb-Noun', and this is done to reduce language complexity.
-He noticed a significant difference in the use of formulaic sentences between AD and Healthy Controls.
-\par
-\section{Pragmatic Language}
-The pragmatic language domain refers to the social rules for language for the purposes of communication including, using language to achieve goals, using information from the context to achieve these goals and using the interaction between people to initiate, maintain and terminate conversations.
-\subsection{Coherence}
-Coherence, in lay terms, can be defined as the ability to maintain awareness of the topic at hand.
-It can be separated into local coherence which is related to themes of the immediately preceding utterance and global coherence which looks at how closely an utterance is related to the topic currently being discussed.
-Chapman et al used picture descriptions of Norman Rockwell prints within a frame analysis, with frames being defined as the context in which the picture is viewed \cite{Chapman1998}.
-The authors identified aspects of content, including whether the frames of interpretation being offered were typical, atypical, incorrect or had no frame.
-They also looked propositions supporting frames and propositions disrupting frames as measures of coherence.
-They examined these variables with early stage AD, old-elderly and normal controls.
-Healthy older adults and normal controls produced significantly more typical frames and more frame supporting information than the AD group.
-The authors attributed AD patients' difficulties to memory deficits, attentional deficits, visual perceptual problems, disruption of internalized frame representation, or failure to access frame knowledge.
-\par
-\subsection{Perseveration}
-Perseveration can be broadly defined as the repetition of a response regardless of the absence or cessation of stimulus which could have generated an appropriate response in the first place.
-A typical example could be the idea of conversation moving from introductions to a more general conversation, someone who has difficulties with perseveration will struggle with the move between social contexts and repeat language that would be more used in an introduction context.
-\par 
-One study examined verbal preservation in the description of Norman Rockwell prints \cite{Bayles198}.
-The presented participants with a number of similar pictures that had significantly different contexts and asked participants to describe these prints.
-They divided the total number of words within perseverations by total number of words in the speech sample.
-The authors also calculated rate of perseveration on two other language tasks: confrontation naming and generative naming.
-In all tasks, the AD group produced significantly more perseverations than controls but there were no significant differences between the two groups in the picture task in isolation.
-The authors felt that this was because picture description was an easier task, as it was a visual task in contrast to the other tasks which tapped into other cognitive processes.
-\par
-\subsection{Empty Speech}
-Verbal fluency is a term used in neuropsychological contexts generally referring to timed, word-generation tasks, while in speech-language pathology contexts, "fluency disorders" are defined as interruptions in the flow of speaking characterized by atypical rate, rhythm and repetitions in sounds, syllables, words and phrases.
-"Fluency", in the literature of discourse of adults with AD, typically refers to the smoothness or flow of spoken language.
-Abnormalities of fluency in this population are typically characterised by filled and unfilled pauses, word repetitions, circumlocutions, and revisions.
-\par
-The study of "empty speech" by Nicholas et al was one of the first to examine aspects of fluency in the connected speech of persons with AD \cite{Nicholas1985}.
-They found that adults with AD had significantly more repetitions than controls.
-Similarly, Bayles and Tomoeda found more aborted phrases, revisions and ideational repetitions in the AD group than in controls \cite{Bayles1983}.
-Several other studies support the idea that in AD population there are a greater number of repetitions and revisions than in healthy controls.
-However, there are some studies which contradict these findings\cite{Ahmed2013}.
-\subsection{Conclusions}
-This section has looked at a number of different attributes which researchers have cited as being important in the detection of MCI and AD.
-Whilst in some cases there is a broad consensus about a given feature in the vast majority of cases there is some uncertainty.
-Some of the difficulty lies in different approaches used particularly around the criteria for experimental groups which means that it is difficult to be able to compare studies, like for like, as the populations of the participants being studied vary in subtle or obvious ways.
-It is important for any studies moving forward to use comparable standards when it comes to their experimental groups, such that the picture can be made clearer.
+\chapter{Discussion and Future Work}
 
-\section{Mild Cognitive Impairment}
-% Mild cognitive impairment as a diagnostic entity. Petersen (2004)
-Research has noted that there is an observable transitional period between normal ageing and the point where it is possible to diagnose probable AD.
-This transitional period has been described in a range of different terms including Mild Cognitive Impairment (MCI), prodromal dementia and Mild Neurocognitive Disorder (mNCD).
-Traditionally, a diagnosis of probable dementia can only be given once cognitive deterioration has reached a certain point, for example - when a patient scores between 20-24 on the MMSE, this has been given the term Mild Alzheimer's Disease.
-\par 
-In some ways this transitional period is difficult to pinpoint.
-At the start of this transitional period, this may manifest in extremely subtle ways such that it might not be differentiable from cognitive decline associated with aging.
-Equally there is an overlap at the end of this transitional period and the very early stages of clinically diagnosable probably AD.
-\par
-The American Psychiatric Association (APA) recently published the DSM-5 and added the term Mild Neurocognitive Disorder
+\chapter{Conclusions}
 
-% Mild Cognitive Impairment. Petersen (2016)
+\bibliographystyle{ieeetr}
+\bibliography{ThesisRefs}
 
-\subsection{Criteria for Mild Cognitive Impairment due to Alzheimer Disease}
-
-\subsection{}
-
-\subsection{Rates of conversion from MCI to Dementia}
-The American Academy of Neurology (AAN) concluded that MCI is a useful marker to identify since the condition significantly increases the risk of progression to Dementia.
-
-% A Longitudinal Follow-Up of 550 Mild Cognitive Impairment Patients: Evidence for Large Conversion to Dementia Rates and Detection of Major Risk Factors - Espinosa et al (2013)
-The annual conversion of amnestic MCI to AD reported in different studies ranges from 10\% to 40\%.
-
-\section{Online collection of Data}
-
-\section{State of literature into Machine Learning and Natural Language processing techniques}
-Diagnosing dementia through language analysis has a large background in terms of psychological research.
-Increasingly current research has called for the use of machine learning as a way of assisting in the process of diagnosis \cite{Orimaye2017, Boschi2017}.
-This section will look at what areas of natural language processing (NLP) and machine learning (ML) could potentially be used as tools to help in this domain as well as any research that has applied machine learning to this problem.
-\subsection{Natural Language Processing}
-Natural Language Processing as a topic can be defined as the intersection between Machine Learning and Linguistics and looks at a number of language tasks, particularly important in this domain is how to enable a computer to process and analyse large amounts of language data.
-Some other tasks include speech recognition, natural language understanding and natural language generation.
-Generally speaking, Natural Language Processing (NLP) systems mimic the semiotic perspective on language.
-That is to say that we can subdivide NLP processes into different components (see Figure 2.3).
-\begin{enumerate}
-	\item Morphological/Lexical: providing the basic language elements or vocabulary, such as words, their roots and inflections.
-	\item Syntactic: for grouping and sequencing elements within samples of language (usually sentences).
-	\item Semantic: for knowing the meaning of an utterance, usually defined in terms of the 'truth value' of the logical propositions that are thought to be expressed by sentences.
-	\item Pragmatic: for understanding the context and purpose of an utterance.
-\end{enumerate}
-\begin{figure}[H]
-\centering
-\includegraphics[width=240px, height=150px]{images/workflow.jpg}
-\caption{How Natural Language Processing tasks are subdivided.\label{white}}
-\end{figure}
-In order to facilitate these and other tasks, a number of frameworks have been developed that automate some of the more common processing tasks. A brief review of these frameworks follows. 
-\subsubsection{Natural Language Toolkit}
-The Natural Language Toolkit (NLTK) was originally designed as part of a computational linguistics course at the University of Pennsylvania but has evolved to become an open-source framework which includes methods and modules for a wide array of Natural Language tasks (see Table 2.2 for a full list) \cite{Bird2009}.
-\begin{table}
-	\begin{tabular}{ | p{4cm} | p{7cm} | p{1cm} |}
-		\hline
-		\textbf{Language processing task} & \textbf{Functionality} \\ \hline
-		Accessing corpora & standardized interfaces to corpora and lexicons \\ \hline
-		String processing & tokenizers, sentence tokenizers, stemmers \\ \hline
-		Collocation discovery & t-test, chi-squared, point-wise mutual information \\ \hline
-		Part-of-Speech tagging & n-gram, backoff, Brill, HMM, TnT \\ \hline
-		Machine Learning & decision tree, maximum entropy, naive Bayes, EM, k-means \\ \hline
-		Chunking & regular expression, n-gram, named-entity \\ \hline
-		Parsing & chart, feature-based, unification, probabilistic, dependency \\ \hline
-		Semantic interpretation & lambda calculus, first-order logic, model checking \\ \hline
-		Evaluation metrics & precision, recall, agreement coefficients \\ \hline
-		Probability and estimation & frequency distributions, smoothed probability distributions \\ \hline
-		Applications & Graphical concordance, parsers, WordNet browser, chatbots \\ \hline
-		Linguistic Framework & manipulate data in SIL Toolbox format \\ \hline
-	\end{tabular}
-		\caption{\label{tab:table-name}NLTK tasks and functionality}
-\end{table}
-The NLTK has been successfully used in both teaching and research purposes and as part of this provides access to large array of databases and corpora which have potential uses in common tasks.
-
-\subsubsection{CoreNLP}
-Stanford CoreNLP provides a set of human language technology tools.
-This comes as an installable package that requires JAVA and can be interacted with through the command line or by an accompanying API.
-Whilst NLTK is much more oriented to parsing text and generating features, CoreNLP does this and adds more functionality such as a Named Entity Recogniser and sentiment analysis functions.
-At it's centre is the Stanford parser, which parses text and can give the base forms of words, their parts of speech, whether they are names of companies, people, etc., normalize dates, times, and numeric quantities, mark up the structure of sentences in terms of phrases and syntactic dependencies, indicate which noun phrases refer to the same entities, indicate sentiment, extract particular or open-class relations between entity mentions, get the quotes people said, etc.
-\par
-\begin{figure}[H]
-\centering
-\includegraphics[width=350px, height=200px]{images/CoreNLP.png}
-\caption{Depiction of how Core NLP marks up a sentence}
-\end{figure}
-It's intended for CoreNLP to provide tools and frameworks that will enable higher level language analysis whilst remaining domain agnostic.
-
-\subsubsection{Linguistic Inquiry and Word Count}
-The Linguistic Inquiry and Word Count (LIWC) is a piece of software that counts words and assigns these counts to categories \cite{Pennebaker2015}.
-The rationale behind this is that the words we speak inform an observer of the state of mind of an individual.
-A typical example of this might be someone with depression who may typically be more inwardly focused and therefore talking in the first person is common\cite{Al-Mosaiwi2018}.
-The 2015 version of the LIWC has over 90 different categories \cite{Pennebaker2015} and some supporting research suggests that it is effective at identifying themes and associating these themes with mental health difficulties such as depression and anxiety \cite{Sonnenschein2018}.
-\par
-Whilst not necessarily a tool that can parse data and look at trends from a linguistic perspective, a framework like this could be used as another tool to explore themes within a persons speech from a more qualitative angle.
-To the authors knowledge, there is no research that looks at speech changes from this perspective.
-
-\subsection{Traditional methods of Machine Learning}
-In terms of Machine Learning research in this area, a number of researchers have used transcripts based on picture description tasks \cite{Zimmerer2016, Orimaye2017, Mueller2018a, Fraser2015} and have successfully extracted linguistic features that could differentiate between AD and controls.
-\par
-One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the most frequent class.
-They detail several lexical approaches to the problem of detecting and rating AD.
-The approaches they looked at relied primarily on character n-gram techniques but also explore correlation of usage of frequency of different types of speech.
-Their results act as a proof in concept of the utility of using a pure computational approach to the diagnosis of dementia using spontaneous speech.
-The were able to obtain 95\% accuracy dementia vs controls, 70\% accuracy classifying dementia into two categories and 50\% accuracy classifying dementia into four categories.
-They suggest further exploration of characteristics to classify \cite{Thomas2005}.
-\par 
-% Cite CRUM Population-based norms for the Mini-Mental Status Examination by age and educational level
-Bucks et al \cite{Bucks2000} took a more traditional approach using frequency rates per 100 words of various word types such as nouns, verbs, adjectives and pronouns and some measures of lexical richness to attempt to discriminate between healthy older participants and participants with probable AD.
-They found that AD patients had higher mean pronoun rate, adjective rate and verb rate but lower noun rate vs normal older controls \cite{Bucks2000}.
-It's important to note that the clinical group averaged a score on the MMSE of 15 (sd=6.8) which puts the clinical group in the 'moderately impaired range'.
-At this stage, the two groups should be reasonably easy to differentiate in terms of language and other cognitive abilities.
-\par 
-Fraser et al \cite{Fraser2015} built upon the work by Bucks by adding an acoustic analysis element to her analysis using the DementiaBank corpus, as well as significantly expanding the list of linguistic features they used.
-They used features generated from the Stanford parser (Now known as Core NLP) and computed the frequency of the occurrence of different parts of speech which they then normalised by the total number of words in each utterance.
-They also computed ratios between these parts of speech such as pronouns to nouns.
-They also included measures of syntactic complexity such as mean length of sentences and measures of lexical richness such as TTR.
-Their analysis showed was able to achieve 81\% accuracy in distinguishing individuals with AD from controls based on short samples of speech.
-They found that there are four factors which they felt were important in discriminating between these two groups: semantic impairment, acoustic abnormality, syntactic impairment and information impairment.
-\par 
-Zimmerer et al \cite{Zimmerer2016} took a slightly different approach in that they developed the Frequency in Language Analysis Tool (FLAT) which looked at the degree of formulaicity in an individual sample.
-The rationale behind this is that people with impaired cognitive function are able to depend on formulaic sentences for language generation to ease the burden of language generation on cognitive symptoms.
-They were able find that those with probable AD produced more formulaic language than controls.
-\par 
-Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls.
-Their main finding supported n-grams as the most significant predictor.
-One of the criticisms is the use of picture description tasks and n-grams.
-Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
-\par
-
-\subsection{The case for Deep Learning}
-One of the criticisms of traditional learning models is it's reliance on features that are generated for them.
-As we can see from the brief look at some of the research in this field, researchers have taken a number of different approaches with success.
-However, there does not seem to be a consensus.
-As machine learning in general moves away from traditional machine learning models to deep learning techniques, it poses the question can deep learning assist with this problem?
-\par 
-One of the main benefits of deep learning is that it does not rely on pre defined features being fed to the model.
-Instead, deep learning models take raw data in with some amount of preprocessing and generates it's own features.
-This move away from relying on features solves one of the difficulties we have with the current psychological literature, namely that there are some disagreement on the which
-\par 
-One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
-This is the case with all machine learning techniques, but more so with deep learning.
-There have been ways to 'create' more data such as data augmentation.
-
-\subsection{Conclusions}
-We can see that both Natural Language Processing and Machine Learning techniques have a lot to offer, Indeed there has been a lot of research which have used pre-existing datasets to explore this area with promising results.
-One of the difficulties with the current research is the approach of trying to discriminate between those AD and healthy controls.
-This is not necessarily a problem in the real world as it is trivially easy to do for trained clinicians.
-A more interesting, but potentially harder to problem to solve is to discriminate between those with MCI and healthy controls, and more importantly to track their decline over time.
-Further, to date the authors are unaware of any research in which these techniques are applied to newly created samples of language.
-
-\subsection{One Class Classification}
-The prevalence of Dementia in the UK's population is roughly 1 in 14 or 7\%.
-The prevalence of MCI has only briefly been explored, however some research by a team at the Univesity of Leeds indicate that the prevalence of MCI is between 3.2 - 24\%, and found that the prevalence of MCI in people who have self-reported memory problems is almost double (18\%) the amount of do not self report.
-Whilst this means that it is entirely possible to recruit a cohort with MCI and AD, it is comparatively costly in terms of time and resources.
-Given that, from a statistical perspective, this is a highly imbalanced population we can look at the problem as a kind of anomaly detection.
-\par 
-This has some precedent in the field of . In 
-\par 
-One-class classifiers analyze only examples of the majority class (in this case - Healthy Older Adults) to learn a classification boundary that excludes outlier examples.
-When dealing with heavily imbalanced datasets with text classifcation as an objective, the one-class classifier approach has been demonstrated to outperform conventional two-class classification.
-
-\section{Discussion}
-Current state-of-the-art diagnostic measures of AD are invasive (CSF analysis), expensive (neuroimaging), and time-consuming (neuropsychological assessment).
-Furthermore, these measures are limited to speciality clinics and thus have limited accessibility as frontline screening and diagnostic tools for AD.
-More importantly, nonspecialists are often inaccurate at identifying early AD and MCI.
-Thus, there is an increasing need for additional noninvasive and/or cost-effective tools, allowing effective frontline identification of subjects in the preclinical or early clinical stages of AD who could be suitable for monitoring in speciality clinics and for early treatment.
-Implementation of effective screening instruments will allow diagnosis earlier in the course of dementia, even at the point when memory function is still essentially within the normal range.
-This strategy would enable an earlier, and potentially more effective, prevention and treatment of AD with a special focus to preserve cognitive functions.
-\par 
-However the literature has identified a number of challenges when approaching this problem.
-Firstly, the clinical features that combine to meet the diagnostic criteria for Dementia or it's variants are continuous in nature and heterogenous between patients and are also impacted by other variables.
-For example as shown above, cognitive performance is affected in part by a patients educational attainment and a patients ability to live independently is impacted by a their physical health as much as their cognitive health.
-The challenge therefore is to find features that are minimally impacted by other factors, or that can be controlled for by a experimental design such as a matched pairs design to control for educational attainment.
-\par 
-Another challenge lies with the recruitment of suitable individuals who may notice a decline in cognition to the point where we might classify them as having MCI, but these individuals deduce that there is little to no value to admitting there is a problem and seeking help whilst their symptoms are 'manageable'.
-\par
-Finally there is a large amount of variability in the presentations of those with MCI and early dementia, and this is compounded by an similar amount of variability in the criteria researchers have used for experimental groups and the approaches researchers have used to tackle this problem.
-This had led to a confused literature.
-Recently a call for research that has consistent inclusion / exclusion criteria has been made along with some proposed definitions of MCI and it's subgroups \cite{Petersen2014}.
-Researchers have identified the analysis of language impairment as an area of promise to explore in the diagnosis of MCI and early AD and recent developments in natural language processing and machine learning techniques have the potential to assist in this research.
-Indeed, given the increased burden on the diagnosis of MCI and AD on professionals there has been a call to use technology to potentially ease this burden \cite{Boschi2017}.
-A small but growing amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
-\par 
-One of the areas for research to study is a careful examination of the features that are being used to measure language deterioration.
-For example, Zimmerer (2016) \cite{Zimmerer2016} describes connectivity in such as a way that correlates directly with what Mueller (2018) \cite{Mueller2018a} calls Fluency.
-Whilst these are very nuanced measures which differ slightly in the form they take, the sheer range of measures and features being produced make it difficult to organise and explore what is truly going on in those with MCI and early AD..
-Some work needs to be done in producing a consistent list of measures that are validated using existing datasets and can be used for future research moving forward.
-\newline
-\par
-Teng et al suggests that work should focus on the MCI population and concentrate on developing a consensus neuropsychological battery that could yield predictable rates of progression to AD \cite{Teng2009}.
-This, in conjunction with the development of a model of language, sensitive enough to detect subtle deterioration in language use to act as an additional cognitive marker to aid diagnosis could potentially move some way to providing this.
-\par 
-Future research should also be directed towards developing non-intrusive ways of detecting subtle changes in natural language such that any perceived deterioration that could indicate the presence of MCI or AD could be flagged up early.
-Machine learning approaches seem to be the most logical approach for achieving this aim as language could be collected in non-intrusive ways and passed to a machine learning algorithm for preliminary classification.
-Despite the excellent quality of datasets, for example the DementiaBank dataset, being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which some researchers could claim are limited in scope.
-Finally, the recent resurgence in the use of neural networks and deep learning could provide the answer to the confused literature in terms of features.
-A key benefit of deep learning is it's ability to automate the process of feature engineering.
-So there is an opportunity to explore the use of deep learning, to not only develop new features but also validate existing features independently.
-\par
-This area of research is extremely promising in its early results and the impact of successful research would be life changing for both individuals and the health of the worlds aging population in general.
-
-\chapter{Feature Engineering and Classification of DementiaBank Transcripts and Audio}
-
-\chapter{Comparison of different speech to language applications}
-
-\chapter{Development of an application}
-
-\chapter{Techniques in feature engineering - What features matter? Comparing language generated from picture description task to language generated from our task}
-
-\chapter{Conclusions and Future Work}
-With the older population of the world continuing to grow year upon year, Dementia in all it's forms is also going to continue to grow.
-Indeed it is already one of the most common causes of death among the elderly \cite{Prince2015}.
-Whilst a lot of research gone into Alzheimer's disease and other forms of dementia, this research thus far has not yielded positive results in terms of our understanding of the disease or of a potential cure.
-Focus therefore should be on the early, pre-clinical stages of dementia to increase our understanding of the disease and to research potential early interventions in those who are likely to go on to develop dementia.
-\par 
-We looked at areas of research in terms of what is available for testing for MCI at the moment and what language markers have been identified as potentially significant.
-We also reviewed the contributions that Natural Language Processing and Machine Learning techniques could make to the field as well as a brief look at some of the key pieces of research that have used techniques.
-The picture is confusing as there has not been much in the way of consistency in researchers results in either of these areas for those with MCI.
-We propose that the problem is due to the heterogeneity of the MCI population and that simple differentiation of MCI or Healthy may not be possible.
-\par 
-We then looked at a potential longitudinal approach, originally done by Berisha and Liss \cite{Berisha2015}.
-We not only replicated a lot of their findings, despite slightly differing methodologies, but we found other language factors which change with the progression of MCI in our results.
-This is significant as it proves that we can use an individual as his or her own control in the measure of language deterioration and this goes some way to solving the problem of heterogeneity in this population.
-\par
-We then go on to propose more research looking at other available longitudinal data (Task 3), as well as proposing a methodology for collecting language data for analysis (Task 2, 5).
-Finally, we propose a longitudinal study (Task 4) which aims to use our newly created methodology for collecting language data in healthy controls, those with probable AD, those subjective cognitive impairment and those with a diagnosis of MCI to see whether we can detect signs of MCI through language analysis.
-\par
-If this work is successful, we have the opportunity to increase our understanding of the very early stages of AD as well as potentially having a new approach to the diagnosis of MCI and AD in the future.
-
-\bibliographystyle{unsrt}
-\bibliography{Bibliography}
-
-\end {document}
+\end{document}

### 2019-06-20 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
reorganised files

\documentclass[12pt]{article}

\title{Exploring the Presidents Corpus}

\author{
	Jomar Alcantara \\
	Department of Computer Science\\
	School of Engineering and Applied Sciences\\
	Aston University\\
	\texttt{alcantaj@aston.ac.uk} \\
		\and
	Peter Sawyer \\
	Department of Computer Science\\
	School of Engineering and Applied Sciences\\
	Aston University\\
	\texttt{p.sawyer@aston.ac.uk}
		\and
	George Vogiatzis\\
	Department of Computer Science\\
	School of Engineering and Applied Sciences\\
	Aston University \\
	\texttt{g.vogiatzis@aston.ac.uk} \\
		\and
	Cristina Romani\\
	Department of Psychology\\
	School of Life and Health Sciences\\
	Aston University \\
	\texttt{c.romani@aston.ac.uk}
}

\date{\today}

\usepackage{arxiv}

\begin{document}
\maketitle

\keywords{MCI \and Alzheimer's Disease \and Machine Learning \and Diagnosis} 
\bigskip
\begin{abstract}
This is the paper's abstract \ldots
\end{abstract}

\section{Introduction}\label{introduction}

\section{Methodology}\label{methodology}

\section{Results}\label{results}

\section{Discussion}\label{discussion}

\section{Conclusions}\label{conclusions}

\bibliographystyle{unsrt}
\bibliography{main}

\end{document}

### 2019-06-20 Source/Paper 2 - Systematic Review/Systematic Review.tex
changed documents
reorganised files

--- 2019-05-17 Source/Systematic Review/Systematic Review.tex
@@
-\title{Systematic Review}
+\documentclass[12pt]{article}
+
+\title{Applying Natural Language Processing and Machine Learning techniquess to aid in the diagnosis of Mild Cognitive Impairment and Early Dementia - A Systematic Review of the current literature}
+
 \author{
@@
 	School of Engineering and Applied Sciences\\
-	Aston University
+	Aston University\\
+	\texttt{alcantaj@aston.ac.uk} \\
 		\and
@@
 	School of Engineering and Applied Sciences\\
-	Aston University
+	Aston University\\
+	\texttt{p.sawyer@aston.ac.uk}
 		\and
@@
 	School of Engineering and Applied Sciences\\
-	Aston University}
+	Aston University \\
+	\texttt{g.vogiatzis@aston.ac.uk} \\
+		\and
+	Cristina Romani\\
+	Department of Psychology\\
+	School of Life and Health Sciences\\
+	Aston University \\
+	\texttt{c.romani@aston.ac.uk}
+}
 
@@
 
-\documentclass[12pt]{article}
-
-\usepackage[utf8]{inputenc}
-\usepackage{tikz}
-\usetikzlibrary{shapes.geometric, arrows}
-
-\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
-\tikzstyle{io} = [trapezium, trapezium left angle=70, trapezium right angle=110, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=blue!30]
-\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, text width=3cm, draw=black, fill=orange!30]
-\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]
-\tikzstyle{arrow} = [thick,->,>=stealth]
+\usepackage{arxiv}
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{lipsum}
 
@@
 
+\keywords{MCI \and Alzheimer's Disease \and Machine Learning \and Diagnosis} 
+\bigskip
 \begin{abstract}
@@
 Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
+%% This is an important point to fit in.
+ However, the absence of a single test and the complexity of AD create delays in diagnosis.
 
@@
 \section{Results}\label{results}
-In this section we describe the results.
+\subsection{Features of Language}
+\subsubsection{Type token ratio(TTR)}
+Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
+\begin{equation} \label{x1}
+TTR = numberOfUniqueWords / totalNumberOfWords.
+\end{equation}
+\subsubsection{Brunet's Index(W)} % Don't forget to cite the original article.
+Brunet's Index (W) differentiates itself for TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
+\begin{equation} \label{x2}
+W = N^{V(-0.165)}
+\end{equation}
+where N is the total length of the utterance being measured and V is equal to the total vocabulary being used by the subject.
+Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers.
+\newline
+\par 
+\subsubsection{Honore's Statistic (R)} % Don't forget to cite the original article.
+Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation: 
+\begin{equation} \label{x3}
+%% Check to see if this is accurate.
+R = (100 \log N) / (1 - V1/V)
+\end{equation}
+where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
+\subsection{Quantity - Total number of words}
+Several studies report that adults with moderate AD produce fewer words than controls on picture description, however other studies found no differences in total words among groups of controls and patients with MCI or AD.
+Murray and Nicholas et al investigated normal controls, patients with AD and older adults with depression and found no group differences in total words.
+In contrast, Lira 2014 found that controls produced more total words than patients with AD but found no difference between mild and moderate groups.
+\newline
+\par
+\subsection{Syntax and Morphology (Language Form)}
+Syntax can be defined as the rules that govern how words can be combined to form sentences, whilst Morphology is the system that governs the structure of words and the construction of word forms.
+Multiple studies of language decline in dementia included at least one measure of syntax or syntactic complexity.
+Common constructs included words per clause, grammatical form (measures of an appropriate use of syntactic conjections, tenses, conditionals, subordinate clauses and passive constructions), measures of phrase length and proportions of words in sentences.
+Some researchers have explored the use of formulaic language in those with dementia, the theory being that well practiced phrases are less effortful and therefore place low load on the cognitive abilities of those with AD.
+The general hypothesis motivating these studies is that either working memory limitations or semantic memory limitations in AD affect one's ability to use complex constructions.
+\newline
+\par
+\subsection{N-grams}
+One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
+An n-gram is a contiguous sequence of n items from a given sample of text or speech.
+The items can be phonemes, syllables, letters, words or base pairs according to the application.
+For example, given the sequence of words "to be or not to be", this extract is said to contain six 1-gram sequences (to, be, or, not, to, be), five 2-gram sequences (to be, be or, or not, not to, to be), four 3-gram sequences(to be or, be or not, or not to, not to be) and so on.
+This is useful as, given a large portion of text or speech, we can predict the probability of a word being close by to a given word.
+A number of researchers have used n-grams as features.
+One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class.
+Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls.
+Their main finding supported n-grams as the most significant predictor.
+One of the criticisms is the use of picture description tasks and n-grams.
+Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
+\newline
+\par
+Asgari, Kaye and Dodge (2017) \cite{Asgari2017} used another form of word frequency measurement.
+Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words \cite{Pennebaker2015}.
+They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
+\newline
+\par
+\subsection{Mean length of utterance (MLU)}
+Murray found that MLU was not a distinguishing factor among health adults, adults with depression and adults with AD.
+Ripich et al found a decrease in MLU in adults with severe AD over time, and this was supported by findings of Le et al in their studies of authors \cite{Le2012} \newline
+\par 
+\subsection{Proportion of verbs to nouns plus verbs}
+Kave and Levy used a verb index to capture syntactic complexity and found that adults with AD expressed the same amount of verbs to nouns plus verbs as adult controls. \newline
+\par 
+\subsection{Syntactic Complexity - Composite measures of MLU, syntactic errors and verbs}
+Ahmed et al, and Ahmed, Haigh et al found differences in syntactic complexity between adults with MCI and controls, and between MCI and moderate AD stages.
+The differences in syntactic complexity were not significant when individual measures were tested, but were apparent using a composite score consisting of MLU, words in sentences, syntactic errors, nouns with determiners, and verbs with inflections.
+\newline
+\par 
+\subsection{Formulaic Language}
+Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
+They found that there were four factors which informed the classification of participants as either healthy or AD.
+These four factors were semantic impairment, acoustic abnormality, syntactic impairment and information impairment and were based on existing measures of semantic and syntactic complexity.
+Zimmerer (2016) \cite{Zimmerer2016} looked at whether language was more formulaic in those suffering from AD.
+He proposed that those who suffer from AD rely on formulatic sentences, for example 'Noun-Verb-Noun', and this is done to reduce language complexity.
+He noticed a significant difference in the use of formulaic sentences between AD and Healthy Controls.
+\newline
+\subsection{N-grams}
+One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
+An n-gram is a contiguous sequence of n items from a given sample of text or speech.
+The items can be phonemes, syllables, letters, words or base pairs according to the application.
+For example, given the sequence of words "to be or not to be", this extract is said to contain six 1-gram sequences (to, be, or, not, to, be), five 2-gram sequences (to be, be or, or not, not to, to be), four 3-gram sequences(to be or, be or not, or not to, not to be) and so on.
+This is useful as, given a large portion of text or speech, we can predict the probability of a word being close by to a given word.
+A number of researchers have used n-grams as features.
+One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class.
+Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls.
+Their main finding supported n-grams as the most significant predictor.
+One of the criticisms is the use of picture description tasks and n-grams.
+Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
+\newline
+\par
+Asgari, Kaye and Dodge (2017) \cite{Asgari2017} used another form of word frequency measurement.
+Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words \cite{Pennebaker2015}.
+They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
+\newline
+
+\section{Machine Learning methods}
+\subsection{Traditional Machine Learning methods}
+\subsection{Deep Learning methods}
+\subsubsection{Deep language space neural network for classifying mild cognitive impairment and Alzheimer-type dementia - Orimaye, Wong and Wong (2018)}
+In this paper, Orimaye et al use deep-deep neural networks language models (D2NNLM) to learn linguistic changes that distinguish the language of patients with MCI and AD-type dementia from the healthy controls using higher order n-grams.
+An ordinary DNNLM uses lower order n-gram N-dimensional sparse vectors as descrete feature representations to train the neural network with multiple hidden layers.
+
+\subsubsection{Deep Neural Network Language Models}
+
+\subsection{Aided Diagnosis of Dementia Type through Computer-Based Analysis of Spontaneous Speech - Jarrold, Peintner, Wilkins, Vergryi, Richey, Gorno-Tempini and Ogar}
 
@@
 
-\subsection{Deep language space neural network for classifying mild cognitive impairment and Alzheimer-type dementia - Orimaye, Wong and Wong (2018)}
-In this paper, Orimaye et al use deep-deep neural networks language models (D2NNLM) to learn linguistic changes that distinguish the language of patients with MCI and AD-type dementia from the healthy controls using higher order n-grams.
-An ordinary DNNLM uses lower order n-gram N-dimensional sparse vectors as descrete feature representations to train the neural network with multiple hidden layers.
-
-\subsubsection{Deep Neural Network Language Models}
-
 \section{Discussion}\label{discussion}
+\subsection{Discussion of the current evidence}
+\subsection{Methodological Issues}
+\subsection{Limitations}
+\subsection{The future of the field}
 
@@
 
-\bibliographystyle{abbrv}
+\bibliographystyle{unsrt}
 \bibliography{main}

### 2019-06-20 Source/Thesis/Chapter1.tex
reorganised files

--- 2019-05-28 Source/Thesis/Thesis.tex
@@
-\title{Early Detection of MCI that progresses to dementia - An Interdisciplinary Approach}
-\author{
-        Jomar Alcantara \\
-        Department of Computer Science \\
-        School of Engineering and Applied Sciences \\
-        Aston University\\
-        Birmingham, United Kingdom
-}
-\date{\today}
-
-\documentclass[twopages, a4paper]{book}
-
-\begin{document}
-\maketitle
-\newpage
-% Abstract
-\section*{Abstract}
-\newpage
-% Acknowledgements
-\section*{Acknowledgements}
-\newpage
-\tableofcontents
-\newpage
-\listoffigures
-\listoftables
-\chapter{Introduction}
-Dementia has been identified as one of the one of the biggest global health and social care challenges facing the world tday.
+\section{Introduction}
+Dementia has been identified as one of the one of the biggest global health and social care challenges facing the world today.
 A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
@@
 \par
-\chapter{Background and Related Work}
 
-\chapter{Systematic Review}
+\begin{equation}
+    \label{simple_equation}
+    \alpha = \sqrt{ \beta }
+\end{equation}
 
-\chapter{Language Data Collection - Developing a consensus}
+\subsection{Subsection Heading Here}
+Write your subsection text here.
 
-\chapter{Pipeline Development}
-
-\chapter{Analysis of DementiaBank dataset}
-
-\chapter{Pilot Study}
-
-\chapter{Discussion and Future Work}
-
-\chapter{Conclusions}
-
-\bibliographystyle{ieeetr}
-\bibliography{ThesisRefs}
-
-\end{document}
+\section{Conclusion}
+Write your conclusion here.

### 2019-06-20 Source/Thesis/Chapter2.tex
reorganised files

--- 2019-06-20 Source/Thesis/Chapter1.tex
@@
 \section{Introduction}
-Dementia has been identified as one of the one of the biggest global health and social care challenges facing the world today.
-A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
-The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
-\par
-In 2009, the UK's Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
-A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two distinct areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
-As our aging population increases, the potential burden on health care and social services will increase and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
-At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017}.
-I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
-\par
+Here is the text of your introduction.
 

### 2019-06-20 Source/Thesis/Thesis.tex
thesis update
reorganised files

@@
+\documentclass[twopages, a4paper,11pt]{book}
+
 \title{Early Detection of MCI that progresses to dementia - An Interdisciplinary Approach}
@@
 
-\documentclass[twopages, a4paper]{book}
+\begin{document}
 
-\begin{document}
 \maketitle
@@
 \listoftables
+
 \chapter{Introduction}
-Dementia has been identified as one of the one of the biggest global health and social care challenges facing the world tday.
-A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
-The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
-\par
-In 2009, the UK's Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
-A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two distinct areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
-As our aging population increases, the potential burden on health care and social services will increase and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
-At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017}.
-I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
-\par
+\include{Chapter1}
+
 \chapter{Background and Related Work}
+\include{Chapter2}
 
-\chapter{Systematic Review}
+\chapter{Systematic Review of NLP and Machine Learning Research}
+\include{Chapter3}
 
-\chapter{Language Data Collection - Developing a consensus}
+\chapter{Delphi Methodology and developing consensus on how best to collect language samples using technology}
+\include{Chapter4}
 
-\chapter{Pipeline Development}
+\chapter{Development of a pipeline that processes language data accurately}
+\include{Chapter5}
 
 \chapter{Analysis of DementiaBank dataset}
+\include{Chapter6}
 
-\chapter{Pilot Study}
+\chapter{Pilot study of the methodology developed}
+\include{Chapter7}
 
-\chapter{Discussion and Future Work}
+\chapter{General Discussion}
+\include{Chapter8}
 
 \chapter{Conclusions}
-
-\bibliographystyle{ieeetr}
-\bibliography{ThesisRefs}
+\include{Chapter9}
+\bibliographystyle{unsrt}
+\bibliography{main.bib}
 

### 2019-07-08 Source/Defense/defense.tex
organised and updated

--- 2018-04-22 Defense/defense.tex
@@
 	Answer - Research from the University of Lancaster suggested that this was the most effective test for differentiating between controls, MCI and full blown AD.
+
+	Question - Why Free Speech?
+	Answer - Use of Speech Analyses within a Mobile Application for the Assessment of Cognitive Impairment in Elderly People by Toth suggests that Tthe fluency and free speech tasks obtain the highest accuracy rates of classifying AD vs.
+MD vs.
+MCI vs.
+SCI.
+(AD - Alzheimer's Disease, MD - Mixed Dementia, MCI - Mild Cognitive Impairment, SCI - Subjective Cognitive Impairment) Using the data, we demonstrated classification accuracy as follows: SCI vs.
+AD = 92\% accuracy; SCI vs.
+MD = 92\% accuracy; SCI vs.
+MCI = 86\% accuracy and MCI vs.
+AD = 86\%.
 \end{document}

### 2019-07-08 Source/Paper 2 - Systematic Review/Systematic Review - Protocol.tex
changed documents
organised and updated

--- 2019-05-09 Source/Systematic Review/Systematic Review - Protocol.tex
@@
 \section{Introduction}
-Dementia has been identified as one of those fast growing difficulties facing the world.
+Dementia has been identified as one of the fastest growing difficulties facing the world.
 A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
@@
 \begin{enumerate}
-	\item RQ1: Which ML and MS techniques are being used in the dementia and comorbidities research?
-	\item RQ2: What data characteristics (variables, determinants and indicators) are being considered when applying the ML or/and MS techniques (physiological, demographic/social, genetics, lifestyle etc)?
-	\item RQ3: What are the goals of the studies that employ ML or MS techniques for prognosis of dementia and comorbidities?
-	\item RQ4: How is data censoring being handled in the studies?
+	\item RQ1: Which NLP and ML techniques are being used in dementia research?
+	\item RQ2: What features / data characteristics of text (variables, determinants and indicators) that are considered when applying the ML techniques (n-grams, PoS Tagging etc)?
+	\item RQ3: What are the goals of the studies that employ NLP / ML techniques for prognosis of dementia?
+	\item RQ4: Do the studies focus on time as factor??
 	\item RQ5: Do the studies focus on individuals or populations?

### 2019-07-08 Source/Paper 2 - Systematic Review/Systematic Review.tex
organised and updated

@@
 
+This yielded 25 studies for review, 43 papers were reviewed.
+
 \subsection{Study selection}
-A total of 2599 papers were identified through the searches conducted above.
-Papers that were duplicated were removed and Papers were then evaluated by the JA based on the inclusion and exclusion criteria defined previously (see Table 6).
+A total of 1490 unique papers were identified through the searches conducted above.
+Each paper was initially reviewed by just the title and the abstract.
+This yielded a total of 67 potential papers.
+Papers were then evaluated by the JA based on the inclusion and exclusion criteria defined previously (see Table 6).
 Where JA could not reach a decision, PS and GV were consulted and majority vote on inclusion was conducted.
-Each paper was initially presented with just the title and the abstract.
 

### 2019-07-08 Source/Thesis/Chapter1.tex
organised and updated

@@
 \section{Introduction}
-Dementia has been identified as one of the one of the biggest global health and social care challenges facing the world today.
-A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
-The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
+Alzheimer's Disease (AD) is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
+Whilst a definitive diagnosis of Dementia can only be produced at post-mortem, there are a number of clinical indicators from psychological perspective that can indicate dementia is present.
+From a clinical psychology perspective, those who have dementia demonstrate cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}.
+In addition, these symptoms are often accompanied by emotional problems such as depression and behavioural difficulties.
+\par
+AD and other forms of dementia affect a significant proportion of the geriatric population in the world today and is currently the sixth leading cause of death in the US and was named the leading killer of women in the UK.
+According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
+This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population \cite{AlzheimersSociety2014}.
+From a financial perspective, they estimate an annual spend of £4.3 billion of which approximately £85 million is spent solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
+Globally, this picture is a lot bleaker.
+Another report by Alzheimer's Disease International suggests that in 2015 there were 46 million people with a diagnosis of dementia globally and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
+\par
+Despite this growing problem, at present there are no drugs that improve the prognosis of those suffering with AD.
+All the drugs that are on the market are designed to manage symptoms.
+Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{AlzheimersSociety2014}.
+Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place.
+It is therefore important to focus AD at it's earliest stages which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.
+\par
+Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent.
+Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time and alter cognitive function until a threshold is reached and clinical symptoms become more obvious \cite{Nestor2006}.
+The case of Iris Murdoch, who had a confirmed diagnosis of dementia, illustrates this theory well.
+Le et al \cite{Le2011} found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel when, it is theorized, the symptoms of AD manifested themselves more significantly.
+If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
+One of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of dementia.
+Albert and his team have worked to define clinical criteria which professionals can use to diagnose MCI due to AD and differentiate this from age-associated memory impairment and age-associated cognitive decline.
+They note that the diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}.
+\begin{enumerate}
+	\item Concern regarding a change in cognition - A person or an informant should express concern that there is a change in cognitive ability in comparison to previous level of performance.
+	\item Impairment in one or more cognitive domains - There should be evidence of lower performance in one or more cognitive domains beyond what would be expected of a person given their age and education.
+	\item Preservation of independence in functional - Whilst persons with MCI are expected to be able to maintain independence, it is common to experience mild problems in complex functional tasks which they may have been able to perform previously.
+This might mean that they take more time or be less efficient at completing these tasks, or it may be that may make more mistakes.
+	\item Not demented - The deterioration should be mild to the point that there is no significant loss of functioning in social or occupational contexts.
+\end{enumerate}
+In addition to meeting the above criteria, a clinician must rule out other conditions or factors that could account for the decline in cognition with the goal to increase the likelihood that the underlying cause of this decline is dementia.
+\par 
+Many researchers have studied the early detection of MCI.
+These studies usually follow two main approaches.
+The analysis of biomarkers and the examination of patients who have demonstrated decreasing cognitive abilities.
+The first approach yields reliable results in the detection of AD in its moderate and advanced states but does not perform well during the early stages of the disease.
+The second approach has gained more attention in recent years, due to the fact that in clinical practice it has shown promise in the early detection of AD.
+In addition, the analysis of the decline of cognitive abilities is comparatively inexpensive and less invasive that the first approach which commonly requires the collection of a sample of cerebro-spinal fluid which is painful for the patient involved.
+\par
+One of the most common ways in which clinicians traditionally make an early diagnosis of dementia is through the use of the Mini Mental State Examination (MMSE) \cite{Folstein1975}.
+The MMSE is a brief questionnaire consisting of eleven questions which tests cognitive aspects of mental function and requires only 5-10 minutes to administer \cite{Folstein1975}.
+The MMSE is chosen due to it's effectiveness at assessing a person's cognitive mental state at a specific point in time, as well as being as sensitive to changes as a more detailed and complex assessment such as the Wechler Adult Intelligence Scale \cite{Folstein1975}.
+Whilst the MMSE is useful as a brief screening tool it has it's limitations.
+The MMSE was not specifically created to screen for dementias and therefore does not interrogate key aspects of cognitive impairment known to be affected in dementia.
+It also has limited value in assessing under-educated subjects and a meta-analysis on the effectiveness of the MMSE as a diagnostic tool for dementia showed that it's accuracy was low (sensitivity between 78.4\% and 85.1\% and specificity between 81.3\% and 87.8\%).
+As the MMSE has been shown to have low accuracy specifically in the diagnosis of dementia, it becomes necessary for professionals to employ the use of other tools or measures such as the Free Cued Selective Reminding Test (FCSRT) \cite{Grober2010} or the Montreal Cognitive Assessment (MoCA) \cite{Davis2015}.
+\par 
+These tests have the benefits of being much more accurate at diagnosing cognitive impairment and discriminating between dementia and other types of cognitive impairment at the cost time and training of psychological professionals such as clinical psychologists in administering these tests.
+However, the utility of diagnosing dementia at the point where clinical intervention is warranted is limited because at this stage both psychological and pharmacological interventions have been shown to not be effective \cite{Prince2015}.
+In order to further our understanding of the progression of dementia it is important to detect the signs of dementia before they are clinically apparent.
+\par
+The two main ways in which diagnosis is performed is through assessment of memory and language.
+Tests of memory are classically among the most accurate ways of diagnosing dementia, however these tests suffer from the same reliance on clinicians to administer these tests in a clinical setting.
+Language however is a lot easier to collect and can be done in more naturalistic settings.
+As with memory, these tests can be done over time and would be able to chart a patients language degeneration over time.
+Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on measure decline in the use of language in those with AD.
+There are a number of difficulties to watch out for with this approach.
+There are a wide number of factors that are involved in language degeneration in the elderly, and consequently there will be an expected amount of variability between subjects.
+The administration of such tests may induce nervousness and discomfort which may impact performance, and also repeatedly administering the same language tests for differences over time may be confounded by improved performance at tasks via practice effects.
+However, there is enough promise in this approach such that it could help further our understanding of the disease, it's progression and the parts of the brain affected in the early stages.
+\par
+According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with mild dementia suffer from noticeable word finding difficulty.
+They may substitute general terms for more specific terms and may avoid the use of specific names of acquaintances.
+There may be grammatical errors involving subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc.
+Those who have progressed from Mild to Major depression also have difficulties with expressive or receptive language.
+They will often use general-use phrases such as "that thing" and "you know what I mean" and prefers general pronouns rather than names.
+With severe impairment, sufferers may not even recall names of closer friends and family.
+Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur.
+Echolalia (meaningless repetition of another person's spoken words) and automatic speech typically precede mutism.
+With the wide range of deficits someone with AD can suffer, it makes sense to try to categorise these deficits in some way.
+\par
+One of the most famous pieces of research on the topic of language decline in dementia was by Berisha and Liss who examined speeches and public interviews of former US president Ronald Reagan \cite{Berisha2015}.
+They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease.
+It turned out later to be the case that he had Alzheimer's Disease.
+Another classical study by Snowdon et al looked at whether linguistic ability in early life was associated with cognitive function and AD in later life \cite{Snowdon1996} .
+They found that idea density (defined as the number of expressed propositions divided by the number of words) was a key predictor in predicting whether nuns would go on to develop AD in later life.
+They found that those who would go on to develop AD all had low idea density in early life and they found no AD present in those with high idea density in early life.
+As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are extremely variable and can differ from patient to patient as the disease progresses.
+The consensus among researchers that this language degeneration is typically accelerated by the presence of dementia \cite{Berisha2015} and that a potential indicator of dementia is the rate of change in which the decline occurs relative to a fixed point in time rather than a comparison across a cohort of individuals.
+\par
+Emery \cite{Emery2000} completed a literature review looking at all the potential language deficits that could exist in those with AD and / or MCI.
+She divided these deficits into four levels of language: Phonology, Morphology, Syntax and Semantics.
+She proposed that language and the processes involved in language are hierarchical in nature and that language moves from simple units of construction (Phonology and Morphology), and build layers of complexity and sophistication (Syntax and Semantics).
+She found that people with AD generally had intact Phonology and Morphology but more impaired Syntax and Semantics.
+She asserted that the language forms we learn last are the first to deteriorate as we generally learn language in small simple units initially and build syntax and complexity as we are more comfortable with language.
+However it is important to note that different variants of dementia show different deficits in terms of language productions.
+Regardless, dividing language in this way is useful as it allows us to detect deteriorations in different parts of language usage and therefore may provide a way of discriminating between different forms of dementia.
+\par
+It is clear from both the clinical diagnostic criteria and supporting research that language is impacted in those with AD.
+However, whilst there is a move towards research aimed at looking more specifically at MCI we currently lack the measures that are sensitive enough to detect MCI.
+Given that we know language is affected before a clinical diagnosis of dementia is usually made \cite{Berisha2015, Snowdon1996, Le2011}, it makes sense to explore whether language on it's own can provide markers that may indicate a cognitive impairment that could progress to dementia.
+The field of machine learning and natural language processing has been suggested as a way to improve the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline \cite{Boschi2017}.
 \par 
 In 2009, the UK's Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
-A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two distinct areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
-As our aging population increases, the potential burden on health care and social services will increase and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
-At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017}.
+A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
+As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages \cite{Prince2015} and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
+At the present time diagnostic assessment is typically conducted at memory clinics by trained clinicians \cite{Boschi2017} and whilst it is not possible to provide a definitive diagnosis until after death.
+
 I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
 \par
-
-\begin{equation}
-    \label{simple_equation}
-    \alpha = \sqrt{ \beta }
-\end{equation}
-
-\subsection{Subsection Heading Here}
-Write your subsection text here.
-
-\section{Conclusion}
-Write your conclusion here.
+There is a large body of research that looks at the decline in language in those with MCI and AD \cite{Taler2008, Boschi2017}.
+However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD \cite{Taler2008, Boschi2017}.
+Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
+Another area of focus for research of this nature is the process of collecting appropriate language samples.
+Whilst collecting samples of language is comparatively unintrusive, researchers recognise that these samples require a rich sample of language that potentially cannot be generated by tasks such as the picture description task.
+Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD.
+There is some evidence to support this.
+Berisha et al \cite{Berisha2015}, has shown through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control.
+\par
+The potential impact of this research in this area is immense.
+Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease.
+Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
+\par
+In chapter 2 Background and Related Work
+\par 
+In chapter 3 Systematic Review of NLP and Machine Learning Research
+\par 
+In chapter 4 Delphi Methodology and developing consensus on how best to collect language samples using technology
+\par 
+In chapter 5 Development of a pipeline that processes language data accurately
+\par 
+In chapter 6 Analysis of DementiaBank dataset
+\par
+In chapter 7 Pilot study of the methodology developed
+\par 
+Finally, in chapter 8.
+I discuss conduct a general discussion of the results of my research.
+I also think about the strengths and weaknesses of my work and suggest ways in which the research have been improved.
+Lastly, I look at a number of areas for future work which can build on this research.

### 2019-07-08 Source/Thesis/Chapter2.tex
organised and updated

@@
-\section{Introduction}
-Here is the text of your introduction.
+\section{Existing Neuropsychological Measures of Cognitive Impairment, and Repeatable Battery of Neuropsychological Tests}
+In terms of standardised cognitive tests there are two main aims.
+Does the test distinguish accurately between normal aging, MCI and AD (diagnostic utility) and does the test distinguish between those individuals with MCI who will then go on to develop AD and those individuals with MCI who don't then go on to develop AD (prognostic utility).
+This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their performance in terms of both diagnostic and prognostic utility.
+\par 
+\subsection{Repeatable Battery for the Assessment of Neuropsychological Status}
+The Repeatable Battery for the Assessment of Neuropsychological Status (RBANS) was originally developed as an assessment tool for dementia, specifically looking at detecting a characterizing very mild dementia \cite{Randolph1998}.
+The authors felt that there was a shortfall of appropriate measures that were sensitive enough to milder impairments as well as a number of other shortcomings of existing tests.
+They met a number of design goals for this new battery of tests that addressed these shortcomings.
+The RBANS consists of a number of sub-tests across five distinct domains.
 
-\begin{equation}
-    \label{simple_equation}
-    \alpha = \sqrt{ \beta }
+\begin{enumerate}
+	\item Immediate Memory 
+	\begin{enumerate}
+		\item{List Learning}
+		\item{Story Memory}
+	\end{enumerate}
+	\item Visuospatial / Constructional
+	\begin{enumerate}
+		\item{Figure Copy}
+		\item{Line Orientation}
+	\end{enumerate}
+	\item Language
+	\begin{enumerate}
+		\item{Picture Naming}
+		\item{Semantic Fluency}
+	\end{enumerate}
+	\item Attention
+	\begin{enumerate}
+		\item{Digit Span}
+		\item{Digit Coding}
+	\end{enumerate}
+	\item Delayed Memory
+	\begin{enumerate}
+		\item{List Recall}
+		\item{List Recognitions}
+		\item{Story Memory}
+		\item{Figure Recall}
+	\end{enumerate}
+\end{enumerate}
+
+However, whilst the authors claim that the RBANS is adequately sensitive in person's with MCI \cite{Randolph1998} and there is some research to support this assertion \cite{Karantzoulis2013}, other research points out that the RBANS has poor sensitivity in detecting MCI \cite{Duff2010}.
+Another drawback of this battery is the lack of executive function measures and object naming tasks.
+Research has shown that the RBANS has good test-retest reliability and convergent validity.
+However, Duff et al warned that caution should be exercised when using the RBANS in a MCI population as it has lower sensitivity in this population \cite{Duff2010}.
+
+\subsection{Digit Span Test}
+The Digit Span Test (DST) is predominantly used to measure a person's working memory capacity, specifically the capacity used to store and recall numbers.
+A participant is presented with a series of numbers of fixed length and is asked to recall those numbers in normal or reverse order.
+The series of numbers gets progressively longer until such time as a participant fails three or more times out of eight presentations.
+\par
+Research has consistently shown that MCI patients have a significantly lower digit span score in both normal and reverse order versions of this test and a study by Muangpaisan has shown that the reverse order version of this test can, to some degree, predict a diagnosis of MCI \cite{Muangpaisan2008}.
+Muangpaisan also found that Age, Gender and Education have an impact on the performance of the tests.
+Emrani et al found that the DST revealed no differences between specifically amnestic MCI and controls, but could differentiate between Mixed MCI and the other groups with mixed MCI recalling fewer correct responses than other groups.
+Notably, there was an attenuated recency effect in those with mixed MCI.
+\cite{Emrani2018}.
+Kessels identifies that that there are working memory deficits in MCI patients and these worsen with AD patients.
+\cite{Kessels2011} He also identified that both MCI and AD have impaired performance on all three conditions of the digit span test.
+No differences were found between forward and backward conditions in any of the groups.
+However, available tests may not detect subtle impairments \cite{Kessels2015}.
+\subsection{Rey Auditory-Verbal learning test}
+Rey's Auditory Verbal Learning Test (RAVLT+) looks a wide range of neuropsychological processes including short-term auditory-verbal memory, learning and retention of information.
+Participants are given a list of 15 unrelated words, repeated over five different trials and are asked to repeat.
+Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then repeat it again after 30 minutes \cite{Schmidt1996}.
+\par
+Several studies have shown that an impairment in RAVLT score reflect well the underlying pathology caused by AD.
+Thus making the RAVLT an effective early marker to detect AD in persons with memory complaints.
+Moradi investigated to what extent the RAVLT scores are predictable based on MRI data using machine learning approaches, as well as to find out what the most important brain regions are for the estimation of RAVLT scores \cite{Moradi2017}.
+They found a highly significant cross validated correlation between the estimated and observed RAVLT immediate and RAVLT Percent Forgetting.
+Further, they found that the conversion of MCI subjects to AD in 3-years could be predicted based on either observed or estimated RAVLT scores with an accuracy comparible to MRI-based biomarkers \cite{Moradi2017}.
+Another study by Schoenberg found that RAVLT to best distinguish patients suspected of Alzheimer's disease from the psychiatric comparison group \cite{Schoenberg2006}.
+\subsection{Digit Symbol Substitution Test}
+The Digit Symbol Substitution Test (DSST) involves a key consisting of the numbers 1-9 and a corresponding unique symbol.
+Below this key is a series of numbers from 1-9 in a randomized order and repeated multiple times.
+The participant is asked to fill in the corresponding symbol for each number.
+The task requires that the participant move between the key and the randomized sequence such that  they may retrieve the correct answer from the key, hold this in short-term memory and transcribe the key in the appropriate place.
+\par 
+Among those with no disorder in cognition, mobility and mood, being in the lowest DSST quartile compared to the highest was associated with nearly twice the odds of developing one or more clinical or subclinical disorders.
+Associations were stronger for incident clinical disorders in cognition.
+Slower psychomotor speed may serve as a biomarker of risk of clinical disorders, mobility and mood.
+While in part attributable to vascular brain disease, other potentially modifiable contributors may be present \cite{Rosano2016}.
+Further studying the causes of psychomotor slowing with ageing might provide novel insights into age-related brain disorders.
+Pascoe compared patients with PD with Normal Cognition (PD-N) with those with PD and MCI (PD-MCI) and healthy participants.
+PD-MCI participants achieved significantly lower scores than other groups in the DSST task \cite{Pascoe2018}.
+\subsection{Verbal Fluency}
+There are a number of tests that characterise this category of verbal fluency, but generally fall into two categories.
+Letter (Phonemic) fluency involves the generation of as many words as possible which begin with a specified letter.
+Category fluency involves the generation of as many words as possible that fall into a specfied category.
+Both these tasks impose demands on a number of different cognitive processes namely, executive function, verbal retrieval and recall, giving appropriate answers while monitoring previous answers and inhibit inappropriate responses.
+However, these two tasks require different strategies when attempting them.
+Letter fluency relies on search strategies based on lexical representations whereas category fluency requires a search for semantic extensions of a superordinate term, meaning that semantic associations within the lexicon must be intact in order for the task to be carried out successfully.
+\par  
+There have been numerous studies which have documented the impact that AD has on verbal fluency tests in both categories.
+A review carried out by Henry et al \cite{Henry2004}, found that performance in both letter fluency and category fluency was impaired in those with AD vs controls, but found a larger effect for tests of semantic fluency.
+\subsection{Naming Tests}
+Word-finding difficulty is a common symptom of AD and these deficits usually occur during the early stages of the disease progression.
+As such , a test of a patients ability to find words (known as confrontation naming) is a common way to measure cognitive decline.
+One common way to do this is the Boston Naming Test (BNT) which comprises 60 items on a spectrum of very frequent to very infrequent.
+This has been reduced subsequently to two thirty item versions and four fifteen item versions, which correlate significantly with the original sixty item version and the benefit of the shorter versions of the test is that it facilitates testing of individuals with AD who may suffer from fatigue or limited attention span \cite{Williams1989}.
+\par 
+Willers et al \cite{Willers2008} studied Twenty aMCI patients, twenty AD and 21 controls matched by age, sex and education level.
+They found AD patients obtained significantly lower total scores on the BNT than aMCI patients and controls.
+aMCI patients did not obtain significant differences in total scores but showed significantly higher semantic errors compared to controls.
+Semantic processing is impaired during confrontation naming in aMCI.
+\par 
+Vadikolias investigated the impact of education on naming tests \cite{Vadikolias2012} and found that higher educational attainment in aMCI subjects were correlated with better performance in verbal and non-verbal tasks during repeated examinations over 1-year.
+Subjects with a lower level of education performed worse than patients with a high level of education who presented a more stable clinical score.
+The explanation for this is the idea of a 'cognitive reserve' in participants with a higher education such that this provides a buffer that, while not preventing the physiological symptoms of AD, can potentially delay the clinical onset of cognitive symptoms that characterise AD.
+This theory is supported by Snowdon who found a relationship between early life linguistic ability and the density of neurofibrillary tangles in his nun study \cite{Snowdon1996}.
+Whilst these studies and tests provide evidence that there are word finding difficulties in those with MCI, on their own they do not provide sufficient ability to diagnose MCI or provide an prognosis of disease progression.
+\subsection{Rey-Osterrieth Complex Figure Task}
+The Rey-Osterrieth Complex Figure (ROCF) is a task widely as a test of visuo-spatial skill and visual memory.
+The task, which was originally designed by Rey (1944) and standardised by Osterrieth (1944) \cite{Rey1941, Osterrieth1944}, asks a participant to copy a complex geometrical figure (known as the immediate copy condition and to recall and reproduce the figure from memory without warning (known as the delayed recall condition).
+In the immediate copy condition, the complexity of the figure requires an integrative cognitive ability.
+The reproduction of such a complex structure involves processes such as planning and organizational strategies that are related to executive functions.
+In order to make this task repeatable, Taylor et al designed a comparable set of figures which have proven to be of equivalent difficulty \cite{Taylor1969, Strauss1990}.
+\par
+Salvadori found that patients with vascular MCI had a worse performance in the immediate copy of the ROCF compared to individuals with degenerative MCI, despite their significant impairment in terms of general cognitive status and visual memory \cite{Salvadori2018}.
+Evidence shows that patients with disorders that possibly involve attention and executive functions are characterized by a more disorganised approach when copying the ROCF compared to controls.
+One of the difficulties with the presentation of this task in a repeated battery will be the fact that it turns from an incidental memory task (it is a memory task but the participant is not forewarned that they will be required to memorise the picture) into an intentional memory task (given the previous exposure to the task, it would be expected that a participant would preempt the delayed recall portion of the test and spend more time attempting to memorise the details) \cite{Teng2009}.
+\subsection{Hayling Sentence Completion Test}
+Inhibitory deficits are a common in all stages of dementia.
+This is usually tested using Stroop test, however this has the drawback of lacking ecological validity.
+Therefore researchers have moved towards using the Hayling Sentence Completion Test (HCST) which uses skills such as word retrieval as well as the ability to inhibit ones responses where appropriate and is therefore a much more ecologically valid task.
+There are two parts to the HCST.
+In the first part, participants have to complete a sentence by providing a word that best fits the given sentence (this is known as the initiation condition).
+In this second part, participants have to complete sentences by inhibiting an impulse to give the word that best first the sentence as in the first part, and producing a semantically unconnected word.
+Performance in both conditions is measured by the time taken for the participant to initiate a response, and in the inhibition condition also by the correctness of the word.
+\par 
+Martyr et al compared healthy controls with patients with dementia and patients with Parkinson's disease \cite{Martyr2017}.
+They found that a high proportion of Category A errors (producing a word that fits the sentence when instructed otherwise) was a factor in performance loss for participants with dementia.
+Findings suggest that the HSCT may be sensitive to verbal suppression deficits and may provide insight into inhibitory control in participants with dementia.
+Patients with Dementia were significantly slower than controls in the initiation and inhibition conditions vs healthy controls, and slower than patients with Parkinson's disease in the initiation but not the inhibition condition.
+\subsection{Grooved Pegboard Test}
+The Grooved Pegboard Test (GPT) was originally designed to cover a variety of different psycho-motor functions including hand-eye coordination and motor speed.
+However, some studies have shown a correlation of performance in the GPT and measures of cognitive performance such as the Montreal Cognitive Assessment.
+\par 
+Bezdicek found that the GPT predicted performance on the MoCA and concluded that in addition to being a measure of motor skills, there were results that showed that the GPT could also provide information about a participants cognitive skills \cite{Bezdicek2014}.
+Whilst this was specifically using a population of patients with Parkinson's disease, given the MoCA and GPT have shown degree of correlation, there remains an opportunity for research that explores the use of the GPT with MCI and Early AD populations.
+\par 
+Darweesh et al, used the Purdue Pegboard Test (PPT) to assess manual dexterity in a healthy older adult population and followed their participants for between eight and twelve years until an indication of the onset of a neurodegenerative disease \cite{Darweesh2017}.
+In this time 227 (4\%) of their participants went on to develop a diagnosis of dementia.
+They noted that higher PPT scores were associated with lower risk of incident neurodegnerative disease and noted significant associations of PPT scores with all forms of dementia and this potentially highlights a deterioration of motor function in the pre-clinical phase of dementia.
+\subsection{Visual Search}
+The visual search task requires participants to be watching a computerised display in which a target appears either by itself or surrounded by other elements which are used as a distraction.
+These other elements can be vary in similarity to the target.
+This is inherently a measure of attention shifting and processing of various similarity and as one would expect, the visual search time is increased in the presence of distractors.
+This effect is magnified in participants who have a cognitive impairment.
+\par
+Research has shown there to be deficits in both AD and MCI populations in this task \cite{Tales2005}, and those with MCI exhibited less severe deficits compabut that these deficits were not as apparent in MCI populations, there remained a significant enough difference that could be used to differentiate healthy controls from those with MCI.
+In a comparison of visual search performance between those with aMCI and healthy controls, Tales noted significantly poorer performance in the aMCI group.
+However, she also noted a good deal of heterogeneity in the aMCI group which illustrates that whilst the aMCI group have essentially the same condition, presentations within this population can differ markedly.
+The results from this study also illustrates the use of a non-memory task as a means of diagnosing dementia as patients who went on to be diagnosed with dementia whilst this study was in progress had significantly poorer visual search performance.
+\subsection{Free Cued Selective Reminding Test}
+The Free Cued Selective Reminding Test (FCSRT) was borne out of the premise that by controlling the conditions of learning, a measure of memory is possible that is not confounded by normal age related changes in cognition.
+Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cognitive decline due to a cognitive impairment.
+The FCSRT starts with a learning phase in which participants are asked to look at a card containing four pictures (e.g., grapes) for an item that belongs to a named category(e.g., fruit).
+After these four items are identified, immediate cued recall of these four items is tested and this is repeated for a total of 16 items.
+Following this learning phase is the recall phase in which participants are asked to recall all the items identified without cues and any items which are not recalled are then cued.
+There are three outcome measures for this test, free recall (the number of items recalled without cues), total recall (the number of items recalled with or without cues) and cue efficiency which is identified as follows \cite{Grober2010}.
+\begin{equation} \label{x1}
+CueEfficiency = (totalRecall-freeRecall)/48-freeRecall, range 0.0-1.0)
 \end{equation}
+Grober found that patients with impaired free recall with four times more likely to develop dementia than those with intact free recall \cite{Grober2010}.
+This test has also been shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the researchers to used the test to categorise prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT score \cite{Sarazin2007}.
+\subsection{Conclusions}
+I have looked at a number of different cognitive tests and a battery of tests that aim to have high diagnostic and prognostic capabilities in the MCI population.
+However, particularly with the RBANS, there lacks sufficient sensitivity in differentiating those with MCI from healthy elderly individuals such that this tool could be used with a level of confidence in a clinical setting.
+In regard to the tests, a common theme is a lack of studies and therefore evidence into the utility of these tests with an MCI population.
+However if, as researcher, we aim to investigate this population further then a benchmark battery of tests which is sensitive enough in both diagnostic and/or prognostic utility should be a goal.
+\par 
+Another criticism of the current literature is the lack of consistency with regard to the experimental groups.
+For example, some studies focus on differentiating between an MCI group, a AD group and a healthy controls group whereas other studies may further subdivide the MCI category according to a number of factors such as Amnesic or Non-Amnesic MCI, Mixed MCI (sometimes called Multi-domain MCI in the literature).
+Given the inconsistency in defining the experimental groups the confusing and often conflicting results that researchers produce is to be expected.
+Future research should look at the standardization of the operational definition of cognitive impairment in MCI may result in more consistent predictions of progression to AD.
+\par 
+Finally, it can be said that in testing for MCI and comparing these results with a cohort of those with a similar diagnosis has limited utility.
+This is because there is huge variability in the presentation of MCI within participants that even if controlled for with a matched pairs design for age, gender and education will produce inconsistent results.
+There is an argument therefore for the use of longitudinal studies, where the comparisons on the performance of these tasks are with the participants themselves.
+There are not many studies which use a longitudinal approach, although those that do show promising results.
 
-\subsection{Subsection Heading Here}
-Write your subsection text here.
+\section{Types of Language Assessment}
+One of the key debates when looking at how to analyse language is the type of task provided to elicit language production in participants.
+In the literature researchers have primarily focused on Picture Description tasks but have also suggested other ways in which we might collect data.
+\par
+\subsection{Picture Description Tasks}
+One of the most commonly used tasks to measure language is the Picture Description task.
+An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Borod1980}.
+The Cookie Theft picture (pictured below) depicts a scene of a home typical of the period of time when it was created and would generally not require participants to use any complicated vocabulary to describe.
+In this task participants are asked to describe the picture presented to them in as much detail as possible.
+This task was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}
+\begin{figure}[H]
+\centering
+\includegraphics[width=240px, height=150px]{images/BCTPicture.png}
+\caption{Cookie Theft Picture - From Kaplan and Goodglass (1983)}
+\end{figure}
+\par
+The picture description task does a fine job of eliciting descriptive language however because of the specific content the language produce could be considered quite limited.
+There is some disagreement as to the benefits of this using this methodology.
+This task is reported as being useful to lexico-semantic disorders \cite{Boschi2017, Sajjadi2012} as the language being generated is primarly nouns and deixis (words to identify items and words to put those items into context).
+However, Ash \cite{Ash2012}felt that there was no difference in using this task vs Story Narration (described below).
+In explaining the differences, it is worth noting that these researchers were using differing variables and this could explain their different perspectives.
+\par
+\subsection{Narrative description task}
+The story narration task is designed to study a participant's ability to describe and elaborate on a story which is depicted using a series of pictures.
+The stories depicted are usually based on children's books or famous stories with the Cinderella being the one most typically used \cite{Fraser2014}.
+This task requires ordering the story in a structured and coherent framework.
+It also requires comprehension and understanding of the stories characters and the events depicted, as well as an awareness of a character's actions, motivations and internal reactions to given events.
+This task is particularly useful as the procedure reduces the demands on memory, due to the participant being able to access the picture book during the description and is therefore able rule out memory as a confounding variable for any results observed.
+As noted above, Ash \cite{Ash2012} felt that this task was interchangable with the Picture Description task.
+However, other research felt that this was a studier test of lexical and semantic abilities as well as syntactic complexity because this task requires interpretation and elaboration in additional to a simple description \cite{DeLira2011}.
+\par
+Given the relative strengths of the Narrative description task vs Picture description task, there are few pieces of research that have used Machine Learning to analyse features from Narrative picture tasks \cite{Fraser2014}.
+This could be due to the availability of data and the absence of any meaningful sets of transcripts of participants performing this task.
+However, this could be an interesting direction to take research in the future to see if features generated from this task could be used to predict MCI or AD.
+\par
+\subsection{Interviews}
+Interviews can also be used to elicit language in a more natural way by asking questions to guide a conversation between speakers.
+There are three types of interviews: unstructured, structured and semi-structured.
+Structured interviews tend to produce very limited speech and therefore has never been used in this area \cite{Boschi2017}.
+Unstructured interviews are open ended and generally do not conform to any particular pattern.
+They use generic themes such as family or hobbies to guide the conversation.
+Whilst this is the most ecologically valid form of conversation and therefore language generation, it's unstructured nature means that the protocol cannot be consistent and therefore reproduced.
+Semi-structured interviews are therefore preferred over other forms of interview as a middle ground.
+The semi structured nature of these interviews means that there is some replicability but does not constrain the participant in answering questions.
+\par
+The analysis of interviews can be difficult to analyse as both the content can vary even between participants, although it can be argued that content should not affect the type of language being generated unless it is narrow topic or the participant is constrained in how they answer a given question.
+It is also difficult to measure as there are no pre-defined task goals in comparison to the other two methods.
+Nevertheless, this is the most naturalistic setting for looking at language production and can be used to look at the syntactic and semantic parts of language generation \cite{Sajjadi2012}.
+There have been some attempts to use interviews to assess language production in AD with promising results \cite{Asgari2017, Guinn2015}.
+\par
+\subsection{Conclusions}
+One can view the different types of tasks above as a continuum where picture tasks represent a much more controllable task with a lot of supporting research but which generates a much more constrained set of language that is atypical of normal speech in terms of the cognitive functions used.
 
-\section{Conclusion}
-Write your conclusion here.
+\section{How do we analyse language, issues and debates}
+There are a wide variety of approaches that we can take in the analysis of language and there is no real consensus on the best approach to this particular problem.
+This section looks at the different ways in which researchers have looked at the problem as well as discussing some of the areas of contention.
+\subsection{Single Word Language tasks vs Connected Language tasks}
+Part of the reason we need to pay attention to how we ask participants to generate data is understanding how we wish to analyse the data afterwards.
+As discussed above, the different methodologies to collect data generate different types of language.
+There are two main approaches which we have looked at to analyse language, using frequencies of words and combinations of words and measures of syntax and semantics.
+There are other less common methods of analysing language but these are beyond the scope of this review.
+\par
+Single word tasks such as the Boston Naming Test and other such standardized language tests generally target a participants word production where this is defined as the ability to form and express words in accordance with certain criteria (see above for a full description of verbal fluency as task).
+There are a number of benefits to using single word tasks.
+From a research methodology perspective, using a standardized test allows researchers to target a very specific process in language generation and isolate factors that impact performance in language well.
+However, this approach does not take into account other cognitive processes that are used in lengthier speech tasks.
+More specifically, single word language tasks do not look at the interaction between language, executive functions and reasoning abilities.
+They also do not require the logical and efficient organisation of ideas.
+\par
+Connected language tasks, such as the picture description task and interviews described above are much more reflective of the processes involved in natural language generation.
+They are able to be relatively constrained in the language available to be used, for example in the picture description task, or they can be unconstrained in the form of interviews.
+But regardless of how they are framed, they are potentially much more useful in terms looking at all the processes involved in language generation with the drawback that you will not be able to isolate specific parts of the language generation process.
+A review by Boschi et al concludes that analysis of connected speech is potentially useful in guiding clinicians to identify language disorders \cite{Boschi2017} and also highlights the role of NLP and Machine learning in assisting in this endeavour.
+\par
+\subsection{Semantics vs Pragmatics}
+When navigating the English language it is necessary to distinguish between what a sentence says in both semantic and pragmatic terms.
+Semantic meaning refers to the meaning of the words in a sentence local only to the given sentence.
+Another way to put this is, semantics considers the meaning of words without taking into account the context in which these words are spoken.
+Pragmatic meaning refers looks at the same sentence in terms of words and grammar but takes into account the situation or context in which these words are spoken.
+Emery in her literature review looks at all levels of language tasks except for pragmatics, however this is an area that should not be ignored.
+Whilst the study of language from a pragmatic perspective is much more complex, it is perhaps one of the most vital areas to study because of the number of different cognitive processes involved.
+\par
+\subsection{Semantic Content}
+Another approach to linguistic analysis in this field is the idea of measuring semantic content and complexity.
+According to Emery (2000) \cite{Emery2000} in which she states that Semantic and Syntactic skills deteriorate first in people with MCI and AD.
+If this is true, then psychological measures of semantic and syntactic skills should be able to pick up signs of deterioration and act as markers for possible MCI and AD.
+An example of a semantic complexity measure is the concept of idea density.
+Formally, idea density is defined as the average number of propositions per sentence \cite{Keenan1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}.
+An example of semantic content measures is Type to Token Ratio (described below) which is used to measure the lexical diversity of a given piece of text and/or utterance.
+This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2013}.
+\par
+\subsubsection{Type token ratio(TTR)}
+Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
+\begin{equation} \label{x2}
+TTR = numberOfUniqueWords / totalNumberOfWords.
+\end{equation}
+\subsection{Thematic and Content elements in relation to the Picture description task}
+A number of studies looked at the accuracy of picture descriptions using counts of thematic and content elements within a picture.
+Whilst called diffferent names such as 'pictorial themes', 'relevant observations and 'semantic units', they all represented the same idea.
+The only difference between the studies was the number of thematic elements that 'scored' correctly.
+Nicholas et al identified eight thematic elements of the Cookie Theft picture and used the number of elements as an outcome measure in different groups.
+He found that patients with AD expressed significantly fewer content elements than controls.
+\par
+Hier, Hagenlocker and Shindler assessed content using a similar list of thematic elements \cite{Hier1985}.
+They divided their participants into early-stage and late-stage AD, as well as including a control group.
+The late-stage AD group produced significantly fewer relevant observations than the early stage group, and the AD group combined produced fewer relevant observations than controls.
+This study was replicated by Lukatela et al \cite{Lukatela1998}.
+\par
+Smith, Chenery and Murdoch applied Hier's methodology for constructing pictorial 'themes' with the Picnic Scene from the Western Aphasia Battery (WAB) with a control and patients with moderate to moderately severe AD \cite{Smith1989}.
+The authors found no difference in the number of semantic elements produced but did not that the group with moderate to moderately severe AD took more time and more syllables to communicate these elements.
+\par
+\begin{figure}[H]
+\centering
+\includegraphics[width=240px, height=150px]{images/picnic-scene.jpg}
+\caption{Picnic Scene taken from the Western Aphasia Battery (WAB).}
+\end{figure}
+Sajjadi et al examined 10 pictorial themes in picture description the Comprehensive Aphasia Test and found that the group with mild AD produced similar themes than controls \cite{Sajjadi2012}.
+Bschor et al.
+(2001) examined Cookie Theft picture descriptions at four stages of AD.
+They found that whilst each AD group differed significantly from the others and also from controls, the measures were not able to distinguish between MCI and normal controls \cite{Bschor2001}.
+\par
+Finally, a number of studies used composite measures which contained thematic elements and other unspecified information units resulting in a list of 23 possible information units of the Cookie Theft picture.
+The authors felt that this provided a wider, more liberal range of relevant content and thus subtler differences could be noted.
+Studies using these features found some differences between AD and controls, and some could differentiate between different stages of AD.
+\subsection{General Information Units or Content Information Units}
+Some studies used a more general concept of content, using terms such as "general information units" or "content information units" and this could be defined as "the smallest non redundant meaningful fact or inference,"  and was counted whether or not the information conveyed was specific to the context in which the conversation happened.
+Giles et al for example studied adults with minimal, mild or moderate AD vs controls and found that adults with AD produced fewer overall information units than controls \cite{Giles1996}.
+\subsection{Conciseness of information}1`
+Conciseness has been defined as the number of words a speaker uses to express ideas.
+The theory is that people with AD would need more words to convey ideas because of difficulties with word-finding and compensatory  behaviours such as circumlocutions and repetitions.
+Conciseness has previously been calculated by dividing the number of ideas expressed by the total number of words in a measure commonly referred to as idea density but also known as lexical index, information content and information unit conciseness index.
+Snowdon examined written discourse from the Nun study and found that low idea density in early life was associated with reduced cognitive performance in later life \cite{Snowdon1996}.
+Riley et al extended these findings by concluding that early-life idea density was associated with lower brain weight, higher degree of cerebral atrophy and increased neurofibrillary pathology in later life \cite{Riley2005}.
+\par
+Ahmed, de Jager et al examined idea density with patients who had confirmed AD post mortem \cite{Ahmed2013}.
+They found that those with AD produced fewer total semantic units than controls but there was no significant difference between the groups with regards to idea density.
+The study of "empty speech" by Nicholas et al examined conciseness and specifically looked at empty phrases (defined as common utterances which contribute no relevant information), deictic terms (e.g.
+"this", "that" without referents), indefinite terms (e.g.
+"thing" or "stuff"), pronouns without proper noun antecedants, and repetitions.
+In their study they found that AD patients produced more of these than did controls.
+\par
+\subsection{Efficiency}
+% CITE MURRAY!
+Efficiency is the rate at which meaningful information is conveyed over time,  and can be calculated by dividing the total number of information units by the duration in seconds of the speech sample.
+Smith et al, 1989 found that 18 adults AD produced fewer content units over time on average than controls, he attributed these differences to increased circumlocutions and repetitions in the AD group \cite{Smith1989}.
+Murray used an similar measure in which fillers, irrelevant words, revisions or false stars, vague or non-specific vocabulary and inaccurate output were group together as 'performance deviations" and were divided by the total number of minutes in the sample.
+This measure was lower for those with AD than those with depression, and also healthy controls.
+The authors suggested that discourse information measures may help disentangle the similarities in symptoms of early AD versus depression in older adults.
+Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in dialogue in which a speaker provides responses do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls.
+\subsection{Total number of words}
+Several studies report that adults with moderate AD produce fewer words than controls on picture description, however other studies found no differences in total words among groups of controls and patients with MCI or AD.
+Murray and Nicholas et al investigated normal controls, patients with AD and older adults with depression and found no group differences in total words \cite{Nicholas1985}.
+In contrast, Lira et al found that controls produced more total words than patients with AD but found no difference between mild and moderate groups \cite{Lira2014}.
+\par
+\section{Syntax and Morphology (Language Form)}
+Syntax can be defined as the rules that govern how words can be combined to form sentences, whilst Morphology is the system that governs the structure of words and the construction of word forms.
+Multiple studies of language decline in dementia included at least one measure of syntax or syntactic complexity \cite{Orimaye2017}.
+Common constructs included words per clause, grammatical form (measures of an appropriate use of syntactic conjections, tenses, conditionals, subordinate clauses and passive constructions), measures of phrase length and proportions of words in sentences.
+Some researchers have explored the use of formulaic language in those with dementia, the theory being that well practiced phrases are less effortful and therefore place low load on the cognitive abilities of those with AD.
+The general hypothesis motivating these studies is that either working memory limitations or semantic memory limitations in AD affect one's ability to use complex constructions.
+\par
+\subsubsection{N-grams}
+One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
+An n-gram is a contiguous sequence of n items from a given sample of text or speech.
+The items can be phonemes, syllables, letters, words or base pairs according to the application.
+For example, given the sequence of words "to be or not to be", this extract is said to contain six 1-gram sequences (to, be, or, not, to, be), five 2-gram sequences (to be, be or, or not, not to, to be), four 3-gram sequences(to be or, be or not, or not to, not to be) and so on.
+This is useful as, given a large portion of text or speech, we can predict the probability of a word being close by to a given word.
+A number of researchers have used n-grams as features.
+\par
+Asgari, Kaye and Dodge (2017) \cite{Asgari2017} used another form of word frequency measurement.
+Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words \cite{Pennebaker2015}.
+They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
+\par
+\subsection{Formulaic Language}
+Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
+They found that there were four factors which they identified as important in the classification of participants as either healthy or AD.
+These four factors were semantic impairment, acoustic abnormality, syntactic impairment and information impairment and were based on existing measures of semantic and syntactic complexity.
+Zimmerer (2016) \cite{Zimmerer2016} looked at whether language was more formulaic in those suffering from AD.
+He proposed that those who suffer from AD rely on formulatic sentences, for example 'Noun-Verb-Noun', and this is done to reduce language complexity.
+He noticed a significant difference in the use of formulaic sentences between AD and Healthy Controls.
+\par
+\section{Pragmatic Language}
+The pragmatic language domain refers to the social rules for language for the purposes of communication including, using language to achieve goals, using information from the context to achieve these goals and using the interaction between people to initiate, maintain and terminate conversations.
+\subsection{Coherence}
+Coherence, in lay terms, can be defined as the ability to maintain awareness of the topic at hand.
+It can be separated into local coherence which is related to themes of the immediately preceding utterance and global coherence which looks at how closely an utterance is related to the topic currently being discussed.
+Chapman et al used picture descriptions of Norman Rockwell prints within a frame analysis, with frames being defined as the context in which the picture is viewed \cite{Chapman1998}.
+The authors identified aspects of content, including whether the frames of interpretation being offered were typical, atypical, incorrect or had no frame.
+They also looked propositions supporting frames and propositions disrupting frames as measures of coherence.
+They examined these variables with early stage AD, old-elderly and normal controls.
+Healthy older adults and normal controls produced significantly more typical frames and more frame supporting information than the AD group.
+The authors attributed AD patients' difficulties to memory deficits, attentional deficits, visual perceptual problems, disruption of internalized frame representation, or failure to access frame knowledge.
+\par
+\subsection{Perseveration}
+Perseveration can be broadly defined as the repetition of a response regardless of the absence or cessation of stimulus which could have generated an appropriate response in the first place.
+A typical example could be the idea of conversation moving from introductions to a more general conversation, someone who has difficulties with perseveration will struggle with the move between social contexts and repeat language that would be more used in an introduction context.
+\par 
+One study examined verbal preservation in the description of Norman Rockwell prints \cite{Bayles198}.
+The presented participants with a number of similar pictures that had significantly different contexts and asked participants to describe these prints.
+They divided the total number of words within perseverations by total number of words in the speech sample.
+The authors also calculated rate of perseveration on two other language tasks: confrontation naming and generative naming.
+In all tasks, the AD group produced significantly more perseverations than controls but there were no significant differences between the two groups in the picture task in isolation.
+The authors felt that this was because picture description was an easier task, as it was a visual task in contrast to the other tasks which tapped into other cognitive processes.
+\par
+\subsection{Empty Speech}
+Verbal fluency is a term used in neuropsychological contexts generally referring to timed, word-generation tasks, while in speech-language pathology contexts, "fluency disorders" are defined as interruptions in the flow of speaking characterized by atypical rate, rhythm and repetitions in sounds, syllables, words and phrases.
+"Fluency", in the literature of discourse of adults with AD, typically refers to the smoothness or flow of spoken language.
+Abnormalities of fluency in this population are typically characterised by filled and unfilled pauses, word repetitions, circumlocutions, and revisions.
+\par
+The study of "empty speech" by Nicholas et al was one of the first to examine aspects of fluency in the connected speech of persons with AD \cite{Nicholas1985}.
+They found that adults with AD had significantly more repetitions than controls.
+Similarly, Bayles and Tomoeda found more aborted phrases, revisions and ideational repetitions in the AD group than in controls \cite{Bayles1983}.
+Several other studies support the idea that in AD population there are a greater number of repetitions and revisions than in healthy controls.
+However, there are some studies which contradict these findings\cite{Ahmed2013}.
+\subsection{Conclusions}
+This section has looked at a number of different attributes which researchers have cited as being important in the detection of MCI and AD.
+Whilst in some cases there is a broad consensus about a given feature in the vast majority of cases there is some uncertainty.
+Some of the difficulty lies in different approaches used particularly around the criteria for experimental groups which means that it is difficult to be able to compare studies, like for like, as the populations of the participants being studied vary in subtle or obvious ways.
+It is important for any studies moving forward to use comparable standards when it comes to their experimental groups, such that the picture can be made clearer.
+
+\section{State of literature into Machine Learning and Natural Language processing techniques}
+Diagnosing dementia through language analysis has a large background in terms of psychological research.
+Increasingly current research has called for the use of machine learning as a way of assisting in the process of diagnosis \cite{Orimaye2017, Boschi2017}.
+This section will look at what areas of natural language processing (NLP) and machine learning (ML) could potentially be used as tools to help in this domain as well as any research that has applied machine learning to this problem.
+\subsection{Natural Language Processing}
+Natural Language Processing as a topic can be defined as the intersection between Machine Learning and Linguistics and looks at a number of language tasks, particularly important in this domain is how to enable a computer to process and analyse large amounts of language data.
+Some other tasks include speech recognition, natural language understanding and natural language generation.
+Generally speaking, Natural Language Processing (NLP) systems mimic the semiotic perspective on language.
+That is to say that we can subdivide NLP processes into different components (see Figure 2.3).
+\begin{enumerate}
+	\item Morphological/Lexical: providing the basic language elements or vocabulary, such as words, their roots and inflections.
+	\item Syntactic: for grouping and sequencing elements within samples of language (usually sentences).
+	\item Semantic: for knowing the meaning of an utterance, usually defined in terms of the 'truth value' of the logical propositions that are thought to be expressed by sentences.
+	\item Pragmatic: for understanding the context and purpose of an utterance.
+\end{enumerate}
+\begin{figure}[H]
+\centering
+\includegraphics[width=240px, height=150px]{images/workflow.jpg}
+\caption{How Natural Language Processing tasks are subdivided.\label{white}}
+\end{figure}
+In order to facilitate these and other tasks, a number of frameworks have been developed that automate some of the more common processing tasks. A brief review of these frameworks follows. 
+\subsubsection{Natural Language Toolkit}
+The Natural Language Toolkit (NLTK) was originally designed as part of a computational linguistics course at the University of Pennsylvania but has evolved to become an open-source framework which includes methods and modules for a wide array of Natural Language tasks (see Table 2.2 for a full list) \cite{Bird2009}.
+\begin{table}
+	\begin{tabular}{ | p{4cm} | p{7cm} | p{1cm} |}
+		\hline
+		\textbf{Language processing task} & \textbf{Functionality} \\ \hline
+		Accessing corpora & standardized interfaces to corpora and lexicons \\ \hline
+		String processing & tokenizers, sentence tokenizers, stemmers \\ \hline
+		Collocation discovery & t-test, chi-squared, point-wise mutual information \\ \hline
+		Part-of-Speech tagging & n-gram, backoff, Brill, HMM, TnT \\ \hline
+		Machine Learning & decision tree, maximum entropy, naive Bayes, EM, k-means \\ \hline
+		Chunking & regular expression, n-gram, named-entity \\ \hline
+		Parsing & chart, feature-based, unification, probabilistic, dependency \\ \hline
+		Semantic interpretation & lambda calculus, first-order logic, model checking \\ \hline
+		Evaluation metrics & precision, recall, agreement coefficients \\ \hline
+		Probability and estimation & frequency distributions, smoothed probability distributions \\ \hline
+		Applications & Graphical concordance, parsers, WordNet browser, chatbots \\ \hline
+		Linguistic Framework & manipulate data in SIL Toolbox format \\ \hline
+	\end{tabular}
+		\caption{\label{tab:table-name}NLTK tasks and functionality}
+\end{table}
+The NLTK has been successfully used in both teaching and research purposes and as part of this provides access to large array of databases and corpora which have potential uses in common tasks.
+
+\subsubsection{CoreNLP}
+Stanford CoreNLP provides a set of human language technology tools.
+This comes as an installable package that requires JAVA and can be interacted with through the command line or by an accompanying API.
+Whilst NLTK is much more oriented to parsing text and generating features, CoreNLP does this and adds more functionality such as a Named Entity Recogniser and sentiment analysis functions.
+At it's centre is the Stanford parser, which parses text and can give the base forms of words, their parts of speech, whether they are names of companies, people, etc., normalize dates, times, and numeric quantities, mark up the structure of sentences in terms of phrases and syntactic dependencies, indicate which noun phrases refer to the same entities, indicate sentiment, extract particular or open-class relations between entity mentions, get the quotes people said, etc.
+\par
+\begin{figure}[H]
+\centering
+\includegraphics[width=350px, height=200px]{images/CoreNLP.png}
+\caption{Depiction of how Core NLP marks up a sentence}
+\end{figure}
+It's intended for CoreNLP to provide tools and frameworks that will enable higher level language analysis whilst remaining domain agnostic.
+
+\subsubsection{Linguistic Inquiry and Word Count}
+The Linguistic Inquiry and Word Count (LIWC) is a piece of software that counts words and assigns these counts to categories \cite{Pennebaker2015}.
+The rationale behind this is that the words we speak inform an observer of the state of mind of an individual.
+A typical example of this might be someone with depression who may typically be more inwardly focused and therefore talking in the first person is common\cite{Al-Mosaiwi2018}.
+The 2015 version of the LIWC has over 90 different categories \cite{Pennebaker2015} and some supporting research suggests that it is effective at identifying themes and associating these themes with mental health difficulties such as depression and anxiety \cite{Sonnenschein2018}.
+\par
+Whilst not necessarily a tool that can parse data and look at trends from a linguistic perspective, a framework like this could be used as another tool to explore themes within a persons speech from a more qualitative angle.
+To the authors knowledge, there is no research that looks at speech changes from this perspective.
+
+\subsection{Traditional methods of Machine Learning}
+In terms of Machine Learning research in this area, a number of researchers have used transcripts based on picture description tasks \cite{Zimmerer2016, Orimaye2017, Mueller2018a, Fraser2015} and have successfully extracted linguistic features that could differentiate between AD and controls.
+\par
+One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the most frequent class.
+They detail several lexical approaches to the problem of detecting and rating AD.
+The approaches they looked at relied primarily on character n-gram techniques but also explore correlation of usage of frequency of different types of speech.
+Their results act as a proof in concept of the utility of using a pure computational approach to the diagnosis of dementia using spontaneous speech.
+The were able to obtain 95\% accuracy dementia vs controls, 70\% accuracy classifying dementia into two categories and 50\% accuracy classifying dementia into four categories.
+They suggest further exploration of characteristics to classify \cite{Thomas2005}.
+\par 
+% Cite CRUM Population-based norms for the Mini-Mental Status Examination by age and educational level
+Bucks et al \cite{Bucks2000} took a more traditional approach using frequency rates per 100 words of various word types such as nouns, verbs, adjectives and pronouns and some measures of lexical richness to attempt to discriminate between healthy older participants and participants with probable AD.
+They found that AD patients had higher mean pronoun rate, adjective rate and verb rate but lower noun rate vs normal older controls \cite{Bucks2000}.
+It's important to note that the clinical group averaged a score on the MMSE of 15 (sd=6.8) which puts the clinical group in the 'moderately impaired range'.
+At this stage, the two groups should be reasonably easy to differentiate in terms of language and other cognitive abilities.
+\par 
+Fraser et al \cite{Fraser2015} built upon the work by Bucks by adding an acoustic analysis element to her analysis using the DementiaBank corpus, as well as significantly expanding the list of linguistic features they used.
+They used features generated from the Stanford parser (Now known as Core NLP) and computed the frequency of the occurrence of different parts of speech which they then normalised by the total number of words in each utterance.
+They also computed ratios between these parts of speech such as pronouns to nouns.
+They also included measures of syntactic complexity such as mean length of sentences and measures of lexical richness such as TTR.
+Their analysis showed was able to achieve 81\% accuracy in distinguishing individuals with AD from controls based on short samples of speech.
+They found that there are four factors which they felt were important in discriminating between these two groups: semantic impairment, acoustic abnormality, syntactic impairment and information impairment.
+\par 
+Zimmerer et al \cite{Zimmerer2016} took a slightly different approach in that they developed the Frequency in Language Analysis Tool (FLAT) which looked at the degree of formulaicity in an individual sample.
+The rationale behind this is that people with impaired cognitive function are able to depend on formulaic sentences for language generation to ease the burden of language generation on cognitive symptoms.
+They were able find that those with probable AD produced more formulaic language than controls.
+\par 
+Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls.
+Their main finding supported n-grams as the most significant predictor.
+One of the criticisms is the use of picture description tasks and n-grams.
+Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
+\par
+
+\subsection{The case for Deep Learning}
+One of the criticisms of traditional learning models is it's reliance on features that are generated for them.
+As we can see from the brief look at some of the research in this field, researchers have taken a number of different approaches with success.
+However, there does not seem to be a consensus.
+As machine learning in general moves away from traditional machine learning models to deep learning techniques, it poses the question can deep learning assist with this problem?
+\par 
+One of the main benefits of deep learning is that it does not rely on pre defined features being fed to the model.
+Instead, deep learning models take raw data in with some amount of preprocessing and generates it's own features.
+This move away from relying on features solves one of the difficulties we have with the current psychological literature, namely that there are some disagreement on the which
+\par 
+One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
+This is the case with all machine learning techniques, but more so with deep learning.
+There have been ways to 'create' more data such as data augmentation.
+
+\subsection{Conclusions}
+We can see that both Natural Language Processing and Machine Learning techniques have a lot to offer, Indeed there has been a lot of research which have used pre-existing datasets to explore this area with promising results.
+One of the difficulties with the current research is the approach of trying to discriminate between those AD and healthy controls.
+This is not necessarily a problem in the real world as it is trivially easy to do for trained clinicians.
+A more interesting, but potentially harder to problem to solve is to discriminate between those with MCI and healthy controls, and more importantly to track their decline over time.
+Further, to date the authors are unaware of any research in which these techniques are applied to newly created samples of language.
+
+\section{Discussion}
+Current state-of-the-art diagnostic measures of AD are invasive (CSF analysis), expensive (neuroimaging), and time-consuming (neuropsychological assessment).
+Furthermore, these measures are limited to speciality clinics and thus have limited accessibility as frontline screening and diagnostic tools for AD.
+More importantly, nonspecialists are often inaccurate at identifying early AD and MCI.
+Thus, there is an increasing need for additional noninvasive and/or cost-effective tools, allowing effective frontline identification of subjects in the preclinical or early clinical stages of AD who could be suitable for monitoring in speciality clinics and for early treatment.
+Implementation of effective screening instruments will allow diagnosis earlier in the course of dementia, even at the point when memory function is still essentially within the normal range.
+This strategy would enable an earlier, and potentially more effective, prevention and treatment of AD with a special focus to preserve cognitive functions.
+\par 
+However the literature has identified a number of challenges when approaching this problem.
+Firstly, the clinical features that combine to meet the diagnostic criteria for Dementia or it's variants are continuous in nature and heterogenous between patients and are also impacted by other variables.
+For example as shown above, cognitive performance is affected in part by a patients educational attainment and a patients ability to live independently is impacted by a their physical health as much as their cognitive health.
+The challenge therefore is to find features that are minimally impacted by other factors, or that can be controlled for by a experimental design such as a matched pairs design to control for educational attainment.
+\par 
+Another challenge lies with the recruitment of suitable individuals who may notice a decline in cognition to the point where we might classify them as having MCI, but these individuals deduce that there is little to no value to admitting there is a problem and seeking help whilst their symptoms are 'manageable'.
+\par
+Finally there is a large amount of variability in the presentations of those with MCI and early dementia, and this is compounded by an similar amount of variability in the criteria researchers have used for experimental groups and the approaches researchers have used to tackle this problem.
+This had led to a confused literature.
+Recently a call for research that has consistent inclusion / exclusion criteria has been made along with some proposed definitions of MCI and it's subgroups \cite{Petersen2014}.
+Researchers have identified the analysis of language impairment as an area of promise to explore in the diagnosis of MCI and early AD and recent developments in natural language processing and machine learning techniques have the potential to assist in this research.
+Indeed, given the increased burden on the diagnosis of MCI and AD on professionals there has been a call to use technology to potentially ease this burden \cite{Boschi2017}.
+A small but growing amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
+
+\subsection{Future Work}
+One of the areas for research to study is a careful examination of the features that are being used to measure language deterioration.
+For example, Zimmerer (2016) \cite{Zimmerer2016} describes connectivity in such as a way that correlates directly with what Mueller (2018) \cite{Mueller2018a} calls Fluency.
+Whilst these are very nuanced measures which differ slightly in the form they take, the sheer range of measures and features being produced make it difficult to organise and explore what is truly going on in those with MCI and early AD..
+Some work needs to be done in producing a consistent list of measures that are validated using existing datasets and can be used for future research moving forward.
+\newline
+\par
+Teng et al suggests that work should focus on the MCI population and concentrate on developing a consensus neuropsychological battery that could yield predictable rates of progression to AD \cite{Teng2009}.
+This, in conjunction with the development of a model of language, sensitive enough to detect subtle deterioration in language use to act as an additional cognitive marker to aid diagnosis could potentially move some way to providing this.
+\par 
+Future research should also be directed towards developing non-intrusive ways of detecting subtle changes in natural language such that any perceived deterioration that could indicate the presence of MCI or AD could be flagged up early.
+Machine learning approaches seem to be the most logical approach for achieving this aim as language could be collected in non-intrusive ways and passed to a machine learning algorithm for preliminary classification.
+Despite the excellent quality of datasets, for example the DementiaBank dataset, being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which some researchers could claim are limited in scope.
+Finally, the recent resurgence in the use of neural networks and deep learning could provide the answer to the confused literature in terms of features.
+A key benefit of deep learning is it's ability to automate the process of feature engineering.
+So there is an opportunity to explore the use of deep learning, to not only develop new features but also validate existing features independently.
+\par
+This area of research is extremely promising in its early results and the impact of successful research would be life changing for both individuals and the health of the worlds aging population in general.

### 2019-07-08 Source/Thesis/Chapter5.tex
reorganised files
organised and updated

--- 2019-06-20 Source/Thesis/Chapter2.tex
@@
-\section{Introduction}
-Here is the text of your introduction.
+\section{Background}
+There has been a significant research in the area of language deterioration as a means of detecting Alzheimer's Disease.
+This usually takes the form analysis of speech recorded as part of a cognitive assessment such as the Picture Description Task \cite{Orimaye2014,Fraser2015}.
+Given that language samples are relatively easy to collect, research has moved towards analysis of spontaneous speech.
+An good example of this type of research is the study conducted by Berisha and Liss which looked at the differences in language use between two US presidents, Ronald Reagan (who would go on to receive a diagnosis of Dementia) and George H.W.
+Bush who acted as a matched control based on Age \cite{Berisha2015}.
+They found several differences in language use which they felt acted as indicators of Reagan's difficulties with language due to dementia.
+These significant differences were in the number of unique words used per speech, the use of non-specific nouns and fillers and low-imageability verbs \cite{Berisha2015}.
+\par   
+This study replicates work done by Berisha and Liss and extends this by adding Donald Trump as an alternative, more appropriate comparison to Ronald Reagan as he is much closer in age than George H.W.
+Bush.
+This experiment will look at the features originally identified by Berisha and Liss, as well as any others that have potential as discussed in the literature review above.
+\section{Methods}
+I took 46 transcripts of Ronald Reagan’s (RR) press conferences from 1981 to 1988 and compared them with 134 press conferences by George H.
+W.
+Bush (GHWB) and 29 press conferences conducted by Donald J.
+Trump (DJT).
+I analyzed transcripts for lexical features shown to change longitudinally with dementia  (for a comprehensive review of these, see the literature review above).
+For this collection of documents, I generated a number of features which looked at a number of different aspects of each document.
+These features encompassed, word level, sentence level and document level features and included a number of features contained in the study by Berisha and Liss with the aim of replicating and extending on their findings.
+These findings were.
+number of unique words, non-specific nouns and fillers and low imageability (LI) verbs.
+Imageability is characterized, according to Berisha and Liss, as the ease with which a term gives rise to a sensory .mental image.
+I compared the trends described in the transcripts of RR and GHWB, but also included DJT.
+Berisha and Liss originally made the comparison as it GHWB (GHWB - age at the start of presidency - 64 years, 222 days) was the closest match to RR in terms of age (RR  - age at the start of presidency, 69 yeaars and 349 days).
+However, with the inauguration of Trump, he now is the closest comparible president in terms of age (DJT - age at the start of presidency - 70 years, 220 days).
+It would be interesting to look at a comparison of RR and DJT to see whether the comparisons made by Berisha and Liss hold true with this more appropriate match (in terms of age).
+DJT as with GHWB has no known diagnosis of AD.
+I used the press conference transcripts in the American Presidency Project (APP) archive as a data source for this project.
+The APP is a comprehensive and organized searchable database of presidential documents, including transcripts of speeches, transcripts of news conferences, and other public documents.
+\subsection{Pre-processing}
+To generate the files necessary for analysis, I downloaded each transcript and performed the following changes.
+I omitted the prepared statement by the president and any speech by other individuals.
+I started each transcript at the beginning of the first answer to a question.
+I filtered any annotations that were added to the transcript, including any references or clarifications, and any laughter.
+It's worth noting that there appears to be a difference in how 'hesitations' were marked down between each president, for RR hesitations were marked by a single hyphen whereas for GHWB hesitations are marked by a double hyphen.
+In order to maintain consistency when parsing through the documents, I have changed both types of hesitation to be marked by a single hyphen.
+I also omitted one word sentences as this data would, from a theoretical perspective, not be relevant for language analysis.
+I did not control for the length of the document, but generated features which would normalise by the length of the document.
+I therefore was able to include all press conferences by both Ronald Reagan and George H.W.
+Bush where there was a question and answer session conducted at least in part by the sitting president (2 press conferences of GHWB were ommited due to a lack of a question and answer session).
+\subsection{Feature Selection}
+We calculated the following features for each transcript in turn using the NLTK (see section 2 for a description) \cite{Bird2009} and Python. 
+\par 
+\subsubsection{Measures of lexical variation}
+We constructed two features of lexical variation.
+Firstly we looked at the number of unique words.
+To do this we were able to split each transcript into individual words and changed them to lowercase using NLTK and were then count the number of unique words that appeared in each transcript.
+We also used the TTR formula (see section 2 for a description) for a feature that measures lexical diversity independent of sample size \cite{Richards1987}.
+\subsubsection{Fillers, Non-Specific Nouns and LI Verbs}
+For these features, we counted the number of occurrences for different categories (see table for list of categories tracked and the words counted).
+The features were used by Berisha and Liss in their research \cite{Berisha2015} and were taken from work done by Bird et al \cite{Bird2000}.
 
-\begin{equation}
-    \label{simple_equation}
-    \alpha = \sqrt{ \beta }
-\end{equation}
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{6cm} |}
+		\hline
+		Category & Words \\ \hline
+		Fillers & \textit{"well", "so", "basically", "actually", "literally", "um", "ah"} \\ \hline
+		Non Specific Nouns & \textit{"something", "anything", "thing", "everything"} \\ \hline
+		LI Verbs & \textit{"be", "come", "do", "get", "give", go", "know", "look", "make", "see", "tell", "think", "want"} \\ \hline
 
-\subsection{Subsection Heading Here}
-Write your subsection text here.
+	\end{tabular}
+	\caption{\label{tab:table-name}Categories and Words Counted}
+	\end{center} 
+\end{table}
 
-\section{Conclusion}
-Write your conclusion here.
+\subsubsection{Usage of parts of speech}
+This section involves using a Part of Speech tagger (PoS) which analyses a sentence and assigns a 'tag' to each word based on the function the word has in a sentence.
+At a basic level this can be divided into the eight defined parts of speech: 'nouns', 'pronouns', verbs', 'adjectives', 'adverbs', 'conjunctions', 'prepositions' and 'interjections' but can be further subcategorised.
+We used the PoS tagger built into NLTK to tag each transcript in turn and used these the counts from each of these eight categories in our analysis.
+In addition to frequency counts we also normalised these features by dividing the frequency count by the number of words in the document to take into account transcript length.
+
+\section{Results}
+One of the most important thing to note is the wide variety of samples between the three presidents and also the varying timescales.
+RR participated in 46 press conferences over eight years (an average of 5.75 a year) which is the fewest number of press conferences given by an American president during their term of office.
+GHWB participated in 136 press conferences over four years (an average of  34 a year) and DJT participated in 29 press conferences to date (an average of 19.3 per year).
+Equally, there are variances in the average number of words.
+RR produced an average of 3424 words per conference compared to 2608 by GHWB (unpaired t = 4.434, p\textless 0.001) and DJT at 1849 words (unpaired t = 6.524, p\textless 0.001).
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR & GHWB & DJT \\ \hline
+		Total Words & 3423.91 (416.42) & 2607.72 (1210.38) & 1848.65 (1549.38) \\ \hline
+		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
+		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
+		LI Verbs & 124.22 (17.89)& 103.45 (51.87) & 84.48 (75.78) \\ \hline
+	\end{tabular}
+	\caption{\label{tab:table-name}Means and Standard Deviations of important features}
+	\end{center} 
+\end{table}
+
+In terms of unique words, we found that RR used significantly more unique words, non-specific nouns and low imageability verbs than GHWB and DJT (see Table 3.3).
+Some of these differences are due to the length of the sample, particularly in the case of DJT where his average sample is almost half the sample of RR.
+It could also be said that this could be down to differences in linguistic abilities or speaking style \cite{Berisha2015, Le2011}.
+However, we can certainly see that as controls GWHB and DJT are comparative in relation to non-specific nouns and LI verbs.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR v GHWB & RR v DJT & GHWB v DJT \\ \hline
+		Total Words & \textbf{4.434***} & \textbf{6.524***} & \textbf{2.899**} \\ \hline
+		Unique Words & \textbf{6.832***} & \textbf{11.403***} & \textbf{4.137***} \\ \hline
+		Non Specific Nouns & \textbf{7.877***} & \textbf{3.426**} & -0.574 \\ \hline
+		LI Verbs & \textbf{2.656**} & \textbf{3.420***} & 1.628 \\ \hline
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+	\end{tabular}
+	\caption{\label{tab:table-name}RR T-tests vs GWB and DJT}
+	\end{center} 
+\end{table}
+
+We then looked at the data from a longitudinal perspective as we are interested seeing whether we can track various language variables and their progress over time.
+We ran a number of Pearsons correlations with transcript index number as a time reference and the dependant variables (Table 3.4).
+For our controls, we found them to be stable for the most part with the main highlights being a decrease in Adverb usage for DJT (R=-0.36, p=0.049) and a steady but not severe decline in a number of variables for GWHB, namely total word count, unique words, low imageability words and verb usage.
+\par 
+For RR, his decline is more marked and more widespread through his language use.
+We noticed an significant increase in adverb (R=0.41, p=0.004) and pronoun usage (R=0.65, p\textless0.001), as well as a slight usage increase in Non-specific nouns(R=0.30, p=0.03).
+There was a highly significant decrease in number of unique words (R=-0.56, p\textless0.001) and noun usage (R=-0.70, p\textless0.001).
+Also very significant decrease in adjective usage (R=-0.40, p=0.005) and a significant decrease in total word count (R=-0.31, p=0.03).
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | l | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR & GHWB & DJT \\ \hline
+		Word Count & \textbf{-0.31*} & \textbf{-0.21*} & 0.08 \\ \hline 
+		Unique Words & \textbf{-0.56***} & \textbf{-0.25**} & 0.16 \\ \hline
+		Non Specific Nouns & \textbf{0.30*} & -0.08 & -0.03 \\ \hline
+		LI Verbs & -0.19 & \textbf{-0.20**} & 0.02 \\ \hline
+		Nouns Normalised & \textbf{-0.70***} & -0.03 & 0.14 \\ \hline
+		Verbs Normalised & \textbf{0.36**} & \textbf{0.24***} & -0.03 \\ \hline
+		Adjectives Normalised & \textbf{-0.40**} & 0.08 & -0.34 \\ \hline
+		Adverbs Normalised & \textbf{0.41***} & 0.02 & \textbf{-0.36*} \\ \hline
+		Pronouns Normalised & \textbf{0.65***} & 0.13 & 0.07 \\ \hline
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+	\end{tabular}
+	\caption{\label{tab:table-name}Pearson Correlations for Features}
+	\end{center} 
+\end{table}
+
+\section{Discussion}
+President Reagan received his diagnosis of AD in August 1994 but using transcripts of speeches he made in his two terms as President (January 1981 - January 1989) we have be able to identify certain changes in his use of language that we might ascribe to the onset of MCI and early AD.
+Despite differences in our methodology, our research supports the findings of Berisha and Liss in that we both find a significant decrease in unique words over time and an increase in non-specifc noun usage.
+Compared to our controls (GWHB andDJT), we find some slight trends with GWHB but no such trends with DJT in his speech albeit his samples of speech span a shorter amount of time.
+\begin{figure}[H]
+	\centering
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/RRUniqueWords.png}
+		\caption{Ronald Reagan - Unique Words over time}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/GWHBUniqueWords.png}
+		\caption{George H.W. Bush - Unique Words over time}
+	\end{minipage}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/RRNSNouns.png}
+		\caption{Ronald Reagan - Non-specifc Nouns over time}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/DJTNSNouns.png}
+		\caption{Donald J. Trump - Non-specific Nouns over time}
+	\end{minipage}
+\end{figure}
+
+A criticism of Berisha and Liss's work is the problems they had with normalising the transcripts in terms of length.
+This was also a problem in the work of Garrard et al \cite{Garrard2005, Le2011}.
+Whilst it is important to control for outliers, there are other ways in which we can control for length of sample.
+\par 
+Interestingly, when we normalised the various types of words used by the presidents we found some interesting patterns that further differentiated RR from the controls.
+Whilst Non-specific nouns increased over time, we found that noun usage in general significantly decreased and pronouns increased similarly significantly.
+The increase in pronoun for those with early AD has been identified in literature, although there are only a few studies that explore this \cite{Almor1999, Wendelstein2015}.
+Wendlestein et al propose that the increased used of pronouns is an expression of an impaired ability to adapt language to the listener's needs \cite{Wendelstein2015}.
+Almor et al attributed this reliance on pronouns due to a impaired working memory \cite{Almor1999}.
+\par 
+The decrease in overall noun usage has also been identified as a feature.
+Jarrold et al found that AD patients would use more pronouns, verbs and fewer nouns than controls \cite{Jarrold2014}.
+Wendlestein in their investigations into noun usage found that decreased later on in AD progression and was unaffected in the pre-clinical stages of AD \cite{Wendelstein2014}.
+Our results are supported by existing literature and this potentially means that language analysis in the way we have structured it may have diagnostic or prognostic properties.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=240px, height=150px]{images/RRNounsNormalised.png}
+	\caption{Ronald Reagan - Nouns Normalised over time}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=240px, height=150px]{images/RRPronouns.png}
+	\caption{Ronald Reagan - Pronouns Normalised over time}
+\end{figure}
+
+There are limitations of this research.
+Whilst in terms of age, DJT is certainly more suitable as a control to match with RR, in some ways they held very different styles of press conferences in that RR preferred to do solo press conferences and DJT has shown a preference for doing joint press conferences which have an impact on the amount of language produced.
+This artifact of the data is in itself notable as it illustrates the problems we may have with smaller amounts of speech.
+Also, the problem of finding an appropriate control is a common one in this domain.
+Given that those with MCI and early dementia have such variable presentations, it might prove of limited value in matched pairs design.
+\par 
+With further work, it is not feasible to the vast array of samples over a timeframe, as we have had with the president corpus and so it would be worth exploring how the quality of these predictions might lessen when faced with considerably fewer samples and over a smaller time period.
+It would also be worth extending this research further to encompass more of the linguistic features Fraser used in her work \cite{Fraser2015} to see if there are any further insights to be gained.
+In addition, this replication and extension has demonstrated the potential utility of using longitudinal data as a means of comparing language use of a person at two or more time periods and using this information as a diagnostic aid for MCI and therefore more work would be helpful from a longitudinal perspective to see if this approach may be valid in moving towards a solution for this particular problem.
+\par 
+The results of this work show that we can track a person's use of language through time in a number of ways, and it is possible for an individual to be his or her own control.
+This is important as it means the heterogenous nature of the MCI population does not impact results as much as if we were comparing those with MCI to controls.
+Equally, it would be helpful to have controls to ascertain what would be usual to expect in the decline of language in a healthy older adult.

### 2019-07-22 Source/Thesis/Chapter1.tex
updated Chapter 1

@@
 All the drugs that are on the market are designed to manage symptoms.
-Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{AlzheimersSociety2014}.
-Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place.
+Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{Cummings2014}.
+Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place \cite{Cummings2014}.
 It is therefore important to focus AD at it's earliest stages which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.
 \par
-Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent.
-Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time and alter cognitive function until a threshold is reached and clinical symptoms become more obvious \cite{Nestor2006}.
-The case of Iris Murdoch, who had a confirmed diagnosis of dementia, illustrates this theory well.
-Le et al \cite{Le2011} found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel when, it is theorized, the symptoms of AD manifested themselves more significantly.
-If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
-One of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of dementia.
+One of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of cognitive impairment or dementia.
 Albert and his team have worked to define clinical criteria which professionals can use to diagnose MCI due to AD and differentiate this from age-associated memory impairment and age-associated cognitive decline.
 They note that the diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}.
+The criteria for MCI is detailed below \cite{Albert2011}.
 \begin{enumerate}
-	\item Concern regarding a change in cognition - A person or an informant should express concern that there is a change in cognitive ability in comparison to previous level of performance.
+	\item Concern regarding a change in cognition - A person or an informant should express concern that there is a change in cognitive ability in comparison to a previous level of performance.
 	\item Impairment in one or more cognitive domains - There should be evidence of lower performance in one or more cognitive domains beyond what would be expected of a person given their age and education.
-	\item Preservation of independence in functional - Whilst persons with MCI are expected to be able to maintain independence, it is common to experience mild problems in complex functional tasks which they may have been able to perform previously.
+	\item Preservation of independence in functional abilities - Whilst persons with MCI are expected to be able to maintain independence, it is common to experience mild problems in complex functional tasks which they may have been able to perform previously.
 This might mean that they take more time or be less efficient at completing these tasks, or it may be that may make more mistakes.
@@
 \par
-Many researchers have studied the early detection of MCI.
-These studies usually follow two main approaches.
-The analysis of biomarkers and the examination of patients who have demonstrated decreasing cognitive abilities.
+% Do more on the state of the art in terms of the analysis of biomarkers for MCI.
+Studies that explore ways in which we can diagnose MCI / AD generally follow one of two main approaches, the analysis of biomarkers such as concentrations of amyloid-β 1-42 (Aβ42) in cerebrospinal fluid (CSF) and analysis of cognitive abilities.
 The first approach yields reliable results in the detection of AD in its moderate and advanced states but does not perform well during the early stages of the disease.
-The second approach has gained more attention in recent years, due to the fact that in clinical practice it has shown promise in the early detection of AD.
+The second approach, that of analysing the cognitive abilities of patients in memory clinics, has gained more attention in recent years due to the fact that in clinical practice it has shown promise in the early detection of AD.
 In addition, the analysis of the decline of cognitive abilities is comparatively inexpensive and less invasive that the first approach which commonly requires the collection of a sample of cerebro-spinal fluid which is painful for the patient involved.
+This has a number of benefits for countries with less developed healthcare systems, or where the burden of healthcare is more extreme.
 \par
-One of the most common ways in which clinicians traditionally make an early diagnosis of dementia is through the use of the Mini Mental State Examination (MMSE) \cite{Folstein1975}.
+One of the most common ways in which clinicians traditionally use the analysis of cognitive abilities to make an early diagnosis of dementia is through the use of the Mini Mental State Examination (MMSE) \cite{Folstein1975}.
 The MMSE is a brief questionnaire consisting of eleven questions which tests cognitive aspects of mental function and requires only 5-10 minutes to administer \cite{Folstein1975}.
@@
 These tests have the benefits of being much more accurate at diagnosing cognitive impairment and discriminating between dementia and other types of cognitive impairment at the cost time and training of psychological professionals such as clinical psychologists in administering these tests.
-However, the utility of diagnosing dementia at the point where clinical intervention is warranted is limited because at this stage both psychological and pharmacological interventions have been shown to not be effective \cite{Prince2015}.
+However, the utility of diagnosing dementia at the point where clinical intervention is warranted is limited because at this stage both psychological and pharmacological interventions have been shown to not be effective \cite{Prince2015, Cummings2014}.
 In order to further our understanding of the progression of dementia it is important to detect the signs of dementia before they are clinically apparent.
+\par 
+Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent.
+Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time and alter cognitive function until a threshold is reached and clinical symptoms become more obvious \cite{Nestor2006}.
+The case of Iris Murdoch, who had a confirmed diagnosis of dementia, illustrates this theory well.
+Le et al \cite{Le2011} found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel when, it is theorized, the symptoms of AD manifested themselves more significantly.
+If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
 \par
@@
 \par
-According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with mild dementia suffer from noticeable word finding difficulty.
-They may substitute general terms for more specific terms and may avoid the use of specific names of acquaintances.
+McKhann and his team were tasked with developing diagnostic guidelines for dementia in such as way that the criteria were flexible enough to be used by general healthcare providers without access to specialist medical equipment.the DSM 5 \cite{McKhann2011}.
+They state those suffering with mild dementia generally encounter impaired language functions (speaking, reading, writing)––symptoms include: difficulty thinking of common words while speaking, hesitations ;speech, spelling, and writing errors.
+More specifically, they may substitute general terms for more specific terms and may avoid the use of specific names of acquaintances.
 There may be grammatical errors involving subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc.
@@
 The field of machine learning and natural language processing has been suggested as a way to improve the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline \cite{Boschi2017}.
-\par 
-In 2009, the UK's Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
-A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
-As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages \cite{Prince2015} and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
-At the present time diagnostic assessment is typically conducted at memory clinics by trained clinicians \cite{Boschi2017} and whilst it is not possible to provide a definitive diagnosis until after death.
-
 I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
@@
 \par
+The overall objective of the research described in this thesis is to investigate the deterioration of language in people diagnosed with MCI and the early stages of AD.
+In order to do this I aim to answer the following research questions
+\begin{enumerate}
+	\item What research has been conducted in this field so far?
+	\item How can we use technology to best enable the collection of language samples?
+	\item Can we develop a complete data processing pipeline that processes data efficiently
+	\item What can machine learning tell us about how language affects us?
+\end{enumerate} 
+\par 
+This thesis' original contributions are an increased understanding of how language deteriorates over time in those with MCI and early AD, the exploration via a delphi method of ideas for the collection of language features via smartphone or other internet of things devices and the development and execution of a pilot study to test the feasibility of collecting high quality language samples via in-home technology.
+I also contribute to the research into which language features are important in the diagnosis of MCI and early AD over time.
+Finally, I explore the potential of one class classification as a new approach to classifying those with MCI and Early AD vs controls.
 The potential impact of this research in this area is immense.
@@
 \par
-In chapter 2 Background and Related Work
+There is a lot research in this domain from a psychological perspective.
+In chapter 2, I look at the background and related work in the domain of language deterioration in those with MCI and AD from the psychological perspective.
+This includes looking at the current ways in which psychologists collect and analyse language samples and exploration of the features of language that psychologists feel are important in this area.
+In addition, I take a systematic look at the work of the experts in the fields of machine learning and natural language processing in the growing area of research for NLP and Machine Learning and it's application to this problem.
 \par 
-In chapter 3 Systematic Review of NLP and Machine Learning Research
+The start of my research aims to answer questions around how we can use technology to aid us in collecting language samples that may facilitate the process of diagnosis.
+Currently, language samples are collected in memory clinics and physician's offices and is quite a time consuming process.
+More recently, work has been done to automate this process via a web based app called Talk2Me \cite{Komeili2019}, which seeks to use different psychological tests to assess cognitive decline through analysis of language.
+There are some potential flaws in this method of data collection.
+In other areas, the use of smartphone technology has allowed the collection of language for the diagnosis of disorders such as Parkinson's disease and so it appears plausible to be able to use this technology and potentially other technologies such as Amazon Echo to facilitate the collection of language samples.
+The work described in chapter 3 therefore describes an exploration of experts opinions' on the following questions.
+\begin{enumerate}
+	\item How should we collect these language samples - Smartphones? IoT Devices?
+	\item What tasks should we use to elicit these language samples
+	\item At what frequency should we collect these language samples at
+\end{enumerate}
 \par 
-In chapter 4 Delphi Methodology and developing consensus on how best to collect language samples using technology
+An adapted Delphi survey methodology is used to develop consensus on the answers to these questions and the results of these are documented in this chapter.
+The resulting pilot study is described in Chapter 6.
 \par 
-In chapter 5 Development of a pipeline that processes language data accurately
+Once we have a way to collect language samples, the question to consider is how we may best analyse this data.
+There are a number of factors that one must consider when answering the question.
+We look at the effectiveness of automatic speech recognition at translating speech into language, as well as the minimum quality of data that needs to be passed to any pipepline we develop.
+The Talk2Me app collects and then generates approximately 2000 lexico-syntactic, acoustic and semantic features (\cite Kommeili2019} as part of their data collection framework.
+Chapter 4 discusses the challenges connected to constructing an effective data pipeline and contains two experiments.
 \par 
-In chapter 6 Analysis of DementiaBank dataset
+Whilst Chapter 4 is more theoretical in it's content, Chapter 5 documents the first experiments that use this pipeline on three existing datasets.
+These datasets are, the Presidents Press Conferences, The Novels of three authors and the DementiaBank dataset.
 \par
-In chapter 7 Pilot study of the methodology developed
+In chapter 6 Pilot study of the methodology developed
 \par 
-Finally, in chapter 8.
+Finally, in chapter 7.
 I discuss conduct a general discussion of the results of my research.

### 2019-07-22 Source/Thesis/Chapter4.tex
updated Chapter 1

\section{Background}
As Komeili in her paper points out, machine learning is a tool may allow earlier detection and management of change in language \cite{Komeili2019}. But as with all machine learning models, the benefits of accuracy only come when large amounts of data is used to train the model. With our target population, this problem is further exacerbated due to the relatively low incidence of people with MCI or early AD and the relatively high cost of collecting this data in this particular population. 
\par 
One potential way forward is to look not at the ill but at the healthy. This has inverts the problem of scarcity and cost of data collection, and so we now have an abundance of data that we can now use. This, in theory, would allow us to look specifically at any deviations from the healthy. This process is known as one-class classification.
\par

\section{One-class classification}
In traditional machine learning, there are numerous classification algorithms that are widely used. Their utility is classifying instances of data into one (Binary classification) or more (multiclass classification) categories. This ia a fine way to look at classification in the vast majority of cases, but some difficulties do arise under certain circumstances.

### 2019-07-22 Source/Thesis/Thesis.tex
organised and updated
updated Chapter 1

@@
-\documentclass[twopages, a4paper,11pt]{book}
+%%% PREAMBLE
+\documentclass[10pt, letterpaper, twoside, openany]{book}
+
+\usepackage{setspace}
+\doublespacing
+
+\usepackage[utf8]{inputenc}
+\usepackage{fdsymbol} 
+\usepackage{graphicx}
+\usepackage{float}
+
+\DeclareUnicodeCharacter{03B2}{\ensuremath{\beta}}
 
@@
 \section*{Abstract}
+My original contribution to knowledge is....
 \newpage
@@
 \section*{Acknowledgements}
+Please note: An editor has not been used in the construction of this thesis.
 \newpage
@@
 
-\chapter{Background and Related Work}
+\chapter{Background and Related Work, Systematic Review of NLP and Machine Learning Research}
 \include{Chapter2}
 
-\chapter{Systematic Review of NLP and Machine Learning Research}
+\chapter{Delphi Methodology and developing consensus on how best to collect language samples using technology}
 \include{Chapter3}
 
-\chapter{Delphi Methodology and developing consensus on how best to collect language samples using technology}
+\chapter{Development of a pipeline that processes language data accurately}
 \include{Chapter4}
 
-\chapter{Development of a pipeline that processes language data accurately}
+\chapter{Analysis of DementiaBank dataset}
 \include{Chapter5}
 
-\chapter{Analysis of DementiaBank dataset}
+\chapter{Pilot study of the methodology developed}
 \include{Chapter6}
 
-\chapter{Pilot study of the methodology developed}
+\chapter{General Discussion, Conclusions and Future Work}
 \include{Chapter7}
 
-\chapter{General Discussion}
-\include{Chapter8}
-
-\chapter{Conclusions}
-\include{Chapter9}
 \bibliographystyle{unsrt}

### 2019-08-03 Source/Paper 2 - Systematic Review/Systematic Review - Protocol.tex
updated Paper3\

@@
 \documentclass[12pt]{article}
+
+\title{Applying Natural Language Processing and Machine Learning techniques to aid in the diagnosis of Mild Cognitive Impairment and Early Dementia - A Systematic Review}
+
 \usepackage{graphicx}
 \usepackage{float}
+
 \begin{document}
 \tableofcontents
-% https://github.com/anadallora/SNAC/blob/master/protocol%20v4.pdf
-\section{Introduction}
+\section{Background}
 Dementia has been identified as one of the fastest growing difficulties facing the world.
@@
 \par
-In 2009, the Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
 A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
 As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages \cite{Prince2015} and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
-At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017}.
+At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017} but there is an ever increasing amount of research looking at how smartphones and internet of things (IoT) devices can be used to collect data in homes.
+
  I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
@@
 Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
+\par 
+The aim of this review is 
+\par 
+This will allow for
 
-The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to to time-consuming and expensive psychological tests.
-There is a lot of research into the analysis of language as a bio-marker for MCI and Early Dementia.
-Given that sampling a person's language is relatively effortless, my research looks at whether we can find bio-markers of MCI and Early Dementia in natural language.\newline
-Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
-I'm not going to control for memory problems as a potential confound, but does this weaken the research?
-How do I defend this?
-\newline
-\par 
-Introduction to the problem of dementia in the context of the wider world including quality of life and financial implications.
-Exploration of dementia as a syndrome rather than a disease, and a look at the different variants of dementia.
-A look at the rationale behind research into the early diagnosis of dementia as well as a brief look at what has been done in the area (wide context, so pharmacological and psychological).
-\newline
+\section{Research Protocol}
+The main objective of this systematic review is to explore how Natural Language Processing and Machine Learning techniques have been used to aid the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease.
 
-The question to be addressed in this systematic review is how has the field of machine learning and natural language processing addresses language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease.
-\section{Research Protocol}
 \subsection{Research Question}
-The main research question this SLR aims to address is: “How has the field of machine learning and natural language processing addresses language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease?”.
-This main question was decomposed further into five research questions:
+The main research question this SLR aims to address is: “How has the field of machine learning and natural language processing addressed language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease?”.
+This main question can be decomposed further into four research questions:
 \begin{enumerate}
@@
 	\item RQ3: What are the goals of the studies that employ NLP / ML techniques for prognosis of dementia?
-	\item RQ4: Do the studies focus on time as factor??
-	\item RQ5: Do the studies focus on individuals or populations?
+	\item RQ4: Do the studies focus on time as factor?
 \end{enumerate}
+
 \subsection{Search Process}
@@
 	Are the measures used in the study clearly defined? & Yes=1; Partly=0.5; No=0 \\ \hline
-	Are the measuses used in the study valid? & Yes=1; Partly=0.5; No=0 \\ \hline
+	Are the measures used in the study valid? & Yes=1; Partly=0.5; No=0 \\ \hline
 	Are the measures used in the study clearly described? & Yes=1; Partly=0.5; No=0 \\ \hline
@@
 
-\appendix
-\section{\\Appendix A - Change Log}
+\bibliographystyle{unsrt}
+\bibliography{main}
 
-\section{\\Appendix B - Searches}
-\begin{table}
-	\begin{tabular}{ p{12cm} }
-	\hline
-	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
-	Searched - 4th April 2019 - Generated 1257 Articles \\
-	\hline
-	\end{tabular}
-	\caption[Table caption text]{Search Terms for Scopus database}
-	\label{table:name}
-\end{table}
+%\appendix
+%\section{\\Appendix A - Change Log}
+%
+%\section{\\Appendix B - Searches}
+%\begin{table}
+%	\begin{tabular}{ p{12cm} }
+%	\hline
+%	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+%	Searched - 4th April 2019 - Generated 1257 Articles \\
+%	\hline
+%	\end{tabular}
+%	\caption[Table caption text]{Search Terms for Scopus database}
+%	\label{table:name}
+%\end{table}
+%
+%\begin{table}
+%	\begin{tabular}{ p{12cm} }
+%	\hline
+%	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+%	Searched - 4th April 2019 - Generated 1257 Articles \\
+%	\hline
+%	\end{tabular}
+%	\caption[Table caption text]{Search Terms for ProQuest database}
+%	\label{table:name}
+%\end{table}
+%
+%\begin{table}
+%	\begin{tabular}{ p{12cm} }
+%	\hline
+%	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+%	Searched - 4th April 2019 - Generated 1257 Articles \\
+%	\hline
+%	\end{tabular}
+%	\caption[Table caption text]{Search Terms for IEEE Xplore database}
+%	\label{table:name}
+%\end{table}
 
-\begin{table}
-	\begin{tabular}{ p{12cm} }
-	\hline
-	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
-	Searched - 4th April 2019 - Generated 1257 Articles \\
-	\hline
-	\end{tabular}
-	\caption[Table caption text]{Search Terms for ProQuest database}
-	\label{table:name}
-\end{table}
-
-\begin{table}
-	\begin{tabular}{ p{12cm} }
-	\hline
-	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
-	Searched - 4th April 2019 - Generated 1257 Articles \\
-	\hline
-	\end{tabular}
-	\caption[Table caption text]{Search Terms for IEEE Xplore database}
-	\label{table:name}
-\end{table}
 \end{document}

### 2019-08-03 Source/Paper 2 - Systematic Review/Systematic Review.tex
updated Chapter 1
updated Paper3\

@@
 
-\title{Applying Natural Language Processing and Machine Learning techniquess to aid in the diagnosis of Mild Cognitive Impairment and Early Dementia - A Systematic Review of the current literature}
+\title{Applying Natural Language Processing and Machine Learning techniques to aid in the diagnosis of Mild Cognitive Impairment and Early Dementia - A Systematic Review}
 
@@
 	\texttt{alcantaj@aston.ac.uk} \\
-		\and
+		\AND
 	Peter Sawyer \\
@@
 	\texttt{p.sawyer@aston.ac.uk}
-		\and
+		\AND
 	George Vogiatzis\\
@@
 	\texttt{g.vogiatzis@aston.ac.uk} \\
-		\and
+		\AND
 	Cristina Romani\\
@@
 \usepackage{arxiv}
+\usepackage{float} % allows precise placement of tables
 \usepackage[utf8]{inputenc} % allow utf-8 input
@@
 \par
-In 2009, the Department of Health designed it's National Dementia strategy and as part of this made early diagnosis and support one of it's key priorities \cite{England2009}.
 A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
@@
 \par
+The question to be addressed in this systematic review is how has the field of machine learning and natural language processing addresses language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease.
 The potential impact of this research in this area is immense.
-Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease.
+Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease however the absence of a single test and the complexity of AD can create significant delays in diagnosis.
 Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
-%% This is an important point to fit in.
- However, the absence of a single test and the complexity of AD create delays in diagnosis.
-
-The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to to time-consuming and expensive psychological tests.
-There is a lot of research into the analysis of language as a bio-marker for MCI and Early Dementia.
-Given that sampling a person's language is relatively effortless, my research looks at whether we can find bio-markers of MCI and Early Dementia in natural language.\newline
-Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other.
-I'm not going to control for memory problems as a potential confound, but does this weaken the research?
-How do I defend this?
-\newline
 \par 
-Introduction to the problem of dementia in the context of the wider world including quality of life and financial implications.
-Exploration of dementia as a syndrome rather than a disease, and a look at the different variants of dementia.
-A look at the rationale behind research into the early diagnosis of dementia as well as a brief look at what has been done in the area (wide context, so pharmacological and psychological).
-\newline
-
-The question to be addressed in this systematic review is how has the field of machine learning and natural language processing addresses language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease.
-
 The remainder of this article is organized as follows.
 Section~\ref{methodology} gives an account of the process of this Systematic Review.
-Our new and exciting results are described in Section~\ref{results}.
+Our results are described in Section~\ref{results}.
 We discuss the results and implications in Section~\ref{discussion}.
@@
 \section{Methodology}\label{methodology}
-
 A systematic literature review (SLR) describes a process which aims to identify, evaluate and interpret the research and literature in a given area.
@@
 \par 
-Common motivations for conducting a SLR are: 
+Common motivations for conducting an SLR are: 
 \begin{enumerate}
@@
 
-Performing a SLR comprises the following steps: 
+Performing an SLR comprises the following steps:
+
 \begin{enumerate}
-	\item (i) identify the need for performing the SLR.
-	\item (ii) formulate research questions.
-	\item (iii) execute a comprehensive search and selection of primary studies.
-	\item (iv) assess the quality and extract data from the studies.
-	\item (v) interpret the results.
-	\item (vi) report the SLR.
+	\item identify the need for performing the SLR.
+	\item formulate research questions.
+	\item execute a comprehensive search and selection of primary studies.
+	\item assess the quality and extract data from the studies.
+	\item interpret the results.
+	\item report the SLR.
 \end{enumerate}
 
-The main research question this SLR aims to address is: “How has the field of machine learning and natural language processing addresses language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease?”.
-This main question was decomposed further into five research questions:
-\begin{enumerate}
-	\item RQ1: Which ML and MS techniques are being used in the dementia and comorbidities research?
-	\item RQ2: What data characteristics (variables, determinants and indicators) are being considered when applying the ML or/and MS techniques (physiological, demographic/social, genetics, lifestyle etc)?
-	\item RQ3: What are the goals of the studies that employ ML or MS techniques for prognosis of dementia and comorbidities?
-	\item RQ4: How is data censoring being handled in the studies?
-	\item RQ5: Do the studies focus on individuals or populations?
-\end{enumerate}
+The main research question this SLR aims to address is: “How has the field of machine learning and natural language processing addressed language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease?”.
+This main question can be decomposed further into four research questions:
+\begin{itemize}
+	\item RQ1: Which NLP and ML techniques are being used in dementia research?
+	\item RQ2: What features / data characteristics of text (variables, determinants and indicators) that are considered when applying the ML techniques (n-grams, PoS Tagging etc)?
+	\item RQ3: What are the goals of the studies that employ NLP / ML techniques for prognosis of dementia?
+	\item RQ4: Do the studies focus on time as factor?
+\end{itemize}
 
@@
 \subsection{Search strategy}
-To address the research questions, a search string was defined using the PICO approach, which decomposes the main question into four parts: population, intervention, comparison and outcome.
-The comparison component was discarded because the SLR was mainly concerned with a characterization.
-For each of the remaining components, keywords were derived and their rationale can be represented as follows:
+To address the research questions, a search string was defined using the PICO approach, which decomposes the main question into four parts: 
 \begin{enumerate}
-	\item Population: Studies that present research on dementia and comorbidities.
-Dementia’s keywords were selected from the “Systematized Nomenclature of Medicine–Clinical Terms” and selected by A4.
-Comorbidities’ keywords were extracted from the Marengoni et al.
-SLR in this topic.
-	\item Intervention: ML or MS techniques.
+	\item Population - Studies that present research on mild cognitive impairment and dementia.
+Mild Cognitive Impairment and Dementia keywords were selected from the Systematized Nomenclature of Medicine–Clinical Terms (SNOMED-CT).
+	\item Intervention - Intervention: ML or MS techniques.
 The ML keywords were selected from the branch “Machine Learning Approaches” of the “2012 ACM Computing Classification System”.
 The MS keywords were selected by A2.
-	\item Outcome: Prognosis on dementia and comorbidies. The prognosis keywords were provided
+	\item Comparison
+	\item Outcome - Outcome: Prognosis on dementia and comorbidies. The prognosis keywords were provided
 by A4.
@@
 
-The automated searches were performed in the Pubmed, Web of Science and Scopus databases.
+The automated searches were performed in the Pubmed, Web of Science, Scopus and IEEE databases.
 Table 1 shows the search string used for the Pubmed automated search, but note that this search string was adapted to each of the other databases’ search context.
 
-\begin{table}
+\begin{table}[H]
+	\begin{center}
 	\begin{tabular}{ p{12cm} }
@@
 	\end{tabular}
+	\end{center}
 	\caption[Table caption text]{Search Terms for Web of Science database}
@@
 
-\begin{table}
-	\begin{tabular}{ p{12cm} }
-	\hline
-	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
-	Searched - 4th April 2019 - Generated 1257 Articles \\
-	\hline
-	\end{tabular}
-	\caption[Table caption text]{Search Terms for Scopus database}
-	\label{table:name}
-\end{table}
-
-\begin{table}
-	\begin{tabular}{ p{12cm} }
-	\hline
-	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
-	Searched - 4th April 2019 - Generated 1257 Articles \\
-	\hline
-	\end{tabular}
-	\caption[Table caption text]{Search Terms for ProQuest database}
-	\label{table:name}
-\end{table}
-
-\begin{table}
-	\begin{tabular}{ p{12cm} }
-	\hline
-	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
-	Searched - 4th April 2019 - Generated 1257 Articles \\
-	\hline
-	\end{tabular}
-	\caption[Table caption text]{Search Terms for IEEE Xplore database}
-	\label{table:name}
-\end{table}
-
-\begin{table}
+\begin{table}[H]
+	\begin{center}
 	\begin{tabular}{ p{3cm} | p{3cm} | p{3cm} | p{3cm} }
@@
 	\end{tabular}
+	\end{center}
 	\caption[Table caption text]{Totals from each database}
@@
 Where JA could not reach a decision, PS and GV were consulted and majority vote on inclusion was conducted.
-
+\par 
+% Check these numbers
 After these evaluations, 37 papers were selected.
 Then a one-iteration backward snowballing process was carried out looking for possible additional studies.
-The 1199 new identified studies were assessed analogously as the previous ones, resulting in 41 new selected papers.
-Throughout the whole selection process, A3 and A4 acted in conflict resolution in the case where A1 and A2 couldn’t reach an agreement.
+There were 1199 newly identified studies were assessed analogously as the previous ones, resulting in 41 new selected papers.
+Throughout the whole selection process, PS and GV acted as additional assessors in the case where there was uncertainty about whether a paper should be included.
 \par 
 In total, 78 papers were selected to be fully read and assessed regarding its eligibility.
-The ones that successfully passed the established criteria previously defined in the protocol, had their relevant data extracted.
+The ones that successfully passed the filtering criteria described earlier had their relevant data extracted.
 In order to minimize the chance of selecting studies with bias evidence, a quality assessment questionnaire was used.
@@
 If the grading attributed to a paper fell below 8 points (out of a total of 12), it would be rejected for quality reasons.
-The 8-point threshold was
-decided in the research group discussions involving all the authors.
+The 8-point threshold was decided in the research group discussions involving all the authors.
 In this phase, a paper could also be rejected due to inclusion and exclusion criteria because the selection process adopted an inclusive approach.
@@
 
-\begin{table}
+\begin{table}[H]
+	\begin{center}
 	\begin{tabular}{ p{6cm} | p{6cm} }
@@
 	\label{table:name}
+	\end{center}
 \end{table}
 
-\begin{table}
+\begin{table}[H]
+	\begin{center}
 	\begin{tabular}{ p{6cm} | p{6cm} }
@@
 	Dataset Categories & Classes in which the data units were divided into. \\ \hline
-	Handling of Censored Data & Description of the way in which censored data was handled. \\ \hline
 	Follow-up period & Period of time, which the data units were followed. \\ \hline
-	Data Analysis Techniques & ML or MS techniques that were used to build the prognostic models. \\ \hline 
-	Model Variables & The variables used in building the prognostic models. \\ \hline
-	Aim of the Study & The goal of the built prognostic models. \\ \hline 
-	Focus of the Study & If the built prognostic models aim its predictions on an individual or population level. \\
+	Techniques used & Natural Language techniques AND/OR ML techniques that were used to build the diagnostic models. \\ \hline 
+	Features generated & NLP generated features used in building the diagnostic models. \\ \hline
+	Aim of the Study & The goal of the built diagnostic models. \\
 	\hline 
 	\end{tabular}
-
+	\end{center}
 \end{table}
@@
 \section{Results}\label{results}
+\subsection{Machine Learning methods}
+Intro - Addresses Research Question 1
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ p{6cm} | p{6cm} }
+	\hline
+	Machine Learning technique used & Paper Number \\ \hline
+	Logistic Regression & Orimaye \\ \hline
+	Support Vector Machines & Orimaye \\ \hline
+	Linear Discriminant Analysis & Orimaye \\ \hline
+	Decision Trees & Orimaye \\ \hline	
+	\hline 
+	\end{tabular}
+	\end{center}
+\end{table}
+
+\subsection{Traditional Machine Learning methods}
+\subsubsection{Logistic Regression}
+
+\subsubsection{Support Vector Machines}
+
+\subsubsection{Linear Discriminant Analysis}
+
+\subsubsection{Decision Trees}
+
+Curry, Singer and Habash - Alz (66.7 - Pres, 66.7 - Recall) and Control (67.9 - Pres, 67.9 Recall)
+
+\subsubsection{Naive Bayes}
+
+Curry, Singer and Habash - Alz(80.8 - Pres, 0.75 - Recall) and Control (79.3 - Pres, 82.1 Recall) - In the naive bayes classifier, they identified that pauses, go-ahead, fillers and incomplete words as the most significant features.
+%Why is Recall at 0.75???????
+
+\subsection{Deep Learning methods}
+\subsubsection{Deep language space neural network for classifying mild cognitive impairment and Alzheimer-type dementia - Orimaye, Wong and Wong (2018)}
+In this paper, Orimaye et al use deep-deep neural networks language models (D2NNLM) to learn linguistic changes that distinguish the language of patients with MCI and AD-type dementia from the healthy controls using higher order n-grams.
+An ordinary DNNLM uses lower order n-gram N-dimensional sparse vectors as descrete feature representations to train the neural network with multiple hidden layers.
+
 \subsection{Features of Language}
-\subsubsection{Type token ratio(TTR)}
+Intro - Addresses Research Question 2
+
+\par
+Asgari, Kaye and Dodge (2017) \cite{Asgari2017} used another form of word frequency measurement.
+Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words \cite{Pennebaker2015}.
+They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
+\subsubsection{Measures of Semantic Complexity}
+Intro!
+
 Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
@@
 \end{equation}
-\subsubsection{Brunet's Index(W)} % Don't forget to cite the original article.
+\par 
 Brunet's Index (W) differentiates itself for TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
@@
 \par 
-\subsubsection{Honore's Statistic (R)} % Don't forget to cite the original article.
 Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation: 
@@
 where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
+\par 
+Guinn, Singer and Habash \cite{Guinn2015}, they found in their corpus that there was no significant difference between interviewers and those with dementia when applying these measures.
+However, when they compared these results with a control dataset, they did find a significant difference in Honore's statistic.
+They explained this interesting results by suggesting that the interviewer used was trying to match (intentionally or unintentionally) the lexical richness of the person they were interviewing.
+In comparing healthy older adults with those suffering from dementia, they found that there was an increase in lexical diversity which is contrary to other research.
+They did note that conversations involving those with dementia contained roughtly 50\% less speech than those with controls, and that TTR in particular is not a suitable measure because it is more sensitive to length.
+They also note that Brunet's Index and Honore's Statistic are better statistics that control for the total length of the conversation and controls had statistically greater lexical diversity on those measures.
+
 \subsection{Quantity - Total number of words}
@@
 The general hypothesis motivating these studies is that either working memory limitations or semantic memory limitations in AD affect one's ability to use complex constructions.
-\newline
 \par
-\subsection{N-grams}
+\subsection{N-grams and skip-grams}
 One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
@@
 Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
-\newline
 \par
-Asgari, Kaye and Dodge (2017) \cite{Asgari2017} used another form of word frequency measurement.
-Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words \cite{Pennebaker2015}.
-They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
-\newline
+Skip-grams are a variant of n-grams in which word tokens are skipped intermittently while creating n-grams.
+For example, take the sentence 'I am going to London', there are four conventional bigrams: 'I am', 'am going', 'going to', and 'to London' - using skip-grams, we might skip a word to create additional bigrams such as: 'I going' and 'going London'.
+Orimaye defined k-skip-n-grams as a set of n-gram tokens with the following equation, where n is the specified n-gram (e.g.
+2 for a bigram and 3 for a trigram), m is the number of tokens in a given sentence, k is the number of word skip between n-grams given that k < m and a = \{1,..., m-n\}
 \par 
+One problem with this approach is existing research currently uses language generated from picture description tasks.
+Given the nature of these tasks, the language generated is relatively constrained in comparison to language generated spontaneously.
+% Question How does using ngrams with constrained language differ from ngrams with general speech?
+
+\begin{equation} \label{x4}
+T_{n-gram} = {W_{a},...,W_{a+n-k},..., W_{a+n},....,W_{m-n},...,W_{(m-n)+n-k},...,W_{m}}
+\end{equation}
+
+Thus for the sentence 'I am going to London', 1-skip-2-grams will give \{I going, am to, going London\} and 1-skip-3-grams will give \{I going to, I am to, am to London, am going London\}
+
 \subsection{Mean length of utterance (MLU)}
@@
 \par
+Lu's Syntactic Complexity Analyser
+Ygnve measure
+
+\subsection{Semantic features}.
+We compute semantic similarity using the average and minimum cosine distance between each pair of one-hot embeddings of utterances, and the cosine cutoff (i.e., the number of pairs of utterances whose the cosine distance is below a certain threshold).
+We compute word specificity and ambiguity based on tree depth and the number of senses in WordNet [54].
+We also extract multiple WordNet measures of similarity: Resnik [68], Jiang-Coranth [69], Lin [47], Leacock-Chodorow [70], and Wu-Palmer [71].
+
+\subsection{Syntactic features}
+Guinn, Singer and Habash \cite{Guinn2015} found that syntactic features such as Noun rate, verb rate, adjective rate and pronoun rate were not significantly different between interviewers and those with dementia in their dataset, although they did note that there was a slightly higher but non significant use of pronouns.
+
+\subsection{Pragmatic features.} We train a general 100-topic latent Dirichlet allocation (LDA) model [72] on the Wikipedia corpus for generalizability.
+LDA is a generative statistical model used to determine unlabeled topics in a document.
+For each transcript, we extract the probabilities of each LDA topic.
+Next, we extract features related to rhetorical structure theory (RST), which is a classic framework for discourse parsing in which partitions of text are arranged in a tree structure by pragmatic relations such as Elaboration or Contrast [73].
+
+\subsubsection{Go-ahead Utterances}
+Go-ahead utterances are defined as short one or two syllable responses that do not contribute to the conversation beyond a minimal response.
+The function of these go-ahead utterances can be to validate what the other person is saying, or to agree / disagree with what is being said.
+Another function can be that they wish speaker is indicating that they have nothing futher to add to the conversation and a signal to the other speaker to continue with what they are saying.
+In Curry, Singer and Habash's research, they found that in comparing interviewers and those with dementia, that interviewers used significantly fewer go-ahead utterances.
+In comparing controls and those with dementia, they found that there was a relative lack of go-ahead utterances in the controls which implies that controls had a lot more contributions to make in their conversations than those with dementia.
+\par
+
 \subsection{Formulaic Language}
@@
 \newline
-\subsection{N-grams}
-One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
-An n-gram is a contiguous sequence of n items from a given sample of text or speech.
-The items can be phonemes, syllables, letters, words or base pairs according to the application.
-For example, given the sequence of words "to be or not to be", this extract is said to contain six 1-gram sequences (to, be, or, not, to, be), five 2-gram sequences (to be, be or, or not, not to, to be), four 3-gram sequences(to be or, be or not, or not to, not to be) and so on.
-This is useful as, given a large portion of text or speech, we can predict the probability of a word being close by to a given word.
-A number of researchers have used n-grams as features.
-One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class.
-Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls.
-Their main finding supported n-grams as the most significant predictor.
-One of the criticisms is the use of picture description tasks and n-grams.
-Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
-\newline
-\par
-Asgari, Kaye and Dodge (2017) \cite{Asgari2017} used another form of word frequency measurement.
-Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words \cite{Pennebaker2015}.
-They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
-\newline
 
-\section{Machine Learning methods}
-\subsection{Traditional Machine Learning methods}
-\subsection{Deep Learning methods}
-\subsubsection{Deep language space neural network for classifying mild cognitive impairment and Alzheimer-type dementia - Orimaye, Wong and Wong (2018)}
-In this paper, Orimaye et al use deep-deep neural networks language models (D2NNLM) to learn linguistic changes that distinguish the language of patients with MCI and AD-type dementia from the healthy controls using higher order n-grams.
-An ordinary DNNLM uses lower order n-gram N-dimensional sparse vectors as descrete feature representations to train the neural network with multiple hidden layers.
+\subsection{Number of syllables and Characters}
 
-\subsubsection{Deep Neural Network Language Models}
+\subsection{Number of fillers}
 
-\subsection{Aided Diagnosis of Dementia Type through Computer-Based Analysis of Spontaneous Speech - Jarrold, Peintner, Wilkins, Vergryi, Richey, Gorno-Tempini and Ogar}
+\subsection{Readability}
+Flesch reading score, Flesch-kincaid grade level
 
-\subsection{Language Analysis of Speakers with Dementia of the Alzheimer's Type - Guinn and Habash (2012)}
+\subsection{Polarity}
 
-\subsection{Features and Machine Learning Classification of Connected Speech Samples from patients with Autopsy Proven Alzheimer's Disease with and without additional Vascular Pathology - Rentoumi, Raoufian, Ahmed, de Jager, Garrard (2014)}
+\subsection{Frequency}
+Mean values of frequency, age of acquisition, imageability, familiarity, arousal, dominance and valence based on lexical norms
 
-\subsection{Learning Predictive Linguistic Features for Alzheimer's Disease and related Dementias using Verbal Utterances - Orimaye, Wong and Golden (2014)}
+\subsection{Dysfluencies}
+Curry, Singer and Habash noted that in comparing controls and those with Dementia, that those with Dementia had a significantly higher number of pauses per word and a much higher incidence of words that were truncated in mid-speech.
+In comparing interviewers with those with dementia, they also showed other signs of difficulties with fluency with higher rates of incomplete words, filler words and repeated words.
 
-\subsection{Learning Lingustic Biomarkers for Predicting Mild Cognitive Impairment using Compound Skip-grams - Orimaye, Wong, Wong and Tai (2015)}
+\subsection{What are the goals of the studies that employ ML or MS techniques for prognosis of dementia and comorbidities?}
 
-\subsection{Diagnosing People with Dementia using Automatic Conversation Analysis - Mirheidari, Reuber, Blackburn, Walker (2016)}
-
-\subsection{Formulaic Language in People with Probably Alzheimer's Disease: A Frequency-Based Approach - Zimmerer, Wibrow and Varley (2016)}
-
-\subsection{Linguistic Features Identify Alzheimer's Disease in Narrative Speech - Fraser, Melzer and Rudzicz (2016}
-
-\subsection{Declines in Connected Language Are Associated with Very Early Mild Cognitive Impairment: Results from the Wisconsin Registry for Alzheimer's Prevention - Mueller, Koscik, Hermann, Johnson and Turkstra (2018)}
-
-\section{Discussion}\label{discussion}
+\section{Discussion and conclusions}\label{discussion}
 \subsection{Discussion of the current evidence}
@@
 
+\appendix
+\section{Article Table}
+\begin{table}
+	\begin{tabular}{ p{1cm} | p{6cm} | p{6cm}}
+	\hline
+	Paper Number & Paper Title & Authors \\ \hline
+	1 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	2 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	3 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	4 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	5 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	6 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	7 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	8 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	9 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	10 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	11 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	12 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	13 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	14 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	15 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	16 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	17 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	18 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	19 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	20 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	21 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	22 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	23 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	24 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	25 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	26 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	27 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	28 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	29 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	30 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	31 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	32 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	33 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	34 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	35 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	36 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	37 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	39 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	40 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	41 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
+	\end{tabular}
+\end{table}
 \end{document}
+
+%\begin{table}
+%	\begin{tabular}{ p{12cm} }
+%	\hline
+%	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+%	Searched - 4th April 2019 - Generated 1257 Articles \\
+%	\hline
+%	\end{tabular}
+%	\caption[Table caption text]{Search Terms for Scopus database}
+%	\label{table:name}
+%\end{table}
+%
+%\begin{table}
+%	\begin{tabular}{ p{12cm} }
+%	\hline
+%	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+%	Searched - 4th April 2019 - Generated 1257 Articles \\
+%	\hline
+%	\end{tabular}
+%	\caption[Table caption text]{Search Terms for ProQuest database}
+%	\label{table:name}
+%\end{table}
+%
+%\begin{table}
+%	\begin{tabular}{ p{12cm} }
+%	\hline
+%	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+%	Searched - 4th April 2019 - Generated 1257 Articles \\
+%	\hline
+%	\end{tabular}
+%	\caption[Table caption text]{Search Terms for IEEE Xplore database}
+%	\label{table:name}
+%\end{table}

### 2019-08-03 Source/Thesis/Chapter1.tex
updated Paper3\

@@
 Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD.
-There is some evidence to support this.
-Berisha et al \cite{Berisha2015}, has shown through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control.
+As is the case with other quantitative measures of cognitive ability, contrasting individuals with MCI and Early AD is challenging due to the variation in an individual's baseline speech capacity.
+It is therefore prudent to measure cognitive decline over time as suggested by Albert et al \cite{Albert2011}.
+There is some evidence to support this approach, Berisha et al \cite{Berisha2015} demonstrated through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control.
 \par
@@
 We look at the effectiveness of automatic speech recognition at translating speech into language, as well as the minimum quality of data that needs to be passed to any pipepline we develop.
-The Talk2Me app collects and then generates approximately 2000 lexico-syntactic, acoustic and semantic features (\cite Kommeili2019} as part of their data collection framework.
+The Talk2Me app collects and then generates approximately 2000 lexico-syntactic, acoustic and semantic features {\cite Komeili2019} as part of their data collection framework.
 Chapter 4 discusses the challenges connected to constructing an effective data pipeline and contains two experiments.

### 2019-08-06 Source/Paper 2 - Systematic Review/Systematic Review.tex
working on paper

@@
 \usepackage{microtype}      % microtypography
-\usepackage{lipsum}
+\usepackage{longtable}		% for use in table in Appendix A
 
@@
 
+\subsection{Search strategy}
 The main research question this SLR aims to address is: “How has the field of machine learning and natural language processing addressed language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease?”.
-This main question can be decomposed further into four research questions:
-\begin{itemize}
-	\item RQ1: Which NLP and ML techniques are being used in dementia research?
-	\item RQ2: What features / data characteristics of text (variables, determinants and indicators) that are considered when applying the ML techniques (n-grams, PoS Tagging etc)?
-	\item RQ3: What are the goals of the studies that employ NLP / ML techniques for prognosis of dementia?
-	\item RQ4: Do the studies focus on time as factor?
-\end{itemize}
 
-The present paper builds upon these questions and additionally presents the results of the other two additional research questions.
-Further, the key terms related to comorbidities were included in the search string to ensure that relevant studies about ML or MS for the prognosis of a disease, where dementia is considered a comorbidity to that disease would also be retrieved from the database searches, even when the term dementia was not mentioned in the paper’s title or abstract.
-
-\subsection{Search strategy}
-To address the research questions, a search string was defined using the PICO approach, which decomposes the main question into four parts: 
+This SLR builds upon these questions and additionally presents the results of the other two additional research questions.
+Further, the key terms related to MCI and the other names by which it is known were included in the search string to ensure that relevant studies about machine learning and the diagnosis of a disease were also retrieved, even if not specifically mentioned in the paper's title or abstract.
+\par 
+To address the research questions, a search string was defined using the PICO approach, which decomposes the main research question into four parts: 
 \begin{enumerate}
@@
 
+Using this process, the main research question was decomposed  into four research questions:
+\begin{itemize}
+	\item RQ1: Which NLP and ML techniques are being used in dementia research?
+	\item RQ2: What features / data characteristics of text (variables, determinants and indicators) that are considered when applying the ML techniques (n-grams, PoS Tagging etc)?
+	\item RQ3: What are the goals of the studies that employ NLP / ML techniques for prognosis of dementia?
+	\item RQ4: Do the studies focus on time as factor?
+\end{itemize}
+\par 
 The automated searches were performed in the Pubmed, Web of Science, Scopus and IEEE databases.
@@
 	\end{center}
-	\caption[Table caption text]{Search Terms for Web of Science database}
+	\caption[Table caption text]{Example of Search Terms for Web of Science database}
 	\label{table:name}
@@
 	\end{center}
-	\caption[Table caption text]{Totals from each database}
+	\caption[Table caption text]{Number of articles found from each database - complete search terms appear in Appendix B}
 	\label{table:name}
 \end{table}
-
-This yielded 25 studies for review, 43 papers were reviewed.
 
@@
 Each paper was initially reviewed by just the title and the abstract.
-This yielded a total of 67 potential papers.
-Papers were then evaluated by the JA based on the inclusion and exclusion criteria defined previously (see Table 6).
+This yielded a total of 25 potential papers.
+Papers were then evaluated by the JA based on the inclusion and exclusion criteria  (see Table 3).
 Where JA could not reach a decision, PS and GV were consulted and majority vote on inclusion was conducted.
-\par 
-% Check these numbers
-After these evaluations, 37 papers were selected.
-Then a one-iteration backward snowballing process was carried out looking for possible additional studies.
-There were 1199 newly identified studies were assessed analogously as the previous ones, resulting in 41 new selected papers.
-Throughout the whole selection process, PS and GV acted as additional assessors in the case where there was uncertainty about whether a paper should be included.
-\par 
-In total, 78 papers were selected to be fully read and assessed regarding its eligibility.
-The ones that successfully passed the filtering criteria described earlier had their relevant data extracted.
-In order to minimize the chance of selecting studies with bias evidence, a quality assessment questionnaire was used.
-This questionnaire was adapted from Kitchenham’s guidelines and can be found in the SLR protocol.
-If the grading attributed to a paper fell below 8 points (out of a total of 12), it would be rejected for quality reasons.
-The 8-point threshold was decided in the research group discussions involving all the authors.
-In this phase, a paper could also be rejected due to inclusion and exclusion criteria because the selection process adopted an inclusive approach.
-This means that during the reading of the titles and abstracts, in the case where the information provided was incomplete or too general it was selected to be fully read in the posterior phase.
-A common example is the case when the data analysis technique specified in the abstract was merely “classification”, so it was not possible to know if any machine learning occurred.
-\par
-As in the study selection, a quality assessment evaluation round was performed beforehand to ensure consistency in the evaluations.
-A1, A2 and A5 participated in this task.
-In total, 37 studies composed the final set of included primary studies and had their relevant data extracted, 7 papers were rejected due quality reasons, and 34 papers were rejected due to failing the inclusion and exclusion criteria.
-One reason for the high number of the latter was the decision to exclude the papers that used solely statistical methods as data analysis tech- niques to build the prognostic models.
-The selected studies were also assessed for the risk of cumulative evidence bias.
-This was done by checking, in the case of the same research group with different studies in the final set of included primary studies, if it was justified having both studies (i.e different samples).
-
-A search of the literature was conducted using ProQuest (PsychArticles), SCOPUS, Web of Science.
-The following results were found (Table 1).
-All papers were then reviewed for relevance by reading the abstract and full text where appropriate and a shortlist was compiled.
-An additional search through references of shortlisted papers was also conducted and any papers who upon further review appeared relevant were added to the shortlist.
-Papers were included where researchers used machine learning to classify participants as MCI, AD or Healthy using language.
-We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English.
-
-\subsection{Data collection}
-For the data collection, a base extraction form was defined in the protocol, but later in the study it was evolved based on the research group discussions.
-Table 3 lists and defines the collected variables.
 
@@
 
+After the identification of these papers, a one-iteration backward snowballing process was carried out using the reference lists of the original set of papers looking for studies that were missed in the original searches.
+This resulted in 41 additionally identified papers.
+Throughout the whole selection process, PS and GV acted as additional assessors in the case where there was uncertainty about whether a paper should be included.
+\par 
+In total, 66 papers were selected to be fully read.
+A quality assessment questionnaire (see Table 4) was developed based on Kitchenham's guidelines and was used to minimize the chance of bias in the selection process.
+Studies were graded on a 12 point scale and any studies which scored less than 8 points were excluded for quality reasons.
+The ones that successfully passed the filtering criteria described earlier had their relevant data extracted.
+
 \begin{table}[H]
@@
 	\end{tabular}
+	\caption[Table caption text]{Quality Assessment Questionnaire}
+	\label{table:name}
 	\end{center}
 \end{table}
+
+In this phase, a paper could also be rejected due to inclusion and exclusion criteria because the selection process adopted an inclusive approach.
+This means that during the reading of the titles and abstracts, in the case where the information provided was incomplete or too general it was selected to be fully read in the posterior phase.
+A common example is the case when the data analysis technique specified in the abstract was merely “classification”, so it was not possible to know if any machine learning occurred.
+\par
+% NUMBERS!
+In total, 37 studies composed the final set of included primary studies and had their relevant data extracted, 7 papers were rejected due quality reasons, and 34 papers were rejected due to failing the inclusion and exclusion criteria.
+One reason for the high number of the latter was the decision to exclude the papers that used solely statistical methods as data analysis techniques to build the prognostic models.
+The selected studies were also assessed for the risk of cumulative evidence bias.
+This was done by checking, in the case of the same research group with different studies in the final set of included primary studies, if it was justified having both studies (i.e different samples).
+
+\subsection{Data collection}
+For the data collection, a base extraction form was defined in the protocol, but later in the study it was evolved based on the research group discussions.
+Table 4 lists and defines the collected variables.
 

### 2019-08-14 Source/Paper 2 - Systematic Review/Systematic Review.tex
fixed table
updated Paper 2

@@
 \subsubsection{Logistic Regression}
+Logistic Regressions is a probabilistic approach to classification and is dependant on the assumption that your input space can be separated into two distinct regions, one for each class, by a linear boundary.
+In 2D space, this boundary is a line and in 3D space, this boundary is defined as a plane.
+\par 
+In the results of this survey.
 
 \subsubsection{Support Vector Machines}
+Support Vector Machine (SVM) was originally proposed as an algorithm for classification problems; it is a relatively new technique compared to the other ML approaches. The classification
+process consists of mapping the data points (usually the study subjects) into a feature space composed of the variables that characterize these data points, except for the outcome variable.
+Then, the algorithm finds patterns in this feature space by defining the maximum separation between two or more classes, depending on the problem to be solved. Contrary
+to some regression techniques, SVMs are not dependent on a pre-determined model for data fitting, although there are still algorithm specifications to be considered (e.g. choice of a kernel
+function); instead, it is a data-driven algorithm that can work relatively well in a scenario where sample sizes are small compared to the number of variables, reason why it has been widely employed by diagnostic studies in tasks related to the automated classification of diseases.
+\par 
+Regarding the SLR results, SVMs were present in ??/41 selected studies, in 38 proposed models, and being by far the most used machine learning technique in the dementia diagnosis research.
+These numbers account for the traditional SVM and variations.
+In all of the ??
+selected studies the SVMs focused at binary classifications where the task was to discriminate mild cognitively impaired (MCI) patients that will or will not develop Alzheimer's Disease (AD).
+In the general case, the problem is posed as either MCI converters versus MCI non-converters, or progressive MCI versus stable MCI classification.
+This outlines a situation in which a regression problem (when will the MCI patients convert to AD?) is formulated as a classification problem (which MCI patients will convert to AD in X months?) to be solved.
+Reasons for this could be due to limitations in the data used, i.e.
+the limited follow-up periods of the subjects included in the studies.
 
@@
 \subsubsection{Decision Trees}
+A Decision Tree (DT) is a classification algorithm in which the learned knowledge is represented in a tree structure that can be translated to if-then rules. DT's learning process is recursive
+and starts by testing each input variable as to how well each of them, alone, can classify the labeled examples. The best one is selected as a root node for the tree and its descendant nodes
+are defined as the possible values (or relevant ratios) of the selected input variable.
+The training set is then classified between the descendant nodes according to the values of the selected input variable.
+This process is repeated recursively until no more splits in the tree are possible.
+Like SVMs, DTs do not depend on a pre-defined model and are mostly used
+to find important interactions between variables. Being intuitive and easy to interpret, DTs have been used in prognostic studies as a tool for determining prognostic subgroups.
+In this SLR, DTs were the second most frequently used ML technique, present in 6/37 selected studies and proposed in 7 models.
+It was employed for the same reason as SVM; except for one study that investigated the evolution of patients diagnosed with cognitive impairment no dementia (CIND) to AD.
 
@@
 %Why is Recall at 0.75???????
+
+\subsubsection{Artificial Neural Networks}
+An Artificial Neural Network (ANN) is a methodology that performs multifactorial analyses, which is desirable in the health area as medical decision-making problems are usually dependent of many factors.
+An ANN is composed of nodes connected by weighted edges in a multi-layer architecture that comprises: an input layer, one or more hidden layers and an output layer.
+In the training process, inputs and outputs values are known to the network, while the weights are incrementally adjusted so that the outputs of the network are approximate to the known outputs.
+Despite being a powerful predictor, ANNs are `black boxes', which means that they are not able to explain their predictions in an intuitive way, contrary to DTs or BNs.
+Also, they require the specification of the architecture to be used beforehand (i.e.
+the number of hidden layers).
+
+\subsubsection{K-nearest neighbours}
+K Nearest Neighbors (KNN) is a classification algorithm that takes a data point from an unknown class and assigns it as an input vector in the feature space.
+Then, the classification process follows by assigning the unknown class data point to the class in which the majority of the K nearest data points belong to.
+The distance between data points is usually measured by Euclidean distance, but it is possible to employ other measures.
+KNN is one of the simplest ML classification algorithms and have been used in a wide range of applications; however, it can be computationally expensive in a highly dimensional scenario.
+Further, it considers all features to be equally weighted, which can be a problem if the data has superfluous attributes.
 
@@
 \subsection{Discussion of the current evidence}
+One of the criticisms of traditional learning models is it's reliance on features that are generated for them.
+As we can see from the brief look at some of the research in this field, researchers have taken a number of different approaches with success.
+However, there does not seem to be a consensus.
+As machine learning in general moves away from traditional machine learning models to deep learning techniques, it poses the question can deep learning assist with this problem?
+\par 
+One of the main benefits of deep learning is that it does not rely on pre defined features being fed to the model.
+Instead, deep learning models take raw data in with some amount of preprocessing and generates it's own features.
+This move away from relying on features solves one of the difficulties we have with the current psychological literature, namely that there are some disagreement on the which
+\par 
+One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
+This is the case with all machine learning techniques, but more so with deep learning.
+There have been ways to 'create' more data such as data augmentation.
+
+Current state-of-the-art diagnostic measures of AD are invasive (CSF analysis), expensive (neuroimaging), and time-consuming (neuropsychological assessment).
+Furthermore, these measures are limited to speciality clinics and thus have limited accessibility as frontline screening and diagnostic tools for AD.
+More importantly, nonspecialists are often inaccurate at identifying early AD and MCI.
+Thus, there is an increasing need for additional noninvasive and/or cost-effective tools, allowing effective frontline identification of subjects in the preclinical or early clinical stages of AD who could be suitable for monitoring in speciality clinics and for early treatment.
+Implementation of effective screening instruments will allow diagnosis earlier in the course of dementia, even at the point when memory function is still essentially within the normal range.
+This strategy would enable an earlier, and potentially more effective, prevention and treatment of AD with a special focus to preserve cognitive functions.
+\par 
+However the literature has identified a number of challenges when approaching this problem.
+Firstly, the clinical features that combine to meet the diagnostic criteria for Dementia or it's variants are continuous in nature and heterogenous between patients and are also impacted by other variables.
+For example as shown above, cognitive performance is affected in part by a patients educational attainment and a patients ability to live independently is impacted by a their physical health as much as their cognitive health.
+The challenge therefore is to find features that are minimally impacted by other factors, or that can be controlled for by a experimental design such as a matched pairs design to control for educational attainment.
+\par 
+Another challenge lies with the recruitment of suitable individuals who may notice a decline in cognition to the point where we might classify them as having MCI, but these individuals deduce that there is little to no value to admitting there is a problem and seeking help whilst their symptoms are 'manageable'.
+\par
+Finally there is a large amount of variability in the presentations of those with MCI and early dementia, and this is compounded by an similar amount of variability in the criteria researchers have used for experimental groups and the approaches researchers have used to tackle this problem.
+This had led to a confused literature.
+Recently a call for research that has consistent inclusion / exclusion criteria has been made along with some proposed definitions of MCI and it's subgroups \cite{Petersen2014}.
+Researchers have identified the analysis of language impairment as an area of promise to explore in the diagnosis of MCI and early AD and recent developments in natural language processing and machine learning techniques have the potential to assist in this research.
+Indeed, given the increased burden on the diagnosis of MCI and AD on professionals there has been a call to use technology to potentially ease this burden \cite{Boschi2017}.
+A small but growing amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
+
 \subsection{Methodological Issues}
 \subsection{Limitations}
+
 \subsection{The future of the field}
+One of the areas for research to study is a careful examination of the features that are being used to measure language deterioration.
+For example, Zimmerer (2016) \cite{Zimmerer2016} describes connectivity in such as a way that correlates directly with what Mueller (2018) \cite{Mueller2018a} calls Fluency.
+Whilst these are very nuanced measures which differ slightly in the form they take, the sheer range of measures and features being produced make it difficult to organise and explore what is truly going on in those with MCI and early AD..
+Some work needs to be done in producing a consistent list of measures that are validated using existing datasets and can be used for future research moving forward.
+\newline
+\par
+Teng et al suggests that work should focus on the MCI population and concentrate on developing a consensus neuropsychological battery that could yield predictable rates of progression to AD \cite{Teng2009}.
+This, in conjunction with the development of a model of language, sensitive enough to detect subtle deterioration in language use to act as an additional cognitive marker to aid diagnosis could potentially move some way to providing this.
+\par 
+Future research should also be directed towards developing non-intrusive ways of detecting subtle changes in natural language such that any perceived deterioration that could indicate the presence of MCI or AD could be flagged up early.
+Machine learning approaches seem to be the most logical approach for achieving this aim as language could be collected in non-intrusive ways and passed to a machine learning algorithm for preliminary classification.
+Despite the excellent quality of datasets, for example the DementiaBank dataset, being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which some researchers could claim are limited in scope.
+Finally, the recent resurgence in the use of neural networks and deep learning could provide the answer to the confused literature in terms of features.
+A key benefit of deep learning is it's ability to automate the process of feature engineering.
+So there is an opportunity to explore the use of deep learning, to not only develop new features but also validate existing features independently.
+\par
+This area of research is extremely promising in its early results and the impact of successful research would be life changing for both individuals and the health of the worlds aging population in general.
 
 \section{Conclusions}\label{conclusions}
-We worked hard, and achieved very little.
+We can see that both Natural Language Processing and Machine Learning techniques have a lot to offer, Indeed there has been a lot of research which have used pre-existing datasets to explore this area with promising results.
+One of the difficulties with the current research is the approach of trying to discriminate between those AD and healthy controls.
+This is not necessarily a problem in the real world as it is trivially easy to do for trained clinicians.
+A more interesting, but potentially harder to problem to solve is to discriminate between those with MCI and healthy controls, and more importantly to track their decline over time.
+Further, to date the authors are unaware of any research in which these techniques are applied to newly created samples of language.
 
@@
 \section{Article Table}
-\begin{table}
-	\begin{tabular}{ p{1cm} | p{6cm} | p{6cm}}
+\begin{longtable}{ p{1cm} | p{6cm} | p{6cm}}
 	\hline
@@
 	1 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	2 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	3 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	4 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	5 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	6 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	7 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	8 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	9 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	10 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	11 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	12 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	13 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	14 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	15 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	16 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	17 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	18 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	19 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	20 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	21 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	22 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	23 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	24 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	25 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	26 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	27 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	28 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	29 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	30 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	31 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	32 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	33 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	34 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	35 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	36 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	37 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	39 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	40 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	41 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	\end{tabular}
-\end{table}
+	2 & Abnormalities of Connected Speech in Semantic Dementia vs Alzheimer’s diseases & Authors \\ \hline
+	3 & Aided Diagnosis of Dementia Type through Computer-Based Analysis of Spontaneous Speech & Authors \\ \hline
+	4 & Alignment of spoken narratives for automated neuropsychological assessment & Authors \\ \hline
+	5 & Automatic detection and rating of dementia of alzheimer type through lexical analysis of spontaneous speech & Authors \\ \hline
+	6 & Changes in Style in Authors with Alzheimer’s Disease & Authors \\ \hline
+	7 & Computer-based evaluation of Alzheimer's disease and mild cognitive impairment patients during a picture description task & Authors \\ \hline
+	8 & Computerized assessment of syntactic complexity in Alzheimer's disease: A case study of Iris Murdoch's writing & Authors \\ \hline
+	9 & Computerized neuropsychological assessment in mild cognitive impairment based on natural language processing-oriented feature extraction & Authors \\ \hline
+	10 & Contrastive conversational analysis of language production by Alzheimer's and control people& Authors \\ \hline
+	11 & Declines in Connected Language Are Associated with Very Early Mild Cognitive Impairment: Results from the Wisconsin Registry for Alzheimer’s Prevention & Authors \\ \hline
+	12 & Deep-Deep Neural Network Language Models for Predicting Mild Cognitive Impairment & Authors \\ \hline
+	13 & Domain Adaptation for Detecting Mild Cognitive Impairment & Authors \\ \hline
+	14 & Diagnosing people with dementia using automatic Conversation Analysis & Authors \\ \hline
+	15 & Deep language space neural network for classifying mild cognitive impairment and Alzheimer-type dementia & Authors \\ \hline
+	16 & Enriching Complex Networks withWord Embeddings for Detecting Mild Cognitive Impairment from Speech Transcripts & Authors \\ \hline
+	17 & Features and Machine Learning Classification of Connected Speech Samples from Patients with Autopsy Proven Alzheimer’s Disease with and without Additional Vascular Pathology & Authors \\ \hline
+	18 & Formulaic Language in People with Probable Alzheimer’s Disease: A Frequency-Based Approach & Authors \\ \hline
+	19 & Language Analysis of Speakers with Dementia of the Alzheimer’s Type & Authors \\ \hline
+	20 & Language analytics for assessing brain health: Cognitive impairment, depression and pre-symptomatic Alzheimer's disease & Authors \\ \hline
+	21 & Learning Linguistic Biomarkers for Predicting Mild Cognitive Impairment using Compound Skip-grams & Authors \\ \hline
+	22 & Learning Predictive Linguistic Features for Alzheimer’s Disease and related Dementias using Verbal Utterances & Authors \\ \hline
+	23 & Linguistic Features Identify Alzheimer's Disease in Narrative Speech & Authors \\ \hline
+	24 & Longitudinal detection of dementia through lexical and syntactic changes in writing: a case study of three British novelists & Authors \\ \hline
+	25 & Microlinguistic aspects of the oral narrative in patients with Alzheimer’s disease & Authors \\ \hline
+	26 & Multilingual word embeddings for the assessment of narrative speech in mild cognitive impairment & Authors \\ \hline
+	27 & Natural Language Features for Detection of Alzheimer's Disease in Conversational Speech & Authors \\ \hline
+	28 & NLP-Oriented Contrastive Study of Linguistic Productions of Alzheimer's and Control People & Authors \\ \hline
+	29 & Predicting mild cognitive impairment from spontaneous spoken utterances & Authors \\ \hline
+	30 & Predicting probable Alzheimer's disease using linguistic deficits and biomarkers & Authors \\ \hline
+	31 & Propositional Idea Density in women’s written language over the lifespan: Computerized analysis & Authors \\ \hline
+	32 & Sentence segmentation in narrative transcripts from neuropsychological tests using recurrent convolutional neural networks & Authors \\ \hline
+	33 & Speech processing approach for diagnosing dementia in an early stage & Authors \\ \hline
+	34 & Spoken language biomarkers for detecting cognitive impairment & Authors \\ \hline
+	35 & Talk2Me: Automated linguistic data collection for personal assessment & Authors \\ \hline
+	36 & The Effect of Heterogeneous Data for Alzheimer’s Disease Detection from Speech & Authors \\ \hline
+	37 & Toward the Automation of Diagnostic Conversation Analysis in Patients with Memory Complaintss & Authors \\ \hline
+	39 & Tracking Discourse Complexity Preceding Alzheimer's Disease Diagnosis: A Case Study Comparing the Press Conferences of Presidents Ronald Reagan and George Herbert Walker Bush & Authors \\ \hline
+	40 & Vector-space topic models for detecting Alzheimer's disease & Authors \\ \hline
+	41 & Vocabulary Size in Speech May Be an Early Indicator of Cognitive Impairment. & Authors \\ \hline
+\end{longtable}
 \end{document}

### 2019-08-22 Source/Paper 2 - Systematic Review/Systematic Review.tex
updated Thesis

@@
 Mild Cognitive Impairment and Dementia keywords were selected from the Systematized Nomenclature of Medicine–Clinical Terms (SNOMED-CT).
-	\item Intervention - Intervention: ML or MS techniques.
+	\item Intervention - Intervention: ML or Statistical Modelling techniques that focus on classification.
 The ML keywords were selected from the branch “Machine Learning Approaches” of the “2012 ACM Computing Classification System”.
@@
 \begin{itemize}
-	\item RQ1: Which NLP and ML techniques are being used in dementia research?
-	\item RQ2: What features / data characteristics of text (variables, determinants and indicators) that are considered when applying the ML techniques (n-grams, PoS Tagging etc)?
+	\item RQ1: What features / data characteristics of text (variables, determinants and indicators) that are considered when applying the ML techniques (n-grams, PoS Tagging etc)?
+	\item RQ2: Which NLP and ML or Statistical Learning techniques are being used in dementia research?
 	\item RQ3: What are the goals of the studies that employ NLP / ML techniques for prognosis of dementia?
@@
 	\begin{center}
-	\begin{tabular}{ p{12cm} }
+	\begin{tabular}{ | p{12cm} | }
 	\hline
@@
 	\begin{center}
-	\begin{tabular}{ p{3cm} | p{3cm} | p{3cm} | p{3cm} }
+	\begin{tabular}{ | p{3cm} | p{3cm} | p{3cm} | p{3cm} | }
 	\hline
@@
 	\begin{center}
-	\begin{tabular}{ p{6cm} | p{6cm} }
+	\begin{tabular}{ | p{6cm} | p{6cm} | }
 	\hline
@@
 \section{Results}\label{results}
-\subsection{Machine Learning methods}
-Intro - Addresses Research Question 1
-
-\begin{table}[H]
-	\begin{center}
-	\begin{tabular}{ p{6cm} | p{6cm} }
-	\hline
-	Machine Learning technique used & Paper Number \\ \hline
-	Logistic Regression & Orimaye \\ \hline
-	Support Vector Machines & Orimaye \\ \hline
-	Linear Discriminant Analysis & Orimaye \\ \hline
-	Decision Trees & Orimaye \\ \hline	
-	\hline 
-	\end{tabular}
-	\end{center}
-\end{table}
-
-\subsection{Traditional Machine Learning methods}
-\subsubsection{Logistic Regression}
-Logistic Regressions is a probabilistic approach to classification and is dependant on the assumption that your input space can be separated into two distinct regions, one for each class, by a linear boundary.
-In 2D space, this boundary is a line and in 3D space, this boundary is defined as a plane.
-\par 
-In the results of this survey.
-
-\subsubsection{Support Vector Machines}
-Support Vector Machine (SVM) was originally proposed as an algorithm for classification problems; it is a relatively new technique compared to the other ML approaches. The classification
-process consists of mapping the data points (usually the study subjects) into a feature space composed of the variables that characterize these data points, except for the outcome variable.
-Then, the algorithm finds patterns in this feature space by defining the maximum separation between two or more classes, depending on the problem to be solved. Contrary
-to some regression techniques, SVMs are not dependent on a pre-determined model for data fitting, although there are still algorithm specifications to be considered (e.g. choice of a kernel
-function); instead, it is a data-driven algorithm that can work relatively well in a scenario where sample sizes are small compared to the number of variables, reason why it has been widely employed by diagnostic studies in tasks related to the automated classification of diseases.
-\par 
-Regarding the SLR results, SVMs were present in ??/41 selected studies, in 38 proposed models, and being by far the most used machine learning technique in the dementia diagnosis research.
-These numbers account for the traditional SVM and variations.
-In all of the ??
-selected studies the SVMs focused at binary classifications where the task was to discriminate mild cognitively impaired (MCI) patients that will or will not develop Alzheimer's Disease (AD).
-In the general case, the problem is posed as either MCI converters versus MCI non-converters, or progressive MCI versus stable MCI classification.
-This outlines a situation in which a regression problem (when will the MCI patients convert to AD?) is formulated as a classification problem (which MCI patients will convert to AD in X months?) to be solved.
-Reasons for this could be due to limitations in the data used, i.e.
-the limited follow-up periods of the subjects included in the studies.
-
-\subsubsection{Linear Discriminant Analysis}
-
-\subsubsection{Decision Trees}
-A Decision Tree (DT) is a classification algorithm in which the learned knowledge is represented in a tree structure that can be translated to if-then rules. DT's learning process is recursive
-and starts by testing each input variable as to how well each of them, alone, can classify the labeled examples. The best one is selected as a root node for the tree and its descendant nodes
-are defined as the possible values (or relevant ratios) of the selected input variable.
-The training set is then classified between the descendant nodes according to the values of the selected input variable.
-This process is repeated recursively until no more splits in the tree are possible.
-Like SVMs, DTs do not depend on a pre-defined model and are mostly used
-to find important interactions between variables. Being intuitive and easy to interpret, DTs have been used in prognostic studies as a tool for determining prognostic subgroups.
-In this SLR, DTs were the second most frequently used ML technique, present in 6/37 selected studies and proposed in 7 models.
-It was employed for the same reason as SVM; except for one study that investigated the evolution of patients diagnosed with cognitive impairment no dementia (CIND) to AD.
-
-Curry, Singer and Habash - Alz (66.7 - Pres, 66.7 - Recall) and Control (67.9 - Pres, 67.9 Recall)
-
-\subsubsection{Naive Bayes}
-
-Curry, Singer and Habash - Alz(80.8 - Pres, 0.75 - Recall) and Control (79.3 - Pres, 82.1 Recall) - In the naive bayes classifier, they identified that pauses, go-ahead, fillers and incomplete words as the most significant features.
-%Why is Recall at 0.75???????
-
-\subsubsection{Artificial Neural Networks}
-An Artificial Neural Network (ANN) is a methodology that performs multifactorial analyses, which is desirable in the health area as medical decision-making problems are usually dependent of many factors.
-An ANN is composed of nodes connected by weighted edges in a multi-layer architecture that comprises: an input layer, one or more hidden layers and an output layer.
-In the training process, inputs and outputs values are known to the network, while the weights are incrementally adjusted so that the outputs of the network are approximate to the known outputs.
-Despite being a powerful predictor, ANNs are `black boxes', which means that they are not able to explain their predictions in an intuitive way, contrary to DTs or BNs.
-Also, they require the specification of the architecture to be used beforehand (i.e.
-the number of hidden layers).
-
-\subsubsection{K-nearest neighbours}
-K Nearest Neighbors (KNN) is a classification algorithm that takes a data point from an unknown class and assigns it as an input vector in the feature space.
-Then, the classification process follows by assigning the unknown class data point to the class in which the majority of the K nearest data points belong to.
-The distance between data points is usually measured by Euclidean distance, but it is possible to employ other measures.
-KNN is one of the simplest ML classification algorithms and have been used in a wide range of applications; however, it can be computationally expensive in a highly dimensional scenario.
-Further, it considers all features to be equally weighted, which can be a problem if the data has superfluous attributes.
-
-\subsection{Deep Learning methods}
-\subsubsection{Deep language space neural network for classifying mild cognitive impairment and Alzheimer-type dementia - Orimaye, Wong and Wong (2018)}
-In this paper, Orimaye et al use deep-deep neural networks language models (D2NNLM) to learn linguistic changes that distinguish the language of patients with MCI and AD-type dementia from the healthy controls using higher order n-grams.
-An ordinary DNNLM uses lower order n-gram N-dimensional sparse vectors as descrete feature representations to train the neural network with multiple hidden layers.
-
 \subsection{Features of Language}
-Intro - Addresses Research Question 2
+One of the first steps in building an accurate diagnostic model using language is to analyse the effectiveness of the various language features that are used.
 
@@
 
-\subsection{What are the goals of the studies that employ ML or MS techniques for prognosis of dementia and comorbidities?}
+\section{Machine Learning methods}
+The next stage of building statistical or Machine Learning models is to take some features and use these feature to build models able to classify participants into various categories.
+There are various types of classification models that have been used to tackle this problem, this section looks which models have been used in the current literature.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ p{6cm} | p{6cm} }
+	\hline
+	Machine Learning technique used & Paper Number \\ \hline
+	Logistic Regression & Orimaye \\ \hline
+	Support Vector Machines & Orimaye \\ \hline
+	Linear Discriminant Analysis & Orimaye \\ \hline
+	Decision Trees & Orimaye \\ \hline	
+	\hline 
+	\end{tabular}
+	\end{center}
+\end{table}
+
+\subsection{Traditional Machine Learning methods}
+\subsubsection{Logistic Regression}
+Logistic Regressions is a probabilistic approach to classification and is dependant on the assumption that your input space can be separated into two distinct regions, one for each class, by a linear boundary.
+In 2D space, this boundary is a line and in 3D space, this boundary is defined as a plane.
+\par 
+In the results of this survey.
+
+\subsubsection{Support Vector Machines}
+Support Vector Machine (SVM) was originally proposed as an algorithm for classification problems; it is a relatively new technique compared to the other ML approaches.
+The classification process consists of mapping the data points (usually the study subjects) into a feature space composed of the variables that characterize these data points, except for the outcome variable.
+Then, the algorithm finds patterns in this feature space by defining the maximum separation between two or more classes, depending on the problem to be solved.
+Contrary to some regression techniques, SVMs are not dependent on a pre-determined model for data fitting, although there are still algorithm specifications to be considered (e.g.
+choice of a kernel function); instead, it is a data-driven algorithm that can work relatively well in a scenario where sample sizes are small compared to the number of variables, reason why it has been widely employed by diagnostic studies in tasks related to the automated classification of diseases.
+\par 
+Regarding the SLR results, SVMs were present in ??/41 selected studies, in 38 proposed models, and being by far the most used machine learning technique in the dementia diagnosis research.
+These numbers account for the traditional SVM and variations.
+In all of the ??
+selected studies the SVMs focused at binary classifications where the task was to discriminate mild cognitively impaired (MCI) patients that will or will not develop Alzheimer's Disease (AD).
+In the general case, the problem is posed as either MCI converters versus MCI non-converters, or progressive MCI versus stable MCI classification.
+This outlines a situation in which a regression problem (when will the MCI patients convert to AD?) is formulated as a classification problem (which MCI patients will convert to AD in X months?) to be solved.
+Reasons for this could be due to limitations in the data used, i.e.
+the limited follow-up periods of the subjects included in the studies.
+
+\subsubsection{Linear Discriminant Analysis}
+
+\subsubsection{Decision Trees}
+A Decision Tree (DT) is a classification algorithm in which the learned knowledge is represented in a tree structure that can be translated to if-then rules.
+DT's learning process is recursive and starts by testing each input variable as to how well each of them, alone, can classify the labeled examples.
+The best one is selected as a root node for the tree and its descendant nodes are defined as the possible values (or relevant ratios) of the selected input variable.
+The training set is then classified between the descendant nodes according to the values of the selected input variable.
+This process is repeated recursively until no more splits in the tree are possible.
+Like SVMs, DTs do not depend on a pre-defined model and are mostly used to find important interactions between variables.
+Being intuitive and easy to interpret, DTs have been used in prognostic studies as a tool for determining prognostic subgroups.
+\par 
+In this SLR, DTs were the second most frequently used ML technique, present in 6/37 selected studies and proposed in 7 models.
+It was employed for the same reason as SVM; except for one study that investigated the evolution of patients diagnosed with cognitive impairment no dementia (CIND) to AD.
+
+Curry, Singer and Habash - Alz (66.7 - Pres, 66.7 - Recall) and Control (67.9 - Pres, 67.9 Recall)
+
+\subsubsection{Naive Bayes}
+
+Curry, Singer and Habash - Alz(80.8 - Pres, 0.75 - Recall) and Control (79.3 - Pres, 82.1 Recall) - In the naive bayes classifier, they identified that pauses, go-ahead, fillers and incomplete words as the most significant features.
+%Why is Recall at 0.75???????
+
+\subsubsection{Artificial Neural Networks}
+An Artificial Neural Network (ANN) is a methodology that performs multifactorial analyses, which is desirable in the health area as medical decision-making problems are usually dependent of many factors.
+An ANN is composed of nodes connected by weighted edges in a multi-layer architecture that comprises: an input layer, one or more hidden layers and an output layer.
+In the training process, inputs and outputs values are known to the network, while the weights are incrementally adjusted so that the outputs of the network are approximate to the known outputs.
+Despite being a powerful predictor, ANNs are `black boxes', which means that they are not able to explain their predictions in an intuitive way, contrary to DTs or BNs.
+Also, they require the specification of the architecture to be used beforehand (i.e.
+the number of hidden layers).
+
+\subsubsection{K-nearest neighbours}
+K Nearest Neighbors (KNN) is a classification algorithm that takes a data point from an unknown class and assigns it as an input vector in the feature space.
+Then, the classification process follows by assigning the unknown class data point to the class in which the majority of the K nearest data points belong to.
+The distance between data points is usually measured by Euclidean distance, but it is possible to employ other measures.
+KNN is one of the simplest ML classification algorithms and have been used in a wide range of applications; however, it can be computationally expensive in a highly dimensional scenario.
+Further, it considers all features to be equally weighted, which can be a problem if the data has superfluous attributes.
+
+\subsection{Deep Learning methods}
+\subsubsection{Deep language space neural network for classifying mild cognitive impairment and Alzheimer-type dementia - Orimaye, Wong and Wong (2018)}
+In this paper, Orimaye et al use deep-deep neural networks language models (D2NNLM) to learn linguistic changes that distinguish the language of patients with MCI and AD-type dementia from the healthy controls using higher order n-grams.
+An ordinary DNNLM uses lower order n-gram N-dimensional sparse vectors as descrete feature representations to train the neural network with multiple hidden layers.
+
+\subsection{What are the goals of the studies that employ ML or Statistical Learning techniques for diagnosis of MCI or AD?}
+
+\subsection{Do the studies focus on a one point in time or looking at cognitive deterioration over time?}
 

### 2019-08-22 Source/Thesis/Chapter1.tex
updated Thesis

@@
-\section{Introduction}
-Alzheimer's Disease (AD) is a neurodegenerative disease in which, from a physiological perspective, the brain develops neurofibrillary tangles and neuritic plaques along with the deterioration and loss of cortical neurons and synapses.
-Whilst a definitive diagnosis of Dementia can only be produced at post-mortem, there are a number of clinical indicators from psychological perspective that can indicate dementia is present.
-From a clinical psychology perspective, those who have dementia demonstrate cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}.
-In addition, these symptoms are often accompanied by emotional problems such as depression and behavioural difficulties.
-\par
-AD and other forms of dementia affect a significant proportion of the geriatric population in the world today and is currently the sixth leading cause of death in the US and was named the leading killer of women in the UK.
+\section{Background}
+Alzheimer's Disease (AD) and other forms of dementia affect a significant proportion of the geriatric population in the world today and is currently the sixth leading cause of death in the US and was named the leading killer of women in the UK.
 According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
@@
 \par 
-Despite this growing problem, at present there are no drugs that improve the prognosis of those suffering with AD.
-All the drugs that are on the market are designed to manage symptoms.
+AD is a neurodegenerative disease in which a definitive diagnosis can only be produced at post-mortem.
+However, there are a number of psychological and physiological indicators that can indicate that dementia is present.
+From a physiological perspective, researchers have identified two proteins called beta-amyloid and tau.
+In a typical case, tau accumulates and eventually forms tangles inside neurons and beta-amyloid clumps into plaques which slowly builds up between neurons.
+At a certain point, the levels of beta-amyloid rise and trigger a more rapid spread of tau throughout the brain.
+Eventually, due to this and other changes, neurons lose their ability to communicate and the brain starts to shrink.
+This leads to some of the more psychological symptoms, those who have dementia demonstrate cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}.
+In addition, these symptoms are often accompanied by emotional problems such as depression and behavioural difficulties.
+As more neurons die throughout the brain, a person with Alzheimer's gradually loses the ability to think, remember, make decisions and function independently.
+\par
+Despite this growing problem and an increasing understanding about how AD affects the brain there are no medications that improve the prognosis of those with AD.
+All the medications that are currently on the market are designed to manage symptoms.
 Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{Cummings2014}.
 Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place \cite{Cummings2014}.
-It is therefore important to focus AD at it's earliest stages which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.
+Research has started to be more focused on AD at it's earliest stages which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.
 \par
 One of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of cognitive impairment or dementia.
+This challenge is often complicated further due to the large variation in the cognitive abilities and educational background of individuals.
 Albert and his team have worked to define clinical criteria which professionals can use to diagnose MCI due to AD and differentiate this from age-associated memory impairment and age-associated cognitive decline.
-They note that the diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}.
+One of the most important observations from this piece of work is that a diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}, and this is essentially to place more importance on the trajectory of a person's cognitive abilities rather than a person's cognitive ability in general.
 The criteria for MCI is detailed below \cite{Albert2011}.
@@
 \par
+Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease.
+Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
 % Do more on the state of the art in terms of the analysis of biomarkers for MCI.
+\section{Statement of the Problem}
 Studies that explore ways in which we can diagnose MCI / AD generally follow one of two main approaches, the analysis of biomarkers such as concentrations of amyloid-β 1-42 (Aβ42) in cerebrospinal fluid (CSF) and analysis of cognitive abilities.
@@
 \par
+\section{Research Questions}
 The overall objective of the research described in this thesis is to investigate the deterioration of language in people diagnosed with MCI and the early stages of AD.
@@
 The potential impact of this research in this area is immense.
-Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease.
-Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
-\par
+\section{Structure of thesis}
 There is a lot research in this domain from a psychological perspective.

### 2019-09-16 Source/Paper 2 - Systematic Review/Systematic Review.tex
updated SR

@@
 \begin{abstract}
-This is the paper's abstract \ldots
+Dementia is a growing problem which faces the world today with the total worldwide cost of dementia last year projected to be in the region of one trillion US dollars.
+Given the huge impact Dementia has, a lot of research has been focused on finding ways to improve the early diagnosis of dementia and Mild Cognitive Impairment.
+Equally, our ability to process language and find patterns of decline using the techniques of Natural Language Processing and Machine Learning has vastly increased over the recent times.
+This represents an opportunity to harness these techniques to find patterns in cognitive decline that could indicate the presence of the early stages of dementia.
+This paper looks at body of research focused on the analysis of cognitive decline in those suspected to have dementia.
+Specifically,   \ldots
 \end{abstract}
@@
 \section{Introduction}
-Dementia has been identified as one of those fast growing difficulties facing the world.
+Dementia has been identified as one of the fast growing difficulties facing the world.
 A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
@@
 A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
-As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages \cite{Prince2015} and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
+As described above \cite{Prince2015}, the numbers of those suffering from AD and MCI are going to increase as the population ages and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
 At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017}.
@@
 \par
-There is a large body of research that looks at the decline in language in those with MCI and AD \cite{Taler2008, Boschi2017}.
-However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD \cite{Taler2008, Boschi2017}.
-Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
+There is a large body of research that looks at language deterioration in those suspected to have MCI or AD \cite{Taler2008, Boschi2017}.
+However there is conflicting evidence in these studies about which declining language factors are associated with MCI and AD \cite{Taler2008, Boschi2017}.
+Research therefore, should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
 Another area of focus for research of this nature is the process of collecting appropriate language samples.
@@
 	Paper Number & Paper Title & Authors \\ \hline
-	1 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
-	2 & Abnormalities of Connected Speech in Semantic Dementia vs Alzheimer’s diseases & Authors \\ \hline
-	3 & Aided Diagnosis of Dementia Type through Computer-Based Analysis of Spontaneous Speech & Authors \\ \hline
-	4 & Alignment of spoken narratives for automated neuropsychological assessment & Authors \\ \hline
-	5 & Automatic detection and rating of dementia of alzheimer type through lexical analysis of spontaneous speech & Authors \\ \hline
-	6 & Changes in Style in Authors with Alzheimer’s Disease & Authors \\ \hline
-	7 & Computer-based evaluation of Alzheimer's disease and mild cognitive impairment patients during a picture description task & Authors \\ \hline
-	8 & Computerized assessment of syntactic complexity in Alzheimer's disease: A case study of Iris Murdoch's writing & Authors \\ \hline
-	9 & Computerized neuropsychological assessment in mild cognitive impairment based on natural language processing-oriented feature extraction & Authors \\ \hline
-	10 & Contrastive conversational analysis of language production by Alzheimer's and control people& Authors \\ \hline
-	11 & Declines in Connected Language Are Associated with Very Early Mild Cognitive Impairment: Results from the Wisconsin Registry for Alzheimer’s Prevention & Authors \\ \hline
-	12 & Deep-Deep Neural Network Language Models for Predicting Mild Cognitive Impairment & Authors \\ \hline
-	13 & Domain Adaptation for Detecting Mild Cognitive Impairment & Authors \\ \hline
-	14 & Diagnosing people with dementia using automatic Conversation Analysis & Authors \\ \hline
-	15 & Deep language space neural network for classifying mild cognitive impairment and Alzheimer-type dementia & Authors \\ \hline
-	16 & Enriching Complex Networks withWord Embeddings for Detecting Mild Cognitive Impairment from Speech Transcripts & Authors \\ \hline
-	17 & Features and Machine Learning Classification of Connected Speech Samples from Patients with Autopsy Proven Alzheimer’s Disease with and without Additional Vascular Pathology & Authors \\ \hline
-	18 & Formulaic Language in People with Probable Alzheimer’s Disease: A Frequency-Based Approach & Authors \\ \hline
-	19 & Language Analysis of Speakers with Dementia of the Alzheimer’s Type & Authors \\ \hline
-	20 & Language analytics for assessing brain health: Cognitive impairment, depression and pre-symptomatic Alzheimer's disease & Authors \\ \hline
-	21 & Learning Linguistic Biomarkers for Predicting Mild Cognitive Impairment using Compound Skip-grams & Authors \\ \hline
-	22 & Learning Predictive Linguistic Features for Alzheimer’s Disease and related Dementias using Verbal Utterances & Authors \\ \hline
-	23 & Linguistic Features Identify Alzheimer's Disease in Narrative Speech & Authors \\ \hline
-	24 & Longitudinal detection of dementia through lexical and syntactic changes in writing: a case study of three British novelists & Authors \\ \hline
-	25 & Microlinguistic aspects of the oral narrative in patients with Alzheimer’s disease & Authors \\ \hline
-	26 & Multilingual word embeddings for the assessment of narrative speech in mild cognitive impairment & Authors \\ \hline
-	27 & Natural Language Features for Detection of Alzheimer's Disease in Conversational Speech & Authors \\ \hline
-	28 & NLP-Oriented Contrastive Study of Linguistic Productions of Alzheimer's and Control People & Authors \\ \hline
-	29 & Predicting mild cognitive impairment from spontaneous spoken utterances & Authors \\ \hline
-	30 & Predicting probable Alzheimer's disease using linguistic deficits and biomarkers & Authors \\ \hline
-	31 & Propositional Idea Density in women’s written language over the lifespan: Computerized analysis & Authors \\ \hline
-	32 & Sentence segmentation in narrative transcripts from neuropsychological tests using recurrent convolutional neural networks & Authors \\ \hline
-	33 & Speech processing approach for diagnosing dementia in an early stage & Authors \\ \hline
-	34 & Spoken language biomarkers for detecting cognitive impairment & Authors \\ \hline
-	35 & Talk2Me: Automated linguistic data collection for personal assessment & Authors \\ \hline
-	36 & The Effect of Heterogeneous Data for Alzheimer’s Disease Detection from Speech & Authors \\ \hline
-	37 & Toward the Automation of Diagnostic Conversation Analysis in Patients with Memory Complaintss & Authors \\ \hline
-	39 & Tracking Discourse Complexity Preceding Alzheimer's Disease Diagnosis: A Case Study Comparing the Press Conferences of Presidents Ronald Reagan and George Herbert Walker Bush & Authors \\ \hline
-	40 & Vector-space topic models for detecting Alzheimer's disease & Authors \\ \hline
-	41 & Vocabulary Size in Speech May Be an Early Indicator of Cognitive Impairment. & Authors \\ \hline
+	1 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & C. Guinn, B. Singer, A. Habash \\ \hline
+	2 & Aided Diagnosis of Dementia Type through Computer-Based Analysis of Spontaneous Speech & W. Jarrold, B. Peintner, D. Wilkins, D. Vergryi, C. Richey, M.L. Gorno-Tempini, J. Ogar \\ \hline
+	3 & Automatic detection and rating of dementia of alzheimer type through lexical analysis of spontaneous speech & C. Thomas, V. Keselj, N. Cercone, K. Rockwood, E. Asp \\ \hline
+	4 & Changes in Style in Authors with Alzheimer’s Disease & G. Hurst, V.W. Feng \\ \hline
+	5 & Computer-based evaluation of Alzheimer's disease and mild cognitive impairment patients during a picture description task & L.
+Hernandez-Dominguez, S.
+Ratte, G.
+Sierra-Martinez, A.
+Roche-Bergua \\ \hline
+	6 & Computerized neuropsychological assessment in mild cognitive impairment based on natural language processing-oriented feature extraction & C.
+Fang, P.
+Janwattanapong, H.
+Martin, M.
+Cabrerizo, A.
+Barreto, D.
+Loewenstein, R.
+Duara, M.
+Adjouadi \\ \hline
+	7 & Deep-Deep Neural Network Language Models for Predicting Mild Cognitive Impairment & S.O. Orimaye, J. S-M. Wong, J. S. G. Fernandez \\ \hline
+	8 & Deep language space neural network for classifying mild cognitive impairment and Alzheimer-type dementia & S.O. Orimaye, J. S-M. Wong, C.P. Wong \\ \hline 
+	9 & Diagnosing people with dementia using automatic Conversation Analysis & B. Mirheidari, D. Blackburn, M. Reuber, T. Walker, H. Christensen \\ \hline
+	10 & Enriching Complex Networks withWord Embeddings for Detecting Mild Cognitive Impairment from Speech Transcripts & L.
+B.
+dos Santos, E.
+A.
+Correa Jr, O.
+N.
+Oliveira Jr, D.R.
+Amancio, L.
+L.
+Mansur, S.
+M.
+Alusio \\ \hline
+	11 & Features and Machine Learning Classification of Connected Speech Samples from Patients with Autopsy Proven Alzheimer’s Disease with and without Additional Vascular Pathology & V.
+Rentoumi, L.
+Raoufian, S.
+Ahmed, C.
+A.
+de Jager, P.
+Garrard \\ \hline
+	12 & Language Analysis of Speakers with Dementia of the Alzheimer’s Type & C. Guinn, A. Habash \\ \hline
+	13 & Learning Linguistic Biomarkers for Predicting Mild Cognitive Impairment using Compound Skip-grams & S. O. Orimaye, J. S-M. Wong, K. Y. Tai, C. P. Wong \\ \hline
+	14 & Learning Predictive Linguistic Features for Alzheimer’s Disease and related Dementias using Verbal Utterances & S.O. Orimaye, J. S-M. Wong, K. J. Golden \\ \hline
+	15 & Linguistic Features Identify Alzheimer's Disease in Narrative Speech & K. C. Fraser, J. A. Meltzer, F. Rudzicz \\ \hline
+	16 & Longitudinal detection of dementia through lexical and syntactic changes in writing: a case study of three British novelists & X. Le, I. Lancashire, G. Hirst, R. Jokel \\ \hline
+	17 & Multilingual word embeddings for the assessment of narrative speech in mild cognitive impairment &  K. C. Fraser, K. L. Fors, D. Kokkinakis \\ \hline
+	18 & Predicting mild cognitive impairment from spontaneous spoken utterances & M. Asgari, J. Kaye, H. Dodge \\ \hline
+	19 & Predicting probable Alzheimer's disease using linguistic deficits and biomarkers & S. O. Orimaye, J. S-M. Wong, K. J. Golden, C. P. Wong, I. N. Soyiri \\ \hline
+	20 & Sentence segmentation in narrative transcripts from neuropsychological tests using recurrent convolutional neural networks & M. V. Treviso, C. Shulby, S. M. Alusio \\ \hline
+	21 & Social Markers of Mild Cognitive Impairment: Proportion of Word Counts in Free Conversational Speech & H.
+H.
+Dodge, N.
+Mattek, M.
+Gregor, M.
+Bowman, A.
+Seelye, O.
+Ybarra, M.
+Asgari, J.
+A.
+Kaye \\ \hline
+	22 & Spoken language biomarkers for detecting cognitive impairment & T. Alhanai, R. Au, J. Glass \\ \hline
+	23 & Spoken Language Derivied Measures for Detecting Mild Cognitive Impairment & B. Roark, M. Mitchell, J-P. Hosom, K. Hollingshead, J. A. Kaye \\ \hline
+	24 & Talk2Me: Automated linguistic data collection for personal assessment & M. Komeili, C. Pou-Prom, D. Liaqat, K. C. Fraser, M. Yancheva, F. Rudzicz \\ \hline
+	25 & The Effect of Heterogeneous Data for Alzheimer’s Disease Detection from Speech & A. Balagopalan, J. Novikova, F. Rudzicz, M. Ghassemi \\ \hline
+	26 & Toward the Automation of Diagnostic Conversation Analysis in Patients with Memory Complaintss & B.
+Mirheidari, D.
+Blackburn, K.
+Harkness, T.
+Walker, A.
+Venneri, M.
+Reuber, H.
+Cristensen \\ \hline
+	27 & Tracking Discourse Complexity Preceding Alzheimer's Disease Diagnosis: A Case Study Comparing the Press Conferences of Presidents Ronald Reagan and George Herbert Walker Bush & V.
+Berisha, S.
+Wang, A.
+LaCross, J.
+Liss \\ \hline
+	28 & Vector-space topic models for detecting Alzheimer's disease & M. Yancheva, F. Rudzicz \\ \hline
 \end{longtable}

### 2019-09-19 Source/Paper 2 - Systematic Review/Systematic Review.tex
updated SR

@@
 	Machine Learning technique used & Paper Number \\ \hline
-	Logistic Regression & Orimaye \\ \hline
-	Support Vector Machines & Orimaye \\ \hline
-	Linear Discriminant Analysis & Orimaye \\ \hline
-	Decision Trees & Orimaye \\ \hline	
+	Logistic Regression & 2 \\ \hline
+	Support Vector Machines & 4 \\ \hline
+	Naive Bayes & 1 \\ \hline
+	Decision Trees & 1, 2 \\ \hline
+	Multi-layered perceptrons & 2 \\ \hline
 	\hline 
@@
 
+\subsubsection{Best Results}
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ p{6cm} | p{6cm} }
+	\hline
+	Machine Learning technique used & Paper Number \\ \hline
+	Logistic Regression & 2 \\ \hline
+	Support Vector Machines & 4, 5 \\ \hline
+	Naive Bayes & 1 \\ \hline
+	Decision Trees and Random Forest Classifiers & 1, 2, 5 \\ \hline
+	Multi-layered perceptrons & 2 \\ \hline
+	\hline 
+	\end{tabular}
+	\end{center}
+\end{table}
+
 \subsection{Deep Learning methods}
@@
 An ordinary DNNLM uses lower order n-gram N-dimensional sparse vectors as descrete feature representations to train the neural network with multiple hidden layers.
+
+\subsection{What type of data is used by the studies?}
+One of the key debates when looking at how to analyse language is the type of task provided to elicit language production in participants.
+In the literature researchers have primarily focused on Picture Description tasks but have also suggested other ways in which we might collect data.
+\par
+\subsection{Picture Description Tasks}
+One of the most commonly used tasks to measure language is the Picture Description task.
+An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Borod1980}.
+The Cookie Theft picture (pictured below) depicts a scene of a home typical of the period of time when it was created and would generally not require participants to use any complicated vocabulary to describe.
+In this task participants are asked to describe the picture presented to them in as much detail as possible.
+This task was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}
+\begin{figure}[H]
+\centering
+\includegraphics[width=240px, height=150px]{images/BCTPicture.png}
+\caption{Cookie Theft Picture - From Kaplan and Goodglass (1983)}
+\end{figure}
+\par
+The picture description task does a fine job of eliciting descriptive language however because of the specific content the language produce could be considered quite limited.
+There is some disagreement as to the benefits of this using this methodology.
+This task is reported as being useful to lexico-semantic disorders \cite{Boschi2017, Sajjadi2012} as the language being generated is primarly nouns and deixis (words to identify items and words to put those items into context).
+However, Ash \cite{Ash2012}felt that there was no difference in using this task vs Story Narration (described below).
+In explaining the differences, it is worth noting that these researchers were using differing variables and this could explain their different perspectives.
+\par 
+There are a number of existing corpus which use this tasks as the foundation of their data.
+The most well known is the Pitt Corpus of the Dementia Bank database.
+The participants of the corpus are mainly healthy controls, those with MCI and those with probable and/or possible AD.
+\subsection{Narrative description task}
+The story narration task is designed to study a participant's ability to describe and elaborate on a story which is depicted using a series of pictures.
+The stories depicted are usually based on children's books or famous stories with the Cinderella being the one most typically used \cite{Fraser2014}.
+This task requires ordering the story in a structured and coherent framework.
+It also requires comprehension and understanding of the stories characters and the events depicted, as well as an awareness of a character's actions, motivations and internal reactions to given events.
+This task is particularly useful as the procedure reduces the demands on memory, due to the participant being able to access the picture book during the description and is therefore able rule out memory as a confounding variable for any results observed.
+As noted above, Ash \cite{Ash2012} felt that this task was interchangable with the Picture Description task.
+However, other research felt that this was a studier test of lexical and semantic abilities as well as syntactic complexity because this task requires interpretation and elaboration in additional to a simple description \cite{DeLira2011}.
+\par
+Given the relative strengths of the Narrative description task vs Picture description task, there are few pieces of research that have used Machine Learning to analyse features from Narrative picture tasks \cite{Fraser2014}.
+This could be due to the availability of data and the absence of any meaningful sets of transcripts of participants performing this task.
+However, this could be an interesting direction to take research in the future to see if features generated from this task could be used to predict MCI or AD.
+\par
+\subsection{Interviews}
+Interviews can also be used to elicit language in a more natural way by asking questions to guide a conversation between speakers.
+There are three types of interviews: unstructured, structured and semi-structured.
+Structured interviews tend to produce very limited speech and therefore has never been used in this area \cite{Boschi2017}.
+Unstructured interviews are open ended and generally do not conform to any particular pattern.
+They use generic themes such as family or hobbies to guide the conversation.
+Whilst this is the most ecologically valid form of conversation and therefore language generation, it's unstructured nature means that the protocol cannot be consistent and therefore reproduced.
+Semi-structured interviews are therefore preferred over other forms of interview as a middle ground.
+The semi structured nature of these interviews means that there is some replicability but does not constrain the participant in answering questions.
+\par
+The analysis of interviews can be difficult to analyse as both the content can vary even between participants, although it can be argued that content should not affect the type of language being generated unless it is narrow topic or the participant is constrained in how they answer a given question.
+It is also difficult to measure as there are no pre-defined task goals in comparison to the other two methods.
+Nevertheless, this is the most naturalistic setting for looking at language production and can be used to look at the syntactic and semantic parts of language generation \cite{Sajjadi2012}.
+There have been some attempts to use interviews to assess language production in AD with promising results \cite{Asgari2017, Guinn2015}.
+\par
+\subsection{Conclusions}
+One can view the different types of tasks above as a continuum where picture tasks represent a much more controllable task with a lot of supporting research but which generates a much more constrained set of language that is atypical of normal speech in terms of the cognitive functions used.
 
@@
 	25 & The Effect of Heterogeneous Data for Alzheimer’s Disease Detection from Speech & A. Balagopalan, J. Novikova, F. Rudzicz, M. Ghassemi \\ \hline
-	26 & Toward the Automation of Diagnostic Conversation Analysis in Patients with Memory Complaintss & B.
-Mirheidari, D.
-Blackburn, K.
-Harkness, T.
-Walker, A.
-Venneri, M.
-Reuber, H.
-Cristensen \\ \hline
+	26 & Toward the Automation of Diagnostic Conversation Analysis in Patients with Memory Complaints & B. Mirheidari, D. Blackburn, K. Harkness, T. Walker, A. Venneri, M. Reuber, H. Cristensen \\ \hline
 	27 & Tracking Discourse Complexity Preceding Alzheimer's Disease Diagnosis: A Case Study Comparing the Press Conferences of Presidents Ronald Reagan and George Herbert Walker Bush & V.

### 2019-10-11 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
edited PresidentCorpus

@@
 	\texttt{g.vogiatzis@aston.ac.uk} \\
-		\and
-	Cristina Romani\\
-	Department of Psychology\\
-	School of Life and Health Sciences\\
-	Aston University \\
-	\texttt{c.romani@aston.ac.uk}
 }
@@
 \usepackage{arxiv}
+\usepackage{float}
 
@@
 
-\keywords{MCI \and Alzheimer's Disease \and Machine Learning \and Diagnosis} 
+\keywords{Mild Cognitive Impairment \and Alzheimer's Disease \and Machine Learning \and Diagnosis \and Natural Language Processing} 
 \bigskip
@@
 \section{Introduction}\label{introduction}
+Alzheimer's Disease (AD) and other forms of dementia affect a significant proportion of the older age population in the world today and is currently the sixth leading cause of death in the US and was named the leading killer of women in the UK.
+According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
+This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population \cite{AlzheimersSociety2014}.
+From a financial perspective, they estimate an annual spend of £4.3 billion of which approximately £85 million is spent solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
+Globally, this picture is a lot bleaker.
+Another report by Alzheimer's Disease International suggests that in 2015 there were 46 million people with a diagnosis of dementia globally and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
+\par 
+AD is a neurodegenerative disease in which a definitive diagnosis can only be produced at post-mortem.
+However, there are a number of psychological and physiological indicators that can indicate that dementia is present.
+From a physiological perspective, researchers have identified two proteins called beta-amyloid and tau.
+In a typical case, tau accumulates and eventually forms tangles inside neurons and beta-amyloid clumps into plaques which slowly builds up between neurons.
+At a certain point, the levels of beta-amyloid rise and trigger a more rapid spread of tau throughout the brain.
+Eventually, due to this and other changes, neurons lose their ability to communicate and the brain starts to shrink.
+This leads to some of the more psychological symptoms, those who have dementia demonstrate cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}.
+In addition, these symptoms are often accompanied by emotional problems such as depression and behavioural difficulties.
+As more neurons die throughout the brain, a person with Alzheimer's gradually loses the ability to think, remember, make decisions and function independently.
+\par
+Despite this growing problem and an increasing understanding about how AD affects the brain there are no medications that improve the prognosis of those with AD.
+All the medications that are currently on the market are designed to manage symptoms.
+Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{Cummings2014}.
+Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place \cite{Cummings2014}.
+Research has started to be more focused on AD at it's earliest stages which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.
+\par
+One of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of cognitive impairment or dementia.
+This challenge is often complicated further due to the large variation in the cognitive abilities and educational background of individuals.
+Albert and his team have worked to define clinical criteria which professionals can use to diagnose MCI due to AD and differentiate this from age-associated memory impairment and age-associated cognitive decline.
+One of the most important observations from this piece of work is that a diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}, and this is essentially to place more importance on the trajectory of a person's cognitive abilities rather than a person's cognitive ability in general.
+\par 
+There has been a significant research in the area of language deterioration as a means of detecting Alzheimer's Disease.
+This usually takes the form analysis of speech recorded as part of a cognitive assessment such as the Picture Description Task \cite{Orimaye2014,Fraser2015}.
+Given that language samples are relatively easy to collect, research has moved towards analysis of spontaneous speech.
+An good example of this type of research is the study conducted by Berisha and Liss which looked at the differences in language use between two US presidents, Ronald Reagan (RR), who would go on to receive a diagnosis of Dementia and George H.
+W.
+Bush (GHWB) who acted as a matched control based on Age \cite{Berisha2015}.
+They found several differences in language use over time which they felt acted as indicators of RR's difficulties with language due to dementia.
+These significant differences were in the number of unique words used per speech, the use of non-specific nouns and fillers and low-imageability verbs \cite{Berisha2015}.
+\par   
+This study replicates work done by Berisha and Liss and extends this by adding Donald Trump (DJT) as an alternative, more appropriate comparison to RR as he is much closer in age than GHWB.
+This work contributes to our understanding of the pre-diagnosis phase of alzheimer's disease and it's effect on language.
+We explore a wider range of features than previously looked at with this dataset and our hypothesis is that there will be features of RR's language which change more noticably over time when compared with the two closest appropriate controls.
 
 \section{Methodology}\label{methodology}
+We took 46 transcripts of Ronald Reagan’s (RR) press conferences from 1981 to 1988 and compared them with 134 press conferences by George H.
+W.
+Bush (GHWB) and 29 press conferences conducted by Donald J.
+Trump (DJT).
+We analyzed transcripts for lexical features shown to change longitudinally with dementia.
+These features encompassed, word level, sentence level and document level features and included a number of features contained in the study by Berisha and Liss with the aim of replicating and extending on their findings.
+We compared the trends described in the transcripts of RR and GHWB, but also included DJT.
+Berisha and Liss originally made the comparison as it GHWB (GHWB - age at the start of presidency - 64 years, 222 days) was the closest match to RR in terms of age (RR  - age at the start of presidency, 69 yeaars and 349 days).
+However, with the inauguration of Trump, he now is the closest comparible president in terms of age (DJT - age at the start of presidency - 70 years, 220 days).
+It would be interesting to look at a comparison of RR and DJT to see whether the comparisons made by Berisha and Liss hold true with this more appropriate match (in terms of age).
+DJT as with GHWB has no known diagnosis of AD.
+I used the press conference transcripts in the American Presidency Project (APP) archive as a data source for this project.
+The APP is a comprehensive and organized searchable database of presidential documents, including transcripts of speeches, transcripts of news conferences, and other public documents.
+
+\subsection{Pre-processing}
+To generate the files necessary for analysis, I downloaded each transcript and performed the following changes.
+I omitted the prepared statement by the president and any speech by other individuals.
+I started each transcript at the beginning of the first answer to a question.
+I filtered any annotations that were added to the transcript, including any references or clarifications, and any laughter.
+It's worth noting that there appears to be a difference in how 'hesitations' were marked down between each president, for RR hesitations were marked by a single hyphen whereas for GHWB hesitations are marked by a double hyphen.
+In order to maintain consistency when parsing through the documents, We have changed both types of hesitation to be marked by a single hyphen.
+We also omitted one word sentences as this data would, from a theoretical perspective, not be relevant for language analysis.
+We did not control for the length of the document, but generated features which would normalise by the length of the document.
+We therefore was able to include all press conferences by both Ronald Reagan and George H.W.
+Bush where there was a question and answer session conducted at least in part by the sitting president (2 press conferences of GHWB were omitted due to a lack of a question and answer session).
+
+\subsection{Feature Selection}
+We calculated the following features for each transcript in turn using the NLTK (see section 2 for a description) \cite{Bird2009}, Linguistic Inquiry and Word Count (LIWC) and Python. 
+\par 
+\subsubsection{Measures of lexical variation}
+Lexical variation is a measure of the variety of language used within a given document.
+A document is said to have high lexical variation if the number of unique words is large.
+We constructed four features of lexical variation.
+Firstly we looked at the number of unique words.
+To do this we were able to split each transcript into individual words and changed them to lowercase using NLTK and were then count the number of unique words that appeared in each transcript.
+We also used the TTR formula, Brunet's Index and Honore's Statistic as other measures of lexical diversity \cite{Richards1987}.
+
+Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
+\begin{equation} \label{x1}
+TTR = numberOfUniqueWords / totalNumberOfWords.
+\end{equation}
+\par 
+Brunet's Index (W) differentiates itself for TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
+\begin{equation} \label{x2}
+W = N^{V(-0.165)}
+\end{equation}
+where N is the total length of the utterance being measured and V is equal to the total vocabulary being used by the subject.
+Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers.
+\newline
+\par 
+Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation: 
+\begin{equation} \label{x3}
+%% Check to see if this is accurate.
+R = (100 \log N) / (1 - V1/V)
+\end{equation}
+where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
+\subsubsection{Fillers, Non-Specific Nouns and LI Verbs}
+For these features, we counted the number of occurrences for different categories (see table for list of categories tracked and the words counted).
+The features were used by Berisha and Liss in their research \cite{Berisha2015} and were taken from work done by Bird et al \cite{Bird2000}.
+These findings were.
+number of unique words, non-specific nouns and fillers and low imageability (LI) verbs.
+Imageability is characterized, according to Berisha and Liss, as the ease with which a term gives rise to a sensory .mental image.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{6cm} |}
+		\hline
+		Category & Words \\ \hline
+		Fillers & \textit{"well", "so", "basically", "actually", "literally", "um", "ah"} \\ \hline
+		Non Specific Nouns & \textit{"something", "anything", "thing", "everything"} \\ \hline
+		LI Verbs & \textit{"be", "come", "do", "get", "give", go", "know", "look", "make", "see", "tell", "think", "want"} \\ \hline
+
+	\end{tabular}
+	\caption{\label{tab:table-name}Categories and Words Counted}
+	\end{center} 
+\end{table}
+
+\subsubsection{Usage of parts of speech}
+This section involves using a Part of Speech tagger (PoS) which analyses a sentence and assigns a 'tag' to each word based on the function the word has in a sentence.
+At a basic level this can be divided into the eight defined parts of speech: 'nouns', 'pronouns', verbs', 'adjectives', 'adverbs', 'conjunctions', 'prepositions' and 'interjections' but can be further subcategorised.
+We used the PoS tagger built into NLTK to tag each transcript in turn and used these the counts from each of these eight categories in our analysis.
+In addition to frequency counts we also normalised these features by dividing the frequency count by the number of words in the document to take into account transcript length.
+
+\subsubsection{Linguistic Inquiry and Word Count(LIWC)}
+The LIWC
 
 \section{Results}\label{results}
+One of the most important thing to note is the wide variety of samples between the three presidents and also the varying timescales.
+RR participated in 46 press conferences over eight years (an average of 5.75 a year) which is the fewest number of press conferences given by an American president during their term of office.
+GHWB participated in 136 press conferences over four years (an average of  34 a year) and DJT participated in 29 press conferences to date (an average of 19.3 per year).
+Equally, there are variances in the average number of words.
+RR produced an average of 3424 words per conference compared to 2608 by GHWB (unpaired t = 4.434, p\textless 0.001) and DJT at 1849 words (unpaired t = 6.524, p\textless 0.001).
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR & GHWB & DJT \\ \hline
+		Total Words & 3423.91 (416.42) & 2607.72 (1210.38) & 1848.65 (1549.38) \\ \hline
+		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
+		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
+		LI Verbs & 124.22 (17.89)& 103.45 (51.87) & 84.48 (75.78) \\ \hline
+	\end{tabular}
+	\caption{\label{tab:table-name}Means and Standard Deviations of important features}
+	\end{center} 
+\end{table}
+
+In terms of unique words, we found that RR used significantly more unique words, non-specific nouns and low imageability verbs than GHWB and DJT (see Table 3.3).
+Some of these differences are due to the length of the sample, particularly in the case of DJT where his average sample is almost half the sample of RR.
+It could also be said that this could be down to differences in linguistic abilities or speaking style \cite{Berisha2015, Le2011}.
+However, we can certainly see that as controls GWHB and DJT are comparative in relation to non-specific nouns and LI verbs.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR v GHWB & RR v DJT & GHWB v DJT \\ \hline
+		Total Words & \textbf{4.434***} & \textbf{6.524***} & \textbf{2.899**} \\ \hline
+		Unique Words & \textbf{6.832***} & \textbf{11.403***} & \textbf{4.137***} \\ \hline
+		Non Specific Nouns & \textbf{7.877***} & \textbf{3.426**} & -0.574 \\ \hline
+		LI Verbs & \textbf{2.656**} & \textbf{3.420***} & 1.628 \\ \hline
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+	\end{tabular}
+	\caption{\label{tab:table-name}RR T-tests vs GWB and DJT}
+	\end{center} 
+\end{table}
+
+We then looked at the data from a longitudinal perspective as we are interested seeing whether we can track various language variables and their progress over time.
+We ran a number of Pearsons correlations with transcript index number as a time reference and the dependant variables (Table 3.4).
+For our controls, we found them to be stable for the most part with the main highlights being a decrease in Adverb usage for DJT (R=-0.36, p=0.049) and a steady but not severe decline in a number of variables for GWHB, namely total word count, unique words, low imageability words and verb usage.
+\par 
+For RR, his decline is more marked and more widespread through his language use.
+We noticed an significant increase in adverb (R=0.41, p=0.004) and pronoun usage (R=0.65, p\textless0.001), as well as a slight usage increase in Non-specific nouns(R=0.30, p=0.03).
+There was a highly significant decrease in number of unique words (R=-0.56, p\textless0.001) and noun usage (R=-0.70, p\textless0.001).
+Also very significant decrease in adjective usage (R=-0.40, p=0.005) and a significant decrease in total word count (R=-0.31, p=0.03).
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | l | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR & GHWB & DJT \\ \hline
+		Word Count & \textbf{-0.31*} & \textbf{-0.21*} & 0.08 \\ \hline 
+		Unique Words & \textbf{-0.56***} & \textbf{-0.25**} & 0.16 \\ \hline
+		Non Specific Nouns & \textbf{0.30*} & -0.08 & -0.03 \\ \hline
+		LI Verbs & -0.19 & \textbf{-0.20**} & 0.02 \\ \hline
+		Nouns Normalised & \textbf{-0.70***} & -0.03 & 0.14 \\ \hline
+		Verbs Normalised & \textbf{0.36**} & \textbf{0.24***} & -0.03 \\ \hline
+		Adjectives Normalised & \textbf{-0.40**} & 0.08 & -0.34 \\ \hline
+		Adverbs Normalised & \textbf{0.41***} & 0.02 & \textbf{-0.36*} \\ \hline
+		Pronouns Normalised & \textbf{0.65***} & 0.13 & 0.07 \\ \hline
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+	\end{tabular}
+	\caption{\label{tab:table-name}Pearson Correlations for Features}
+	\end{center} 
+\end{table}
+
+Given the number of hypotheses being tested, it is necessary to apply some correction for family wise error rate and therefore a Hochberg step-up procedure was applied to the results of the analysis.
+Because features have been generated from multiple sources, there are some cases where there is significant overlap between the features.
+Where possible the most appropriate feature has been selected, for example in the case of the raw count of pronouns and pronouns normalised by the length of the document, the latter has been selected.
+Where there is doubt as to the most appropriate feature, all features have been included.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | l | p{2.5cm} | p{2.5cm} | p{2.5cm} |}
+		\hline
+		& RR R-Squared  & GWB R-Squared & DJT R-Squared \\ \hline
+		ppron & \textbf{0.700***} & 0.2 & 0.35 \\ \hline
+		social & \textbf{0.698***} & 0.2 & 0.35 \\ \hline
+		NounsNormalised & \textbf{-0.689***} & 0.2 & 0.35 \\ \hline
+		function & \textbf{0.670***} & 0.2 & 0.35 \\ \hline
+		conj &\textbf{0.644***} & 0.2 & 0.35 \\ \hline
+		PronounsNormalised & \textbf{0.631***} & 0.2 & 0.35 \\ \hline
+		Analytic & \textbf{-0.626***} & 0.2 & 0.35 \\ \hline
+		NN & \textbf{-0.601***} & 0.2 & 0.35 \\ \hline
+		male & \textbf{0.585***} & 0.2 & 0.35 \\ \hline
+		UniqueWords & \textbf{-0.578***} & 0.2 & 0.35 \\ \hline
+		WDT & \textbf{-0.577***} & 0.2 & 0.35 \\ \hline
+		shehe & \textbf{0.529*} & 0.2 & 0.35 \\ \hline
+		VBZ & \textbf{-0.521*} & 0.2 & 0.35 \\ \hline
+		JJ & \textbf{-0.518*} & 0.2 & 0.35 \\ \hline
+		Fillers & \textbf{-0.518*} & \textbf{-0.518*}  & 0.35 \\ \hline	
+		EX & \textbf{-0.518*} & \textbf{-0.518*}  & 0.35 \\ \hline	
+		achieve & \textbf{-0.518*} & \textbf{-0.518*}  & 0.35 \\ \hline		
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+
+	\end{tabular}
+	\caption{\label{tab:table-name}Pearson Correlations for Features}
+	\end{center} 
+\end{table}
 
 \section{Discussion}\label{discussion}
+\section{Discussion}
+President Reagan received his diagnosis of AD in August 1994 but using transcripts of speeches he made in his two terms as President (January 1981 - January 1989) we have be able to identify certain changes in his use of language that we might ascribe to the onset of MCI and early AD.
+Despite differences in our methodology, our research supports the findings of Berisha and Liss in that we both find a significant decrease in unique words over time and an increase in non-specifc noun usage.
+Compared to our controls (GWHB andDJT), we find some slight trends with GWHB but no such trends with DJT in his speech albeit his samples of speech span a shorter amount of time.
+
+A criticism of Berisha and Liss's work is the problems they had with normalising the transcripts in terms of length.
+This was also a problem in the work of Garrard et al \cite{Garrard2005, Le2011}.
+Whilst it is important to control for outliers, there are other ways in which we can control for length of sample.
+\par 
+Interestingly, when we normalised the various types of words used by the presidents we found some interesting patterns that further differentiated RR from the controls.
+Whilst Non-specific nouns increased over time, we found that noun usage in general significantly decreased and pronouns increased similarly significantly.
+The increase in pronoun for those with early AD has been identified in literature, although there are only a few studies that explore this \cite{Almor1999, Wendelstein2015}.
+Wendlestein et al propose that the increased used of pronouns is an expression of an impaired ability to adapt language to the listener's needs \cite{Wendelstein2015}.
+Almor et al attributed this reliance on pronouns due to a impaired working memory \cite{Almor1999}.
+\par 
+The decrease in overall noun usage has also been identified as a feature.
+Jarrold et al found that AD patients would use more pronouns, verbs and fewer nouns than controls \cite{Jarrold2014}.
+Wendlestein in their investigations into noun usage found that decreased later on in AD progression and was unaffected in the pre-clinical stages of AD \cite{Wendelstein2014}.
+Our results are supported by existing literature and this potentially means that language analysis in the way we have structured it may have diagnostic or prognostic properties.
+
+There are limitations of this research.
+Whilst in terms of age, DJT is certainly more suitable as a control to match with RR, in some ways they held very different styles of press conferences in that RR preferred to do solo press conferences and DJT has shown a preference for doing joint press conferences which have an impact on the amount of language produced.
+This artifact of the data is in itself notable as it illustrates the problems we may have with smaller amounts of speech.
+Also, the problem of finding an appropriate control is a common one in this domain.
+Given that those with MCI and early dementia have such variable presentations, it might prove of limited value in matched pairs design.
+\par 
+With further work, it is not feasible to the vast array of samples over a timeframe, as we have had with the president corpus and so it would be worth exploring how the quality of these predictions might lessen when faced with considerably fewer samples and over a smaller time period.
+It would also be worth extending this research further to encompass more of the linguistic features Fraser used in her work \cite{Fraser2015} to see if there are any further insights to be gained.
+In addition, this replication and extension has demonstrated the potential utility of using longitudinal data as a means of comparing language use of a person at two or more time periods and using this information as a diagnostic aid for MCI and therefore more work would be helpful from a longitudinal perspective to see if this approach may be valid in moving towards a solution for this particular problem.
+\par
 
 \section{Conclusions}\label{conclusions}
+The results of this work show that we can track a person's use of language through time in a number of ways, and it is possible for an individual to be his or her own control.
+This is important as it means the heterogenous nature of the MCI population does not impact results as much as if we were comparing those with MCI to controls.
+Equally, it would be helpful to have controls to ascertain what would be usual to expect in the decline of language in a healthy older adult.
 

### 2019-10-11 Source/Paper 2 - Systematic Review/Systematic Review.tex
edited PresidentCorpus

@@
 \subsection{The future of the field}
+One of the main patterns
+
+One-class classification is a traditional machine learning model in which the object is to recognize instances of a concept by only using examples of the same concept.
+In the training of this model, instances of only a single object class are available during training.
+In this context,all other classes (any instances that deviate from the single object class) are referred to as 'alien' classes.
+During the testing of the model, it may encounter alien classes and the job of the model is to distinguish between objects of the known class from objects belonging to any 'alien' classes.
+This approach to classification has broad applications and is mainly used when there is a high cost of obtaining samples of alien classes.
+\par 
+In this context, we can see from
+
 One of the areas for research to study is a careful examination of the features that are being used to measure language deterioration.

### 2019-10-16 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
updated presidents corpus

@@
 		\and
-	Peter Sawyer \\
+	\textbf{Peter Sawyer} \\
 	Department of Computer Science\\
@@
 	Aston University\\
-	\texttt{p.sawyer@aston.ac.uk}
-		\and
+	\texttt{p.sawyer@aston.ac.uk} \\
+		\AND
 	George Vogiatzis\\
@@
 	\texttt{g.vogiatzis@aston.ac.uk} \\
+		\and
+	\textbf{Felipe Campelo Franca Pinto}\\
+	Department of Computer Science\\
+	School of Engineering and Applied Sciences\\
+	Aston University\\
+	\texttt{f.campelo@aston.ac.uk}
 }
@@
 This study replicates work done by Berisha and Liss and extends this by adding Donald Trump (DJT) as an alternative, more appropriate comparison to RR as he is much closer in age than GHWB.
-This work contributes to our understanding of the pre-diagnosis phase of alzheimer's disease and it's effect on language.
+This work contributes to our understanding of the pre-diagnosis phase of Alzheimer's Disease and it's effect on language.
 We explore a wider range of features than previously looked at with this dataset and our hypothesis is that there will be features of RR's language which change more noticably over time when compared with the two closest appropriate controls.
+In addition, our assumption is that any language deterioration found will be non-linear and so we will be looking at non linear and linear correlations within the data to see if decline is constant over time
 
@@
 DJT as with GHWB has no known diagnosis of AD.
-I used the press conference transcripts in the American Presidency Project (APP) archive as a data source for this project.
+We used the press conference transcripts in the American Presidency Project (APP) archive as a data source for this project.
 The APP is a comprehensive and organized searchable database of presidential documents, including transcripts of speeches, transcripts of news conferences, and other public documents.
@@
 \subsection{Pre-processing}
-To generate the files necessary for analysis, I downloaded each transcript and performed the following changes.
-I omitted the prepared statement by the president and any speech by other individuals.
-I started each transcript at the beginning of the first answer to a question.
-I filtered any annotations that were added to the transcript, including any references or clarifications, and any laughter.
-It's worth noting that there appears to be a difference in how 'hesitations' were marked down between each president, for RR hesitations were marked by a single hyphen whereas for GHWB hesitations are marked by a double hyphen.
+To generate the files necessary for analysis, we downloaded each transcript and performed the following changes.
+We omitted the prepared statement by the president and any speech by other individuals and started each transcript at the beginning of the first answer to a question by a member of the press.
+We filtered any annotations that were added to the transcript, including any references or clarifications, and any laughter.
+It's worth noting that there appears to be a difference in how 'hesitations' were marked down between each president, for RR \& DJT hesitations were marked by a single hyphen whereas for GHWB hesitations are marked by a double hyphen.
 In order to maintain consistency when parsing through the documents, We have changed both types of hesitation to be marked by a single hyphen.
@@
 We did not control for the length of the document, but generated features which would normalise by the length of the document.
-We therefore was able to include all press conferences by both Ronald Reagan and George H.W.
-Bush where there was a question and answer session conducted at least in part by the sitting president (2 press conferences of GHWB were omitted due to a lack of a question and answer session).
+We therefore was able to include all press conferences by all presidents analysed where there was a question and answer session conducted at least in part by the president in question (2 press conferences of GHWB were omitted due to a lack of a question and answer session).
 
-\subsection{Feature Selection}
-We calculated the following features for each transcript in turn using the NLTK (see section 2 for a description) \cite{Bird2009}, Linguistic Inquiry and Word Count (LIWC) and Python. 
+\subsection{Feature Generation}
+We calculated the following features for each transcript in turn using the Natural Language ToolKit (NLTK) package \cite{Bird2009}, Linguistic Inquiry and Word Count (LIWC) \cite{Pennebaker2015} and Python.
+Words were stemmed to their root form using the Porter2 (Snowball) stemmer algorithm.
 \par 
-\subsubsection{Measures of lexical variation}
-Lexical variation is a measure of the variety of language used within a given document.
-A document is said to have high lexical variation if the number of unique words is large.
+\subsubsection{Measures of lexical diversity}
+Lexical diversity is a measure of the variety of language used within a given document.
+A document is said to have high lexical diversity if the number of unique words is large.
 We constructed four features of lexical variation.
@@
 We also used the TTR formula, Brunet's Index and Honore's Statistic as other measures of lexical diversity \cite{Richards1987}.
-
+\par 
 Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
@@
 \par 
-Brunet's Index (W) differentiates itself for TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
+Brunet's Index (W) differentiates itself from TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
 \begin{equation} \label{x2}
@@
 where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
-\subsubsection{Fillers, Non-Specific Nouns and LI Verbs}
-For these features, we counted the number of occurrences for different categories (see table for list of categories tracked and the words counted).
-The features were used by Berisha and Liss in their research \cite{Berisha2015} and were taken from work done by Bird et al \cite{Bird2000}.
-These findings were.
-number of unique words, non-specific nouns and fillers and low imageability (LI) verbs.
-Imageability is characterized, according to Berisha and Liss, as the ease with which a term gives rise to a sensory .mental image.
+\subsubsection{Fillers, Non-Specific Nouns and Low Imageability Verbs}
+Fillers, Non-Specific Nouns and Low Imageability Verbs were features used by Berisha and Liss in their research \cite{Berisha2015} and were taken from work done by Bird et al \cite{Bird2000}.
+Fillers can be described as a potentially meaningless word that marks a pause or hesitation in speech.
+In those with MCI and AD, these words can be used to temporarily disguise problems in thought processes or word finding difficulties.
+Non-Specific Nouns refer to a category or an unspecified member of a given category, once again this can be characterised as a compensatory strategy for word finding difficulties.
+Imageability is characterized, according to Berisha and Liss, as the ease with which a word provokes a mental image of what the word describes.
 
@@
 	\begin{center}
-	\begin{tabular}{ | p{3cm} | p{6cm} |}
+	\begin{tabular}{ | p{3cm} | p{10cm} |}
 		\hline
 		Category & Words \\ \hline
-		Fillers & \textit{"well", "so", "basically", "actually", "literally", "um", "ah"} \\ \hline
+		Fillers & \textit{"um", "uh", "er", "ah", "like", "okay", "right",  "you know", "well", "so", "basically", "actually", "literally"} \\ \hline
 		Non Specific Nouns & \textit{"something", "anything", "thing", "everything"} \\ \hline
 		LI Verbs & \textit{"be", "come", "do", "get", "give", go", "know", "look", "make", "see", "tell", "think", "want"} \\ \hline
-
 	\end{tabular}
-	\caption{\label{tab:table-name}Categories and Words Counted}
+	\caption{\label{tab:table-name}Examples of words belonging to the categories Fillers, Non-Specific Nouns and Low Imageability Verbs}
 	\end{center} 
@@
 \subsubsection{Usage of parts of speech}
-This section involves using a Part of Speech tagger (PoS) which analyses a sentence and assigns a 'tag' to each word based on the function the word has in a sentence.
+Using a Part of Speech tagger (PoS) on each transcript analyses each sentence within the transcript and assigns a 'tag' to each word based on the function the word has in a sentence.
 At a basic level this can be divided into the eight defined parts of speech: 'nouns', 'pronouns', verbs', 'adjectives', 'adverbs', 'conjunctions', 'prepositions' and 'interjections' but can be further subcategorised.
@@
 \subsubsection{Linguistic Inquiry and Word Count(LIWC)}
-The LIWC
+The LIWC is a text processor which analysis words within a given a document and compares the words in the LIWC dictionary file.
+The LIWC dictionary file contains over 6000 words which are categorised into approximately 90 different types.
+For example the word 'cried' is part of five word categories: sadness, negative emotion, overall affect, verbs and past focus \cite{Pennebaker2015}.
+Each word in the document or set of documents is searched for in the dictionary file and if found, the appropriate type is incremented by 1.
+What is output is an analysis of word usage in reference to each category.
+This is particularly useful in analysing both structural language changes but also the content or themes of language for a given transcript or set of transcripts.
+\par 
+As with Berisha and Liss, we analysed the transcripts over time in order to determine if there was an underlying temporal trend that may indicate that different parts of language increase or decrease over time.
+In AD, cognitive abilities are said to deteriorate over time and therefore it may be useful to analyse temporal trends that indicate a potential deterioration of cognition.
 
@@
 \section{Discussion}\label{discussion}
-\section{Discussion}
 President Reagan received his diagnosis of AD in August 1994 but using transcripts of speeches he made in his two terms as President (January 1981 - January 1989) we have be able to identify certain changes in his use of language that we might ascribe to the onset of MCI and early AD.
-Despite differences in our methodology, our research supports the findings of Berisha and Liss in that we both find a significant decrease in unique words over time and an increase in non-specifc noun usage.
-Compared to our controls (GWHB andDJT), we find some slight trends with GWHB but no such trends with DJT in his speech albeit his samples of speech span a shorter amount of time.
+Despite differences in our methodology, our research supports some of the findings of Berisha and Liss in that we find an increase in non-specifc noun usage.
+Compared to our controls (GWHB and DJT), we find some slight trends with GWHB but no such trends with DJT in his speech albeit his samples of speech span a shorter amount of time.
+\par
 
@@
 Our results are supported by existing literature and this potentially means that language analysis in the way we have structured it may have diagnostic or prognostic properties.
+One of the most striking differences between this study and the original is when looking at Unique Words.
+Whilst Unique Words as a feature is significant, which was a finding in the original study, when we controlled via normalisation, we found that this was not significant.
 
@@
 In addition, this replication and extension has demonstrated the potential utility of using longitudinal data as a means of comparing language use of a person at two or more time periods and using this information as a diagnostic aid for MCI and therefore more work would be helpful from a longitudinal perspective to see if this approach may be valid in moving towards a solution for this particular problem.
-\par
 

### 2019-10-16 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
Presidents Corpus - Intro and Method

@@
 GHWB participated in 136 press conferences over four years (an average of 34 a year) and DJT participated in 29 press conferences to date (an average of 19.3 per year).
-Equally, there are variances in the average number of words.
+Equally, there are differences in the average number of words.
 RR produced an average of 3424 words per conference compared to 2608 by GHWB (unpaired t = 4.434, p\textless 0.001) and DJT at 1849 words (unpaired t = 6.524, p\textless 0.001).
@@
 		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
+		Mean Length of Utterance & 23.17 (1.402) & 18.71 (2.067) & 481.82 (221.29) \\ \hline
 		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
@@
 		Total Words & \textbf{4.434***} & \textbf{6.524***} & \textbf{2.899**} \\ \hline
-		Unique Words & \textbf{6.832***} & \textbf{11.403***} & \textbf{4.137***} \\ \hline
+		Unique Stems & \textbf{10.878***} & \textbf{10.111***} & \textbf{4.148***} \\ \hline
+		Mean Length of Utterance & \textbf{16.175***} & \textbf{25.084***} & \textbf{13.484***} \\ \hline	
 		Non Specific Nouns & \textbf{7.877***} & \textbf{3.426**} & -0.574 \\ \hline

### 2019-10-17 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
update on Presidents paper

@@
 \usepackage{float}
+\usepackage{graphicx}
 
@@
 \section{Introduction}\label{introduction}
-Alzheimer's Disease (AD) and other forms of dementia affect a significant proportion of the older age population in the world today and is currently the sixth leading cause of death in the US and was named the leading killer of women in the UK.
-According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
-This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population \cite{AlzheimersSociety2014}.
-From a financial perspective, they estimate an annual spend of £4.3 billion of which approximately £85 million is spent solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
+Alzheimer's Disease (AD) and other forms of dementia affect a significant proportion of the global older adult population and the impact on morbidity and mortality rates is considerable.
+Dementia and AD are currently the leading cause of death in England and Wales and the sixth leading cause of death in the United States (US).
+A recent report commissioned by the Alzheimer's Society in 2015 estimated that in the UK by 2015 there would be approximately 855,000 people rising to 1 million by 2021.
+This represents 1 in 79 of the general population rising to 1 in 14 of those aged 65 or over \cite{AlzheimersSociety2014}.
+Worldwide, there are 46 million people with a diagnosis of dementia globally and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+From a financial perspective, they estimated annual spend on dementia healthcare is £4.3 billion of which approximately £85 million is spent on diagnosis.
+The total financial burden of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually.
 Globally, this picture is a lot bleaker.
-Another report by Alzheimer's Disease International suggests that in 2015 there were 46 million people with a diagnosis of dementia globally and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
-The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
+Another report by Alzheimer's Disease International suggests that in 2015 there were  The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
 \par 
@@
 In AD, cognitive abilities are said to deteriorate over time and therefore it may be useful to analyse temporal trends that indicate a potential deterioration of cognition.
+In order to get a accurate measure of deterioration over time, we time stamped each transcript in relation to number of days from the first transcript to the date of the transcript being analysed as the press conferences were not evenly spaced out during the presidencies.
 
@@
 RR produced an average of 3424 words per conference compared to 2608 by GHWB (unpaired t = 4.434, p\textless 0.001) and DJT at 1849 words (unpaired t = 6.524, p\textless 0.001).
+
 \begin{table}[H]
@@
 		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
-		Mean Length of Utterance & 23.17 (1.402) & 18.71 (2.067) & 481.82 (221.29) \\ \hline
-		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
-		LI Verbs & 124.22 (17.89)& 103.45 (51.87) & 84.48 (75.78) \\ \hline
+		Mean Length of Utterance & 23.17 (1.402) & 18.71 (2.067) & 13.84 (1.619) \\ \hline
 	\end{tabular}
-	\caption{\label{tab:table-name}Means and Standard Deviations of important features}
+	\caption{\label{tab:table-name}Means and Standard Deviations of general features for each set of transcripts}
 	\end{center} 
@@
 
-In terms of unique words, we found that RR used significantly more unique words, non-specific nouns and low imageability verbs than GHWB and DJT (see Table 3.3).
+In terms of more specific language differences between the presidents, we found that RR used significantly more unique words, non-specific nouns and low imageability verbs than GHWB and DJT (see Table 3.3).
+The mean length of utterance for RR was significantly greater than that of GHWB and DJT.
 Some of these differences are due to the length of the sample, particularly in the case of DJT where his average sample is almost half the sample of RR.
@@
 
+\subsection{Longitudinal Analysis}
 We then looked at the data from a longitudinal perspective as we are interested seeing whether we can track various language variables and their progress over time.
@@
 Given the number of hypotheses being tested, it is necessary to apply some correction for family wise error rate and therefore a Hochberg step-up procedure was applied to the results of the analysis.
-Because features have been generated from multiple sources, there are some cases where there is significant overlap between the features.
-Where possible the most appropriate feature has been selected, for example in the case of the raw count of pronouns and pronouns normalised by the length of the document, the latter has been selected.
-Where there is doubt as to the most appropriate feature, all features have been included.
+We also used the Benjamini-Yekutieli procedure for controlling for False Discovery rate and came up with a final list of important features for each president.
+For RR, this was 22 features, for GHWB this was 14 features and there were no significant features for DJT after applying these methods.
+The features were generated from multiple sources and therefore there was some significant overlap between the features.
+In these cases, the normalised features were prefered to raw count features.
+Where there was doubt as to the most appropriate feature, all features have been included.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | l | p{2.5cm} | p{2.5cm} | p{2.5cm} |}
+		\hline
+		& RR R-Squared  & GWB R-Squared & DJT R-Squared \\ \hline
+		ppron & \textbf{0.700***} & 0.2 & 0.35 \\ \hline
+		social & \textbf{0.698***} & 0.2 & 0.35 \\ \hline
+		NounsNormalised & \textbf{-0.689***} & 0.2 & 0.35 \\ \hline
+		function & \textbf{0.670***} & 0.2 & 0.35 \\ \hline
+		conj &\textbf{0.644***} & 0.2 & 0.35 \\ \hline
+		PronounsNormalised & \textbf{0.631***} & 0.2 & 0.35 \\ \hline
+		Analytic & \textbf{-0.626***} & 0.2 & 0.35 \\ \hline
+		NN & \textbf{-0.601***} & 0.2 & 0.35 \\ \hline
+		male & \textbf{0.585***} & 0.2 & 0.35 \\ \hline
+		UniqueWords & \textbf{-0.578***} & 0.2 & 0.35 \\ \hline
+		WDT & \textbf{-0.577***} & 0.2 & 0.35 \\ \hline
+		shehe & \textbf{0.529*} & 0.2 & 0.35 \\ \hline
+		VBZ & \textbf{-0.521*} & 0.2 & 0.35 \\ \hline
+		JJ & \textbf{-0.518*} & 0.2 & 0.35 \\ \hline
+		Fillers & \textbf{-0.518*} & \textbf{-0.518*}  & 0.35 \\ \hline	
+		EX & \textbf{-0.518*} & \textbf{-0.518*}  & 0.35 \\ \hline	
+		achieve & \textbf{-0.518*} & \textbf{-0.518*}  & 0.35 \\ \hline		
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+
+	\end{tabular}
+	\caption{\label{tab:table-name}Pearson Correlations for Features}
+	\end{center} 
+\end{table}
 
@@
 \par 
-
 A criticism of Berisha and Liss's work is the problems they had with normalising the transcripts in terms of length.
@@
 Whilst it is important to control for outliers, there are other ways in which we can control for length of sample.
+In this paper, we controlled for transcript length by dividing any features that were raw counts by the total length of the transcript.
+Given the increased amounts of hypothesis testing, it became necessary to control for false discoveries.
 \par 
@@
 Whilst Unique Words as a feature is significant, which was a finding in the original study, when we controlled via normalisation, we found that this was not significant.
+\par
+Another observation is that when conducting fitting a linear model to this data does not tell the whole story.
+For example, consider the following two plots.
+One shows a model fit to a linear  model, and the other shows a model fit using local polynomial regression model which aims to model the data more closely.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=300px, height=300px]{RRppromLM.jpg}
+\caption{RR Personal Pronouns - Linear Model}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=300px, height=300px]{RRppromLM.jpg}
+\caption{RR Personal Pronouns - Local Regression Model}
+\end{figure}
 

### 2019-10-18 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
updated pres paper
updated changes from HV

@@
 \section{Introduction}\label{introduction}
-Alzheimer's Disease (AD) and other forms of dementia affect a significant proportion of the global older adult population and the impact on morbidity and mortality rates is considerable.
-Dementia and AD are currently the leading cause of death in England and Wales and the sixth leading cause of death in the United States (US).
-A recent report commissioned by the Alzheimer's Society in 2015 estimated that in the UK by 2015 there would be approximately 855,000 people rising to 1 million by 2021.
-This represents 1 in 79 of the general population rising to 1 in 14 of those aged 65 or over \cite{AlzheimersSociety2014}.
+The diagnosis of dementia is generally made when there is a decline in brain function due to physical changes in the brain \cite{Albert2011}.
+It affects a significant proportion of the global older adult population and the impact on morbidity and mortality rates is considerable.
+Dementia is currently the leading cause of death in England and Wales and the sixth leading cause of death in the United States (US).
+A 2014 report commissioned by the Alzheimer's Society estimated that in the UK by 2015 there would be approximately 855,000 people rising to 1 million by 2021 \cite{}.
+This represents 1 in 79 of the total UK population rising to 1 in 14 of those aged 65 or over \cite{AlzheimersSociety2014}.
 Worldwide, there are 46 million people with a diagnosis of dementia globally and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
-From a financial perspective, they estimated annual spend on dementia healthcare is £4.3 billion of which approximately £85 million is spent on diagnosis.
+From a financial perspective, the cost burden is also significant.
+The estimated annual spend on dementia healthcare in the UK is £4.3 billion of which approximately £85 million is spent on diagnosis.
 The total financial burden of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually.
 Globally, this picture is a lot bleaker.
-Another report by Alzheimer's Disease International suggests that in 2015 there were  The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
+Another report by Alzheimer's Disease International suggests that in 2015 there were  The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars \cite{}.
 \par 
-AD is a neurodegenerative disease in which a definitive diagnosis can only be produced at post-mortem.
+There are different types of dementia including Alzheimer's Disease (AD), vascular dementia, dementia with Lewy bodies and fronto-temporal dementia, all of which are currently incurable.
+AD is a progressive neurodegenerative disease and is the most common type of dementia, responsible for approximately XX\% of all cases \cite{}.
+Currently, a definitive diagnosis for AD can only be produced at post-mortem.
 However, there are a number of psychological and physiological indicators that can indicate that dementia is present.
-From a physiological perspective, researchers have identified two proteins called beta-amyloid and tau.
-In a typical case, tau accumulates and eventually forms tangles inside neurons and beta-amyloid clumps into plaques which slowly builds up between neurons.
-At a certain point, the levels of beta-amyloid rise and trigger a more rapid spread of tau throughout the brain.
-Eventually, due to this and other changes, neurons lose their ability to communicate and the brain starts to shrink.
-This leads to some of the more psychological symptoms, those who have dementia demonstrate cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}.
-In addition, these symptoms are often accompanied by emotional problems such as depression and behavioural difficulties.
-As more neurons die throughout the brain, a person with Alzheimer's gradually loses the ability to think, remember, make decisions and function independently.
+From a physiological perspective, researchers have identified two proteins called beta-amyloid and tau have been identified as two of the key proteins that have a role in initiation and progression of AD.
+In a typical case beta-amyloid clumps into plaques which slowly build up between neurons with tau proteins accumulating and eventually forming tangles inside neurons.
+At a certain point, the levels of beta-amyloid rise triggering a more rapid spread of tau throughout the brain.
+Eventually, due to these and other changes, neurons lose their ability to communicate and the brain starts to shrink.
+This leads to the development of some of the psychological symptoms associated with AD, primarily cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}.
+These cognitive symptoms are often accompanied by emotional problems such as depression and behavioural difficulties.
+As more neurons die throughout the brain, a person with AD gradually loses the ability to think, remember, make decisions and function independently.
 \par
-Despite this growing problem and an increasing understanding about how AD affects the brain there are no medications that improve the prognosis of those with AD.
+Despite the increasing prevalence of AD and an improved understanding about how it affects the brain there are no medications that improve prognosis.
 All the medications that are currently on the market are designed to manage symptoms.
-Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{Cummings2014}.
-Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place \cite{Cummings2014}.
-Research has started to be more focused on AD at it's earliest stages which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.
+Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage of these drugs fail in clinical trial stage of the drug discovery process ((99.6\% failure rate vs 80\% for systemic anti-cancer drugs) \cite{Cummings2014}.
+Cummings et al proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already occurred \cite{Cummings2014}.
+Research focus has now started to shift to the earlier stages of AD (i.e.
+symptomatic pre-dementia phase of AD) which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.
 \par
-One of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of cognitive impairment or dementia.
+One of the challenges associated with the early detection of AD is differentiating natural age associated memory impairment and cognitive decline due to aging from with decline due to AD.
 This challenge is often complicated further due to the large variation in the cognitive abilities and educational background of individuals.
-Albert and his team have worked to define clinical criteria which professionals can use to diagnose MCI due to AD and differentiate this from age-associated memory impairment and age-associated cognitive decline.
+The work of Albert et al helps to address this by the development of clinical criteria which professionals can use to diagnose MCI due to AD.
 One of the most important observations from this piece of work is that a diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}, and this is essentially to place more importance on the trajectory of a person's cognitive abilities rather than a person's cognitive ability in general.
 \par 
-There has been a significant research in the area of language deterioration as a means of detecting Alzheimer's Disease.
-This usually takes the form analysis of speech recorded as part of a cognitive assessment such as the Picture Description Task \cite{Orimaye2014,Fraser2015}.
+There has been a significant research in the area of language deterioration as a method of detecting AD at an earlier stage.
+This usually takes the form of recording speech whilst patients' undertake a cognitive assessment such as the Picture Description Task \cite{Orimaye2014,Fraser2015}.
 Given that language samples are relatively easy to collect, research has moved towards analysis of spontaneous speech.
-An good example of this type of research is the study conducted by Berisha and Liss which looked at the differences in language use between two US presidents, Ronald Reagan (RR), who would go on to receive a diagnosis of Dementia and George H.
+The work of Berisha and Liss is a good example of this, this study compared the differences in language use between two US presidents, Ronald Reagan (RR), who would go on to receive a diagnosis of Dementia and George H.
 W.
 Bush (GHWB) who acted as a matched control based on Age \cite{Berisha2015}.
-They found several differences in language use over time which they felt acted as indicators of RR's difficulties with language due to dementia.
-These significant differences were in the number of unique words used per speech, the use of non-specific nouns and fillers and low-imageability verbs \cite{Berisha2015}.
+They found several differences in language use over time which they felt acted as indicators of RR's difficulties with language due to AD.
+The statistically significant differences identified included the number of unique words used per speech, the use of non-specific nouns and fillers and low imageability verbs \cite{Berisha2015}.
 \par   
-This study replicates work done by Berisha and Liss and extends this by adding Donald Trump (DJT) as an alternative, more appropriate comparison to RR as he is much closer in age than GHWB.
-This work contributes to our understanding of the pre-diagnosis phase of Alzheimer's Disease and it's effect on language.
-We explore a wider range of features than previously looked at with this dataset and our hypothesis is that there will be features of RR's language which change more noticably over time when compared with the two closest appropriate controls.
-In addition, our assumption is that any language deterioration found will be non-linear and so we will be looking at non linear and linear correlations within the data to see if decline is constant over time
+This study replicates the research of Berisha and Liss and extends this by adding Donald Trump (DJT) as an additional comparison to RR.
+This study aims to contribute to our understanding of the pre-diagnosis phase of Alzheimer's Disease and it's effect on language.
+Our hypothesis is that there will be features of RR's language which change more noticably over time when compared with the two closest appropriate controls.
+Compared to the previous research study, we explored a wider range of features within this dataset.
+In addition, our assumption was that any language deterioration found would be non-linear and so we will be looking at non linear and linear correlations within the data were looked at to see if language decline is constant over time
 
 \section{Methodology}\label{methodology}
-We took 46 transcripts of Ronald Reagan’s (RR) press conferences from 1981 to 1988 and compared them with 134 press conferences by George H.
-W.
-Bush (GHWB) and 29 press conferences conducted by Donald J.
-Trump (DJT).
-We analyzed transcripts for lexical features shown to change longitudinally with dementia.
-These features encompassed, word level, sentence level and document level features and included a number of features contained in the study by Berisha and Liss with the aim of replicating and extending on their findings.
-We compared the trends described in the transcripts of RR and GHWB, but also included DJT.
-Berisha and Liss originally made the comparison as it GHWB (GHWB - age at the start of presidency - 64 years, 222 days) was the closest match to RR in terms of age (RR  - age at the start of presidency, 69 yeaars and 349 days).
-However, with the inauguration of Trump, he now is the closest comparible president in terms of age (DJT - age at the start of presidency - 70 years, 220 days).
-It would be interesting to look at a comparison of RR and DJT to see whether the comparisons made by Berisha and Liss hold true with this more appropriate match (in terms of age).
-DJT as with GHWB has no known diagnosis of AD.
+We took 46 transcripts of press conferences given by RR from 1981 to 1988 and compared them with 134 press conferences from 1989 to 1993 given by GHWB and 29 press conferences from 2016 to 2019 conducted by DJT.
+We analyzed transcripts for lexical markers (described below) shown to change longitudinally with AD.
+These lexical features encompassed, word level, sentence level and document level features and included a number of features contained in the study by Berisha and Liss.
+\par 
+In the original study, GHWB was selected as the comparator president as he was the closest match in terms of age to RR (GHWB - age at the start of presidency: 64 years and 222 days vs RR: 69 years and 349 days).
+However, since his inauguration, DJT is now the closest comparible president in terms of age (DJT - age at the start of presidency: 70 years and 220 days).
+We included DJT who like GHWB has no known diagnosis of AD to determine whether the comparisons made by Berisha and Liss hold true with this closer presidential match in terms of age.
+\par 
 We used the press conference transcripts in the American Presidency Project (APP) archive as a data source for this project.
 The APP is a comprehensive and organized searchable database of presidential documents, including transcripts of speeches, transcripts of news conferences, and other public documents.
+These documents are open access and can be downloaded at any time from the APP archive.
 
@@
 
-\begin{figure}[H]
+\begin{figure}[b]
 	\centering
-\includegraphics[width=300px, height=300px]{RRppromLM.jpg}
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=150px, height=150px,  natwidth=300, natheight=300]{RRppronLM.jpg}
 		\caption{RR Personal Pronouns - Linear Model}
-\end{figure}
+	\end{minipage}
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=300px, height=300px]{RRppromLM.jpg}
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=150px, height=150px, natwidth=300, natheight=300]{RRppronLOESS.jpg}
 		\caption{RR Personal Pronouns - Local Regression Model}
+	\end{minipage}		
 \end{figure}
@@
 \bibliographystyle{unsrt}
-\bibliography{main}
+\bibliography{main3}
 

### 2019-10-21 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
updated Presidents paper

@@
 Dementia is currently the leading cause of death in England and Wales and the sixth leading cause of death in the United States (US).
-A 2014 report commissioned by the Alzheimer's Society estimated that in the UK by 2015 there would be approximately 855,000 people rising to 1 million by 2021 \cite{}.
+A 2014 report commissioned by the Alzheimer's Society estimated that in the UK by 2015 there would be approximately 855,000 people rising living with dementia increasing to 1 million by 2021 \cite{}.
 This represents 1 in 79 of the total UK population rising to 1 in 14 of those aged 65 or over \cite{AlzheimersSociety2014}.
@@
 Globally, this picture is a lot bleaker.
-Another report by Alzheimer's Disease International suggests that in 2015 there were  The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars \cite{}.
+The worldwide cost of dementia in 2018 was estimated to be in the region of one trillon US dollars \cite{}.
 \par 
@@
 They found several differences in language use over time which they felt acted as indicators of RR's difficulties with language due to AD.
-The statistically significant differences identified included the number of unique words used per speech, the use of non-specific nouns and fillers and low imageability verbs \cite{Berisha2015}.
+Differences in features of language identified to be statistically significant included the number of unique words used per speech, the use of non-specific nouns and fillers and low imageability verbs \cite{Berisha2015}.
 \par   
@@
 Our hypothesis is that there will be features of RR's language which change more noticably over time when compared with the two closest appropriate controls.
-Compared to the previous research study, we explored a wider range of features within this dataset.
+Compared to the previous research study, we explored a wider range of language features within this dataset.
 In addition, our assumption was that any language deterioration found would be non-linear and so we will be looking at non linear and linear correlations within the data were looked at to see if language decline is constant over time
@@
 \section{Methodology}\label{methodology}
-We took 46 transcripts of press conferences given by RR from 1981 to 1988 and compared them with 134 press conferences from 1989 to 1993 given by GHWB and 29 press conferences from 2016 to 2019 conducted by DJT.
-We analyzed transcripts for lexical markers (described below) shown to change longitudinally with AD.
-These lexical features encompassed, word level, sentence level and document level features and included a number of features contained in the study by Berisha and Liss.
+We took 46 transcripts of press conferences given by RR (from 1981 to 1988) and compared them with 134 press conferences (from 1989 to 1993) given by GHWB and 29 press conferences (from 2016 to 2019) conducted by DJT.
+We analysed transcripts for language features (described below) shown to change longitudinally with AD.
+These language features are analysed at the word level, sentence level and document level.
 \par 
@@
 
-\bibliographystyle{unsrt}
+\bibliographystyle{apalike}
 \bibliography{main3}

### 2019-10-23 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
updated Presidents Corpus
updated Pres Paper

@@
 \par   
-This study replicates the research of Berisha and Liss and extends this by adding Donald Trump (DJT) as an additional comparison to RR.
-This study aims to contribute to our understanding of the pre-diagnosis phase of Alzheimer's Disease and it's effect on language.
-Our hypothesis is that there will be features of RR's language which change more noticably over time when compared with the two closest appropriate controls.
-Compared to the previous research study, we explored a wider range of language features within this dataset.
-In addition, our assumption was that any language deterioration found would be non-linear and so we will be looking at non linear and linear correlations within the data were looked at to see if language decline is constant over time
+Berisha and Liss have developed some very interesting ideas about how we might track changes over time in various lexical features which have been associated with the development of Alzheimer's Disease.
+We refine these ideas with the aim of exploring the potential for a protocol that can be used for analyzing language deterioration.
+This study extends the work of Berisha and Liss in a number of ways.
+Our hypotheses are:
+\begin{enumerate}
+	\item Language use of RR will be significantly different to our controls (GHWB, DJT).
+	\item Additional lexical features will illustrate significant cognitive decline in RR.
+	\item Correlations of lexical features over time will be a significantly improved using non-linear models compared to linear models.
+\end{enumerate}
 
@@
 
-\begin{figure}[b]
+\begin{figure}
 	\centering
-	\begin{minipage}[b]{0.4\textwidth}
-		\includegraphics[width=150px, height=150px,  natwidth=300, natheight=300]{RRppronLM.jpg}
-		\caption{RR Personal Pronouns - Linear Model}
+	\begin{minipage}[b]{.45\textwidth}
+		\includegraphics[width=6cm, height=5cm]{comp1.png}
 	\end{minipage}
-
-	\begin{minipage}[b]{0.4\textwidth}
-		\includegraphics[width=150px, height=150px, natwidth=300, natheight=300]{RRppronLOESS.jpg}
-		\caption{RR Personal Pronouns - Local Regression Model}
+	\hfill
+		\begin{minipage}[b]{.45\textwidth}
+		\includegraphics[width=6cm, height=5cm]{comp2.png}
 	\end{minipage}
@@
 
-\bibliographystyle{apalike}
+\bibliographystyle{ieeetr}
 \bibliography{main3}

### 2019-11-07 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
Added abstract

@@
 \begin{abstract}
-This is the paper's abstract \ldots
+This paper looks at evidence of cognitive decline in three presidents speeches.
+Ronald Reagan, who would later go on to receive a diagnosis of Alzheimer's Disease and both George Herbert Walker Bush and Donald Trump who have no known diagnosis of dementia.
+This study used the methodology used by Berisha and Liss as a starting off point and offers some refinements in how the data was pre-processed, how the features were generated and how the analysis was conducted.
+We hypothesized that there would be features identified which would decline over time for RR, but not for our controls.
+We also hypothesised that a non-linear model would map decline more accurately than linear models.
+We collected data from the Question and Answer portions of the presidents press conferences and used Natural Language Processing to generate features for analysis.
+We found that there were a number of lexical features that significantly changed over time for RR but did not change for either GHWB or DJT.
+These features have a foundation in psychological literature and have been linked to cognitive decline in Alzheimer's Disease.
+Also, we found that non-linear models were able to more accurately map declines in RR's language over time.
 \end{abstract}
@@
 	\item Language use of RR will be significantly different to our controls (GHWB, DJT).
-	\item Additional lexical features will illustrate significant cognitive decline in RR.
-	\item Correlations of lexical features over time will be a significantly improved using non-linear models compared to linear models.
+	\item Modelling the dynamics of lexical features over time can be improved using non-linear models compared to linear models.
 \end{enumerate}
@@
 This is particularly useful in analysing both structural language changes but also the content or themes of language for a given transcript or set of transcripts.
-\par 
+
+\subsubsection{Longitudinal Analysis}
 As with Berisha and Liss, we analysed the transcripts over time in order to determine if there was an underlying temporal trend that may indicate that different parts of language increase or decrease over time.
@@
 Also, the problem of finding an appropriate control is a common one in this domain.
+This variability in the number and quality of the transcriptions raises some doubts as to the results of DJT.
+Is it the lack of instances of data, or that the number of words per transcript is significantly less than with RR or GHWB.
+This does have further implications for using a protocol such as this in a more general way, as it will impact how data is collected.
+\par
+
 Given that those with MCI and early dementia have such variable presentations, it might prove of limited value in matched pairs design. 
@@
 Equally, it would be helpful to have controls to ascertain what would be usual to expect in the decline of language in a healthy older adult.
+\par 
+The results of this work also identify some clues as to what
 

### 2019-11-07 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
added Abstract

@@
 \begin{abstract}
-This paper looks at evidence of cognitive decline in three presidents speeches.
+Several research efforts aimed at improving the early diagnosis of Alzheimer's Disease have focused on the longitudinal analysis of lexical features which have been implicated in the pre-clinical development of Alzheimer's Disease.
+We build on previous studies on this and continue to refine a method of extracting and analyzing language from transcripts with the long-term goal of developing this method for use in clinical practice.
+We evaluate this method on the unscripted question and answer portions of news conferences from three Presidents.
 Ronald Reagan, who would later go on to receive a diagnosis of Alzheimer's Disease and both George Herbert Walker Bush and Donald Trump who have no known diagnosis of dementia.
-This study used the methodology used by Berisha and Liss as a starting off point and offers some refinements in how the data was pre-processed, how the features were generated and how the analysis was conducted.
-We hypothesized that there would be features identified which would decline over time for RR, but not for our controls.
-We also hypothesised that a non-linear model would map decline more accurately than linear models.
-We collected data from the Question and Answer portions of the presidents press conferences and used Natural Language Processing to generate features for analysis.
-We found that there were a number of lexical features that significantly changed over time for RR but did not change for either GHWB or DJT.
-These features have a foundation in psychological literature and have been linked to cognitive decline in Alzheimer's Disease.
-Also, we found that non-linear models were able to more accurately map declines in RR's language over time.
+We found that there were several lexical features that significantly changed over time for President Reagan but did not change for either President Bush or President Trump.
+These features include unique words and noun usage which significantly decrease over time and pronouns (particularly personal pronouns), conjunctions and words which are related to social processes significantly increase over time.
+No significant trends were found in the analysis of the transcripts of President Bush or President Trump.
+These features have support in psychological literature and have been linked to cognitive decline in Alzheimer's Disease.
+We also found that non-linear models were able to more accurately map declines in RR's language over time in comparison to linear models.
+This work points the way to developing a comparatively non-invasive protocol that can be used in clinical settings that may be able to spot atypical declines in language use over time that may indicate the presence of Alzheimer's Disease.
+
 \end{abstract}
@@
 Berisha and Liss have developed some very interesting ideas about how we might track changes over time in various lexical features which have been associated with the development of Alzheimer's Disease.
+However we feel that cognitive decline in pre-clinical AD is not accurately modeled as a linear process and therefore we explored the application of generalised additive models (GAMs) to this data which have no assumed understanding of the distribution of the data.
 We refine these ideas with the aim of exploring the potential for a protocol that can be used for analyzing language deterioration.
-This study extends the work of Berisha and Liss in a number of ways.
-Our hypotheses are:
+
+This study extends the work of Berisha and Liss in a number of ways. Our hypotheses are:
 \begin{enumerate}
@@
 In order to get a accurate measure of deterioration over time, we time stamped each transcript in relation to number of days from the first transcript to the date of the transcript being analysed as the press conferences were not evenly spaced out during the presidencies.
+\par
+We looked at 
+\par 
+In order to compare the fit of linear models with
 

### 2019-11-07 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
added abstract

--- 2019-11-07 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
@@
 \usepackage{graphicx}
-
 \begin{document}
@@
 \begin{abstract}
-This paper looks at evidence of cognitive decline in three presidents speeches.
-Ronald Reagan, who would later go on to receive a diagnosis of Alzheimer's Disease and both George Herbert Walker Bush and Donald Trump who have no known diagnosis of dementia.
-This study used the methodology used by Berisha and Liss as a starting off point and offers some refinements in how the data was pre-processed, how the features were generated and how the analysis was conducted.
-We hypothesized that there would be features identified which would decline over time for RR, but not for our controls.
-We also hypothesised that a non-linear model would map decline more accurately than linear models.
-We collected data from the Question and Answer portions of the presidents press conferences and used Natural Language Processing to generate features for analysis.
-We found that there were a number of lexical features that significantly changed over time for RR but did not change for either GHWB or DJT.
-These features have a foundation in psychological literature and have been linked to cognitive decline in Alzheimer's Disease.
-Also, we found that non-linear models were able to more accurately map declines in RR's language over time.
+Several research efforts aimed at improving the early diagnosis of Alzheimer's Disease have focused on the longitudinal analysis of lexical features which have been implicated in the pre-clinical progression of this disease.
+We build on these studies and continue to refine a method for extracting and analysing language from transcripts.
+This method is based on a statistical evaluation of the most relevant features, as well as on the use of Generalised Additive Models instead of Linear Regression as a way to more adequately model the declines in language over time.
+We evaluate this method using the unscripted question and answer portions of news conferences from three U.S Presidents: Ronald Reagan (RR), who would later go on to receive a diagnosis of Alzheimer's Disease; George H.
+W.
+Bush (GB) and Donald Trump (DT), who have no known diagnosis of dementia.
+Our analysis uncovered several lexical features that did not change significantly for GB and DT, but presented significant changes for RR, including significant decreases in unique words and noun usage, and significant increases in pronouns (particularly personal pronouns), conjunctions and words which are related to social processes significantly increase over time.
+These findings are consistent with the psychological literature, which documents a link between these features and cognitive decline in Alzheimer's Disease.
+This work points the way to developing a comparatively non-invasive protocol that could be used in clinical settings that help to spot atypical declines in language use over time, potentially contributing to earlier, more accurate diagnoses of Alzheimer's Disease.
+
 \end{abstract}

### 2019-11-13 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
updated abstract
updated discussion

@@
 
-\title{Exploring the Presidents Corpus}
+\title{Using language as a indicator of cognitive decline in pre-clinical Alheimer's Disease}
 
@@
 This work points the way to developing a comparatively non-invasive protocol that could be used in clinical settings that help to spot atypical declines in language use over time, potentially contributing to earlier, more accurate diagnoses of Alzheimer's Disease.
-
 \end{abstract}
@@
 One shows a model fit to a linear  model, and the other shows a model fit using local polynomial regression model which aims to model the data more closely.
+It is clear to see that both models can track the movements of language use over time however, the linear model is quite a rigid model which does not allow for different inflection points.
+It is not the case that language declines in a linear way, as demonstrated by figure  and whilst we have a number of data points that we can use to model RR's decline, in a clinical setting we will often not have this luxury.
+\par
 
@@
 \par 
-The results of this work also identify some clues as to what
+The results of this work also identify some clues as to what could work in a clinical setting, or how we might be able to collect data in a way that accurately tracks language decline without making incorrect assumptions.
 

### 2019-11-25 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
added to PresidentsCorpus

--- 2019-11-07 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
@@
 Bush (GB) and Donald Trump (DT), who have no known diagnosis of dementia.
-Our analysis uncovered several lexical features that did not change significantly for GB and DT, but presented significant changes for RR, including significant decreases in unique words and noun usage, and significant increases in pronouns (particularly personal pronouns), conjunctions and words which are related to social processes significantly increase over time.
+Our controls were the closest in  age to RR when they became president.
+Our analysis uncovered several lexical features that did not change significantly for GB and DT, but presented significant changes for RR, including significant decreases in unique words and noun usage, and significant increases in pronouns, conjunctions and words which are related to social processes significantly increase over time.
+We also show that non-linear models provide a more appropriate way to model these changes over time.
 These findings are consistent with the psychological literature, which documents a link between these features and cognitive decline in Alzheimer's Disease.
@@
 \section{Conclusions}\label{conclusions}
-The results of this work show that we can track a person's use of language through time in a number of ways, and it is possible for an individual to be his or her own control.
-This is important as it means the heterogenous nature of the MCI population does not impact results as much as if we were comparing those with MCI to controls.
+The results of this work show that we can track a person's use of language through time in a number of ways and that it is possible for an individual to be his or her own control.
+This is important as it means the heterogenous nature of the MCI population does not impact results as much as if we were comparing those with group of MCI patients with a group of controls.
 Equally, it would be helpful to have controls to ascertain what would be usual to expect in the decline of language in a healthy older adult.

### 2019-11-25 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
added to PresidentsCorpus

@@
 \par 
-With further work, it is not feasible to the vast array of samples over a timeframe, as we have had with the president corpus and so it would be worth exploring how the quality of these predictions might lessen when faced with considerably fewer samples and over a smaller time period.
+With further work, it is not feasible to the vast array of samples over a timeframe, as we have had with the president corpus and so it would be worth exploring how the quality of these predictions may lessen when faced with considerably fewer samples and over a smaller time period.
 It would also be worth extending this research further to encompass more of the linguistic features Fraser used in her work \cite{Fraser2015} to see if there are any further insights to be gained.
 In addition, this replication and extension has demonstrated the potential utility of using longitudinal data as a means of comparing language use of a person at two or more time periods and using this information as a diagnostic aid for MCI and therefore more work would be helpful from a longitudinal perspective to see if this approach may be valid in moving towards a solution for this particular problem.
+\par 
+Given the findings of this particular study, a natural progression would be to explore whether smaller numbers of transcripts and with smaller number of words in each transcript would be just as robust in detecting changes in language as a clinical setting would not have the luxury of collecting similar quantities of transcripts at the quality we have for this dataset.
 
@@
 \par 
-The results of this work also identify some clues as to what
-
+The results of this work also identify some clues as to what could work in a clinical setting, or how we might be able to collect data in a way that accurately tracks language decline without making incorrect assumptions.
+\par
+From a clinical perspective, we can see that using samples (albeit for this dataset) is able to track languages changes over a given time frame.
+This potentially means that we can use a similar methodology to collect regular language samples in settings such as memory clinics and GP's surgeries and even potentially in the people's homes and that these language samples, have the potential to act as an early warning sign for Mild Cognitive Impairment that potentially will identify patients at risk of developing Alzheimer's Disease.
 \bibliographystyle{ieeetr}

### 2019-12-03 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
organised paper 1

@@
 
-\title{Exploring the Presidents Corpus}
+\title{Using language as an indicator of cognitive decline in pre-clinical Alzheimer's Disease.}
 
@@
 	\texttt{p.sawyer@aston.ac.uk} \\
-		\AND
+		\and
 	George Vogiatzis\\
@@
 }
-
 \date{\today}
 
-\usepackage{arxiv}
 \usepackage{float}
 \usepackage{graphicx}
+\usepackage{color}
+
 \begin{document}
-\maketitle
 
-\keywords{Mild Cognitive Impairment \and Alzheimer's Disease \and Machine Learning \and Diagnosis \and Natural Language Processing} 
 \bigskip
@@
 
+\maketitle
 \section{Introduction}\label{introduction}
-The diagnosis of dementia is generally made when there is a decline in brain function due to physical changes in the brain \cite{Albert2011}.
+A diagnosis of dementia is generally made when there is a decline in brain function due to physical changes in the brain \cite{Albert2011}.
 It affects a significant proportion of the global older adult population and the impact on morbidity and mortality rates is considerable.
 Dementia is currently the leading cause of death in England and Wales and the sixth leading cause of death in the United States (US).
-A 2014 report commissioned by the Alzheimer's Society estimated that in the UK by 2015 there would be approximately 855,000 people rising living with dementia increasing to 1 million by 2021 \cite{}.
+A 2014 report commissioned by the Alzheimer's Society estimated that in the UK by 2015 there would be approximately 855,000 people rising living with dementia increasing to 1 million by 2021 \cite{AlzheimersSociety2014}.
 This represents 1 in 79 of the total UK population rising to 1 in 14 of those aged 65 or over \cite{AlzheimersSociety2014}.
@@
 Globally, this picture is a lot bleaker.
-The worldwide cost of dementia in 2018 was estimated to be in the region of one trillon US dollars \cite{}.
+The worldwide cost of dementia in 2018 was estimated to be in the region of one trillon US dollars \cite{Prince2015}.
 \par 
 There are different types of dementia including Alzheimer's Disease (AD), vascular dementia, dementia with Lewy bodies and fronto-temporal dementia, all of which are currently incurable.
-AD is a progressive neurodegenerative disease and is the most common type of dementia, responsible for approximately XX\% of all cases \cite{}.
+AD is a progressive neurodegenerative disease and is the most common type of dementia, responsible for approximately 60\% to 80\% of all cases \cite{Duong2017}.
 Currently, a definitive diagnosis for AD can only be produced at post-mortem.
@@
 At a certain point, the levels of beta-amyloid rise triggering a more rapid spread of tau throughout the brain.
-Eventually, due to these and other changes, neurons lose their ability to communicate and the brain starts to shrink.
+Eventually, due to these and other changes, neurons lose their ability to communicate and the brain starts to shrink\cite{Duong2017}.
 This leads to the development of some of the psychological symptoms associated with AD, primarily cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}.
@@
 There has been a significant research in the area of language deterioration as a method of detecting AD at an earlier stage.
-This usually takes the form of recording speech whilst patients' undertake a cognitive assessment such as the Picture Description Task \cite{Orimaye2014,Fraser2015}.
+This usually takes the form of recording speech whilst patients' undertake a cognitive assessment such as the Picture Description Task \cite{Fraser2015}.
 Given that language samples are relatively easy to collect, research has moved towards analysis of spontaneous speech.
@@
 In order to maintain consistency when parsing through the documents, We have changed both types of hesitation to be marked by a single hyphen.
-We also omitted one word sentences as this data would, from a theoretical perspective, not be relevant for language analysis.
-We did not control for the length of the document, but generated features which would normalise by the length of the document.
+We also omitted one word answers to questions as this data would, from a theoretical perspective, not be relevant for language analysis.
+We did not make any alterations to the length of the document, but instead generated features which would normalise by the length of the document.
 We therefore was able to include all press conferences by all presidents analysed where there was a question and answer session conducted at least in part by the president in question (2 press conferences of GHWB were omitted due to a lack of a question and answer session).
@@
 \subsection{Feature Generation}
-We calculated the following features for each transcript in turn using the Natural Language ToolKit (NLTK) package \cite{Bird2009}, Linguistic Inquiry and Word Count (LIWC) \cite{Pennebaker2015} and Python.
-Words were stemmed to their root form using the Porter2 (Snowball) stemmer algorithm.
+Features were generated by running each transcript through a number of natural language processing libraries.
+We used the Natural Language ToolKit (NLTK) package \cite{Bird2009} in python to stem each transcript using the Snowball Stemmer and also completed a part of speech (POS) tagging process.
+We also ran each transcript through Linguistic Inquiry and Word Count (LIWC) \cite{Pennebaker2015} software.
+We completed all NLP tasks in Python and completed all statistical analysis in R..
 \par 
@@
 \begin{equation} \label{x3}
-%% Check to see if this is accurate.
 R = (100 \log N) / (1 - V1/V)
@@
 \par
-We looked at 
-\par 
-In order to compare the fit of linear models with
+We looked also compared the results of linear models and a non linear model in terms of best fit.
+The non linear model we used was the generalised additive model \cite{Hastie1986}.
+The selection of this non linear model was due to the authors not wanting to assume an underlying distribution of the data and therefore this was the most appropriate model to use.
 
@@
 
-In terms of more specific language differences between the presidents, we found that RR used significantly more unique words, non-specific nouns and low imageability verbs than GHWB and DJT (see Table 3.3).
+In terms of more specific language differences between the presidents, we found that RR used significantly more unique words, non-specific nouns and low imageability verbs than GHWB and DJT (see Table 3).
 The mean length of utterance for RR was significantly greater than that of GHWB and DJT.
@@
 	\begin{center}
-	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+	\begin{tabular}{ | p{3cm} | p{2.2cm} | p{2.2cm} | p{2.2cm} |}
 		\hline
@@
 \subsection{Longitudinal Analysis}
-We then looked at the data from a longitudinal perspective as we are interested seeing whether we can track various language variables and their progress over time.
-We ran a number of Pearsons correlations with transcript index number as a time reference and the dependant variables (Table 3.4).
+We then looked at the data from a longitudinal perspective as we were interested seeing whether we can track various language variables and their progress over time.
+We ran a number of Pearsons correlations with transcript index number as a time reference and the dependant variables (Table 4).
 For our controls, we found them to be stable for the most part with the main highlights being a decrease in Adverb usage for DJT (R=-0.36, p=0.049) and a steady but not severe decline in a number of variables for GWHB, namely total word count, unique words, low imageability words and verb usage.
@@
 
-Given the number of hypotheses being tested, it is necessary to apply some correction for family wise error rate and therefore a Hochberg step-up procedure was applied to the results of the analysis.
+Given the number of hypotheses being tested, it was necessary to apply some correction for family wise error rate and therefore a Hochberg step-up procedure was applied to the results of the analysis.
 We also used the Benjamini-Yekutieli procedure for controlling for False Discovery rate and came up with a final list of important features for each president.
@@
 The features were generated from multiple sources and therefore there was some significant overlap between the features.
-In these cases, the normalised features were prefered to raw count features.
+In these cases, the normalised features were preferred to raw count features.
 Where there was doubt as to the most appropriate feature, all features have been included.
@@
     	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
-
 	\end{tabular}
@@
 
+In terms of comparing the general additive model and linear models.
+We calculated the predicted residual error sum of squares (PRESS) statistic for each feature is a form of cross-validation used in regression analysis to provide a summary measure of the fit of a model to a sample of observations that were not themselves used to estimate the model.
+There were no significant differences (paired t-test: t=1.8875, df  = 21, p-value = 0.07299.
+
 \begin{table}[H]
 	\begin{center}
-	\begin{tabular}{ | l | p{2.5cm} | p{2.5cm} | p{2.5cm} |}
+	\begin{tabular}{ | p{4cm} | p{3.2cm} | p{3.2cm} |}
 		\hline
-		& RR R-Squared  & GWB R-Squared & DJT R-Squared \\ \hline
-		ppron & \textbf{0.700***} & 0.2 & 0.35 \\ \hline
-		social & \textbf{0.698***} & 0.2 & 0.35 \\ \hline
-		NounsNormalised & \textbf{-0.689***} & 0.2 & 0.35 \\ \hline
-		function & \textbf{0.670***} & 0.2 & 0.35 \\ \hline
-		conj &\textbf{0.644***} & 0.2 & 0.35 \\ \hline
-		PronounsNormalised & \textbf{0.631***} & 0.2 & 0.35 \\ \hline
-		Analytic & \textbf{-0.626***} & 0.2 & 0.35 \\ \hline
-		NN & \textbf{-0.601***} & 0.2 & 0.35 \\ \hline
-		male & \textbf{0.585***} & 0.2 & 0.35 \\ \hline
-		UniqueWords & \textbf{-0.578***} & 0.2 & 0.35 \\ \hline
-		WDT & \textbf{-0.577***} & 0.2 & 0.35 \\ \hline
-		shehe & \textbf{0.529*} & 0.2 & 0.35 \\ \hline
-		VBZ & \textbf{-0.521*} & 0.2 & 0.35 \\ \hline
-		JJ & \textbf{-0.518*} & 0.2 & 0.35 \\ \hline
-		Fillers & \textbf{-0.518*} & \textbf{-0.518*}  & 0.35 \\ \hline	
-		EX & \textbf{-0.518*} & \textbf{-0.518*}  & 0.35 \\ \hline	
-		achieve & \textbf{-0.518*} & \textbf{-0.518*}  & 0.35 \\ \hline		
-		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
-    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
-    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
-
+		Feature & GAM Press Statistic & LM Press Statistic \\ \hline
+		ppron & 18.591104613 & 20.001055542 \\ \hline
+		social & 21.681732112 & 24.906646115  \\ \hline
+		NounsNormalised & 0.002818941 & 0.002797798\\ \hline
+		function & 46.546071705 & 46.031993602 \\ \hline
+		conj & 8.812330188 & 8.354974826 \\ \hline
+		PronounsNormalised & 0.002284437 & 0.002356129\\ \hline
+		Analytic & 2272.200519272 & 2239.224863610 \\ \hline
+		Pronoun & 44.557282749 & 46.894576916 \\ \hline		
+		NN & 101685.870974277 & 96485.575598505 \\ \hline
+		male & 5.191815443 & 5.278559152 \\ \hline
+		UniqueWords & 255114.636660673 & 238042.427437454 \\ \hline
+		WDT & 1987.149134694 & 1958.781236655 \\ \hline
+		nouns & 276723.222762842 & 262192.120091097 \\ \hline
+		nouns/100 & 27.670762078 & 26.219212009 \\ \hline
+		UniqueStems & 170450.645283172 & 166302.725530118 \\ \hline
+		shehe & 5.188136314 & 5.265385185 \\ \hline
+		VBZ & 13889.871831536 & 13721.175422577 \\ \hline
+		JJ & 24950.440222526 & 24692.043386181 \\ \hline
+		article & 13.500676675 & 13.468083961 \\ \hline	
+		Adjectives & 30210.006270270 & 29865.240707991 \\ \hline	
+		Adjectives.100 & 3.020898615 & 2.986524071 \\ \hline
+		Dic & 40.099289973 & 38.501060715 \\ \hline	
 	\end{tabular}
-	\caption{\label{tab:table-name}Pearson Correlations for Features}
+	\caption{\label{tab:table-name}Comparison of GAM and Linear Model using the PRESS statistic}
 	\end{center} 
@@
 	\begin{minipage}[b]{.45\textwidth}
-		\includegraphics[width=6cm, height=5cm]{comp1.png}
+		\includegraphics[width=6cm, height=5cm]{plots/comp1.png}
 	\end{minipage}
@@
 		\begin{minipage}[b]{.45\textwidth}
-		\includegraphics[width=6cm, height=5cm]{comp2.png}
+		\includegraphics[width=6cm, height=5cm]{plots/comp2.png}
 	\end{minipage}
@@
 \bibliographystyle{ieeetr}
-\bibliography{main3}
+\bibliography{main}
 

### 2019-12-04 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
updated first draft

@@
 To do this we were able to split each transcript into individual words and changed them to lowercase using NLTK and were then count the number of unique words that appeared in each transcript.
-We also used the TTR formula, Brunet's Index and Honore's Statistic as other measures of lexical diversity \cite{Richards1987}.
+We also used the TTR formula, Brunet's Index and Honore's Statistic as other measures of lexical diversity \cite{Bucks2000}.
 \par 
@@
 Non-Specific Nouns refer to a category or an unspecified member of a given category, once again this can be characterised as a compensatory strategy for word finding difficulties.
-Imageability is characterized, according to Berisha and Liss, as the ease with which a word provokes a mental image of what the word describes.
+Imageability is characterized, according to Berisha and Liss\cite{Berisha2015}, as the ease with which a word provokes a mental image of what the word describes.
 
@@
 \subsubsection{Longitudinal Analysis}
-As with Berisha and Liss, we analysed the transcripts over time in order to determine if there was an underlying temporal trend that may indicate that different parts of language increase or decrease over time.
+As with Berisha and Liss\cite{Berisha2015}, we analysed the transcripts over time in order to determine if there was an underlying temporal trend that may indicate that different parts of language increase or decrease over time.
 In AD, cognitive abilities are said to deteriorate over time and therefore it may be useful to analyse temporal trends that indicate a potential deterioration of cognition.
@@
 
-Given the number of hypotheses being tested, it was necessary to apply some correction for family wise error rate and therefore a Hochberg step-up procedure was applied to the results of the analysis.
+To get a better idea of decline over time, we also ran Pearson's correlations using time elapsed in days from the first transcript instead a transcript index.
+This was to control for variation in the number of days between transcripts.
+Given the number of hypotheses being tested, we felt it was necessary to apply some correction for family wise error rate and therefore a Hochberg step-up procedure was applied to the results of the analysis.
 We also used the Benjamini-Yekutieli procedure for controlling for False Discovery rate and came up with a final list of important features for each president.
 For RR, this was 22 features, for GHWB this was 14 features and there were no significant features for DJT after applying these methods.
-The features were generated from multiple sources and therefore there was some significant overlap between the features.
-In these cases, the normalised features were preferred to raw count features.
-Where there was doubt as to the most appropriate feature, all features have been included.
+The features were generated from multiple sources and therefore there was some significant overlap between the features but we have included all features in Table 6.
 
@@
 	\begin{center}
-	\begin{tabular}{ | l | p{2.5cm} | p{2.5cm} | p{2.5cm} |}
+	\begin{tabular}{ | l | c | c | c |}
 		\hline
 		& RR R-Squared  & GWB R-Squared & DJT R-Squared \\ \hline
-		ppron & \textbf{0.700***} & 0.2 & 0.35 \\ \hline
-		social & \textbf{0.698***} & 0.2 & 0.35 \\ \hline
-		NounsNormalised & \textbf{-0.689***} & 0.2 & 0.35 \\ \hline
-		function & \textbf{0.670***} & 0.2 & 0.35 \\ \hline
-		conj &\textbf{0.644***} & 0.2 & 0.35 \\ \hline
-		PronounsNormalised & \textbf{0.631***} & 0.2 & 0.35 \\ \hline
-		Analytic & \textbf{-0.626***} & 0.2 & 0.35 \\ \hline
-		NN & \textbf{-0.601***} & 0.2 & 0.35 \\ \hline
-		male & \textbf{0.585***} & 0.2 & 0.35 \\ \hline
-		UniqueWords & \textbf{-0.578***} & 0.2 & 0.35 \\ \hline
-		WDT & \textbf{-0.577***} & 0.2 & 0.35 \\ \hline
-		shehe & \textbf{0.529*} & 0.2 & 0.35 \\ \hline
-		VBZ & \textbf{-0.521*} & 0.2 & 0.35 \\ \hline
-		JJ & \textbf{-0.518*} & 0.2 & 0.35 \\ \hline
-		Fillers & \textbf{-0.518*} & \textbf{-0.518*}  & 0.35 \\ \hline	
-		EX & \textbf{-0.518*} & \textbf{-0.518*}  & 0.35 \\ \hline	
-		achieve & \textbf{-0.518*} & \textbf{-0.518*}  & 0.35 \\ \hline		
+		ppron & \textbf{0.700***} & 0.007 & -0.196 \\ \hline
+		social & \textbf{0.698***} & 0.204 & 0.032 \\ \hline
+		NounsNormalised & \textbf{-0.689***} & -0.023 & 0.194 \\ \hline
+		function & \textbf{0.670***} & -0.169 & -0.243 \\ \hline
+		conj &\textbf{0.644***} & -0.452 & -0.187 \\ \hline
+		PronounsNormalised & \textbf{0.631***} & 0.131 & 0.002 \\ \hline
+		Analytic & \textbf{-0.626***} & -0.013 & 0.366 \\ \hline
+		NN & \textbf{-0.601***} & -0.197 & 0.116 \\ \hline
+		male & \textbf{0.585***} & 0.024 & 0.042 \\ \hline
+		UniqueWords & \textbf{-0.578***} & -0.257 & 0.204 \\ \hline
+		WDT & \textbf{-0.577***} & -0.173 & -0.105 \\ \hline
+		shehe & \textbf{0.529*} & -0.005 & 0.064 \\ \hline
+		VBZ & \textbf{-0.521*} & -0.157 & 0.136 \\ \hline
+		JJ & \textbf{-0.518*} & -0.206 & 0.062 \\ \hline
+		Fillers & -0.076 & \textbf{-0.359**}  & -0.032 \\ \hline	
+		EX & -0.193* & \textbf{-0.337**}  & 0.339 \\ \hline	
+		achieve & -0.248 & \textbf{0.334**}  & 0.169 \\ \hline		
 		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
@@
 We calculated the predicted residual error sum of squares (PRESS) statistic for each feature is a form of cross-validation used in regression analysis to provide a summary measure of the fit of a model to a sample of observations that were not themselves used to estimate the model.
-There were no significant differences (paired t-test: t=1.8875, df  = 21, p-value = 0.07299.
+There were no significant differences (paired t-test: t=1.8875, df  = 21, p-value = 0.07299).
 
@@
 	\begin{center}
-	\begin{tabular}{ | p{4cm} | p{3.2cm} | p{3.2cm} |}
+	\begin{tabular}{ | l | c | c |}
 		\hline
@@
 Compared to our controls (GWHB and DJT), we find some slight trends with GWHB but no such trends with DJT in his speech albeit his samples of speech span a shorter amount of time.
-\par 
-A criticism of Berisha and Liss's work is the problems they had with normalising the transcripts in terms of length.
-This was also a problem in the work of Garrard et al \cite{Garrard2005, Le2011}.
-Whilst it is important to control for outliers, there are other ways in which we can control for length of sample.
-In this paper, we controlled for transcript length by dividing any features that were raw counts by the total length of the transcript.
-Given the increased amounts of hypothesis testing, it became necessary to control for false discoveries.
-\par 
 Interestingly, when we normalised the various types of words used by the presidents we found some interesting patterns that further differentiated RR from the controls.
@@
 Our results are supported by existing literature and this potentially means that language analysis in the way we have structured it may have diagnostic or prognostic properties.
-One of the most striking differences between this study and the original is when looking at Unique Words.
-Whilst Unique Words as a feature is significant, which was a finding in the original study, when we controlled via normalisation, we found that this was not significant.
 \par 
-Another observation is that when conducting fitting a linear model to this data does not tell the whole story.
-For example, consider the following two plots.
-One shows a model fit to a linear  model, and the other shows a model fit using local polynomial regression model which aims to model the data more closely.
+A criticism of Berisha and Liss's work is the problems they had with normalising the transcripts in terms of length.
+This was also a problem in the work of Garrard et al \cite{Garrard2005, Le2011}.
+Whilst it is important to control for outliers, there are other ways in which we can control for length of sample.
+In this paper, we controlled for transcript length by dividing any features that were raw counts by the total length of the transcript.
+When we did this, we found that there was a significant decrease in the number of unique words (R=-0.56, p\textless0.001) used however when we controlled via normalisation, we found that this was not a significant feature (R=-0.172, p=0.25).
+\par 
+Overall, we are able to show differences in the a number of language features between RR, GHWB and DJT over time and the psychological literature confirms our findings in reference to these changes.
+\par
+Our next hypothesis involved exploring whether linear models were the most appropriate way to track this longitudinal data.
+In our analysis, we found there was no significant difference in terms of mean sum of squares errors between the linear model and the generalised additive model however this does not mean that these are equivalent.
+For example, considering the plot below.
+We compare a model fit to a linear model with a generalised additive model which aims to model the data more closely.
 It is clear to see that both models can track the movements of language use over time however, the linear model is quite a rigid model which does not allow for different inflection points.
-It is not the case that language declines in a linear way, as demonstrated by figure  and whilst we have a number of data points that we can use to model RR's decline, in a clinical setting we will often not have this luxury.
+We can see that RR's decline is not linear, and that there are periods of relative stability and some periods where the decline is more severe.
+We argue that these more nuanced changes over time can only be modelled by a model such as a GAM.
 \par
-
-\begin{figure}
+\begin{figure}[H]
 	\centering
-	\begin{minipage}[b]{.45\textwidth}
-		\includegraphics[width=6cm, height=5cm]{plots/comp1.png}
-	\end{minipage}
-	\hfill
-		\begin{minipage}[b]{.45\textwidth}
-		\includegraphics[width=6cm, height=5cm]{plots/comp2.png}
-	\end{minipage}
+	\includegraphics[width=6cm, height=5cm]{plots/comparison1.png}
+	\caption{Comparing a linear model with a generalised additive model}
 \end{figure}
 
+In her paper on developing diagnositic criteria for Mild Cognitive Impairment, Albert et all \cite{Albert2011} states 'it is important to obtain longitudinal assessments of cognition, whenever possible' and 'obtaining objective evidence of progressive declines in cognition over time is important for establishing the accuracy of the diagnosis, as well as for
+assessing any potential treatment response.' As we have shown, it is not the case that language declines in a linear way and whilst we have a number of data points that we can use to model RR's decline, in a clinical setting it would be impractical to have upwards of 40 data collection points.
+We feel that  it is important that we have more than just two.
+In the figures below, we can see that if we measured language at Reagan's first transcript to a point after 700 days the vast majority of points would point to a decline in language.
+However, the next figure shows what would happen if we measured language from Reagan's second transcript, just 34 days later.
+In this case, we can see that in some cases we would mark Reagan as improving when it is clear that he is not.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=6cm, height=5cm]{plots/comp1.png}
+	\caption{Measuring decline from Reagan's first transcript to all other transcripts greater than 700 days later}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=6cm, height=5cm]{plots/comp2.png}
+	\caption{Measuring decline from Reagan's first transcript to all other transcripts greater than 700 days later}
+\end{figure}
+\par 
 There are limitations of this research.
 Whilst in terms of age, DJT is certainly more suitable as a control to match with RR, in some ways they held very different styles of press conferences in that RR preferred to do solo press conferences and DJT has shown a preference for doing joint press conferences which have an impact on the amount of language produced.
-This artifact of the data is in itself notable as it illustrates the problems we may have with smaller amounts of speech.
-Also, the problem of finding an appropriate control is a common one in this domain.
-This variability in the number and quality of the transcriptions raises some doubts as to the results of DJT.
+This artefact of the data is in itself notable as it illustrates the problems we may have with smaller amounts of speech and the problem of finding an appropriate control is a common one in this domain, given the considerable variation in factors such as age, language ability and education.
+As mentioned above, we are lucky enough to have numerous samples of data collected over a long period of time and this is not something that can be easily translated to a clinical protocol.
+Finally, the variability in the number and quality of the transcriptions raises some doubts as to the results of DJT.
 Is it the lack of instances of data, or that the number of words per transcript is significantly less than with RR or GHWB.
-This does have further implications for using a protocol such as this in a more general way, as it will impact how data is collected.
-\par
-
-Given that those with MCI and early dementia have such variable presentations, it might prove of limited value in matched pairs design. 
+This does have further implications for using a protocol such as this in a more general way as it will impact how data is collected.
 \par 
@@
 In addition, this replication and extension has demonstrated the potential utility of using longitudinal data as a means of comparing language use of a person at two or more time periods and using this information as a diagnostic aid for MCI and therefore more work would be helpful from a longitudinal perspective to see if this approach may be valid in moving towards a solution for this particular problem.
-\par 
-Given the findings of this particular study, a natural progression would be to explore whether smaller numbers of transcripts and with smaller number of words in each transcript would be just as robust in detecting changes in language as a clinical setting would not have the luxury of collecting similar quantities of transcripts at the quality we have for this dataset.
 

### 2019-12-05 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
updated Pres

@@
 Whilst Non-specific nouns increased over time, we found that noun usage in general significantly decreased and pronouns increased similarly significantly.
-The increase in pronoun for those with early AD has been identified in literature, although there are only a few studies that explore this \cite{Almor1999, Wendelstein2015}.
+The increase in pronoun for those with early AD has been identified in literature, although there are only a few studies that explore this \cite{Wendelstein2015}.
 Wendlestein et al propose that the increased used of pronouns is an expression of an impaired ability to adapt language to the listener's needs \cite{Wendelstein2015}.

### 2019-12-06 Source/Paper 3 - Delphi Review/DelphiEmail.tex
update Thesis and Delphi Questionnaire

\documentclass{article}

\begin{document}

Dear XXXXXX,

Thank you for agreeing to take part in this Delphi study which will form part of my Ph.D research project. Your participation will be completely confidential and you will remain completely anonymous throughout this process. It is intended that the findings of this study will submitted for publication.

My Ph.D project aims to explore the potential for language to aid in the early detection of MCI that progresses to dementia. As part of this process, I have identified a need for an smartphone application and associated protocol to aid in the collection of language samples. It is therefore the goal of this study to identify what the application might look like and the best way to collect the data. The results of this Delphi study will inform a pilot study that will test the protocol created by this study.

The following survey is stage 1 of a Delphi questionnaire. This is designed to obtain your personal opinions relating to the problem described above. The Delphi process involves obtaining responses from you on three separate occasions and the we estimate that it should take no more than 30 minutes to respond:

Round One: Some general open ended questions will be submitted to you requiring your response. These are below for you to reply to now.

Round 2: Your answers (and those from the other panellists) from round 1 will be summarised and formulated into a series of more specific questions that you will be asked to respond to.

Round 3: Round 2’s questions will be submitted to you again but this time you will also be able to see the average reply of the other panellists and you will then be asked if you would like to adjust your answer from the second round or not. The identity of all panelists will remain confidential at all times.

The 4 questions listed below are designed to seek your personal opinion. Please reply to each one but please do not feel limited in the length or style of your answers. A reply to these by email is fine.

1) It is estimated that between five and ten minutes of spontaneous speech would be an ideal balance between data quality and ease of providing a sample. 
	a) Do you agree with this? if not, please state why?
	b) What task(s) do you think would be useful for eliciting an appropriate amount of speech?
	c) How often should data be collected that will provide a balance between accurate tracking of language use over time whilst reducing intrusiveness?

2) The aim of the end application is to provide a way for users to provide multiple samples over a period of time, however as with all longitudinal data collection methods will be subject to participant attrition.
	a) In your opinion, how might this application encourage continued submissions of language samples?

3) The aim of creating this mobile application is to facilitate the ease in which a large number of samples can be collected? 
	a) In your opinion, what problems do you foresee happening with this collection method? 
	b) Are there any potential solutions to these problems that spring to mind?

Thanks in advance for your time,

Jomar Alcantara
Aston University
\end{document}

### 2019-12-06 Source/Thesis/Chapter2.tex
update Thesis and Delphi Questionnaire

@@
 A review by Boschi et al concludes that analysis of connected speech is potentially useful in guiding clinicians to identify language disorders \cite{Boschi2017} and also highlights the role of NLP and Machine learning in assisting in this endeavour.
-\par
 \subsection{Semantics vs Pragmatics}
@@
 Whilst the study of language from a pragmatic perspective is much more complex, it is perhaps one of the most vital areas to study because of the number of different cognitive processes involved.
-\par
 \subsection{Semantic Content}
@@
 This area of research is extremely promising in its early results and the impact of successful research would be life changing for both individuals and the health of the worlds aging population in general.
+
+\section{Introduction}
+Dementia has been identified as one of the fast growing difficulties facing the world.
+A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
+\par
+A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
+As described above \cite{Prince2015}, the numbers of those suffering from AD and MCI are going to increase as the population ages and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
+At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017}.
+I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
+\par
+There is a large body of research that looks at language deterioration in those suspected to have MCI or AD \cite{Taler2008, Boschi2017}.
+However there is conflicting evidence in these studies about which declining language factors are associated with MCI and AD \cite{Taler2008, Boschi2017}.
+Research therefore, should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
+Another area of focus for research of this nature is the process of collecting appropriate language samples.
+Whilst collecting samples of language is comparatively unintrusive, researchers recognise that these samples require a rich sample of language that potentially cannot be generated by tasks such as the picture description task.
+Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD.
+There is some evidence to support this.
+Berisha et al \cite{Berisha2015}, has shown through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control.
+\par
+The question to be addressed in this systematic review is how has the field of machine learning and natural language processing addresses language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease.
+The potential impact of this research in this area is immense.
+Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease however the absence of a single test and the complexity of AD can create significant delays in diagnosis.
+Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
+\par 
+The remainder of this article is organized as follows.
+Section~\ref{methodology} gives an account of the process of this Systematic Review.
+Our results are described in Section~\ref{results}.
+We discuss the results and implications in Section~\ref{discussion}.
+Finally, Section~\ref{conclusions} gives the conclusions.
+
+\section{Methodology}\label{methodology}
+A systematic literature review (SLR) describes a process which aims to identify, evaluate and interpret the research and literature in a given area.
+They are designed to provide a complete and exhaustive summary of the current evidence relevant to an identified research question.
+SLR's conduct a thorough search of all literature following a pre-defined protocol that specifies focused research questions, identifies criteria for the selection of studies and assessment of their quality, and forms to execute the data extraction and synthesis of results.
+\par 
+Common motivations for conducting an SLR are: 
+\begin{enumerate}
+	\item to summary all the evidence about a topic.
+	\item find gaps in the research. 
+	\item to provide a ground for a fundament to new research.
+	\item and to examine how the current research supports a hypothesis. 
+\end{enumerate}
+
+Performing an SLR comprises the following steps:
+
+\begin{enumerate}
+	\item identify the need for performing the SLR.
+	\item formulate research questions.
+	\item execute a comprehensive search and selection of primary studies.
+	\item assess the quality and extract data from the studies.
+	\item interpret the results.
+	\item report the SLR.
+\end{enumerate}
+
+\subsection{Search strategy}
+The main research question this SLR aims to address is: “How has the field of machine learning and natural language processing addressed language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease?”.
+
+This SLR builds upon these questions and additionally presents the results of the other two additional research questions.
+Further, the key terms related to MCI and the other names by which it is known were included in the search string to ensure that relevant studies about machine learning and the diagnosis of a disease were also retrieved, even if not specifically mentioned in the paper's title or abstract.
+\par 
+To address the research questions, a search string was defined using the PICO approach, which decomposes the main research question into four parts: 
+\begin{enumerate}
+	\item Population - Studies that present research on mild cognitive impairment and dementia.
+Mild Cognitive Impairment and Dementia keywords were selected from the Systematized Nomenclature of Medicine–Clinical Terms (SNOMED-CT).
+	\item Intervention - Intervention: ML or Statistical Modelling techniques that focus on classification.
+The ML keywords were selected from the branch “Machine Learning Approaches” of the “2012 ACM Computing Classification System”.
+The MS keywords were selected by A2.
+	\item Comparison
+	\item Outcome - Outcome: Prognosis on dementia and comorbidies. The prognosis keywords were provided
+by A4.
+\end{enumerate}
+
+Using this process, the main research question was decomposed  into four research questions:
+\begin{itemize}
+	\item RQ1: What features / data characteristics of text (variables, determinants and indicators) that are considered when applying the ML techniques (n-grams, PoS Tagging etc)?
+	\item RQ2: Which NLP and ML or Statistical Learning techniques are being used in dementia research?
+	\item RQ3: What are the goals of the studies that employ NLP / ML techniques for prognosis of dementia?
+	\item RQ4: Do the studies focus on time as factor?
+\end{itemize}
+\par 
+The automated searches were performed in the Pubmed, Web of Science, Scopus and IEEE databases.
+Table 1 shows the search string used for the Pubmed automated search, but note that this search string was adapted to each of the other databases’ search context.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{12cm} | }
+	\hline
+	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+	Searched - 4th April 2019 - Generated 1257 Articles \\
+	\hline
+	\end{tabular}
+	\end{center}
+	\caption[Table caption text]{Example of Search Terms for Web of Science database}
+	\label{table:name}
+\end{table}
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{3cm} | p{3cm} | p{3cm} | }
+	\hline
+	Scopus & Web of Science & PubMed & IEEE Xplore \\ \hline
+	1002 & 991 & 376 & 230 \\ 
+	\hline
+	\end{tabular}
+	\end{center}
+	\caption[Table caption text]{Number of articles found from each database - complete search terms appear in Appendix B}
+	\label{table:name}
+\end{table}
+
+\subsection{Study selection} 
+A total of 1490 unique papers were identified through the searches conducted above.
+Each paper was initially reviewed by just the title and the abstract.
+This yielded a total of 25 potential papers.
+Papers were then evaluated by the JA based on the inclusion and exclusion criteria  (see Table 3).
+Where JA could not reach a decision, PS and GV were consulted and majority vote on inclusion was conducted.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{6cm} | p{6cm} | }
+	\hline
+	Inclusion Criteria & Exclusion Criteria \\ \hline
+	Be a primary study in English; AND address research on dementia and comorbidities; AND address at least one ML or MS technique; AND address a prognosis related to dementia and comorbidities; AND use cognition or language decline as a factor for analysis.
+&
+	Be a secondary or tertiary study; OR be written in another language other than English; OR do not address a research on dementia and comorbidities; OR do not address at least one ML or MS technique; OR do not address a prognosis related to dementia and comorbidities.
+\\
+	\hline 
+	\end{tabular}
+	\caption[Table caption text]{Inclusion and Exclusion Criteria}
+	\label{table:name}
+	\end{center}
+\end{table}
+
+After the identification of these papers, a one-iteration backward snowballing process was carried out using the reference lists of the original set of papers looking for studies that were missed in the original searches.
+This resulted in 41 additionally identified papers.
+Throughout the whole selection process, PS and GV acted as additional assessors in the case where there was uncertainty about whether a paper should be included.
+\par 
+In total, 66 papers were selected to be fully read.
+A quality assessment questionnaire (see Table 4) was developed based on Kitchenham's guidelines and was used to minimize the chance of bias in the selection process.
+Studies were graded on a 12 point scale and any studies which scored less than 8 points were excluded for quality reasons.
+The ones that successfully passed the filtering criteria described earlier had their relevant data extracted.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{6cm} | p{6cm} | }
+	\hline
+	Variable & Definition \\ \hline
+	Conditions Studied & For which dementia disorder is the study deriving a prognosis. \\ \hline
+	Database used in the study & Name and origin of the data source used to derive the prognosis of the studied dementia. \\ \hline
+	Dataset Categories & Classes in which the data units were divided into. \\ \hline
+	Follow-up period & Period of time, which the data units were followed. \\ \hline
+	Techniques used & Natural Language techniques AND/OR ML techniques that were used to build the diagnostic models. \\ \hline 
+	Features generated & NLP generated features used in building the diagnostic models. \\ \hline
+	Aim of the Study & The goal of the built diagnostic models. \\
+	\hline 
+	\end{tabular}
+	\caption[Table caption text]{Quality Assessment Questionnaire}
+	\label{table:name}
+	\end{center}
+\end{table}
+
+In this phase, a paper could also be rejected due to inclusion and exclusion criteria because the selection process adopted an inclusive approach.
+This means that during the reading of the titles and abstracts, in the case where the information provided was incomplete or too general it was selected to be fully read in the posterior phase.
+A common example is the case when the data analysis technique specified in the abstract was merely “classification”, so it was not possible to know if any machine learning occurred.
+\par
+% NUMBERS!
+In total, 37 studies composed the final set of included primary studies and had their relevant data extracted, 7 papers were rejected due quality reasons, and 34 papers were rejected due to failing the inclusion and exclusion criteria.
+One reason for the high number of the latter was the decision to exclude the papers that used solely statistical methods as data analysis techniques to build the prognostic models.
+The selected studies were also assessed for the risk of cumulative evidence bias.
+This was done by checking, in the case of the same research group with different studies in the final set of included primary studies, if it was justified having both studies (i.e different samples).
+
+\subsection{Data collection}
+For the data collection, a base extraction form was defined in the protocol, but later in the study it was evolved based on the research group discussions.
+Table 4 lists and defines the collected variables.
+
+In addition to these variables other basic data about the studies was collected, these were: title, authors, journal/source, year and type of publication.
+No summary measures were used.
+Summary tables were used for the synthesis of results and no additional analyses were carried out
+
+\section{Results}\label{results}
+\subsection{Features of Language}
+One of the first steps in building an accurate diagnostic model using language is to analyse the effectiveness of the various language features that are used.
+
+\par
+Asgari, Kaye and Dodge (2017) \cite{Asgari2017} used another form of word frequency measurement.
+Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words \cite{Pennebaker2015}.
+They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
+\subsubsection{Measures of Semantic Complexity}
+Intro!
+
+Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
+\begin{equation} \label{x1}
+TTR = numberOfUniqueWords / totalNumberOfWords.
+\end{equation}
+\par 
+Brunet's Index (W) differentiates itself for TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
+\begin{equation} \label{x2}
+W = N^{V(-0.165)}
+\end{equation}
+where N is the total length of the utterance being measured and V is equal to the total vocabulary being used by the subject.
+Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers.
+\newline
+\par 
+Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation: 
+\begin{equation} \label{x3}
+%% Check to see if this is accurate.
+R = (100 \log N) / (1 - V1/V)
+\end{equation}
+where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
+\par 
+Guinn, Singer and Habash \cite{Guinn2015}, they found in their corpus that there was no significant difference between interviewers and those with dementia when applying these measures.
+However, when they compared these results with a control dataset, they did find a significant difference in Honore's statistic.
+They explained this interesting results by suggesting that the interviewer used was trying to match (intentionally or unintentionally) the lexical richness of the person they were interviewing.
+In comparing healthy older adults with those suffering from dementia, they found that there was an increase in lexical diversity which is contrary to other research.
+They did note that conversations involving those with dementia contained roughtly 50\% less speech than those with controls, and that TTR in particular is not a suitable measure because it is more sensitive to length.
+They also note that Brunet's Index and Honore's Statistic are better statistics that control for the total length of the conversation and controls had statistically greater lexical diversity on those measures.
+
+\subsection{Quantity - Total number of words}
+Several studies report that adults with moderate AD produce fewer words than controls on picture description, however other studies found no differences in total words among groups of controls and patients with MCI or AD.
+Murray and Nicholas et al investigated normal controls, patients with AD and older adults with depression and found no group differences in total words.
+In contrast, Lira 2014 found that controls produced more total words than patients with AD but found no difference between mild and moderate groups.
+\newline
+\par
+\subsection{Syntax and Morphology (Language Form)}
+Syntax can be defined as the rules that govern how words can be combined to form sentences, whilst Morphology is the system that governs the structure of words and the construction of word forms.
+Multiple studies of language decline in dementia included at least one measure of syntax or syntactic complexity.
+Common constructs included words per clause, grammatical form (measures of an appropriate use of syntactic conjections, tenses, conditionals, subordinate clauses and passive constructions), measures of phrase length and proportions of words in sentences.
+Some researchers have explored the use of formulaic language in those with dementia, the theory being that well practiced phrases are less effortful and therefore place low load on the cognitive abilities of those with AD.
+The general hypothesis motivating these studies is that either working memory limitations or semantic memory limitations in AD affect one's ability to use complex constructions.
+\par
+\subsection{N-grams and skip-grams}
+One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
+An n-gram is a contiguous sequence of n items from a given sample of text or speech.
+The items can be phonemes, syllables, letters, words or base pairs according to the application.
+For example, given the sequence of words "to be or not to be", this extract is said to contain six 1-gram sequences (to, be, or, not, to, be), five 2-gram sequences (to be, be or, or not, not to, to be), four 3-gram sequences(to be or, be or not, or not to, not to be) and so on.
+This is useful as, given a large portion of text or speech, we can predict the probability of a word being close by to a given word.
+A number of researchers have used n-grams as features.
+One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class.
+Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls.
+Their main finding supported n-grams as the most significant predictor.
+One of the criticisms is the use of picture description tasks and n-grams.
+Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
+\par
+Skip-grams are a variant of n-grams in which word tokens are skipped intermittently while creating n-grams.
+For example, take the sentence 'I am going to London', there are four conventional bigrams: 'I am', 'am going', 'going to', and 'to London' - using skip-grams, we might skip a word to create additional bigrams such as: 'I going' and 'going London'.
+Orimaye defined k-skip-n-grams as a set of n-gram tokens with the following equation, where n is the specified n-gram (e.g.
+2 for a bigram and 3 for a trigram), m is the number of tokens in a given sentence, k is the number of word skip between n-grams given that k < m and a = \{1,..., m-n\}
+\par 
+One problem with this approach is existing research currently uses language generated from picture description tasks.
+Given the nature of these tasks, the language generated is relatively constrained in comparison to language generated spontaneously.
+% Question How does using ngrams with constrained language differ from ngrams with general speech?
+
+\begin{equation} \label{x4}
+T_{n-gram} = {W_{a},...,W_{a+n-k},..., W_{a+n},....,W_{m-n},...,W_{(m-n)+n-k},...,W_{m}}
+\end{equation}
+
+Thus for the sentence 'I am going to London', 1-skip-2-grams will give \{I going, am to, going London\} and 1-skip-3-grams will give \{I going to, I am to, am to London, am going London\}
+
+\subsection{Mean length of utterance (MLU)}
+Murray found that MLU was not a distinguishing factor among health adults, adults with depression and adults with AD.
+Ripich et al found a decrease in MLU in adults with severe AD over time, and this was supported by findings of Le et al in their studies of authors \cite{Le2012} \newline
+\par 
+\subsection{Proportion of verbs to nouns plus verbs}
+Kave and Levy used a verb index to capture syntactic complexity and found that adults with AD expressed the same amount of verbs to nouns plus verbs as adult controls. \newline
+\par 
+\subsection{Syntactic Complexity - Composite measures of MLU, syntactic errors and verbs}
+Ahmed et al, and Ahmed, Haigh et al found differences in syntactic complexity between adults with MCI and controls, and between MCI and moderate AD stages.
+The differences in syntactic complexity were not significant when individual measures were tested, but were apparent using a composite score consisting of MLU, words in sentences, syntactic errors, nouns with determiners, and verbs with inflections.
+\newline
+\par
+Lu's Syntactic Complexity Analyser
+Ygnve measure
+
+\subsection{Semantic features}.
+We compute semantic similarity using the average and minimum cosine distance between each pair of one-hot embeddings of utterances, and the cosine cutoff (i.e., the number of pairs of utterances whose the cosine distance is below a certain threshold).
+We compute word specificity and ambiguity based on tree depth and the number of senses in WordNet [54].
+We also extract multiple WordNet measures of similarity: Resnik [68], Jiang-Coranth [69], Lin [47], Leacock-Chodorow [70], and Wu-Palmer [71].
+
+\subsection{Syntactic features}
+Guinn, Singer and Habash \cite{Guinn2015} found that syntactic features such as Noun rate, verb rate, adjective rate and pronoun rate were not significantly different between interviewers and those with dementia in their dataset, although they did note that there was a slightly higher but non significant use of pronouns.
+
+\subsection{Pragmatic features.} We train a general 100-topic latent Dirichlet allocation (LDA) model [72] on the Wikipedia corpus for generalizability.
+LDA is a generative statistical model used to determine unlabeled topics in a document.
+For each transcript, we extract the probabilities of each LDA topic.
+Next, we extract features related to rhetorical structure theory (RST), which is a classic framework for discourse parsing in which partitions of text are arranged in a tree structure by pragmatic relations such as Elaboration or Contrast [73].
+
+\subsubsection{Go-ahead Utterances}
+Go-ahead utterances are defined as short one or two syllable responses that do not contribute to the conversation beyond a minimal response.
+The function of these go-ahead utterances can be to validate what the other person is saying, or to agree / disagree with what is being said.
+Another function can be that they wish speaker is indicating that they have nothing futher to add to the conversation and a signal to the other speaker to continue with what they are saying.
+In Curry, Singer and Habash's research, they found that in comparing interviewers and those with dementia, that interviewers used significantly fewer go-ahead utterances.
+In comparing controls and those with dementia, they found that there was a relative lack of go-ahead utterances in the controls which implies that controls had a lot more contributions to make in their conversations than those with dementia.
+\par
+
+\subsection{Formulaic Language}
+Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
+They found that there were four factors which informed the classification of participants as either healthy or AD.
+These four factors were semantic impairment, acoustic abnormality, syntactic impairment and information impairment and were based on existing measures of semantic and syntactic complexity.
+Zimmerer (2016) \cite{Zimmerer2016} looked at whether language was more formulaic in those suffering from AD.
+He proposed that those who suffer from AD rely on formulatic sentences, for example 'Noun-Verb-Noun', and this is done to reduce language complexity.
+He noticed a significant difference in the use of formulaic sentences between AD and Healthy Controls.
+\newline
+
+\subsection{Number of syllables and Characters}
+
+\subsection{Number of fillers}
+
+\subsection{Readability}
+Flesch reading score, Flesch-kincaid grade level
+
+\subsection{Polarity}
+
+\subsection{Frequency}
+Mean values of frequency, age of acquisition, imageability, familiarity, arousal, dominance and valence based on lexical norms
+
+\subsection{Dysfluencies}
+Curry, Singer and Habash noted that in comparing controls and those with Dementia, that those with Dementia had a significantly higher number of pauses per word and a much higher incidence of words that were truncated in mid-speech.
+In comparing interviewers with those with dementia, they also showed other signs of difficulties with fluency with higher rates of incomplete words, filler words and repeated words.
+
+\section{Machine Learning methods}
+The next stage of building statistical or Machine Learning models is to take some features and use these feature to build models able to classify participants into various categories.
+There are various types of classification models that have been used to tackle this problem, this section looks which models have been used in the current literature.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ p{6cm} | p{6cm} }
+	\hline
+	Machine Learning technique used & Paper Number \\ \hline
+	Logistic Regression & 2 \\ \hline
+	Support Vector Machines & 4 \\ \hline
+	Naive Bayes & 1 \\ \hline
+	Decision Trees & 1, 2 \\ \hline
+	Multi-layered perceptrons & 2 \\ \hline
+	\hline 
+	\end{tabular}
+	\end{center}
+\end{table}
+
+\subsection{Traditional Machine Learning methods}
+\subsubsection{Logistic Regression}
+Logistic Regressions is a probabilistic approach to classification and is dependant on the assumption that your input space can be separated into two distinct regions, one for each class, by a linear boundary.
+In 2D space, this boundary is a line and in 3D space, this boundary is defined as a plane.
+\par 
+In the results of this survey.
+
+\subsubsection{Support Vector Machines}
+Support Vector Machine (SVM) was originally proposed as an algorithm for classification problems; it is a relatively new technique compared to the other ML approaches.
+The classification process consists of mapping the data points (usually the study subjects) into a feature space composed of the variables that characterize these data points, except for the outcome variable.
+Then, the algorithm finds patterns in this feature space by defining the maximum separation between two or more classes, depending on the problem to be solved.
+Contrary to some regression techniques, SVMs are not dependent on a pre-determined model for data fitting, although there are still algorithm specifications to be considered (e.g.
+choice of a kernel function); instead, it is a data-driven algorithm that can work relatively well in a scenario where sample sizes are small compared to the number of variables, reason why it has been widely employed by diagnostic studies in tasks related to the automated classification of diseases.
+\par 
+Regarding the SLR results, SVMs were present in ??/41 selected studies, in 38 proposed models, and being by far the most used machine learning technique in the dementia diagnosis research.
+These numbers account for the traditional SVM and variations.
+In all of the ??
+selected studies the SVMs focused at binary classifications where the task was to discriminate mild cognitively impaired (MCI) patients that will or will not develop Alzheimer's Disease (AD).
+In the general case, the problem is posed as either MCI converters versus MCI non-converters, or progressive MCI versus stable MCI classification.
+This outlines a situation in which a regression problem (when will the MCI patients convert to AD?) is formulated as a classification problem (which MCI patients will convert to AD in X months?) to be solved.
+Reasons for this could be due to limitations in the data used, i.e.
+the limited follow-up periods of the subjects included in the studies.
+
+\subsubsection{Linear Discriminant Analysis}
+
+\subsubsection{Decision Trees}
+A Decision Tree (DT) is a classification algorithm in which the learned knowledge is represented in a tree structure that can be translated to if-then rules.
+DT's learning process is recursive and starts by testing each input variable as to how well each of them, alone, can classify the labeled examples.
+The best one is selected as a root node for the tree and its descendant nodes are defined as the possible values (or relevant ratios) of the selected input variable.
+The training set is then classified between the descendant nodes according to the values of the selected input variable.
+This process is repeated recursively until no more splits in the tree are possible.
+Like SVMs, DTs do not depend on a pre-defined model and are mostly used to find important interactions between variables.
+Being intuitive and easy to interpret, DTs have been used in prognostic studies as a tool for determining prognostic subgroups.
+\par 
+In this SLR, DTs were the second most frequently used ML technique, present in 6/37 selected studies and proposed in 7 models.
+It was employed for the same reason as SVM; except for one study that investigated the evolution of patients diagnosed with cognitive impairment no dementia (CIND) to AD.
+
+Curry, Singer and Habash - Alz (66.7 - Pres, 66.7 - Recall) and Control (67.9 - Pres, 67.9 Recall)
+
+\subsubsection{Naive Bayes}
+
+Curry, Singer and Habash - Alz(80.8 - Pres, 0.75 - Recall) and Control (79.3 - Pres, 82.1 Recall) - In the naive bayes classifier, they identified that pauses, go-ahead, fillers and incomplete words as the most significant features.
+%Why is Recall at 0.75???????
+
+\subsubsection{Artificial Neural Networks}
+An Artificial Neural Network (ANN) is a methodology that performs multifactorial analyses, which is desirable in the health area as medical decision-making problems are usually dependent of many factors.
+An ANN is composed of nodes connected by weighted edges in a multi-layer architecture that comprises: an input layer, one or more hidden layers and an output layer.
+In the training process, inputs and outputs values are known to the network, while the weights are incrementally adjusted so that the outputs of the network are approximate to the known outputs.
+Despite being a powerful predictor, ANNs are `black boxes', which means that they are not able to explain their predictions in an intuitive way, contrary to DTs or BNs.
+Also, they require the specification of the architecture to be used beforehand (i.e.
+the number of hidden layers).
+
+\subsubsection{K-nearest neighbours}
+K Nearest Neighbors (KNN) is a classification algorithm that takes a data point from an unknown class and assigns it as an input vector in the feature space.
+Then, the classification process follows by assigning the unknown class data point to the class in which the majority of the K nearest data points belong to.
+The distance between data points is usually measured by Euclidean distance, but it is possible to employ other measures.
+KNN is one of the simplest ML classification algorithms and have been used in a wide range of applications; however, it can be computationally expensive in a highly dimensional scenario.
+Further, it considers all features to be equally weighted, which can be a problem if the data has superfluous attributes.
+
+\subsubsection{Best Results}
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ p{6cm} | p{6cm} }
+	\hline
+	Machine Learning technique used & Paper Number \\ \hline
+	Logistic Regression & 2 \\ \hline
+	Support Vector Machines & 4, 5 \\ \hline
+	Naive Bayes & 1 \\ \hline
+	Decision Trees and Random Forest Classifiers & 1, 2, 5 \\ \hline
+	Multi-layered perceptrons & 2 \\ \hline
+	\hline 
+	\end{tabular}
+	\end{center}
+\end{table}
+
+\subsection{Deep Learning methods}
+\subsubsection{Deep language space neural network for classifying mild cognitive impairment and Alzheimer-type dementia - Orimaye, Wong and Wong (2018)}
+In this paper, Orimaye et al use deep-deep neural networks language models (D2NNLM) to learn linguistic changes that distinguish the language of patients with MCI and AD-type dementia from the healthy controls using higher order n-grams.
+An ordinary DNNLM uses lower order n-gram N-dimensional sparse vectors as descrete feature representations to train the neural network with multiple hidden layers.
+
+\subsection{What type of data is used by the studies?}
+One of the key debates when looking at how to analyse language is the type of task provided to elicit language production in participants.
+In the literature researchers have primarily focused on Picture Description tasks but have also suggested other ways in which we might collect data.
+\par
+\subsection{Picture Description Tasks}
+One of the most commonly used tasks to measure language is the Picture Description task.
+An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Borod1980}.
+The Cookie Theft picture (pictured below) depicts a scene of a home typical of the period of time when it was created and would generally not require participants to use any complicated vocabulary to describe.
+In this task participants are asked to describe the picture presented to them in as much detail as possible.
+This task was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}
+\begin{figure}[H]
+\centering
+\includegraphics[width=240px, height=150px]{images/BCTPicture.png}
+\caption{Cookie Theft Picture - From Kaplan and Goodglass (1983)}
+\end{figure}
+\par
+The picture description task does a fine job of eliciting descriptive language however because of the specific content the language produce could be considered quite limited.
+There is some disagreement as to the benefits of this using this methodology.
+This task is reported as being useful to lexico-semantic disorders \cite{Boschi2017, Sajjadi2012} as the language being generated is primarly nouns and deixis (words to identify items and words to put those items into context).
+However, Ash \cite{Ash2012}felt that there was no difference in using this task vs Story Narration (described below).
+In explaining the differences, it is worth noting that these researchers were using differing variables and this could explain their different perspectives.
+\par 
+There are a number of existing corpus which use this tasks as the foundation of their data.
+The most well known is the Pitt Corpus of the Dementia Bank database.
+The participants of the corpus are mainly healthy controls, those with MCI and those with probable and/or possible AD.
+\subsection{Narrative description task}
+The story narration task is designed to study a participant's ability to describe and elaborate on a story which is depicted using a series of pictures.
+The stories depicted are usually based on children's books or famous stories with the Cinderella being the one most typically used \cite{Fraser2014}.
+This task requires ordering the story in a structured and coherent framework.
+It also requires comprehension and understanding of the stories characters and the events depicted, as well as an awareness of a character's actions, motivations and internal reactions to given events.
+This task is particularly useful as the procedure reduces the demands on memory, due to the participant being able to access the picture book during the description and is therefore able rule out memory as a confounding variable for any results observed.
+As noted above, Ash \cite{Ash2012} felt that this task was interchangable with the Picture Description task.
+However, other research felt that this was a studier test of lexical and semantic abilities as well as syntactic complexity because this task requires interpretation and elaboration in additional to a simple description \cite{DeLira2011}.
+\par
+Given the relative strengths of the Narrative description task vs Picture description task, there are few pieces of research that have used Machine Learning to analyse features from Narrative picture tasks \cite{Fraser2014}.
+This could be due to the availability of data and the absence of any meaningful sets of transcripts of participants performing this task.
+However, this could be an interesting direction to take research in the future to see if features generated from this task could be used to predict MCI or AD.
+\par
+\subsection{Interviews}
+Interviews can also be used to elicit language in a more natural way by asking questions to guide a conversation between speakers.
+There are three types of interviews: unstructured, structured and semi-structured.
+Structured interviews tend to produce very limited speech and therefore has never been used in this area \cite{Boschi2017}.
+Unstructured interviews are open ended and generally do not conform to any particular pattern.
+They use generic themes such as family or hobbies to guide the conversation.
+Whilst this is the most ecologically valid form of conversation and therefore language generation, it's unstructured nature means that the protocol cannot be consistent and therefore reproduced.
+Semi-structured interviews are therefore preferred over other forms of interview as a middle ground.
+The semi structured nature of these interviews means that there is some replicability but does not constrain the participant in answering questions.
+\par
+The analysis of interviews can be difficult to analyse as both the content can vary even between participants, although it can be argued that content should not affect the type of language being generated unless it is narrow topic or the participant is constrained in how they answer a given question.
+It is also difficult to measure as there are no pre-defined task goals in comparison to the other two methods.
+Nevertheless, this is the most naturalistic setting for looking at language production and can be used to look at the syntactic and semantic parts of language generation \cite{Sajjadi2012}.
+There have been some attempts to use interviews to assess language production in AD with promising results \cite{Asgari2017, Guinn2015}.
+\par
+\subsection{Conclusions}
+One can view the different types of tasks above as a continuum where picture tasks represent a much more controllable task with a lot of supporting research but which generates a much more constrained set of language that is atypical of normal speech in terms of the cognitive functions used.
+
+\subsection{What are the goals of the studies that employ ML or Statistical Learning techniques for diagnosis of MCI or AD?}
+
+\subsection{Do the studies focus on a one point in time or looking at cognitive deterioration over time?}
+
+\section{Discussion and conclusions}\label{discussion}
+\subsection{Discussion of the current evidence}
+One of the criticisms of traditional learning models is it's reliance on features that are generated for them.
+As we can see from the brief look at some of the research in this field, researchers have taken a number of different approaches with success.
+However, there does not seem to be a consensus.
+As machine learning in general moves away from traditional machine learning models to deep learning techniques, it poses the question can deep learning assist with this problem?
+\par 
+One of the main benefits of deep learning is that it does not rely on pre defined features being fed to the model.
+Instead, deep learning models take raw data in with some amount of preprocessing and generates it's own features.
+This move away from relying on features solves one of the difficulties we have with the current psychological literature, namely that there are some disagreement on the which
+\par 
+One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
+This is the case with all machine learning techniques, but more so with deep learning.
+There have been ways to 'create' more data such as data augmentation.
+
+Current state-of-the-art diagnostic measures of AD are invasive (CSF analysis), expensive (neuroimaging), and time-consuming (neuropsychological assessment).
+Furthermore, these measures are limited to speciality clinics and thus have limited accessibility as frontline screening and diagnostic tools for AD.
+More importantly, nonspecialists are often inaccurate at identifying early AD and MCI.
+Thus, there is an increasing need for additional noninvasive and/or cost-effective tools, allowing effective frontline identification of subjects in the preclinical or early clinical stages of AD who could be suitable for monitoring in speciality clinics and for early treatment.
+Implementation of effective screening instruments will allow diagnosis earlier in the course of dementia, even at the point when memory function is still essentially within the normal range.
+This strategy would enable an earlier, and potentially more effective, prevention and treatment of AD with a special focus to preserve cognitive functions.
+\par 
+However the literature has identified a number of challenges when approaching this problem.
+Firstly, the clinical features that combine to meet the diagnostic criteria for Dementia or it's variants are continuous in nature and heterogenous between patients and are also impacted by other variables.
+For example as shown above, cognitive performance is affected in part by a patients educational attainment and a patients ability to live independently is impacted by a their physical health as much as their cognitive health.
+The challenge therefore is to find features that are minimally impacted by other factors, or that can be controlled for by a experimental design such as a matched pairs design to control for educational attainment.
+\par 
+Another challenge lies with the recruitment of suitable individuals who may notice a decline in cognition to the point where we might classify them as having MCI, but these individuals deduce that there is little to no value to admitting there is a problem and seeking help whilst their symptoms are 'manageable'.
+\par
+Finally there is a large amount of variability in the presentations of those with MCI and early dementia, and this is compounded by an similar amount of variability in the criteria researchers have used for experimental groups and the approaches researchers have used to tackle this problem.
+This had led to a confused literature.
+Recently a call for research that has consistent inclusion / exclusion criteria has been made along with some proposed definitions of MCI and it's subgroups \cite{Petersen2014}.
+Researchers have identified the analysis of language impairment as an area of promise to explore in the diagnosis of MCI and early AD and recent developments in natural language processing and machine learning techniques have the potential to assist in this research.
+Indeed, given the increased burden on the diagnosis of MCI and AD on professionals there has been a call to use technology to potentially ease this burden \cite{Boschi2017}.
+A small but growing amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
+
+\subsection{Methodological Issues}
+\subsection{Limitations}
+
+\subsection{The future of the field}
+One of the main patterns
+
+One-class classification is a traditional machine learning model in which the object is to recognize instances of a concept by only using examples of the same concept.
+In the training of this model, instances of only a single object class are available during training.
+In this context,all other classes (any instances that deviate from the single object class) are referred to as 'alien' classes.
+During the testing of the model, it may encounter alien classes and the job of the model is to distinguish between objects of the known class from objects belonging to any 'alien' classes.
+This approach to classification has broad applications and is mainly used when there is a high cost of obtaining samples of alien classes.
+\par 
+In this context, we can see from
+
+One of the areas for research to study is a careful examination of the features that are being used to measure language deterioration.
+For example, Zimmerer (2016) \cite{Zimmerer2016} describes connectivity in such as a way that correlates directly with what Mueller (2018) \cite{Mueller2018a} calls Fluency.
+Whilst these are very nuanced measures which differ slightly in the form they take, the sheer range of measures and features being produced make it difficult to organise and explore what is truly going on in those with MCI and early AD..
+Some work needs to be done in producing a consistent list of measures that are validated using existing datasets and can be used for future research moving forward.
+\newline
+\par
+Teng et al suggests that work should focus on the MCI population and concentrate on developing a consensus neuropsychological battery that could yield predictable rates of progression to AD \cite{Teng2009}.
+This, in conjunction with the development of a model of language, sensitive enough to detect subtle deterioration in language use to act as an additional cognitive marker to aid diagnosis could potentially move some way to providing this.
+\par 
+Future research should also be directed towards developing non-intrusive ways of detecting subtle changes in natural language such that any perceived deterioration that could indicate the presence of MCI or AD could be flagged up early.
+Machine learning approaches seem to be the most logical approach for achieving this aim as language could be collected in non-intrusive ways and passed to a machine learning algorithm for preliminary classification.
+Despite the excellent quality of datasets, for example the DementiaBank dataset, being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which some researchers could claim are limited in scope.
+Finally, the recent resurgence in the use of neural networks and deep learning could provide the answer to the confused literature in terms of features.
+A key benefit of deep learning is it's ability to automate the process of feature engineering.
+So there is an opportunity to explore the use of deep learning, to not only develop new features but also validate existing features independently.
+\par
+This area of research is extremely promising in its early results and the impact of successful research would be life changing for both individuals and the health of the worlds aging population in general.
+
+\section{Conclusions}\label{conclusions}
+We can see that both Natural Language Processing and Machine Learning techniques have a lot to offer, Indeed there has been a lot of research which have used pre-existing datasets to explore this area with promising results.
+One of the difficulties with the current research is the approach of trying to discriminate between those AD and healthy controls.
+This is not necessarily a problem in the real world as it is trivially easy to do for trained clinicians.
+A more interesting, but potentially harder to problem to solve is to discriminate between those with MCI and healthy controls, and more importantly to track their decline over time.
+Further, to date the authors are unaware of any research in which these techniques are applied to newly created samples of language.

### 2019-12-06 Source/Thesis/Chapter5.tex
update Thesis and Delphi Questionnaire

@@
 Equally, it would be helpful to have controls to ascertain what would be usual to expect in the decline of language in a healthy older adult.
+
+\section{Introduction}\label{introduction}
+A diagnosis of dementia is generally made when there is a decline in brain function due to physical changes in the brain \cite{Albert2011}.
+It affects a significant proportion of the global older adult population and the impact on morbidity and mortality rates is considerable.
+Dementia is currently the leading cause of death in England and Wales and the sixth leading cause of death in the United States (US).
+A 2014 report commissioned by the Alzheimer's Society estimated that in the UK by 2015 there would be approximately 855,000 people rising living with dementia increasing to 1 million by 2021 \cite{AlzheimersSociety2014}.
+This represents 1 in 79 of the total UK population rising to 1 in 14 of those aged 65 or over \cite{AlzheimersSociety2014}.
+Worldwide, there are 46 million people with a diagnosis of dementia globally and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+From a financial perspective, the cost burden is also significant.
+The estimated annual spend on dementia healthcare in the UK is £4.3 billion of which approximately £85 million is spent on diagnosis.
+The total financial burden of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually.
+Globally, this picture is a lot bleaker.
+The worldwide cost of dementia in 2018 was estimated to be in the region of one trillon US dollars \cite{Prince2015}.
+\par 
+There are different types of dementia including Alzheimer's Disease (AD), vascular dementia, dementia with Lewy bodies and fronto-temporal dementia, all of which are currently incurable.
+AD is a progressive neurodegenerative disease and is the most common type of dementia, responsible for approximately 60\% to 80\% of all cases \cite{Duong2017}.
+Currently, a definitive diagnosis for AD can only be produced at post-mortem.
+However, there are a number of psychological and physiological indicators that can indicate that dementia is present.
+From a physiological perspective, researchers have identified two proteins called beta-amyloid and tau have been identified as two of the key proteins that have a role in initiation and progression of AD.
+In a typical case beta-amyloid clumps into plaques which slowly build up between neurons with tau proteins accumulating and eventually forming tangles inside neurons.
+At a certain point, the levels of beta-amyloid rise triggering a more rapid spread of tau throughout the brain.
+Eventually, due to these and other changes, neurons lose their ability to communicate and the brain starts to shrink\cite{Duong2017}.
+This leads to the development of some of the psychological symptoms associated with AD, primarily cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}.
+These cognitive symptoms are often accompanied by emotional problems such as depression and behavioural difficulties.
+As more neurons die throughout the brain, a person with AD gradually loses the ability to think, remember, make decisions and function independently.
+\par
+Despite the increasing prevalence of AD and an improved understanding about how it affects the brain there are no medications that improve prognosis.
+All the medications that are currently on the market are designed to manage symptoms.
+Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage of these drugs fail in clinical trial stage of the drug discovery process ((99.6\% failure rate vs 80\% for systemic anti-cancer drugs) \cite{Cummings2014}.
+Cummings et al proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already occurred \cite{Cummings2014}.
+Research focus has now started to shift to the earlier stages of AD (i.e.
+symptomatic pre-dementia phase of AD) which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.
+\par
+One of the challenges associated with the early detection of AD is differentiating natural age associated memory impairment and cognitive decline due to aging from with decline due to AD.
+This challenge is often complicated further due to the large variation in the cognitive abilities and educational background of individuals.
+The work of Albert et al helps to address this by the development of clinical criteria which professionals can use to diagnose MCI due to AD.
+One of the most important observations from this piece of work is that a diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}, and this is essentially to place more importance on the trajectory of a person's cognitive abilities rather than a person's cognitive ability in general.
+\par 
+There has been a significant research in the area of language deterioration as a method of detecting AD at an earlier stage.
+This usually takes the form of recording speech whilst patients' undertake a cognitive assessment such as the Picture Description Task \cite{Fraser2015}.
+Given that language samples are relatively easy to collect, research has moved towards analysis of spontaneous speech.
+The work of Berisha and Liss is a good example of this, this study compared the differences in language use between two US presidents, Ronald Reagan (RR), who would go on to receive a diagnosis of Dementia and George H.
+W.
+Bush (GHWB) who acted as a matched control based on Age \cite{Berisha2015}.
+They found several differences in language use over time which they felt acted as indicators of RR's difficulties with language due to AD.
+Differences in features of language identified to be statistically significant included the number of unique words used per speech, the use of non-specific nouns and fillers and low imageability verbs \cite{Berisha2015}.
+\par   
+Berisha and Liss have developed some very interesting ideas about how we might track changes over time in various lexical features which have been associated with the development of Alzheimer's Disease.
+However we feel that cognitive decline in pre-clinical AD is not accurately modeled as a linear process and therefore we explored the application of generalised additive models (GAMs) to this data which have no assumed understanding of the distribution of the data.
+We refine these ideas with the aim of exploring the potential for a protocol that can be used for analyzing language deterioration.
+
+This study extends the work of Berisha and Liss in a number of ways. Our hypotheses are:
+\begin{enumerate}
+	\item Language use of RR will be significantly different to our controls (GHWB, DJT).
+	\item Modelling the dynamics of lexical features over time can be improved using non-linear models compared to linear models.
+\end{enumerate}
+
+\section{Methodology}\label{methodology}
+We took 46 transcripts of press conferences given by RR (from 1981 to 1988) and compared them with 134 press conferences (from 1989 to 1993) given by GHWB and 29 press conferences (from 2016 to 2019) conducted by DJT.
+We analysed transcripts for language features (described below) shown to change longitudinally with AD.
+These language features are analysed at the word level, sentence level and document level.
+\par 
+In the original study, GHWB was selected as the comparator president as he was the closest match in terms of age to RR (GHWB - age at the start of presidency: 64 years and 222 days vs RR: 69 years and 349 days).
+However, since his inauguration, DJT is now the closest comparible president in terms of age (DJT - age at the start of presidency: 70 years and 220 days).
+We included DJT who like GHWB has no known diagnosis of AD to determine whether the comparisons made by Berisha and Liss hold true with this closer presidential match in terms of age.
+\par 
+We used the press conference transcripts in the American Presidency Project (APP) archive as a data source for this project.
+The APP is a comprehensive and organized searchable database of presidential documents, including transcripts of speeches, transcripts of news conferences, and other public documents.
+These documents are open access and can be downloaded at any time from the APP archive.
+
+\subsection{Pre-processing}
+To generate the files necessary for analysis, we downloaded each transcript and performed the following changes.
+We omitted the prepared statement by the president and any speech by other individuals and started each transcript at the beginning of the first answer to a question by a member of the press.
+We filtered any annotations that were added to the transcript, including any references or clarifications, and any laughter.
+It's worth noting that there appears to be a difference in how 'hesitations' were marked down between each president, for RR \& DJT hesitations were marked by a single hyphen whereas for GHWB hesitations are marked by a double hyphen.
+In order to maintain consistency when parsing through the documents, We have changed both types of hesitation to be marked by a single hyphen.
+We also omitted one word answers to questions as this data would, from a theoretical perspective, not be relevant for language analysis.
+We did not make any alterations to the length of the document, but instead generated features which would normalise by the length of the document.
+We therefore was able to include all press conferences by all presidents analysed where there was a question and answer session conducted at least in part by the president in question (2 press conferences of GHWB were omitted due to a lack of a question and answer session).
+
+\subsection{Feature Generation}
+Features were generated by running each transcript through a number of natural language processing libraries.
+We used the Natural Language ToolKit (NLTK) package \cite{Bird2009} in python to stem each transcript using the Snowball Stemmer and also completed a part of speech (POS) tagging process.
+We also ran each transcript through Linguistic Inquiry and Word Count (LIWC) \cite{Pennebaker2015} software.
+We completed all NLP tasks in Python and completed all statistical analysis in R..
+\par 
+\subsubsection{Measures of lexical diversity}
+Lexical diversity is a measure of the variety of language used within a given document.
+A document is said to have high lexical diversity if the number of unique words is large.
+We constructed four features of lexical variation.
+Firstly we looked at the number of unique words.
+To do this we were able to split each transcript into individual words and changed them to lowercase using NLTK and were then count the number of unique words that appeared in each transcript.
+We also used the TTR formula, Brunet's Index and Honore's Statistic as other measures of lexical diversity \cite{Bucks2000}.
+\par 
+Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
+\begin{equation} \label{x1}
+TTR = numberOfUniqueWords / totalNumberOfWords.
+\end{equation}
+\par 
+Brunet's Index (W) differentiates itself from TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
+\begin{equation} \label{x2}
+W = N^{V(-0.165)}
+\end{equation}
+where N is the total length of the utterance being measured and V is equal to the total vocabulary being used by the subject.
+Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers.
+\newline
+\par 
+Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation: 
+\begin{equation} \label{x3}
+R = (100 \log N) / (1 - V1/V)
+\end{equation}
+where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
+\subsubsection{Fillers, Non-Specific Nouns and Low Imageability Verbs}
+Fillers, Non-Specific Nouns and Low Imageability Verbs were features used by Berisha and Liss in their research \cite{Berisha2015} and were taken from work done by Bird et al \cite{Bird2000}.
+Fillers can be described as a potentially meaningless word that marks a pause or hesitation in speech.
+In those with MCI and AD, these words can be used to temporarily disguise problems in thought processes or word finding difficulties.
+Non-Specific Nouns refer to a category or an unspecified member of a given category, once again this can be characterised as a compensatory strategy for word finding difficulties.
+Imageability is characterized, according to Berisha and Liss\cite{Berisha2015}, as the ease with which a word provokes a mental image of what the word describes.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{10cm} |}
+		\hline
+		Category & Words \\ \hline
+		Fillers & \textit{"um", "uh", "er", "ah", "like", "okay", "right",  "you know", "well", "so", "basically", "actually", "literally"} \\ \hline
+		Non Specific Nouns & \textit{"something", "anything", "thing", "everything"} \\ \hline
+		LI Verbs & \textit{"be", "come", "do", "get", "give", go", "know", "look", "make", "see", "tell", "think", "want"} \\ \hline
+	\end{tabular}
+	\caption{\label{tab:table-name}Examples of words belonging to the categories Fillers, Non-Specific Nouns and Low Imageability Verbs}
+	\end{center} 
+\end{table}
+
+\subsubsection{Usage of parts of speech}
+Using a Part of Speech tagger (PoS) on each transcript analyses each sentence within the transcript and assigns a 'tag' to each word based on the function the word has in a sentence.
+At a basic level this can be divided into the eight defined parts of speech: 'nouns', 'pronouns', verbs', 'adjectives', 'adverbs', 'conjunctions', 'prepositions' and 'interjections' but can be further subcategorised.
+We used the PoS tagger built into NLTK to tag each transcript in turn and used these the counts from each of these eight categories in our analysis.
+In addition to frequency counts we also normalised these features by dividing the frequency count by the number of words in the document to take into account transcript length.
+
+\subsubsection{Linguistic Inquiry and Word Count(LIWC)}
+The LIWC is a text processor which analysis words within a given a document and compares the words in the LIWC dictionary file.
+The LIWC dictionary file contains over 6000 words which are categorised into approximately 90 different types.
+For example the word 'cried' is part of five word categories: sadness, negative emotion, overall affect, verbs and past focus \cite{Pennebaker2015}.
+Each word in the document or set of documents is searched for in the dictionary file and if found, the appropriate type is incremented by 1.
+What is output is an analysis of word usage in reference to each category.
+This is particularly useful in analysing both structural language changes but also the content or themes of language for a given transcript or set of transcripts.
+
+\subsubsection{Longitudinal Analysis}
+As with Berisha and Liss\cite{Berisha2015}, we analysed the transcripts over time in order to determine if there was an underlying temporal trend that may indicate that different parts of language increase or decrease over time.
+In AD, cognitive abilities are said to deteriorate over time and therefore it may be useful to analyse temporal trends that indicate a potential deterioration of cognition.
+In order to get a accurate measure of deterioration over time, we time stamped each transcript in relation to number of days from the first transcript to the date of the transcript being analysed as the press conferences were not evenly spaced out during the presidencies.
+\par
+We looked also compared the results of linear models and a non linear model in terms of best fit.
+The non linear model we used was the generalised additive model \cite{Hastie1986}.
+The selection of this non linear model was due to the authors not wanting to assume an underlying distribution of the data and therefore this was the most appropriate model to use.
+
+\section{Results}\label{results}
+One of the most important thing to note is the wide variety of samples between the three presidents and also the varying timescales.
+RR participated in 46 press conferences over eight years (an average of 5.75 a year) which is the fewest number of press conferences given by an American president during their term of office.
+GHWB participated in 136 press conferences over four years (an average of 34 a year) and DJT participated in 29 press conferences to date (an average of 19.3 per year).
+Equally, there are differences in the average number of words.
+RR produced an average of 3424 words per conference compared to 2608 by GHWB (unpaired t = 4.434, p\textless 0.001) and DJT at 1849 words (unpaired t = 6.524, p\textless 0.001).
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR & GHWB & DJT \\ \hline
+		Total Words & 3423.91 (416.42) & 2607.72 (1210.38) & 1848.65 (1549.38) \\ \hline
+		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
+		Mean Length of Utterance & 23.17 (1.402) & 18.71 (2.067) & 13.84 (1.619) \\ \hline
+	\end{tabular}
+	\caption{\label{tab:table-name}Means and Standard Deviations of general features for each set of transcripts}
+	\end{center} 
+\end{table}
+
+In terms of more specific language differences between the presidents, we found that RR used significantly more unique words, non-specific nouns and low imageability verbs than GHWB and DJT (see Table 3).
+The mean length of utterance for RR was significantly greater than that of GHWB and DJT.
+Some of these differences are due to the length of the sample, particularly in the case of DJT where his average sample is almost half the sample of RR.
+It could also be said that this could be down to differences in linguistic abilities or speaking style \cite{Berisha2015, Le2011}.
+However, we can certainly see that as controls GWHB and DJT are comparative in relation to non-specific nouns and LI verbs.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{2.2cm} | p{2.2cm} | p{2.2cm} |}
+		\hline
+		& RR v GHWB & RR v DJT & GHWB v DJT \\ \hline
+		Total Words & \textbf{4.434***} & \textbf{6.524***} & \textbf{2.899**} \\ \hline
+		Unique Stems & \textbf{10.878***} & \textbf{10.111***} & \textbf{4.148***} \\ \hline
+		Mean Length of Utterance & \textbf{16.175***} & \textbf{25.084***} & \textbf{13.484***} \\ \hline	
+		Non Specific Nouns & \textbf{7.877***} & \textbf{3.426**} & -0.574 \\ \hline
+		LI Verbs & \textbf{2.656**} & \textbf{3.420***} & 1.628 \\ \hline
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+	\end{tabular}
+	\caption{\label{tab:table-name}RR T-tests vs GWB and DJT}
+	\end{center} 
+\end{table}
+
+\subsection{Longitudinal Analysis}
+We then looked at the data from a longitudinal perspective as we were interested seeing whether we can track various language variables and their progress over time.
+We ran a number of Pearsons correlations with transcript index number as a time reference and the dependant variables (Table 4).
+For our controls, we found them to be stable for the most part with the main highlights being a decrease in Adverb usage for DJT (R=-0.36, p=0.049) and a steady but not severe decline in a number of variables for GWHB, namely total word count, unique words, low imageability words and verb usage.
+\par 
+For RR, his decline is more marked and more widespread through his language use.
+We noticed an significant increase in adverb (R=0.41, p=0.004) and pronoun usage (R=0.65, p\textless0.001), as well as a slight usage increase in Non-specific nouns(R=0.30, p=0.03).
+There was a highly significant decrease in number of unique words (R=-0.56, p\textless0.001) and noun usage (R=-0.70, p\textless0.001).
+Also very significant decrease in adjective usage (R=-0.40, p=0.005) and a significant decrease in total word count (R=-0.31, p=0.03).
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | l | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR & GHWB & DJT \\ \hline
+		Word Count & \textbf{-0.31*} & \textbf{-0.21*} & 0.08 \\ \hline 
+		Unique Words & \textbf{-0.56***} & \textbf{-0.25**} & 0.16 \\ \hline
+		Non Specific Nouns & \textbf{0.30*} & -0.08 & -0.03 \\ \hline
+		LI Verbs & -0.19 & \textbf{-0.20**} & 0.02 \\ \hline
+		Nouns Normalised & \textbf{-0.70***} & -0.03 & 0.14 \\ \hline
+		Verbs Normalised & \textbf{0.36**} & \textbf{0.24***} & -0.03 \\ \hline
+		Adjectives Normalised & \textbf{-0.40**} & 0.08 & -0.34 \\ \hline
+		Adverbs Normalised & \textbf{0.41***} & 0.02 & \textbf{-0.36*} \\ \hline
+		Pronouns Normalised & \textbf{0.65***} & 0.13 & 0.07 \\ \hline
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+	\end{tabular}
+	\caption{\label{tab:table-name}Pearson Correlations for Features}
+	\end{center} 
+\end{table}
+
+To get a better idea of decline over time, we also ran Pearson's correlations using time elapsed in days from the first transcript instead a transcript index.
+This was to control for variation in the number of days between transcripts.
+Given the number of hypotheses being tested, we felt it was necessary to apply some correction for family wise error rate and therefore a Hochberg step-up procedure was applied to the results of the analysis.
+We also used the Benjamini-Yekutieli procedure for controlling for False Discovery rate and came up with a final list of important features for each president.
+For RR, this was 22 features, for GHWB this was 14 features and there were no significant features for DJT after applying these methods.
+The features were generated from multiple sources and therefore there was some significant overlap between the features but we have included all features in Table 6.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | l | c | c | c |}
+		\hline
+		& RR R-Squared  & GWB R-Squared & DJT R-Squared \\ \hline
+		ppron & \textbf{0.700***} & 0.007 & -0.196 \\ \hline
+		social & \textbf{0.698***} & 0.204 & 0.032 \\ \hline
+		NounsNormalised & \textbf{-0.689***} & -0.023 & 0.194 \\ \hline
+		function & \textbf{0.670***} & -0.169 & -0.243 \\ \hline
+		conj &\textbf{0.644***} & -0.452 & -0.187 \\ \hline
+		PronounsNormalised & \textbf{0.631***} & 0.131 & 0.002 \\ \hline
+		Analytic & \textbf{-0.626***} & -0.013 & 0.366 \\ \hline
+		NN & \textbf{-0.601***} & -0.197 & 0.116 \\ \hline
+		male & \textbf{0.585***} & 0.024 & 0.042 \\ \hline
+		UniqueWords & \textbf{-0.578***} & -0.257 & 0.204 \\ \hline
+		WDT & \textbf{-0.577***} & -0.173 & -0.105 \\ \hline
+		shehe & \textbf{0.529*} & -0.005 & 0.064 \\ \hline
+		VBZ & \textbf{-0.521*} & -0.157 & 0.136 \\ \hline
+		JJ & \textbf{-0.518*} & -0.206 & 0.062 \\ \hline
+		Fillers & -0.076 & \textbf{-0.359**}  & -0.032 \\ \hline	
+		EX & -0.193* & \textbf{-0.337**}  & 0.339 \\ \hline	
+		achieve & -0.248 & \textbf{0.334**}  & 0.169 \\ \hline		
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+	\end{tabular}
+	\caption{\label{tab:table-name}Pearson Correlations for Features}
+	\end{center} 
+\end{table}
+
+In terms of comparing the general additive model and linear models.
+We calculated the predicted residual error sum of squares (PRESS) statistic for each feature is a form of cross-validation used in regression analysis to provide a summary measure of the fit of a model to a sample of observations that were not themselves used to estimate the model.
+There were no significant differences (paired t-test: t=1.8875, df  = 21, p-value = 0.07299).
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | l | c | c |}
+		\hline
+		Feature & GAM Press Statistic & LM Press Statistic \\ \hline
+		ppron & 18.591104613 & 20.001055542 \\ \hline
+		social & 21.681732112 & 24.906646115  \\ \hline
+		NounsNormalised & 0.002818941 & 0.002797798\\ \hline
+		function & 46.546071705 & 46.031993602 \\ \hline
+		conj & 8.812330188 & 8.354974826 \\ \hline
+		PronounsNormalised & 0.002284437 & 0.002356129\\ \hline
+		Analytic & 2272.200519272 & 2239.224863610 \\ \hline
+		Pronoun & 44.557282749 & 46.894576916 \\ \hline		
+		NN & 101685.870974277 & 96485.575598505 \\ \hline
+		male & 5.191815443 & 5.278559152 \\ \hline
+		UniqueWords & 255114.636660673 & 238042.427437454 \\ \hline
+		WDT & 1987.149134694 & 1958.781236655 \\ \hline
+		nouns & 276723.222762842 & 262192.120091097 \\ \hline
+		nouns/100 & 27.670762078 & 26.219212009 \\ \hline
+		UniqueStems & 170450.645283172 & 166302.725530118 \\ \hline
+		shehe & 5.188136314 & 5.265385185 \\ \hline
+		VBZ & 13889.871831536 & 13721.175422577 \\ \hline
+		JJ & 24950.440222526 & 24692.043386181 \\ \hline
+		article & 13.500676675 & 13.468083961 \\ \hline	
+		Adjectives & 30210.006270270 & 29865.240707991 \\ \hline	
+		Adjectives.100 & 3.020898615 & 2.986524071 \\ \hline
+		Dic & 40.099289973 & 38.501060715 \\ \hline	
+	\end{tabular}
+	\caption{\label{tab:table-name}Comparison of GAM and Linear Model using the PRESS statistic}
+	\end{center} 
+\end{table}
+
+\section{Discussion}\label{discussion}
+President Reagan received his diagnosis of AD in August 1994 but using transcripts of speeches he made in his two terms as President (January 1981 - January 1989) we have be able to identify certain changes in his use of language that we might ascribe to the onset of MCI and early AD.
+Despite differences in our methodology, our research supports some of the findings of Berisha and Liss in that we find an increase in non-specifc noun usage.
+Compared to our controls (GWHB and DJT), we find some slight trends with GWHB but no such trends with DJT in his speech albeit his samples of speech span a shorter amount of time.
+Interestingly, when we normalised the various types of words used by the presidents we found some interesting patterns that further differentiated RR from the controls.
+Whilst Non-specific nouns increased over time, we found that noun usage in general significantly decreased and pronouns increased similarly significantly.
+The increase in pronoun for those with early AD has been identified in literature, although there are only a few studies that explore this \cite{Wendelstein2015}.
+Wendlestein et al propose that the increased used of pronouns is an expression of an impaired ability to adapt language to the listener's needs \cite{Wendelstein2015}.
+Almor et al attributed this reliance on pronouns due to a impaired working memory \cite{Almor1999}.
+\par 
+The decrease in overall noun usage has also been identified as a feature.
+Jarrold et al found that AD patients would use more pronouns, verbs and fewer nouns than controls \cite{Jarrold2014}.
+Wendlestein in their investigations into noun usage found that decreased later on in AD progression and was unaffected in the pre-clinical stages of AD \cite{Wendelstein2014}.
+Our results are supported by existing literature and this potentially means that language analysis in the way we have structured it may have diagnostic or prognostic properties.
+\par 
+A criticism of Berisha and Liss's work is the problems they had with normalising the transcripts in terms of length.
+This was also a problem in the work of Garrard et al \cite{Garrard2005, Le2011}.
+Whilst it is important to control for outliers, there are other ways in which we can control for length of sample.
+In this paper, we controlled for transcript length by dividing any features that were raw counts by the total length of the transcript.
+When we did this, we found that there was a significant decrease in the number of unique words (R=-0.56, p\textless0.001) used however when we controlled via normalisation, we found that this was not a significant feature (R=-0.172, p=0.25).
+\par 
+Overall, we are able to show differences in the a number of language features between RR, GHWB and DJT over time and the psychological literature confirms our findings in reference to these changes.
+\par
+Our next hypothesis involved exploring whether linear models were the most appropriate way to track this longitudinal data.
+In our analysis, we found there was no significant difference in terms of mean sum of squares errors between the linear model and the generalised additive model however this does not mean that these are equivalent.
+For example, considering the plot below.
+We compare a model fit to a linear model with a generalised additive model which aims to model the data more closely.
+It is clear to see that both models can track the movements of language use over time however, the linear model is quite a rigid model which does not allow for different inflection points.
+We can see that RR's decline is not linear, and that there are periods of relative stability and some periods where the decline is more severe.
+We argue that these more nuanced changes over time can only be modelled by a model such as a GAM.
+\par
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=6cm, height=5cm]{plots/comparison1.png}
+	\caption{Comparing a linear model with a generalised additive model}
+\end{figure}
+
+In her paper on developing diagnositic criteria for Mild Cognitive Impairment, Albert et all \cite{Albert2011} states 'it is important to obtain longitudinal assessments of cognition, whenever possible' and 'obtaining objective evidence of progressive declines in cognition over time is important for establishing the accuracy of the diagnosis, as well as for
+assessing any potential treatment response.' As we have shown, it is not the case that language declines in a linear way and whilst we have a number of data points that we can use to model RR's decline, in a clinical setting it would be impractical to have upwards of 40 data collection points.
+We feel that  it is important that we have more than just two.
+In the figures below, we can see that if we measured language at Reagan's first transcript to a point after 700 days the vast majority of points would point to a decline in language.
+However, the next figure shows what would happen if we measured language from Reagan's second transcript, just 34 days later.
+In this case, we can see that in some cases we would mark Reagan as improving when it is clear that he is not.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=6cm, height=5cm]{plots/comp1.png}
+	\caption{Measuring decline from Reagan's first transcript to all other transcripts greater than 700 days later}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=6cm, height=5cm]{plots/comp2.png}
+	\caption{Measuring decline from Reagan's first transcript to all other transcripts greater than 700 days later}
+\end{figure}
+\par 
+There are limitations of this research.
+Whilst in terms of age, DJT is certainly more suitable as a control to match with RR, in some ways they held very different styles of press conferences in that RR preferred to do solo press conferences and DJT has shown a preference for doing joint press conferences which have an impact on the amount of language produced.
+This artefact of the data is in itself notable as it illustrates the problems we may have with smaller amounts of speech and the problem of finding an appropriate control is a common one in this domain, given the considerable variation in factors such as age, language ability and education.
+As mentioned above, we are lucky enough to have numerous samples of data collected over a long period of time and this is not something that can be easily translated to a clinical protocol.
+Finally, the variability in the number and quality of the transcriptions raises some doubts as to the results of DJT.
+Is it the lack of instances of data, or that the number of words per transcript is significantly less than with RR or GHWB.
+This does have further implications for using a protocol such as this in a more general way as it will impact how data is collected.
+\par 
+With further work, it is not feasible to the vast array of samples over a timeframe, as we have had with the president corpus and so it would be worth exploring how the quality of these predictions may lessen when faced with considerably fewer samples and over a smaller time period.
+It would also be worth extending this research further to encompass more of the linguistic features Fraser used in her work \cite{Fraser2015} to see if there are any further insights to be gained.
+In addition, this replication and extension has demonstrated the potential utility of using longitudinal data as a means of comparing language use of a person at two or more time periods and using this information as a diagnostic aid for MCI and therefore more work would be helpful from a longitudinal perspective to see if this approach may be valid in moving towards a solution for this particular problem.
+
+\section{Conclusions}\label{conclusions}
+The results of this work show that we can track a person's use of language through time in a number of ways and that it is possible for an individual to be his or her own control.
+This is important as it means the heterogenous nature of the MCI population does not impact results as much as if we were comparing those with group of MCI patients with a group of controls.
+Equally, it would be helpful to have controls to ascertain what would be usual to expect in the decline of language in a healthy older adult.
+\par 
+The results of this work also identify some clues as to what could work in a clinical setting, or how we might be able to collect data in a way that accurately tracks language decline without making incorrect assumptions.
+\par
+From a clinical perspective, we can see that using samples (albeit for this dataset) is able to track languages changes over a given time frame.
+This potentially means that we can use a similar methodology to collect regular language samples in settings such as memory clinics and GP's surgeries and even potentially in the people's homes and that these language samples, have the potential to act as an early warning sign for Mild Cognitive Impairment that potentially will identify patients at risk of developing Alzheimer's Disease.

### 2019-12-06 Source/Paper 3 - Delphi Review/DelphiEmail.tex
updated final Delphi Email

@@
-\documentclass{article}
+\documentclass{letter}
+\usepackage{enumerate}
 
@@
 
-The following survey is stage 1 of a Delphi questionnaire.
+The following survey is stage one of a Delphi questionnaire.
 This is designed to obtain your personal opinions relating to the problem described above.
-The Delphi process involves obtaining responses from you on three separate occasions and the we estimate that it should take no more than 30 minutes to respond:
+The Delphi process involves obtaining responses from you on three separate occasions and we estimate that it should take no more than 30 minutes to respond at each stage:
 
@@
 
-Round 2: Your answers (and those from the other panellists) from round 1 will be summarised and formulated into a series of more specific questions that you will be asked to respond to.
+Round Two: Your answers (and those from the other panellists) from round 1 will be summarised and formulated into a series of more specific questions that you will be asked to respond to.
 
-Round 3: Round 2’s questions will be submitted to you again but this time you will also be able to see the average reply of the other panellists and you will then be asked if you would like to adjust your answer from the second round or not.
-The identity of all panelists will remain confidential at all times.
+Round Three: Round Two’s questions will be submitted to you again but this time you will also be able to see the average reply of the other panellists and you will then be asked if you would like to adjust your answer from the second round or not.
+The identity of all panellists will remain confidential at all times.
 
@@
 
-1) It is estimated that between five and ten minutes of spontaneous speech would be an ideal balance between data quality and ease of providing a sample. 
-	a) Do you agree with this? if not, please state why?
-	b) What task(s) do you think would be useful for eliciting an appropriate amount of speech?
-	c) How often should data be collected that will provide a balance between accurate tracking of language use over time whilst reducing intrusiveness?
+1) Literature suggests that between five and ten minutes of spontaneous speech would be an ideal balance between data quality and ease of providing a sample. 
+\begin{enumerate}[(a)]
+	\item Do you agree with this? if not, please state why.
+	\item What task(s) do you think would be useful for eliciting an appropriate amount of speech? For example. Picture Description task
+	\item How often should data be collected that will provide a balance between accurate tracking of language use over time whilst reducing intrusiveness?
+\end{enumerate}
 
-2) The aim of the end application is to provide a way for users to provide multiple samples over a period of time, however as with all longitudinal data collection methods will be subject to participant attrition.
-	a) In your opinion, how might this application encourage continued submissions of language samples?
+2) The aim of the end application is to provide a way for users to provide multiple samples over a period of time, however as with all longitudinal data collection methods it will be subject to participant attrition.
+\begin{enumerate}[(a)]
+	\item In your opinion, how might we encourage continued submissions of language samples?
+\end{enumerate}
 
-3) The aim of creating this mobile application is to facilitate the ease in which a large number of samples can be collected? 
-	a) In your opinion, what problems do you foresee happening with this collection method? 
-	b) Are there any potential solutions to these problems that spring to mind?
-
+3) The aim of creating this mobile application is to facilitate the ease with which a large number of samples can be collected? 
+\begin{enumerate}[(a)]
+	\item In your opinion, what problems do you foresee happening with this collection method? 
+	\item Are there any potential solutions to these problems that spring to mind?
+\end{enumerate}
 Thanks in advance for your time,
@@
 Jomar Alcantara
+
 Aston University

### 2020-01-06 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
updated Presidents Paper

@@
 It affects a significant proportion of the global older adult population and the impact on morbidity and mortality rates is considerable.
-Dementia is currently the leading cause of death in England and Wales and the sixth leading cause of death in the United States (US).
+Dementia including Alzheimer's Disease is currently the leading cause of death in England and Wales \cite{Patel2018} and the sixth leading cause of death in the United States (US) accounting for 32\% of all adult deaths in the US and projected to rise to 43\% by 2050 \cite{Weuve2014} .
 A 2014 report commissioned by the Alzheimer's Society estimated that in the UK by 2015 there would be approximately 855,000 people rising living with dementia increasing to 1 million by 2021 \cite{AlzheimersSociety2014}.
@@
 Currently, a definitive diagnosis for AD can only be produced at post-mortem.
-However, there are a number of psychological and physiological indicators that can indicate that dementia is present.
-From a physiological perspective, researchers have identified two proteins called beta-amyloid and tau have been identified as two of the key proteins that have a role in initiation and progression of AD.
-In a typical case beta-amyloid clumps into plaques which slowly build up between neurons with tau proteins accumulating and eventually forming tangles inside neurons.
-At a certain point, the levels of beta-amyloid rise triggering a more rapid spread of tau throughout the brain.
-Eventually, due to these and other changes, neurons lose their ability to communicate and the brain starts to shrink\cite{Duong2017}.
-This leads to the development of some of the psychological symptoms associated with AD, primarily cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}.
+However, there are a number of psychological and physiological indicators that can indicate that AD is is present.
+These physiological changes lead to the development of some of the psychological symptoms associated with AD, primarily cognitive deficits such as problems with episodic and semantic memory, organizing and planning and other problems with executive function, difficulties with language, and visuospatial deficits \cite{McKhann2011}.
 These cognitive symptoms are often accompanied by emotional problems such as depression and behavioural difficulties.
@@
 All the medications that are currently on the market are designed to manage symptoms.
-Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage of these drugs fail in clinical trial stage of the drug discovery process ((99.6\% failure rate vs 80\% for systemic anti-cancer drugs) \cite{Cummings2014}.
-Cummings et al proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already occurred \cite{Cummings2014}.
+Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage of these drugs fail in the clinical trial stage of the drug discovery process (99.6\% failure rate vs 80\% for systemic anti-cancer drugs) \cite{Cummings2014}.
+Cummings et al proposed that a possible reason for the lack of success is that the drug treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already occurred \cite{Cummings2014}.
 Research focus has now started to shift to the earlier stages of AD (i.e.
@@
 \par
-One of the challenges associated with the early detection of AD is differentiating natural age associated memory impairment and cognitive decline due to aging from with decline due to AD.
-This challenge is often complicated further due to the large variation in the cognitive abilities and educational background of individuals.
+One of the challenges associated with the early detection of AD is differentiating natural age associated memory impairment and cognitive decline due to aging from decline due to AD \cite{Lo2017}.
+This challenge is often complicated further due to the large variation in the cognitive abilities and educational background of individuals \cite{Harada2013}.
 The work of Albert et al helps to address this by the development of clinical criteria which professionals can use to diagnose MCI due to AD.
-One of the most important observations from this piece of work is that a diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}, and this is essentially to place more importance on the trajectory of a person's cognitive abilities rather than a person's cognitive ability in general.
+One of the most important observations from this piece of work is that a diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}.This is essentially to place more importance on the trajectory of a person's cognitive abilities rather than a person's cognitive ability in general.
+% Should we say something about the US presidents project.
+I know you descript it properly in a page or two but the point here is that although speeches easily collected, there is little in the piblic domain that is of good fidelity and that improptu speech of someone who went on to develop AD.
+
 \par 
-There has been a significant research in the area of language deterioration as a method of detecting AD at an earlier stage.
+There has been significant research in the area of language deterioration as a method of detecting AD at an earlier stage.
 This usually takes the form of recording speech whilst patients' undertake a cognitive assessment such as the Picture Description Task \cite{Fraser2015}.
 Given that language samples are relatively easy to collect, research has moved towards analysis of spontaneous speech.
-The work of Berisha and Liss is a good example of this, this study compared the differences in language use between two US presidents, Ronald Reagan (RR), who would go on to receive a diagnosis of Dementia and George H.
+The work of Berisha and Liss is a good example of this, their study compared the differences in language use between two US presidents, Ronald Reagan (RR), who would go on to receive a diagnosis of Dementia and George H.
 W.
@@
 \par   
-Berisha and Liss have developed some very interesting ideas about how we might track changes over time in various lexical features which have been associated with the development of Alzheimer's Disease.
-However we feel that cognitive decline in pre-clinical AD is not accurately modeled as a linear process and therefore we explored the application of generalised additive models (GAMs) to this data which have no assumed understanding of the distribution of the data.
+Berisha and Liss have developed interesting ideas about how we might track changes over time in various lexical features which have been associated with the development of Alzheimer's Disease.
+However we feel that cognitive decline in pre-clinical AD is not accurately modeled as a linear process.
+
+ and therefore we explored the application of generalised additive models (GAMs) to this data which have no assumed understanding of the distribution of the data.
 We refine these ideas with the aim of exploring the potential for a protocol that can be used for analyzing language deterioration.
@@
 \section{Methodology}\label{methodology}
-We took 46 transcripts of press conferences given by RR (from 1981 to 1988) and compared them with 134 press conferences (from 1989 to 1993) given by GHWB and 29 press conferences (from 2016 to 2019) conducted by DJT.
+We took the transcripts of 46 press conferences given by RR (from 1981 to 1988) and compared them with 134 press conferences (from 1989 to 1993) given by GHWB and 29 press conferences (from 2016 to 2019) conducted by DJT.
 We analysed transcripts for language features (described below) shown to change longitudinally with AD.
@@
 \subsection{Pre-processing}
-To generate the files necessary for analysis, we downloaded each transcript and performed the following changes.
+To generate the files necessary for analysis, we downloaded each transcript and pre-processed each transcript in the following way.
 We omitted the prepared statement by the president and any speech by other individuals and started each transcript at the beginning of the first answer to a question by a member of the press.
 We filtered any annotations that were added to the transcript, including any references or clarifications, and any laughter.
-It's worth noting that there appears to be a difference in how 'hesitations' were marked down between each president, for RR \& DJT hesitations were marked by a single hyphen whereas for GHWB hesitations are marked by a double hyphen.
-In order to maintain consistency when parsing through the documents, We have changed both types of hesitation to be marked by a single hyphen.
+It is worth noting that there appears to be a difference in how 'hesitations' were marked down between each president.
+For RR \& DJT hesitations were marked by a single hyphen whereas for GHWB hesitations are marked by a double hyphen.
+In order to maintain consistency when parsing through the documents, we have changed both types of hesitation to be marked by a single hyphen.
 We also omitted one word answers to questions as this data would, from a theoretical perspective, not be relevant for language analysis.
 We did not make any alterations to the length of the document, but instead generated features which would normalise by the length of the document.
-We therefore was able to include all press conferences by all presidents analysed where there was a question and answer session conducted at least in part by the president in question (2 press conferences of GHWB were omitted due to a lack of a question and answer session).
+We therefore were able to include all press conferences by the presidents analysed where there was a question and answer session conducted at least in part by the president (2 press conferences of GHWB were omitted due to a lack of a question and answer session).
 
@@
 Features were generated by running each transcript through a number of natural language processing libraries.
-We used the Natural Language ToolKit (NLTK) package \cite{Bird2009} in python to stem each transcript using the Snowball Stemmer and also completed a part of speech (POS) tagging process.
+We used the Natural Language ToolKit (NLTK) package \cite{Bird2009} in Python to stem each transcript using the Snowball Stemmer and also completed a part of speech (POS) tagging process.
 We also ran each transcript through Linguistic Inquiry and Word Count (LIWC) \cite{Pennebaker2015} software.
@@
 \subsubsection{Measures of lexical diversity}
-Lexical diversity is a measure of the variety of language used within a given document.
+Lexical diversity is a measure of the variety of language used within a given document and this is commonly used in the literature to measure language declinein those with MCI and AD \cite{Bucks2000, Le2012}.
 A document is said to have high lexical diversity if the number of unique words is large.
@@
 We also used the TTR formula, Brunet's Index and Honore's Statistic as other measures of lexical diversity \cite{Bucks2000}.
+For the formula's below, we use the following notation.
+N = total number of words, V = number of unique words and V1 = the number of words spoken exactly once.
 \par 
@@
 \begin{equation} \label{x1}
-TTR = numberOfUniqueWords / totalNumberOfWords.
+TTR = V / N
 \end{equation}
@@
 \begin{equation} \label{x2}
-W = N^{V(-0.165)}
+W = N^{-0.165V}
 \end{equation}
-where N is the total length of the utterance being measured and V is equal to the total vocabulary being used by the subject.
-Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers.
-\newline
+ Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers. \newline
 \par 
@@
 \end{equation}
-where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
 \subsubsection{Fillers, Non-Specific Nouns and Low Imageability Verbs}
@@
 
-\subsubsection{Longitudinal Analysis}
+\subsection{Longitudinal Analysis}
 As with Berisha and Liss\cite{Berisha2015}, we analysed the transcripts over time in order to determine if there was an underlying temporal trend that may indicate that different parts of language increase or decrease over time.
@@
 \par
-We looked also compared the results of linear models and a non linear model in terms of best fit.
-The non linear model we used was the generalised additive model \cite{Hastie1986}.
-The selection of this non linear model was due to the authors not wanting to assume an underlying distribution of the data and therefore this was the most appropriate model to use.
+We looked also compared the results of linear models and a non linear model in terms of best fit. The non linear model we used was the generalised additive model \cite{Hastie1986}.
+There is evidence to support the idea that language deterioration.
+In figure 1, we look at the difference of between a linear model of Reagan's use of personal pronouns vs a non linear model.
+Whilst they are broadly similar in that they track an increase in his use of personal pronouns, they do not capture the variation in this increase as closely as a non linear model might.
+We therefore decided to apply this to all the features we extracted.
 
 \section{Results}\label{results}
-One of the most important thing to note is the wide variety of samples between the three presidents and also the varying timescales.
+Two of the most important things to note are the wide variety of samples between the three presidents and also the varying timescales.
 RR participated in 46 press conferences over eight years (an average of 5.75 a year) which is the fewest number of press conferences given by an American president during their term of office.
 GHWB participated in 136 press conferences over four years (an average of 34 a year) and DJT participated in 29 press conferences to date (an average of 19.3 per year).
-Equally, there are differences in the average number of words.
-RR produced an average of 3424 words per conference compared to 2608 by GHWB (unpaired t = 4.434, p\textless 0.001) and DJT at 1849 words (unpaired t = 6.524, p\textless 0.001).
 
@@
 		& RR & GHWB & DJT \\ \hline
+		Sample Size & 46 & 136 & 29 \\ \hline
 		Total Words & 3423.91 (416.42) & 2607.72 (1210.38) & 1848.65 (1549.38) \\ \hline
@@
 
-In terms of more specific language differences between the presidents, we found that RR used significantly more unique words, non-specific nouns and low imageability verbs than GHWB and DJT (see Table 3).
+In terms of more specific language differences between the presidents, as we found that RR used significantly more unique words, non-specific nouns and low imageability verbs than GHWB and DJT (see Table 3).
 The mean length of utterance for RR was significantly greater than that of GHWB and DJT.
@@
 
-\subsection{Longitudinal Analysis}
+\subsection{Longitudinal Analysis - Comparison of Ronald Reagan and George H.W. Bush}
 We then looked at the data from a longitudinal perspective as we were interested seeing whether we can track various language variables and their progress over time.
-We ran a number of Pearsons correlations with transcript index number as a time reference and the dependant variables (Table 4).
+We ran a number of Pearsons correlations with number of days as a time reference (we calculated this as the total number of days from the first sample) and the dependant variables.
+We also calculated these Pearsons correlations with transcript index number as a time reference and the dependant variables as a replication of the Berisha and LIss study \cite{Berisha2015} and these results are described in the supplementary material.
+
 For our controls, we found them to be stable for the most part with the main highlights being a decrease in Adverb usage for DJT (R=-0.36, p=0.049) and a steady but not severe decline in a number of variables for GWHB, namely total word count, unique words, low imageability words and verb usage.
@@
 We noticed an significant increase in adverb (R=0.41, p=0.004) and pronoun usage (R=0.65, p\textless0.001), as well as a slight usage increase in Non-specific nouns(R=0.30, p=0.03).
-There was a highly significant decrease in number of unique words (R=-0.56, p\textless0.001) and noun usage (R=-0.70, p\textless0.001).
+There was also a significant decrease in number of unique words (R=-0.56, p\textless0.001) and noun usage (R=-0.70, p\textless0.001).
 Also very significant decrease in adjective usage (R=-0.40, p=0.005) and a significant decrease in total word count (R=-0.31, p=0.03).
-
-\begin{table}[H]
-	\begin{center}
-	\begin{tabular}{ | l | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
-		\hline
-		& RR & GHWB & DJT \\ \hline
-		Word Count & \textbf{-0.31*} & \textbf{-0.21*} & 0.08 \\ \hline 
-		Unique Words & \textbf{-0.56***} & \textbf{-0.25**} & 0.16 \\ \hline
-		Non Specific Nouns & \textbf{0.30*} & -0.08 & -0.03 \\ \hline
-		LI Verbs & -0.19 & \textbf{-0.20**} & 0.02 \\ \hline
-		Nouns Normalised & \textbf{-0.70***} & -0.03 & 0.14 \\ \hline
-		Verbs Normalised & \textbf{0.36**} & \textbf{0.24***} & -0.03 \\ \hline
-		Adjectives Normalised & \textbf{-0.40**} & 0.08 & -0.34 \\ \hline
-		Adverbs Normalised & \textbf{0.41***} & 0.02 & \textbf{-0.36*} \\ \hline
-		Pronouns Normalised & \textbf{0.65***} & 0.13 & 0.07 \\ \hline
-		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
-    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
-    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
-	\end{tabular}
-	\caption{\label{tab:table-name}Pearson Correlations for Features}
-	\end{center} 
-\end{table}
-
-To get a better idea of decline over time, we also ran Pearson's correlations using time elapsed in days from the first transcript instead a transcript index.
-This was to control for variation in the number of days between transcripts.
-Given the number of hypotheses being tested, we felt it was necessary to apply some correction for family wise error rate and therefore a Hochberg step-up procedure was applied to the results of the analysis.
-We also used the Benjamini-Yekutieli procedure for controlling for False Discovery rate and came up with a final list of important features for each president.
-For RR, this was 22 features, for GHWB this was 14 features and there were no significant features for DJT after applying these methods.
-The features were generated from multiple sources and therefore there was some significant overlap between the features but we have included all features in Table 6.
+\par
+Given the number of hypotheses being tested, we felt it was necessary to control for false discovery rate and therefore Benjamii-Yekutieli procedure was applied to the results of the analysis.
+By applying this, we produced a final list of important features for each president.
+For RR, this was 22 features and for GHWB this was 14 features.
+The features were generated from multiple sources and therefore there was some significant overlap between the features but for completeness we have included all features in Table 6.
 
@@
 
+\subsection{Comparing linear models and non-linear models}
 In terms of comparing the general additive model and linear models.
@@
 \par
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=6cm, height=5cm]{plots/comparison1.png}
-	\caption{Comparing a linear model with a generalised additive model}
-\end{figure}
 
@@
 
+%\begin{table}[H]
+%	\begin{center}
+%	\begin{tabular}{ | l | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+%		\hline
+%		& RR & GHWB & DJT \\ \hline
+%		Word Count & \textbf{-0.31*} & \textbf{-0.21*} & 0.08 \\ \hline 
+%		Unique Words & \textbf{-0.56***} & \textbf{-0.25**} & 0.16 \\ \hline
+%		Non Specific Nouns & \textbf{0.30*} & -0.08 & -0.03 \\ \hline
+%		LI Verbs & -0.19 & \textbf{-0.20**} & 0.02 \\ \hline
+%		Nouns Normalised & \textbf{-0.70***} & -0.03 & 0.14 \\ \hline
+%		Verbs Normalised & \textbf{0.36**} & \textbf{0.24***} & -0.03 \\ \hline
+%		Adjectives Normalised & \textbf{-0.40**} & 0.08 & -0.34 \\ \hline
+%		Adverbs Normalised & \textbf{0.41***} & 0.02 & \textbf{-0.36*} \\ \hline
+%		Pronouns Normalised & \textbf{0.65***} & 0.13 & 0.07 \\ \hline
+%		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+%    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+%    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+%	\end{tabular}
+%	\caption{\label{tab:table-name}Pearson Correlations for Features}
+%	\end{center} 
+%\end{table}
+
 \end{document}

### 2020-01-06 Source/Paper 3 - Delphi Review/DelphiEmail.tex
edited Delphi Email

--- 2019-12-06 Source/Paper 3 - Delphi Review/DelphiEmail.tex
@@
 
-Round 2: Your answers (and those from the other panellists) from round 1 will be summarised and formulated into a series of more specific questions that you will be asked to respond to.
+Round Two: Your answers (and those from the other panellists) from round 1 will be summarised and formulated into a series of more specific questions that you will be asked to respond to.
 
-Round 3: Round 2’s questions will be submitted to you again but this time you will also be able to see the average reply of the other panellists and you will then be asked if you would like to adjust your answer from the second round or not.
+Round Three: The questions from the previous round will be submitted to you again but this time you will also be able to see the average reply of the other panellists and you will then be asked if you would like to adjust your answer from the second round or not.
+
 The identity of all panelists will remain confidential at all times.
@@
 A reply to these by email is fine.
-
+\par
+\itshape
 1) It is estimated that between five and ten minutes of spontaneous speech would be an ideal balance between data quality and ease of providing a sample. 
-	a) Do you agree with this? if not, please state why?
-	b) What task(s) do you think would be useful for eliciting an appropriate amount of speech?
-	c) How often should data be collected that will provide a balance between accurate tracking of language use over time whilst reducing intrusiveness?
+\begin{itemize}
+	\item Do you agree with this? if not, please state why?
+	\item What task(s) do you think would be useful for eliciting an appropriate amount of speech?
+	\item How often should data be collected that will provide a balance between accurate tracking of language use over time whilst reducing intrusiveness?
+\end{itemize}
 
 2) The aim of the end application is to provide a way for users to provide multiple samples over a period of time, however as with all longitudinal data collection methods will be subject to participant attrition.
-	a) In your opinion, how might this application encourage continued submissions of language samples?
+\begin{itemize}	
+	\item In your opinion, how might this application encourage continued submissions of language samples?
+\end{itemize}
 
 3) The aim of creating this mobile application is to facilitate the ease in which a large number of samples can be collected? 
-	a) In your opinion, what problems do you foresee happening with this collection method? 
-	b) Are there any potential solutions to these problems that spring to mind?
-
+\begin{itemize}	
+	\item In your opinion, what problems do you foresee happening with this collection method? 
+	\item Are there any potential solutions to these problems that spring to mind?
+\end{itemize}	
+\par 
 Thanks in advance for your time,

### 2020-01-06 Source/Paper 3 - Delphi Review/DelphiEmail.tex

--- 2019-12-06 Source/Paper 3 - Delphi Review/DelphiEmail.tex
@@
 
-Round Three: Round Two’s questions will be submitted to you again but this time you will also be able to see the average reply of the other panellists and you will then be asked if you would like to adjust your answer from the second round or not.
+Round Three: The questions from the previous round will be submitted to you again but this time you will also be able to see the average reply of the other panellists and you will then be asked if you would like to adjust your answer from the second round or not.
 The identity of all panellists will remain confidential at all times.
+
+The identity of all panelists will remain confidential at all times.
 
@@
 A reply to these by email is fine.
+\par
+\itshape
 
@@
 \end{enumerate}
+
 Thanks in advance for your time,

### 2020-01-06 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
edited Presidents Corpus

@@
 		\and
-	\textbf{Felipe Campelo Franca Pinto}\\
+	\textbf{Felipe Campelo}\\
 	Department of Computer Science\\
@@
 		& RR v GHWB & RR v DJT & GHWB v DJT \\ \hline
-		Total Words & \textbf{4.434***} & \textbf{6.524***} & \textbf{2.899**} \\ \hline
-		Unique Stems & \textbf{10.878***} & \textbf{10.111***} & \textbf{4.148***} \\ \hline
-		Mean Length of Utterance & \textbf{16.175***} & \textbf{25.084***} & \textbf{13.484***} \\ \hline	
-		Non Specific Nouns & \textbf{7.877***} & \textbf{3.426**} & -0.574 \\ \hline
-		LI Verbs & \textbf{2.656**} & \textbf{3.420***} & 1.628 \\ \hline
-		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
-		\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
-		\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+		Total Words & 6.6751 (574.8808, 1057.5324) & 5.352 (974.6629, 2175.8489) & 2.4774 (137.5811, 1380.5176) \\ \hline
+		Unique Stems & 10.878 (161.1580, 232.6168) & 10.11 (275.4688, 414.0975) & 4.148 (75.86722, 219.92432) \\ \hline
+		Mean Length of Utterance & 16.175 (3.915869, 5.009114) & 25.084 (8.453404, 9.922328)  & 13.484 (4.021601, 5.429148) \\ \hline	
+		Non Specific Nouns & 9.2052 (5.881447, 9.13725) & 3.8627 (3.275567, 10.474625) & -0.37956 (-4.037218, 2.770237) \\ \hline
+		LI Verbs & 4.0434 (10.84689, 31.53291) & 2.804 (10.91107, 67.41231) & 1.2836 (-11.06486, 49.00844)\\ \hline
 	\end{tabular}
-	\caption{\label{tab:table-name}RR T-tests vs GWB and DJT}
+	\caption{\label{tab:table-name}RR T-tests vs GWB and DJT, T-statistic and 95\%confidence intervals}
 	\end{center} 
@@
 
-\subsection{Longitudinal Analysis - Comparison of Ronald Reagan and George H.W. Bush}
+\subsection{Longitudinal Analysis}
+\subsubsection{Comparison of Ronald Reagan and George H.W. Bush}
 We then looked at the data from a longitudinal perspective as we were interested seeing whether we can track various language variables and their progress over time.
@@
 	\begin{center}
-	\begin{tabular}{ | l | c | c | c |}
+	\begin{tabular}{ | l | c | c |}
 		\hline
-		& RR R-Squared  & GWB R-Squared & DJT R-Squared \\ \hline
-		ppron & \textbf{0.700***} & 0.007 & -0.196 \\ \hline
-		social & \textbf{0.698***} & 0.204 & 0.032 \\ \hline
-		NounsNormalised & \textbf{-0.689***} & -0.023 & 0.194 \\ \hline
-		function & \textbf{0.670***} & -0.169 & -0.243 \\ \hline
-		conj &\textbf{0.644***} & -0.452 & -0.187 \\ \hline
-		PronounsNormalised & \textbf{0.631***} & 0.131 & 0.002 \\ \hline
-		Analytic & \textbf{-0.626***} & -0.013 & 0.366 \\ \hline
-		NN & \textbf{-0.601***} & -0.197 & 0.116 \\ \hline
-		male & \textbf{0.585***} & 0.024 & 0.042 \\ \hline
-		UniqueWords & \textbf{-0.578***} & -0.257 & 0.204 \\ \hline
-		WDT & \textbf{-0.577***} & -0.173 & -0.105 \\ \hline
-		shehe & \textbf{0.529*} & -0.005 & 0.064 \\ \hline
-		VBZ & \textbf{-0.521*} & -0.157 & 0.136 \\ \hline
-		JJ & \textbf{-0.518*} & -0.206 & 0.062 \\ \hline
-		Fillers & -0.076 & \textbf{-0.359**}  & -0.032 \\ \hline	
-		EX & -0.193* & \textbf{-0.337**}  & 0.339 \\ \hline	
-		achieve & -0.248 & \textbf{0.334**}  & 0.169 \\ \hline		
-		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
-    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
-    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+		& RR R-Squared  & GWB R-Squared \\ \hline
+		ppron & \textbf{0.700***} & 0.007 \\ \hline
+		social & \textbf{0.698***} & 0.204 \\ \hline
+		NounsNormalised & \textbf{-0.689***} & -0.023 \\ \hline
+		function & \textbf{0.670***} & -0.169 \\ \hline
+		conj &\textbf{0.644***} & -0.452 \\ \hline
+		PronounsNormalised & \textbf{0.631***} & 0.131 \\ \hline
+		Analytic & \textbf{-0.626***} & -0.013 \\ \hline
+		NN & \textbf{-0.601***} & -0.197 \\ \hline
+		male & \textbf{0.585***} & 0.024 \\ \hline
+		UniqueWords & \textbf{-0.578***} & -0.257 \\ \hline
+		WDT & \textbf{-0.577***} & -0.173 \\ \hline
+		shehe & \textbf{0.529*} & -0.005 \\ \hline
+		VBZ & \textbf{-0.521*} & -0.157 \\ \hline
+		JJ & \textbf{-0.518*} & -0.206 \\ \hline
+		Fillers & -0.076 & \textbf{-0.359**} \\ \hline	
+		EX & -0.193* & \textbf{-0.337**} \\ \hline	
+		achieve & -0.248 & \textbf{0.334**} \\ \hline		
+		\multicolumn{3}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+    	\multicolumn{3}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+    	\multicolumn{3}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+	\end{tabular}
+	\caption{\label{tab:table-name}Pearson Correlations for Features}
+	\end{center} 
+\end{table}
+
+\subsubsection{Comparison of Ronald Reagan and Donald J. Trump}
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | l | c | c |}
+		\hline
+		& RR R-Squared & DJT R-Squared \\ \hline
+		ppron & \textbf{0.700***} & 0.007 \\ \hline
+		social & \textbf{0.698***} & 0.204 \\ \hline
+		NounsNormalised & \textbf{-0.689***} & -0.023 \\ \hline
+		function & \textbf{0.670***} & -0.169 \\ \hline
+		conj &\textbf{0.644***} & -0.452 \\ \hline
+		PronounsNormalised & \textbf{0.631***} & 0.131 \\ \hline
+		Analytic & \textbf{-0.626***} & -0.013 \\ \hline
+		NN & \textbf{-0.601***} & -0.197 \\ \hline
+		male & \textbf{0.585***} & 0.024 \\ \hline
+		UniqueWords & \textbf{-0.578***} & -0.257 \\ \hline
+		WDT & \textbf{-0.577***} & -0.173 \\ \hline
+		shehe & \textbf{0.529*} & -0.005 \\ \hline
+		VBZ & \textbf{-0.521*} & -0.157 \\ \hline
+		JJ & \textbf{-0.518*} & -0.206 \\ \hline
+		Fillers & -0.076 & \textbf{-0.359**} \\ \hline	
+		EX & -0.193* & \textbf{-0.337**} \\ \hline	
+		achieve & -0.248 & \textbf{0.334**} \\ \hline		
+		\multicolumn{3}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+    	\multicolumn{3}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+    	\multicolumn{3}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
 	\end{tabular}
@@
 \section{Discussion}\label{discussion}
-President Reagan received his diagnosis of AD in August 1994 but using transcripts of speeches he made in his two terms as President (January 1981 - January 1989) we have be able to identify certain changes in his use of language that we might ascribe to the onset of MCI and early AD.
+President Reagan received his diagnosis of AD in August 1994 but using transcripts of speeches he made in his two terms as President (January 1981 - January 1989) we have been able to identify a number of changes in his use of language that we might ascribe to the onset of MCI and early AD.
 Despite differences in our methodology, our research supports some of the findings of Berisha and Liss in that we find an increase in non-specifc noun usage.
-Compared to our controls (GWHB and DJT), we find some slight trends with GWHB but no such trends with DJT in his speech albeit his samples of speech span a shorter amount of time.
+Compared to our controls (GWHB and DJT), we find some slight trends with GWHB but no such trends with DJT in his speech albeit his samples of speech span a shorter period of time.
 Interestingly, when we normalised the various types of words used by the presidents we found some interesting patterns that further differentiated RR from the controls.
@@
 Wendlestein et al propose that the increased used of pronouns is an expression of an impaired ability to adapt language to the listener's needs \cite{Wendelstein2015}.
-Almor et al attributed this reliance on pronouns due to a impaired working memory \cite{Almor1999}.
+Almor et al attributed this reliance on pronouns due to an impaired working memory \cite{Almor1999}.
 \par 
 The decrease in overall noun usage has also been identified as a feature.
-Jarrold et al found that AD patients would use more pronouns, verbs and fewer nouns than controls \cite{Jarrold2014}.
-Wendlestein in their investigations into noun usage found that decreased later on in AD progression and was unaffected in the pre-clinical stages of AD \cite{Wendelstein2014}.
+Jarrold et al found that AD patients would use more pronouns, more verbs and fewer nouns than controls \cite{Jarrold2014}.
+Wendlestein in their investigations into noun usage found that noun usage decreased later on in AD progression and was unaffected in the pre-clinical stages of AD \cite{Wendelstein2014}.
 Our results are supported by existing literature and this potentially means that language analysis in the way we have structured it may have diagnostic or prognostic properties.
@@
 In this paper, we controlled for transcript length by dividing any features that were raw counts by the total length of the transcript.
-When we did this, we found that there was a significant decrease in the number of unique words (R=-0.56, p\textless0.001) used however when we controlled via normalisation, we found that this was not a significant feature (R=-0.172, p=0.25).
+When we did this, we found that there was a significant decrease in the number of unique words (R=-0.56, p\textless0.001) used.
+However when we controlled via normalisation as described above, we found that this was not a significant feature (R=-0.172, p=0.25).
 \par 
-Overall, we are able to show differences in the a number of language features between RR, GHWB and DJT over time and the psychological literature confirms our findings in reference to these changes.
+Overall, we are able to show differences in the a number of language features between RR, GHWB and DJT over time and the psychological literature confirms our findings with reference to these changes.
 \par
 Our next hypothesis involved exploring whether linear models were the most appropriate way to track this longitudinal data.
-In our analysis, we found there was no significant difference in terms of mean sum of squares errors between the linear model and the generalised additive model however this does not mean that these are equivalent.
-For example, considering the plot below.
+In our analysis, we found there was no significant difference in terms of mean sum of squares errors between the linear model and the generalised additive model.
+However this does not mean that these are equivalent.
+For example, consider Figure 2.
 We compare a model fit to a linear model with a generalised additive model which aims to model the data more closely.
@@
 \par
-
 In her paper on developing diagnositic criteria for Mild Cognitive Impairment, Albert et all \cite{Albert2011} states 'it is important to obtain longitudinal assessments of cognition, whenever possible' and 'obtaining objective evidence of progressive declines in cognition over time is important for establishing the accuracy of the diagnosis, as well as for
@@
 We feel that  it is important that we have more than just two.
-In the figures below, we can see that if we measured language at Reagan's first transcript to a point after 700 days the vast majority of points would point to a decline in language.
+In the figure 2 and figure 3 below, we can see that if we measured language at Reagan's first transcript to a point after 700 days the vast majority of points would point to a decline in language.
 However, the next figure shows what would happen if we measured language from Reagan's second transcript, just 34 days later.
@@
 \par 
-With further work, it is not feasible to the vast array of samples over a timeframe, as we have had with the president corpus and so it would be worth exploring how the quality of these predictions may lessen when faced with considerably fewer samples and over a smaller time period.
+With further work, it is not feasible to have a vast array of samples over a timeframe, as we have had with the president corpus and so it would be worth exploring how the quality of these predictions may lessen when faced with considerably fewer samples and over a smaller time period.
 It would also be worth extending this research further to encompass more of the linguistic features Fraser used in her work \cite{Fraser2015} to see if there are any further insights to be gained.

### 2020-01-06 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
edited Presidents Corpus

@@
 Several research efforts aimed at improving the early diagnosis of Alzheimer's Disease have focused on the longitudinal analysis of lexical features which have been implicated in the pre-clinical progression of this disease.
-We build on these studies and continue to refine a method for extracting and analysing language from transcripts.
-This method is based on a statistical evaluation of the most relevant features, as well as on the use of Generalised Additive Models instead of Linear Regression as a way to more adequately model the declines in language over time.
-We evaluate this method using the unscripted question and answer portions of news conferences from three U.S Presidents: Ronald Reagan (RR), who would later go on to receive a diagnosis of Alzheimer's Disease; George H.
+We build on these studies to develop a refined method for extracting and analysing language from transcripts.
+Our method is based on a statistical evaluation of the most relevant features, as well as on the use of Generalised Additive Models instead of Linear Regression as a way to more adequately model the declines in language over time.
+We evaluate our method using the unscripted question and answer portions of news conferences from three U.S Presidents: Ronald Reagan (RR), who would later go on to receive a diagnosis of Alzheimer's Disease; George H.
 W.
@@
 Our controls were the closest in age to RR when they became president.
-Our analysis uncovered several lexical features that did not change significantly for GB and DT, but presented significant changes for RR, including significant decreases in unique words and noun usage, and significant increases in pronouns, conjunctions and words which are related to social processes significantly increase over time.
-We also show that non-linear models provide a more appropriate way to model these changes over time.
-These findings are consistent with the psychological literature, which documents a link between these features and cognitive decline in Alzheimer's Disease.
-This work points the way to developing a comparatively non-invasive protocol that could be used in clinical settings that help to spot atypical declines in language use over time, potentially contributing to earlier, more accurate diagnoses of Alzheimer's Disease.
+Our analysis uncovered several lexical features that did not change significantly for GB and DT, but presented significant changes for RR.
+These included decreases in unique words and noun usage, and significant increases in pronouns and conjunctions and also words which are related to social processes significantly increase over time.
+We also show that non-linear models provide a more detailed way to model these changes over time that is sensitive to variability in the degrees of change that is typical of AD.
+Our findings are consistent with the psychological literature, which documents a link between these features and cognitive decline in Alzheimer's Disease.
+This work demonstrates that there are potentially a wider number of language markers that researchers can track, which may indicate the presence of mild cognitive impairment as well as illustrating a more nuanced way of modelling these changes via Generalised Additive Models.
 \end{abstract}
@@
 AD is a progressive neurodegenerative disease and is the most common type of dementia, responsible for approximately 60\% to 80\% of all cases \cite{Duong2017}.
-Currently, a definitive diagnosis for AD can only be produced at post-mortem.
+Currently, a definitive diagnosis for AD can only be produced post-mortem.
 However, there are a number of psychological and physiological indicators that can indicate that AD is is present.
@@
 One of the most important observations from this piece of work is that a diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}.This is essentially to place more importance on the trajectory of a person's cognitive abilities rather than a person's cognitive ability in general.
-% Should we say something about the US presidents project.
-I know you descript it properly in a page or two but the point here is that although speeches easily collected, there is little in the piblic domain that is of good fidelity and that improptu speech of someone who went on to develop AD.
-
+Berisha and Liss used the transcripts of press conferences of Ronald Reagan who would go on to be diagnosed with AD, and George Bush Sr who had no diagnosis.
+These speeches are in the public domain and provide a good opportunity of analysing the spontaneous language use of someone with AD, before a diagnosis was made.
 \par 
@@
 \par
+% Stronger motivation
 Berisha and Liss have developed interesting ideas about how we might track changes over time in various lexical features which have been associated with the development of Alzheimer's Disease.
-However we feel that cognitive decline in pre-clinical AD is not accurately modeled as a linear process.
-
- and therefore we explored the application of generalised additive models (GAMs) to this data which have no assumed understanding of the distribution of the data.
+However we feel that cognitive decline in pre-clinical AD is not accurately modeled as a linear process and therefore we explored the application of generalised additive models (GAMs) to this data which have no assumed understanding of the distribution of the data.
 We refine these ideas with the aim of exploring the potential for a protocol that can be used for analyzing language deterioration.
@@
 For RR \& DJT hesitations were marked by a single hyphen whereas for GHWB hesitations are marked by a double hyphen.
-In order to maintain consistency when parsing through the documents, we have changed both types of hesitation to be marked by a single hyphen.
+In order to maintain consistency when parsing through the documents we changed both types of hesitation to be marked by a single hyphen.
 We also omitted one word answers to questions as this data would, from a theoretical perspective, not be relevant for language analysis.
 We did not make any alterations to the length of the document, but instead generated features which would normalise by the length of the document.
-We therefore were able to include all press conferences by the presidents analysed where there was a question and answer session conducted at least in part by the president (2 press conferences of GHWB were omitted due to a lack of a question and answer session).
+We were therefore able to include all press conferences by the presidents analysed where there was a question and answer session conducted at least in part by the president (2 press conferences of GHWB were omitted due to a lack of a question and answer session).
 
 \subsection{Feature Generation}
+% We selected the type of processing as
 Features were generated by running each transcript through a number of natural language processing libraries.
@@
 Firstly we looked at the number of unique words.
-To do this we were able to split each transcript into individual words and changed them to lowercase using NLTK and were then count the number of unique words that appeared in each transcript.
+To do this we split each transcript into individual words and changed them to lowercase using the NLTK package and were then count the number of unique words that appeared in each transcript.
 We also used the TTR formula, Brunet's Index and Honore's Statistic as other measures of lexical diversity \cite{Bucks2000}.
@@
 At a basic level this can be divided into the eight defined parts of speech: 'nouns', 'pronouns', verbs', 'adjectives', 'adverbs', 'conjunctions', 'prepositions' and 'interjections' but can be further subcategorised.
-We used the PoS tagger built into NLTK to tag each transcript in turn and used these the counts from each of these eight categories in our analysis.
+We used the PoS tagger built into NLTK to tag each transcript in turn and used the counts from each of these eight categories in our analysis.
 In addition to frequency counts we also normalised these features by dividing the frequency count by the number of words in the document to take into account transcript length.
@@
 \subsubsection{Linguistic Inquiry and Word Count(LIWC)}
-The LIWC is a text processor which analysis words within a given a document and compares the words in the LIWC dictionary file.
+The LIWC is a text processor / semantic tagger which analysis words within a given a document and compares the words in the LIWC dictionary file.
 The LIWC dictionary file contains over 6000 words which are categorised into approximately 90 different types.
 For example the word 'cried' is part of five word categories: sadness, negative emotion, overall affect, verbs and past focus \cite{Pennebaker2015}.
-Each word in the document or set of documents is searched for in the dictionary file and if found, the appropriate type is incremented by 1.
+Each word in the document or set of documents is searched for in the dictionary file and if found, the appropriate types is incremented by 1.
 What is output is an analysis of word usage in reference to each category.
@@
 Some of these differences are due to the length of the sample, particularly in the case of DJT where his average sample is almost half the sample of RR.
-It could also be said that this could be down to differences in linguistic abilities or speaking style \cite{Berisha2015, Le2011}.
+It could also be said that this could be due to differences in linguistic abilities or speaking style \cite{Berisha2015, Le2011}.
 However, we can certainly see that as controls GWHB and DJT are comparative in relation to non-specific nouns and LI verbs.
@@
 \subsubsection{Comparison of Ronald Reagan and George H.W. Bush}
-We then looked at the data from a longitudinal perspective as we were interested seeing whether we can track various language variables and their progress over time.
-We ran a number of Pearsons correlations with number of days as a time reference (we calculated this as the total number of days from the first sample) and the dependant variables.
+We then looked at the data from a longitudinal perspective as we were interested in seeing whether we can track various language features and their progress over time.
+We ran a number of Pearsons correlations with number of days as a time reference (we calculated this as the total number of days from the first sample) and the dependent variables.
 We also calculated these Pearsons correlations with transcript index number as a time reference and the dependant variables as a replication of the Berisha and LIss study \cite{Berisha2015} and these results are described in the supplementary material.
@@
 There was also a significant decrease in number of unique words (R=-0.56, p\textless0.001) and noun usage (R=-0.70, p\textless0.001).
-Also very significant decrease in adjective usage (R=-0.40, p=0.005) and a significant decrease in total word count (R=-0.31, p=0.03).
+Also there was a significant decrease in adjective usage (R=-0.40, p=0.005) and a significant decrease in total word count (R=-0.31, p=0.03).
 \par
@@
 \end{table}
+% Needs explanation
 

### 2020-02-20 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
updated Presidentspaper

@@
 \par
-% Stronger motivation
 Berisha and Liss have developed interesting ideas about how we might track changes over time in various lexical features which have been associated with the development of Alzheimer's Disease.
@@
 We refine these ideas with the aim of exploring the potential for a protocol that can be used for analyzing language deterioration.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=6cm, height=5cm]{plots/comparison1.png}
+	\caption{Comparing linear model with a non-linear model}
+\end{figure}
 
@@
 In order to get a accurate measure of deterioration over time, we time stamped each transcript in relation to number of days from the first transcript to the date of the transcript being analysed as the press conferences were not evenly spaced out during the presidencies.
+We then applied a pearsons product moment correlation to examine whether these language features increased or decreased significantly over time.
+Finally, for the features we found significant for RR we compared these results with GHWB and DJT using ???
+to look at whether the difference was significant given the sample size.
 \par
@@
 	\begin{center}
-	\begin{tabular}{ | p{3cm} | p{2.2cm} | p{2.2cm} | p{2.2cm} |}
+	\begin{tabular}{ | p{3cm} | p{3cm} | p{3cm} | p{3cm} |}
 		\hline
 		& RR v GHWB & RR v DJT & GHWB v DJT \\ \hline
-		Total Words & 6.6751 (574.8808, 1057.5324) & 5.352 (974.6629, 2175.8489) & 2.4774 (137.5811, 1380.5176) \\ \hline
-		Unique Stems & 10.878 (161.1580, 232.6168) & 10.11 (275.4688, 414.0975) & 4.148 (75.86722, 219.92432) \\ \hline
-		Mean Length of Utterance & 16.175 (3.915869, 5.009114) & 25.084 (8.453404, 9.922328)  & 13.484 (4.021601, 5.429148) \\ \hline	
-		Non Specific Nouns & 9.2052 (5.881447, 9.13725) & 3.8627 (3.275567, 10.474625) & -0.37956 (-4.037218, 2.770237) \\ \hline
-		LI Verbs & 4.0434 (10.84689, 31.53291) & 2.804 (10.91107, 67.41231) & 1.2836 (-11.06486, 49.00844)\\ \hline
+		Total Words & 6.6751 (479.77, 1134.64) & 5.352 (766.95, 2383.60) & 2.4774 (-74.34, 1592.44) \\ \hline
+		Unique Stems & 10.878 (149.73, 244.04) & 10.11 (251.17, 437.85) & 4.148 (51.53, 244.26) \\ \hline
+		Mean Length of Utterance & 16.175 (3.73, 5.19) & 25.084 (8.21, 10.17)  & 13.484 (3.78, 5.66) \\ \hline	
+		Non Specific Nouns & 9.2052 (5.35, 9.66) & 3.8627 (2.06, 11.69) & -0.37956 (-5.21, 3.95) \\ \hline
+		LI Verbs & 4.0434 (7.54, 34.84) & 2.804 (0.78, 79.55) & 1.2836 (-21.35, 59.30)\\ \hline
 	\end{tabular}
-	\caption{\label{tab:table-name}RR T-tests vs GWB and DJT, T-statistic and 95\%confidence intervals}
+	\caption{\label{tab:table-name}RR T-tests vs GWB and DJT, T-statistic and 99\% confidence intervals rounded to 2.d.p}
 	\end{center} 
@@
 We also calculated these Pearsons correlations with transcript index number as a time reference and the dependant variables as a replication of the Berisha and LIss study \cite{Berisha2015} and these results are described in the supplementary material.
-
+\par 
 For our controls, we found them to be stable for the most part with the main highlights being a decrease in Adverb usage for DJT (R=-0.36, p=0.049) and a steady but not severe decline in a number of variables for GWHB, namely total word count, unique words, low imageability words and verb usage.
@@
 \par
-Given the number of hypotheses being tested, we felt it was necessary to control for false discovery rate and therefore Benjamii-Yekutieli procedure was applied to the results of the analysis.
+Given the number of features being compared, we felt it was necessary to control for false discovery rate and therefore Benjamii-Yekutieli procedure was applied to the results of the analysis.
 By applying this, we produced a final list of important features for each president.
 For RR, this was 22 features and for GHWB this was 14 features.
-The features were generated from multiple sources and therefore there was some significant overlap between the features but for completeness we have included all features in Table 6.
+The features were generated from multiple sources and therefore there was some significant overlap between the features.
+Where this was the case, for example total nouns and nouns normalised,  we selected selected the feature that best controlled for document length (see Table 6).
+We also removed any features which calculated the use of punctuation as this was added later by the transcriptionist.
+This reduced the number of features that were significant for RR to 14 and for GHWB to 3.
+The majority of these features was significantly different in terms of movement over time when comparing RR to GHWB.
+The three features that were significant for GHWB in terms of correlation coefficient were not significant when comparing them with RR.
 
@@
 		& RR R-Squared  & GWB R-Squared \\ \hline
-		ppron & \textbf{0.700***} & 0.007 \\ \hline
-		social & \textbf{0.698***} & 0.204 \\ \hline
-		NounsNormalised & \textbf{-0.689***} & -0.023 \\ \hline
-		function & \textbf{0.670***} & -0.169 \\ \hline
-		conj &\textbf{0.644***} & -0.452 \\ \hline
-		PronounsNormalised & \textbf{0.631***} & 0.131 \\ \hline
-		Analytic & \textbf{-0.626***} & -0.013 \\ \hline
-		NN & \textbf{-0.601***} & -0.197 \\ \hline
-		male & \textbf{0.585***} & 0.024 \\ \hline
-		UniqueWords & \textbf{-0.578***} & -0.257 \\ \hline
-		WDT & \textbf{-0.577***} & -0.173 \\ \hline
+		ppron & \textbf{0.700*} & 0.007 \\ \hline
+		social & \textbf{0.698*} & 0.204 \\ \hline
+		NounsNormalised & \textbf{-0.689*} & -0.023 \\ \hline
+		function & \textbf{0.670*} & -0.169 \\ \hline
+		conj &\textbf{0.644*} & -0.452 \\ \hline
+		PronounsNormalised & \textbf{0.631*} & 0.131 \\ \hline
+		Analytic & \textbf{-0.626*} & -0.013 \\ \hline
+		NN & \textbf{-0.601*} & -0.197 \\ \hline
+		male & \textbf{0.585*} & 0.024 \\ \hline
+		UniqueWords & \textbf{-0.578*} & -0.257 \\ \hline
+		WDT & \textbf{-0.577*} & -0.173 \\ \hline
 		shehe & \textbf{0.529*} & -0.005 \\ \hline
@@
 		JJ & \textbf{-0.518*} & -0.206 \\ \hline
-		Fillers & -0.076 & \textbf{-0.359**} \\ \hline	
-		EX & -0.193* & \textbf{-0.337**} \\ \hline	
-		achieve & -0.248 & \textbf{0.334**} \\ \hline		
+		Fillers & -0.076 & \textbf{-0.359*} \\ \hline	
+		EX & -0.193* & \textbf{-0.337*} \\ \hline	
+		achieve & -0.248 & \textbf{0.334*} \\ \hline		
 		\multicolumn{3}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
-    	\multicolumn{3}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
-    	\multicolumn{3}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
 	\end{tabular}
-	\caption{\label{tab:table-name}Pearson Correlations for Features}
+	\caption{\label{tab:table-name}Pearson Correlations for Features for RR and GHWB}
 	\end{center} 
@@
 \subsubsection{Comparison of Ronald Reagan and Donald J. Trump}
+We then used DJT as another control to see our results from our comparison with GHWB held true.
+We found that the same 53 features were significantly different between RR and DJT and all of the features which we deemed significant in terms of increase or decrease were included.
+This confirms our results from our original comparison.
+
 \begin{table}[H]
@@
 		& RR R-Squared & DJT R-Squared \\ \hline
-		ppron & \textbf{0.700***} & 0.007 \\ \hline
-		social & \textbf{0.698***} & 0.204 \\ \hline
-		NounsNormalised & \textbf{-0.689***} & -0.023 \\ \hline
-		function & \textbf{0.670***} & -0.169 \\ \hline
-		conj &\textbf{0.644***} & -0.452 \\ \hline
-		PronounsNormalised & \textbf{0.631***} & 0.131 \\ \hline
-		Analytic & \textbf{-0.626***} & -0.013 \\ \hline
-		NN & \textbf{-0.601***} & -0.197 \\ \hline
-		male & \textbf{0.585***} & 0.024 \\ \hline
-		UniqueWords & \textbf{-0.578***} & -0.257 \\ \hline
-		WDT & \textbf{-0.577***} & -0.173 \\ \hline
-		shehe & \textbf{0.529*} & -0.005 \\ \hline
-		VBZ & \textbf{-0.521*} & -0.157 \\ \hline
-		JJ & \textbf{-0.518*} & -0.206 \\ \hline
-		Fillers & -0.076 & \textbf{-0.359**} \\ \hline	
-		EX & -0.193* & \textbf{-0.337**} \\ \hline	
-		achieve & -0.248 & \textbf{0.334**} \\ \hline		
+		ppron & \textbf{0.700*} & -0.196 \\ \hline
+		social & \textbf{0.698*} & 0.032 \\ \hline
+		NounsNormalised & \textbf{-0.689*} & 0.194 \\ \hline
+		function & \textbf{0.670*} & -0.244 \\ \hline
+		conj &\textbf{0.644*} & -0.187 \\ \hline
+		PronounsNormalised & \textbf{0.631*} & 0.002 \\ \hline
+		Analytic & \textbf{-0.626*} & 0.367 \\ \hline
+		NN & \textbf{-0.601*} & 0.116 \\ \hline
+		male & \textbf{0.585*} & 0.043 \\ \hline
+		UniqueWords & \textbf{-0.578*} & 0.204 \\ \hline
+		WDT & \textbf{-0.577*} & -0.106 \\ \hline
+		shehe & \textbf{0.529*} & 0.065 \\ \hline
+		VBZ & \textbf{-0.521*} & 0.136 \\ \hline
+		JJ & \textbf{-0.518*} & 0.063 \\ \hline	
 		\multicolumn{3}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
-    	\multicolumn{3}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
-    	\multicolumn{3}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
 	\end{tabular}
-	\caption{\label{tab:table-name}Pearson Correlations for Features}
+	\caption{\label{tab:table-name}Pearson Correlations for Features for RR and DJT}
 	\end{center} 
@@
 \subsection{Comparing linear models and non-linear models}
-In terms of comparing the general additive model and linear models.
-We calculated the predicted residual error sum of squares (PRESS) statistic for each feature is a form of cross-validation used in regression analysis to provide a summary measure of the fit of a model to a sample of observations that were not themselves used to estimate the model.
-There were no significant differences (paired t-test: t=1.8875, df  = 21, p-value = 0.07299).
+In terms of comparing the general additive model and linear models. We calculated the predicted residual error sum of squares (PRESS) statistic for each feature. The PRESS statistic
+is a form of cross-validation used in regression analysis to provide a summary measure of the fit of a model to a sample of observations that were not themselves used to estimate the model.
+There were no significant differences when comparing the models (paired t-test, t = 1.8875, df = 21, p = 0.07299).
 
 \begin{table}[H]
-	\begin{center}
-	\begin{tabular}{ | l | c | c |}
+\centering
+\begin{tabular}{lrr}
   \hline
-		Feature & GAM Press Statistic & LM Press Statistic \\ \hline
-		ppron & 18.591104613 & 20.001055542 \\ \hline
-		social & 21.681732112 & 24.906646115  \\ \hline
-		NounsNormalised & 0.002818941 & 0.002797798\\ \hline
-		function & 46.546071705 & 46.031993602 \\ \hline
-		conj & 8.812330188 & 8.354974826 \\ \hline
-		PronounsNormalised & 0.002284437 & 0.002356129\\ \hline
-		Analytic & 2272.200519272 & 2239.224863610 \\ \hline
-		Pronoun & 44.557282749 & 46.894576916 \\ \hline		
-		NN & 101685.870974277 & 96485.575598505 \\ \hline
-		male & 5.191815443 & 5.278559152 \\ \hline
-		UniqueWords & 255114.636660673 & 238042.427437454 \\ \hline
-		WDT & 1987.149134694 & 1958.781236655 \\ \hline
-		nouns & 276723.222762842 & 262192.120091097 \\ \hline
-		nouns/100 & 27.670762078 & 26.219212009 \\ \hline
-		UniqueStems & 170450.645283172 & 166302.725530118 \\ \hline
-		shehe & 5.188136314 & 5.265385185 \\ \hline
-		VBZ & 13889.871831536 & 13721.175422577 \\ \hline
-		JJ & 24950.440222526 & 24692.043386181 \\ \hline
-		article & 13.500676675 & 13.468083961 \\ \hline	
-		Adjectives & 30210.006270270 & 29865.240707991 \\ \hline	
-		Adjectives.100 & 3.020898615 & 2.986524071 \\ \hline
-		Dic & 40.099289973 & 38.501060715 \\ \hline	
+feature & gamPRESS & lmPRESS \\ 
+  \hline
+ppron & 18.59 & 20.00 \\ 
+  social & 21.68 & 24.91 \\ 
+  NounsNormalised & 0.00 & 0.00 \\ 
+  function. & 46.55 & 46.03 \\ 
+  conj & 8.81 & 8.35 \\ 
+  PronounsNormalised & 0.00 & 0.00 \\ 
+  Analytic & 2272.20 & 2239.22 \\ 
+  NN & 101685.87 & 96485.58 \\ 
+  male & 5.19 & 5.28 \\ 
+  UniqueWords & 255114.64 & 238042.43 \\ 
+  WDT & 1987.15 & 1958.78 \\ 
+  Nouns/100 & 27.67 & 26.22 \\ 
+  UniqueStems & 170450.65 & 166302.73 \\ 
+  shehe & 5.19 & 5.27 \\ 
+  VBZ & 13889.87 & 13721.18 \\ 
+  JJ & 24950.44 & 24692.04 \\ 
+  article & 13.50 & 13.47 \\ 
+  Adjectives.100 & 3.02 & 2.99 \\ 
+  Dic & 40.10 & 38.50 \\ 
+   \hline
 \end{tabular}
-	\caption{\label{tab:table-name}Comparison of GAM and Linear Model using the PRESS statistic}
-	\end{center} 
 \end{table}
-
 \section{Discussion}\label{discussion}
@@
 \par 
+Another we did was to look at the difference in decline between the presidents.
+We compared both RR v GHWB and RR v DJT.
+We found that for those features that were significant in terms of language decline for RR, they were significantly different for both our controls.
+This does appear to mean that for RR, his language changed in a considerably different way as he got older and we could attribute this decline to a pre-clinical stage of alzheimer's disease.
+\par
 Overall, we are able to show differences in the a number of language features between RR, GHWB and DJT over time and the psychological literature confirms our findings with reference to these changes.
@@
 We argue that these more nuanced changes over time can only be modelled by a model such as a GAM.
+In Le's et al's paper, he looks at the written work of Iris Murdoch who would go on to receive a diganosis of Alzheimer's disease.
+He describes something called 'Murdoch's trough' which is a period of decline in her late 40's and early 50's, however there was a significant period following this of an improvement in her linguistic ability before a subsequent severe decline in her last novel.
+Figure 2 illustrates this, as well as what a non-linear model would look like vs a linear model.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=6cm, height=5cm]{plots/IMStems.png}
+	\caption{Measuring unique word stems in the novels of Iris Murdoch over time}
+\end{figure}
+We can clearly see that a linear model is insufficient for modelling this particular case and whilst Iris Murdoch may be an atypical case, we can see the value in using non-linear models that may be able to map these unusual cases whilst retaining the ability to illustrate fluctuations in language use.
 \par
@@
 Equally, it would be helpful to have controls to ascertain what would be usual to expect in the decline of language in a healthy older adult.
+We were able to identify a number of linguistic features that merit further exploration in those with MCI or AD.
 \par 

### 2020-04-06 Source/Paper 1 - Presidents Corpus Paper/PresidentsCorpus.tex
updated Pres

@@
 Cummings et al proposed that a possible reason for the lack of success is that the drug treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already occurred \cite{Cummings2014}.
-Research focus has now started to shift to the earlier stages of AD (i.e.
-symptomatic pre-dementia phase of AD) which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.
+\par
+
+Research focus has now started to shift to the earlier stages of AD (i.e. symptomatic pre-dementia phase of AD) which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.
 \par
@@
 	\centering
-	\includegraphics[width=6cm, height=5cm]{plots/comparison1.png}
+	\includegraphics[width=12cm, height=9cm]{plots/comparison1.png}
 	\caption{Comparing linear model with a non-linear model}
@@
 \subsection{Comparing linear models and non-linear models}
-In terms of comparing the general additive model and linear models. We calculated the predicted residual error sum of squares (PRESS) statistic for each feature. The PRESS statistic
-is a form of cross-validation used in regression analysis to provide a summary measure of the fit of a model to a sample of observations that were not themselves used to estimate the model.
+In terms of comparing the general additive model and linear models.
+We calculated the predicted residual error sum of squares (PRESS) statistic for each feature.
+The PRESS statistic is a form of leave one out cross-validation (LOOCV) used in regression analysis to provide a summary measure of the fit of a model to a sample of observations that were not themselves used to estimate the model.
 There were no significant differences when comparing the models (paired t-test, t = 1.8875, df = 21, p = 0.07299).
@@
 \centering
-\begin{tabular}{lrr}
+\begin{tabular}{| l | r | r | r |}
   \hline
-feature & gamPRESS & lmPRESS \\ 
+feature & gamPRESS & lmPRESS & Best Model\\ 
   \hline
-ppron & 18.59 & 20.00 \\ 
-  social & 21.68 & 24.91 \\ 
-  NounsNormalised & 0.00 & 0.00 \\ 
-  function. & 46.55 & 46.03 \\ 
-  conj & 8.81 & 8.35 \\ 
-  PronounsNormalised & 0.00 & 0.00 \\ 
-  Analytic & 2272.20 & 2239.22 \\ 
-  NN & 101685.87 & 96485.58 \\ 
-  male & 5.19 & 5.28 \\ 
-  UniqueWords & 255114.64 & 238042.43 \\ 
-  WDT & 1987.15 & 1958.78 \\ 
-  Nouns/100 & 27.67 & 26.22 \\ 
-  UniqueStems & 170450.65 & 166302.73 \\ 
-  shehe & 5.19 & 5.27 \\ 
-  VBZ & 13889.87 & 13721.18 \\ 
-  JJ & 24950.44 & 24692.04 \\ 
-  article & 13.50 & 13.47 \\ 
-  Adjectives.100 & 3.02 & 2.99 \\ 
-  Dic & 40.10 & 38.50 \\ 
+ppron & 18.59 & 20.00 & GAM\\ 
+  social & 21.68 & 24.91 & GAM\\ 
+  NounsNormalised & 0.00281 & 0.00279 & LM\\ 
+  function. & 46.55 & 46.03 & LM\\ 
+  conj & 8.81 & 8.35 & LM \\ 
+  PronounsNormalised & 0.00228 & 0.00235 & GAM \\ 
+  Analytic & 2272.20 & 2239.22 & LM\\ 
+  NN & 101685.87 & 96485.58 & LM\\ 
+  male & 5.19 & 5.28 & GAM\\ 
+  UniqueWords & 255114.64 & 238042.43 & LM\\ 
+  WDT & 1987.15 & 1958.78 & LM \\ 
+  Nouns/100 & 27.67 & 26.22 & LM\\ 
+  UniqueStems & 170450.65 & 166302.73 & LM \\ 
+  shehe & 5.19 & 5.27 & GAM\\ 
+  VBZ & 13889.87 & 13721.18 & LM\\ 
+  JJ & 24950.44 & 24692.04 & LM\\ 
+  article & 13.50 & 13.47 & LM\\ 
+  Adjectives.100 & 3.02 & 2.99 & LM\\ 
+  Dic & 40.10 & 38.50 & LM\\ 
    \hline
@@
 We found that for those features that were significant in terms of language decline for RR, they were significantly different for both our controls.
-This does appear to mean that for RR, his language changed in a considerably different way as he got older and we could attribute this decline to a pre-clinical stage of alzheimer's disease.
+This does appear to mean that for RR, his language changed in a considerably different way as he got older and we could attribute this decline to a pre-clinical stage of AD.
 \par
@@
 \end{figure}
-We can clearly see that a linear model is insufficient for modelling this particular case and whilst Iris Murdoch may be an atypical case, we can see the value in using non-linear models that may be able to map these unusual cases whilst retaining the ability to illustrate fluctuations in language use.
+We can clearly see that a linear model is insufficient for modeling this particular case and whilst Iris Murdoch may be an atypical case, we can see the value in using non-linear models that may be able to map these unusual cases whilst retaining the ability to illustrate fluctuations in language use.
 \par
-In her paper on developing diagnositic criteria for Mild Cognitive Impairment, Albert et all \cite{Albert2011} states 'it is important to obtain longitudinal assessments of cognition, whenever possible' and 'obtaining objective evidence of progressive declines in cognition over time is important for establishing the accuracy of the diagnosis, as well as for
+In her paper on developing diagnostic criteria for Mild Cognitive Impairment, Albert et al \cite{Albert2011} states 'it is important to obtain longitudinal assessments of cognition, whenever possible' and 'obtaining objective evidence of progressive declines in cognition over time is important for establishing the accuracy of the diagnosis, as well as for
 assessing any potential treatment response.' As we have shown, it is not the case that language declines in a linear way and whilst we have a number of data points that we can use to model RR's decline, in a clinical setting it would be impractical to have upwards of 40 data collection points.

### 2020-04-06 Source/Thesis/Thesis.tex
updated Thesis
update Thesis and Delphi Questionnaire
updated Pres

@@
 %%% PREAMBLE
-\documentclass[10pt, letterpaper, twoside, openany]{book}
+\documentclass[10pt, letterpaper, twoside, openany]{thesis}
 
@@
 \usepackage[utf8]{inputenc}
-\usepackage{fdsymbol} 
 \usepackage{graphicx}
@@
 \chapter{Introduction}
-\include{Chapter1}
+\section{Background}
+Alzheimer's Disease (AD) and other forms of dementia affect a significant proportion of the geriatric population in the world today and is currently the sixth leading cause of death in the US and was named the leading killer of women in the UK.
+According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people.
+This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population \cite{AlzheimersSociety2014}.
+From a financial perspective, they estimate an annual spend of £4.3 billion of which approximately £85 million is spent solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually.
+Globally, this picture is a lot bleaker.
+Another report by Alzheimer's Disease International suggests that in 2015 there were 46 million people with a diagnosis of dementia globally and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
+\par 
+AD is a neurodegenerative disease in which a definitive diagnosis can only be produced at post-mortem.
+However, there are a number of psychological and physiological indicators that can indicate that dementia is present.
+From a physiological perspective, researchers have identified two proteins called beta-amyloid and tau.
+In a typical case, tau accumulates and eventually forms tangles inside neurons and beta-amyloid clumps into plaques which slowly builds up between neurons.
+At a certain point, the levels of beta-amyloid rise and trigger a more rapid spread of tau throughout the brain.
+Eventually, due to this and other changes, neurons lose their ability to communicate and the brain starts to shrink.
+This leads to some of the more psychological symptoms, those who have dementia demonstrate cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}.
+In addition, these symptoms are often accompanied by emotional problems such as depression and behavioural difficulties.
+As more neurons die throughout the brain, a person with Alzheimer's gradually loses the ability to think, remember, make decisions and function independently.
+\par
+Despite this growing problem and an increasing understanding about how AD affects the brain there are no medications that improve the prognosis of those with AD.
+All the medications that are currently on the market are designed to manage symptoms.
+Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{Cummings2014}.
+Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place \cite{Cummings2014}.
+\par
+We can characterise a person's progression through AD in three stages.
+Whilst these stages are called different things in the literature, we will stick to the following naming convention throughout this thesis
+\begin{enumerate}
+	\item[a)] Stage 1: Preclinical AD
+	\item[b)] Stage 2: Mild Cognitive Impairment due to AD
+	\item[c)] Stage 3: Dementia due to AD
+\end{enumerate}
+Research has started to be more focused on stage 2 however a problem with focusing here is there are no clear defined boundaries between these three stages.
+\par
+One of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of cognitive impairment or dementia.
+This challenge is often complicated further due to the large variation in the cognitive abilities and educational background of individuals.
+Albert and his team have worked to define clinical criteria which professionals can use to diagnose MCI due to AD and differentiate this from age-associated memory impairment and age-associated cognitive decline.
+One of the most important observations from this piece of work is that a diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}, and this is essentially to place more importance on the trajectory of a person's cognitive abilities rather than a person's cognitive ability in general.
+The criteria for MCI is detailed below \cite{Albert2011}.
+\begin{enumerate}
+	\item Concern regarding a change in cognition - A person or an informant should express concern that there is a change in cognitive ability in comparison to a previous level of performance.
+	\item Impairment in one or more cognitive domains - There should be evidence of lower performance in one or more cognitive domains beyond what would be expected of a person given their age and education.
+	\item Preservation of independence in functional abilities - Whilst persons with MCI are expected to be able to maintain independence, it is common to experience mild problems in complex functional tasks which they may have been able to perform previously.
+This might mean that they take more time or be less efficient at completing these tasks, or it may be that may make more mistakes.
+	\item Not demented - The deterioration should be mild to the point that there is no significant loss of functioning in social or occupational contexts.
+\end{enumerate}
+In addition to meeting the above criteria, a clinician must rule out other conditions or factors that could account for the decline in cognition with the goal to increase the likelihood that the underlying cause of this decline is dementia.
+\par
+Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease.
+Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
+% Do more on the state of the art in terms of the analysis of biomarkers for MCI.
+\section{Statement of the Problem}
+Studies that explore ways in which we can diagnose MCI / AD generally follow one of two main approaches, the analysis of biomarkers such as concentrations of amyloid-β 1-42 (Aβ42) in cerebrospinal fluid (CSF) and analysis of cognitive abilities.
+The first approach yields reliable results in the detection of AD in its moderate and advanced states but does not perform well during the early stages of the disease.
+The second approach, that of analysing the cognitive abilities of patients in memory clinics, has gained more attention in recent years due to the fact that in clinical practice it has shown promise in the early detection of AD.
+In addition, the analysis of the decline of cognitive abilities is comparatively inexpensive and less invasive that the first approach which commonly requires the collection of a sample of cerebro-spinal fluid which is painful for the patient involved.
+This has a number of benefits for countries with less developed healthcare systems, or where the burden of healthcare is more extreme.
+\par
+One of the most common ways in which clinicians traditionally use the analysis of cognitive abilities to make an early diagnosis of dementia is through the use of the Mini Mental State Examination (MMSE) \cite{Folstein1975}.
+The MMSE is a brief questionnaire consisting of eleven questions which tests cognitive aspects of mental function and requires only 5-10 minutes to administer \cite{Folstein1975}.
+The MMSE is chosen due to it's effectiveness at assessing a person's cognitive mental state at a specific point in time, as well as being as sensitive to changes as a more detailed and complex assessment such as the Wechler Adult Intelligence Scale \cite{Folstein1975}.
+Whilst the MMSE is useful as a brief screening tool it has it's limitations.
+The MMSE was not specifically created to screen for dementias and therefore does not interrogate key aspects of cognitive impairment known to be affected in dementia.
+It also has limited value in assessing under-educated subjects and a meta-analysis on the effectiveness of the MMSE as a diagnostic tool for dementia showed that it's accuracy was low (sensitivity between 78.4\% and 85.1\% and specificity between 81.3\% and 87.8\%).
+As the MMSE has been shown to have low accuracy specifically in the diagnosis of dementia, it becomes necessary for professionals to employ the use of other tools or measures such as the Free Cued Selective Reminding Test (FCSRT) \cite{Grober2010} or the Montreal Cognitive Assessment (MoCA) \cite{Davis2015}.
+\par 
+These tests have the benefits of being much more accurate at diagnosing cognitive impairment and discriminating between dementia and other types of cognitive impairment at the cost time and training of psychological professionals such as clinical psychologists in administering these tests.
+However, the utility of diagnosing dementia at the point where clinical intervention is warranted is limited because at this stage both psychological and pharmacological interventions have been shown to not be effective \cite{Prince2015, Cummings2014}.
+In order to further our understanding of the progression of dementia it is important to detect the signs of dementia before they are clinically apparent.
+\par 
+Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent.
+Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time and alter cognitive function until a threshold is reached and clinical symptoms become more obvious \cite{Nestor2006}.
+The case of Iris Murdoch, who had a confirmed diagnosis of dementia, illustrates this theory well.
+Le et al \cite{Le2011} found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel when, it is theorized, the symptoms of AD manifested themselves more significantly.
+If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
+\par
+The two main ways in which diagnosis is performed is through assessment of memory and language.
+Tests of memory are classically among the most accurate ways of diagnosing dementia, however these tests suffer from the same reliance on clinicians to administer these tests in a clinical setting.
+Language however is a lot easier to collect and can be done in more naturalistic settings.
+As with memory, these tests can be done over time and would be able to chart a patients language degeneration over time.
+Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on measure decline in the use of language in those with AD.
+There are a number of difficulties to watch out for with this approach.
+There are a wide number of factors that are involved in language degeneration in the elderly, and consequently there will be an expected amount of variability between subjects.
+The administration of such tests may induce nervousness and discomfort which may impact performance, and also repeatedly administering the same language tests for differences over time may be confounded by improved performance at tasks via practice effects.
+However, there is enough promise in this approach such that it could help further our understanding of the disease, it's progression and the parts of the brain affected in the early stages.
+\par
+McKhann and his team were tasked with developing diagnostic guidelines for dementia in such as way that the criteria were flexible enough to be used by general healthcare providers without access to specialist medical equipment.the DSM 5 \cite{McKhann2011}.
+They state those suffering with mild dementia generally encounter impaired language functions (speaking, reading, writing)––symptoms include: difficulty thinking of common words while speaking, hesitations ;speech, spelling, and writing errors.
+More specifically, they may substitute general terms for more specific terms and may avoid the use of specific names of acquaintances.
+There may be grammatical errors involving subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc.
+Those who have progressed from Mild to Major depression also have difficulties with expressive or receptive language.
+They will often use general-use phrases such as "that thing" and "you know what I mean" and prefers general pronouns rather than names.
+With severe impairment, sufferers may not even recall names of closer friends and family.
+Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur.
+Echolalia (meaningless repetition of another person's spoken words) and automatic speech typically precede mutism.
+With the wide range of deficits someone with AD can suffer, it makes sense to try to categorise these deficits in some way.
+\par
+One of the most famous pieces of research on the topic of language decline in dementia was by Berisha and Liss who examined speeches and public interviews of former US president Ronald Reagan \cite{Berisha2015}.
+They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease.
+It turned out later to be the case that he had Alzheimer's Disease.
+Another classical study by Snowdon et al looked at whether linguistic ability in early life was associated with cognitive function and AD in later life \cite{Snowdon1996} .
+They found that idea density (defined as the number of expressed propositions divided by the number of words) was a key predictor in predicting whether nuns would go on to develop AD in later life.
+They found that those who would go on to develop AD all had low idea density in early life and they found no AD present in those with high idea density in early life.
+As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are extremely variable and can differ from patient to patient as the disease progresses.
+The consensus among researchers that this language degeneration is typically accelerated by the presence of dementia \cite{Berisha2015} and that a potential indicator of dementia is the rate of change in which the decline occurs relative to a fixed point in time rather than a comparison across a cohort of individuals.
+\par
+Emery \cite{Emery2000} completed a literature review looking at all the potential language deficits that could exist in those with AD and / or MCI.
+She divided these deficits into four levels of language: Phonology, Morphology, Syntax and Semantics.
+She proposed that language and the processes involved in language are hierarchical in nature and that language moves from simple units of construction (Phonology and Morphology), and build layers of complexity and sophistication (Syntax and Semantics).
+She found that people with AD generally had intact Phonology and Morphology but more impaired Syntax and Semantics.
+She asserted that the language forms we learn last are the first to deteriorate as we generally learn language in small simple units initially and build syntax and complexity as we are more comfortable with language.
+However it is important to note that different variants of dementia show different deficits in terms of language productions.
+Regardless, dividing language in this way is useful as it allows us to detect deteriorations in different parts of language usage and therefore may provide a way of discriminating between different forms of dementia.
+\par
+It is clear from both the clinical diagnostic criteria and supporting research that language is impacted in those with AD.
+However, whilst there is a move towards research aimed at looking more specifically at MCI we currently lack the measures that are sensitive enough to detect MCI.
+Given that we know language is affected before a clinical diagnosis of dementia is usually made \cite{Berisha2015, Snowdon1996, Le2011}, it makes sense to explore whether language on it's own can provide markers that may indicate a cognitive impairment that could progress to dementia.
+The field of machine learning and natural language processing has been suggested as a way to improve the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline \cite{Boschi2017}.
+I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
+\par
+There is a large body of research that looks at the decline in language in those with MCI and AD \cite{Taler2008, Boschi2017}.
+However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD \cite{Taler2008, Boschi2017}.
+Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
+Another area of focus for research of this nature is the process of collecting appropriate language samples.
+Whilst collecting samples of language is comparatively unintrusive, researchers recognise that these samples require a rich sample of language that potentially cannot be generated by tasks such as the picture description task.
+Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD.
+As is the case with other quantitative measures of cognitive ability, contrasting individuals with MCI and Early AD is challenging due to the variation in an individual's baseline speech capacity.
+It is therefore prudent to measure cognitive decline over time as suggested by Albert et al \cite{Albert2011}.
+There is some evidence to support this approach, Berisha et al \cite{Berisha2015} demonstrated through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control.
+\par
+\section{Research Questions}
+The overall objective of the research described in this thesis is to investigate the deterioration of language in people diagnosed with MCI and the early stages of AD.
+In order to do this I aim to answer the following research questions
+\begin{enumerate}
+	\item What research has been conducted in this field so far?
+	\item How can we use technology to best enable the collection of language samples?
+	\item Can we develop a complete data processing pipeline that processes data efficiently
+	\item What can machine learning tell us about how language affects us?
+\end{enumerate} 
+\par 
+This thesis' original contributions are an increased understanding of how language deteriorates over time in those with MCI and early AD, the exploration via a delphi method of ideas for the collection of language features via smartphone or other internet of things devices and the development and execution of a pilot study to test the feasibility of collecting high quality language samples via in-home technology.
+I also contribute to the research into which language features are important in the diagnosis of MCI and early AD over time.
+Finally, I explore the potential of one class classification as a new approach to classifying those with MCI and Early AD vs controls.
+The potential impact of this research in this area is immense.
+\par 
+A number of research themes combine both analysis of language and analysis of the acoustic component of speech.
+Whilst we acknowledge the existence of this large body of research and it's potential contribution to this problem, we have chosen to focus and attempt to understand the language analysis portion of this problem.
+
+\section{Structure of thesis}
+There is a lot research in this domain from a psychological perspective.
+In chapter 2, I look at the background and related work in the domain of language deterioration in those with MCI and AD from the psychological perspective.
+This includes looking at the current ways in which psychologists collect and analyse language samples and exploration of the features of language that psychologists feel are important in this area.
+In addition, I take a systematic look at the work of the experts in the fields of machine learning and natural language processing in the growing area of research for NLP and Machine Learning and it's application to this problem.
+\par 
+The start of my research aims to answer questions around how we can use technology to aid us in collecting language samples that may facilitate the process of diagnosis.
+Currently, language samples are collected in memory clinics and physician's offices and is quite a time consuming process.
+More recently, work has been done to automate this process via a web based app called Talk2Me \cite{Komeili2019}, which seeks to use different psychological tests to assess cognitive decline through analysis of language.
+There are some potential flaws in this method of data collection.
+In other areas, the use of smartphone technology has allowed the collection of language for the diagnosis of disorders such as Parkinson's disease and so it appears plausible to be able to use this technology and potentially other technologies such as Amazon Echo to facilitate the collection of language samples.
+The work described in chapter 3 therefore describes an exploration of experts opinions' on the following questions.
+\begin{enumerate}
+	\item How should we collect these language samples - Smartphones? IoT Devices?
+	\item What tasks should we use to elicit these language samples
+	\item At what frequency should we collect these language samples at
+\end{enumerate}
+\par 
+An adapted Delphi survey methodology is used to develop consensus on the answers to these questions and the results of these are documented in this chapter.
+The resulting pilot study is described in Chapter 6.
+\par 
+Once we have a way to collect language samples, the question to consider is how we may best analyse this data.
+There are a number of factors that one must consider when answering the question.
+We look at the effectiveness of automatic speech recognition at translating speech into language, as well as the minimum quality of data that needs to be passed to any pipepline we develop.
+The Talk2Me app collects and then generates approximately 2000 lexico-syntactic, acoustic and semantic features {\cite Komeili2019} as part of their data collection framework.
+Chapter 4 discusses the challenges connected to constructing an effective data pipeline and contains two experiments.
+\par 
+Whilst Chapter 4 is more theoretical in it's content, Chapter 5 documents the first experiments that use this pipeline on three existing datasets.
+These datasets are, the Presidents Press Conferences, The Novels of three authors and the DementiaBank dataset.
+\par
+In chapter 6 Pilot study of the methodology developed
+\par 
+Finally, in chapter 7.
+I discuss conduct a general discussion of the results of my research.
+I also think about the strengths and weaknesses of my work and suggest ways in which the research have been improved.
+Lastly, I look at a number of areas for future work which can build on this research.
 
 \chapter{Background and Related Work, Systematic Review of NLP and Machine Learning Research}
-\include{Chapter2}
+\section{Existing Neuropsychological Measures of Cognitive Impairment, and Repeatable Battery of Neuropsychological Tests}
+In terms of standardised cognitive tests there are two main aims.
+Does the test distinguish accurately between normal aging, MCI and AD (diagnostic utility) and does the test distinguish between those individuals with MCI who will then go on to develop AD and those individuals with MCI who don't then go on to develop AD (prognostic utility).
+This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their performance in terms of both diagnostic and prognostic utility.
+\par 
+\subsection{Repeatable Battery for the Assessment of Neuropsychological Status}
+The Repeatable Battery for the Assessment of Neuropsychological Status (RBANS) was originally developed as an assessment tool for dementia, specifically looking at detecting a characterizing very mild dementia \cite{Randolph1998}.
+The authors felt that there was a shortfall of appropriate measures that were sensitive enough to milder impairments as well as a number of other shortcomings of existing tests.
+They met a number of design goals for this new battery of tests that addressed these shortcomings.
+The RBANS consists of a number of sub-tests across five distinct domains.
+
+\begin{enumerate}
+	\item Immediate Memory 
+	\begin{enumerate}
+		\item{List Learning}
+		\item{Story Memory}
+	\end{enumerate}
+	\item Visuospatial / Constructional
+	\begin{enumerate}
+		\item{Figure Copy}
+		\item{Line Orientation}
+	\end{enumerate}
+	\item Language
+	\begin{enumerate}
+		\item{Picture Naming}
+		\item{Semantic Fluency}
+	\end{enumerate}
+	\item Attention
+	\begin{enumerate}
+		\item{Digit Span}
+		\item{Digit Coding}
+	\end{enumerate}
+	\item Delayed Memory
+	\begin{enumerate}
+		\item{List Recall}
+		\item{List Recognitions}
+		\item{Story Memory}
+		\item{Figure Recall}
+	\end{enumerate}
+\end{enumerate}
+
+However, whilst the authors claim that the RBANS is adequately sensitive in person's with MCI \cite{Randolph1998} and there is some research to support this assertion \cite{Karantzoulis2013}, other research points out that the RBANS has poor sensitivity in detecting MCI \cite{Duff2010}.
+Another drawback of this battery is the lack of executive function measures and object naming tasks.
+Research has shown that the RBANS has good test-retest reliability and convergent validity.
+However, Duff et al warned that caution should be exercised when using the RBANS in a MCI population as it has lower sensitivity in this population \cite{Duff2010}.
+
+\subsection{Digit Span Test}
+The Digit Span Test (DST) is predominantly used to measure a person's working memory capacity, specifically the capacity used to store and recall numbers.
+A participant is presented with a series of numbers of fixed length and is asked to recall those numbers in normal or reverse order.
+The series of numbers gets progressively longer until such time as a participant fails three or more times out of eight presentations.
+\par
+Research has consistently shown that MCI patients have a significantly lower digit span score in both normal and reverse order versions of this test and a study by Muangpaisan has shown that the reverse order version of this test can, to some degree, predict a diagnosis of MCI \cite{Muangpaisan2008}.
+Muangpaisan also found that Age, Gender and Education have an impact on the performance of the tests.
+Emrani et al found that the DST revealed no differences between specifically amnestic MCI and controls, but could differentiate between Mixed MCI and the other groups with mixed MCI recalling fewer correct responses than other groups.
+Notably, there was an attenuated recency effect in those with mixed MCI.
+\cite{Emrani2018}.
+Kessels identifies that that there are working memory deficits in MCI patients and these worsen with AD patients.
+\cite{Kessels2011} He also identified that both MCI and AD have impaired performance on all three conditions of the digit span test.
+No differences were found between forward and backward conditions in any of the groups.
+However, available tests may not detect subtle impairments \cite{Kessels2015}.
+\subsection{Rey Auditory-Verbal learning test}
+Rey's Auditory Verbal Learning Test (RAVLT+) looks a wide range of neuropsychological processes including short-term auditory-verbal memory, learning and retention of information.
+Participants are given a list of 15 unrelated words, repeated over five different trials and are asked to repeat.
+Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then repeat it again after 30 minutes \cite{Schmidt1996}.
+\par
+Several studies have shown that an impairment in RAVLT score reflect well the underlying pathology caused by AD.
+Thus making the RAVLT an effective early marker to detect AD in persons with memory complaints.
+Moradi investigated to what extent the RAVLT scores are predictable based on MRI data using machine learning approaches, as well as to find out what the most important brain regions are for the estimation of RAVLT scores \cite{Moradi2017}.
+They found a highly significant cross validated correlation between the estimated and observed RAVLT immediate and RAVLT Percent Forgetting.
+Further, they found that the conversion of MCI subjects to AD in 3-years could be predicted based on either observed or estimated RAVLT scores with an accuracy comparible to MRI-based biomarkers \cite{Moradi2017}.
+Another study by Schoenberg found that RAVLT to best distinguish patients suspected of Alzheimer's disease from the psychiatric comparison group \cite{Schoenberg2006}.
+\subsection{Digit Symbol Substitution Test}
+The Digit Symbol Substitution Test (DSST) involves a key consisting of the numbers 1-9 and a corresponding unique symbol.
+Below this key is a series of numbers from 1-9 in a randomized order and repeated multiple times.
+The participant is asked to fill in the corresponding symbol for each number.
+The task requires that the participant move between the key and the randomized sequence such that  they may retrieve the correct answer from the key, hold this in short-term memory and transcribe the key in the appropriate place.
+\par 
+Among those with no disorder in cognition, mobility and mood, being in the lowest DSST quartile compared to the highest was associated with nearly twice the odds of developing one or more clinical or subclinical disorders.
+Associations were stronger for incident clinical disorders in cognition.
+Slower psychomotor speed may serve as a biomarker of risk of clinical disorders, mobility and mood.
+While in part attributable to vascular brain disease, other potentially modifiable contributors may be present \cite{Rosano2016}.
+Further studying the causes of psychomotor slowing with ageing might provide novel insights into age-related brain disorders.
+Pascoe compared patients with PD with Normal Cognition (PD-N) with those with PD and MCI (PD-MCI) and healthy participants.
+PD-MCI participants achieved significantly lower scores than other groups in the DSST task \cite{Pascoe2018}.
+\subsection{Verbal Fluency}
+There are a number of tests that characterise this category of verbal fluency, but generally fall into two categories.
+Letter (Phonemic) fluency involves the generation of as many words as possible which begin with a specified letter.
+Category fluency involves the generation of as many words as possible that fall into a specfied category.
+Both these tasks impose demands on a number of different cognitive processes namely, executive function, verbal retrieval and recall, giving appropriate answers while monitoring previous answers and inhibit inappropriate responses.
+However, these two tasks require different strategies when attempting them.
+Letter fluency relies on search strategies based on lexical representations whereas category fluency requires a search for semantic extensions of a superordinate term, meaning that semantic associations within the lexicon must be intact in order for the task to be carried out successfully.
+\par  
+There have been numerous studies which have documented the impact that AD has on verbal fluency tests in both categories.
+A review carried out by Henry et al \cite{Henry2004}, found that performance in both letter fluency and category fluency was impaired in those with AD vs controls, but found a larger effect for tests of semantic fluency.
+\subsection{Naming Tests}
+Word-finding difficulty is a common symptom of AD and these deficits usually occur during the early stages of the disease progression.
+As such , a test of a patients ability to find words (known as confrontation naming) is a common way to measure cognitive decline.
+One common way to do this is the Boston Naming Test (BNT) which comprises 60 items on a spectrum of very frequent to very infrequent.
+This has been reduced subsequently to two thirty item versions and four fifteen item versions, which correlate significantly with the original sixty item version and the benefit of the shorter versions of the test is that it facilitates testing of individuals with AD who may suffer from fatigue or limited attention span \cite{Williams1989}.
+\par 
+Willers et al \cite{Willers2008} studied Twenty aMCI patients, twenty AD and 21 controls matched by age, sex and education level.
+They found AD patients obtained significantly lower total scores on the BNT than aMCI patients and controls.
+aMCI patients did not obtain significant differences in total scores but showed significantly higher semantic errors compared to controls.
+Semantic processing is impaired during confrontation naming in aMCI.
+\par 
+Vadikolias investigated the impact of education on naming tests \cite{Vadikolias2012} and found that higher educational attainment in aMCI subjects were correlated with better performance in verbal and non-verbal tasks during repeated examinations over 1-year.
+Subjects with a lower level of education performed worse than patients with a high level of education who presented a more stable clinical score.
+The explanation for this is the idea of a 'cognitive reserve' in participants with a higher education such that this provides a buffer that, while not preventing the physiological symptoms of AD, can potentially delay the clinical onset of cognitive symptoms that characterise AD.
+This theory is supported by Snowdon who found a relationship between early life linguistic ability and the density of neurofibrillary tangles in his nun study \cite{Snowdon1996}.
+Whilst these studies and tests provide evidence that there are word finding difficulties in those with MCI, on their own they do not provide sufficient ability to diagnose MCI or provide an prognosis of disease progression.
+\subsection{Rey-Osterrieth Complex Figure Task}
+The Rey-Osterrieth Complex Figure (ROCF) is a task widely as a test of visuo-spatial skill and visual memory.
+The task, which was originally designed by Rey (1944) and standardised by Osterrieth (1944) \cite{Rey1941, Osterrieth1944}, asks a participant to copy a complex geometrical figure (known as the immediate copy condition and to recall and reproduce the figure from memory without warning (known as the delayed recall condition).
+In the immediate copy condition, the complexity of the figure requires an integrative cognitive ability.
+The reproduction of such a complex structure involves processes such as planning and organizational strategies that are related to executive functions.
+In order to make this task repeatable, Taylor et al designed a comparable set of figures which have proven to be of equivalent difficulty \cite{Taylor1969, Strauss1990}.
+\par
+Salvadori found that patients with vascular MCI had a worse performance in the immediate copy of the ROCF compared to individuals with degenerative MCI, despite their significant impairment in terms of general cognitive status and visual memory \cite{Salvadori2018}.
+Evidence shows that patients with disorders that possibly involve attention and executive functions are characterized by a more disorganised approach when copying the ROCF compared to controls.
+One of the difficulties with the presentation of this task in a repeated battery will be the fact that it turns from an incidental memory task (it is a memory task but the participant is not forewarned that they will be required to memorise the picture) into an intentional memory task (given the previous exposure to the task, it would be expected that a participant would preempt the delayed recall portion of the test and spend more time attempting to memorise the details) \cite{Teng2009}.
+\subsection{Hayling Sentence Completion Test}
+Inhibitory deficits are a common in all stages of dementia.
+This is usually tested using Stroop test, however this has the drawback of lacking ecological validity.
+Therefore researchers have moved towards using the Hayling Sentence Completion Test (HCST) which uses skills such as word retrieval as well as the ability to inhibit ones responses where appropriate and is therefore a much more ecologically valid task.
+There are two parts to the HCST.
+In the first part, participants have to complete a sentence by providing a word that best fits the given sentence (this is known as the initiation condition).
+In this second part, participants have to complete sentences by inhibiting an impulse to give the word that best first the sentence as in the first part, and producing a semantically unconnected word.
+Performance in both conditions is measured by the time taken for the participant to initiate a response, and in the inhibition condition also by the correctness of the word.
+\par 
+Martyr et al compared healthy controls with patients with dementia and patients with Parkinson's disease \cite{Martyr2017}.
+They found that a high proportion of Category A errors (producing a word that fits the sentence when instructed otherwise) was a factor in performance loss for participants with dementia.
+Findings suggest that the HSCT may be sensitive to verbal suppression deficits and may provide insight into inhibitory control in participants with dementia.
+Patients with Dementia were significantly slower than controls in the initiation and inhibition conditions vs healthy controls, and slower than patients with Parkinson's disease in the initiation but not the inhibition condition.
+\subsection{Grooved Pegboard Test}
+The Grooved Pegboard Test (GPT) was originally designed to cover a variety of different psycho-motor functions including hand-eye coordination and motor speed.
+However, some studies have shown a correlation of performance in the GPT and measures of cognitive performance such as the Montreal Cognitive Assessment.
+\par 
+Bezdicek found that the GPT predicted performance on the MoCA and concluded that in addition to being a measure of motor skills, there were results that showed that the GPT could also provide information about a participants cognitive skills \cite{Bezdicek2014}.
+Whilst this was specifically using a population of patients with Parkinson's disease, given the MoCA and GPT have shown degree of correlation, there remains an opportunity for research that explores the use of the GPT with MCI and Early AD populations.
+\par 
+Darweesh et al, used the Purdue Pegboard Test (PPT) to assess manual dexterity in a healthy older adult population and followed their participants for between eight and twelve years until an indication of the onset of a neurodegenerative disease \cite{Darweesh2017}.
+In this time 227 (4\%) of their participants went on to develop a diagnosis of dementia.
+They noted that higher PPT scores were associated with lower risk of incident neurodegnerative disease and noted significant associations of PPT scores with all forms of dementia and this potentially highlights a deterioration of motor function in the pre-clinical phase of dementia.
+\subsection{Visual Search}
+The visual search task requires participants to be watching a computerised display in which a target appears either by itself or surrounded by other elements which are used as a distraction.
+These other elements can be vary in similarity to the target.
+This is inherently a measure of attention shifting and processing of various similarity and as one would expect, the visual search time is increased in the presence of distractors.
+This effect is magnified in participants who have a cognitive impairment.
+\par
+Research has shown there to be deficits in both AD and MCI populations in this task \cite{Tales2005}, and those with MCI exhibited less severe deficits compabut that these deficits were not as apparent in MCI populations, there remained a significant enough difference that could be used to differentiate healthy controls from those with MCI.
+In a comparison of visual search performance between those with aMCI and healthy controls, Tales noted significantly poorer performance in the aMCI group.
+However, she also noted a good deal of heterogeneity in the aMCI group which illustrates that whilst the aMCI group have essentially the same condition, presentations within this population can differ markedly.
+The results from this study also illustrates the use of a non-memory task as a means of diagnosing dementia as patients who went on to be diagnosed with dementia whilst this study was in progress had significantly poorer visual search performance.
+\subsection{Free Cued Selective Reminding Test}
+The Free Cued Selective Reminding Test (FCSRT) was borne out of the premise that by controlling the conditions of learning, a measure of memory is possible that is not confounded by normal age related changes in cognition.
+Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cognitive decline due to a cognitive impairment.
+The FCSRT starts with a learning phase in which participants are asked to look at a card containing four pictures (e.g., grapes) for an item that belongs to a named category(e.g., fruit).
+After these four items are identified, immediate cued recall of these four items is tested and this is repeated for a total of 16 items.
+Following this learning phase is the recall phase in which participants are asked to recall all the items identified without cues and any items which are not recalled are then cued.
+There are three outcome measures for this test, free recall (the number of items recalled without cues), total recall (the number of items recalled with or without cues) and cue efficiency which is identified as follows \cite{Grober2010}.
+\begin{equation} \label{x1}
+CueEfficiency = (totalRecall-freeRecall)/48-freeRecall, range 0.0-1.0)
+\end{equation}
+Grober found that patients with impaired free recall with four times more likely to develop dementia than those with intact free recall \cite{Grober2010}.
+This test has also been shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the researchers to used the test to categorise prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT score \cite{Sarazin2007}.
+\subsection{Conclusions}
+I have looked at a number of different cognitive tests and a battery of tests that aim to have high diagnostic and prognostic capabilities in the MCI population.
+However, particularly with the RBANS, there lacks sufficient sensitivity in differentiating those with MCI from healthy elderly individuals such that this tool could be used with a level of confidence in a clinical setting.
+In regard to the tests, a common theme is a lack of studies and therefore evidence into the utility of these tests with an MCI population.
+However if, as researcher, we aim to investigate this population further then a benchmark battery of tests which is sensitive enough in both diagnostic and/or prognostic utility should be a goal.
+\par 
+Another criticism of the current literature is the lack of consistency with regard to the experimental groups.
+For example, some studies focus on differentiating between an MCI group, a AD group and a healthy controls group whereas other studies may further subdivide the MCI category according to a number of factors such as Amnesic or Non-Amnesic MCI, Mixed MCI (sometimes called Multi-domain MCI in the literature).
+Given the inconsistency in defining the experimental groups the confusing and often conflicting results that researchers produce is to be expected.
+Future research should look at the standardization of the operational definition of cognitive impairment in MCI may result in more consistent predictions of progression to AD.
+\par 
+Finally, it can be said that in testing for MCI and comparing these results with a cohort of those with a similar diagnosis has limited utility.
+This is because there is huge variability in the presentation of MCI within participants that even if controlled for with a matched pairs design for age, gender and education will produce inconsistent results.
+There is an argument therefore for the use of longitudinal studies, where the comparisons on the performance of these tasks are with the participants themselves.
+There are not many studies which use a longitudinal approach, although those that do show promising results.
+
+\section{Types of Language Assessment}
+One of the key debates when looking at how to analyse language is the type of task provided to elicit language production in participants.
+In the literature researchers have primarily focused on Picture Description tasks but have also suggested other ways in which we might collect data.
+\par
+\subsection{Picture Description Tasks}
+One of the most commonly used tasks to measure language is the Picture Description task.
+An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Borod1980}.
+The Cookie Theft picture (pictured below) depicts a scene of a home typical of the period of time when it was created and would generally not require participants to use any complicated vocabulary to describe.
+In this task participants are asked to describe the picture presented to them in as much detail as possible.
+This task was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}
+\begin{figure}[H]
+\centering
+\includegraphics[width=240px, height=150px]{images/BCTPicture.png}
+\caption{Cookie Theft Picture - From Kaplan and Goodglass (1983)}
+\end{figure}
+\par
+The picture description task does a fine job of eliciting descriptive language however because of the specific content the language produce could be considered quite limited.
+There is some disagreement as to the benefits of this using this methodology.
+This task is reported as being useful to lexico-semantic disorders \cite{Boschi2017, Sajjadi2012} as the language being generated is primarly nouns and deixis (words to identify items and words to put those items into context).
+However, Ash \cite{Ash2012}felt that there was no difference in using this task vs Story Narration (described below).
+In explaining the differences, it is worth noting that these researchers were using differing variables and this could explain their different perspectives.
+\par
+\subsection{Narrative description task}
+The story narration task is designed to study a participant's ability to describe and elaborate on a story which is depicted using a series of pictures.
+The stories depicted are usually based on children's books or famous stories with the Cinderella being the one most typically used \cite{Fraser2014}.
+This task requires ordering the story in a structured and coherent framework.
+It also requires comprehension and understanding of the stories characters and the events depicted, as well as an awareness of a character's actions, motivations and internal reactions to given events.
+This task is particularly useful as the procedure reduces the demands on memory, due to the participant being able to access the picture book during the description and is therefore able rule out memory as a confounding variable for any results observed.
+As noted above, Ash \cite{Ash2012} felt that this task was interchangable with the Picture Description task.
+However, other research felt that this was a studier test of lexical and semantic abilities as well as syntactic complexity because this task requires interpretation and elaboration in additional to a simple description \cite{DeLira2011}.
+\par
+Given the relative strengths of the Narrative description task vs Picture description task, there are few pieces of research that have used Machine Learning to analyse features from Narrative picture tasks \cite{Fraser2014}.
+This could be due to the availability of data and the absence of any meaningful sets of transcripts of participants performing this task.
+However, this could be an interesting direction to take research in the future to see if features generated from this task could be used to predict MCI or AD.
+\par
+\subsection{Interviews}
+Interviews can also be used to elicit language in a more natural way by asking questions to guide a conversation between speakers.
+There are three types of interviews: unstructured, structured and semi-structured.
+Structured interviews tend to produce very limited speech and therefore has never been used in this area \cite{Boschi2017}.
+Unstructured interviews are open ended and generally do not conform to any particular pattern.
+They use generic themes such as family or hobbies to guide the conversation.
+Whilst this is the most ecologically valid form of conversation and therefore language generation, it's unstructured nature means that the protocol cannot be consistent and therefore reproduced.
+Semi-structured interviews are therefore preferred over other forms of interview as a middle ground.
+The semi structured nature of these interviews means that there is some replicability but does not constrain the participant in answering questions.
+\par
+The analysis of interviews can be difficult to analyse as both the content can vary even between participants, although it can be argued that content should not affect the type of language being generated unless it is narrow topic or the participant is constrained in how they answer a given question.
+It is also difficult to measure as there are no pre-defined task goals in comparison to the other two methods.
+Nevertheless, this is the most naturalistic setting for looking at language production and can be used to look at the syntactic and semantic parts of language generation \cite{Sajjadi2012}.
+There have been some attempts to use interviews to assess language production in AD with promising results \cite{Asgari2017, Guinn2015}.
+\par
+\subsection{Conclusions}
+One can view the different types of tasks above as a continuum where picture tasks represent a much more controllable task with a lot of supporting research but which generates a much more constrained set of language that is atypical of normal speech in terms of the cognitive functions used.
+
+\section{How do we analyse language, issues and debates}
+There are a wide variety of approaches that we can take in the analysis of language and there is no real consensus on the best approach to this particular problem.
+This section looks at the different ways in which researchers have looked at the problem as well as discussing some of the areas of contention.
+\subsection{Single Word Language tasks vs Connected Language tasks}
+Part of the reason we need to pay attention to how we ask participants to generate data is understanding how we wish to analyse the data afterwards.
+As discussed above, the different methodologies to collect data generate different types of language.
+There are two main approaches which we have looked at to analyse language, using frequencies of words and combinations of words and measures of syntax and semantics.
+There are other less common methods of analysing language but these are beyond the scope of this review.
+\par
+Single word tasks such as the Boston Naming Test and other such standardized language tests generally target a participants word production where this is defined as the ability to form and express words in accordance with certain criteria (see above for a full description of verbal fluency as task).
+There are a number of benefits to using single word tasks.
+From a research methodology perspective, using a standardized test allows researchers to target a very specific process in language generation and isolate factors that impact performance in language well.
+However, this approach does not take into account other cognitive processes that are used in lengthier speech tasks.
+More specifically, single word language tasks do not look at the interaction between language, executive functions and reasoning abilities.
+They also do not require the logical and efficient organisation of ideas.
+\par
+Connected language tasks, such as the picture description task and interviews described above are much more reflective of the processes involved in natural language generation.
+They are able to be relatively constrained in the language available to be used, for example in the picture description task, or they can be unconstrained in the form of interviews.
+But regardless of how they are framed, they are potentially much more useful in terms looking at all the processes involved in language generation with the drawback that you will not be able to isolate specific parts of the language generation process.
+A review by Boschi et al concludes that analysis of connected speech is potentially useful in guiding clinicians to identify language disorders \cite{Boschi2017} and also highlights the role of NLP and Machine learning in assisting in this endeavour.
+\subsection{Semantics vs Pragmatics}
+When navigating the English language it is necessary to distinguish between what a sentence says in both semantic and pragmatic terms.
+Semantic meaning refers to the meaning of the words in a sentence local only to the given sentence.
+Another way to put this is, semantics considers the meaning of words without taking into account the context in which these words are spoken.
+Pragmatic meaning refers looks at the same sentence in terms of words and grammar but takes into account the situation or context in which these words are spoken.
+Emery in her literature review looks at all levels of language tasks except for pragmatics, however this is an area that should not be ignored.
+Whilst the study of language from a pragmatic perspective is much more complex, it is perhaps one of the most vital areas to study because of the number of different cognitive processes involved.
+\subsection{Semantic Content}
+Another approach to linguistic analysis in this field is the idea of measuring semantic content and complexity.
+According to Emery (2000) \cite{Emery2000} in which she states that Semantic and Syntactic skills deteriorate first in people with MCI and AD.
+If this is true, then psychological measures of semantic and syntactic skills should be able to pick up signs of deterioration and act as markers for possible MCI and AD.
+An example of a semantic complexity measure is the concept of idea density.
+Formally, idea density is defined as the average number of propositions per sentence \cite{Keenan1973} and this was used to successfully differentiate between people who would later go on to develop AD \cite{Snowdon1996}.
+An example of semantic content measures is Type to Token Ratio (described below) which is used to measure the lexical diversity of a given piece of text and/or utterance.
+This has also shown to be effective in differentiating between MCI, AD and Controls, with those with language impairments \cite{Bucks2000} and this has carried through in research involving machine learning \cite{Wang2016, Thomas2013}.
+\par
+\subsubsection{Type token ratio(TTR)}
+Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
+\begin{equation} \label{x2}
+TTR = numberOfUniqueWords / totalNumberOfWords.
+\end{equation}
+\subsection{Thematic and Content elements in relation to the Picture description task}
+A number of studies looked at the accuracy of picture descriptions using counts of thematic and content elements within a picture.
+Whilst called diffferent names such as 'pictorial themes', 'relevant observations and 'semantic units', they all represented the same idea.
+The only difference between the studies was the number of thematic elements that 'scored' correctly.
+Nicholas et al identified eight thematic elements of the Cookie Theft picture and used the number of elements as an outcome measure in different groups.
+He found that patients with AD expressed significantly fewer content elements than controls.
+\par
+Hier, Hagenlocker and Shindler assessed content using a similar list of thematic elements \cite{Hier1985}.
+They divided their participants into early-stage and late-stage AD, as well as including a control group.
+The late-stage AD group produced significantly fewer relevant observations than the early stage group, and the AD group combined produced fewer relevant observations than controls.
+This study was replicated by Lukatela et al \cite{Lukatela1998}.
+\par
+Smith, Chenery and Murdoch applied Hier's methodology for constructing pictorial 'themes' with the Picnic Scene from the Western Aphasia Battery (WAB) with a control and patients with moderate to moderately severe AD \cite{Smith1989}.
+The authors found no difference in the number of semantic elements produced but did not that the group with moderate to moderately severe AD took more time and more syllables to communicate these elements.
+\par
+\begin{figure}[H]
+\centering
+\includegraphics[width=240px, height=150px]{images/picnic-scene.jpg}
+\caption{Picnic Scene taken from the Western Aphasia Battery (WAB).}
+\end{figure}
+Sajjadi et al examined 10 pictorial themes in picture description the Comprehensive Aphasia Test and found that the group with mild AD produced similar themes than controls \cite{Sajjadi2012}.
+Bschor et al.
+(2001) examined Cookie Theft picture descriptions at four stages of AD.
+They found that whilst each AD group differed significantly from the others and also from controls, the measures were not able to distinguish between MCI and normal controls \cite{Bschor2001}.
+\par
+Finally, a number of studies used composite measures which contained thematic elements and other unspecified information units resulting in a list of 23 possible information units of the Cookie Theft picture.
+The authors felt that this provided a wider, more liberal range of relevant content and thus subtler differences could be noted.
+Studies using these features found some differences between AD and controls, and some could differentiate between different stages of AD.
+\subsection{General Information Units or Content Information Units}
+Some studies used a more general concept of content, using terms such as "general information units" or "content information units" and this could be defined as "the smallest non redundant meaningful fact or inference,"  and was counted whether or not the information conveyed was specific to the context in which the conversation happened.
+Giles et al for example studied adults with minimal, mild or moderate AD vs controls and found that adults with AD produced fewer overall information units than controls \cite{Giles1996}.
+\subsection{Conciseness of information}1`
+Conciseness has been defined as the number of words a speaker uses to express ideas.
+The theory is that people with AD would need more words to convey ideas because of difficulties with word-finding and compensatory  behaviours such as circumlocutions and repetitions.
+Conciseness has previously been calculated by dividing the number of ideas expressed by the total number of words in a measure commonly referred to as idea density but also known as lexical index, information content and information unit conciseness index.
+Snowdon examined written discourse from the Nun study and found that low idea density in early life was associated with reduced cognitive performance in later life \cite{Snowdon1996}.
+Riley et al extended these findings by concluding that early-life idea density was associated with lower brain weight, higher degree of cerebral atrophy and increased neurofibrillary pathology in later life \cite{Riley2005}.
+\par
+Ahmed, de Jager et al examined idea density with patients who had confirmed AD post mortem \cite{Ahmed2013}.
+They found that those with AD produced fewer total semantic units than controls but there was no significant difference between the groups with regards to idea density.
+The study of "empty speech" by Nicholas et al examined conciseness and specifically looked at empty phrases (defined as common utterances which contribute no relevant information), deictic terms (e.g.
+"this", "that" without referents), indefinite terms (e.g.
+"thing" or "stuff"), pronouns without proper noun antecedants, and repetitions.
+In their study they found that AD patients produced more of these than did controls.
+\par
+\subsection{Efficiency}
+% CITE MURRAY!
+Efficiency is the rate at which meaningful information is conveyed over time,  and can be calculated by dividing the total number of information units by the duration in seconds of the speech sample.
+Smith et al, 1989 found that 18 adults AD produced fewer content units over time on average than controls, he attributed these differences to increased circumlocutions and repetitions in the AD group \cite{Smith1989}.
+Murray used an similar measure in which fillers, irrelevant words, revisions or false stars, vague or non-specific vocabulary and inaccurate output were group together as 'performance deviations" and were divided by the total number of minutes in the sample.
+This measure was lower for those with AD than those with depression, and also healthy controls.
+The authors suggested that discourse information measures may help disentangle the similarities in symptoms of early AD versus depression in older adults.
+Guinn (2012, 2015) \cite{Guinn2012, Guinn2015} found that 'Go-ahead utterances' - instances in dialogue in which a speaker provides responses do not add anything in a conversation beyond a minimal response, were significantly more frequent in those with AD than healthy controls.
+\subsection{Total number of words}
+Several studies report that adults with moderate AD produce fewer words than controls on picture description, however other studies found no differences in total words among groups of controls and patients with MCI or AD.
+Murray and Nicholas et al investigated normal controls, patients with AD and older adults with depression and found no group differences in total words \cite{Nicholas1985}.
+In contrast, Lira et al found that controls produced more total words than patients with AD but found no difference between mild and moderate groups \cite{Lira2014}.
+\par
+\section{Syntax and Morphology (Language Form)}
+Syntax can be defined as the rules that govern how words can be combined to form sentences, whilst Morphology is the system that governs the structure of words and the construction of word forms.
+Multiple studies of language decline in dementia included at least one measure of syntax or syntactic complexity \cite{Orimaye2017}.
+Common constructs included words per clause, grammatical form (measures of an appropriate use of syntactic conjections, tenses, conditionals, subordinate clauses and passive constructions), measures of phrase length and proportions of words in sentences.
+Some researchers have explored the use of formulaic language in those with dementia, the theory being that well practiced phrases are less effortful and therefore place low load on the cognitive abilities of those with AD.
+The general hypothesis motivating these studies is that either working memory limitations or semantic memory limitations in AD affect one's ability to use complex constructions.
+\par
+\subsubsection{N-grams}
+One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
+An n-gram is a contiguous sequence of n items from a given sample of text or speech.
+The items can be phonemes, syllables, letters, words or base pairs according to the application.
+For example, given the sequence of words "to be or not to be", this extract is said to contain six 1-gram sequences (to, be, or, not, to, be), five 2-gram sequences (to be, be or, or not, not to, to be), four 3-gram sequences(to be or, be or not, or not to, not to be) and so on.
+This is useful as, given a large portion of text or speech, we can predict the probability of a word being close by to a given word.
+A number of researchers have used n-grams as features.
+\par
+Asgari, Kaye and Dodge (2017) \cite{Asgari2017} used another form of word frequency measurement.
+Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words \cite{Pennebaker2015}.
+They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
+\par
+\subsection{Formulaic Language}
+Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
+They found that there were four factors which they identified as important in the classification of participants as either healthy or AD.
+These four factors were semantic impairment, acoustic abnormality, syntactic impairment and information impairment and were based on existing measures of semantic and syntactic complexity.
+Zimmerer (2016) \cite{Zimmerer2016} looked at whether language was more formulaic in those suffering from AD.
+He proposed that those who suffer from AD rely on formulatic sentences, for example 'Noun-Verb-Noun', and this is done to reduce language complexity.
+He noticed a significant difference in the use of formulaic sentences between AD and Healthy Controls.
+\par
+\section{Pragmatic Language}
+The pragmatic language domain refers to the social rules for language for the purposes of communication including, using language to achieve goals, using information from the context to achieve these goals and using the interaction between people to initiate, maintain and terminate conversations.
+\subsection{Coherence}
+Coherence, in lay terms, can be defined as the ability to maintain awareness of the topic at hand.
+It can be separated into local coherence which is related to themes of the immediately preceding utterance and global coherence which looks at how closely an utterance is related to the topic currently being discussed.
+Chapman et al used picture descriptions of Norman Rockwell prints within a frame analysis, with frames being defined as the context in which the picture is viewed \cite{Chapman1998}.
+The authors identified aspects of content, including whether the frames of interpretation being offered were typical, atypical, incorrect or had no frame.
+They also looked propositions supporting frames and propositions disrupting frames as measures of coherence.
+They examined these variables with early stage AD, old-elderly and normal controls.
+Healthy older adults and normal controls produced significantly more typical frames and more frame supporting information than the AD group.
+The authors attributed AD patients' difficulties to memory deficits, attentional deficits, visual perceptual problems, disruption of internalized frame representation, or failure to access frame knowledge.
+\par
+\subsection{Perseveration}
+Perseveration can be broadly defined as the repetition of a response regardless of the absence or cessation of stimulus which could have generated an appropriate response in the first place.
+A typical example could be the idea of conversation moving from introductions to a more general conversation, someone who has difficulties with perseveration will struggle with the move between social contexts and repeat language that would be more used in an introduction context.
+\par 
+One study examined verbal preservation in the description of Norman Rockwell prints \cite{Bayles198}.
+The presented participants with a number of similar pictures that had significantly different contexts and asked participants to describe these prints.
+They divided the total number of words within perseverations by total number of words in the speech sample.
+The authors also calculated rate of perseveration on two other language tasks: confrontation naming and generative naming.
+In all tasks, the AD group produced significantly more perseverations than controls but there were no significant differences between the two groups in the picture task in isolation.
+The authors felt that this was because picture description was an easier task, as it was a visual task in contrast to the other tasks which tapped into other cognitive processes.
+\par
+\subsection{Empty Speech}
+Verbal fluency is a term used in neuropsychological contexts generally referring to timed, word-generation tasks, while in speech-language pathology contexts, "fluency disorders" are defined as interruptions in the flow of speaking characterized by atypical rate, rhythm and repetitions in sounds, syllables, words and phrases.
+"Fluency", in the literature of discourse of adults with AD, typically refers to the smoothness or flow of spoken language.
+Abnormalities of fluency in this population are typically characterised by filled and unfilled pauses, word repetitions, circumlocutions, and revisions.
+\par
+The study of "empty speech" by Nicholas et al was one of the first to examine aspects of fluency in the connected speech of persons with AD \cite{Nicholas1985}.
+They found that adults with AD had significantly more repetitions than controls.
+Similarly, Bayles and Tomoeda found more aborted phrases, revisions and ideational repetitions in the AD group than in controls \cite{Bayles1983}.
+Several other studies support the idea that in AD population there are a greater number of repetitions and revisions than in healthy controls.
+However, there are some studies which contradict these findings\cite{Ahmed2013}.
+\subsection{Conclusions}
+This section has looked at a number of different attributes which researchers have cited as being important in the detection of MCI and AD.
+Whilst in some cases there is a broad consensus about a given feature in the vast majority of cases there is some uncertainty.
+Some of the difficulty lies in different approaches used particularly around the criteria for experimental groups which means that it is difficult to be able to compare studies, like for like, as the populations of the participants being studied vary in subtle or obvious ways.
+It is important for any studies moving forward to use comparable standards when it comes to their experimental groups, such that the picture can be made clearer.
+\section{Argument for one class classification}
+
+\section{State of literature into Machine Learning and Natural Language processing techniques}
+Diagnosing dementia through language analysis has a large background in terms of psychological research.
+Increasingly current research has called for the use of machine learning as a way of assisting in the process of diagnosis \cite{Orimaye2017, Boschi2017}.
+This section will look at what areas of natural language processing (NLP) and machine learning (ML) could potentially be used as tools to help in this domain as well as any research that has applied machine learning to this problem.
+\subsection{Natural Language Processing}
+Natural Language Processing as a topic can be defined as the intersection between Machine Learning and Linguistics and looks at a number of language tasks, particularly important in this domain is how to enable a computer to process and analyse large amounts of language data.
+Some other tasks include speech recognition, natural language understanding and natural language generation.
+Generally speaking, Natural Language Processing (NLP) systems mimic the semiotic perspective on language.
+That is to say that we can subdivide NLP processes into different components (see Figure 2.3).
+\begin{enumerate}
+	\item Morphological/Lexical: providing the basic language elements or vocabulary, such as words, their roots and inflections.
+	\item Syntactic: for grouping and sequencing elements within samples of language (usually sentences).
+	\item Semantic: for knowing the meaning of an utterance, usually defined in terms of the 'truth value' of the logical propositions that are thought to be expressed by sentences.
+	\item Pragmatic: for understanding the context and purpose of an utterance.
+\end{enumerate}
+\begin{figure}[H]
+\centering
+\includegraphics[width=240px, height=150px]{images/workflow.jpg}
+\caption{How Natural Language Processing tasks are subdivided.\label{white}}
+\end{figure}
+In order to facilitate these and other tasks, a number of frameworks have been developed that automate some of the more common processing tasks. A brief review of these frameworks follows. 
+\subsubsection{Natural Language Toolkit}
+The Natural Language Toolkit (NLTK) was originally designed as part of a computational linguistics course at the University of Pennsylvania but has evolved to become an open-source framework which includes methods and modules for a wide array of Natural Language tasks (see Table 2.2 for a full list) \cite{Bird2009}.
+\begin{table}
+	\begin{tabular}{ | p{4cm} | p{7cm} | p{1cm} |}
+		\hline
+		\textbf{Language processing task} & \textbf{Functionality} \\ \hline
+		Accessing corpora & standardized interfaces to corpora and lexicons \\ \hline
+		String processing & tokenizers, sentence tokenizers, stemmers \\ \hline
+		Collocation discovery & t-test, chi-squared, point-wise mutual information \\ \hline
+		Part-of-Speech tagging & n-gram, backoff, Brill, HMM, TnT \\ \hline
+		Machine Learning & decision tree, maximum entropy, naive Bayes, EM, k-means \\ \hline
+		Chunking & regular expression, n-gram, named-entity \\ \hline
+		Parsing & chart, feature-based, unification, probabilistic, dependency \\ \hline
+		Semantic interpretation & lambda calculus, first-order logic, model checking \\ \hline
+		Evaluation metrics & precision, recall, agreement coefficients \\ \hline
+		Probability and estimation & frequency distributions, smoothed probability distributions \\ \hline
+		Applications & Graphical concordance, parsers, WordNet browser, chatbots \\ \hline
+		Linguistic Framework & manipulate data in SIL Toolbox format \\ \hline
+	\end{tabular}
+		\caption{\label{tab:table-name}NLTK tasks and functionality}
+\end{table}
+The NLTK has been successfully used in both teaching and research purposes and as part of this provides access to large array of databases and corpora which have potential uses in common tasks.
+
+\subsubsection{CoreNLP}
+Stanford CoreNLP provides a set of human language technology tools.
+This comes as an installable package that requires JAVA and can be interacted with through the command line or by an accompanying API.
+Whilst NLTK is much more oriented to parsing text and generating features, CoreNLP does this and adds more functionality such as a Named Entity Recogniser and sentiment analysis functions.
+At it's centre is the Stanford parser, which parses text and can give the base forms of words, their parts of speech, whether they are names of companies, people, etc., normalize dates, times, and numeric quantities, mark up the structure of sentences in terms of phrases and syntactic dependencies, indicate which noun phrases refer to the same entities, indicate sentiment, extract particular or open-class relations between entity mentions, get the quotes people said, etc.
+\par
+\begin{figure}[H]
+\centering
+\includegraphics[width=350px, height=200px]{images/CoreNLP.png}
+\caption{Depiction of how Core NLP marks up a sentence}
+\end{figure}
+It's intended for CoreNLP to provide tools and frameworks that will enable higher level language analysis whilst remaining domain agnostic.
+
+\subsubsection{Linguistic Inquiry and Word Count}
+The Linguistic Inquiry and Word Count (LIWC) is a piece of software that counts words and assigns these counts to categories \cite{Pennebaker2015}.
+The rationale behind this is that the words we speak inform an observer of the state of mind of an individual.
+A typical example of this might be someone with depression who may typically be more inwardly focused and therefore talking in the first person is common\cite{Al-Mosaiwi2018}.
+The 2015 version of the LIWC has over 90 different categories \cite{Pennebaker2015} and some supporting research suggests that it is effective at identifying themes and associating these themes with mental health difficulties such as depression and anxiety \cite{Sonnenschein2018}.
+\par
+Whilst not necessarily a tool that can parse data and look at trends from a linguistic perspective, a framework like this could be used as another tool to explore themes within a persons speech from a more qualitative angle.
+To the authors knowledge, there is no research that looks at speech changes from this perspective.
+
+\subsection{Traditional methods of Machine Learning}
+In terms of Machine Learning research in this area, a number of researchers have used transcripts based on picture description tasks \cite{Zimmerer2016, Orimaye2017, Mueller2018a, Fraser2015} and have successfully extracted linguistic features that could differentiate between AD and controls.
+\par
+One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the most frequent class.
+They detail several lexical approaches to the problem of detecting and rating AD.
+The approaches they looked at relied primarily on character n-gram techniques but also explore correlation of usage of frequency of different types of speech.
+Their results act as a proof in concept of the utility of using a pure computational approach to the diagnosis of dementia using spontaneous speech.
+The were able to obtain 95\% accuracy dementia vs controls, 70\% accuracy classifying dementia into two categories and 50\% accuracy classifying dementia into four categories.
+They suggest further exploration of characteristics to classify \cite{Thomas2005}.
+\par 
+% Cite CRUM Population-based norms for the Mini-Mental Status Examination by age and educational level
+Bucks et al \cite{Bucks2000} took a more traditional approach using frequency rates per 100 words of various word types such as nouns, verbs, adjectives and pronouns and some measures of lexical richness to attempt to discriminate between healthy older participants and participants with probable AD.
+They found that AD patients had higher mean pronoun rate, adjective rate and verb rate but lower noun rate vs normal older controls \cite{Bucks2000}.
+It's important to note that the clinical group averaged a score on the MMSE of 15 (sd=6.8) which puts the clinical group in the 'moderately impaired range'.
+At this stage, the two groups should be reasonably easy to differentiate in terms of language and other cognitive abilities.
+\par 
+Fraser et al \cite{Fraser2015} built upon the work by Bucks by adding an acoustic analysis element to her analysis using the DementiaBank corpus, as well as significantly expanding the list of linguistic features they used.
+They used features generated from the Stanford parser (Now known as Core NLP) and computed the frequency of the occurrence of different parts of speech which they then normalised by the total number of words in each utterance.
+They also computed ratios between these parts of speech such as pronouns to nouns.
+They also included measures of syntactic complexity such as mean length of sentences and measures of lexical richness such as TTR.
+Their analysis showed was able to achieve 81\% accuracy in distinguishing individuals with AD from controls based on short samples of speech.
+They found that there are four factors which they felt were important in discriminating between these two groups: semantic impairment, acoustic abnormality, syntactic impairment and information impairment.
+\par 
+Zimmerer et al \cite{Zimmerer2016} took a slightly different approach in that they developed the Frequency in Language Analysis Tool (FLAT) which looked at the degree of formulaicity in an individual sample.
+The rationale behind this is that people with impaired cognitive function are able to depend on formulaic sentences for language generation to ease the burden of language generation on cognitive symptoms.
+They were able find that those with probable AD produced more formulaic language than controls.
+\par 
+Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls.
+Their main finding supported n-grams as the most significant predictor.
+One of the criticisms is the use of picture description tasks and n-grams.
+Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
+\par
+
+\subsection{The case for Deep Learning}
+One of the criticisms of traditional learning models is it's reliance on features that are generated for them.
+As we can see from the brief look at some of the research in this field, researchers have taken a number of different approaches with success.
+However, there does not seem to be a consensus.
+As machine learning in general moves away from traditional machine learning models to deep learning techniques, it poses the question can deep learning assist with this problem?
+\par 
+One of the main benefits of deep learning is that it does not rely on pre defined features being fed to the model.
+Instead, deep learning models take raw data in with some amount of preprocessing and generates it's own features.
+This move away from relying on features solves one of the difficulties we have with the current psychological literature, namely that there are some disagreement on the which
+\par 
+One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
+This is the case with all machine learning techniques, but more so with deep learning.
+There have been ways to 'create' more data such as data augmentation.
+
+\subsection{Conclusions}
+We can see that both Natural Language Processing and Machine Learning techniques have a lot to offer, Indeed there has been a lot of research which have used pre-existing datasets to explore this area with promising results.
+One of the difficulties with the current research is the approach of trying to discriminate between those AD and healthy controls.
+This is not necessarily a problem in the real world as it is trivially easy to do for trained clinicians.
+A more interesting, but potentially harder to problem to solve is to discriminate between those with MCI and healthy controls, and more importantly to track their decline over time.
+Further, to date the authors are unaware of any research in which these techniques are applied to newly created samples of language.
+
+\section{Discussion}
+Current state-of-the-art diagnostic measures of AD are invasive (CSF analysis), expensive (neuroimaging), and time-consuming (neuropsychological assessment).
+Furthermore, these measures are limited to speciality clinics and thus have limited accessibility as frontline screening and diagnostic tools for AD.
+More importantly, nonspecialists are often inaccurate at identifying early AD and MCI.
+Thus, there is an increasing need for additional noninvasive and/or cost-effective tools, allowing effective frontline identification of subjects in the preclinical or early clinical stages of AD who could be suitable for monitoring in speciality clinics and for early treatment.
+Implementation of effective screening instruments will allow diagnosis earlier in the course of dementia, even at the point when memory function is still essentially within the normal range.
+This strategy would enable an earlier, and potentially more effective, prevention and treatment of AD with a special focus to preserve cognitive functions.
+\par 
+However the literature has identified a number of challenges when approaching this problem.
+Firstly, the clinical features that combine to meet the diagnostic criteria for Dementia or it's variants are continuous in nature and heterogenous between patients and are also impacted by other variables.
+For example as shown above, cognitive performance is affected in part by a patients educational attainment and a patients ability to live independently is impacted by a their physical health as much as their cognitive health.
+The challenge therefore is to find features that are minimally impacted by other factors, or that can be controlled for by a experimental design such as a matched pairs design to control for educational attainment.
+\par 
+Another challenge lies with the recruitment of suitable individuals who may notice a decline in cognition to the point where we might classify them as having MCI, but these individuals deduce that there is little to no value to admitting there is a problem and seeking help whilst their symptoms are 'manageable'.
+\par
+Finally there is a large amount of variability in the presentations of those with MCI and early dementia, and this is compounded by an similar amount of variability in the criteria researchers have used for experimental groups and the approaches researchers have used to tackle this problem.
+This had led to a confused literature.
+Recently a call for research that has consistent inclusion / exclusion criteria has been made along with some proposed definitions of MCI and it's subgroups \cite{Petersen2014}.
+Researchers have identified the analysis of language impairment as an area of promise to explore in the diagnosis of MCI and early AD and recent developments in natural language processing and machine learning techniques have the potential to assist in this research.
+Indeed, given the increased burden on the diagnosis of MCI and AD on professionals there has been a call to use technology to potentially ease this burden \cite{Boschi2017}.
+A small but growing amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
+
+\subsection{Future Work}
+One of the areas for research to study is a careful examination of the features that are being used to measure language deterioration.
+For example, Zimmerer (2016) \cite{Zimmerer2016} describes connectivity in such as a way that correlates directly with what Mueller (2018) \cite{Mueller2018a} calls Fluency.
+Whilst these are very nuanced measures which differ slightly in the form they take, the sheer range of measures and features being produced make it difficult to organise and explore what is truly going on in those with MCI and early AD..
+Some work needs to be done in producing a consistent list of measures that are validated using existing datasets and can be used for future research moving forward.
+\newline
+\par
+Teng et al suggests that work should focus on the MCI population and concentrate on developing a consensus neuropsychological battery that could yield predictable rates of progression to AD \cite{Teng2009}.
+This, in conjunction with the development of a model of language, sensitive enough to detect subtle deterioration in language use to act as an additional cognitive marker to aid diagnosis could potentially move some way to providing this.
+\par 
+Future research should also be directed towards developing non-intrusive ways of detecting subtle changes in natural language such that any perceived deterioration that could indicate the presence of MCI or AD could be flagged up early.
+Machine learning approaches seem to be the most logical approach for achieving this aim as language could be collected in non-intrusive ways and passed to a machine learning algorithm for preliminary classification.
+Despite the excellent quality of datasets, for example the DementiaBank dataset, being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which some researchers could claim are limited in scope.
+Finally, the recent resurgence in the use of neural networks and deep learning could provide the answer to the confused literature in terms of features.
+A key benefit of deep learning is it's ability to automate the process of feature engineering.
+So there is an opportunity to explore the use of deep learning, to not only develop new features but also validate existing features independently.
+\par
+This area of research is extremely promising in its early results and the impact of successful research would be life changing for both individuals and the health of the worlds aging population in general.
+
+\section{Introduction}
+Dementia has been identified as one of the fast growing difficulties facing the world.
+A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
+\par
+A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}.
+As described above \cite{Prince2015}, the numbers of those suffering from AD and MCI are going to increase as the population ages and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD.
+At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017}.
+I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
+\par
+There is a large body of research that looks at language deterioration in those suspected to have MCI or AD \cite{Taler2008, Boschi2017}.
+However there is conflicting evidence in these studies about which declining language factors are associated with MCI and AD \cite{Taler2008, Boschi2017}.
+Research therefore, should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression.
+Another area of focus for research of this nature is the process of collecting appropriate language samples.
+Whilst collecting samples of language is comparatively unintrusive, researchers recognise that these samples require a rich sample of language that potentially cannot be generated by tasks such as the picture description task.
+Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD.
+There is some evidence to support this.
+Berisha et al \cite{Berisha2015}, has shown through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control.
+\par
+The question to be addressed in this systematic review is how has the field of machine learning and natural language processing addresses language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease.
+The potential impact of this research in this area is immense.
+Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease however the absence of a single test and the complexity of AD can create significant delays in diagnosis.
+Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
+\par 
+The remainder of this article is organized as follows.
+Section~\ref{methodology} gives an account of the process of this Systematic Review.
+Our results are described in Section~\ref{results}.
+We discuss the results and implications in Section~\ref{discussion}.
+Finally, Section~\ref{conclusions} gives the conclusions.
+
+\section{Methodology}\label{methodology}
+A systematic literature review (SLR) describes a process which aims to identify, evaluate and interpret the research and literature in a given area.
+They are designed to provide a complete and exhaustive summary of the current evidence relevant to an identified research question.
+SLR's conduct a thorough search of all literature following a pre-defined protocol that specifies focused research questions, identifies criteria for the selection of studies and assessment of their quality, and forms to execute the data extraction and synthesis of results.
+\par 
+Common motivations for conducting an SLR are: 
+\begin{enumerate}
+	\item to summary all the evidence about a topic.
+	\item find gaps in the research. 
+	\item to provide a ground for a fundament to new research.
+	\item and to examine how the current research supports a hypothesis. 
+\end{enumerate}
+
+Performing an SLR comprises the following steps:
+
+\begin{enumerate}
+	\item identify the need for performing the SLR.
+	\item formulate research questions.
+	\item execute a comprehensive search and selection of primary studies.
+	\item assess the quality and extract data from the studies.
+	\item interpret the results.
+	\item report the SLR.
+\end{enumerate}
+
+\subsection{Search strategy}
+The main research question this SLR aims to address is: “How has the field of machine learning and natural language processing addressed language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease?”.
+
+This SLR builds upon these questions and additionally presents the results of the other two additional research questions.
+Further, the key terms related to MCI and the other names by which it is known were included in the search string to ensure that relevant studies about machine learning and the diagnosis of a disease were also retrieved, even if not specifically mentioned in the paper's title or abstract.
+\par 
+To address the research questions, a search string was defined using the PICO approach, which decomposes the main research question into four parts: 
+\begin{enumerate}
+	\item Population - Studies that present research on mild cognitive impairment and dementia.
+Mild Cognitive Impairment and Dementia keywords were selected from the Systematized Nomenclature of Medicine–Clinical Terms (SNOMED-CT).
+	\item Intervention - Intervention: ML or Statistical Modelling techniques that focus on classification.
+The ML keywords were selected from the branch “Machine Learning Approaches” of the “2012 ACM Computing Classification System”.
+The MS keywords were selected by A2.
+	\item Comparison
+	\item Outcome - Outcome: Prognosis on dementia and comorbidies. The prognosis keywords were provided
+by A4.
+\end{enumerate}
+
+Using this process, the main research question was decomposed  into four research questions:
+\begin{itemize}
+	\item RQ1: What features / data characteristics of text (variables, determinants and indicators) that are considered when applying the ML techniques (n-grams, PoS Tagging etc)?
+	\item RQ2: Which NLP and ML or Statistical Learning techniques are being used in dementia research?
+	\item RQ3: What are the goals of the studies that employ NLP / ML techniques for prognosis of dementia?
+	\item RQ4: Do the studies focus on time as factor?
+\end{itemize}
+\par 
+The automated searches were performed in the Pubmed, Web of Science, Scopus and IEEE databases.
+Table 1 shows the search string used for the Pubmed automated search, but note that this search string was adapted to each of the other databases’ search context.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{12cm} | }
+	\hline
+	(dementia OR MCI OR Mild Cognitive Impairment OR Alzheimer's OR Mild Neurocognitive Disorder OR AD) AND TOPIC: (machine learning OR Data Mining OR Decision Support System OR NLP OR Natural 			Language Processing) AND TOPIC: (prognosis OR prognostic estimate OR predictor OR prediction OR model OR patterns OR diagnosis OR diagnostic OR forecasting OR projection OR Deep Language Model 		OR Deep Neural Network) AND TOPIC: (classification OR regression OR kernel OR support vector machines OR Gaussian Process OR Bayesian Network OR Factor Analysis OR Deep Learning OR Neural 			Networks OR Maximum Likelihood OR Principal Component Analysis OR Markov OR Linear Model OR Mixture Model OR Perceptron Algorithm OR Logical Learning OR relational learning OR Supervised 				Learning OR Unsupervised Learning OR clustering OR Decision Tree) AND TOPIC: (Language OR Cognitive OR Speech OR Conversation OR Connected Speech OR Picture Description OR Discourse Analysis OR 		Verbal Fluency)  \\ \hline
+	Searched - 4th April 2019 - Generated 1257 Articles \\
+	\hline
+	\end{tabular}
+	\end{center}
+	\caption[Table caption text]{Example of Search Terms for Web of Science database}
+	\label{table:name}
+\end{table}
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{3cm} | p{3cm} | p{3cm} | }
+	\hline
+	Scopus & Web of Science & PubMed & IEEE Xplore \\ \hline
+	1002 & 991 & 376 & 230 \\ 
+	\hline
+	\end{tabular}
+	\end{center}
+	\caption[Table caption text]{Number of articles found from each database - complete search terms appear in Appendix B}
+	\label{table:name}
+\end{table}
+
+\subsection{Study selection} 
+A total of 1490 unique papers were identified through the searches conducted above.
+Each paper was initially reviewed by just the title and the abstract.
+This yielded a total of 25 potential papers.
+Papers were then evaluated by the JA based on the inclusion and exclusion criteria  (see Table 3).
+Where JA could not reach a decision, PS and GV were consulted and majority vote on inclusion was conducted.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{6cm} | p{6cm} | }
+	\hline
+	Inclusion Criteria & Exclusion Criteria \\ \hline
+	Be a primary study in English; AND address research on dementia and comorbidities; AND address at least one ML or MS technique; AND address a prognosis related to dementia and comorbidities; AND use cognition or language decline as a factor for analysis.
+&
+	Be a secondary or tertiary study; OR be written in another language other than English; OR do not address a research on dementia and comorbidities; OR do not address at least one ML or MS technique; OR do not address a prognosis related to dementia and comorbidities.
+\\
+	\hline 
+	\end{tabular}
+	\caption[Table caption text]{Inclusion and Exclusion Criteria}
+	\label{table:name}
+	\end{center}
+\end{table}
+
+After the identification of these papers, a one-iteration backward snowballing process was carried out using the reference lists of the original set of papers looking for studies that were missed in the original searches.
+This resulted in 41 additionally identified papers.
+Throughout the whole selection process, PS and GV acted as additional assessors in the case where there was uncertainty about whether a paper should be included.
+\par 
+In total, 66 papers were selected to be fully read.
+A quality assessment questionnaire (see Table 4) was developed based on Kitchenham's guidelines and was used to minimize the chance of bias in the selection process.
+Studies were graded on a 12 point scale and any studies which scored less than 8 points were excluded for quality reasons.
+The ones that successfully passed the filtering criteria described earlier had their relevant data extracted.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{6cm} | p{6cm} | }
+	\hline
+	Variable & Definition \\ \hline
+	Conditions Studied & For which dementia disorder is the study deriving a prognosis. \\ \hline
+	Database used in the study & Name and origin of the data source used to derive the prognosis of the studied dementia. \\ \hline
+	Dataset Categories & Classes in which the data units were divided into. \\ \hline
+	Follow-up period & Period of time, which the data units were followed. \\ \hline
+	Techniques used & Natural Language techniques AND/OR ML techniques that were used to build the diagnostic models. \\ \hline 
+	Features generated & NLP generated features used in building the diagnostic models. \\ \hline
+	Aim of the Study & The goal of the built diagnostic models. \\
+	\hline 
+	\end{tabular}
+	\caption[Table caption text]{Quality Assessment Questionnaire}
+	\label{table:name}
+	\end{center}
+\end{table}
+
+In this phase, a paper could also be rejected due to inclusion and exclusion criteria because the selection process adopted an inclusive approach.
+This means that during the reading of the titles and abstracts, in the case where the information provided was incomplete or too general it was selected to be fully read in the posterior phase.
+A common example is the case when the data analysis technique specified in the abstract was merely “classification”, so it was not possible to know if any machine learning occurred.
+\par
+% NUMBERS!
+In total, 37 studies composed the final set of included primary studies and had their relevant data extracted, 7 papers were rejected due quality reasons, and 34 papers were rejected due to failing the inclusion and exclusion criteria.
+One reason for the high number of the latter was the decision to exclude the papers that used solely statistical methods as data analysis techniques to build the prognostic models.
+The selected studies were also assessed for the risk of cumulative evidence bias.
+This was done by checking, in the case of the same research group with different studies in the final set of included primary studies, if it was justified having both studies (i.e different samples).
+
+\subsection{Data collection}
+For the data collection, a base extraction form was defined in the protocol, but later in the study it was evolved based on the research group discussions.
+Table 4 lists and defines the collected variables.
+
+In addition to these variables other basic data about the studies was collected, these were: title, authors, journal/source, year and type of publication.
+No summary measures were used.
+Summary tables were used for the synthesis of results and no additional analyses were carried out
+
+\section{Results}\label{results}
+\subsection{Features of Language}
+One of the first steps in building an accurate diagnostic model using language is to analyse the effectiveness of the various language features that are used.
+
+\par
+Asgari, Kaye and Dodge (2017) \cite{Asgari2017} used another form of word frequency measurement.
+Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words \cite{Pennebaker2015}.
+They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84\%.
+\subsubsection{Measures of Semantic Complexity}
+Intro!
+
+Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
+\begin{equation} \label{x1}
+TTR = numberOfUniqueWords / totalNumberOfWords.
+\end{equation}
+\par 
+Brunet's Index (W) differentiates itself for TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
+\begin{equation} \label{x2}
+W = N^{V(-0.165)}
+\end{equation}
+where N is the total length of the utterance being measured and V is equal to the total vocabulary being used by the subject.
+Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers.
+\newline
+\par 
+Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation: 
+\begin{equation} \label{x3}
+%% Check to see if this is accurate.
+R = (100 \log N) / (1 - V1/V)
+\end{equation}
+where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
+\par 
+Guinn, Singer and Habash \cite{Guinn2015}, they found in their corpus that there was no significant difference between interviewers and those with dementia when applying these measures.
+However, when they compared these results with a control dataset, they did find a significant difference in Honore's statistic.
+They explained this interesting results by suggesting that the interviewer used was trying to match (intentionally or unintentionally) the lexical richness of the person they were interviewing.
+In comparing healthy older adults with those suffering from dementia, they found that there was an increase in lexical diversity which is contrary to other research.
+They did note that conversations involving those with dementia contained roughtly 50\% less speech than those with controls, and that TTR in particular is not a suitable measure because it is more sensitive to length.
+They also note that Brunet's Index and Honore's Statistic are better statistics that control for the total length of the conversation and controls had statistically greater lexical diversity on those measures.
+
+\subsection{Quantity - Total number of words}
+Several studies report that adults with moderate AD produce fewer words than controls on picture description, however other studies found no differences in total words among groups of controls and patients with MCI or AD.
+Murray and Nicholas et al investigated normal controls, patients with AD and older adults with depression and found no group differences in total words.
+In contrast, Lira 2014 found that controls produced more total words than patients with AD but found no difference between mild and moderate groups.
+\newline
+\par
+\subsection{Syntax and Morphology (Language Form)}
+Syntax can be defined as the rules that govern how words can be combined to form sentences, whilst Morphology is the system that governs the structure of words and the construction of word forms.
+Multiple studies of language decline in dementia included at least one measure of syntax or syntactic complexity.
+Common constructs included words per clause, grammatical form (measures of an appropriate use of syntactic conjections, tenses, conditionals, subordinate clauses and passive constructions), measures of phrase length and proportions of words in sentences.
+Some researchers have explored the use of formulaic language in those with dementia, the theory being that well practiced phrases are less effortful and therefore place low load on the cognitive abilities of those with AD.
+The general hypothesis motivating these studies is that either working memory limitations or semantic memory limitations in AD affect one's ability to use complex constructions.
+\par
+\subsection{N-grams and skip-grams}
+One of the first features discussed as a potential predictor of MCI or AD is the n-gram.
+An n-gram is a contiguous sequence of n items from a given sample of text or speech.
+The items can be phonemes, syllables, letters, words or base pairs according to the application.
+For example, given the sequence of words "to be or not to be", this extract is said to contain six 1-gram sequences (to, be, or, not, to, be), five 2-gram sequences (to be, be or, or not, not to, to be), four 3-gram sequences(to be or, be or not, or not to, not to be) and so on.
+This is useful as, given a large portion of text or speech, we can predict the probability of a word being close by to a given word.
+A number of researchers have used n-grams as features.
+One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class.
+Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls.
+Their main finding supported n-grams as the most significant predictor.
+One of the criticisms is the use of picture description tasks and n-grams.
+Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
+\par
+Skip-grams are a variant of n-grams in which word tokens are skipped intermittently while creating n-grams.
+For example, take the sentence 'I am going to London', there are four conventional bigrams: 'I am', 'am going', 'going to', and 'to London' - using skip-grams, we might skip a word to create additional bigrams such as: 'I going' and 'going London'.
+Orimaye defined k-skip-n-grams as a set of n-gram tokens with the following equation, where n is the specified n-gram (e.g.
+2 for a bigram and 3 for a trigram), m is the number of tokens in a given sentence, k is the number of word skip between n-grams given that k < m and a = \{1,..., m-n\}
+\par 
+One problem with this approach is existing research currently uses language generated from picture description tasks.
+Given the nature of these tasks, the language generated is relatively constrained in comparison to language generated spontaneously.
+% Question How does using ngrams with constrained language differ from ngrams with general speech?
+
+\begin{equation} \label{x4}
+T_{n-gram} = {W_{a},...,W_{a+n-k},..., W_{a+n},....,W_{m-n},...,W_{(m-n)+n-k},...,W_{m}}
+\end{equation}
+
+Thus for the sentence 'I am going to London', 1-skip-2-grams will give \{I going, am to, going London\} and 1-skip-3-grams will give \{I going to, I am to, am to London, am going London\}
+
+\subsection{Mean length of utterance (MLU)}
+Murray found that MLU was not a distinguishing factor among health adults, adults with depression and adults with AD.
+Ripich et al found a decrease in MLU in adults with severe AD over time, and this was supported by findings of Le et al in their studies of authors \cite{Le2012} \newline
+\par 
+\subsection{Proportion of verbs to nouns plus verbs}
+Kave and Levy used a verb index to capture syntactic complexity and found that adults with AD expressed the same amount of verbs to nouns plus verbs as adult controls. \newline
+\par 
+\subsection{Syntactic Complexity - Composite measures of MLU, syntactic errors and verbs}
+Ahmed et al, and Ahmed, Haigh et al found differences in syntactic complexity between adults with MCI and controls, and between MCI and moderate AD stages.
+The differences in syntactic complexity were not significant when individual measures were tested, but were apparent using a composite score consisting of MLU, words in sentences, syntactic errors, nouns with determiners, and verbs with inflections.
+\newline
+\par
+Lu's Syntactic Complexity Analyser
+Ygnve measure
+
+\subsection{Semantic features}.
+We compute semantic similarity using the average and minimum cosine distance between each pair of one-hot embeddings of utterances, and the cosine cutoff (i.e., the number of pairs of utterances whose the cosine distance is below a certain threshold).
+We compute word specificity and ambiguity based on tree depth and the number of senses in WordNet [54].
+We also extract multiple WordNet measures of similarity: Resnik [68], Jiang-Coranth [69], Lin [47], Leacock-Chodorow [70], and Wu-Palmer [71].
+
+\subsection{Syntactic features}
+Guinn, Singer and Habash \cite{Guinn2015} found that syntactic features such as Noun rate, verb rate, adjective rate and pronoun rate were not significantly different between interviewers and those with dementia in their dataset, although they did note that there was a slightly higher but non significant use of pronouns.
+
+\subsection{Pragmatic features.} We train a general 100-topic latent Dirichlet allocation (LDA) model [72] on the Wikipedia corpus for generalizability.
+LDA is a generative statistical model used to determine unlabeled topics in a document.
+For each transcript, we extract the probabilities of each LDA topic.
+Next, we extract features related to rhetorical structure theory (RST), which is a classic framework for discourse parsing in which partitions of text are arranged in a tree structure by pragmatic relations such as Elaboration or Contrast [73].
+
+\subsubsection{Go-ahead Utterances}
+Go-ahead utterances are defined as short one or two syllable responses that do not contribute to the conversation beyond a minimal response.
+The function of these go-ahead utterances can be to validate what the other person is saying, or to agree / disagree with what is being said.
+Another function can be that they wish speaker is indicating that they have nothing futher to add to the conversation and a signal to the other speaker to continue with what they are saying.
+In Curry, Singer and Habash's research, they found that in comparing interviewers and those with dementia, that interviewers used significantly fewer go-ahead utterances.
+In comparing controls and those with dementia, they found that there was a relative lack of go-ahead utterances in the controls which implies that controls had a lot more contributions to make in their conversations than those with dementia.
+\par
+
+\subsection{Formulaic Language}
+Fraser, Meltzer and Rudzicz (2015) \cite{Fraser2015} looked at connected speech using the DementiaBank corpus.
+They found that there were four factors which informed the classification of participants as either healthy or AD.
+These four factors were semantic impairment, acoustic abnormality, syntactic impairment and information impairment and were based on existing measures of semantic and syntactic complexity.
+Zimmerer (2016) \cite{Zimmerer2016} looked at whether language was more formulaic in those suffering from AD.
+He proposed that those who suffer from AD rely on formulatic sentences, for example 'Noun-Verb-Noun', and this is done to reduce language complexity.
+He noticed a significant difference in the use of formulaic sentences between AD and Healthy Controls.
+\newline
+
+\subsection{Number of syllables and Characters}
+
+\subsection{Number of fillers}
+
+\subsection{Readability}
+Flesch reading score, Flesch-kincaid grade level
+
+\subsection{Polarity}
+
+\subsection{Frequency}
+Mean values of frequency, age of acquisition, imageability, familiarity, arousal, dominance and valence based on lexical norms
+
+\subsection{Dysfluencies}
+Curry, Singer and Habash noted that in comparing controls and those with Dementia, that those with Dementia had a significantly higher number of pauses per word and a much higher incidence of words that were truncated in mid-speech.
+In comparing interviewers with those with dementia, they also showed other signs of difficulties with fluency with higher rates of incomplete words, filler words and repeated words.
+
+\section{Machine Learning methods}
+The next stage of building statistical or Machine Learning models is to take some features and use these feature to build models able to classify participants into various categories.
+There are various types of classification models that have been used to tackle this problem, this section looks which models have been used in the current literature.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ p{6cm} | p{6cm} }
+	\hline
+	Machine Learning technique used & Paper Number \\ \hline
+	Logistic Regression & 2 \\ \hline
+	Support Vector Machines & 4 \\ \hline
+	Naive Bayes & 1 \\ \hline
+	Decision Trees & 1, 2 \\ \hline
+	Multi-layered perceptrons & 2 \\ \hline
+	\hline 
+	\end{tabular}
+	\end{center}
+\end{table}
+
+\subsection{Traditional Machine Learning methods}
+\subsubsection{Logistic Regression}
+Logistic Regressions is a probabilistic approach to classification and is dependant on the assumption that your input space can be separated into two distinct regions, one for each class, by a linear boundary.
+In 2D space, this boundary is a line and in 3D space, this boundary is defined as a plane.
+\par 
+In the results of this survey.
+
+\subsubsection{Support Vector Machines}
+Support Vector Machine (SVM) was originally proposed as an algorithm for classification problems; it is a relatively new technique compared to the other ML approaches.
+The classification process consists of mapping the data points (usually the study subjects) into a feature space composed of the variables that characterize these data points, except for the outcome variable.
+Then, the algorithm finds patterns in this feature space by defining the maximum separation between two or more classes, depending on the problem to be solved.
+Contrary to some regression techniques, SVMs are not dependent on a pre-determined model for data fitting, although there are still algorithm specifications to be considered (e.g.
+choice of a kernel function); instead, it is a data-driven algorithm that can work relatively well in a scenario where sample sizes are small compared to the number of variables, reason why it has been widely employed by diagnostic studies in tasks related to the automated classification of diseases.
+\par 
+Regarding the SLR results, SVMs were present in ??/41 selected studies, in 38 proposed models, and being by far the most used machine learning technique in the dementia diagnosis research.
+These numbers account for the traditional SVM and variations.
+In all of the ??
+selected studies the SVMs focused at binary classifications where the task was to discriminate mild cognitively impaired (MCI) patients that will or will not develop Alzheimer's Disease (AD).
+In the general case, the problem is posed as either MCI converters versus MCI non-converters, or progressive MCI versus stable MCI classification.
+This outlines a situation in which a regression problem (when will the MCI patients convert to AD?) is formulated as a classification problem (which MCI patients will convert to AD in X months?) to be solved.
+Reasons for this could be due to limitations in the data used, i.e.
+the limited follow-up periods of the subjects included in the studies.
+
+\subsubsection{Linear Discriminant Analysis}
+
+\subsubsection{Decision Trees}
+A Decision Tree (DT) is a classification algorithm in which the learned knowledge is represented in a tree structure that can be translated to if-then rules.
+DT's learning process is recursive and starts by testing each input variable as to how well each of them, alone, can classify the labeled examples.
+The best one is selected as a root node for the tree and its descendant nodes are defined as the possible values (or relevant ratios) of the selected input variable.
+The training set is then classified between the descendant nodes according to the values of the selected input variable.
+This process is repeated recursively until no more splits in the tree are possible.
+Like SVMs, DTs do not depend on a pre-defined model and are mostly used to find important interactions between variables.
+Being intuitive and easy to interpret, DTs have been used in prognostic studies as a tool for determining prognostic subgroups.
+\par 
+In this SLR, DTs were the second most frequently used ML technique, present in 6/37 selected studies and proposed in 7 models.
+It was employed for the same reason as SVM; except for one study that investigated the evolution of patients diagnosed with cognitive impairment no dementia (CIND) to AD.
+
+Curry, Singer and Habash - Alz (66.7 - Pres, 66.7 - Recall) and Control (67.9 - Pres, 67.9 Recall)
+
+\subsubsection{Naive Bayes}
+
+Curry, Singer and Habash - Alz(80.8 - Pres, 0.75 - Recall) and Control (79.3 - Pres, 82.1 Recall) - In the naive bayes classifier, they identified that pauses, go-ahead, fillers and incomplete words as the most significant features.
+%Why is Recall at 0.75???????
+
+\subsubsection{Artificial Neural Networks}
+An Artificial Neural Network (ANN) is a methodology that performs multifactorial analyses, which is desirable in the health area as medical decision-making problems are usually dependent of many factors.
+An ANN is composed of nodes connected by weighted edges in a multi-layer architecture that comprises: an input layer, one or more hidden layers and an output layer.
+In the training process, inputs and outputs values are known to the network, while the weights are incrementally adjusted so that the outputs of the network are approximate to the known outputs.
+Despite being a powerful predictor, ANNs are `black boxes', which means that they are not able to explain their predictions in an intuitive way, contrary to DTs or BNs.
+Also, they require the specification of the architecture to be used beforehand (i.e.
+the number of hidden layers).
+
+\subsubsection{K-nearest neighbours}
+K Nearest Neighbors (KNN) is a classification algorithm that takes a data point from an unknown class and assigns it as an input vector in the feature space.
+Then, the classification process follows by assigning the unknown class data point to the class in which the majority of the K nearest data points belong to.
+The distance between data points is usually measured by Euclidean distance, but it is possible to employ other measures.
+KNN is one of the simplest ML classification algorithms and have been used in a wide range of applications; however, it can be computationally expensive in a highly dimensional scenario.
+Further, it considers all features to be equally weighted, which can be a problem if the data has superfluous attributes.
+
+\subsubsection{Best Results}
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ p{6cm} | p{6cm} }
+	\hline
+	Machine Learning technique used & Paper Number \\ \hline
+	Logistic Regression & 2 \\ \hline
+	Support Vector Machines & 4, 5 \\ \hline
+	Naive Bayes & 1 \\ \hline
+	Decision Trees and Random Forest Classifiers & 1, 2, 5 \\ \hline
+	Multi-layered perceptrons & 2 \\ \hline
+	\hline 
+	\end{tabular}
+	\end{center}
+\end{table}
+
+\subsection{Deep Learning methods}
+\subsubsection{Deep language space neural network for classifying mild cognitive impairment and Alzheimer-type dementia - Orimaye, Wong and Wong (2018)}
+In this paper, Orimaye et al use deep-deep neural networks language models (D2NNLM) to learn linguistic changes that distinguish the language of patients with MCI and AD-type dementia from the healthy controls using higher order n-grams.
+An ordinary DNNLM uses lower order n-gram N-dimensional sparse vectors as descrete feature representations to train the neural network with multiple hidden layers.
+
+\subsection{What type of data is used by the studies?}
+One of the key debates when looking at how to analyse language is the type of task provided to elicit language production in participants.
+In the literature researchers have primarily focused on Picture Description tasks but have also suggested other ways in which we might collect data.
+\par
+\subsection{Picture Description Tasks}
+One of the most commonly used tasks to measure language is the Picture Description task.
+An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Borod1980}.
+The Cookie Theft picture (pictured below) depicts a scene of a home typical of the period of time when it was created and would generally not require participants to use any complicated vocabulary to describe.
+In this task participants are asked to describe the picture presented to them in as much detail as possible.
+This task was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}
+\begin{figure}[H]
+\centering
+\includegraphics[width=240px, height=150px]{images/BCTPicture.png}
+\caption{Cookie Theft Picture - From Kaplan and Goodglass (1983)}
+\end{figure}
+\par
+The picture description task does a fine job of eliciting descriptive language however because of the specific content the language produce could be considered quite limited.
+There is some disagreement as to the benefits of this using this methodology.
+This task is reported as being useful to lexico-semantic disorders \cite{Boschi2017, Sajjadi2012} as the language being generated is primarly nouns and deixis (words to identify items and words to put those items into context).
+However, Ash \cite{Ash2012}felt that there was no difference in using this task vs Story Narration (described below).
+In explaining the differences, it is worth noting that these researchers were using differing variables and this could explain their different perspectives.
+\par 
+There are a number of existing corpus which use this tasks as the foundation of their data.
+The most well known is the Pitt Corpus of the Dementia Bank database.
+The participants of the corpus are mainly healthy controls, those with MCI and those with probable and/or possible AD.
+\subsection{Narrative description task}
+The story narration task is designed to study a participant's ability to describe and elaborate on a story which is depicted using a series of pictures.
+The stories depicted are usually based on children's books or famous stories with the Cinderella being the one most typically used \cite{Fraser2014}.
+This task requires ordering the story in a structured and coherent framework.
+It also requires comprehension and understanding of the stories characters and the events depicted, as well as an awareness of a character's actions, motivations and internal reactions to given events.
+This task is particularly useful as the procedure reduces the demands on memory, due to the participant being able to access the picture book during the description and is therefore able rule out memory as a confounding variable for any results observed.
+As noted above, Ash \cite{Ash2012} felt that this task was interchangable with the Picture Description task.
+However, other research felt that this was a studier test of lexical and semantic abilities as well as syntactic complexity because this task requires interpretation and elaboration in additional to a simple description \cite{DeLira2011}.
+\par
+Given the relative strengths of the Narrative description task vs Picture description task, there are few pieces of research that have used Machine Learning to analyse features from Narrative picture tasks \cite{Fraser2014}.
+This could be due to the availability of data and the absence of any meaningful sets of transcripts of participants performing this task.
+However, this could be an interesting direction to take research in the future to see if features generated from this task could be used to predict MCI or AD.
+\par
+\subsection{Interviews}
+Interviews can also be used to elicit language in a more natural way by asking questions to guide a conversation between speakers.
+There are three types of interviews: unstructured, structured and semi-structured.
+Structured interviews tend to produce very limited speech and therefore has never been used in this area \cite{Boschi2017}.
+Unstructured interviews are open ended and generally do not conform to any particular pattern.
+They use generic themes such as family or hobbies to guide the conversation.
+Whilst this is the most ecologically valid form of conversation and therefore language generation, it's unstructured nature means that the protocol cannot be consistent and therefore reproduced.
+Semi-structured interviews are therefore preferred over other forms of interview as a middle ground.
+The semi structured nature of these interviews means that there is some replicability but does not constrain the participant in answering questions.
+\par
+The analysis of interviews can be difficult to analyse as both the content can vary even between participants, although it can be argued that content should not affect the type of language being generated unless it is narrow topic or the participant is constrained in how they answer a given question.
+It is also difficult to measure as there are no pre-defined task goals in comparison to the other two methods.
+Nevertheless, this is the most naturalistic setting for looking at language production and can be used to look at the syntactic and semantic parts of language generation \cite{Sajjadi2012}.
+There have been some attempts to use interviews to assess language production in AD with promising results \cite{Asgari2017, Guinn2015}.
+\par
+\subsection{Conclusions}
+One can view the different types of tasks above as a continuum where picture tasks represent a much more controllable task with a lot of supporting research but which generates a much more constrained set of language that is atypical of normal speech in terms of the cognitive functions used.
+
+\subsection{What are the goals of the studies that employ ML or Statistical Learning techniques for diagnosis of MCI or AD?}
+
+\subsection{Do the studies focus on a one point in time or looking at cognitive deterioration over time?}
+
+\section{Discussion and conclusions}\label{discussion}
+\subsection{Discussion of the current evidence}
+One of the criticisms of traditional learning models is it's reliance on features that are generated for them.
+As we can see from the brief look at some of the research in this field, researchers have taken a number of different approaches with success.
+However, there does not seem to be a consensus.
+As machine learning in general moves away from traditional machine learning models to deep learning techniques, it poses the question can deep learning assist with this problem?
+\par 
+One of the main benefits of deep learning is that it does not rely on pre defined features being fed to the model.
+Instead, deep learning models take raw data in with some amount of preprocessing and generates it's own features.
+This move away from relying on features solves one of the difficulties we have with the current psychological literature, namely that there are some disagreement on the which
+\par 
+One of the drawbacks of attempting to use natural language processing and machine learning in this context is the lack of data.
+This is the case with all machine learning techniques, but more so with deep learning.
+There have been ways to 'create' more data such as data augmentation.
+
+Current state-of-the-art diagnostic measures of AD are invasive (CSF analysis), expensive (neuroimaging), and time-consuming (neuropsychological assessment).
+Furthermore, these measures are limited to speciality clinics and thus have limited accessibility as frontline screening and diagnostic tools for AD.
+More importantly, nonspecialists are often inaccurate at identifying early AD and MCI.
+Thus, there is an increasing need for additional noninvasive and/or cost-effective tools, allowing effective frontline identification of subjects in the preclinical or early clinical stages of AD who could be suitable for monitoring in speciality clinics and for early treatment.
+Implementation of effective screening instruments will allow diagnosis earlier in the course of dementia, even at the point when memory function is still essentially within the normal range.
+This strategy would enable an earlier, and potentially more effective, prevention and treatment of AD with a special focus to preserve cognitive functions.
+\par 
+However the literature has identified a number of challenges when approaching this problem.
+Firstly, the clinical features that combine to meet the diagnostic criteria for Dementia or it's variants are continuous in nature and heterogenous between patients and are also impacted by other variables.
+For example as shown above, cognitive performance is affected in part by a patients educational attainment and a patients ability to live independently is impacted by a their physical health as much as their cognitive health.
+The challenge therefore is to find features that are minimally impacted by other factors, or that can be controlled for by a experimental design such as a matched pairs design to control for educational attainment.
+\par 
+Another challenge lies with the recruitment of suitable individuals who may notice a decline in cognition to the point where we might classify them as having MCI, but these individuals deduce that there is little to no value to admitting there is a problem and seeking help whilst their symptoms are 'manageable'.
+\par
+Finally there is a large amount of variability in the presentations of those with MCI and early dementia, and this is compounded by an similar amount of variability in the criteria researchers have used for experimental groups and the approaches researchers have used to tackle this problem.
+This had led to a confused literature.
+Recently a call for research that has consistent inclusion / exclusion criteria has been made along with some proposed definitions of MCI and it's subgroups \cite{Petersen2014}.
+Researchers have identified the analysis of language impairment as an area of promise to explore in the diagnosis of MCI and early AD and recent developments in natural language processing and machine learning techniques have the potential to assist in this research.
+Indeed, given the increased burden on the diagnosis of MCI and AD on professionals there has been a call to use technology to potentially ease this burden \cite{Boschi2017}.
+A small but growing amount of research has gone into the use of machine learning techniques to potentially look at the automated classification of participants with MCI and/or AD, however this is a new area of research and there are some gaps in our knowledge.
+
+\subsection{Methodological Issues}
+\subsection{Limitations}
+
+\subsection{The future of the field}
+One of the main patterns
+
+One-class classification is a traditional machine learning model in which the object is to recognize instances of a concept by only using examples of the same concept.
+In the training of this model, instances of only a single object class are available during training.
+In this context,all other classes (any instances that deviate from the single object class) are referred to as 'alien' classes.
+During the testing of the model, it may encounter alien classes and the job of the model is to distinguish between objects of the known class from objects belonging to any 'alien' classes.
+This approach to classification has broad applications and is mainly used when there is a high cost of obtaining samples of alien classes.
+\par 
+In this context, we can see from
+
+One of the areas for research to study is a careful examination of the features that are being used to measure language deterioration.
+For example, Zimmerer (2016) \cite{Zimmerer2016} describes connectivity in such as a way that correlates directly with what Mueller (2018) \cite{Mueller2018a} calls Fluency.
+Whilst these are very nuanced measures which differ slightly in the form they take, the sheer range of measures and features being produced make it difficult to organise and explore what is truly going on in those with MCI and early AD..
+Some work needs to be done in producing a consistent list of measures that are validated using existing datasets and can be used for future research moving forward.
+\newline
+\par
+Teng et al suggests that work should focus on the MCI population and concentrate on developing a consensus neuropsychological battery that could yield predictable rates of progression to AD \cite{Teng2009}.
+This, in conjunction with the development of a model of language, sensitive enough to detect subtle deterioration in language use to act as an additional cognitive marker to aid diagnosis could potentially move some way to providing this.
+\par 
+Future research should also be directed towards developing non-intrusive ways of detecting subtle changes in natural language such that any perceived deterioration that could indicate the presence of MCI or AD could be flagged up early.
+Machine learning approaches seem to be the most logical approach for achieving this aim as language could be collected in non-intrusive ways and passed to a machine learning algorithm for preliminary classification.
+Despite the excellent quality of datasets, for example the DementiaBank dataset, being used to 'backtest' these algorithms, further research should look at generating additional datasets to increase the validity of the results found so far as well as using other methods to generate data other than Picture Description tasks which some researchers could claim are limited in scope.
+Finally, the recent resurgence in the use of neural networks and deep learning could provide the answer to the confused literature in terms of features.
+A key benefit of deep learning is it's ability to automate the process of feature engineering.
+So there is an opportunity to explore the use of deep learning, to not only develop new features but also validate existing features independently.
+\par
+This area of research is extremely promising in its early results and the impact of successful research would be life changing for both individuals and the health of the worlds aging population in general.
+
+\section{Conclusions}\label{conclusions}
+We can see that both Natural Language Processing and Machine Learning techniques have a lot to offer, Indeed there has been a lot of research which have used pre-existing datasets to explore this area with promising results.
+One of the difficulties with the current research is the approach of trying to discriminate between those AD and healthy controls.
+This is not necessarily a problem in the real world as it is trivially easy to do for trained clinicians.
+A more interesting, but potentially harder to problem to solve is to discriminate between those with MCI and healthy controls, and more importantly to track their decline over time.
+Further, to date the authors are unaware of any research in which these techniques are applied to newly created samples of language.
 
 \chapter{Delphi Methodology and developing consensus on how best to collect language samples using technology}
-\include{Chapter3}
+\section{Introduction}
+Here is the text of your introduction.
+
+\begin{equation}
+    \label{simple_equation}
+    \alpha = \sqrt{ \beta }
+\end{equation}
+
+\subsection{Subsection Heading Here}
+Write your subsection text here.
+
+\section{Conclusion}
+Write your conclusion here.
 
 \chapter{Development of a pipeline that processes language data accurately}
-\include{Chapter4}
+\section{Background}
+As Komeili in her paper points out, machine learning is a tool may allow earlier detection and management of change in language \cite{Komeili2019}.
+But as with all machine learning models, the benefits of accuracy only come when large amounts of data is used to train the model.
+With our target population, this problem is further exacerbated due to the relatively low incidence of people with MCI or early AD and the relatively high cost of collecting this data in this particular population.
+\par 
+One potential way forward is to look not at the ill but at the healthy.
+This has inverts the problem of scarcity and cost of data collection, and so we now have an abundance of data that we can now use.
+This, in theory, would allow us to look specifically at any deviations from the healthy.
+This process is known as one-class classification.
+\par
 
-\chapter{Analysis of DementiaBank dataset}
-\include{Chapter5}
+\section{One-class classification}
+In traditional machine learning, there are numerous classification algorithms that are widely used.
+Their utility is classifying instances of data into one (Binary classification) or more (multiclass classification) categories.
+This ia a fine way to look at classification in the vast majority of cases, but some difficulties do arise under certain circumstances.
+
+\chapter{Analysis of the Presidents Corpus, Three Authors and DementiaBank datasets}
+\section{Background}
+There has been a significant research in the area of language deterioration as a means of detecting Alzheimer's Disease.
+This usually takes the form analysis of speech recorded as part of a cognitive assessment such as the Picture Description Task \cite{Orimaye2014,Fraser2015}.
+Given that language samples are relatively easy to collect, research has moved towards analysis of spontaneous speech.
+An good example of this type of research is the study conducted by Berisha and Liss which looked at the differences in language use between two US presidents, Ronald Reagan (who would go on to receive a diagnosis of Dementia) and George H.W.
+Bush who acted as a matched control based on Age \cite{Berisha2015}.
+They found several differences in language use which they felt acted as indicators of Reagan's difficulties with language due to dementia.
+These significant differences were in the number of unique words used per speech, the use of non-specific nouns and fillers and low-imageability verbs \cite{Berisha2015}.
+\par   
+This study replicates work done by Berisha and Liss and extends this by adding Donald Trump as an alternative, more appropriate comparison to Ronald Reagan as he is much closer in age than George H.W.
+Bush.
+This experiment will look at the features originally identified by Berisha and Liss, as well as any others that have potential as discussed in the literature review above.
+\section{Methods}
+I took 46 transcripts of Ronald Reagan’s (RR) press conferences from 1981 to 1988 and compared them with 134 press conferences by George H.
+W.
+Bush (GHWB) and 29 press conferences conducted by Donald J.
+Trump (DJT).
+I analyzed transcripts for lexical features shown to change longitudinally with dementia  (for a comprehensive review of these, see the literature review above).
+For this collection of documents, I generated a number of features which looked at a number of different aspects of each document.
+These features encompassed, word level, sentence level and document level features and included a number of features contained in the study by Berisha and Liss with the aim of replicating and extending on their findings.
+These findings were.
+number of unique words, non-specific nouns and fillers and low imageability (LI) verbs.
+Imageability is characterized, according to Berisha and Liss, as the ease with which a term gives rise to a sensory .mental image.
+I compared the trends described in the transcripts of RR and GHWB, but also included DJT.
+Berisha and Liss originally made the comparison as it GHWB (GHWB - age at the start of presidency - 64 years, 222 days) was the closest match to RR in terms of age (RR  - age at the start of presidency, 69 yeaars and 349 days).
+However, with the inauguration of Trump, he now is the closest comparible president in terms of age (DJT - age at the start of presidency - 70 years, 220 days).
+It would be interesting to look at a comparison of RR and DJT to see whether the comparisons made by Berisha and Liss hold true with this more appropriate match (in terms of age).
+DJT as with GHWB has no known diagnosis of AD.
+I used the press conference transcripts in the American Presidency Project (APP) archive as a data source for this project.
+The APP is a comprehensive and organized searchable database of presidential documents, including transcripts of speeches, transcripts of news conferences, and other public documents.
+\subsection{Pre-processing}
+To generate the files necessary for analysis, I downloaded each transcript and performed the following changes.
+I omitted the prepared statement by the president and any speech by other individuals.
+I started each transcript at the beginning of the first answer to a question.
+I filtered any annotations that were added to the transcript, including any references or clarifications, and any laughter.
+It's worth noting that there appears to be a difference in how 'hesitations' were marked down between each president, for RR hesitations were marked by a single hyphen whereas for GHWB hesitations are marked by a double hyphen.
+In order to maintain consistency when parsing through the documents, I have changed both types of hesitation to be marked by a single hyphen.
+I also omitted one word sentences as this data would, from a theoretical perspective, not be relevant for language analysis.
+I did not control for the length of the document, but generated features which would normalise by the length of the document.
+I therefore was able to include all press conferences by both Ronald Reagan and George H.W.
+Bush where there was a question and answer session conducted at least in part by the sitting president (2 press conferences of GHWB were ommited due to a lack of a question and answer session).
+\subsection{Feature Selection}
+We calculated the following features for each transcript in turn using the NLTK (see section 2 for a description) \cite{Bird2009} and Python. 
+\par 
+\subsubsection{Measures of lexical variation}
+We constructed two features of lexical variation.
+Firstly we looked at the number of unique words.
+To do this we were able to split each transcript into individual words and changed them to lowercase using NLTK and were then count the number of unique words that appeared in each transcript.
+We also used the TTR formula (see section 2 for a description) for a feature that measures lexical diversity independent of sample size \cite{Richards1987}.
+\subsubsection{Fillers, Non-Specific Nouns and LI Verbs}
+For these features, we counted the number of occurrences for different categories (see table for list of categories tracked and the words counted).
+The features were used by Berisha and Liss in their research \cite{Berisha2015} and were taken from work done by Bird et al \cite{Bird2000}.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{6cm} |}
+		\hline
+		Category & Words \\ \hline
+		Fillers & \textit{"well", "so", "basically", "actually", "literally", "um", "ah"} \\ \hline
+		Non Specific Nouns & \textit{"something", "anything", "thing", "everything"} \\ \hline
+		LI Verbs & \textit{"be", "come", "do", "get", "give", go", "know", "look", "make", "see", "tell", "think", "want"} \\ \hline
+
+	\end{tabular}
+	\caption{\label{tab:table-name}Categories and Words Counted}
+	\end{center} 
+\end{table}
+
+\subsubsection{Usage of parts of speech}
+This section involves using a Part of Speech tagger (PoS) which analyses a sentence and assigns a 'tag' to each word based on the function the word has in a sentence.
+At a basic level this can be divided into the eight defined parts of speech: 'nouns', 'pronouns', verbs', 'adjectives', 'adverbs', 'conjunctions', 'prepositions' and 'interjections' but can be further subcategorised.
+We used the PoS tagger built into NLTK to tag each transcript in turn and used these the counts from each of these eight categories in our analysis.
+In addition to frequency counts we also normalised these features by dividing the frequency count by the number of words in the document to take into account transcript length.
+
+\section{Results}
+One of the most important thing to note is the wide variety of samples between the three presidents and also the varying timescales.
+RR participated in 46 press conferences over eight years (an average of 5.75 a year) which is the fewest number of press conferences given by an American president during their term of office.
+GHWB participated in 136 press conferences over four years (an average of  34 a year) and DJT participated in 29 press conferences to date (an average of 19.3 per year).
+Equally, there are variances in the average number of words.
+RR produced an average of 3424 words per conference compared to 2608 by GHWB (unpaired t = 4.434, p\textless 0.001) and DJT at 1849 words (unpaired t = 6.524, p\textless 0.001).
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR & GHWB & DJT \\ \hline
+		Total Words & 3423.91 (416.42) & 2607.72 (1210.38) & 1848.65 (1549.38) \\ \hline
+		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
+		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
+		LI Verbs & 124.22 (17.89)& 103.45 (51.87) & 84.48 (75.78) \\ \hline
+	\end{tabular}
+	\caption{\label{tab:table-name}Means and Standard Deviations of important features}
+	\end{center} 
+\end{table}
+
+In terms of unique words, we found that RR used significantly more unique words, non-specific nouns and low imageability verbs than GHWB and DJT (see Table 3.3).
+Some of these differences are due to the length of the sample, particularly in the case of DJT where his average sample is almost half the sample of RR.
+It could also be said that this could be down to differences in linguistic abilities or speaking style \cite{Berisha2015, Le2011}.
+However, we can certainly see that as controls GWHB and DJT are comparative in relation to non-specific nouns and LI verbs.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR v GHWB & RR v DJT & GHWB v DJT \\ \hline
+		Total Words & \textbf{4.434***} & \textbf{6.524***} & \textbf{2.899**} \\ \hline
+		Unique Words & \textbf{6.832***} & \textbf{11.403***} & \textbf{4.137***} \\ \hline
+		Non Specific Nouns & \textbf{7.877***} & \textbf{3.426**} & -0.574 \\ \hline
+		LI Verbs & \textbf{2.656**} & \textbf{3.420***} & 1.628 \\ \hline
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+	\end{tabular}
+	\caption{\label{tab:table-name}RR T-tests vs GWB and DJT}
+	\end{center} 
+\end{table}
+
+We then looked at the data from a longitudinal perspective as we are interested seeing whether we can track various language variables and their progress over time.
+We ran a number of Pearsons correlations with transcript index number as a time reference and the dependant variables (Table 3.4).
+For our controls, we found them to be stable for the most part with the main highlights being a decrease in Adverb usage for DJT (R=-0.36, p=0.049) and a steady but not severe decline in a number of variables for GWHB, namely total word count, unique words, low imageability words and verb usage.
+\par 
+For RR, his decline is more marked and more widespread through his language use.
+We noticed an significant increase in adverb (R=0.41, p=0.004) and pronoun usage (R=0.65, p\textless0.001), as well as a slight usage increase in Non-specific nouns(R=0.30, p=0.03).
+There was a highly significant decrease in number of unique words (R=-0.56, p\textless0.001) and noun usage (R=-0.70, p\textless0.001).
+Also very significant decrease in adjective usage (R=-0.40, p=0.005) and a significant decrease in total word count (R=-0.31, p=0.03).
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | l | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR & GHWB & DJT \\ \hline
+		Word Count & \textbf{-0.31*} & \textbf{-0.21*} & 0.08 \\ \hline 
+		Unique Words & \textbf{-0.56***} & \textbf{-0.25**} & 0.16 \\ \hline
+		Non Specific Nouns & \textbf{0.30*} & -0.08 & -0.03 \\ \hline
+		LI Verbs & -0.19 & \textbf{-0.20**} & 0.02 \\ \hline
+		Nouns Normalised & \textbf{-0.70***} & -0.03 & 0.14 \\ \hline
+		Verbs Normalised & \textbf{0.36**} & \textbf{0.24***} & -0.03 \\ \hline
+		Adjectives Normalised & \textbf{-0.40**} & 0.08 & -0.34 \\ \hline
+		Adverbs Normalised & \textbf{0.41***} & 0.02 & \textbf{-0.36*} \\ \hline
+		Pronouns Normalised & \textbf{0.65***} & 0.13 & 0.07 \\ \hline
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+	\end{tabular}
+	\caption{\label{tab:table-name}Pearson Correlations for Features}
+	\end{center} 
+\end{table}
+
+\section{Discussion}
+President Reagan received his diagnosis of AD in August 1994 but using transcripts of speeches he made in his two terms as President (January 1981 - January 1989) we have be able to identify certain changes in his use of language that we might ascribe to the onset of MCI and early AD.
+Despite differences in our methodology, our research supports the findings of Berisha and Liss in that we both find a significant decrease in unique words over time and an increase in non-specifc noun usage.
+Compared to our controls (GWHB andDJT), we find some slight trends with GWHB but no such trends with DJT in his speech albeit his samples of speech span a shorter amount of time.
+\begin{figure}[H]
+	\centering
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/RRUniqueWords.png}
+		\caption{Ronald Reagan - Unique Words over time}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/GWHBUniqueWords.png}
+		\caption{George H.W. Bush - Unique Words over time}
+	\end{minipage}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/RRNSNouns.png}
+		\caption{Ronald Reagan - Non-specifc Nouns over time}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[b]{0.4\textwidth}
+		\includegraphics[width=190px, height=140px]{images/DJTNSNouns.png}
+		\caption{Donald J. Trump - Non-specific Nouns over time}
+	\end{minipage}
+\end{figure}
+
+A criticism of Berisha and Liss's work is the problems they had with normalising the transcripts in terms of length.
+This was also a problem in the work of Garrard et al \cite{Garrard2005, Le2011}.
+Whilst it is important to control for outliers, there are other ways in which we can control for length of sample.
+\par 
+Interestingly, when we normalised the various types of words used by the presidents we found some interesting patterns that further differentiated RR from the controls.
+Whilst Non-specific nouns increased over time, we found that noun usage in general significantly decreased and pronouns increased similarly significantly.
+The increase in pronoun for those with early AD has been identified in literature, although there are only a few studies that explore this \cite{Almor1999, Wendelstein2015}.
+Wendlestein et al propose that the increased used of pronouns is an expression of an impaired ability to adapt language to the listener's needs \cite{Wendelstein2015}.
+Almor et al attributed this reliance on pronouns due to a impaired working memory \cite{Almor1999}.
+\par 
+The decrease in overall noun usage has also been identified as a feature.
+Jarrold et al found that AD patients would use more pronouns, verbs and fewer nouns than controls \cite{Jarrold2014}.
+Wendlestein in their investigations into noun usage found that decreased later on in AD progression and was unaffected in the pre-clinical stages of AD \cite{Wendelstein2014}.
+Our results are supported by existing literature and this potentially means that language analysis in the way we have structured it may have diagnostic or prognostic properties.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=240px, height=150px]{images/RRNounsNormalised.png}
+	\caption{Ronald Reagan - Nouns Normalised over time}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=240px, height=150px]{images/RRPronouns.png}
+	\caption{Ronald Reagan - Pronouns Normalised over time}
+\end{figure}
+
+There are limitations of this research.
+Whilst in terms of age, DJT is certainly more suitable as a control to match with RR, in some ways they held very different styles of press conferences in that RR preferred to do solo press conferences and DJT has shown a preference for doing joint press conferences which have an impact on the amount of language produced.
+This artifact of the data is in itself notable as it illustrates the problems we may have with smaller amounts of speech.
+Also, the problem of finding an appropriate control is a common one in this domain.
+Given that those with MCI and early dementia have such variable presentations, it might prove of limited value in matched pairs design.
+\par 
+With further work, it is not feasible to the vast array of samples over a timeframe, as we have had with the president corpus and so it would be worth exploring how the quality of these predictions might lessen when faced with considerably fewer samples and over a smaller time period.
+It would also be worth extending this research further to encompass more of the linguistic features Fraser used in her work \cite{Fraser2015} to see if there are any further insights to be gained.
+In addition, this replication and extension has demonstrated the potential utility of using longitudinal data as a means of comparing language use of a person at two or more time periods and using this information as a diagnostic aid for MCI and therefore more work would be helpful from a longitudinal perspective to see if this approach may be valid in moving towards a solution for this particular problem.
+\par 
+The results of this work show that we can track a person's use of language through time in a number of ways, and it is possible for an individual to be his or her own control.
+This is important as it means the heterogenous nature of the MCI population does not impact results as much as if we were comparing those with MCI to controls.
+Equally, it would be helpful to have controls to ascertain what would be usual to expect in the decline of language in a healthy older adult.
+
+\section{Introduction}\label{introduction}
+A diagnosis of dementia is generally made when there is a decline in brain function due to physical changes in the brain \cite{Albert2011}.
+It affects a significant proportion of the global older adult population and the impact on morbidity and mortality rates is considerable.
+Dementia is currently the leading cause of death in England and Wales and the sixth leading cause of death in the United States (US).
+A 2014 report commissioned by the Alzheimer's Society estimated that in the UK by 2015 there would be approximately 855,000 people rising living with dementia increasing to 1 million by 2021 \cite{AlzheimersSociety2014}.
+This represents 1 in 79 of the total UK population rising to 1 in 14 of those aged 65 or over \cite{AlzheimersSociety2014}.
+Worldwide, there are 46 million people with a diagnosis of dementia globally and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}.
+From a financial perspective, the cost burden is also significant.
+The estimated annual spend on dementia healthcare in the UK is £4.3 billion of which approximately £85 million is spent on diagnosis.
+The total financial burden of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually.
+Globally, this picture is a lot bleaker.
+The worldwide cost of dementia in 2018 was estimated to be in the region of one trillon US dollars \cite{Prince2015}.
+\par 
+There are different types of dementia including Alzheimer's Disease (AD), vascular dementia, dementia with Lewy bodies and fronto-temporal dementia, all of which are currently incurable.
+AD is a progressive neurodegenerative disease and is the most common type of dementia, responsible for approximately 60\% to 80\% of all cases \cite{Duong2017}.
+Currently, a definitive diagnosis for AD can only be produced at post-mortem.
+However, there are a number of psychological and physiological indicators that can indicate that dementia is present.
+From a physiological perspective, researchers have identified two proteins called beta-amyloid and tau have been identified as two of the key proteins that have a role in initiation and progression of AD.
+In a typical case beta-amyloid clumps into plaques which slowly build up between neurons with tau proteins accumulating and eventually forming tangles inside neurons.
+At a certain point, the levels of beta-amyloid rise triggering a more rapid spread of tau throughout the brain.
+Eventually, due to these and other changes, neurons lose their ability to communicate and the brain starts to shrink\cite{Duong2017}.
+This leads to the development of some of the psychological symptoms associated with AD, primarily cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}.
+These cognitive symptoms are often accompanied by emotional problems such as depression and behavioural difficulties.
+As more neurons die throughout the brain, a person with AD gradually loses the ability to think, remember, make decisions and function independently.
+\par
+Despite the increasing prevalence of AD and an improved understanding about how it affects the brain there are no medications that improve prognosis.
+All the medications that are currently on the market are designed to manage symptoms.
+Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage of these drugs fail in clinical trial stage of the drug discovery process ((99.6\% failure rate vs 80\% for systemic anti-cancer drugs) \cite{Cummings2014}.
+Cummings et al proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already occurred \cite{Cummings2014}.
+Research focus has now started to shift to the earlier stages of AD (i.e.
+symptomatic pre-dementia phase of AD) which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.
+\par
+One of the challenges associated with the early detection of AD is differentiating natural age associated memory impairment and cognitive decline due to aging from with decline due to AD.
+This challenge is often complicated further due to the large variation in the cognitive abilities and educational background of individuals.
+The work of Albert et al helps to address this by the development of clinical criteria which professionals can use to diagnose MCI due to AD.
+One of the most important observations from this piece of work is that a diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}, and this is essentially to place more importance on the trajectory of a person's cognitive abilities rather than a person's cognitive ability in general.
+\par 
+There has been a significant research in the area of language deterioration as a method of detecting AD at an earlier stage.
+This usually takes the form of recording speech whilst patients' undertake a cognitive assessment such as the Picture Description Task \cite{Fraser2015}.
+Given that language samples are relatively easy to collect, research has moved towards analysis of spontaneous speech.
+The work of Berisha and Liss is a good example of this, this study compared the differences in language use between two US presidents, Ronald Reagan (RR), who would go on to receive a diagnosis of Dementia and George H.
+W.
+Bush (GHWB) who acted as a matched control based on Age \cite{Berisha2015}.
+They found several differences in language use over time which they felt acted as indicators of RR's difficulties with language due to AD.
+Differences in features of language identified to be statistically significant included the number of unique words used per speech, the use of non-specific nouns and fillers and low imageability verbs \cite{Berisha2015}.
+\par   
+Berisha and Liss have developed some very interesting ideas about how we might track changes over time in various lexical features which have been associated with the development of Alzheimer's Disease.
+However we feel that cognitive decline in pre-clinical AD is not accurately modeled as a linear process and therefore we explored the application of generalised additive models (GAMs) to this data which have no assumed understanding of the distribution of the data.
+We refine these ideas with the aim of exploring the potential for a protocol that can be used for analyzing language deterioration.
+
+This study extends the work of Berisha and Liss in a number of ways. Our hypotheses are:
+\begin{enumerate}
+	\item Language use of RR will be significantly different to our controls (GHWB, DJT).
+	\item Modelling the dynamics of lexical features over time can be improved using non-linear models compared to linear models.
+\end{enumerate}
+
+\section{Methodology}\label{methodology}
+We took 46 transcripts of press conferences given by RR (from 1981 to 1988) and compared them with 134 press conferences (from 1989 to 1993) given by GHWB and 29 press conferences (from 2016 to 2019) conducted by DJT.
+We analysed transcripts for language features (described below) shown to change longitudinally with AD.
+These language features are analysed at the word level, sentence level and document level.
+\par 
+In the original study, GHWB was selected as the comparator president as he was the closest match in terms of age to RR (GHWB - age at the start of presidency: 64 years and 222 days vs RR: 69 years and 349 days).
+However, since his inauguration, DJT is now the closest comparible president in terms of age (DJT - age at the start of presidency: 70 years and 220 days).
+We included DJT who like GHWB has no known diagnosis of AD to determine whether the comparisons made by Berisha and Liss hold true with this closer presidential match in terms of age.
+\par 
+We used the press conference transcripts in the American Presidency Project (APP) archive as a data source for this project.
+The APP is a comprehensive and organized searchable database of presidential documents, including transcripts of speeches, transcripts of news conferences, and other public documents.
+These documents are open access and can be downloaded at any time from the APP archive.
+
+\subsection{Pre-processing}
+To generate the files necessary for analysis, we downloaded each transcript and performed the following changes.
+We omitted the prepared statement by the president and any speech by other individuals and started each transcript at the beginning of the first answer to a question by a member of the press.
+We filtered any annotations that were added to the transcript, including any references or clarifications, and any laughter.
+It's worth noting that there appears to be a difference in how 'hesitations' were marked down between each president, for RR \& DJT hesitations were marked by a single hyphen whereas for GHWB hesitations are marked by a double hyphen.
+In order to maintain consistency when parsing through the documents, We have changed both types of hesitation to be marked by a single hyphen.
+We also omitted one word answers to questions as this data would, from a theoretical perspective, not be relevant for language analysis.
+We did not make any alterations to the length of the document, but instead generated features which would normalise by the length of the document.
+We therefore was able to include all press conferences by all presidents analysed where there was a question and answer session conducted at least in part by the president in question (2 press conferences of GHWB were omitted due to a lack of a question and answer session).
+
+\subsection{Feature Generation}
+Features were generated by running each transcript through a number of natural language processing libraries.
+We used the Natural Language ToolKit (NLTK) package \cite{Bird2009} in python to stem each transcript using the Snowball Stemmer and also completed a part of speech (POS) tagging process.
+We also ran each transcript through Linguistic Inquiry and Word Count (LIWC) \cite{Pennebaker2015} software.
+We completed all NLP tasks in Python and completed all statistical analysis in R..
+\par 
+\subsubsection{Measures of lexical diversity}
+Lexical diversity is a measure of the variety of language used within a given document.
+A document is said to have high lexical diversity if the number of unique words is large.
+We constructed four features of lexical variation.
+Firstly we looked at the number of unique words.
+To do this we were able to split each transcript into individual words and changed them to lowercase using NLTK and were then count the number of unique words that appeared in each transcript.
+We also used the TTR formula, Brunet's Index and Honore's Statistic as other measures of lexical diversity \cite{Bucks2000}.
+\par 
+Type token ratio (TTR) is the ratio obtained by dividing the types (The total number of different words) by the tokens (the total number of words in an utterance).
+\begin{equation} \label{x1}
+TTR = numberOfUniqueWords / totalNumberOfWords.
+\end{equation}
+\par 
+Brunet's Index (W) differentiates itself from TTR, as it is not impacted by the length of the text itself. Brunet's Index is defined by the following equation:
+\begin{equation} \label{x2}
+W = N^{V(-0.165)}
+\end{equation}
+where N is the total length of the utterance being measured and V is equal to the total vocabulary being used by the subject.
+Brunet's Index usually has a score of between 10 and 20, with high numbers indicating a more rich vocabulary compared to low numbers.
+\newline
+\par 
+Honore's Statistic is based on the idea that vocabulary richness is implied when a speaker uses a greater amount of unique words. This is indicated by the following equation: 
+\begin{equation} \label{x3}
+R = (100 \log N) / (1 - V1/V)
+\end{equation}
+where v1 is equal to the number of unique words, V is the total vocabulary used and N is the total number of words in the utterance being measured.
+\subsubsection{Fillers, Non-Specific Nouns and Low Imageability Verbs}
+Fillers, Non-Specific Nouns and Low Imageability Verbs were features used by Berisha and Liss in their research \cite{Berisha2015} and were taken from work done by Bird et al \cite{Bird2000}.
+Fillers can be described as a potentially meaningless word that marks a pause or hesitation in speech.
+In those with MCI and AD, these words can be used to temporarily disguise problems in thought processes or word finding difficulties.
+Non-Specific Nouns refer to a category or an unspecified member of a given category, once again this can be characterised as a compensatory strategy for word finding difficulties.
+Imageability is characterized, according to Berisha and Liss\cite{Berisha2015}, as the ease with which a word provokes a mental image of what the word describes.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{10cm} |}
+		\hline
+		Category & Words \\ \hline
+		Fillers & \textit{"um", "uh", "er", "ah", "like", "okay", "right",  "you know", "well", "so", "basically", "actually", "literally"} \\ \hline
+		Non Specific Nouns & \textit{"something", "anything", "thing", "everything"} \\ \hline
+		LI Verbs & \textit{"be", "come", "do", "get", "give", go", "know", "look", "make", "see", "tell", "think", "want"} \\ \hline
+	\end{tabular}
+	\caption{\label{tab:table-name}Examples of words belonging to the categories Fillers, Non-Specific Nouns and Low Imageability Verbs}
+	\end{center} 
+\end{table}
+
+\subsubsection{Usage of parts of speech}
+Using a Part of Speech tagger (PoS) on each transcript analyses each sentence within the transcript and assigns a 'tag' to each word based on the function the word has in a sentence.
+At a basic level this can be divided into the eight defined parts of speech: 'nouns', 'pronouns', verbs', 'adjectives', 'adverbs', 'conjunctions', 'prepositions' and 'interjections' but can be further subcategorised.
+We used the PoS tagger built into NLTK to tag each transcript in turn and used these the counts from each of these eight categories in our analysis.
+In addition to frequency counts we also normalised these features by dividing the frequency count by the number of words in the document to take into account transcript length.
+
+\subsubsection{Linguistic Inquiry and Word Count(LIWC)}
+The LIWC is a text processor which analysis words within a given a document and compares the words in the LIWC dictionary file.
+The LIWC dictionary file contains over 6000 words which are categorised into approximately 90 different types.
+For example the word 'cried' is part of five word categories: sadness, negative emotion, overall affect, verbs and past focus \cite{Pennebaker2015}.
+Each word in the document or set of documents is searched for in the dictionary file and if found, the appropriate type is incremented by 1.
+What is output is an analysis of word usage in reference to each category.
+This is particularly useful in analysing both structural language changes but also the content or themes of language for a given transcript or set of transcripts.
+
+\subsubsection{Longitudinal Analysis}
+As with Berisha and Liss\cite{Berisha2015}, we analysed the transcripts over time in order to determine if there was an underlying temporal trend that may indicate that different parts of language increase or decrease over time.
+In AD, cognitive abilities are said to deteriorate over time and therefore it may be useful to analyse temporal trends that indicate a potential deterioration of cognition.
+In order to get a accurate measure of deterioration over time, we time stamped each transcript in relation to number of days from the first transcript to the date of the transcript being analysed as the press conferences were not evenly spaced out during the presidencies.
+\par
+We looked also compared the results of linear models and a non linear model in terms of best fit.
+The non linear model we used was the generalised additive model \cite{Hastie1986}.
+The selection of this non linear model was due to the authors not wanting to assume an underlying distribution of the data and therefore this was the most appropriate model to use.
+
+\section{Results}\label{results}
+One of the most important thing to note is the wide variety of samples between the three presidents and also the varying timescales.
+RR participated in 46 press conferences over eight years (an average of 5.75 a year) which is the fewest number of press conferences given by an American president during their term of office.
+GHWB participated in 136 press conferences over four years (an average of 34 a year) and DJT participated in 29 press conferences to date (an average of 19.3 per year).
+Equally, there are differences in the average number of words.
+RR produced an average of 3424 words per conference compared to 2608 by GHWB (unpaired t = 4.434, p\textless 0.001) and DJT at 1849 words (unpaired t = 6.524, p\textless 0.001).
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR & GHWB & DJT \\ \hline
+		Total Words & 3423.91 (416.42) & 2607.72 (1210.38) & 1848.65 (1549.38) \\ \hline
+		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
+		Mean Length of Utterance & 23.17 (1.402) & 18.71 (2.067) & 13.84 (1.619) \\ \hline
+	\end{tabular}
+	\caption{\label{tab:table-name}Means and Standard Deviations of general features for each set of transcripts}
+	\end{center} 
+\end{table}
+
+In terms of more specific language differences between the presidents, we found that RR used significantly more unique words, non-specific nouns and low imageability verbs than GHWB and DJT (see Table 3).
+The mean length of utterance for RR was significantly greater than that of GHWB and DJT.
+Some of these differences are due to the length of the sample, particularly in the case of DJT where his average sample is almost half the sample of RR.
+It could also be said that this could be down to differences in linguistic abilities or speaking style \cite{Berisha2015, Le2011}.
+However, we can certainly see that as controls GWHB and DJT are comparative in relation to non-specific nouns and LI verbs.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | p{3cm} | p{2.2cm} | p{2.2cm} | p{2.2cm} |}
+		\hline
+		& RR v GHWB & RR v DJT & GHWB v DJT \\ \hline
+		Total Words & \textbf{4.434***} & \textbf{6.524***} & \textbf{2.899**} \\ \hline
+		Unique Stems & \textbf{10.878***} & \textbf{10.111***} & \textbf{4.148***} \\ \hline
+		Mean Length of Utterance & \textbf{16.175***} & \textbf{25.084***} & \textbf{13.484***} \\ \hline	
+		Non Specific Nouns & \textbf{7.877***} & \textbf{3.426**} & -0.574 \\ \hline
+		LI Verbs & \textbf{2.656**} & \textbf{3.420***} & 1.628 \\ \hline
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+	\end{tabular}
+	\caption{\label{tab:table-name}RR T-tests vs GWB and DJT}
+	\end{center} 
+\end{table}
+
+\subsection{Longitudinal Analysis}
+We then looked at the data from a longitudinal perspective as we were interested seeing whether we can track various language variables and their progress over time.
+We ran a number of Pearsons correlations with transcript index number as a time reference and the dependant variables (Table 4).
+For our controls, we found them to be stable for the most part with the main highlights being a decrease in Adverb usage for DJT (R=-0.36, p=0.049) and a steady but not severe decline in a number of variables for GWHB, namely total word count, unique words, low imageability words and verb usage.
+\par 
+For RR, his decline is more marked and more widespread through his language use.
+We noticed an significant increase in adverb (R=0.41, p=0.004) and pronoun usage (R=0.65, p\textless0.001), as well as a slight usage increase in Non-specific nouns(R=0.30, p=0.03).
+There was a highly significant decrease in number of unique words (R=-0.56, p\textless0.001) and noun usage (R=-0.70, p\textless0.001).
+Also very significant decrease in adjective usage (R=-0.40, p=0.005) and a significant decrease in total word count (R=-0.31, p=0.03).
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | l | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
+		\hline
+		& RR & GHWB & DJT \\ \hline
+		Word Count & \textbf{-0.31*} & \textbf{-0.21*} & 0.08 \\ \hline 
+		Unique Words & \textbf{-0.56***} & \textbf{-0.25**} & 0.16 \\ \hline
+		Non Specific Nouns & \textbf{0.30*} & -0.08 & -0.03 \\ \hline
+		LI Verbs & -0.19 & \textbf{-0.20**} & 0.02 \\ \hline
+		Nouns Normalised & \textbf{-0.70***} & -0.03 & 0.14 \\ \hline
+		Verbs Normalised & \textbf{0.36**} & \textbf{0.24***} & -0.03 \\ \hline
+		Adjectives Normalised & \textbf{-0.40**} & 0.08 & -0.34 \\ \hline
+		Adverbs Normalised & \textbf{0.41***} & 0.02 & \textbf{-0.36*} \\ \hline
+		Pronouns Normalised & \textbf{0.65***} & 0.13 & 0.07 \\ \hline
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+	\end{tabular}
+	\caption{\label{tab:table-name}Pearson Correlations for Features}
+	\end{center} 
+\end{table}
+
+To get a better idea of decline over time, we also ran Pearson's correlations using time elapsed in days from the first transcript instead a transcript index.
+This was to control for variation in the number of days between transcripts.
+Given the number of hypotheses being tested, we felt it was necessary to apply some correction for family wise error rate and therefore a Hochberg step-up procedure was applied to the results of the analysis.
+We also used the Benjamini-Yekutieli procedure for controlling for False Discovery rate and came up with a final list of important features for each president.
+For RR, this was 22 features, for GHWB this was 14 features and there were no significant features for DJT after applying these methods.
+The features were generated from multiple sources and therefore there was some significant overlap between the features but we have included all features in Table 6.
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | l | c | c | c |}
+		\hline
+		& RR R-Squared  & GWB R-Squared & DJT R-Squared \\ \hline
+		ppron & \textbf{0.700***} & 0.007 & -0.196 \\ \hline
+		social & \textbf{0.698***} & 0.204 & 0.032 \\ \hline
+		NounsNormalised & \textbf{-0.689***} & -0.023 & 0.194 \\ \hline
+		function & \textbf{0.670***} & -0.169 & -0.243 \\ \hline
+		conj &\textbf{0.644***} & -0.452 & -0.187 \\ \hline
+		PronounsNormalised & \textbf{0.631***} & 0.131 & 0.002 \\ \hline
+		Analytic & \textbf{-0.626***} & -0.013 & 0.366 \\ \hline
+		NN & \textbf{-0.601***} & -0.197 & 0.116 \\ \hline
+		male & \textbf{0.585***} & 0.024 & 0.042 \\ \hline
+		UniqueWords & \textbf{-0.578***} & -0.257 & 0.204 \\ \hline
+		WDT & \textbf{-0.577***} & -0.173 & -0.105 \\ \hline
+		shehe & \textbf{0.529*} & -0.005 & 0.064 \\ \hline
+		VBZ & \textbf{-0.521*} & -0.157 & 0.136 \\ \hline
+		JJ & \textbf{-0.518*} & -0.206 & 0.062 \\ \hline
+		Fillers & -0.076 & \textbf{-0.359**}  & -0.032 \\ \hline	
+		EX & -0.193* & \textbf{-0.337**}  & 0.339 \\ \hline	
+		achieve & -0.248 & \textbf{0.334**}  & 0.169 \\ \hline		
+		\multicolumn{4}{@{}p{1.5in}}{\footnotesize * denotes p\textless 0.05} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize ** denotes p\textless 0.01} \\
+    	\multicolumn{4}{@{}p{1.5in}}{\footnotesize *** denotes p\textless 0.001} \\
+	\end{tabular}
+	\caption{\label{tab:table-name}Pearson Correlations for Features}
+	\end{center} 
+\end{table}
+
+In terms of comparing the general additive model and linear models.
+We calculated the predicted residual error sum of squares (PRESS) statistic for each feature is a form of cross-validation used in regression analysis to provide a summary measure of the fit of a model to a sample of observations that were not themselves used to estimate the model.
+There were no significant differences (paired t-test: t=1.8875, df  = 21, p-value = 0.07299).
+
+\begin{table}[H]
+	\begin{center}
+	\begin{tabular}{ | l | c | c |}
+		\hline
+		Feature & GAM Press Statistic & LM Press Statistic \\ \hline
+		ppron & 18.591104613 & 20.001055542 \\ \hline
+		social & 21.681732112 & 24.906646115  \\ \hline
+		NounsNormalised & 0.002818941 & 0.002797798\\ \hline
+		function & 46.546071705 & 46.031993602 \\ \hline
+		conj & 8.812330188 & 8.354974826 \\ \hline
+		PronounsNormalised & 0.002284437 & 0.002356129\\ \hline
+		Analytic & 2272.200519272 & 2239.224863610 \\ \hline
+		Pronoun & 44.557282749 & 46.894576916 \\ \hline		
+		NN & 101685.870974277 & 96485.575598505 \\ \hline
+		male & 5.191815443 & 5.278559152 \\ \hline
+		UniqueWords & 255114.636660673 & 238042.427437454 \\ \hline
+		WDT & 1987.149134694 & 1958.781236655 \\ \hline
+		nouns & 276723.222762842 & 262192.120091097 \\ \hline
+		nouns/100 & 27.670762078 & 26.219212009 \\ \hline
+		UniqueStems & 170450.645283172 & 166302.725530118 \\ \hline
+		shehe & 5.188136314 & 5.265385185 \\ \hline
+		VBZ & 13889.871831536 & 13721.175422577 \\ \hline
+		JJ & 24950.440222526 & 24692.043386181 \\ \hline
+		article & 13.500676675 & 13.468083961 \\ \hline	
+		Adjectives & 30210.006270270 & 29865.240707991 \\ \hline	
+		Adjectives.100 & 3.020898615 & 2.986524071 \\ \hline
+		Dic & 40.099289973 & 38.501060715 \\ \hline	
+	\end{tabular}
+	\caption{\label{tab:table-name}Comparison of GAM and Linear Model using the PRESS statistic}
+	\end{center} 
+\end{table}
+
+\section{Discussion}\label{discussion}
+President Reagan received his diagnosis of AD in August 1994 but using transcripts of speeches he made in his two terms as President (January 1981 - January 1989) we have be able to identify certain changes in his use of language that we might ascribe to the onset of MCI and early AD.
+Despite differences in our methodology, our research supports some of the findings of Berisha and Liss in that we find an increase in non-specifc noun usage.
+Compared to our controls (GWHB and DJT), we find some slight trends with GWHB but no such trends with DJT in his speech albeit his samples of speech span a shorter amount of time.
+Interestingly, when we normalised the various types of words used by the presidents we found some interesting patterns that further differentiated RR from the controls.
+Whilst Non-specific nouns increased over time, we found that noun usage in general significantly decreased and pronouns increased similarly significantly.
+The increase in pronoun for those with early AD has been identified in literature, although there are only a few studies that explore this \cite{Wendelstein2015}.
+Wendlestein et al propose that the increased used of pronouns is an expression of an impaired ability to adapt language to the listener's needs \cite{Wendelstein2015}.
+Almor et al attributed this reliance on pronouns due to a impaired working memory \cite{Almor1999}.
+\par 
+The decrease in overall noun usage has also been identified as a feature.
+Jarrold et al found that AD patients would use more pronouns, verbs and fewer nouns than controls \cite{Jarrold2014}.
+Wendlestein in their investigations into noun usage found that decreased later on in AD progression and was unaffected in the pre-clinical stages of AD \cite{Wendelstein2014}.
+Our results are supported by existing literature and this potentially means that language analysis in the way we have structured it may have diagnostic or prognostic properties.
+\par 
+A criticism of Berisha and Liss's work is the problems they had with normalising the transcripts in terms of length.
+This was also a problem in the work of Garrard et al \cite{Garrard2005, Le2011}.
+Whilst it is important to control for outliers, there are other ways in which we can control for length of sample.
+In this paper, we controlled for transcript length by dividing any features that were raw counts by the total length of the transcript.
+When we did this, we found that there was a significant decrease in the number of unique words (R=-0.56, p\textless0.001) used however when we controlled via normalisation, we found that this was not a significant feature (R=-0.172, p=0.25).
+\par 
+Overall, we are able to show differences in the a number of language features between RR, GHWB and DJT over time and the psychological literature confirms our findings in reference to these changes.
+\par
+Our next hypothesis involved exploring whether linear models were the most appropriate way to track this longitudinal data.
+In our analysis, we found there was no significant difference in terms of mean sum of squares errors between the linear model and the generalised additive model however this does not mean that these are equivalent.
+For example, considering the plot below.
+We compare a model fit to a linear model with a generalised additive model which aims to model the data more closely.
+It is clear to see that both models can track the movements of language use over time however, the linear model is quite a rigid model which does not allow for different inflection points.
+We can see that RR's decline is not linear, and that there are periods of relative stability and some periods where the decline is more severe.
+We argue that these more nuanced changes over time can only be modelled by a model such as a GAM.
+\par
+In her paper on developing diagnositic criteria for Mild Cognitive Impairment, Albert et all \cite{Albert2011} states 'it is important to obtain longitudinal assessments of cognition, whenever possible' and 'obtaining objective evidence of progressive declines in cognition over time is important for establishing the accuracy of the diagnosis, as well as for
+assessing any potential treatment response.' As we have shown, it is not the case that language declines in a linear way and whilst we have a number of data points that we can use to model RR's decline, in a clinical setting it would be impractical to have upwards of 40 data collection points.
+We feel that  it is important that we have more than just two.
+In the figures below, we can see that if we measured language at Reagan's first transcript to a point after 700 days the vast majority of points would point to a decline in language.
+However, the next figure shows what would happen if we measured language from Reagan's second transcript, just 34 days later.
+In this case, we can see that in some cases we would mark Reagan as improving when it is clear that he is not.
+\par 
+There are limitations of this research.
+Whilst in terms of age, DJT is certainly more suitable as a control to match with RR, in some ways they held very different styles of press conferences in that RR preferred to do solo press conferences and DJT has shown a preference for doing joint press conferences which have an impact on the amount of language produced.
+This artefact of the data is in itself notable as it illustrates the problems we may have with smaller amounts of speech and the problem of finding an appropriate control is a common one in this domain, given the considerable variation in factors such as age, language ability and education.
+As mentioned above, we are lucky enough to have numerous samples of data collected over a long period of time and this is not something that can be easily translated to a clinical protocol.
+Finally, the variability in the number and quality of the transcriptions raises some doubts as to the results of DJT.
+Is it the lack of instances of data, or that the number of words per transcript is significantly less than with RR or GHWB.
+This does have further implications for using a protocol such as this in a more general way as it will impact how data is collected.
+\par 
+With further work, it is not feasible to the vast array of samples over a timeframe, as we have had with the president corpus and so it would be worth exploring how the quality of these predictions may lessen when faced with considerably fewer samples and over a smaller time period.
+It would also be worth extending this research further to encompass more of the linguistic features Fraser used in her work \cite{Fraser2015} to see if there are any further insights to be gained.
+In addition, this replication and extension has demonstrated the potential utility of using longitudinal data as a means of comparing language use of a person at two or more time periods and using this information as a diagnostic aid for MCI and therefore more work would be helpful from a longitudinal perspective to see if this approach may be valid in moving towards a solution for this particular problem.
+
+\section{Conclusions}\label{conclusions}
+The results of this work show that we can track a person's use of language through time in a number of ways and that it is possible for an individual to be his or her own control.
+This is important as it means the heterogenous nature of the MCI population does not impact results as much as if we were comparing those with group of MCI patients with a group of controls.
+Equally, it would be helpful to have controls to ascertain what would be usual to expect in the decline of language in a healthy older adult.
+\par 
+The results of this work also identify some clues as to what could work in a clinical setting, or how we might be able to collect data in a way that accurately tracks language decline without making incorrect assumptions.
+\par
+From a clinical perspective, we can see that using samples (albeit for this dataset) is able to track languages changes over a given time frame.
+This potentially means that we can use a similar methodology to collect regular language samples in settings such as memory clinics and GP's surgeries and even potentially in the people's homes and that these language samples, have the potential to act as an early warning sign for Mild Cognitive Impairment that potentially will identify patients at risk of developing Alzheimer's Disease.
 
 \chapter{Pilot study of the methodology developed}
-\include{Chapter6}
+\section{Introduction}
+Here is the text of your introduction.
+
+\begin{equation}
+    \label{simple_equation}
+    \alpha = \sqrt{ \beta }
+\end{equation}
+
+\subsection{Subsection Heading Here}
+Write your subsection text here.
+
+\section{Conclusion}
+Write your conclusion here.
 
 \chapter{General Discussion, Conclusions and Future Work}
-\include{Chapter7}
+\section{Introduction}
+Here is the text of your introduction.
+
+\begin{equation}
+    \label{simple_equation}
+    \alpha = \sqrt{ \beta }
+\end{equation}
+
+\subsection{Subsection Heading Here}
+Write your subsection text here.
+
+\section{Conclusion}
+Write your conclusion here.
 
